### 2020-02-04 manuscript.Rmd
Add paper draft

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
knit              : worcs::cite_all
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2020-02-04 paper/Manuscript.Rmd
Add paper draft

---
title             : "WORCS: A Workflow for Open Reproducible Code in Science"
shorttitle        : "WORKFLOW FOR OPEN SCIENCE"

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"

affiliation:
  - id            : "1"
    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"

abstract: |
  Abstract here

keywords          : "open science; reproducibility; r; dynamic document generation; version control"
wordcount         : "7356"

bibliography      : ["worcs.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : yes

class             : "man"
classoption       : "noextraspace"
output            : papaja::apa6_pdf
urlcolor          : blue
---

```{r load_packages, include = FALSE}
library("papaja")
library(motley)
#source("rfunctions.R")
#load("results.RData")
#load("table_coefficients.RData")
mask <- TRUE
masked <- function(text){
  ifelse(mask, "[masked]", text)
}
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```

\setlength{\parskip}{0em}

\vspace{\baselineskip}

<!-- Structure
# Introduction
# Defining Open Science
* TOP guidelines
# Goal of this paper
* Grass-roots
* Open science as opportunity ?move
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction
# Introducing the tools
# Setup - do this only once
# WORCS - steps to follow for each project
# Discussion
* advantages
* limitations
* future directions
-->

# Introduction

The social sciences are amidst a paradigm shift towards open science. Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud, increasing awareness of questionable research practices, and the replication crisis. However, Open science is not merely a cure for this crisis - it is also an opportunity. Technological advances now enable us to perform reliable, cumulative, collaborative science more easily. Capitalizing on these advances has the potential to accellerate scientific progress. This paper introduces WORCS: Workflow for Open, Reproducible Code in Science. WORCS is a lightweight approach to open science and computational reproducibility. It is easy to implement, and still ensures reproducibility in most circumstances. Adoption is further facilitated by the availability of a WORCS project template.

# Defining Open Science

One of the most impactful contributions to Open Science has been the development of the TOP-guidelines, and the effort to urge publishers to adopt them. WORCS defines the goals of Open Science in terms of the eight standards described in the TOP-guidelines: Comprehensive citation of literature, data, materials, and methods (1); data sharing (2), sharing code required to reproduce analyses (3), sharing new research materials (4), and design and analysis (5); pre-registration of studies before data collection (6), and pre-registration of the analysis plan prior to analysis (7); and replication of published results (8). WORCS helps researchers meet each of these guidelines, with one exception: We do not address replication of published results, because that pertains to the topic of research, instead of its execution. We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).

## Goal of this paper

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by Journals, funders, and organizations. Yet the value of grassroots change should not be underestimated. Many individual researchers are willing to adopt best practices for Open Science, even in the absence of institutional support. The goal of this paper is therefore to present a workflow that can help individual scientists to meet the principles of Open Science. This workflow can be used regardless of any top-down Open Science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them, and if they do not exist,

## Existing solutions

There have been several previous efforts to promote grass-roots adoption of Open Science principles. Each of these efforts has a different scope, strengths, and limitations that set it apart from WORCS. For instance, @aalbersbergMakingScienceTransparent2018 suggested publishing a "TOP-statement" as supplemental material, which discloses the authors' adherence to Open Science principles. This is very easy to implement, and addresses TOP-guidelines 1-7. A limitation of this approach, which WORCS addresses, is that it does not clarify *how* researchers can meet the TOP-guidelines. One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement. A second relevant source is @peikertReproducibleDataAnalysis2019, who describe a workflow for reproducible analyses. Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5. Peikert and Brandmaier aim to achieve strict computational reproducibility by conducting analyses on a virtual machine, that can be stored like a sort of time capsule and reinstated on different systems, instead of on a private computer. The workflow is excellent, but difficult to implement for non-programmers. WORCS instead uses a much more lightweight approach to safeguard computational reproducibility under *most circumstances*. This should be sufficient for most users, but can also be combined with Peikert and Brandmaier if strict computational reproducibility is essential. WORCS thus addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines. Still, it can be combined with existing approaches.

```{r, echo = FALSE, eval = FALSE}
tab1 <- data.frame(
  Reference = c("Aalbers et al., 2018", "Peikert & Brandmaier, 2019"),
  Scope = c("Disclose adherence to Open Science principles", "Completely reproducible analysis"),
  Difficulty = c("Easy", "Hard"),
  Effectiveness = c("Low", "High"),
  "TOP-guidelines" = c("1-7", "2,3,5")
)
apa_table(tab1)
```

<!-- 
* Grass-roots
* Open science as opportunity
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction
-->

## Introducing the tools

WORCS relies on a few free, open source software solutions that, when combined, facilitate meeting the TOP-guidelines related to citation, sharing, and preregistration. The first tool is *dynamic document generation*. Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses. The text is automatically formatted as an APA-style manuscript, and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables. Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text. Although there is a slight learning curve to transitioning to DDG, the investment really pays off: First, the time saved from painstakingly copy-pasting output and manually formatting text soon outweighs the investment of switching to a new program; second, human error in copy-pasting results is eliminated; third, when revisions require major changes to the analyses, all results, Figures and Tables are automatically updated; and finally, flexibility in output formats means that a manuscript can be rendered to presentation format, or even to a website. In sum, while writing academic papers in a programming environment might seem counter-intuitive at first, this approach is much more amenable to the needs of academics than most word processing software. It prevents mistakes, and saves time.

The second tool is *version control*: Maintaining an indelible log of every change to all project files. Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax. Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions. WORCS instead uses the *explicit* version control software "Git". Git tracks changes to files, and stores these changes when the user makes a "commit". Git retains a complete log of all commits, and users can compare changes between different commits, or go back to a previous version of the code (for example, after making a mistake, or to replicate a previous version of the results). A project version controlled with Git is called a "Repository", or Repo.

<!--In statistical consultations, I often hear that researchers are unable to reproduce specific results because they only worked with a "very specific" version of the syntax - usually a combination of -->

The functionality of Git is amplified by services such as GitHub (other services are available). GitHub is best understood as a cloud storage service with social networking functionality. The "cloud storage" aspect of GitHub works as follows: You can "clone" a local Git repository to the GitHub website, as a backup or research archive. The "social network" aspect comes into play when the project is set to "public": This allows other researchers to peruse the repository on the GitHub website to see how the work was done, clone it to their own computer to replicate the original work or apply the methods to their own data, open Issues to ask questions or give feedback on the project, or even send a "Pull request" with changes to the text or code for your consideration. Git and GitHub shine as tools for collaboration, because different people can simultaneously work on different parts of a project, and their changes can be compared and automatically merged on the website. Even on solo projects, working with Git/GitHub has many benefits: Staying organized, being able to start a new study with a clone of an old, similar repository, or splitting off an "experimental branch" to try something new, while retaining the ability to revert to a previous state of the project, or to merge the experimental branch.

The third tool is *dependency management*: Keeping track of exactly what software was used to conduct the analyses. <!-- Being able to replicate analyses requires not only access to the data and analysis code, but also to the software used to evaluate that code.--> At first glance, it might seem sufficient to state that analyses were conducted in *Program X*. However, every program is susceptible to changes and updates. R-packages in particular are updated frequently, because they are Open Source, and there is an active community of developers contributing functionality and bugfixes. Potentially, any update could change the results of the code. For example, R-version 3.6.0 fixed a bug in the random number generator. This means that any analyses that use random numbers, such as bootstrapping, cannot be strictly replicated from previous versions.

Many solutions exist to ensure computational reproducibility. These solutions typically work by enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs. These solutions differ in user-friendliness and effectiveness. If strict reproducibility is required, one might use the aforementioned workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine. This is quite effective, but very difficult to set up. A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project. This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (in my case, hours) and a lot of hard drive space. Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing all software used, it simply records the calendar date, and then installs the version of all required packages that was available on the central R repository `CRAN` on that date from an archival website. Thus, all that is needed to restore a project is the date on which it was conducted. This is very user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN` packages, and cannot mix package versions from different dates.

A recently developed package strikes a great balance between user-friendliness and effectiveness: `renv`, developed by the team behind Rstudio. This package maintains a text-based log file of which packages are used, in what versions, and where they were installed from (e.g., CRAN, Bioconductor, GitHub). Being text-based, this log can be version controlled with Git. On a local computer, `renv` installs all packages used in a cache that is shared between all projects that use `renv`. Thus, if one of your projects uses a version of a package that is already in the cache, it does not need to be installed again. This overcomes the limitations of long installation times and large space requirements of `renv`. When someone else loads your project, `renv` will install all of the required packages onto a cache on their computer. Whereas `packrat` copies the source code of all packages, `renv` obtains the packages from their original repositories. Thus, as long as these repositories used maintain their archives, this solution is very lightweight and effective. It is not foolproof, but some have argued that all solutions for computational reproducibility have a limited shelf life [@brownHowLearnedStop2017]. Ultimately, the best solution might be to use a good-enough solution, like `renv`, and acknowledge that all code requires some maintenance if you want to reproduce it in the future.

## Additional considerations

A few additional considerations were considered in developing WORCS. The first consideration is to develop a workflow for R, instead of for SPSS or Python. Several arguments support the choice to work in R. First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for Open Science. Second, all of the aforementioned tools for an Open Science workflow are implemented in R. In general, many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones. Third, at the time of writing, R is the second-most cited statistical software package [@PopularityDataScience2012], following SPSS, which has no support for any of the tools discussed in this paper. Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format. Working with R is simplified immensely by using the Rstudio user interface for R. Rstudio integrates many of the tools used in WORCS in a user-friendly manner, and automates and streamlines many tedious or complicated aspects of working with R. This renders Rstudio a catch-all solution for generating open, reproducible code.

Another consideration is to develop a workflow that relies on text-based files only. These files can be read by machines and humans alike, without requiring special (licensed) software. Text-based files are also small, can be version-controlled with Git (binary files should not be version controlled), and can be perused online in repositories such as GitHub.

<!--

- Use RStudio projects to bundle analyses and writing into a self-contained unit
- Use text-based files only
    * `.R` for analyses
    * `.Rmd` for documents
    * `.csv` for data
- Version control all these files with Git
- Ensure full reproducibility of the R environment with the `renv` package, which takes a snapshot of all packages used (and their version numbers), and can restore this configuration on a new system
- Publish the entire self-contained research unit on GitHub
- Connect this GitHub repository to an OSF project

Using text-based files has the advantage that they can be read by machines and humans without requiring special (licensed) software. Text-based files are also small, can be version-controlled with Git, and can be readily uploaded to online repositories.

-->

## Open science as opportunity

Sometimes, Open Science principles are discussed as a kind of "antidote" to scientific misconduct. Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.

## How WORCS meets the goals of Open Science

### Comprehensive citation

TOP encourages comprehensive citation of literature, data, materials, methods, and software. In principle, researchers can meet this requirement by simply citing every reference used. In practice, however, print journals often discourage comprehensive citation due to space constraints. <!--either actively, or passively by including the reference list in the manuscript word count.--> Researchers can nonetheless meet the  comprehensive citation requirement by preparing two versions of the manuscript: A version with comprehensive citations for online dissemination (e.g., through preprint servers), and a print version that includes only essential citations. The Open Science statement in the print version should address the fact that additional references are available in the online version, e.g.:

> In the spirit of Open Science, an online repository is available for this manuscript at XXX, which contains more comprehensive citations, the data, analysis code, new research materials, and pre-registration forms.

The `worcs` package includes functions that can be used render a manuscript with, or without, comprehensive citations. The procedure to use these functions is:

1. During writing, maintain a plain-text `.bib` file with the BibTeX references for all citations.
  + You can make this file by hand; e.g., Figure \@ref(fig:scholarbib) shows how to obtain a BibTeX reference from Google Scholar; simply copy-paste each reference into the `.bib` file
  + You can export a `.bib` file from most reference manager programs; the free, open-source reference manager [Zotero](https://www.zotero.org/download/) is excellent and user-friendly. [Here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-rmarkdown/) is a tutorial for using Zotero with Rmarkdown.
2. To cite a reference, use the `citekey` - the first word in the BibTeX entry for that reference. Insert it in the Rmarkdown file like so: `@yourcitekey2020`. For a parenthesized reference, use `[@citekeyone2020; @citekeytwo2020]`. For more options, see the [Rmarkdown cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html).
3. To indicate a *non-essential* citation, mark it with a double at-symbol: `@@nonessential2020`.
4. When Knitting the document, adapt the `knit` command in the YAML header.  
  `knit: worcs::cite_all` renders all citations, and  
  `knit: worcs::cite_essential` removes all *non-essential* citations.
5. Optional: To be extremely thorough, you could make a "branch" of the Github repository for the print version of the manuscript. Only in this branch, you use the function `knit: worcs::cite_essential`. The procedure is documented in [this tutorial](http://rstudio-pubs-static.s3.amazonaws.com/142364_3b344a38149b465c8ebc9a8cd2eee3aa.html).

```{r scholarbib, echo = FALSE}
knitr::include_graphics("scholar_bib.png")
```

<!--1) use comprehensive citation, 2) document which citations are essential, and 3) publishing a pre-print with comprehensive citations, and 4) filtering out non-essential citations for the print version, and referencing the repository with the pre-print.

To facilitate the citation of data, researchers can publish a data paper in specialized journals, such as the [*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/), or publish their data in an online repository - such as a GitHub repository - and include a text file with the preferred citation for the data, and the license that applies to the data, such as Creative Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or [BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/).

Full version that includes all citations
Branch for publication version, with suppressed non-essential citations

Explanation for citations? How best to include
* Note
* In a function
* In a separate document-->

### Data sharing

Data sharing is important for computational reproducibility and secondary analysis. Computational reproducibility means that a third party can exactly recreate the results from the raw data, using the published analysis code. Secondary analysis means that a third party can conduct sensitivity analyses, expliore alternative explanations, or even use existing data to answer a different research question. From an open science perspective, data sharing is always desirable. From a practical point of view, it is not always possible. Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns. In some cases, it is difficult or impossible to share data.

When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it. Any data processing steps should be documented in the code, instead of applying them to the raw data file. This way, the code will be a pipeline from the raw data to the published results.

When data cannot be shared, researchers should aim to safeguard the potential for computational reproducibility and secondary analysis as much as possible. WORCS uses two solutions to accomplish this goal. The first solution is publishing a *checksum* of the raw data upon collection. A checksum is a 32-character summary of a file of any size. Any change to the underlying file results in a different checksum. Thus, one can use a checksum to verify that the contents of a file remain unchanged. If researchers publish the checksum of their raw data and the complete analysis code, then they can prove to an independent auditer that running the public analysis code on the private dataset with the correct checksum results in the published results of their work.

The second solution is to share a synthetic dataset with similar characteristics to the real data. Synthetic data typically mimic the level of measurement and (conditional) distributions of the real data, but complex multivariate relationships are often lost. Thus, analyses on the synthetic data should not be substantively interpreted, and findings from the real data might not be replicated in the synthetic data. Still, sharing synthetic data allows any third party to 1) verify that the published code runs, 2) debug the code, and 3) write valid code for alternative analyses. A third party can send this code to the original authors, who evaluate it on the real data, and send back the results. The ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) functionality on GitHub is a particularly user-friendly way to do this.

The `worcs` package includes three convenience functions to facilitate working with open and closed data. These functions are:

* `open_data(data)`: This function on a data object (a `data.frame` or `matrix`) generates a file called `data.csv`, updates the `.gitignore` file to stop ignoring this specific `.csv` file, and writes a checksum for this data file to `checksums.csv`.
* `closed_data(data)`: This function generates a file called `data.csv` and a synthetic dataset called `synthetic_data.csv`, updates the `.gitignore` file to stop ignoring `synthetic_data.csv` file, and writes a checksum for the real data file to `checksums.csv`.
* `load_data()`: This function loads the real data if it is available (i.e., the file `data.csv` exists, and its checksum is correct), and otherwise, loads the synthetic data.

The WORCS-template additionally creates a file called `prepare_data.R`. This file is intended to document any steps that must be taken to get your data *into* R, and that are not intended to be repeated when re-running the analyses. Use this file to report any preprocessing required to obtain a neat `data.frame`. For example, if the data was originally in SPSS format, this file might just contain the code required to read the SPSS file. Typically, the last line of the `prepare_data.R` file will be `open_data()` or `closed_data()`. The proper analysis code will then start with `load_data()`.

### Sharing code, research materials, design and analysis

(3), sharing new research materials (4), and design and analysis (5); pre-registration of studies before data collection (6), and pre-registration of the analysis plan prior to analysis

### Preregistration

Lindsay [-@lindsayResearchPreregistration1012016] defines preregistration as "creating a permanent record of your study plans before you look at the data. The plan is stored in a date-stamped, uneditable file in a secure online archive." Two such archives are well-known in the social sciences: AsPredicted.org, and OSF.io. However, GitHub also conforms to these standards. Thus, it is possible to preregister a study simply by pushing a text document with the study plans to GitHub. If the GitHub repository is Public, the preregistration is immediately visible to the world, and Reviewers can submit comments through pull requests. If the repository is Private, Reviewers can be invited as "Collaborators", although this removes their anonymity. Private repositories can be made public at a later date, along with their entire time-stamped history.

Which studies can be preregistered is a matter of some debate. Some have taken the position that studies can only be preregistered prior to data collection. This is the only way to ensure that a researcher is not "HARKing": Hypothesizing after the results are known. From this point of view, it makes little

#### Preregistering study plans

Several excellent preregistration templates are available in the R-package `prereg` [@austPreregMarkdownTemplates2019], including templates from AsPredicted.org, OSF.io, and from individual researchers [e.g., @vantveerPreregistrationSocialPsychology2016]. When opening a new Rstudio project with the WORCS project template, it is possible to select one of these preregistration templates, which will generate a file called `preregistration.Rmd`. This file should be used to document study plans, ideally prior to data collection.

#### Preregistering analyses

The ideal preregistered analysis consists of a complete analysis script that can be evaluated once the data are obtained. This ideal is often unattainable, because the data present researchers with unanticipated challenges; e.g., assumptions are violated, or analyses work differently than expected. Some of these challenges can be avoided by simulating the data one expects to obtain, and writing the analysis syntax based on the simulated data. This topic is beyond the scope of the present paper, but many user-friendly methods for simulating data are available in `R`, including the package [`simstudy`](https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html). When using the WORCS-workflow, it is recommended to preregister, if possible, a `.R` script with the planned analyses, along with a verbal, conceptual description. If any changes must be made to the R-script after obtaining the data, one can refer to the conceptual description to explain why these changes are necessary.

WORCS uses the following approach to preregistration:

1. Document study plans in a `.Rmd` file, and planned analyses in a `.R` file.
2. Commit these documents to the local Git repository, and push them to GitHub
3. On GitHub, [tag the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration". A tagged release helps others retrieve this commit.
4. Optional: Upload the files as attachments to another pre-registration repository, e.g., AsPredicted.org or OSF.io.
  + The `.Rmd` file can be Knitted to a PDF form before uploading to a repository

<!--
3. Write conceptual preregistration
* https://osf.io/zab38/wiki/home/ Templates of OSF Registration Forms
* R package prereg
* mention other examples and cite van 't Veer, A. E., & Giner-Sorolla, R. (2016). Pre-registration in social psychology---A discussion and suggested template. Journal of Experimental Social Psychology, 67, 2--12. doi: 10.1016/j.jesp.2016.03.004
3. Commit, push, and tag th as preregistration
Preregistration: tag on github
https://stackoverflow.com/questions/18216991/create-a-tag-in-a-github-repository

4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
4. Collect data and either add to repository, or register hash of raw datafile
5. Conduct analyses and write; commit regularly (ideally, after completing a small and clearly defined task), with informative commit messages. Call renv::snapshot() to save the state of the project library to the lockfile (called renv.lock),
6. Create OSF project, connect to GitHub, mention OSF page in manuscript
7. Optional: Publish preprint in a not-for-profit preprint repository (check Sherpa Romeo) http://sherpa.ac.uk/romeo/index.php
7. Submit
8. Make repository public -->

WORCS helps researchers meet each of these guidelines, with one exception: We do not address replication of published results, because that pertains to the topic of research, instead of its execution. We group the remaining seven guidelines into three categories: Citation (1), sharing (2-4), and preregistration (6-7).

# Workflow for Open Reproducible Code in Science

## Preparing your system

Before you can use the WORCS workflow, you have to install the required software, and register for a GitHub account. The setup process is documented in a vignette included with the `worcs` R-package, and also [available online](https://cjvanlissa.github.io/worcs/articles/setup.html).

## Phase 1: Study design

1. Create a new Private repository on github, copy the https:// link to clipboard  
  The link should look something like https://github.com/yourname/yourrepo.git
2. In Rstudio, click File > New Project > New directory > WORCS Project Template
    a. Paste the GitHub Repository address in the textbox
    b. Keep the checkbox for `renv` checked if you want to document all dependencies (recommended)
    c. Select a preregistration template
3. Write the preregistration `.Rmd`
4. In the top-right corner of Rstudio, select the Git tab, select the checkboxes next to all files, and click the Commit button. Write an informative message for the commit, e.g., "Preregistration", again click Commit, and then click the green Push arrow to send your commit to GitHub
5. Go to the GitHub repository for this project, and tag the Commit as a preregistration
6. Optional: Render the preregistration to PDF, and upload it to AsPredicted.org or OSF.io as an attachment
7. Optional: Add study Materials (to which you own the rights) to the repository. It is possible to solicit feedback (by opening a GitHub Issue) and acknowledge outside contributions (by accepting Pull requests)

## Phase 2: Data analysis

8. Read the data into R, and document this procedure in `prepare_data.R`
9. Use `open_data()` or `closed_data()` to store the data
10. Write the manuscript in `Manuscript.Rmd`, using code chunks to perform the analyses.
11. Regularly commit your progress to the Git repository; ideally, after completing each small and clearly defined task. Use informative commit messages. Push the commits to GitHub.
12. Cite essential references with one at-symbol (`[@essentialref2020]`), and non-essential references with a double at-symbol (`[@@nonessential2020]`).

## Phase 3: Paper submission

13. To save the state of the project library (all packages used), call `renv::snapshot()`. This updates the lockfile, `renv.lock`.
14. To render the paper with essential citations only for submission, change the line `knit: worcs::cite_all` to `knit: worcs::cite_essential`. Then, press the Knit button to generate a PDF

## Phase 4: Publication

Note that some researchers take these steps at an earlier stage. However, for newcomers to Open Science, who are unsure about potential implications for the publication of their work, it might be prudent to perform these actions after acceptance, and before publication:

13. Make the GitHub repository public
14. [Create an OSF project](https://help.osf.io/hc/en-us/articles/360019737594-Create-a-Project); although you may have already done this in Step 6.
15. [Connect your GitHub repository to the OSF project](https://help.osf.io/hc/en-us/articles/360019929813-Connect-GitHub-to-a-Project)
16. Add an Open Science statement to the manuscript, with a link to the OSF project
17. Optional: [Publish preprint in a not-for-profit preprint repository such as PsyArchiv, and connect it to your existing OSF project](https://help.osf.io/hc/en-us/articles/360019930533-Upload-a-Preprint)
    + Check [Sherpa Romeo](http://sherpa.ac.uk/romeo/index.php) to be sure that your intended outlet allows the publication of preprints; many journals do, nowadays - and if they do not, it is worth considering other outlets.

<!-- Now, the worcs package adds a template to Rstudio

Template repository 
Template repositories let users generate new repositories with the same directory structure and files. Indicate if cjvanlissa/tidySEM can be used as a template for creating other repositories.

You’ve registered a free GitHub account (chapter 4).
You’ve installed/updated R and RStudio (chapter 5).
You’ve installed Git (chapter 6).
You’ve introduced yourself to Git (chapter 7).
You’ve confirmed that you can push to / pull from GitHub from the command line (chapter 9).

3. Write conceptual preregistration
* https://osf.io/zab38/wiki/home/ Templates of OSF Registration Forms
* R package prereg
* mention other examples and cite van 't Veer, A. E., & Giner-Sorolla, R. (2016). Pre-registration in social psychology---A discussion and suggested template. Journal of Experimental Social Psychology, 67, 2--12. doi: 10.1016/j.jesp.2016.03.004
3. Commit, push, and tag as preregistration
<!--Preregistration: tag on github
https://stackoverflow.com/questions/18216991/create-a-tag-in-a-github-repository

4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
-->

# Discussion

* Dealing with collaborators
    - Add to Repo
    - Add Issues for tasks that still need to be accomplished, and tag Collaborators to address them
    - Handle pull requests from collaborators
* advantages
* limitations
* future directions

\newpage

# References
```{r create_r-references}
#r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
\setlength{\parskip}{0em}

<div id = "refs"></div>
\endgroup

### 2020-02-04 paper/notes.Rmd
Add paper draft

<!-- End; below here are notes -->

<!-- We will introduce these tools, before explaining how they adress the TOP-guidelines.

-->

### Dynamic Document Generation

<!--are essential o Here, we combine four software tools, whose interplay can guarantee full computational reproducibility of data analyses and their reporting. There are various ideas on how to enhance reproducibility [@piccoloToolsTechniquesComputational2016], four of which we believe to be particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016], and containerization [@clyburne-sherinComputationalReproducibilityContainers2018]. We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report. Various implementations of these concepts exist, but we consider the following four best suited for the R environment [@r2019]: for literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014]. Each of these software solutions serves a valuable meta-scientific goal (reproducibility) and increases the researches' productivity. They all are highly flexible and powerful, so that mastering them fully, requires a considerable amount of practice. However, for our purposes, it is sufficient to master a valuable minimal subset of functions to ensure reproducibility of scientific analyses.
-->

# Citation

Sometimes have to shorten list of citations for publication; keep online version of document with full citations for every publication, package, etc
Include a document with a list of every citation, and which information from the original source substantiates the claim

## Data

The requirement of open data can be easily met by including the data in a Git repository. When doing so, it is best to use a text-based format (like `.csv`): Git is designed for tracking changes in text-based; `.csv` is displayed in tabular format on GitHub, and anyone can open text files on any computer without requiring licensed software.

From a reproducibility point-of-view, it is desirable to share data in the rawest form possible. Pre-processing, data cleaning, and handling of missing values should all be part of the reproducible workflow, documented in a file called `data_cleaning.R`.
The only action that should always be performed prior to data sharing is anonymization.

Anonymization is often a strict requirement for sharing primary data, so make sure that any non-anonymized data are never tracked in your git repository. To prevent accidental sharing of non-anonymized data, the `.gitignore` file in the WORCS template ignores all commonly used data-related filetypes. When saving data using `open_data()`, a file called `data.csv` is generated, and added as an exception to `.gitignore`. The `closed_data()` function also saves the data, but does not alter the `.gitignore` file. The recommended procedure is thus to first fully anonymize the raw data, then save the anonymized file using `open_data()` or `closed_data()` and start tracking it in Git, and finally, document all preprocessing steps in `data_cleaning.R`.

### When data cannot be shared

Sometimes primary data cannot be shared, due to privacy constraints, ethical concerns, legal limitations (e.g., using controlled-acces data), or the volume of data. It is important to note that when data is not shared, the reproducibility and transparancy of the research is impeded. Here, discuss best practices for open science with closed data.

#### Validating files

When data are closed, and not version controlled in Git, there is a risk that the local file might be changed, thereby rendering the entire analysis non-reproducible. To ensure that all analyses are conducted on the original, unchanged file, one can generate a "checksum" when saving `data.csv` [REF Peikert]. A checksum is a highly compressed, 32-character representation of the file. If the contents of the file change, the checksum changes too. The functions `open_data()` and `closed_data()` in the `worcs` package generate such a checksum along with a `data.csv` file, and the function `load_data()` validates the checksum before loading the file.

#### Synthetic data

If the original data cannot be shared, it is good practice to include a simulated dataset with similar properties to the original data. Others can use the simulated dataset to review the original analysis code or try out alternative analyses. This will result in code that works with the original data as well, which can be submitted to the original authors through a [pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) on GitHub. Using a pull request has the benefit that anyone can see that additional analyses have been requested, thereby providing an incentive for the original authors to run these analyses on the original dataset, and share the outcomes. The function `closed_data` generates a synthetic dataset using the `synthpop` package, with rudimentary default settings. For a customized synthetic dataset, it is possible to follow [a tutorial](https://www.r-bloggers.com/generating-synthetic-data-sets-with-synthpop-in-r/). I recommend using `method = "ranger"`, as the ranger algorithm is much faster than the default (REF ranger).

Sometimes external requirements make it impossible fot the data to be stored and shared together with the scripts. In most of the cases we have seen, these are either space constraints or privacy considerations. In these cases, of course, unrestricted reproducibility is not guaranteed. If dividing data and scripts is unavoidable, we recommend validating all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them. To do this, a checksum must be created and stored at the time of the original analysis. At the time of reproduction, the current checksum must then be compared with the stored checksum to ensure data integrity.

Create checksum for raw datafile
Publish anonymized data in repository
Synthesize anonymized datafile and share that
Check checksum in analyses

## Code

Publish all code and track changes to the analysis process

## Materials

Publish

## Design

Publish

## Pre-registration of studies

Push to private git repository and tag commit
Later, when git repository is made public, people can verify the commit

## Pre-registration of analysis plan

Push to private git repository and tag commit
Later, when git repository is made public, people can verify the commit

## Replication

focus on incentives to motivate researchers to adopt open science principles; this checklist focuses on *means* for adopting these principles. It aims to lower the threshold for compliance by providing simple solutions
* TOP is geared towards Journals, funders, and societies. Acknowledges that Open Science needs grassroots support CODES empowers individual researchers to comply with open science principles even in the absence of top-down requirements.

## Means vs ends

Contribution vis-a-vis TOP-guidelines

## Keynote lectures and special events focused on Open Science

Efficiency productivity transparency response to interdisciplinary research needs
Make research, data, and publications to all levels of iinquiring society

## Transparency

1. Preregistration
2. Version control

Van Aken: We have to invest in open data, and learn to use each other's data (presidential address ECDP 2019)

## Data anonymization

<!-- https://piwik.pro/blog/the-ultimate-guide-to-data-anonymization-in-analytics/ -->

## Replication

https://journals.sagepub.com/doi/abs/10.1177/0956797617747090

"Recently, theory-of-mind research has been revolutionized by findings from novel implicit tasks suggesting that at least some aspects of false-belief reasoning develop earlier in ontogeny than previously assumed and operate automatically throughout adulthood. Although these findings are the empirical basis for far-reaching theories, systematic replications are still missing. This article reports a preregistered large-scale attempt to replicate four influential anticipatory-looking implicit theory-of-mind tasks using original stimuli and procedures. Results showed that only one of the four paradigms was reliably replicated. A second set of studies revealed, further, that this one paradigm was no longer replicated once confounds were removed, which calls its validity into question. There were also no correlations between paradigms, and thus, no evidence for their convergent validity. In conclusion, findings from anticipatory-looking false-belief paradigms seem less reliable and valid than previously assumed, thus limiting the conclusions that can be drawn from them."

Text book mentions replication: Developmental Research Methods By Scott A. Miller

Conceptual replication more feasible; says something about *generalizability* of findings (see also Coyne)

## Review
@rooyenEffectPeerReview2010:
Conclusion Telling peer reviewers that their signed reviews might be available in the public domain on the BMJ’s website had no important effect on review quality. Although the possibility of posting reviews online was associated with a high refusal rate among potential peer reviewers and an increase in the amount of time taken to write a review, we believe that the ethical arguments in favour of open peer review more than outweigh these disadvantages.

@walshOpenPeerReview2000:
A total of 245 reviewers (76%) agreed to sign. Signed reviews were of higher quality, were more courteous and took longer to complete than unsigned reviews. Reviewers who signed were more likely to recommend publication.

<!--@ross-hellauerWhatOpenPeer2017:
OPR definition: Open peer review is an umbrella term for a number of overlapping ways that peer review models can be adapted in line with the aims of Open Science, including making reviewer and author identities open, publishing review reports and enabling greater participation in the peer review process. The full list of traits is:

Open identities: Authors and reviewers are aware of each other’s identity

Open reports: Review reports are published alongside the relevant article.

Open participation: The wider community to able to contribute to the review process.

Open interaction: Direct reciprocal discussion between author(s) and reviewers, and/or between reviewers, is allowed and encouraged.

Open pre-review manuscripts: Manuscripts are made immediately available (e.g., via pre-print servers like arXiv) in advance of any formal peer review procedures.

Open final-version commenting: Review or commenting on final “version of record” publications.

Open platforms (“decoupled review”): Review is facilitated by a different organizational entity than the venue of publication.-->

Signed review: Accountability
Publons (credit for review)
Give permission to include review as supplementary material or upload on osf

Post-publication review: OSF has comments feature; github issues

I sign my reviews consistently, and include the following text in the review and in the message to the Editor:

In the spirit of open science, transparency, and accountability, I sign my reviews. In the   
context of the replication crisis, I consider these principles to be more important than the
anonimity of blind review. I have reviewed this paper in good faith, to the best of my ability,
and with the sole intention of providing helpful suggestions that might improve the quality of
the work. If my comments cause offense, please know that this was unintentional. This review is
licensed under a Creative Commons Attribution-NoDerivatives 4.0 International
License. The terms of this license are available at
http://creativecommons.org/licenses/by-nd/4.0/ Practically speaking: You can use my review 
as usual, but the publisher cannot assert copyright over it. You can share (parts of) my
review with attribution, if you so choose. Best of luck!

Sincerely,
<name and affiliation>

As of yet, I have not received any comments from Editors about the note or the fact that I sign my reviews.

## TOP Guidelines

Transparency and openness promotion

### How top is implemented

TOP Statements are standardized tools for disclosing research outputs such as datasets.
Registered Reports protect research against biased analysis and publication.
Open Science Badges signal transparent research.

### 8 modular standards:

### Citation

Under Open Science principles, citation standards are expanded to give credit to the authors of data and software used for analysis.

Many developmental psychological studies are secondary analyses of existing, large, longitudinal datasets. It is not uncommon for such papers to accrue several coauthors who were responsible for initial data collection, but whose contribution to the paper was limited or non-existent. This practice is in violation of some journals' publication guidelines, although it does not conflict with the APA guidelines, which define an author as "anyone involved with initial research design, data collection and analysis, manuscript drafting, and final approval" (American Psychological Association, 2019, 10/25/2019).

Many ethical guidelines hold that co-authors do not merely share credit for a publication, but are also all responsible for its content - although it can be debated whether this is reasonable and fair [@helgessonResponsibilityScientificMisconduct2018]. A practical solution, already implemented by many journals, is to provide a footnote specifying each author's contributions [@national2009being]. If a journal does not request this information by default, authors can still choose to include it in a footnote. Finally, using Git for version control has the added advantage that every author's contributions are fully traceable.

Given that co-authors incur some responsibility over the work, it may be desireable for all parties involved to separate authorship from credit given for data collection. This can be accomplished by publishing "data papers" that describe the design and procedures of the study. Some new journals focus only on the publication of data papers, such as the *Journal of Open Psychology Data*. Needless to say, these journals are also a great way to learn about data that could be used for secondary analysis.

One common practice in our field

Open citation standards require   citation for data and materials

(1); transparant disclosure or sharing of data (2), methods, including the code required to reproduce them (3), research materials (4), and design and analysis (5); pre-registration of studies before data collection (6), and pre-registration of the analysis plan prior to analysis (7); and replication of published results (8).

Adequate citation of statistical software is important, because open source software is an important type of research output for statisticians. An easy way to cite all packages used is to write an APA-style paper in Rmarkdown using the `papaja` package. Cite all packages with the function `cite_r()`, and add the BibTeX entries for these packages to the reference file using `r_refs("filename.bib")`. However, it might not be feasible or desireable (e.g., due to journal limits on the number of references) to cite all packages. Therefore, a more practical solution might be to cite only the most important packages. The citation for each package can be obtained by running `citation("packagename")`. If this approach is combined with a fully reproducible R-environment (using the `renv` package) and code sharing (e.g., on Github or the OSF), then the full list of package dependencies is documented automatically in the `.lock` file.

### Citation software

It is worth mentioning the free, open-source reference manager Zotero, because it can be used as easily in conjunction with Rstudio as with Word; the plugin BetterBibTex can export continuously updated `.bib` files for citation in Rmarkdown. Finally, it is worth noting that, as of June 2019, Zotero automatically flags retracted papers in your database. This might stop discredited papers from unnecessarily muddling the literature.
<!--Retracted papers

Retracted papers are notoriously persistent in the literature, continuing to accumulate citations long after their findings have been debunked. (See “The Zombie Literature,” The Scientist, May 2016.) The UCLA group’s Oncogene paper, for example, was cited at least 15 times between being flagged on PubPeer in 2014 and being retracted two years later. Moreover, retraction notices themselves are often opaque, making it unclear what exactly led to a paper’s retraction, or how authors behaved during the process.
-->

## Open code

Sharing code is essential for computational reproducibility. The easiest way to share fully reproducible code is to work with an Rstudio project that is version controlled using Git, combined with the `renv` package to ensure all package dependencies are tracked. Such a project can be uploaded ("pushed") wholesale to a code repository, such as GitHub. Projects on GitHub can be publically accessible, so that other researchers can browse the code or copy ("clone") the entire project. It is also possible to make the project private until the manuscript is accepted, which has the benefit that one can show the entire evolution of the analyses from conception to publication.

3. Design and analysis
Sets standards for research design disclosures

4. Pre-registration of analysis plans
Specification of analytical details before data collection

5. Data
Disclose, require, or verify shared data

6. Research materials
Disclose, require, or verify shared materials

7. Pre-registration of studies
Specification of study details before data collection

8. Replication
Encourages publication of replication studies

### Across 3 tiers:

1. Disclosure: The  article  must  disclose whether  or  not  materials are available.
2. Requirement: The  article  must  share materials when possible.
3. Verification: Third party must verify that the standard is being met.

Transparency, open sharing, and reproducibility are core values of science, but not always part of daily practice. Journals, funders, and societies can increase research reproducibility by adopting the TOP Guidelines.

### 2020-02-05 paper/Manuscript.Rmd
Writing 5-2-2020

@@
 However, Open science is not merely a cure for this crisis - it is also an opportunity.
-Technological advances now enable us to perform reliable, cumulative, collaborative science more easily.
+Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science.
 Capitalizing on these advances has the potential to accellerate scientific progress.
-This paper introduces WORCS: Workflow for Open, Reproducible Code in Science.
+
+This paper introduces WORCS: A Workflow for Open Reproducible Code in Science.
 WORCS is a lightweight approach to open science and computational reproducibility.
-It is easy to implement, and still ensures reproducibility in most circumstances.
-Adoption is further facilitated by the availability of a WORCS project template.
+It provides a step-by-step procedure that researchers can follow to make an entire research project Open and Reproducible.
+WORCS is based on best practices, and can be used either in parallel to, or in absence of, top-down requirements by journals.
+The main goal of WORCS is to lower the threshold for adopting an Open Science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an Rstudio project template.
+
+<!--
+thereby lowering the threshold for adoption.
+, and still ensures reproducibility in most circumstances. 
+ It is easy to implement, thereby lowering the threshold
+
+* Grass-roots
+* Open science as opportunity
+    - Easier work
+    - Less mistakes
+    - Better collaboration
+    - Cumulative science
+    - Error correction
+
+Benefits to individual scientists: Automation (after a learning curve), thus more productive, avoid mistakes, join projects, clone projects and make them your own.
+Collective benefits: Cumulative science, -->
 
@@
 One of the most impactful contributions to Open Science has been the development of the TOP-guidelines, and the effort to urge publishers to adopt them.
-WORCS defines the goals of Open Science in terms of the eight standards described in the TOP-guidelines: Comprehensive citation of literature, data, materials, and methods (1); data sharing (2), sharing code required to reproduce analyses (3), sharing new research materials (4), and design and analysis (5); pre-registration of studies before data collection (6), and pre-registration of the analysis plan prior to analysis (7); and replication of published results (8).
-WORCS helps researchers meet each of these guidelines, with one exception: We do not address replication of published results, because that pertains to the topic of research, instead of its execution.
+These guidelines describe eight standards for Open Science: 1) Comprehensive citation of literature, data, materials, and methods; sharing 2) data, 3) code required to reproduce analyses, 4) new research materials, and 5) details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results.
+WORCS defines the goals of Open Science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the topic of a research project, instead of its execution.
 We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).
@@
 The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by Journals, funders, and organizations.
-Yet the value of grassroots change should not be underestimated.
+Yet there is also substantial momentum for grassroots change.
 Many individual researchers are willing to adopt best practices for Open Science, even in the absence of institutional support.
-The goal of this paper is therefore to present a workflow that can help individual scientists to meet the principles of Open Science.
-This workflow can be used regardless of any top-down Open Science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them, and if they do not exist,
+The formation of "Open Science Communities" at universities around the globe illustrates this movement.
+This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of Open Science.
+This workflow can be used regardless of any top-down Open Science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them, and if they do not exist, facilitates meeting them.
 
@@
 ```
-
-<!-- 
-* Grass-roots
-* Open science as opportunity
-    - Easier work
-    - Less mistakes
-    - Better collaboration
-    - Cumulative science
-    - Error correction
--->
 
@@
 Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.
+
 Working with R is simplified immensely by using the Rstudio user interface for R.
@@
 This renders Rstudio a catch-all solution for generating open, reproducible code.
+One important feature of Rstudio is project management.
+A project bundles writing, analyses, data, references, etcetera, into a self-contained unit that can be uploaded to GitHub and downloaded by future users.
+The `worcs` R-package installs a new Rstudio project template, that performs much of the bookkeeping required to set up an Open Science project automatically  behind the scenes.
 
-Another consideration is to develop a workflow that relies on text-based files only.
-These files can be read by machines and humans alike, without requiring special (licensed) software.
-Text-based files are also small, can be version-controlled with Git (binary files should not be version controlled), and can be perused online in repositories such as GitHub.
+Another key consideration is what filetypes to use.
+A case can be made to use text-based files only (or as much as possible), instead of binary files.
+Text-based files can be read by machines and humans alike, without requiring special (licensed) software.
+Binary files must be decoded by software first.
+Git is designed to keep track of changes to the content of text-based files.
+Each change reflects a word or line of code added or removed.
+Binary files such as `.docx` or `.sav`, on the other hand, should not be version controlled.
+The changes in the encrypted file are uninterpretable.
+Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub.
+Uploading large binary files is frowned upon.
+GitHub also allows visitors to peruse text-based files online, and conveniently renders certain filetypes.
+For example, GitHub highlights syntax in `.R` and `.Rmd` files, and displays `.csv` files as spreadsheets.
 
@@
 - Connect this GitHub repository to an OSF project
-
-Using text-based files has the advantage that they can be read by machines and humans without requiring special (licensed) software.
-Text-based files are also small, can be version-controlled with Git, and can be readily uploaded to online repositories.
-
 -->
@@
 In principle, researchers can meet this requirement by simply citing every reference used.
-In practice, however, print journals often discourage comprehensive citation due to space constraints.
-<!--either actively, or passively by including the reference list in the manuscript word count.--> Researchers can nonetheless meet the  comprehensive citation requirement by preparing two versions of the manuscript: A version with comprehensive citations for online dissemination (e.g., through preprint servers), and a print version that includes only essential citations.
-The Open Science statement in the print version should address the fact that additional references are available in the online version, e.g.:
+Unfortunately, citation of data and software is less commonplace than citation of literature and materials.
+Crediting these resources is important, because it incentivizes data sharing and the development of open-source software, and supports the open science efforts of others.
 
-> In the spirit of Open Science, an online repository is available for this manuscript at XXX, which contains more comprehensive citations, the data, analysis code, new research materials, and pre-registration forms.
+<!--Various initiatives are underway to innovate how scientific output is measured and valued. https://www.nwo.nl/beleid/statement+waarderen+en+belonen+van+wetenschappers --> 
+To facilitate citing datasets, researchers sometimes publish *data papers*; documents that detail the procedure, sample, and codebook.
+Specialized journals, such as the [*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/), aid in the publication of these data papers.
+For smaller projects, researchers sometimes publish their data in an online repository - like GitHub - along with a text file with the preferred citation for the data (which can be a substantive paper), and the license that applies to the data, such as Creative Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or [BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/).
+When in doubt, one can always contact the data creators and ask what the preferred citation is.
 
-The `worcs` package includes functions that can be used render a manuscript with, or without, comprehensive citations. The procedure to use these functions is:
+When using R, software (R packages) can be conveniently cited by calling `citation("packagename")`.
+This returns an APA-style reference, and a BibTeX entry.
+However, users of the WORCS project template for Rstudio do not need to manually cite R-packages.
+WORCS generates a template manuscript in APA6 style, based on the `papaja` package [@austPapajaPrepareReproducible2020], which automatically cites all packages used.
+
+One important impediment to comprehensive citation is the fact that print journals operate with space constraints.
+Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count.
+Researchers can overcome this impediment by preparing two versions of the manuscript: One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
+The print version should reference the online version in the print version, so interested readers can find the comprehensive reference list.
+Step 20 of the WORCS procedure suggests uploading the online version to a preprint server.
+This is important because most preprint servers - [at least the ones hosted by OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
+This means that authors will receive credit for cited work; even if they are cited only in the online version.
+Moreover, preprint servers ensure that the online version will have a persistent DOI, and will remain reliably accessible, just like the print version.
+
+It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication.
+Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference.
+Users can render a manuscript to PDF either with, or without, comprehensive citations.
+
+The procedure for citation in WORCS is as follows:
 
@@
   + You can make this file by hand; e.g., Figure \@ref(fig:scholarbib) shows how to obtain a BibTeX reference from Google Scholar; simply copy-paste each reference into the `.bib` file
-  + You can export a `.bib` file from most reference manager programs; the free, open-source reference manager [Zotero](https://www.zotero.org/download/) is excellent and user-friendly.
+  + You can export a `.bib` file from most reference manager programs; the free, open-source reference manager [Zotero](https://www.zotero.org/download/) is excellent and user-friendly, and highly interoperable with other commercial reference managers.
 [Here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-rmarkdown/) is a tutorial for using Zotero with Rmarkdown.
@@
 
-<!--1) use comprehensive citation, 2) document which citations are essential, and 3) publishing a pre-print with comprehensive citations, and 4) filtering out non-essential citations for the print version, and referencing the repository with the pre-print.
-
-To facilitate the citation of data, researchers can publish a data paper in specialized journals, such as the [*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/), or publish their data in an online repository - such as a GitHub repository - and include a text file with the preferred citation for the data, and the license that applies to the data, such as Creative Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or [BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/).
-
-Full version that includes all citations
-Branch for publication version, with suppressed non-essential citations
-
-Explanation for citations? How best to include
+<!--
+Future: Explain why each reference is cited? How best to include this?
 * Note
@@
 Computational reproducibility means that a third party can exactly recreate the results from the raw data, using the published analysis code.
-Secondary analysis means that a third party can conduct sensitivity analyses, expliore alternative explanations, or even use existing data to answer a different research question.
+Secondary analysis means that a third party can conduct sensitivity analyses, explore alternative explanations, or even use existing data to answer a different research question.
 From an open science perspective, data sharing is always desirable.
@@
 When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it.
-Any data processing steps should be documented in the code, instead of applying them to the raw data file.
+Any data processing steps should be documented in the code, instead of applied to the raw data.
 This way, the code will be a pipeline from the raw data to the published results.
@@
 The second solution is to share a synthetic dataset with similar characteristics to the real data.
-Synthetic data typically mimic the level of measurement and (conditional) distributions of the real data, but complex multivariate relationships are often lost.
-Thus, analyses on the synthetic data should not be substantively interpreted, and findings from the real data might not be replicated in the synthetic data.
-Still, sharing synthetic data allows any third party to 1) verify that the published code runs, 2) debug the code, and 3) write valid code for alternative analyses.
-A third party can send this code to the original authors, who evaluate it on the real data, and send back the results.
+Synthetic data mimic the level of measurement and (conditional) distributions of the real data.
+Sharing synthetic data allows any third party to 1) verify that the published code runs, 2) debug the code, and 3) write valid code for alternative analyses.
+However, it is important to know that complex multivariate relationships present in the real data are often lost in synthetic data.
+Thus, findings from the real data might not be replicated in the synthetic data, and findings in the synthetic data *should not be substantively interpreted*.
+Still, sharing synthetic data facilitates data requests from third parties.
+A third party can write analysis code based on the synthetic data, and send it to the authors who evaluate it on the real data and send back the results.
 The ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) functionality on GitHub is a particularly user-friendly way to do this.
 
-The `worcs` package includes three convenience functions to facilitate working with open and closed data. These functions are:
-
-* `open_data(data)`: This function on a data object (a `data.frame` or `matrix`) generates a file called `data.csv`, updates the `.gitignore` file to stop ignoring this specific `.csv` file, and writes a checksum for this data file to `checksums.csv`.
-* `closed_data(data)`: This function generates a file called `data.csv` and a synthetic dataset called `synthetic_data.csv`, updates the `.gitignore` file to stop ignoring `synthetic_data.csv` file, and writes a checksum for the real data file to `checksums.csv`.
-* `load_data()`: This function loads the real data if it is available (i.e., the file `data.csv` exists, and its checksum is correct), and otherwise, loads the synthetic data.
-
-The WORCS-template additionally creates a file called `prepare_data.R`.
+The procedure for data sharing in WORCS is as follows.
+First, researchers load the data into R.
+This process should be documented in the file `prepare_data.R`, which is automatically created by the WORCS-template.
 This file is intended to document any steps that must be taken to get your data *into* R, and that are not intended to be repeated when re-running the analyses.
@@
 For example, if the data was originally in SPSS format, this file might just contain the code required to read the SPSS file.
-Typically, the last line of the `prepare_data.R` file will be `open_data()` or `closed_data()`.
-The proper analysis code will then start with `load_data()`.
+
+Then, researchers must decide whether to use open or closed data, and call one of these convenience functions:
+
+* `open_data(data)`: This function takes a data object (a `data.frame` or `matrix`), writes it to a file called `data.csv`, updates the `.gitignore` file to stop ignoring this specific `.csv` file, and writes a checksum for this data file to `checksums.csv`.
+The data can now be committed and pushed to GitHub, where they will be open once the GitHub repository is made public.
+* `closed_data(data)`: This function generates a local file called `data.csv`, and a synthetic dataset called `synthetic_data.csv`, updates the `.gitignore` file to stop ignoring `synthetic_data.csv` file, and writes a checksum for the real data file to `checksums.csv`.
+The synthetic data can now be committed and pushed to GitHub, and the real data will be ignored by Git.
+
+Typically, the last line of the `prepare_data.R` file will thus be `open_data()` or `closed_data()`.
+The code for analysis of the data, on the other hand, will typically begin with `load_data()`.
+This function loads the real data if it is available (i.e., the file `data.csv` exists, and its checksum is correct), and otherwise, loads the synthetic data.
+This will have the effect that third party users who copy the GitHub repository will automatically load the synthetic dataset, whereas the study authors, who have the real data stored locally, will automatically load the real data.
 
@@
 
-(3), sharing new research materials (4), and design and analysis (5); pre-registration of studies before data collection (6), and pre-registration of the analysis plan prior to analysis
+When writing a manuscript in Rmarkdown, analysis code is embedded in the prose of the paper.
+Thus, the TOP-guideline of sharing analysis code can be met simply by committing the source code of the manuscript to GitHub, and making this GitHub repository Public upon publication.
+If authors additionally use open data and a reproducible environment (default in WORCS), then a third party can simply replicate all analyses by copying the entire GitHub repository, and Knitting the manuscript on their local computer.
+
+Aside from analysis code, the TOP-guidelines also encourage sharing new research materials, and details of the study design and analysis.
+These goals can be accomplished simply by placing any relevant documents in the Rstudio project folder, committing them to the Git repository, and pushing to GitHub.
+As with any document version controlled in this way, it is advisable to use plain text only.
 
@@
 
-Which studies can be preregistered is a matter of some debate.
+<!--Which studies can be preregistered is a matter of some debate.
 Some have taken the position that studies can only be preregistered prior to data collection.
 This is the only way to ensure that a researcher is not "HARKing": Hypothesizing after the results are known.
-From this point of view, it makes little
+From this point of view, it makes little -->
+
+WORCS uses the following approach to preregistration:
+
+1. Document study plans in a `preregistration.Rmd` file, and planned analyses in a `.R` file.
+2. Commit these documents to the local Git repository, and push them to GitHub
+3. On GitHub, [tag the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration". A tagged release helps others retrieve this commit.
+4. Optional: Upload the files as attachments to another pre-registration repository, e.g., AsPredicted.org or OSF.io.
+  + The `.Rmd` file can be Knitted to a PDF form before uploading to a repository
 
@@
 If any changes must be made to the R-script after obtaining the data, one can refer to the conceptual description to explain why these changes are necessary.
-
-WORCS uses the following approach to preregistration:
-
-1. Document study plans in a `.Rmd` file, and planned analyses in a `.R` file.
-2. Commit these documents to the local Git repository, and push them to GitHub
-3. On GitHub, [tag the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration". A tagged release helps others retrieve this commit.
-4. Optional: Upload the files as attachments to another pre-registration repository, e.g., AsPredicted.org or OSF.io.
-  + The `.Rmd` file can be Knitted to a PDF form before uploading to a repository
 
@@
 
-WORCS helps researchers meet each of these guidelines, with one exception: We do not address replication of published results, because that pertains to the topic of research, instead of its execution.
-We group the remaining seven guidelines into three categories: Citation (1), sharing (2-4), and preregistration (6-7).
-
 # Workflow for Open Reproducible Code in Science
@@
     a. Paste the GitHub Repository address in the textbox
-    b. Keep the checkbox for `renv` checked if you want to document all dependencies (recommended)
+    b.
+Keep the checkbox for `renv` checked if you want to document all dependencies (recommended).
+The only reason not to do this is if you have dependencies outside of the R environment, or if you are using a more comprehensive solution for computational reproducibility, e.g., @peikertReproducibleDataAnalysis2019.
     c. Select a preregistration template
+    d. Click Open in new session, and Create project
 3. Write the preregistration `.Rmd`
@@
 
-13. To save the state of the project library (all packages used), call `renv::snapshot()`. This updates the lockfile, `renv.lock`.
-14. To render the paper with essential citations only for submission, change the line `knit: worcs::cite_all` to `knit: worcs::cite_essential`. Then, press the Knit button to generate a PDF
+13. Save the state of the project library (all packages used), call `renv::snapshot()`. This updates the lockfile, `renv.lock`.
+14.
+Add an Open Science statement to the manuscript, which references the online repository for the project.
+This reference can be masked for blind review.
+The Open Science statement should indicate whether additional references are available in the online repository; whether data, code, materials, and study design details are shared, and whether a pre-registration is available.
+For further guidance, see @aalbersbergMakingScienceTransparent2018.
+Example:
+    > In the spirit of Open Science, an online repository is available at XXX, which contains more comprehensive citations, [the data/a synthetic data file], analysis code, the research materials used, details about the study design, and a tagged commit with a preregistration.
+
+15. To render the paper with essential citations only for submission, change the line `knit: worcs::cite_all` to `knit: worcs::cite_essential`. Then, press the Knit button to generate a PDF
 
@@
 
-13. Make the GitHub repository public
-14. [Create an OSF project](https://help.osf.io/hc/en-us/articles/360019737594-Create-a-Project); although you may have already done this in Step 6.
-15. [Connect your GitHub repository to the OSF project](https://help.osf.io/hc/en-us/articles/360019929813-Connect-GitHub-to-a-Project)
-16. Add an Open Science statement to the manuscript, with a link to the OSF project
-17.
+16. Make the GitHub repository public
+17. [Create an OSF project](https://help.osf.io/hc/en-us/articles/360019737594-Create-a-Project); although you may have already done this in Step 6.
+18. [Connect your GitHub repository to the OSF project](https://help.osf.io/hc/en-us/articles/360019929813-Connect-GitHub-to-a-Project)
+19. Add an Open Science statement to the manuscript, with a link to the OSF project
+20.
 Optional: [Publish preprint in a not-for-profit preprint repository such as PsyArchiv, and connect it to your existing OSF project](https://help.osf.io/hc/en-us/articles/360019930533-Upload-a-Preprint)
@@
 
-<!-- Now, the worcs package adds a template to Rstudio
-
-Template repository 
-Template repositories let users generate new repositories with the same directory structure and files. Indicate if cjvanlissa/tidySEM can be used as a template for creating other repositories.
-
-You’ve registered a free GitHub account (chapter 4).
-You’ve installed/updated R and RStudio (chapter 5).
-You’ve installed Git (chapter 6).
-You’ve introduced yourself to Git (chapter 7).
-You’ve confirmed that you can push to / pull from GitHub from the command line (chapter 9).
-
-3. Write conceptual preregistration
-* https://osf.io/zab38/wiki/home/ Templates of OSF Registration Forms
-* R package prereg
-* mention other examples and cite van 't Veer, A.
-E., & Giner-Sorolla, R.
-(2016).
-Pre-registration in social psychology---A discussion and suggested template.
-Journal of Experimental Social Psychology, 67, 2--12.
-doi: 10.1016/j.jesp.2016.03.004
-3. Commit, push, and tag as preregistration
-<!--Preregistration: tag on github
-https://stackoverflow.com/questions/18216991/create-a-tag-in-a-github-repository
-
+<!-- 
 4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
@@
 
-* Dealing with collaborators
+In this tutorial paper, we have presented a workflow for open reproducible code in science.
+The workflow aims to lower the threshold for grass-roots adoption of open science principles.
+It is accompanied by an R-package with convenience functions, and a WORCS project template for Rstudio.
+This relatively light-weight workflow meets most of the requirements for Open Science, as detailed in the TOP-guidelines.
+However, several limitations and issues for future development remain.
+
+One important remaining issue is how to deal with collaborators when using WORCS.
+Several different scenarios exist.
+In one scenario, only the lead author uses WORCS.
+In this case, it is possible to Knit the manuscript to Microsoft Word (.docx), by changing the line `output: papaja::apa6_pdf` in the manuscript's YAML header to `output: papaja::apa6_docx`.
+There are some limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
+When soliting feedback from co-authors, ask them to use Tracked Changes and comment bubbles in Word.
+Then, manually incorporate their changes into the `manuscript.Rmd` file.
+In most cases, this is the most user-friendly approach, as most lead authors would review changes by co-authors anyway.
+
+A second approach is to send the `manuscript.Rmd` file, and ask them to open (and save) it in Word or Notepad as a plain text file.
+When they send it back, make sure any changes to your local file are committed, and then simply overwrite your local version with their file.
+In Rstudio, select the file in the Git tab, and click the Diff button to examine what changes the collaborator has made relative to your last committed version.
+
+A third scenario is when all collaborators are committed to using WORCS.
+In this case, they can [Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo) from the lead author on GitHub, [clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository), make their own changes, and [send you a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) to incorporate their changes.
+Working this way is extremely conducive to scientific collaboration [@ramGitCanFacilitate2013].
+Moreover, being familiar with this workflow opens doors to new forms of collaboration.
+For example, I became a co-author on several R-packages after submitting pull requests with bug fixes or additional functionality.
+
+Finally, @peikertReproducibleDataAnalysis2019 mention that the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] might be useful for collaborating with Word users.
+This package enables bidirectional conversion between Word and Rmarkdown, with support for tracked changes.
+However, a simpler solution (i.e., asking collaborators to work in plain text) might be sufficient.
+
     - Add to Repo

### 2020-02-05 paper/Manuscript.Rmd
Update manuscript

@@
   - name          : "Caspar J. van Lissa"
-    affiliation   : "1"
+    affiliation   : "1,2"
     corresponding : yes    # Define only one corresponding author
@@
     institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
+  - id            : "2"
+    institution   : "Open Science Community Utrecht"
 
@@
 Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud, increasing awareness of questionable research practices, and the replication crisis.
-However, Open science is not merely a cure for this crisis - it is also an opportunity.
-Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science.
-Capitalizing on these advances has the potential to accellerate scientific progress.
+However, Open science is not merely a cure for this crisis - it is also an opportunity [@@adolphOpenBehavioralScience2012].
+Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012].
+Capitalizing on these advances has the potential to accellerate scientific progress [see also @coyneReplicationInitiativesWill2016].
 
@@
 The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by Journals, funders, and organizations.
-Yet there is also substantial momentum for grassroots change.
+However, large organizations are slow to change.
+For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @@klebelPeerReviewPreprint2020].
+<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 --> At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
 Many individual researchers are willing to adopt best practices for Open Science, even in the absence of institutional support.
@@
 Each of these efforts has a different scope, strengths, and limitations that set it apart from WORCS.
-For instance, @aalbersbergMakingScienceTransparent2018 suggested publishing a "TOP-statement" as supplemental material, which discloses the authors' adherence to Open Science principles.
-This is very easy to implement, and addresses TOP-guidelines 1-7.
-A limitation of this approach, which WORCS addresses, is that it does not clarify *how* researchers can meet the TOP-guidelines.
+For example, there are "signalling solutions"; guidelines to structure disclosure about Open Science practices.
+Specifically, @aalbersbergMakingScienceTransparent2018 suggested publishing a "TOP-statement" as supplemental material, which discloses the authors' adherence to Open Science principles.
+Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based Transparency Checklist that authors can complete online to generate a report.
+Such signalling solutions are very easy to adopt, and they address TOP-guidelines 1-7.
+A limitation of these approaches, that WORCS addresses, is that they does not clarify *how* researchers can meet the TOP-guidelines.
 One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.
@@
 Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses.
-The text is automatically formatted as an APA-style manuscript, and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables.
+The text is automatically formatted as an APA-style manuscript [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables.
 Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text.
@@
 The second tool is *version control*: Maintaining an indelible log of every change to all project files.
-Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax.
+Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax [see also @blischakQuickIntroductionVersion2016].
 Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions.
@@
 First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for Open Science.
-Second, all of the aforementioned tools for an Open Science workflow are implemented in R.
-In general, many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones.
-Third, at the time of writing, R is the second-most cited statistical software package [@PopularityDataScience2012], following SPSS, which has no support for any of the tools discussed in this paper.
+Second, all of the aforementioned tools for an Open Science workflow are already implemented in R.
+Many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones.
+Wrappers in R are often available for tools developed in other programming languages.
+Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the Open Science tools discussed in this paper.
 Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.
@@
 Working with R is simplified immensely by using the Rstudio user interface for R.
-Rstudio integrates many of the tools used in WORCS in a user-friendly manner, and automates and streamlines many tedious or complicated aspects of working with R.
-This renders Rstudio a catch-all solution for generating open, reproducible code.
-One important feature of Rstudio is project management.
-A project bundles writing, analyses, data, references, etcetera, into a self-contained unit that can be uploaded to GitHub and downloaded by future users.
-The `worcs` R-package installs a new Rstudio project template, that performs much of the bookkeeping required to set up an Open Science project automatically  behind the scenes.
+Many of the tools used in WORCS are embedded directly into the visual user interface of Rstudio.
+Rstudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
+This makes Rstudio a comprehensive solution for Open Science research projects.
+Another important feature of Rstudio is project management.
+A project bundles writing, analyses, data, references, etcetera, into a self-contained folder, that can be uploaded entirely to GitHub and downloaded by future users.
+The `worcs` R-package installs a new Rstudio project template.
+When a new project is initialized from this template, much of the bookkeeping required to set up an Open Science project is performed automatically.
 
@@
 
-<!--
-
-- Use RStudio projects to bundle analyses and writing into a self-contained unit
-- Use text-based files only
-    * `.R` for analyses
-    * `.Rmd` for documents
-    * `.csv` for data
-- Version control all these files with Git
-- Ensure full reproducibility of the R environment with the `renv` package, which takes a snapshot of all packages used (and their version numbers), and can restore this configuration on a new system
-- Publish the entire self-contained research unit on GitHub
-- Connect this GitHub repository to an OSF project
--->
-
 ## Open science as opportunity
@@
 
-## How WORCS meets the goals of Open Science
+# How WORCS relates to the TOP-guidelines
 
-### Comprehensive citation
+## Comprehensive citation
 
@@
 
-### Data sharing
+## Data sharing
 
@@
 
-### Sharing code, research materials, design and analysis
+## Sharing code, research materials, design and analysis
 
@@
 
-### Preregistration
+## Preregistration
 
@@
 
-#### Preregistering study plans
+### Preregistering study plans
 
-Several excellent preregistration templates are available in the R-package `prereg` [@austPreregMarkdownTemplates2019], including templates from AsPredicted.org, OSF.io, and from individual researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
+Several excellent preregistration templates are available in the R-package `prereg` [@austPreregMarkdownTemplates2019], including templates from organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
 When opening a new Rstudio project with the WORCS project template, it is possible to select one of these preregistration templates, which will generate a file called `preregistration.Rmd`.
@@
 
-#### Preregistering analyses
+### Preregistering analyses
 
@@
 
-# Workflow for Open Reproducible Code in Science
+# The workflow
 
@@
 The setup process is documented in a vignette included with the `worcs` R-package, and also [available online](https://cjvanlissa.github.io/worcs/articles/setup.html).
+After preparing your system, you can follow the step-by-step procedure outlined below.
 
@@
 
-1. Create a new Private repository on github, copy the https:// link to clipboard  
-  The link should look something like https://github.com/yourname/yourrepo.git
+1. Create a new Private repository on github, and copy the repository link to the clipboard. The link will look like https://github.com/yourname/yourrepo.git
 2. In Rstudio, click File > New Project > New directory > WORCS Project Template
@@
 
-One important remaining issue is how to deal with collaborators when using WORCS.
-Several different scenarios exist.
-In one scenario, only the lead author uses WORCS.
-In this case, it is possible to Knit the manuscript to Microsoft Word (.docx), by changing the line `output: papaja::apa6_pdf` in the manuscript's YAML header to `output: papaja::apa6_docx`.
+One potential challenge is the learning curve associated with the tools outlined in this paper.
+Learning to work with R, Rmarkdown, and Git/GitHub requires an initial time investment.
+The amount of time required is reduced by the availability of tutorials, such as this one, and other tutorials cited throughout this document.
+Moreover, the time investment tends to pay off.
+Working with R opens the door to many cutting edge analysis techniques.
+Working with Rmarkdown saves time and prevents mistakes by avoiding tedious copying of results into a text document.
+Working with Git/GitHub keeps projects organized, prevents accidental loss of work, enables integrating changes by collaborators in a non-destructive manner, and ensures that entire research projects are archived and can be accessed or copied by third parties.
+Thus, the time investment is eminently worthwhile.
+
+Another important challenge is managing collaborations when only the lead author uses Rmarkdown, and the coauthors use Word.
+In this case, it is possible to Knit the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf` in the manuscript's YAML header to `output: papaja::apa6_docx`.
 There are some limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
@@
 In most cases, this is the most user-friendly approach, as most lead authors would review changes by co-authors anyway.
-
-A second approach is to send the `manuscript.Rmd` file, and ask them to open (and save) it in Word or Notepad as a plain text file.
+A second approach is to ask collaborators to work in plain text.
+In this case, send collaborators the `manuscript.Rmd` file, and ask them to open (and save) it in Word or Notepad as a plain text file.
 When they send it back, make sure any changes to your local file are committed, and then simply overwrite your local version with their file.
 In Rstudio, select the file in the Git tab, and click the Diff button to examine what changes the collaborator has made relative to your last committed version.
-
-A third scenario is when all collaborators are committed to using WORCS.
-In this case, they can [Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo) from the lead author on GitHub, [clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository), make their own changes, and [send you a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) to incorporate their changes.
-Working this way is extremely conducive to scientific collaboration [@ramGitCanFacilitate2013].
-Moreover, being familiar with this workflow opens doors to new forms of collaboration.
-For example, I became a co-author on several R-packages after submitting pull requests with bug fixes or additional functionality.
-
 Finally, @peikertReproducibleDataAnalysis2019 mention that the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] might be useful for collaborating with Word users.
@@
 
-    - Add to Repo
-    - Add Issues for tasks that still need to be accomplished, and tag Collaborators to address them
-    - Handle pull requests from collaborators
-* advantages
-* limitations
-* future directions
+If all collaborators are committed to using WORCS, they can [Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo) from the lead author on GitHub, [clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository), make their own changes, and [send you a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) to incorporate their changes.
+Working this way is extremely conducive to scientific collaboration [@ramGitCanFacilitate2013].
+Moreover, being familiar with GitHub opens doors to new forms of collaboration: In the open source software community, continuous peer review and voluntary collaborative acts by strangers who are interested in a project are commonplace [see @adolphOpenBehavioralScience2012].
+This kind of collaboration is budding in scientific software development as well; for example, I became a co-author on several R-packages after submitting pull requests with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018].
+It is also possible to invite such collaboration by [opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues) for tasks that still need to be accomplished, and tag known Collaborators to address them, or invite external collaborators to contribute their expertise.
+
+## Conclusion
+
+WORCS offers a workflow for open reproducible code in science.
+The step-by-step procedure outlined in this tutorial helps researchers make an entire research project Open and Reproducible.
+The accompanying R-package provides user-friendly support functions for several steps in the workflow, and an Rstudio project template to get the project set up correctly.
 
@@
 # References
-```{r create_r-references}
-#r_refs(file = "r-references.bib")
-```
 

### 2020-02-05 R/worcs_template.R

@@ -68,6 +68,9 @@ worcs_template <- function(path, ...) {
             "*.xlsx",
             "*.xls",
             "*.pdf",
+            "*.fff",
+            "*.log",
+            "*.tex",
             "!checksums.csv"),
           file = file.path(norm_path, ".gitignore"), append = TRUE)
     if(grepl("^https://github.com/.+?/.+?\\.git$", remote_repo)){

### 2020-02-08 paper/Manuscript.Rmd
First full draft

@@
 ---
-title             : "WORCS: A Workflow for Open Reproducible Code in Science"
+title: |
+    | WORCS: A Workflow for Open Reproducible Code in Science
+    | Version 0.1.0 (Feb 8, 2020)
+
 shorttitle        : "WORKFLOW FOR OPEN SCIENCE"
@@
 abstract: |
-  Abstract here
+  The social sciences are amidst a paradigm shift towards open science. In part,
+  this transition has been fueled by cases of scientific fraud and increasing
+  awareness of questionable research practices. However, Open science is not
+  merely a cure (or punishment) for this crisis - it is also an opportunity.
+  Technological advances enable researchers to more easily conduct reliable,
+  cumulative, collaborative science. Capitalizing on these advances has the
+  potential to accelerate scientific progress. This tutorial paper introduces a
+  workflow for open reproducible code in science (WORCS). WORCS is a 
+  step-by-step procedure that researchers can follow to make an entire project
+  Open and Reproducible. It is based on best practices, and can be used either
+  in parallel to, or in absence of, top-down requirements by journals. WORCS
+  lowers the threshold for adopting an Open Science workflow by providing this
+  tutorial, along with an R-package with user-friendly support functions, and an
+  Rstudio project template.
 
+authornote: |
+  This is a preprint, not a peer-reviewed publication. Comments, contributions,
+  and coauthors are welcome. Please contact the lead author.
 keywords          : "open science; reproducibility; r; dynamic document generation; version control"
-wordcount         : "7356"
+wordcount         : "5921"
 
@@
 lineno            : yes
-mask              : yes
+mask              : no
 
@@
 urlcolor          : blue
+knit              : worcs::cite_all
 ---
@@
 library(motley)
-#source("rfunctions.R")
-#load("results.RData")
-#load("table_coefficients.RData")
 mask <- TRUE
@@
 The social sciences are amidst a paradigm shift towards open science.
-Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud, increasing awareness of questionable research practices, and the replication crisis.
-However, Open science is not merely a cure for this crisis - it is also an opportunity [@@adolphOpenBehavioralScience2012].
+Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [e.g., @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
+However, Open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012].
 Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012].
@@
 
-One of the most impactful contributions to Open Science has been the development of the TOP-guidelines, and the effort to urge publishers to adopt them.
+One of the most impactful contributions to Open Science has been the development of the TOP-guidelines [@nosekPromotingOpenResearch2015a], and the effort to urge publishers to adopt them.
 These guidelines describe eight standards for Open Science: 1) Comprehensive citation of literature, data, materials, and methods; sharing 2) data, 3) code required to reproduce analyses, 4) new research materials, and 5) details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results.
@@
 However, large organizations are slow to change.
-For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @@klebelPeerReviewPreprint2020].
+For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
 <!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 --> At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
@@
 Each of these efforts has a different scope, strengths, and limitations that set it apart from WORCS.
-For example, there are "signalling solutions"; guidelines to structure disclosure about Open Science practices.
+For example, there are "signalling solutions"; guidelines to structure and incentivize disclosure about Open Science practices.
 Specifically, @aalbersbergMakingScienceTransparent2018 suggested publishing a "TOP-statement" as supplemental material, which discloses the authors' adherence to Open Science principles.
@@
 Such signalling solutions are very easy to adopt, and they address TOP-guidelines 1-7.
-A limitation of these approaches, that WORCS addresses, is that they does not clarify *how* researchers can meet the TOP-guidelines.
-One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.
-A second relevant source is @peikertReproducibleDataAnalysis2019, who describe a workflow for reproducible analyses.
+Many journals now also offer authors the opportunity to earn "badges" for adhering to Open Science guidelines [@kidwellBadgesAcknowledgeOpen2016].
+These signalling solutions help structure authors' disclosures about, and incentivize adherence to, Open Science practices.
+
+A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of Open Science.
+<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example here is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019.
 Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5.
@@
 It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication.
-Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference.
+Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference.
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
@@
 For more options, see the [Rmarkdown cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html).
-3. To indicate a *non-essential* citation, mark it with a double at-symbol: `@@nonessential2020`.
+3. To indicate a *non-essential* citation, mark it with a double at-symbol: `@nonessential2020`.
 4. When Knitting the document, adapt the `knit` command in the YAML header.  
@@
 
-Lindsay [-@lindsayResearchPreregistration1012016] defines preregistration as "creating a permanent record of your study plans before you look at the data.
+Lindsay and colleagues [-@lindsayResearchPreregistration1012016] define preregistration as "creating a permanent record of your study plans before you look at the data.
 The plan is stored in a date-stamped, uneditable file in a secure online archive." Two such archives are well-known in the social sciences: AsPredicted.org, and OSF.io.
@@
 
-<!--Which studies can be preregistered is a matter of some debate.
-Some have taken the position that studies can only be preregistered prior to data collection.
-This is the only way to ensure that a researcher is not "HARKing": Hypothesizing after the results are known.
-From this point of view, it makes little -->
+The advantages, disadvantages, and pitfalls for preregistering different types of studies have been extensively debated, and summarizing that debate falls outside of the scope of this paper, and has been addressed adequately elsewhere [e.g., @lindsayResearchPreregistration1012016].
+We take the pragmatic position that, in confirmatory (hypothesis-testing) research, it is beneficial to plan research projects before executing them, to preregister these plans, and adhere to them.
+All deviations from this procedure should be disclosed; e.g., prior exposure to the data, whether direct (e.g., by computing summary statistics) or indirect (e.g., by reading papers using the same data); deviations from the analysis plan to handle unforseen contingencies, such as violations of model assumptions; or additional exploratory analyses.
 
@@
 This file should be used to document study plans, ideally prior to data collection.
+
+Because the practice of preregistration is strongly tied to the collection of primary data, it has been a matter of some debate whether secondary data analyses can be preregistered [but see @westonRecommendationsIncreasingTransparency2019 for an excellent discussion of the topic].
+Preregistration is the only way to ensure that a researcher is not "HARKing": Hypothesizing after the results are known [@kerrHARKingHypothesizingResults1998].
+When analyzing existing data, it is difficult to *prove* that a researcher did not have direct (or indirect, through collaborators or by reading studies using the same data) exposure to the data, before composing the preregistration.
+However, the question of proof is only relevant from a perspective of preventing scientific misconduct.
+Good faith preregistration efforts always improve the quality of deductive (theory-testing) research, because they avoid HARKing, ensure reliable significance tests, avoid overfitting noise in the data, and limit the number of forking paths researchers wander during data analysis [@gelmanStatisticalCrisisScience2014].
+In all cases, researchers should minimize exposure to the data, and disclose any exposure.
 
@@
 11. Regularly commit your progress to the Git repository; ideally, after completing each small and clearly defined task. Use informative commit messages. Push the commits to GitHub.
-12. Cite essential references with one at-symbol (`[@essentialref2020]`), and non-essential references with a double at-symbol (`[@@nonessential2020]`).
+12. Cite essential references with one at-symbol (`[@essentialref2020]`), and non-essential references with a double at-symbol (`[@nonessential2020]`).
 
@@
 
+<!-- Open peer review? -->
+
 ## Conclusion

### 2020-02-11 R/github.R
Jeroen Ooms' suggestions: * Use gert instead of git2r * Add papaja to Remotes

@@ -1,8 +1,22 @@
-#' @importFrom git2r default_signature
+#' @importFrom gert libgit2_config git_config_global
 has_git <- function(){
-  out <- system("git", ignore.stdout = TRUE) == 1
-  if(!out){
+  config <- libgit2_config()
+  settings <- git_config_global()
+  name <- subset(settings, name == "user.name")$value
+  email <- subset(settings, name == "user.email")$value
+  hasit <- (any(unlist(config[c("ssh", "https")])) & (length(name) && length(email)))
+  if(!hasit){
     warning("Rstudio is not yet connected to Git. You will not be able to use the worcs package yet.", call. = FALSE)
   }
-  out
+  hasit
+}
+
+#' @importFrom gert git_config_global_set
+git_credentials <- function(name, email){
+  invisible(
+    tryCatch({
+    git_config_global_set(name = "user.name", value = name)
+    git_config_global_set(name = "user.email", value = email)
+  }, error = function(e){warning("Could not set git credentials.", call. = FALSE)})
+  )
 }

### 2020-02-11 R/worcs_template.R

@@ -1,5 +1,5 @@
 #' @importFrom rmarkdown draft
-#' @importFrom git2r init remote_add add commit push
+#' @importFrom gert git_init git_remote_add git_add git_commit git_push
 # @importFrom renv init
 worcs_template <- function(path, ...) {
   # collect inputs
@@ -57,34 +57,28 @@ worcs_template <- function(path, ...) {
 
   #use_git() initialises a Git repository and adds important files to .gitignore. If user consents, it also makes an initial commit.
   #usethis::use_github()
-  git_success <- TRUE
-  if(has_git()){
-    init(path = norm_path)
-    write(c(".Rhistory",
-            ".Rprofile",
-            "*.csv",
-            "*.sav",
-            "*.sas7bdat",
-            "*.xlsx",
-            "*.xls",
-            "*.pdf",
-            "*.fff",
-            "*.log",
-            "*.tex",
-            "!checksums.csv"),
-          file = file.path(norm_path, ".gitignore"), append = TRUE)
-    if(grepl("^https://github.com/.+?/.+?\\.git$", remote_repo)){
-      remote_add(name = "origin", url = remote_repo)
-      add(path = "README.md")
-      commit(message = "worcs template initial commit", all = TRUE)
-      system("git push -u origin master")
-      #push(name = "origin", refspec = "refs/heads/master")
-    } else {
-      git_success <- FALSE
-    }
-
+  write(c(".Rhistory",
+          ".Rprofile",
+          "*.csv",
+          "*.sav",
+          "*.sas7bdat",
+          "*.xlsx",
+          "*.xls",
+          "*.pdf",
+          "*.fff",
+          "*.log",
+          "*.tex",
+          "!checksums.csv"),
+        file = file.path(norm_path, ".gitignore"), append = TRUE)
+  if(grepl("^https://github.com/.+?/.+?\\.git$", remote_repo)){
+    tryCatch({
+      git_init(path = norm_path)
+      git_remote_add(name = "origin", url = remote_repo)
+      git_add(files = "README.md")
+      git_commit(message = "worcs template initial commit")
+      git_push(remote = "origin")
+    }, error = function(e){warning("Could not connect to a remote GitHub repository. You are working with a local git repository only.", call. = FALSE)})
   } else {
-    git_success <- FALSE
+    warning("No valid GitHub address provided. You are working with a local git repository only.", call. = FALSE)
   }
-  if(!git_success) warning("Could not connect to a remote GitHub repository. You are working with a local git repository only.", call. = FALSE)
 }

### 2020-03-11 README.Rmd
Add readme and license

---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->
<!--[![CRAN status](https://www.r-pkg.org/badges/version/tidySEM)](https://cran.r-project.org/package=tidySEM)
[![](https://cranlogs.r-pkg.org/badges/tidySEM)](https://cran.r-project.org/package=tidySEM)-->
<!--[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build
status](https://travis-ci.org/cjvanlissa/tidySEM.svg?branch=master)](https://travis-ci.org/cjvanlissa/tidySEM)
[![Codecov test
coverage](https://codecov.io/gh/cjvanlissa/tidySEM/branch/master/graph/badge.svg)](https://codecov.io/gh/cjvanlissa/tidySEM?branch=master)
<!--[![DOI](http://joss.theoj.org/papers/10.21105/joss.00978/status.svg)](https://doi.org/10.21105/joss.00978)-->
# WORCS

The Workflow for Open Reproducible Code in Science (WORCS) is an easy to adopt approach to ensuring a research project meets the requirements of Open Science from the start. It is based on a "good enough" philosophy, opting for user-friendliness instead of thoroughness. It can be used either in absence of, or in parallel to, existing requirements for Open workflows. It can also be enhanced with more elaborate solutions for specific issues.

## Where do I start?

For most users, the recommended starting point is to read [the preprint paper, available on the Open Science Framework](https://osf.io/zcvbs/).

The paper describes the WORCS workflow, and how the `worcs` package fits into that workflow.

## About this repository

This repository contains the following:

1.  An R-package called `worcs`, with convenience functions to facilitate the WORCS workflow.
2.  In the subfolder `./paper`, the source files for the paper describing the WORCS workflow.

The repository serves two functions: To allow users to install the `worcs` package, and to allow collaborators access to the source code for the package and paper.

## Installing the package

Before installing the package, please read [this vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), which explains how to set up your computer for `worcs`.

After reading [the vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), you can install the development version of the `worcs` package from GitHub with:

``` r
install.packages("devtools")
devtools::install_github("cjvanlissa/tidySEM")
```

## Citing WORCS

You can cite WORCS using the following citation (please use the same citation for either the package, or the paper):

> Van Lissa, C. J. (2020, February 11). WORCS: A Workflow for Open Reproducible Code in Science. https://doi.org/10.17605/OSF.IO/ZCVBS

## Contributing and Contact Information

WORCS is actively recruiting collaborators with valuable expertise, to improve both the workflow and the package. Relevant contributions warrant coauthorship to the paper. Please contact the lead author at [c.j.vanlissa@uu.nl](mailto:c.j.vanlissa@uu.nl), or:

  - File a GitHub issue [here](https://github.com/cjvanlissa/tidySEM)
  - Make a pull request
    [here](https://github.com/cjvanlissa/tidySEM/pulls)

By participating in this project, you agree to abide by the [Contributor
Code of Conduct v2.0](https://www.contributor-covenant.org/).

### 2020-03-11 README.Rmd
Try to add project structure

@@
 
+## Repository structure
+
+```{r echo = FALSE, results='asis'}
+descripts <- c(
+  "_pkgdown.yml" = "YAML for package website (do not edit)",
+  "DESCRIPTION" = "R-package DESCRIPTION (do not edit)",
+  "LICENSE.md" = "Project license (do not edit)",
+  "NAMESPACE" = "R-package namespace (machine-written)",
+  "README.md" = "Read this file to get started!  (do not edit)",
+  "README.Rmd" = "R-markdown source for readme.md (human editable)",
+  "worcs.Rproj" = "Rstudio project file (do not edit)",
+  "docs/" = "Package website (machine-written)",
+  "inst/" = "Rstudio project template files  (human editable)",
+  "man/" = "R-package documentation (do not edit)",
+  "paper/" = "WORCS paper source files (human editable)",
+  "R/" = "R-package source code (human editable)",
+  "vignettes/" = "R-package vignettes (human editable)"
+)
+prioritize <- c(
+  "README.md",
+  "LICENSE.md",
+  grep("^(?:(?!/).)*$", names(descripts), value = TRUE, perl = TRUE)
+)
+prioritize <- unique(prioritize)
+f <- list.files(recursive = T)
+struc <- f[!grepl("/", f)]
+dirs <- f[grepl("/", f)]
+dirs <- unique(gsub("/.*$", "/", dirs))
+
+struc <- c(struc, dirs)
+
+struc <- data.frame(file = struc)
+struc$desc <- ""
+struc$desc[match(struc$file, names(descripts))] <- descripts[match(struc$file, names(descripts))]
+
+prioritize <- match(prioritize, struc$file)
+ord <- c(prioritize, c(1:nrow(struc))[-prioritize])
+struc <- struc[ord, ]
+
+pd <- paste0("%", -1*(max(nchar(levels(struc$file)))+2), "s")
+out <- c("```", ".", apply(struc, 1, function(thisrow){
+  paste0("├── ", sprintf(pd, thisrow[1]), "<- ", thisrow[2])
+}), "```")
+cat(out, sep = "\r\n")
+```
+
 ## Installing the package

### 2020-03-11 README.Rmd
Try fix structure

@@
 
-```{r echo = FALSE, results='asis'}
+```{r echo = FALSE, results = "asis"}
 descripts <- c(
@@
 pd <- paste0("%", -1*(max(nchar(levels(struc$file)))+2), "s")
-out <- c("```", ".", apply(struc, 1, function(thisrow){
+out <- c("````", "```", ".", apply(struc, 1, function(thisrow){
   paste0("├── ", sprintf(pd, thisrow[1]), "<- ", thisrow[2])
-}), "```")
+}), "```", "````")
 cat(out, sep = "\r\n")

### 2020-03-11 README.Rmd
Leave readme like this, for now

@@
 pd <- paste0("%", -1*(max(nchar(levels(struc$file)))+2), "s")
-out <- c("````", "```", ".", apply(struc, 1, function(thisrow){
+#out <- c("````", "```", ".", apply(struc, 1, function(thisrow){
+#  paste0("├── ", sprintf(pd, thisrow[1]), "<- ", thisrow[2])
+#}), "```", "````")
+out <- c("```", ".", apply(struc, 1, function(thisrow){
   paste0("├── ", sprintf(pd, thisrow[1]), "<- ", thisrow[2])
-}), "```", "````")
+}), "```")
 cat(out, sep = "\r\n")

### 2020-03-11 paper/Manuscript.Rmd
Add Jeroen Ooms to acknowledgements
remove motley

@@
   and coauthors are welcome. Please contact the lead author.
+  Acknowledgements: Jeroen Ooms (feedback, suggesting gert, and using Remotes
+  field).
 keywords          : "open science; reproducibility; r; dynamic document generation; version control"
@@
 library("papaja")
-library(motley)
 mask <- TRUE

### 2020-03-11 paper/Manuscript.Rmd
Loek's comments

@@
     email         : "c.j.vanlissa@uu.nl"
+  - name          : "Loek Brinkman"
+    affiliation   : "2,3"
 
@@
     institution   : "Open Science Community Utrecht"
+  - id            : "3"
+    institution   : "UMC Utrecht"
 
@@
   this transition has been fueled by cases of scientific fraud and increasing
-  awareness of questionable research practices. However, Open science is not
+  awareness of questionable research practices and failed replications. However,
+  open science is not
   merely a cure (or punishment) for this crisis - it is also an opportunity.
@@
   in parallel to, or in absence of, top-down requirements by journals. WORCS
-  lowers the threshold for adopting an Open Science workflow by providing this
+  lowers the threshold for adopting an open science workflow by providing this
   tutorial, along with an R-package with user-friendly support functions, and an
@@
 # Introduction
-# Defining Open Science
+# Defining open science
 * TOP guidelines
@@
 * Grass-roots
-* Open science as opportunity ?move
+* open science as opportunity ?move
     - Easier work
@@
 Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [e.g., @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
-However, Open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012].
+However, open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012].
 Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012].
@@
 WORCS is based on best practices, and can be used either in parallel to, or in absence of, top-down requirements by journals.
-The main goal of WORCS is to lower the threshold for adopting an Open Science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an Rstudio project template.
+The main goal of WORCS is to lower the threshold for adopting an open science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an Rstudio project template.
 
@@
 * Grass-roots
-* Open science as opportunity
+* open science as opportunity
     - Easier work
@@
 
-# Defining Open Science
+# Defining open science
 
-One of the most impactful contributions to Open Science has been the development of the TOP-guidelines [@nosekPromotingOpenResearch2015a], and the effort to urge publishers to adopt them.
-These guidelines describe eight standards for Open Science: 1) Comprehensive citation of literature, data, materials, and methods; sharing 2) data, 3) code required to reproduce analyses, 4) new research materials, and 5) details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results.
-WORCS defines the goals of Open Science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the topic of a research project, instead of its execution.
+<!-- LB: Although Open Science is advocated by many, there is no clear definition of Open Science (ref multiple definitions).
+An influential operationalisation of Open Science has been de TOP guidlines.
+-->
+One of the most impactful contributions to open science has been the development of the TOP-guidelines [@nosekPromotingOpenResearch2015a], and the effort to urge publishers to adopt them.
+These guidelines describe eight standards for open science: 1) Comprehensive citation of literature, data, materials, and methods; 2) sharing data, 3) sharing the code required to reproduce analyses, 4) sharing new research materials, and 5) sharing details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results.
+WORCS defines the goals of open science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the subject of a research project, not to its execution.
 We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).
@@
 
-The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by Journals, funders, and organizations.
+The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
 However, large organizations are slow to change.
@@
 <!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 --> At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
-Many individual researchers are willing to adopt best practices for Open Science, even in the absence of institutional support.
-The formation of "Open Science Communities" at universities around the globe illustrates this movement.
-This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of Open Science.
-This workflow can be used regardless of any top-down Open Science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them, and if they do not exist, facilitates meeting them.
+Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support.
+The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement.
+This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
+This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them.
 
@@
 
-There have been several previous efforts to promote grass-roots adoption of Open Science principles.
+There have been several previous efforts to promote grass-roots adoption of open science principles.
 Each of these efforts has a different scope, strengths, and limitations that set it apart from WORCS.
-For example, there are "signalling solutions"; guidelines to structure and incentivize disclosure about Open Science practices.
-Specifically, @aalbersbergMakingScienceTransparent2018 suggested publishing a "TOP-statement" as supplemental material, which discloses the authors' adherence to Open Science principles.
+For example, there are "signalling solutions"; guidelines to structure and incentivize disclosure about open science practices.
+Specifically, @aalbersbergMakingScienceTransparent2018 suggested publishing a "TOP-statement" as supplemental material, which discloses the authors' adherence to open science principles.
 Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based Transparency Checklist that authors can complete online to generate a report.
 Such signalling solutions are very easy to adopt, and they address TOP-guidelines 1-7.
-Many journals now also offer authors the opportunity to earn "badges" for adhering to Open Science guidelines [@kidwellBadgesAcknowledgeOpen2016].
-These signalling solutions help structure authors' disclosures about, and incentivize adherence to, Open Science practices.
+Many journals now also offer authors the opportunity to earn "badges" for adhering to open science guidelines [@kidwellBadgesAcknowledgeOpen2016].
+These signalling solutions help structure authors' disclosures about, and incentivize adherence to, open science practices.
 
-A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of Open Science.
+A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of open science.
 <!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example here is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019.
@@
 WORCS instead uses a much more lightweight approach to safeguard computational reproducibility under *most circumstances*.
-This should be sufficient for most users, but can also be combined with Peikert and Brandmaier if strict computational reproducibility is essential.
+This should be sufficient for most users, but can also be combined with Peikert and Brandmaier's approach if strict computational reproducibility is essential.
 WORCS thus addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines.
@@
   Reference = c("Aalbers et al., 2018", "Peikert & Brandmaier, 2019"),
-  Scope = c("Disclose adherence to Open Science principles", "Completely reproducible analysis"),
+  Scope = c("Disclose adherence to open science principles", "Completely reproducible analysis"),
   Difficulty = c("Easy", "Hard"),
@@
 Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text.
-Although there is a slight learning curve to transitioning to DDG, the investment really pays off: First, the time saved from painstakingly copy-pasting output and manually formatting text soon outweighs the investment of switching to a new program; second, human error in copy-pasting results is eliminated; third, when revisions require major changes to the analyses, all results, Figures and Tables are automatically updated; and finally, flexibility in output formats means that a manuscript can be rendered to presentation format, or even to a website.
+Although there is a slight learning curve to transitioning to DDG, the investment really pays off:
+
+* The time saved from painstakingly copy-pasting output and manually formatting text soon outweighs the investment of switching to a new program
+* Human error in copy-pasting results is eliminated
+* When revisions require major changes to the analyses, all results, Figures and Tables are automatically updated
+* Flexibility in output formats means that a manuscript can be rendered to presentation format, or even to a website. 
+* Reproducibility is guaranteed, and can be verified by reviewers and/or readers, because the code is run each time the document is compiled.
+
 In sum, while writing academic papers in a programming environment might seem counter-intuitive at first, this approach is much more amenable to the needs of academics than most word processing software.
@@
 The "cloud storage" aspect of GitHub works as follows: You can "clone" a local Git repository to the GitHub website, as a backup or research archive.
-The "social network" aspect comes into play when the project is set to "public": This allows other researchers to peruse the repository on the GitHub website to see how the work was done, clone it to their own computer to replicate the original work or apply the methods to their own data, open Issues to ask questions or give feedback on the project, or even send a "Pull request" with changes to the text or code for your consideration.
+The "social network" aspect comes into play when the project is set to "public": This allows other researchers to peruse the repository on the GitHub website to see how the work was done, clone it to their own computer to replicate the original work or apply the methods to their own data, open "Issues" to ask questions or give feedback on the project, or even send a "Pull request" with suggested changes to the text or code for your consideration.
 Git and GitHub shine as tools for collaboration, because different people can simultaneously work on different parts of a project, and their changes can be compared and automatically merged on the website.
@@
 
+<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc.
+This could be a place where readers stumble and stop reading.
+Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective.
+[main issue] it should become clear earlier what the WORCS workflow entails: e.g.
+switching to working with a particular R Project Template.
+Discuss the ‘workflow' section earlier.
+Related: Can you make a schematic overview of the WORCS?overview-->
 Many solutions exist to ensure computational reproducibility.
@@
 These solutions differ in user-friendliness and effectiveness.
-If strict reproducibility is required, one might use the aforementioned workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine.
+If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine.
 This is quite effective, but very difficult to set up.
-A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project.
-This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (in my case, hours) and a lot of hard drive space.
+<!-- Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project.
+This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space.
 Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library.
@@
 Several arguments support the choice to work in R.
-First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for Open Science.
-Second, all of the aforementioned tools for an Open Science workflow are already implemented in R.
+First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science.
+Second, all of the aforementioned tools for an open science workflow are already implemented in R.
 Many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones.
 Wrappers in R are often available for tools developed in other programming languages.
-Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the Open Science tools discussed in this paper.
+<!-- LB: Perhaps give some pointers to learn R (e.g.
+R4DataScience)-->Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
 Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.
@@
 Rstudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
-This makes Rstudio a comprehensive solution for Open Science research projects.
+This makes Rstudio a comprehensive solution for open science research projects.
 Another important feature of Rstudio is project management.
@@
 The `worcs` R-package installs a new Rstudio project template.
-When a new project is initialized from this template, much of the bookkeeping required to set up an Open Science project is performed automatically.
+When a new project is initialized from this template, much of the bookkeeping required to set up an open science project is performed automatically.
 
@@
 Text-based files can be read by machines and humans alike, without requiring special (licensed) software.
-Binary files must be decoded by software first.
+Binary files, such as Word (`.docx`) or SPSS (`.sav`, `.spo`),  must be decoded by software first.
 Git is designed to keep track of changes to the content of text-based files.
 Each change reflects a word or line of code added or removed.
-Binary files such as `.docx` or `.sav`, on the other hand, should not be version controlled.
+Binary files, on the other hand, should not be version controlled.
 The changes in the encrypted file are uninterpretable.
@@
 
-Sometimes, Open Science principles are discussed as a kind of "antidote" to scientific misconduct.
+<!-- LB:  this section comes at a weird place in the flow of the paper.
+-->Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct.
 Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.
@@
 It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication.
-Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference.
+Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference.
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
 
-The procedure for citation in WORCS is as follows:
+The procedure for citation in WORCS is as follows:<!-- LB: this procedure is secondary to the main workflow. Consider changing the order (see main issue)-->
 
@@
 
-When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it.
+<!-- LB: As I reader, I would prefer to hear about data sharing options before the synthetic data part (which is only relevant to those who cannot share their data).
+E.g.
+can I indicate a point in my workflow (e.g.
+after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it.
 Any data processing steps should be documented in the code, instead of applied to the raw data.
@@
     d. Click Open in new session, and Create project
-3. Write the preregistration `.Rmd`
+3. Write the preregistration `preregistration.Rmd`
 4.
@@
 14.
-Add an Open Science statement to the manuscript, which references the online repository for the project.
+Add an open science statement to the manuscript, which references the online repository for the project.
 This reference can be masked for blind review.
-The Open Science statement should indicate whether additional references are available in the online repository; whether data, code, materials, and study design details are shared, and whether a pre-registration is available.
+The open science statement should indicate whether additional references are available in the online repository; whether data, code, materials, and study design details are shared, and whether a pre-registration is available.
 For further guidance, see @aalbersbergMakingScienceTransparent2018.
 Example:
-    > In the spirit of Open Science, an online repository is available at XXX, which contains more comprehensive citations, [the data/a synthetic data file], analysis code, the research materials used, details about the study design, and a tagged commit with a preregistration.
+    > In the spirit of open science, an online repository is available at XXX, which contains more comprehensive citations, [the data/a synthetic data file], analysis code, the research materials used, details about the study design, and a tagged commit with a preregistration.
 
@@
 
-Note that some researchers take these steps at an earlier stage.
-However, for newcomers to Open Science, who are unsure about potential implications for the publication of their work, it might be prudent to perform these actions after acceptance, and before publication:
+Note that some researchers take the following steps at an earlier stage.
+However, for newcomers to open science, who are unsure about potential implications for the publication of their work, it might be prudent to perform these actions after acceptance, and before publication:
 
@@
 18. [Connect your GitHub repository to the OSF project](https://help.osf.io/hc/en-us/articles/360019929813-Connect-GitHub-to-a-Project)
-19. Add an Open Science statement to the manuscript, with a link to the OSF project
+19. Add an open science statement to the manuscript, with a link to the OSF project
 20.
@@
 It is accompanied by an R-package with convenience functions, and a WORCS project template for Rstudio.
-This relatively light-weight workflow meets most of the requirements for Open Science, as detailed in the TOP-guidelines.
+This relatively light-weight workflow meets most of the requirements for open science, as detailed in the TOP-guidelines.
 However, several limitations and issues for future development remain.
@@
 When soliting feedback from co-authors, ask them to use Tracked Changes and comment bubbles in Word.
-Then, manually incorporate their changes into the `manuscript.Rmd` file.
+<!-- LB: The work-around with track-changes in Word files is doable, but not very satisfying.
+I think you can also use Overleaf instead of Word (user friendly and web based) for bidirectional communication, but haven’t tried this myself.
+I didn’t know about ‘redoc’, will check it out.-->Then, manually incorporate their changes into the `manuscript.Rmd` file.
 In most cases, this is the most user-friendly approach, as most lead authors would review changes by co-authors anyway.

### 2020-04-25 R/worcs_template.R
Add licenses and table with project content

@@ -6,6 +6,7 @@ worcs_template <- function(path, ...) {
   # collect inputs
   dots <- list(...)
   prereg_template <- dots[["prereg_template"]]
+  add_license <- dots[["add_license"]]
   use_renv <- dots[["use_renv"]]
   remote_repo <- dots[["remote_repo"]]
 
@@ -44,7 +45,6 @@ worcs_template <- function(path, ...) {
 
   # Begin prereg
   if(!tolower(prereg_template) == "none"){
-
 	draft(
 	  file.path(path, "preregistration.Rmd"),
 	  paste0(tolower(prereg_template), "_prereg"),
@@ -55,6 +55,21 @@ worcs_template <- function(path, ...) {
 
   }
   # End prereg
+
+  # Begin license
+  if(!tolower(add_license) == "none"){
+    dir.create(path, recursive = TRUE, showWarnings = FALSE)
+
+    # copy 'resources' folder to path
+    license_dir = system.file('rstudio', 'templates', 'project', 'licenses', package = 'worcs', mustWork = TRUE)
+    license_file <- file.path(license_dir, paste0(add_license, ".txt"))
+    if(file.exists(license_file)){
+      file.copy(license_file, file.path(path, "LICENSE"))
+    } else {
+      warning("Could not find requested license.")
+    }
+  }
+  # End license
 
 # Use renv ----------------------------------------------------------------
   norm_path <- normalizePath(path)
@@ -78,6 +93,27 @@ worcs_template <- function(path, ...) {
           "*.tex",
           "!checksums.csv"),
         file = file.path(norm_path, ".gitignore"), append = TRUE)
+
+  # Update readme
+  if(file.exists(file.path(norm_path, "README.md"))){
+    cont <- readLines(file.path(norm_path, "README.md"))
+    f <- list.files(norm_path)
+    tab <- matrix(c("File", "Description", "Usage",
+                    "README.md", "Description of project", "Human editable"), nrow = 2, byrow = TRUE)
+
+    if(any(grepl("\\.rproj$", tolower(f)))){
+      cont[grep("You can load this project in Rstudio by opening the file called ", cont)] <- paste0(cont[grep("You can load this project in Rstudio by opening the file called ", cont)], "'", grep("\\.rproj$", tolower(f), value = TRUE)[1], "'.")
+      tab <- describe_file(grep("\\.rproj$", tolower(f), value = TRUE)[1], "Project file", "Loads project", tab, norm_path)
+    }
[…]

### 2020-04-25 README.Rmd
Add travis and testthat

--- 2020-03-11 README.Rmd
@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
-<!--[![CRAN status](https://www.r-pkg.org/badges/version/tidySEM)](https://cran.r-project.org/package=tidySEM)
-[![](https://cranlogs.r-pkg.org/badges/tidySEM)](https://cran.r-project.org/package=tidySEM)-->
-<!--[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
-[![Travis build
-status](https://travis-ci.org/cjvanlissa/tidySEM.svg?branch=master)](https://travis-ci.org/cjvanlissa/tidySEM)
-[![Codecov test
-coverage](https://codecov.io/gh/cjvanlissa/tidySEM/branch/master/graph/badge.svg)](https://codecov.io/gh/cjvanlissa/tidySEM?branch=master)
+<!--[![CRAN status](https://www.r-pkg.org/badges/version/worcs)](https://cran.r-project.org/package=worcs)
+[![](https://cranlogs.r-pkg.org/badges/worcs)](https://cran.r-project.org/package=worcs)-->
+[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
+[![Travis build status](https://travis-ci.org/cjvanlissa/worcs.svg?branch=master)](https://travis-ci.org/cjvanlissa/worcs)
+<!--[![Codecov test
+coverage](https://codecov.io/gh/cjvanlissa/worcs/branch/master/graph/badge.svg)](https://codecov.io/gh/cjvanlissa/worcs?branch=master)-->
 <!--[![DOI](http://joss.theoj.org/papers/10.21105/joss.00978/status.svg)](https://doi.org/10.21105/joss.00978)-->
@@
 
-```{r echo = FALSE, results='asis'}
-descripts <- c(
-  "_pkgdown.yml" = "YAML for package website (do not edit)",
-  "DESCRIPTION" = "R-package DESCRIPTION (do not edit)",
-  "LICENSE.md" = "Project license (do not edit)",
-  "NAMESPACE" = "R-package namespace (machine-written)",
-  "README.md" = "Read this file to get started!  (do not edit)",
-  "README.Rmd" = "R-markdown source for readme.md (human editable)",
-  "worcs.Rproj" = "Rstudio project file (do not edit)",
-  "docs/" = "Package website (machine-written)",
-  "inst/" = "Rstudio project template files  (human editable)",
-  "man/" = "R-package documentation (do not edit)",
-  "paper/" = "WORCS paper source files (human editable)",
-  "R/" = "R-package source code (human editable)",
-  "vignettes/" = "R-package vignettes (human editable)"
-)
-prioritize <- c(
-  "README.md",
-  "LICENSE.md",
-  grep("^(?:(?!/).)*$", names(descripts), value = TRUE, perl = TRUE)
-)
-prioritize <- unique(prioritize)
-f <- list.files(recursive = T)
-struc <- f[!grepl("/", f)]
-dirs <- f[grepl("/", f)]
-dirs <- unique(gsub("/.*$", "/", dirs))
-
-struc <- c(struc, dirs)
-
-struc <- data.frame(file = struc)
-struc$desc <- ""
-struc$desc[match(struc$file, names(descripts))] <- descripts[match(struc$file, names(descripts))]
-
-prioritize <- match(prioritize, struc$file)
-ord <- c(prioritize, c(1:nrow(struc))[-prioritize])
-struc <- struc[ord, ]
-
-pd <- paste0("%", -1*(max(nchar(levels(struc$file)))+2), "s")
-out <- c("```", ".", apply(struc, 1, function(thisrow){
-  paste0("├── ", sprintf(pd, thisrow[1]), "<- ", thisrow[2])
-}), "```")
-cat(out, sep = "\r\n")
+```{r echo = FALSE, results = "asis"}
+descripts <- data.frame(matrix(c(
+  "_pkgdown.yml", "YAML for package website", "do not edit",
+  "DESCRIPTION", "R-package DESCRIPTION", "do not edit",
+  "LICENSE.md", "Project license", "do not edit",
+  "NAMESPACE", "R-package namespace", "machine-written",
+  "README.md", "Read this file to get started! ", "do not edit",
+  "README.Rmd", "R-markdown source for readme.md", "human editable",
+  "worcs.Rproj", "Rstudio project file", "do not edit",
+  "docs/", "Package website", "machine-written",
+  "inst/", "Rstudio project template files ", "human editable",
+  "man/", "R-package documentation", "do not edit",
+  "paper/", "WORCS paper source files", "human editable",
+  "R/", "R-package source code", "human editable",
+  "vignettes/", "R-package vignettes", "human editable"
+), ncol = 3, byrow = TRUE))
+names(descripts) <- c("File", "Description", "Usage")
+knitr::kable(descripts)
 ```
@@
 install.packages("devtools")
-devtools::install_github("cjvanlissa/tidySEM")
+devtools::install_github("crsh/papaja", dependencies = TRUE)
+devtools::install_github("cjvanlissa/worcs", dependencies = TRUE)
+tinytex::install_tinytex()
+worcs::git_credentials("your_name", "your_email")
 ```
@@
 
-  - File a GitHub issue [here](https://github.com/cjvanlissa/tidySEM)
+  - File a GitHub issue [here](https://github.com/cjvanlissa/worcs)
   - Make a pull request
-    [here](https://github.com/cjvanlissa/tidySEM/pulls)
+    [here](https://github.com/cjvanlissa/worcs/pulls)
 

### 2020-04-25 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(worcs)
+
+test_check("worcs")

### 2020-04-26 README.Rmd
Add worcs badge

@@
 ---
+
+# WORCS <a href='https://osf.io/zcvbs/'><img src='paper/worcs_icon.png' align="right" height="139" /></a>
+
 <!-- README.md is generated from README.Rmd. Please edit that file -->
@@
 Code of Conduct v2.0](https://www.contributor-covenant.org/).
+
+## Acknowledgements
+
+The worcs logo makes use of the gear, services, gears, preferences, settings icon, made by MD Badsha Meah from www.freeicons.io.

### 2020-04-26 README.Rmd
Update readme

@@
 <!--[![DOI](http://joss.theoj.org/papers/10.21105/joss.00978/status.svg)](https://doi.org/10.21105/joss.00978)-->
-# WORCS
 
 The Workflow for Open Reproducible Code in Science (WORCS) is an easy to adopt approach to ensuring a research project meets the requirements of Open Science from the start.
-It is based on a "good enough" philosophy, opting for user-friendliness instead of thoroughness.
+It is based on a "good enough" philosophy, prioritizing user-friendliness over exhaustiveness.
 It can be used either in absence of, or in parallel to, existing requirements for Open workflows.

### 2020-04-26 README.Rmd
Add examples to functions

@@
 
-The worcs logo makes use of the gear, services, gears, preferences, settings icon, made by MD Badsha Meah from www.freeicons.io.
+The worcs logo is inspired by the Open Science Badges by the Center for Open Science (CC-BY-4.0), and makes use of the gear, services, gears, preferences, settings icon, made by MD Badsha Meah from www.freeicons.io (CC-BY-3.0).

### 2020-04-26 R/citations.R

@@ -7,6 +7,32 @@
 #' @param inputFile The file passed to \code{\link[rmarkdown]{render}}.
 #' @param encoding Ignored. The encoding is always assumed to be UTF-8.
 #' @export
+#' @examples
+#' \dontshow{
+#' library(rmarkdown)
+#' old_wd <- getwd()
+#' dir.create(file.path(tempdir(), "citeall"))
+#' setwd(file.path(tempdir(), "citeall"))
+#' draft("test.Rmd", template = "github_document", package = "rmarkdown",
+#'       create_dir = FALSE, edit = FALSE)
+#' write(c("", "Optional reference: @@reference2020"),
+#'       file = "test.Rmd", append = TRUE)
+#' cite_all("test.Rmd", "UTF-8")
+#' setwd(old_wd)
+#' unlink(file.path(tempdir(), "citeall"))
+#' }
+#' \donttest{
+#' # NOTE: Do not use this function interactively, as in the example below.
+#' # Only specify it as custom knit function in an Rmarkdown file, like so:
+#' # knit: worcs::cite_all
+#'
+#' library(rmarkdown)
+#' draft("test.Rmd", template = "github_document", package = "rmarkdown",
+#'       create_dir = FALSE, edit = FALSE)
+#' write(c("", "Optional reference: @@reference2020"),
+#'       file = "test.Rmd", append = TRUE)
+#' cite_all("test.Rmd", "UTF-8")
+#' }
 cite_all <- function(inputFile, encoding){
   Args <- list(
     inputFile = inputFile,
@@ -20,11 +46,37 @@ cite_all <- function(inputFile, encoding){
 #'
 #' This is a wrapper for \code{\link[rmarkdown]{render}}. First, this function
 #' parses the citations in the document, removing citations
-#' marked with double at sign, e.g.: \code{@@@@reference2020}.  Then, it renders
+#' marked with double at sign, e.g.: \code{@@@@reference2020}. Then, it renders
 #' the file.
 #' @param inputFile The file passed to \code{\link[rmarkdown]{render}}.
 #' @param encoding Ignored. The encoding is always assumed to be UTF-8.
 #' @export
+#' @examples
+#' \dontshow{
+#' library(rmarkdown)
+#' old_wd <- getwd()
+#' dir.create(file.path(tempdir(), "citeessential"))
+#' setwd(file.path(tempdir(), "citeessential"))
[…]

### 2020-04-26 R/save_load.R

@@ -7,8 +7,17 @@
 #' Set to \code{NULL} to avoid creating a codebook. Defaults to \code{"html"},
 #' other options documented in \code{\link[dataMaid]{makeDataReport}}.
 #' @examples
-#' \dontrun{
-#'   open_data(iris)
+#' \dontshow{
+#' the_test <- "opendata"
+#' old_wd <- getwd()
+#' dir.create(file.path(tempdir(), the_test))
+#' setwd(file.path(tempdir(), the_test))
+#' open_data(iris[1:5, ], codebook = NULL)
+#' setwd(old_wd)
+#' unlink(file.path(tempdir(), the_test))
+#' }
+#' \donttest{
+#' open_data(iris)
 #' }
 #' @rdname open_data
 #' @seealso closed_data
@@ -29,7 +38,16 @@ open_data <- function(data, codebook = "html"){
 #' Set to \code{NULL} to avoid creating a codebook. Defaults to \code{"html"},
 #' other options documented in \code{\link[dataMaid]{makeDataReport}}.
 #' @examples
-#' \dontrun{
+#' \dontshow{
+#' the_test <- "closeddata"
+#' old_wd <- getwd()
+#' dir.create(file.path(tempdir(), the_test))
+#' setwd(file.path(tempdir(), the_test))
+#' closed_data(iris[1:5, ], codebook = NULL)
+#' setwd(old_wd)
+#' unlink(file.path(tempdir(), the_test))
+#' }
+#' \donttest{
 #' closed_data(iris)
 #' }
 #' @rdname closed_data
@@ -50,7 +68,8 @@ save_data <- function(data, open, codebook = NULL){
     stop("Argument 'data' must be a data.frame, matrix, or inherit from these classes.")
   }
   if(!file.exists(".gitignore")){
-    stop("Could not find .gitignore file. You might not be working from the main project directory.")
+    message("Could not find .gitignore file. You might not be working from the main project directory. Created .gitignore file in folder: ", getwd())
+    file.create(".gitignore")
   }
   data <- as.data.frame(data)
   if(!is.null(codebook)){
@@ -89,8 +108,18 @@ save_data <- function(data, open, codebook = NULL){
 #' the synthetic data.
 #' @return A data.frame.
 #' @examples
-#' \dontrun{
-#' open_data(iris)
+#' \dontshow{
+#' the_test <- "loaddata"
+#' old_wd <- getwd()
[…]

### 2020-04-26 R/syn_ranger.R

@@ -3,9 +3,11 @@ syn.smooth <-
   getFromNamespace("syn.smooth", "synthpop")
 
 #' @title Generating synthetic data sets using ranger
-#' @description This function is for internal use only.
-#' Generates synthetic version(s) of a data set using the fast
-#' implementation of random forests in \code{\link[ranger]{ranger}}.
+#' @description This function is for internal use only. It extends the
+#' functionality of \code{\link[synthpop]{syn}} in the \code{synthpop} package,
+#' which generates synthetic version(s) of a data set. This function uses the
+#' fast implementation of random forests in \code{\link[ranger]{ranger}} to
+#' generate synthetic datasets.
 #' @param y Vector of the outcome variable.
 #' @param x Matrix of x variables.
 #' @param xp Matrix of x variables for prediction.
@@ -14,11 +16,8 @@ syn.smooth <-
 #' @param ... Additional arguments passed to and from functions.
 #' @return List with predictions and fit model.
 #' @examples
-#' \dontrun{
-#' if(interactive()){
-#'  #EXAMPLE1
-#'  }
-#' }
+#' df <- iris[1:10, ]
+#' res <- syn.ranger(df$Species, df[, 1:4], df[, 1:4], smoothing = FALSE)
 #' @rdname syn.ranger
 #' @importFrom ranger ranger
 #' @importFrom stats predict

### 2020-04-27 paper/Manuscript.Rmd
Correct spelling for RStudio

@@
   tutorial, along with an R-package with user-friendly support functions, and an
-  Rstudio project template.
+  RStudio project template.
 
@@
 WORCS is based on best practices, and can be used either in parallel to, or in absence of, top-down requirements by journals.
-The main goal of WORCS is to lower the threshold for adopting an open science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an Rstudio project template.
+The main goal of WORCS is to lower the threshold for adopting an open science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an RStudio project template.
 
@@
 
-A recently developed package strikes a great balance between user-friendliness and effectiveness: `renv`, developed by the team behind Rstudio.
+A recently developed package strikes a great balance between user-friendliness and effectiveness: `renv`, developed by the team behind RStudio.
 This package maintains a text-based log file of which packages are used, in what versions, and where they were installed from (e.g., CRAN, Bioconductor, GitHub).
@@
 
-Working with R is simplified immensely by using the Rstudio user interface for R.
-Many of the tools used in WORCS are embedded directly into the visual user interface of Rstudio.
-Rstudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
-This makes Rstudio a comprehensive solution for open science research projects.
-Another important feature of Rstudio is project management.
+Working with R is simplified immensely by using the RStudio user interface for R.
+Many of the tools used in WORCS are embedded directly into the visual user interface of RStudio.
+RStudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
+This makes RStudio a comprehensive solution for open science research projects.
+Another important feature of RStudio is project management.
 A project bundles writing, analyses, data, references, etcetera, into a self-contained folder, that can be uploaded entirely to GitHub and downloaded by future users.
-The `worcs` R-package installs a new Rstudio project template.
+The `worcs` R-package installs a new RStudio project template.
 When a new project is initialized from this template, much of the bookkeeping required to set up an open science project is performed automatically.
@@
 This returns an APA-style reference, and a BibTeX entry.
-However, users of the WORCS project template for Rstudio do not need to manually cite R-packages.
+However, users of the WORCS project template for RStudio do not need to manually cite R-packages.
 WORCS generates a template manuscript in APA6 style, based on the `papaja` package [@austPapajaPrepareReproducible2020], which automatically cites all packages used.
@@
 Aside from analysis code, the TOP-guidelines also encourage sharing new research materials, and details of the study design and analysis.
-These goals can be accomplished simply by placing any relevant documents in the Rstudio project folder, committing them to the Git repository, and pushing to GitHub.
+These goals can be accomplished simply by placing any relevant documents in the RStudio project folder, committing them to the Git repository, and pushing to GitHub.
 As with any document version controlled in this way, it is advisable to use plain text only.
@@
 Several excellent preregistration templates are available in the R-package `prereg` [@austPreregMarkdownTemplates2019], including templates from organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
-When opening a new Rstudio project with the WORCS project template, it is possible to select one of these preregistration templates, which will generate a file called `preregistration.Rmd`.
+When opening a new RStudio project with the WORCS project template, it is possible to select one of these preregistration templates, which will generate a file called `preregistration.Rmd`.
 This file should be used to document study plans, ideally prior to data collection.
@@
 1. Create a new Private repository on github, and copy the repository link to the clipboard. The link will look like https://github.com/yourname/yourrepo.git
-2. In Rstudio, click File > New Project > New directory > WORCS Project Template
+2. In RStudio, click File > New Project > New directory > WORCS Project Template
     a. Paste the GitHub Repository address in the textbox
@@
 4.
-In the top-right corner of Rstudio, select the Git tab, select the checkboxes next to all files, and click the Commit button.
+In the top-right corner of RStudio, select the Git tab, select the checkboxes next to all files, and click the Commit button.
 Write an informative message for the commit, e.g., "Preregistration", again click Commit, and then click the green Push arrow to send your commit to GitHub
@@
 The workflow aims to lower the threshold for grass-roots adoption of open science principles.
-It is accompanied by an R-package with convenience functions, and a WORCS project template for Rstudio.
+It is accompanied by an R-package with convenience functions, and a WORCS project template for RStudio.
 This relatively light-weight workflow meets most of the requirements for open science, as detailed in the TOP-guidelines.
@@
 When they send it back, make sure any changes to your local file are committed, and then simply overwrite your local version with their file.
-In Rstudio, select the file in the Git tab, and click the Diff button to examine what changes the collaborator has made relative to your last committed version.
+In RStudio, select the file in the Git tab, and click the Diff button to examine what changes the collaborator has made relative to your last committed version.
 Finally, @peikertReproducibleDataAnalysis2019 mention that the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] might be useful for collaborating with Word users.
@@
 The step-by-step procedure outlined in this tutorial helps researchers make an entire research project Open and Reproducible.
-The accompanying R-package provides user-friendly support functions for several steps in the workflow, and an Rstudio project template to get the project set up correctly.
+The accompanying R-package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.
 

### 2020-04-27 paper/Manuscript.Rmd
CRAN spell check

@@
 5.
-Optional: To be extremely thorough, you could make a "branch" of the Github repository for the print version of the manuscript.
+Optional: To be extremely thorough, you could make a "branch" of the GitHub repository for the print version of the manuscript.
 Only in this branch, you use the function `knit: worcs::cite_essential`.
@@
 3. Commit, push, and tag th as preregistration
-Preregistration: tag on github
+Preregistration: tag on GitHub
 https://stackoverflow.com/questions/18216991/create-a-tag-in-a-github-repository
@@
 
-1. Create a new Private repository on github, and copy the repository link to the clipboard. The link will look like https://github.com/yourname/yourrepo.git
+1. Create a new Private repository on GitHub, and copy the repository link to the clipboard. The link will look like https://github.com/yourname/yourrepo.git
 2. In RStudio, click File > New Project > New directory > WORCS Project Template

### 2020-04-27 paper/notes.Rmd
CRAN spell check

@@
 
-Post-publication review: OSF has comments feature; github issues
+Post-publication review: OSF has comments feature; GitHub issues
 
@@
 The citation for each package can be obtained by running `citation("packagename")`.
-If this approach is combined with a fully reproducible R-environment (using the `renv` package) and code sharing (e.g., on Github or the OSF), then the full list of package dependencies is documented automatically in the `.lock` file.
+If this approach is combined with a fully reproducible R-environment (using the `renv` package) and code sharing (e.g., on GitHub or the OSF), then the full list of package dependencies is documented automatically in the `.lock` file.
 

### 2020-04-27 R/synthetic.R

@@ -37,7 +37,7 @@
 #' \item Using \code{model_expression}, a predictive model is built for column
 #' c, for c along 2:j, with c predicted from columns 1:(c-1) of the original
 #' data.
-#' \item Using \code{predict_expression}, columns 1:(c-1) of the syntethic data
+#' \item Using \code{predict_expression}, columns 1:(c-1) of the synthetic data
 #' are used to predict synthetic values for column c.
 #' }
 #' Variables are thus imputed in order of occurrence in the \code{data.frame}.

### 2020-04-27 R/worcs_template.R

@@ -79,7 +79,6 @@ worcs_template <- function(path, ...) {
   }
 
   #use_git() initialises a Git repository and adds important files to .gitignore. If user consents, it also makes an initial commit.
-  #usethis::use_github()
   write(c(".Rhistory",
           ".Rprofile",
           "*.csv",

### 2020-05-04 R/codebook.R
Address CRAN feedback by Martina Schmirl

@@ -1,11 +1,13 @@
 #' @title Create codebook for a dataset
-#' @description Creates a rudimentary Rmarkdown codebook for a dataset, and
-#' renders it to markdown for GitHub. Users can customize the Rmarkdown
+#' @description Creates a rudimentary 'Rmarkdown' codebook for a dataset, and
+#' renders it to markdown for 'GitHub'. Users can customize the 'Rmarkdown'
 #' document, for example, by adding a column with variable descriptions, or a
 #' paragraph with details on the data collection procedure.
 #' @param data A data.frame for which to create a codebook.
 #' @param render_file Logical. Whether or not to render the document.
-#' @return NULL
+#' @return \code{Logical}, indicating whether or not the operation was
+#' succesful. This function is mostly called for its side effect of rendering an
+#' \code{rmarkdown} codebook.
 #' @examples
 #' library(rmarkdown)
 #' library(knitr)
@@ -21,6 +23,7 @@
 #' @importFrom stats median var
 #' @importFrom utils capture.output
 make_codebook <- function(data, render_file = TRUE){
+  function_success <- TRUE
   data_types <- sapply(data, function(x){paste0(class(x), collapse = ", ")})
   summaries <- descfun(data)
   summaries <- cbind(
@@ -43,46 +46,63 @@ make_codebook <- function(data, render_file = TRUE){
   file_contents <- file_contents[1:(grep("^##", file_contents)[1]-1)]
   dm <- dim(data)
   sum_tab <- capture.output(dput(summaries))
-  write(c(file_contents,
-          "## Dataset description",
-          paste0("The data contains ", dm[1], " cases and ", dm[2], " variables."),
-          "",
-          "## Codebook",
-          "",
-          "```{r}",
-          "summaries <- ",
-          sum_tab,
-          "options(knitr.kable.NA = '')",
-          "knitr::kable(summaries, row.names = FALSE, digits = 2)",
-          "```",
-          "",
-          "### Legend",
-          "",
-          "* __Name__: Variable name",
-          "* __type__: Data type of the variable",
-          "* __missing__: Proportion of missing values for this variable",
-          "* __unique__: Number of unique values",
-          "* __mean__: Mean value",
-          "* __median__: Median value",
-          "* __mode__: Most common value (for categorical variables, this shows the frequency of the most common category)",
-          "* **mode_value**: For categorical variables, the value of the most common category",
[…]

### 2020-05-04 R/github.R

@@ -13,6 +13,6 @@ git_credentials <- function(name, email){
     tryCatch({
     git_config_global_set(name = "user.name", value = name)
     git_config_global_set(name = "user.email", value = email)
-  }, error = function(e){warning("Could not set git credentials.", call. = FALSE)})
+  }, error = function(e){warning("Could not set 'Git' credentials.", call. = FALSE)})
   )
 }

### 2020-05-04 R/synthetic.R

@@ -26,7 +26,7 @@
 #' excellent missForest single imputation technique.
 #' @param verbose Logical, Default: TRUE. Whether to show a progress bar while
 #' running the algorithm and provide informative messages.
-#' @return A \code{data.frame}
+#' @return A \code{data.frame} with synthetic data, based on \code{data}.
 #' @details This function uses a simple algorithm to generate a synthetic
 #' dataset with similar characteristics to the original. The algorithm is as
 #' follows:

### 2020-05-04 R/worcs_template.R

@@ -119,9 +119,9 @@ worcs_template <- function(path, ...) {
       git_add(files = "README.md")
       git_commit(message = "worcs template initial commit")
       git_push(remote = "origin")
-    }, error = function(e){warning("Could not connect to a remote GitHub repository. You are working with a local git repository only.", call. = FALSE)})
+    }, error = function(e){warning("Could not connect to a remote 'GitHub' repository. You are working with a local 'Git' repository only.", call. = FALSE)})
   } else {
-    warning("No valid GitHub address provided. You are working with a local git repository only.", call. = FALSE)
+    warning("No valid 'GitHub' address provided. You are working with a local 'Git' repository only.", call. = FALSE)
   }
   if("GCtorture" %in% ls()) rm("GCtorture")
 }

### 2020-05-04 R/zzz.R

@@ -1,9 +1,9 @@
 # @importFrom rstudioapi versionInfo
 .onAttach <- function(libname, pkgname) {
   #correct_version <- versionInfo()$version >= "1.1.28"
-  print_message <- paste0("Welcome to WORCS: Worflow for Open Reproducible Code in Science. Please read the tutorial before using this package, and consider citing it. ",
+  print_message <- paste0("Welcome to WORCS: Worflow for Open Reproducible Code in Science. Please read the tutorial before using this package, and consider citing it: Van Lissa, Brandmaier, Brinkman, & Vreede (2020) <doi:10.17605/OSF.IO/ZCVBS>",
                           ifelse(has_git(), "",
-                                 "Rstudio is not yet connected to Git. You will not be able to use the worcs package yet.")
+                                 "'Rstudio' is not yet connected to 'Git'. You will not be able to use the 'worcs' package yet.")
                           #, ifelse(correct_version, "Haaai", "Rstudio version 1.1.28 or higher is required to use the worcs package.")
   )
   packageStartupMessage(print_message)

### 2020-05-05 README.Rmd

@@
 ``` r
-install.packages("devtools")
-devtools::install_github("crsh/papaja", dependencies = TRUE)
-devtools::install_github("cjvanlissa/worcs", dependencies = TRUE)
+if(!requireNamespace("remotes"))install.packages("remotes")
+remotes::install_github("crsh/papaja", dependencies = TRUE, update = "never")
+remotes::install_github("cjvanlissa/worcs", dependencies = TRUE, update = "never")
 tinytex::install_tinytex()

### 2020-05-05 README.Rmd
Add remotes to Suggests, readme, and vignette Add CRAN version to vignette in preparation for release

--- 2020-04-26 README.Rmd
@@
 ``` r
-install.packages("devtools")
-devtools::install_github("crsh/papaja", dependencies = TRUE)
-devtools::install_github("cjvanlissa/worcs", dependencies = TRUE)
+if(!requireNamespace("remotes"))install.packages("remotes")
+remotes::install_github("cjvanlissa/worcs", dependencies = TRUE, update = "never")
+remotes::install_github("crsh/papaja", dependencies = TRUE, update = "never")
 tinytex::install_tinytex()

### 2020-05-07 paper/Manuscript.Rmd
Fix tests and synthetic

@@
     email         : "c.j.vanlissa@uu.nl"
+  - name          : "Andreas M. Brandmaier"
+    affiliation   : "3,4"
   - name          : "Loek Brinkman"
-    affiliation   : "2,3"
+    affiliation   : "2,5"
 
@@
   - id            : "3"
+    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
+  - id            : "4"
+    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
+  - id            : "5"
     institution   : "UMC Utrecht"
@@
 The social sciences are amidst a paradigm shift towards open science.
-Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [e.g., @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
+Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [as discussed by @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
 However, open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012].
@@
 Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5.
-Peikert and Brandmaier aim to achieve strict computational reproducibility by conducting analyses on a virtual machine, that can be stored like a sort of time capsule and reinstated on different systems, instead of on a private computer.
+Peikert and Brandmaier's approach overlaps, to some extent, with the solution proposed here, but focuses on achieving strict computational reproducibility by conducting analyses on a virtual machine, that can be stored like a sort of time capsule and identically reinstated on different systems.
 The workflow is excellent, but difficult to implement for non-programmers.
@@
 This should be sufficient for most users, but can also be combined with Peikert and Brandmaier's approach if strict computational reproducibility is essential.
-WORCS thus addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines.
-Still, it can be combined with existing approaches.
+WORCS addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines, while being compatible with existing approaches.
 
@@
 
-WORCS relies on a few free, open source software solutions that, when combined, facilitate meeting the TOP-guidelines related to citation, sharing, and preregistration.
-The first tool is *dynamic document generation*.
+WORCS relies on several free, open source software solutions.
+The first of these is *dynamic document generation*.
 Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses.
@@
 
-The second tool is *version control*: Maintaining an indelible log of every change to all project files.
+The second solution is *version control*: Maintaining an indelible log of every change to all project files.
 Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax [see also @blischakQuickIntroductionVersion2016].
@@
 
-The third tool is *dependency management*: Keeping track of exactly what software was used to conduct the analyses.
+The third solution is *dependency management*: Keeping track of exactly what software was used to conduct the analyses.
 <!-- Being able to replicate analyses requires not only access to the data and analysis code, but also to the software used to evaluate that code.--> At first glance, it might seem sufficient to state that analyses were conducted in *Program X*.
@@
 This means that any analyses that use random numbers, such as bootstrapping, cannot be strictly replicated from previous versions.
+Or, for example, R-version 4.0.0 changed a default parameter that affects how data are loaded from files.
 
@@
 If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine.
-This is quite effective, but very difficult to set up.
-<!-- Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project.
+This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data).
+However, it is very difficult to set up.
+<!--For small projects, or novice users || Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project.
 This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space.

### 2020-05-07 R/export.R

@@ -49,7 +49,7 @@ export_project <- function(filename = NULL, open_data = TRUE)
     if (isFALSE(open_data)) {
       hasdata <- endsWith(zip_these, "data.csv")
       if (any(hasdata)) {
-        message("Excluding open data file and generating synthetic data for archive. Ensure that no identifying information is included.")
+        col_message("Excluding open data file and generating synthetic data for archive. Ensure that no identifying information is included.")
         zip_these <- zip_these[!hasdata]
         if(!any(zip_these == "synthetic_data.csv")){
           data <- read.csv(zip_these[endsWith(zip_these, "data.csv")], stringsAsFactors = TRUE)

### 2020-05-07 R/save_load.R

@@ -41,7 +41,7 @@ open_data <- function(data, codebook = TRUE){
 #' old_wd <- getwd()
 #' dir.create(file.path(tempdir(), the_test))
 #' setwd(file.path(tempdir(), the_test))
-#' closed_data(iris[1:10, ], codebook = NULL)
+#' closed_data(iris[1:10, ], codebook = FALSE)
 #' setwd(old_wd)
 #' unlink(file.path(tempdir(), the_test))
 #' @rdname closed_data
@@ -61,35 +61,35 @@ save_data <- function(data, open, codebook = TRUE){
     stop("Argument 'data' must be a data.frame, matrix, or inherit from these classes.")
   }
   if(!file.exists(".gitignore")){
-    message("Could not find .gitignore file. You might not be working from the main project directory. Created .gitignore file in folder: ", getwd())
+    col_message(paste0("Could not find .gitignore file. You might not be working from the main project directory. Created .gitignore file in folder: ", getwd()))
     file.create(".gitignore")
   }
   data <- as.data.frame(data)
   if(codebook){
     make_codebook(data)
   }
-  message("Storing original data in 'data.csv' and updating the checksum in '.worcs'.")
+  col_message("Storing original data in 'data.csv' and updating the checksum in '.worcs'.")
   write.csv(data, "data.csv", row.names = FALSE)
   store_checksum("data.csv")
   gitig <- readLines(".gitignore")
   if(open){
     if(!any(grepl("!data.csv", gitig))){
-      message('Updating ".gitignore".')
+      col_message('Updating ".gitignore".')
       write("!data.csv", file = ".gitignore", append = TRUE)
     }
   } else {
     if(!any(grepl("!synthetic_data.csv", gitig))){
-      message('Updating ".gitignore".')
+      col_message('Updating ".gitignore".')
       write("!synthetic_data.csv", file = ".gitignore", append = TRUE)
     }
-    message("Generating synthetic data for public use. Ensure that no identifying information is included.")
+    col_message("Generating synthetic data for public use. Ensure that no identifying information is included.")
     synth <- synthetic(data, verbose = FALSE)
-    message("Storing synthetic data in 'synthetic_data.csv' and updating the checksum in '.worcs'.")
-    write.csv(synth$syn, "synthetic_data.csv", row.names = FALSE)
[…]

### 2020-05-07 tests/testthat/test-test-closed_data.R

@@ -1,3 +1,59 @@
-test_that("multiplication works", {
-  expect_equal(4+4, 8)
+library(yaml)
+# the_test <- "loaddata"
+# old_wd <- getwd()
+# dir.create(file.path(tempdir(), the_test))
+# setwd(file.path(tempdir(), the_test))
+worcs:::write_worcsfile(".worcs")
+open_data(iris[1:5, ], codebook = FALSE)
+checksums <- read_yaml(".worcs")
+
+test_that(".worcs contains correct checksum", {
+  expect_equivalent(checksums$checksums$data.csv, "576b8c41e688f79720b256ed1420e4e8")
 })
+
+test_that("loading open data works", {
+  expect_error({df <- load_data()}, NA)
+})
+
+df <- load_data()
+
+test_that("loaded data same as original", {
+  expect_equivalent(iris[1:5, ], df)
+})
+
+df <- rbind(df, df[40,])
+write.csv(df, "data.csv", row.names = FALSE)
+
+test_that("loading open data fails when data changed", {
+  expect_error({df <- load_data()})
+})
+
+set.seed(555)
+closed_data(iris, codebook = FALSE)
+checksums <- read_yaml(".worcs")
+tmp <- read.csv("synthetic_data.csv", stringsAsFactors = TRUE)
+
+test_that("synthetic data similar", {
+  expect_equivalent(dim(tmp), dim(iris))
+  expect_equivalent(sapply(iris, class), sapply(tmp, class))
+})
+
+test_that(".worcs contains checksum for synthetic_data.csv", {
+  expect_true(!is.null(checksums$checksums[["synthetic_data.csv"]]))
+})
+
+test_that("loading open data works", {
+  expect_error({df <- load_data()}, NA)
+})
+
+file.remove("data.csv")
+
+df <- load_data()
+
+test_that("loaded synthetic data same as original", {
+  expect_equivalent(tmp, df)
+})
+
+
+# setwd(old_wd)
+# unlink(file.path(tempdir(), the_test))

### 2020-05-07 tests/testthat/test-worcsfile.R

@@ -1,39 +1,0 @@
-library(yaml)
-
-Args <- list(
-  arg1 = "val1", arg2 = list(arg2.1 = "val2.1", arg2.2 = "val2.2"), arg3 = list(arg3.1 = "val3.1", arg3.2 = "val3.2"), arg4 = "hoi"
-)
-arg_list <- c(list(filename = "test.yaml"), Args, list(modify = FALSE))
-do.call(worcs:::write_worcsfile, arg_list)
-
-result <- read_yaml("test.yaml")
-
-test_that("yaml read correctly", {
-  expect_true(all(unlist(Args) == unlist(result)))
-})
-
-worcs:::write_worcsfile("test.yaml", arg4 = "replaced", modify = TRUE)
-result <- read_yaml("test.yaml")
-modified_args <- Args
-modified_args$arg4 <- "replaced"
-
-test_that("can replace one yaml key/value", {
-  expect_true(all(unlist(modified_args) == unlist(result)))
-})
-
-worcs:::write_worcsfile("test.yaml", arg3 = list(arg3.1 = "replaced3.1"), modify = TRUE)
-result <- read_yaml("test.yaml")
-modified_args$arg3$arg3.1 <- "replaced3.1"
-
-test_that("can replace one nested yaml key/value", {
-  expect_true(all(unlist(modified_args) == unlist(result)))
-})
-
-worcs:::write_worcsfile("test.yaml", arg3 = list(arg3.new = "anewvalue"), modify = TRUE)
-result <- read_yaml("test.yaml")
-modified_args$arg3$arg3.new <- "anewvalue"
-
-
-test_that("can append one nested yaml key/value", {
-  expect_true(all(unlist(modified_args) == unlist(result)))
-})

### 2020-05-08 docs/reference/citeall.Rmd
Add tests, update site, bugfixes

---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Optional reference: @reference2020

### 2020-05-08 R/citations.R

@@ -15,18 +15,28 @@
 #' # Only specify it as custom knit function in an 'Rmarkdown' file, like so:
 #' # knit: worcs::cite_all
 #'
-#' library(rmarkdown)
-#' the_test <- "citeall"
-#' old_wd <- getwd()
-#' dir.create(file.path(tempdir(), the_test))
-#' setwd(file.path(tempdir(), "citeall"))
-#' draft("test.Rmd", template = "github_document", package = "rmarkdown",
-#'       create_dir = FALSE, edit = FALSE)
+#' dir_name <- tempdir()
+#' file_name <- file.path(dir_name, "citeall.Rmd")
+# file_name <- tempfile("citeall", fileext = ".Rmd")
+#'
+#' rmarkdown::draft(file_name,
+#'                  template = "github_document",
+#'                  package = "rmarkdown",
+#'                  create_dir = FALSE,
+#'                  edit = FALSE)
 #' write(c("", "Optional reference: @@reference2020"),
-#'       file = "test.Rmd", append = TRUE)
-#' cite_all("test.Rmd", "UTF-8")
-#' setwd(old_wd)
-#' unlink(file.path(tempdir(), the_test))
+#'       file = file_name, append = TRUE)
+#' cite_all(file_name, "UTF-8")
+#' unlink(paste0(normalizePath(tempdir()), "/", dir(tempdir())),
+#'        recursive = TRUE)
+# unlink(list.files(path = dir_name, full.names = TRUE, recursive = TRUE),
+# recursive = TRUE)
+# unlink(c(
+#   file_name,
+#   gsub("\\.Rmd", "\\.md", file_name),
+#   gsub("\\.Rmd", "\\.html", file_name),
+#   gsub("\\.Rmd", "_files", file_name)
+# ), recursive = TRUE)
 cite_all <- function(inputFile, encoding){
   Args <- list(
     inputFile = inputFile,
@@ -53,18 +63,23 @@ cite_all <- function(inputFile, encoding){
 #' # Only specify it as custom knit function in an Rmarkdown file, like so:
 #' # knit: worcs::cite_all
 #'
-#' library(rmarkdown)
-#' the_test <- "citeessential"
-#' old_wd <- getwd()
-#' dir.create(file.path(tempdir(), the_test))
-#' setwd(file.path(tempdir(), "citeessential"))
-#' draft("test.Rmd", template = "github_document", package = "rmarkdown",
-#'       create_dir = FALSE, edit = FALSE)
[…]

### 2020-05-08 R/codebook.R

@@ -5,25 +5,28 @@
 #' paragraph with details on the data collection procedure.
 #' @param data A data.frame for which to create a codebook.
 #' @param render_file Logical. Whether or not to render the document.
+#' @param file_name Character. File name to write the codebook \code{rmarkdown}
+#' file to.
 #' @return \code{Logical}, indicating whether or not the operation was
 #' succesful. This function is mostly called for its side effect of rendering an
 #' \code{rmarkdown} codebook.
 #' @examples
 #' library(rmarkdown)
 #' library(knitr)
-#' the_test <- "codebook"
-#' old_wd <- getwd()
-#' dir.create(file.path(tempdir(), the_test))
-#' setwd(file.path(tempdir(), "citeall"))
-#' make_codebook(iris)
-#' setwd(old_wd)
-#' unlink(file.path(tempdir(), the_test))
+#' file_name <- tempfile("codebook", fileext = ".Rmd")
+#' make_codebook(iris, file_name = file_name)
+#' unlink(c(
+#'   file_name,
+#'   gsub("\\.Rmd", "\\.md", file_name),
+#'   gsub("\\.Rmd", "\\.html", file_name),
+#'   gsub("\\.Rmd", "_files", file_name)
+#' ), recursive = TRUE)
 #' @rdname codebook
 #' @export
 #' @importFrom rmarkdown draft render
 #' @importFrom stats median var
 #' @importFrom utils capture.output
-make_codebook <- function(data, render_file = TRUE){
+make_codebook <- function(data, render_file = TRUE, file_name = "codebook.Rmd"){
   function_success <- TRUE
   data_types <- sapply(data, function(x){paste0(class(x), collapse = ", ")})
   summaries <- descfun(data)
@@ -32,16 +35,16 @@ make_codebook <- function(data, render_file = TRUE){
     type = data_types,
     summaries
   )
-  if(file.exists("codebook.Rmd")){
-    col_message("Removing previous version of 'codebook.Rmd'.")
-    invisible(file.remove("codebook.Rmd"))
+  if(file.exists(file_name)){
+    col_message(paste0("Removing previous version of '", file_name, "'."))
+    invisible(file.remove(file_name))
   }
-  draft("codebook.Rmd",
+  draft(file_name,
                    template = "github_document",
                    package = "rmarkdown",
                    create_dir = FALSE,
                    edit = FALSE)
-  file_contents <- readLines("codebook.Rmd")
[…]

### 2020-05-08 R/export.R

@@ -21,7 +21,7 @@ export_project <- function(filename = NULL, open_data = TRUE)
   # get properties about the project and paths
   base_dir <- normalizePath(".")
   if(!file.exists(".worcs")){
-    message("No '.worcs' file found; not a WORCS project, or the working directory has been changed.")
+    col_message("No '.worcs' file found; not a WORCS project, or the working directory has been changed.", success = FALSE)
     return(invisible(FALSE))
   }
   project_folder <- gsub("^.+\\b(.+?)$", "\\1", base_dir)
@@ -33,7 +33,7 @@ export_project <- function(filename = NULL, open_data = TRUE)
   }
 
   if (!is.character(filename)) {
-    message("Filename must be of type character: ",as.character(filename))
+    col_message(paste0("Filename must be of type character: ",as.character(filename)), success = FALSE)
     return(invisible(FALSE))
   }
 

### 2020-05-08 R/worcs_project.R

@@ -32,6 +32,8 @@
 #' this project. If a valid 'GitHub' repository link is provided, a commit will
 #' be made containing the 'README.md' file, and will be pushed to 'GitHub'.
 #' Default: 'https'. For more information, see <http://github.com/>.
+#' @param verbose Logical. Whether or not to print messages to the console
+#' during project creation. Default: TRUE
 #' @param ... Additional arguments passed to and from functions.
 #' @return No return value. This function is called for its side effects.
 #' @examples
@@ -56,7 +58,7 @@
 #' @importFrom gert git_init git_remote_add git_add git_commit git_push
 #' @importFrom utils installed.packages packageVersion
 # @importFrom renv init
-worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregistration = "COS", add_license = "CC_BY_4.0", use_renv = TRUE, remote_repo = "https", ...) {
+worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregistration = "COS", add_license = "CC_BY_4.0", use_renv = TRUE, remote_repo = "https", verbose = TRUE, ...) {
   # collect inputs
   manuscript <- tolower(manuscript)
   preregistration <- tolower(preregistration)
@@ -70,105 +72,133 @@ worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregist
   if(!use_git){
     message("Could not find a working installation of 'Git', which is required to safeguard the transparency and reproducibility of your project. Please connect 'Git' by following the steps described in this vignette:\n  vignette('setup', package = 'worcs')")
   } else {
+    col_message("Initializing 'Git' repository.")
     git_init(path = path)
   }
 
   # Create .worcs file
-  write_worcsfile(filename = file.path(path, ".worcs"),
-                  worcs_version = as.character(packageVersion("worcs")),
-                  creator = Sys.info()["effective_user"]
-  )
+  tryCatch({
+    write_worcsfile(filename = file.path(path, ".worcs"),
+                    worcs_version = as.character(packageVersion("worcs")),
+                    creator = Sys.info()["effective_user"]
+    )
+    col_message("Writing '.worcs' file.")
+  }, error = function(e){
+    col_message("Writing '.worcs' file.", success = FALSE)
+  })
+
 
   # copy 'resources' folder to path
-  copy_resources(which_files = c(
-    "README.md",
-    "prepare_data.R"
-  ), path = path)
+  tryCatch({
+    copy_resources(which_files = c(
[…]

### 2020-05-08 tests/testthat/test-citations.R

@@ -1,27 +1,34 @@
 library(rmarkdown)
-# the_test <- "citeall"
-# old_wd <- getwd()
-# dir.create(file.path(tempdir(), the_test))
-# setwd(file.path(tempdir(), "citeall"))
-draft("test.Rmd", template = "github_document", package = "rmarkdown",
-      create_dir = FALSE, edit = FALSE)
 
-write(c("", "Optional reference: @@reference2020"),
-      file = "test.Rmd", append = TRUE)
-cite_all("test.Rmd", "UTF-8")
+test_that("documents with spaces in names can be rendered", {
 
-contents <- readLines("test.md")
-test_that("Reference is retained correctly", {
+  skip_on_cran()
+
+  the_test <- "citeall"
+  dir_name <- file.path(tempdir(), the_test)
+  file_name <- file.path(dir_name, "test.Rmd")
+  old_wd <- getwd()
+  dir.create(dir_name)
+  on.exit(unlink(dir_name, recursive = TRUE), add = TRUE)
+
+  setwd(dir_name)
+  draft("test.Rmd", template = "github_document", package = "rmarkdown",
+        create_dir = FALSE, edit = FALSE)
+
+  write(c("", "Optional reference: @@reference2020"),
+        file = "test.Rmd", append = TRUE)
+  cite_all("test.Rmd", "UTF-8")
+
+  contents <- readLines("test.md")
+
   expect_true(any(contents == "Optional reference: @reference2020"))
-})
 
-cite_essential("test.Rmd", "UTF-8")
+  cite_essential("test.Rmd", "UTF-8")
 
-contents <- readLines("test.md")
-test_that("Reference is retained correctly", {
+  contents <- readLines("test.md")
+
   expect_true(!any(contents == "Optional reference: @reference2020"))
   expect_true(any(contents == "Optional reference:"))
-})
 
-# setwd(old_wd)
-# unlink(file.path(tempdir(), the_test))
+  setwd(old_wd)
+})

### 2020-05-08 tests/testthat/test-codebook.R

@@ -1,0 +1,46 @@
+test_that("codebook works for iris", {
+
+  the_test <- "codebook"
+  old_wd <- getwd()
+  test_dir <- file.path(tempdir(), the_test)
+  dir.create(test_dir)
+  setwd(test_dir)
+
+  on.exit(unlink(test_dir, recursive = TRUE), add = TRUE)
+
+  worcs::make_codebook(iris)
+
+  expect_true(file.exists("codebook.Rmd"))
+  expect_true(file.exists("codebook.md"))
+
+  contents <- readLines("codebook.md")
+
+  expect_true(any(contents == "The data contains 150 cases and 5 variables."))
+
+  setwd(old_wd)
+})
+
+
+test_that("codebook works for PlantGrowth with missings", {
+  df <- PlantGrowth
+  df[c(7:8), 1] <- NA
+  df[c(1, 25), 2] <- NA
+  the_test <- "codebook"
+  old_wd <- getwd()
+  test_dir <- file.path(tempdir(), the_test)
+  dir.create(test_dir)
+  setwd(test_dir)
+
+  on.exit(unlink(test_dir, recursive = TRUE), add = TRUE)
+
+  worcs::make_codebook(df)
+
+  expect_true(file.exists("codebook.Rmd"))
+  expect_true(file.exists("codebook.md"))
+
+  contents <- readLines("codebook.md")
+
+  expect_true(any(contents == "The data contains 30 cases and 2 variables."))
+
+  setwd(old_wd)
+})

### 2020-05-08 tests/testthat/test-export_project.R

@@ -34,7 +34,7 @@ test_that("export returned true", {
 })
 
 test_that("exported worcs_project exists", {
-  expect_true("export_test.zip" %in% list.files("../"))
+  expect_true("export_test.zip" %in% list.files())
 })
 }
 # setwd(old_wd)

### 2020-05-08 tests/testthat/test-worcs_project.R

@@ -1,0 +1,29 @@
+library(gert)
+test_that("worcs project can be generated", {
+
+  the_test <- "project"
+  old_wd <- getwd()
+  test_dir <- file.path(tempdir(), the_test)
+  dir.create(test_dir)
+  setwd(test_dir)
+
+  on.exit(unlink(test_dir, recursive = TRUE), add = TRUE)
+
+  worcs_project(
+    path = "worcs_project",
+    manuscript = "github_document",
+    preregistration = "COS",
+    add_license = "CC_BY_4.0",
+    use_renv = FALSE,
+    remote_repo = "bla"
+  )
+  list.files(file.path(test_dir, "worcs_project"))
+  expect_true(file.exists("worcs_project/.worcs"))
+  expect_true(file.exists("worcs_project/LICENSE"))
+  expect_true(file.exists("worcs_project/README.md"))
+  expect_true(file.exists("worcs_project/preregistration.Rmd"))
+  expect_true(file.exists("worcs_project/manuscript/manuscript.Rmd"))
+  expect_error(git_status("worcs_project"), NA)
+
+  setwd(old_wd)
+})

### 2020-05-08 paper/Manuscript.Rmd
Marijn changes

@@
 A project version controlled with Git is called a "Repository", or Repo.
+It is also possible to prevent some files from being version controlled - such as privacy sensitive data.
+The `.gitignore` file is a list of files that should not be version-controlled, and thus exist only on the user's private computer.
 
@@
 Researchers can overcome this impediment by preparing two versions of the manuscript: One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
-The print version should reference the online version in the print version, so interested readers can find the comprehensive reference list.
+The print version should reference the online version, so interested readers can find the comprehensive reference list.
 Step 20 of the WORCS procedure suggests uploading the online version to a preprint server.
@@
 It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication.
-Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference.
+Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference.
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
@@
 For more options, see the [Rmarkdown cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html).
-3. To indicate a *non-essential* citation, mark it with a double at-symbol: `@nonessential2020`.
+3. To indicate a *non-essential* citation, mark it with a double at-symbol: `@@nonessential2020`.
 4. When Knitting the document, adapt the `knit` command in the YAML header.  
@@
 Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns.
-In some cases, it is difficult or impossible to share data.
+For example, the GDPR prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU.
+Different rules may apply to anonymized or pseudonimized data.
+Before sharing any data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.
 
@@
 The first solution is publishing a *checksum* of the raw data upon collection.
-A checksum is a 32-character summary of a file of any size.
+A checksum represents the contents of a file in a 32-character summary.
 Any change to the underlying file results in a different checksum.
@@
 
-* `open_data(data)`: This function takes a data object (a `data.frame` or `matrix`), writes it to a file called `data.csv`, updates the `.gitignore` file to stop ignoring this specific `.csv` file, and writes a checksum for this data file to `checksums.csv`.
-The data can now be committed and pushed to GitHub, where they will be open once the GitHub repository is made public.
-* `closed_data(data)`: This function generates a local file called `data.csv`, and a synthetic dataset called `synthetic_data.csv`, updates the `.gitignore` file to stop ignoring `synthetic_data.csv` file, and writes a checksum for the real data file to `checksums.csv`.
-The synthetic data can now be committed and pushed to GitHub, and the real data will be ignored by Git.
+* `open_data(data)`: This function stores the data object (a `data.frame` or `matrix`) in a spreadsheet file called `data.csv`, stores a checksum for the file, and creates a codebook.
+All of these files should now be (manually) added to the 'Git' repository, committed, and pushed to GitHub.
+Once the GitHub repository is made public, these data will be open to the public.
+* `closed_data(data)`: This function stores the data in a local file called `data.csv`, which is added to the `.gitignore` file so it cannot be accidentally added to the 'Git' repository.
+The function also creates a synthetic dataset called `synthetic_data.csv`.
+The user should commit all changed files and push them to GitHub.
+Once the GitHub repository is made public, people will have access to a checksum for the raw data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
 
-Typically, the last line of the `prepare_data.R` file will thus be `open_data()` or `closed_data()`.
+Some users may intend to share their data through a platform with restricted access (like an institutional repository), but not publically through GitHub.
+In this case, it is recommended to use `closed_data()`, and to manually upload the original data `data.csv` to the restricted access platform.
+If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the GitHub repository.
+
+The `prepare_data.R` file describes the data wrangling process to derive a spreadsheet ready for analysis.
+The final line of this file will often be `open_data()` or `closed_data()`.
 The code for analysis of the data, on the other hand, will typically begin with `load_data()`.
-This function loads the real data if it is available (i.e., the file `data.csv` exists, and its checksum is correct), and otherwise, loads the synthetic data.
+This function loads the real data if it is available (i.e., the file `data.csv` exists, and the checksum is correct), and otherwise, loads the synthetic data.
 This will have the effect that third party users who copy the GitHub repository will automatically load the synthetic dataset, whereas the study authors, who have the real data stored locally, will automatically load the real data.
@@
 
-1. Create a new Private repository on GitHub, and copy the repository link to the clipboard. The link will look like https://github.com/yourname/yourrepo.git
+1. Create a new Private repository on GitHub. Do not change any of the default options. Copy the repository link to the clipboard. This link will look like https://github.com/yourname/yourrepo.git
 2. In RStudio, click File > New Project > New directory > WORCS Project Template
@@
 11. Regularly commit your progress to the Git repository; ideally, after completing each small and clearly defined task. Use informative commit messages. Push the commits to GitHub.
-12. Cite essential references with one at-symbol (`[@essentialref2020]`), and non-essential references with a double at-symbol (`[@nonessential2020]`).
+12. Cite essential references with one at-symbol (`[@essentialref2020]`), and non-essential references with a double at-symbol (`[@@nonessential2020]`).
 

### 2020-05-08 paper/Manuscript.Rmd
Various bug fixes

@@
 
-author: 
-  - name          : "Caspar J. van Lissa"
-    affiliation   : "1,2"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
-    email         : "c.j.vanlissa@uu.nl"
-  - name          : "Andreas M. Brandmaier"
-    affiliation   : "3,4"
-  - name          : "Loek Brinkman"
-    affiliation   : "2,5"
+author: #"`r source('authors_yaml.R')`"
+  - name: "Caspar J. Van Lissa"
+    institution: "1,2"
+    corresponding: yes
+    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
+    email: "c.j.vanlissa@uu.nl"
+  - name: "Andreas M. Brandmaier"
+    institution: "3,4"
+  - name: "Loek Brinkman"
+    institution: "2,5"
+  - name: "Aaron Peikert"
+    institution: "6"
+  - name: "Barbara Vreede"
+    institution: "7"
 
-affiliation:
-  - id            : "1"
-    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
-  - id            : "2"
-    institution   : "Open Science Community Utrecht"
-  - id            : "3"
-    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
-  - id            : "4"
-    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
-  - id            : "5"
-    institution   : "UMC Utrecht"
+affiliation: #"`r source('affiliation_yaml.R')`"
+  - id: "1"
+    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
+  - id: "2"
+    institution: "Open Science Community Utrecht"
+  - id: "3"
+    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
+  - id: "4"
+    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
+  - id: "5"
+    institution: "UMC Utrecht"
+  - id: "6"
+    institution: "Test"
+  - id: "7"
+    institution: "University Library Utrecht"
 

### 2020-05-08 R/codebook.R

@@ -1,12 +1,28 @@
 #' @title Create codebook for a dataset
-#' @description Creates a rudimentary 'Rmarkdown' codebook for a dataset, and
-#' renders it to markdown for 'GitHub'. Users can customize the 'Rmarkdown'
-#' document, for example, by adding a column with variable descriptions, or a
-#' paragraph with details on the data collection procedure.
+#' @description Creates a codebook for a dataset in 'Rmarkdown' format, and
+#' renders it to 'markdown' for 'GitHub'. Users can customize the 'Rmarkdown'
+#' document and re-knit it, for example, to add a paragraph with details on
+#' the data collection procedures. The variable descriptives are stored in
+#' a \code{.csv} file, which can be edited in 'R' or a spreadsheet program.
+#' Columns can be appended, and we encourage users to complete at least the
+#' following two columns in this file:
+#' \itemize{
+#' \item{category} Describe the type of variable in this column. For example:
+#' "morality".
+#' \item{description} Provide a plain-text description of the variable. For
+#' example, the full text of a questionnaire item: "People should be willing to
+#' do anything to help a member of their family".
+#' }
+#' Re-knitting the 'Rmarkdown' file (using \code{\link[rmarkdown]{render}}) will
+#' transfer these changes to the 'markdown' file for 'GitHub'.
 #' @param data A data.frame for which to create a codebook.
 #' @param render_file Logical. Whether or not to render the document.
 #' @param file_name Character. File name to write the codebook \code{rmarkdown}
 #' file to.
+#' @param csv_file Character. File name to write the codebook \code{rmarkdown}
+#' file to. By default, uses the filename stem of the \code{file_name} argument.
+#' Set to \code{NULL} to write the codebook only to the 'Rmarkdown' file, and
+#' not to \code{.csv}.
 #' @return \code{Logical}, indicating whether or not the operation was
 #' succesful. This function is mostly called for its side effect of rendering an
 #' \code{rmarkdown} codebook.
@@ -14,7 +30,7 @@
 #' library(rmarkdown)
 #' library(knitr)
 #' file_name <- tempfile("codebook", fileext = ".Rmd")
-#' make_codebook(iris, file_name = file_name)
+#' make_codebook(iris, file_name = file_name, csv_file = NULL)
 #' unlink(c(
 #'   file_name,
 #'   gsub("\\.Rmd", "\\.md", file_name),
@@ -26,100 +42,155 @@
[…]

### 2020-05-08 tests/testthat/test-codebook.R

@@ -1,8 +1,7 @@
 test_that("codebook works for iris", {
 
-  the_test <- "codebook"
   old_wd <- getwd()
-  test_dir <- file.path(tempdir(), the_test)
+  test_dir <- file.path(tempdir(), "codebook")
   dir.create(test_dir)
   setwd(test_dir)
 
@@ -10,7 +9,7 @@ test_that("codebook works for iris", {
 
   expect_true(file.exists("codebook.Rmd"))
   expect_true(file.exists("codebook.md"))
-
+  expect_true(file.exists("codebook.csv"))
   contents <- readLines("codebook.md")
 
   expect_true(any(contents == "The data contains 150 cases and 5 variables."))
@@ -19,14 +18,14 @@ test_that("codebook works for iris", {
   unlink(test_dir, recursive = TRUE)
 })
 
-
+if(FALSE){
 test_that("codebook works for PlantGrowth with missings", {
-  df <- PlantGrowth
+  df <- CO2
   df[c(7:8), 1] <- NA
   df[c(1, 25), 2] <- NA
-  the_test <- "codebook"
+
   old_wd <- getwd()
-  test_dir <- file.path(tempdir(), the_test)
+  test_dir <- file.path(tempdir(), "codebook")
   dir.create(test_dir)
   setwd(test_dir)
 
@@ -37,8 +36,9 @@ test_that("codebook works for PlantGrowth with missings", {
 
   contents <- readLines("codebook.md")
 
-  expect_true(any(contents == "The data contains 30 cases and 2 variables."))
+  expect_true(any(contents == "The data contains 84 cases and 5 variables.")) #"The data contains 30 cases and 2 variables."))
 
   setwd(old_wd)
   unlink(test_dir, recursive = TRUE)
 })
+}

### 2020-05-08 tests/testthat/test-descriptives.R

@@ -1,16 +1,16 @@
-tmp <- descfun(iris)
+tmp <- descriptives(iris)
 
-test_that("descfun creates correct output", {
+test_that("descriptives creates correct output", {
   expect_true(all(c("missing", "unique", "mean", "median", "mode", "mode_value",
-    "sd", "V", "min", "max", "range", "skew", "skew_2se", "kurt",
+    "sd", "v", "min", "max", "range", "skew", "skew_2se", "kurt",
     "kurt_2se") %in% names(tmp)))
   num_vars <- sapply(iris, class) == "numeric"
   num_cols <- c("missing", "unique", "mean", "median", "mode", "sd", "min", "max", "range", "skew", "skew_2se", "kurt", "kurt_2se")
   expect_true(all(!is.na(tmp[num_vars, num_cols])))
 
-  fac_cols <- c("missing", "unique", "mode", "mode_value", "V")
+  fac_cols <- c("missing", "unique", "mode", "mode_value", "v")
   expect_true(all(!is.na(tmp[!num_vars, fac_cols])))
 
-  expect_true(all(is.na(tmp[num_vars, c("mode_value", "V")])))
+  expect_true(all(is.na(tmp[num_vars, c("mode_value", "v")])))
   expect_true(all(is.na(tmp[!num_vars, c("mean", "median", "sd", "min", "max", "range", "skew", "skew_2se", "kurt", "kurt_2se")])))
 })

### 2020-05-10 paper/Manuscript.Rmd
update aarons affiliations

@@
   - name: "Aaron Peikert"
-    institution: "6"
+    institution: "3,6"
   - name: "Barbara Vreede"
@@
   - id: "6"
-    institution: "Test"
+    institution: "Humboldt-Universität zu Berlin"
   - id: "7"

### 2020-05-11 paper/Manuscript.Rmd

@@
 
-author: #"`r source('authors_yaml.R')`"
+author:
   - name: "Caspar J. Van Lissa"
@@
 
-affiliation: #"`r source('affiliation_yaml.R')`"
+affiliation:
   - id: "1"
@@
   - id: "6"
-    institution: "Humboldt-Universität zu Berlin"
+    institution: "Humboldt-Universitaet zu Berlin"
   - id: "7"

### 2020-05-12 paper/Manuscript.Rmd
highlight compatibility to Peikert & Brandmaier

@@
 Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5.
-Peikert and Brandmaier's approach overlaps, to some extent, with the solution proposed here, but focuses on achieving strict computational reproducibility by conducting analyses on a virtual machine, that can be stored like a sort of time capsule and identically reinstated on different systems.
-The workflow is excellent, but difficult to implement for non-programmers.
-WORCS instead uses a much more lightweight approach to safeguard computational reproducibility under *most circumstances*.
-This should be sufficient for most users, but can also be combined with Peikert and Brandmaier's approach if strict computational reproducibility is essential.
+Peikert and Brandmaier's workflow builds upon the same general principles making it fully compatible with the solution proposed here.
+Their workflow aims to encompass even the most sophisticated analytic strategies and is hence relatively complex itself.
+WORCS greatly simplifies matters by assuming that no other software then R Packages are needed <!-- difference docker -->and that that a single data file is sufficient <!-- difference make -->.
+If both assumptions are met, and they routinely are, this eschews much overhead of Peikert and Brandmaier's workflow resulting in a much more lightweight approach to safeguard computational reproducibility under *most circumstances*.
 WORCS addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines, while being compatible with existing approaches.
@@
 These solutions differ in user-friendliness and effectiveness.
-If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine.
+If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine, that can be stored like a sort of time capsule and identically reinstated on different systems.
 This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data).
-However, it is very difficult to set up.
 <!--For small projects, or novice users || Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project.
@@
     a. Paste the GitHub Repository address in the textbox
-    b.
-Keep the checkbox for `renv` checked if you want to document all dependencies (recommended).
-The only reason not to do this is if you have dependencies outside of the R environment, or if you are using a more comprehensive solution for computational reproducibility, e.g., @peikertReproducibleDataAnalysis2019.
+    b. Keep the checkbox for `renv` checked if you want to document all R-package dependencies, and have them automatically loaded on any system (recommended).
     c. Select a preregistration template

### 2020-05-12 R/citations.R
all checks passing

@@ -74,21 +74,21 @@ comprehensive_cite <- function(input, encoding = "UTF-8", ..., citeall = TRUE) {
 
 cleancitations <- function(text){
   out <- paste0(c(text, " onzin"), collapse = "\n")
-  out <- as.list(strsplit(out, "\\[")[[1]])
+  out <- as.list(strsplit(out, "[", fixed = TRUE)[[1]])
   out <- lapply(out, function(x){
-    if(grepl("@@", x)){
-      ref_sec <- unlist(strsplit(x, split = "\\]"))
+    if(grepl("@@", x, fixed = TRUE)){
+      ref_sec <- unlist(strsplit(x, split = "]", fixed = TRUE))
       if(length(ref_sec) > 1){
-        if(grepl("@@", ref_sec[1])){
+        if(grepl("@@", ref_sec[1], fixed = TRUE)){
           each_ref_sec <- unlist(strsplit(ref_sec[1], split = ";"))
-          ref_sec[1] <- paste0(each_ref_sec[!grepl("@@", each_ref_sec)], collapse = ";")
+          ref_sec[1] <- paste0(each_ref_sec[!grepl("@@", each_ref_sec, fixed = TRUE)], collapse = ";")
           if(ref_sec[1] == "") ref_sec[1] <- "XXXXXDELETEMEXXXXX"
         }
-        if(grepl("@@", ref_sec[2])){
-          ref_sec[2] <- gsub(" -?@@.+? ", " ", ref_sec[2])
+        if(grepl("@@", ref_sec[2], fixed = TRUE)){
+          ref_sec[2] <- gsub("\\s{0,1}-?@@.+?\\b\\s{0,1}", " ", ref_sec[2])
         }
       } else {
-        ref_sec[1] <- gsub(" -?@@.+? ", " ", ref_sec[1])
+        ref_sec[1] <- gsub("\\s{0,1}-?@@.+?\\b\\s{0,1}", " ", ref_sec[1])
       }
       paste0(ref_sec, collapse = "]")
     } else {

### 2020-05-12 R/export.R

@@ -15,7 +15,7 @@
 #' @return Logical, indicating the success of the operation. This function is
 #' called for its side effect of creating a \code{.zip} file.
 #' @examples
-#' export_project()
+#' export_project(worcs_directory = tempdir())
 #' @importFrom utils tail zip
 #' @importFrom gert git_ls
 #' @export
@@ -25,28 +25,31 @@ export_project <- function(zipfile = NULL, worcs_directory = ".", open_data = TR
   #worcs_directory <- normalizePath(worcs_directory)
   worcsfile <- tryCatch(read_yaml(file.path(worcs_directory, ".worcs")), error = function(e){
     col_message("No '.worcs' file found; not a WORCS project, or the working directory has been changed.", success = FALSE)
-    return(invisible(FALSE))
+    FALSE
   })
+  if(isFALSE(worcsfile)) return(invisible(FALSE))
 
   zip_these <- tryCatch({
     git_ls(repo = worcs_directory)$path
   }, error = function(e){
     col_message("Could not find 'Git' repository.", success = FALSE)
-    return(invisible(FALSE))
+    FALSE
   })
+  if(isFALSE(zip_these)) return(invisible(FALSE))
   project_folder <- basename(normalizePath(worcs_directory))
 
   # if no zipfile is given, export to a zip file with
   # the name of the project folder
-  tryCatch({
+  if(isFALSE(tryCatch({
     if(is.null(zipfile)) {
       zipfile <- file.path(dirname(normalizePath(worcs_directory)), paste0(project_folder, ".zip"))
     }
     if(!is.character(zipfile)) stop()
   }, error = function(e){
     col_message("Could not create zipfile.", success = FALSE)
-    return(invisible(FALSE))
-  })
+    FALSE
+  }))) return(invisible(FALSE))
+
   if (file.exists(zipfile)) {
     stop("Could not write to '", zipfile, "' because the file already exists.")
     return(invisible(FALSE))
@@ -54,23 +57,22 @@ export_project <- function(zipfile = NULL, worcs_directory = ".", open_data = TR
 
   # Use this to decide which files to ZIP, but always add data.csv
   # if the user specifies open_data = TRUE
-  data_original <- data_synthetic <- NULL
   if(!is.null(worcsfile[["data"]])){
     data_original <- names(worcsfile[["data"]])
[…]

### 2020-05-12 R/misc.R

@@ -46,7 +46,7 @@ write_as_utf <- function(x, con, append = FALSE, encoding = "UTF-8", ...) {
 }
 
 read_as_utf <- function(..., encoding = "UTF-8"){
-  Args <- as.list(match.call()[-1])
+  Args <- list(...)
   if(is.null(Args[["encoding"]])) Args[["encoding"]] <- encoding
   do.call(readLines, Args)
 }

### 2020-05-12 R/worcs_project.R

@@ -47,7 +47,7 @@ recommend_data <- c('library("worcs")',
 #' the_test <- "worcs_template"
 #' old_wd <- getwd()
 #' dir.create(file.path(tempdir(), the_test))
-#' do.call(git_signature, worcs:::get_credentials())
+#' do.call(git_credentials, worcs:::get_credentials())
 #' worcs_project(file.path(tempdir(), the_test, "worcs_project"),
 #'               manuscript = "github_document",
 #'               preregistration = "None",

### 2020-05-12 tests/testthat/test-export_project.R

@@ -12,7 +12,7 @@
 #   }
 # }
 
-do.call(git_signature, worcs:::get_signature)
+do.call(git_credentials, worcs:::get_credentials())
 
 testdir <- file.path(tempdir(), "export")
 testzip <- tempfile(fileext = ".zip")
@@ -28,6 +28,10 @@ suppressWarnings(
   )
 )
 
+test_that("worcs_project created successfully", {
+  expect_true(file.exists(file.path(testdir, ".worcs")))
+})
+
 result <- export_project(zipfile = testzip, worcs_directory = testdir, open_data = FALSE)
 
 test_that("export returned true", {
@@ -39,4 +43,4 @@ test_that("exported worcs_project exists", {
 })
 
 # setwd(old_wd)
-#unlink(c(file.path(testdir, "export.zip"), testdir), recursive = TRUE)
+#unlink(c(testzip, testdir), recursive = TRUE)

### 2020-05-12 README.Rmd
spell check for CRAN, rename functions

--- 2020-04-26 README.Rmd
@@
   "README.Rmd", "R-markdown source for readme.md", "human editable",
-  "worcs.Rproj", "Rstudio project file", "do not edit",
+  "worcs.Rproj", "RStudio project file", "do not edit",
   "docs/", "Package website", "machine-written",
-  "inst/", "Rstudio project template files ", "human editable",
+  "inst/", "RStudio project template files ", "human editable",
   "man/", "R-package documentation", "do not edit",
@@
 ``` r
-install.packages("devtools")
-devtools::install_github("crsh/papaja", dependencies = TRUE)
-devtools::install_github("cjvanlissa/worcs", dependencies = TRUE)
+if(!requireNamespace("remotes"))install.packages("remotes")
+remotes::install_github("cjvanlissa/worcs", dependencies = TRUE, update = "never")
+remotes::install_github("crsh/papaja", dependencies = TRUE, update = "never")
 tinytex::install_tinytex()
-worcs::git_credentials("your_name", "your_email")
+worcs::git_user("your_name", "your_email")
 ```

### 2020-05-12 paper/notes.Rmd
spell check for CRAN, rename functions

@@
 
-It is worth mentioning the free, open-source reference manager Zotero, because it can be used as easily in conjunction with Rstudio as with Word; the plugin BetterBibTex can export continuously updated `.bib` files for citation in Rmarkdown.
+It is worth mentioning the free, open-source reference manager Zotero, because it can be used as easily in conjunction with RStudio as with Word; the plugin BetterBibTex can export continuously updated `.bib` files for citation in Rmarkdown.
 Finally, it is worth noting that, as of June 2019, Zotero automatically flags retracted papers in your database.
@@
 Sharing code is essential for computational reproducibility.
-The easiest way to share fully reproducible code is to work with an Rstudio project that is version controlled using Git, combined with the `renv` package to ensure all package dependencies are tracked.
+The easiest way to share fully reproducible code is to work with an RStudio project that is version controlled using Git, combined with the `renv` package to ensure all package dependencies are tracked.
 Such a project can be uploaded ("pushed") wholesale to a code repository, such as GitHub.

### 2020-05-12 R/citations.R

@@ -8,10 +8,10 @@
 #' @export
 #' @return Returns \code{NULL} invisibly. This
 #' function is called for its side effect of rendering an
-#' \code{rmarkdown} file.
+#' 'R Markdown' file.
 #' @examples
 #' # NOTE: Do not use this function interactively, as in the example below.
-#' # Only specify it as custom knit function in an 'Rmarkdown' file, like so:
+#' # Only specify it as custom knit function in an 'R Markdown' file, like so:
 #' # knit: worcs::cite_all
 #'
 #' file_name <- file.path(tempdir(), "citeall.Rmd")
@@ -37,10 +37,10 @@ cite_all <- function(...){
 #' @export
 #' @return Returns \code{NULL} invisibly. This
 #' function is called for its side effect of rendering an
-#' \code{rmarkdown} file.
+#' 'R Markdown' file.
 #' @examples
 #' # NOTE: Do not use this function interactively, as in the example below.
-#' # Only specify it as custom knit function in an Rmarkdown file, like so:
+#' # Only specify it as custom knit function in an R Markdown file, like so:
 #' # knit: worcs::cite_all
 #'
 #' file_name <- tempfile("citeessential", fileext = ".Rmd")

### 2020-05-12 R/codebook.R

@@ -1,6 +1,6 @@
 #' @title Create codebook for a dataset
-#' @description Creates a codebook for a dataset in 'Rmarkdown' format, and
-#' renders it to 'markdown' for 'GitHub'. Users can customize the 'Rmarkdown'
+#' @description Creates a codebook for a dataset in 'R Markdown' format, and
+#' renders it to 'markdown' for 'GitHub'. Users can customize the 'R Markdown'
 #' document and re-knit it, for example, to add a paragraph with details on
 #' the data collection procedures. The variable descriptives are stored in
 #' a \code{.csv} file, which can be edited in 'R' or a spreadsheet program.
@@ -13,7 +13,7 @@
 #' example, the full text of a questionnaire item: "People should be willing to
 #' do anything to help a member of their family".
 #' }
-#' Re-knitting the 'Rmarkdown' file (using \code{\link[rmarkdown]{render}}) will
+#' Re-knitting the 'R Markdown' file (using \code{\link[rmarkdown]{render}}) will
 #' transfer these changes to the 'markdown' file for 'GitHub'.
 #' @param data A data.frame for which to create a codebook.
 #' @param render_file Logical. Whether or not to render the document.
@@ -21,11 +21,11 @@
 #' file to.
 #' @param csv_file Character. File name to write the codebook \code{rmarkdown}
 #' file to. By default, uses the filename stem of the \code{filename} argument.
-#' Set to \code{NULL} to write the codebook only to the 'Rmarkdown' file, and
+#' Set to \code{NULL} to write the codebook only to the 'R Markdown' file, and
 #' not to \code{.csv}.
 #' @return \code{Logical}, indicating whether or not the operation was
-#' succesful. This function is mostly called for its side effect of rendering an
-#' \code{rmarkdown} codebook.
+#' successful. This function is mostly called for its side effect of rendering
+#' an 'R Markdown' codebook.
 #' @examples
 #' library(rmarkdown)
 #' library(knitr)

### 2020-05-12 R/github.R

@@ -17,11 +17,11 @@ has_git <- function(){
 #' @param email Character. The email address you want to use with 'Git'.
 #' @return No return value. This function is called for its side effects.
 #' @examples
-#' git_credentials("myname", "my@email.com")
-#' @rdname git_credentials
+#' git_user("myname", "my@email.com")
+#' @rdname git_user
 #' @export
 #' @importFrom gert git_config_global_set
-git_credentials <- function(name, email){
+git_user <- function(name, email){
   invisible(
     tryCatch({
       do.call(git_config_global_set, list(
@@ -38,7 +38,7 @@ git_credentials <- function(name, email){
 }
 
 #' @importFrom gert git_config_global
-get_credentials <- function(){
+get_user <- function(){
   Args <- list(
     name = "yourname",
     email = "yourname@email.com"

### 2020-05-12 R/save_load.R

@@ -1,11 +1,11 @@
 #' @title Use open data in WORCS project
 #' @description This function saves a data.frame as a \code{.csv} file (using
 #' \code{\link[utils]{write.csv}}), stores a checksum in '.worcs',
-#' and amends the .gitignore file to exclude \code{filename}.
+#' and amends the \code{.gitignore} file to exclude \code{filename}.
 #' @param data A data.frame to save.
 #' @param filename Character, naming the file data should be written to.
 #' @param codebook Character, naming the file the codebook should be written to.
-#' An 'rmarkdown' codebook will be created and rendered to
+#' An 'R Markdown' codebook will be created and rendered to
 #' \code{\link[rmarkdown]{github_document}} ('markdown' for 'GitHub').
 #' Defaults to 'codebook.Rmd'. Set to \code{NULL} to avoid creating a codebook.
 #' @return Returns \code{NULL} invisibly. This
@@ -30,9 +30,9 @@ open_data <- function(data, filename = "data.csv", codebook = "codebook.Rmd"){
 #' @title Use closed data in WORCS project
 #' @description This function saves a data.frame as a \code{.csv} file (using
 #' \code{\link[utils]{write.csv}}), stores a checksum in '.worcs',
-#' appends the .gitignore file to exclude \code{filename}, and saves a synthetic
-#' copy of \code{data} for public use. To generate these synthetic data, the
-#' function \code{\link{synthetic}} is used.
+#' appends the \code{.gitignore} file to exclude \code{filename}, and saves a
+#' synthetic copy of \code{data} for public use. To generate these synthetic
+#' data, the function \code{\link{synthetic}} is used.
 #' @inheritParams open_data
 #' @return Returns \code{NULL} invisibly. This
 #' function is called for its side effects.

### 2020-05-12 R/worcs_project.R

@@ -12,7 +12,7 @@ recommend_data <- c('library("worcs")',
 #' @param path Character, indicating the directory in which to create the
 #' 'worcs' project. Default: 'worcs_project'.
 #' @param manuscript Character, indicating what template to use for the
-#' 'rmarkdown' manuscript. Default: 'APA6'. Available choices include:
+#' 'R Markdown' manuscript. Default: 'APA6'. Available choices include:
 #' \code{"APA6", "github_document", "None", "ams_article", "asa_article",
 #' "biometrics_article", "copernicus_article", "ctex", "elsevier_article",
 #' "frontiers_article", "ieee_article", "joss_article", "jss_article",
@@ -47,7 +47,7 @@ recommend_data <- c('library("worcs")',
 #' the_test <- "worcs_template"
 #' old_wd <- getwd()
 #' dir.create(file.path(tempdir(), the_test))
-#' do.call(git_credentials, worcs:::get_credentials())
+#' do.call(git_user, worcs:::get_user())
 #' worcs_project(file.path(tempdir(), the_test, "worcs_project"),
 #'               manuscript = "github_document",
 #'               preregistration = "None",
@@ -230,7 +230,7 @@ worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregist
     tab <- matrix(c("File", "Description", "Usage",
                     "README.md", "Description of project", "Human editable"), nrow = 2, byrow = TRUE)
     rproj_name <- paste0(gsub("^.+\\b(.+)$", "\\1", path), ".Rproj")
-    cont[grep("You can load this project in Rstudio by opening the file called ", cont)] <- paste0(grep("You can load this project in Rstudio by opening the file called ", cont, value = TRUE), "'", rproj_name, "'.")
+    cont[grep("You can load this project in RStudio by opening the file called ", cont)] <- paste0(grep("You can load this project in Rstudio by opening the file called ", cont, value = TRUE), "'", rproj_name, "'.")
     tab <- rbind(tab, c(rproj_name, "Project file", "Loads project"))
     tab <- describe_file("LICENSE", "User permissions", "Read only", tab, path)
     tab <- describe_file(".worcs", "WORCS metadata YAML", "Read only", tab, path)

### 2020-05-12 R/zzz.R

@@ -5,6 +5,6 @@
   if(!has_git()){
     print_message <- paste0(print_message, "\n\033[0;31mCould not find a working installation of 'Git', which is required to safeguard the transparency and reproducibility of your project. Please connect 'Git' by following the steps described in this vignette:\n  vignette('setup', package = 'worcs')\033[0m")
   }
-  #if(!correct_version){ "Rstudio version 1.1.28 or higher is required to use the worcs package."))
+  #if(!correct_version){ "RStudio version 1.1.28 or higher is required to use the worcs package."))
   packageStartupMessage(print_message)
 }

### 2020-05-12 tests/testthat/test-citations.R

@@ -49,7 +49,7 @@ test_that("cite_* work with umlaute", {
   sink_out <- capture_output(cite_all(file_name))
   md_file <- force(gsub("Rmd$", "md", file_name))
   contents <- worcs:::read_as_utf(md_file)
-  print(tail(contents))
+
   expect_true(any(contents == "Optional räference: @reference2020"))
 
   sink_out <- capture_output(cite_essential(file_name))

### 2020-05-12 tests/testthat/test-export_project.R

@@ -12,35 +12,47 @@
 #   }
 # }
 
-do.call(git_credentials, worcs:::get_credentials())
+# See if ZIP works at all
+run_test <- tryCatch({
+  testfile <- tempfile(fileext = ".txt")
+  writeLines("test me", testfile)
+  zipfile <- tempfile(fileext = ".zip")
+  res <- zip(zipfile = zipfile, files = testfile, flags="-jrq")
+  res == 0
+}, error = function(e){FALSE})
 
-testdir <- file.path(tempdir(), "export")
-testzip <- tempfile(fileext = ".zip")
-dir.create(testdir)
-suppressWarnings(
-  worcs_project(
-    testdir,
-    manuscript = "None",
-    preregistration = "None",
-    add_license = "None",
-    use_renv = FALSE,
-    remote_repo = "https"
+if(run_test){
+  do.call(git_user, worcs:::get_user())
+
+  testdir <- file.path(tempdir(), "export")
+  testzip <- tempfile(fileext = ".zip")
+  dir.create(testdir)
+  suppressWarnings(
+    worcs_project(
+      testdir,
+      manuscript = "None",
+      preregistration = "None",
+      add_license = "None",
+      use_renv = FALSE,
+      remote_repo = "https"
+    )
   )
-)
 
-test_that("worcs_project created successfully", {
-  expect_true(file.exists(file.path(testdir, ".worcs")))
-})
+  test_that("worcs_project created successfully", {
+    expect_true(file.exists(file.path(testdir, ".worcs")))
+  })
 
-result <- export_project(zipfile = testzip, worcs_directory = testdir, open_data = FALSE)
+  result <- export_project(zipfile = testzip, worcs_directory = testdir, open_data = FALSE)
 
-test_that("export returned true", {
-  expect_true(result)
-})
+  test_that("export returned true", {
+    expect_true(result)
+  })
 
-test_that("exported worcs_project exists", {
-  expect_true(file.exists(testzip))
-})
+  test_that("exported worcs_project exists", {
+    expect_true(file.exists(testzip))
+  })
 
-# setwd(old_wd)
-#unlink(c(testzip, testdir), recursive = TRUE)
+  # setwd(old_wd)
+  #unlink(c(testzip, testdir), recursive = TRUE)
+
+}

### 2020-05-12 README.Rmd
update 2020-05-12 19:48:02

@@
 
-# WORCS <a href='https://osf.io/zcvbs/'><img src='paper/worcs_icon.png' align="right" height="139" /></a>
+# WORCS <a href='https://osf.io/zcvbs/'><img src='worcs_icon.png' align="right" height="139" /></a>
 
@@
 [![](https://cranlogs.r-pkg.org/badges/worcs)](https://cran.r-project.org/package=worcs)-->
-[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
+[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
 [![Travis build status](https://travis-ci.org/cjvanlissa/worcs.svg?branch=master)](https://travis-ci.org/cjvanlissa/worcs)

### 2020-05-13 paper/Manuscript.Rmd
Updat Manuscript
Added some info on data sharing and personal/pseudononymized data

@@
 Step 20 of the WORCS procedure suggests uploading the online version to a preprint server.
-This is important because most preprint servers - [at least the ones hosted by OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
+This is important because most preprint servers - [at least the ones hosted by the Open Science Framework OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
 This means that authors will receive credit for cited work; even if they are cited only in the online version.
@@
 Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns.
-For example, the GDPR prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU.
-Different rules may apply to anonymized or pseudonimized data.
-Before sharing any data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.
+When working with data from human participants it is important to delete any sensitive personal data and contact information as soon as data collection has finished.
+The resulting pseudononimized data can often be shared if informed consent was given.
+Nevertheles, it is important to note that the GDPR in europe prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU, unless they offer an "adequate level of protection".
+Although different rules may apply to pseudonimized data, there are repositories with EU servers such as the Open Science Framework and universities might have their own repositories that are complient with local legistation.
+Before sharing any human participant data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.
 
@@
 can I indicate a point in my workflow (e.g.
-after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it.
+after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->When data can be shared, it is important to version control the pseudonymized <!-- MS: it is almost impossible to anonymize data, so it's preferred to use pseudononymized ---> raw data upon collection, and avoid changing it.
 Any data processing steps should be documented in the code, instead of applied to the raw data.

### 2020-05-16 paper/Manuscript.Rmd
edits and suggestion by Anna-Lena

--- 2020-05-12 paper/Manuscript.Rmd
@@
 title: |
-    | WORCS: A Workflow for Open Reproducible Code in Science
+    | WORCS: A Workflow for Open Reproducible Code in the Social Sciences and Beyond <!-- AL: I strongly recommend to make the title more specific.
+"Science" is very general, and in the Anglo-American understanding actually refers to the beta sciences.
+Also, the abstract and the paper start with talking about the situation in the social sciences and derive the motivaion from there, so it would be good if the title would reflect this.
+The acronym can stay as it is.
+;) -->
     | Version 0.1.0 (Feb 8, 2020)
@@
     institution: "7"
+  - name: "Anna-Lena Lamprecht"
+    institution: "2,8"
 
@@
     institution: "University Library Utrecht"
+  - id: "8"
+    institution: "Utrecht University, Department of Information and Computing Sciences"
 
@@
 
+<!-- AL: I second LB's comment somewhere below that the overall WORCS workflow should be introduced much earlier, maybe already here.
+A diagrammatic overview/flowchart of the principal steps would be great.
+It should catch the reader's eye and give an immediate idea of what WORCS is about.
+Now it is indeed quite buried in the details that follow.
+-->
+
 <!--
@@
 Collective benefits: Cumulative science, -->
+<!-- AL: These "value propositions" would also be good to discuss already in the introduction. Directly after the introduction of the WORCS overview would make sense, I think. -->
+
+The remainder of the paper is structured as follows.
+...
+<!-- AL: Or some variation of this obligatory end-of-intro paragraph.
+As the paper touches on a lot of aspects, from principal to very practical, it would be very useful to provide some guidance here.
+-->
 
@@
 The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement.
-This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
-This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them.
+<!-- AL: Is there a reference for this?
+Maybe a suitable blog post, or just the OSC website?
+--> This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
+The WORCS workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them.
 
@@
 
+<!-- AL: Similar to my comment on the title, in this section it needs to make clear which scope of related work it covers.
+For example, in the life sciences there has also been a lot of work on practices for increasing reproducibility, with automated workflows, containerization, Research Objects, provenance protocols, and a lot more.
+I don't think it's necessary for this paper to discuss this, but that is under the provision that it defines its scope better.
+-->
+
 ## Introducing the tools
@@
 The first of these is *dynamic document generation*.
-Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses.
+<!-- AL: A bit nitpicky, but: DDG as such is a technique, not a solution.
+Could also be implemented as proprietary, closed source.
+Papaja is the tool/solution.
+The description of formatting as an APA-style manuscript also seems to be tool-specific (generall DDG could also generate any other kind of format), so maybe reorganize this paragraph.
+--> Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses.
 The text is automatically formatted as an APA-style manuscript [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables.
@@
 * Human error in copy-pasting results is eliminated
-* When revisions require major changes to the analyses, all results, Figures and Tables are automatically updated
+* When revisions require major changes to the analyses, all results, figures and tables are automatically updated
 * Flexibility in output formats means that a manuscript can be rendered to presentation format, or even to a website. 
@@
 The second solution is *version control*: Maintaining an indelible log of every change to all project files.
-Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax [see also @blischakQuickIntroductionVersion2016].
+Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax <!-- AL: syntax is also to be respected in e.g.
+interactive R environments, so that alone does not help -  maybe "code files" instead?
+-->  [see also @blischakQuickIntroductionVersion2016].
 Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions.
@@
 These solutions differ in user-friendliness and effectiveness.
-If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine.
+If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a container <!-- AL: Docker containers are not virtual machine.
+Actually that is the whole idea of Docker (and other container technologies), as real virtual machines need way more resources.
+--> .
 This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data).
@@
 Several arguments support the choice to work in R.
-First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science.
+First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science <!-- AL: Python is also open source, so phrase this argument with care.
+--> .
 Second, all of the aforementioned tools for an open science workflow are already implemented in R.
@@
 Wrappers in R are often available for tools developed in other programming languages.
-<!-- LB: Perhaps give some pointers to learn R (e.g.
-R4DataScience)-->Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
-Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.
+<!-- AL: Maybe this should be the first argument.
+While Python is catching up, R still is THE statistical programming environment in many communities, and is thus favorable in many cases.--> <!-- LB: Perhaps give some pointers to learn R (e.g.
+R4DataScience)-->Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012]<!-- AL: If the reference is really from 2012, it is not a good support for "at the time of writing"...
+--> , following SPSS, which has no support for any of the open science tools discussed in this paper.
+Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.<!-- AL: Many of these are also true for Python.
+There, Jupyter notebooks are very popular as export and exchange format (also using MarkDown), might be good to refer to that somewhere.
+I think they also support R.
+-->
+<!-- AL: And maybe this goes a bit too far for this paper, but https://www.scientificpaperofthefuture.org/ also fits in this context. -->
 
@@
 A case can be made to use text-based files only (or as much as possible), instead of binary files.
-Text-based files can be read by machines and humans alike, without requiring special (licensed) software.
-Binary files, such as Word (`.docx`) or SPSS (`.sav`, `.spo`),  must be decoded by software first.
+Text-based files can be read by machines and humans alike (just requiring a plain text editor), without requiring special (often licensed) software.
+Binary files, such as Word documents or (`.docx`) or SPSS objects (`.sav`, `.spo`), must be decoded by specialized software first.
 Git is designed to keep track of changes to the content of text-based files.
@@
 <!-- LB:  this section comes at a weird place in the flow of the paper.
--->Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct.
+--><!-- AL: +1 --> Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct.
 Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.
@@
 
+<!-- AL: Some general things about software citation that could be good to mention here: Tool or software papers are also common, often as short papers, sometimes called "Application Notes".
+The Journal of Open Source Software (JOSS) is a way of turning software directly into a publication, without the need of writing an actual paper about it (the review is about the code).
+https://citation-file-format.github.io/ is a file format to include software citation information into source code repositories.-->
 When using R, software (R packages) can be conveniently cited by calling `citation("packagename")`.
@@
 It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication.
-Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference.
+Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`)<!-- AL: what is the difference?
+do you mean '@@'?
+or is it just showing the same on my machine?-->, used to cite a non-essential reference.
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
@@
 Once the GitHub repository is made public, people will have access to a checksum for the raw data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
+
+<!-- Maybe remark that in the "closed" case, a the data should be backupped somewhere else, as GitHub will not do the job in this setting. -->
 
@@
 # The workflow 
+<!-- AL: Agreeing with LB's ealier comment, it would be good to discuss this overall workflow before going into the details. -->
 
@@
 I think you can also use Overleaf instead of Word (user friendly and web based) for bidirectional communication, but haven’t tried this myself.
-I didn’t know about ‘redoc’, will check it out.-->Then, manually incorporate their changes into the `manuscript.Rmd` file.
+I didn’t know about ‘redoc’, will check it out.-->Then, manually<!--this should really be avoided if possible!
+no fun...
+--> incorporate their changes into the `manuscript.Rmd` file.
 In most cases, this is the most user-friendly approach, as most lead authors would review changes by co-authors anyway.
@@
 
-If all collaborators are committed to using WORCS, they can [Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo) from the lead author on GitHub, [clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository), make their own changes, and [send you a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) to incorporate their changes.
+If all collaborators are committed to using WORCS, they can [Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo) from the lead author on GitHub, [clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository), make their own changes, and [send a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) to incorporate their changes.
 Working this way is extremely conducive to scientific collaboration [@ramGitCanFacilitate2013].
 Moreover, being familiar with GitHub opens doors to new forms of collaboration: In the open source software community, continuous peer review and voluntary collaborative acts by strangers who are interested in a project are commonplace [see @adolphOpenBehavioralScience2012].
-This kind of collaboration is budding in scientific software development as well; for example, I became a co-author on several R-packages after submitting pull requests with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018].
+This kind of collaboration is budding in scientific software development as well; for example, I <!-- AL: which of the meanwhile several authors?
+:) -->  became a co-author on several R-packages after submitting pull requests with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018].
 It is also possible to invite such collaboration by [opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues) for tasks that still need to be accomplished, and tag known Collaborators to address them, or invite external collaborators to contribute their expertise.

### 2020-05-17 paper/Manuscript.Rmd
edits and comments to the manuscript

--- 2020-05-12 paper/Manuscript.Rmd
@@
 
-author:
+author: #"`r source('authors_yaml.R')`"
   - name: "Caspar J. Van Lissa"
     institution: "1,2"
+    orcid: "0000-0002-0808-5024"
     corresponding: yes
@@
     institution: "3,4"
+    orcid: "0000-0001-8765-6982"
   - name: "Loek Brinkman"
     institution: "2,5"
+    orcid: "0000-0003-3997-1173"
   - name: "Aaron Peikert"
-    institution: "3,6"
-  - name: "Barbara Vreede"
+    institution: "6"
+    orcid: "0000-0001-7813-818X"
+  - name: "Barbara M.I. Vreede"
     institution: "7"
+    orcid: "0000-0002-5023-4601"
 
-affiliation:
+affiliation: #"`r source('affiliation_yaml.R')`"
   - id: "1"
@@
   - id: "6"
-    institution: "Humboldt-Universitaet zu Berlin"
+    institution: "Test"
   - id: "7"
@@
 # Introduction
-
-The social sciences are amidst a paradigm shift towards open science.
+<!--
+Barbara:
+A trick I learned in collaborative writing with git is to make sure each new sentence is on a new line.
+This way, you can effectively use a git diff to see changes in individual sentences instead of having to scour through an entire paragraph.
+I'll add this, but only in the paragraphs I've changed, so you do not get inundated with differences!
+-->
+The (social) sciences are amidst a paradigm shift towards open science. 
 Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [as discussed by @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
@@
 
-The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
+The TOP-guidelines are primarily geared towards top-down systemic change by journals, funders, and organizations.
 However, large organizations are slow to change.
 For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
-<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 --> At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
+<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 -->
+At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
 Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support.
@@
 This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
-This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them.
+This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions, and where such requirements do exist, this workflow will help fulfill them.
 
@@
 
+## Prerequisites
+<!--Barbara: I'm moving up this section and editing it a bit, to introduce more explicitly that worcs was developed for R.-->
+WORCS has focused on developing a workflow for R, instead of for SPSS or Python.
+<!-- Barbara: same as for Git, I could not find a citation. Maybe cite the website: www.r-project.org -->
+Several arguments support the choice to work in R.
+First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science.
+Second, many tools for an open science workflow are already implemented in R.
+A vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
+Wrappers in R are often available for tools developed in other programming languages.
+<!-- LB: Perhaps give some pointers to learn R (e.g. R4DataScience)-->
+Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
+Finally, R is highly interoperable:
+Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.
+
+Working with R is simplified immensely by using the RStudio integrated development engine (IDE) for R.
+<!-- Barbara: 
+https://support.rstudio.com/hc/en-us/articles/206212048-Citing-RStudio says:
+To cite RStudio in publications, you can get the latest citation information by running the command RStudio.Version() in a recent version of RStudio IDE.
+
+I did this, which gave:
+
+  @Manual{,
+    title = {RStudio: Integrated Development Environment for R},
+    author = {{RStudio Team}},
+    organization = {RStudio, Inc.},
+    address = {Boston, MA},
+    year = {2016},
+    url = {http://www.rstudio.com/},
+  }
+-->
+Many of the tools used in WORCS are embedded directly into the user interface of RStudio.
+RStudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
+This makes RStudio a comprehensive solution for open science research projects.
+Another important feature of RStudio is project management.
+A project bundles writing, analyses, data, references, etcetera, into a self-contained folder, that can be uploaded entirely to GitHub and downloaded by future users.
+The `worcs` R-package installs a new RStudio project template.
+When a new project is initialized from this template, much of the bookkeeping required to set up an open science project is performed automatically.
+
 ## Introducing the tools
@@
 
-The second solution is *version control*: Maintaining an indelible log of every change to all project files.
+The second solution is *version control*:
+Maintaining an indelible log of every change to all project files.
 Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax [see also @blischakQuickIntroductionVersion2016].
 Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions.
-WORCS instead uses the *explicit* version control software "Git".
-Git tracks changes to files, and stores these changes when the user makes a "commit".
+WORCS instead uses the *explicit* version control software "Git". Git tracks changes to files, and stores these changes when the user makes a "commit".
+<!--Barbara: I have looked for a way to cite Git but did not find one...
+Seems good practice especially in this paper as an example to cite all software, but for Git I just cannot find the way.
+Maybe cite the website: www.git-scm.com -->
 Git retains a complete log of all commits, and users can compare changes between different commits, or go back to a previous version of the code (for example, after making a mistake, or to replicate a previous version of the results).
@@
 
-## Additional considerations
-
-A few additional considerations were considered in developing WORCS.
-The first consideration is to develop a workflow for R, instead of for SPSS or Python.
-Several arguments support the choice to work in R.
-First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science.
-Second, all of the aforementioned tools for an open science workflow are already implemented in R.
-Many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones.
-Wrappers in R are often available for tools developed in other programming languages.
-<!-- LB: Perhaps give some pointers to learn R (e.g.
-R4DataScience)-->Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
-Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.
-
-Working with R is simplified immensely by using the RStudio user interface for R.
-Many of the tools used in WORCS are embedded directly into the visual user interface of RStudio.
-RStudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
-This makes RStudio a comprehensive solution for open science research projects.
-Another important feature of RStudio is project management.
-A project bundles writing, analyses, data, references, etcetera, into a self-contained folder, that can be uploaded entirely to GitHub and downloaded by future users.
-The `worcs` R-package installs a new RStudio project template.
-When a new project is initialized from this template, much of the bookkeeping required to set up an open science project is performed automatically.
-
-Another key consideration is what filetypes to use.
+## Considering filetypes
+A key consideration when developing a research project is what filetypes to use.
 A case can be made to use text-based files only (or as much as possible), instead of binary files.
@@
 Each change reflects a word or line of code added or removed.
-Binary files, on the other hand, should not be version controlled.
-The changes in the encrypted file are uninterpretable.
-Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub.
-Uploading large binary files is frowned upon.
+Binary files, on the other hand, should not be version controlled with Git, as individual changes in the encrypted file are uninterpretable.
+Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub. Uploading large binary files is frowned upon.
 GitHub also allows visitors to peruse text-based files online, and conveniently renders certain filetypes.
 For example, GitHub highlights syntax in `.R` and `.Rmd` files, and displays `.csv` files as spreadsheets.
+
+<!--Barbara: Maybe write here something about writing manuscripts on a line-by-line basis, so that version control can keep track of changes more accurately than if a whole paragraph is identified as altered?-->
 
@@
 <!-- LB:  this section comes at a weird place in the flow of the paper. 
--->Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct.
+Barbara: agreed! I would remove this altogether.-->
+Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct.
 Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.
@@
 Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count.
-Researchers can overcome this impediment by preparing two versions of the manuscript: One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
+Researchers can overcome this impediment by preparing two versions of the manuscript:
+One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
 The print version should reference the online version, so interested readers can find the comprehensive reference list.
 Step 20 of the WORCS procedure suggests uploading the online version to a preprint server.
-This is important because most preprint servers - [at least the ones hosted by OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
+This is important because most preprint servers - [at least the ones listed by OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
 This means that authors will receive credit for cited work; even if they are cited only in the online version.
@@
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
+<!-- Barbara: I see no difference between these two symbols...-->
 
@@
 The first solution is publishing a *checksum* of the raw data upon collection.
-A checksum represents the contents of a file in a 32-character summary.
-Any change to the underlying file results in a different checksum.
-Thus, one can use a checksum to verify that the contents of a file remain unchanged.
-If researchers publish the checksum of their raw data and the complete analysis code, then they can prove to an independent auditer that running the public analysis code on the private dataset with the correct checksum results in the published results of their work.
+A checksum is the result of an algorithm applied to a file, and represents the contents of this file in 32-characters.
+Any change to the underlying file results in a different checksum. Thus, one can use a checksum to verify that the contents of a file remain unchanged.
+If researchers publish the checksum of their raw data, in addition to the complete analysis code, then they can prove to an independent auditer that running the public analysis code on the private dataset with the correct checksum results in the published results of their work.
 
@@
 The ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) functionality on GitHub is a particularly user-friendly way to do this.
+<!--Barbara: I'm not sure this is the case.
+Pull requests are useful if someone wants to change or add to a project, but this would not involve a change to an existing project, but rather a new project/analysis with "old" data.
+I would remove this suggestion altogether.-->
 
@@
 Once the GitHub repository is made public, people will have access to a checksum for the raw data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
+
+<!--Barbara: Is it indeed the case that data by default is loaded as a .csv?
+How about (raw) text data; can worcs accommodate a .txt file, for instance (in case of text analysis)?
+Or should this all be done in prepare_data.R?-->
 
@@
 Private repositories can be made public at a later date, along with their entire time-stamped history.
+<!-- Barbara: Pull requests are not anonymous either, aren't they?-->
 
@@
 This topic is beyond the scope of the present paper, but many user-friendly methods for simulating data are available in `R`, including the package [`simstudy`](https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html).
+<!-- citation:
+
+  @Manual{,
+    title = {simstudy: Simulation of Study Data},
+    author = {Keith Goldfeld},
+    year = {2020},
+    note = {R package version 0.1.16},
+    url = {https://CRAN.R-project.org/package=simstudy},
+  }
+  -->
 When using the WORCS-workflow, it is recommended to preregister, if possible, a `.R` script with the planned analyses, along with a verbal, conceptual description.
@@
 
-Before you can use the WORCS workflow, you have to install the required software, and register for a GitHub account.
-The setup process is documented in a vignette included with the `worcs` R-package, and also [available online](https://cjvanlissa.github.io/worcs/articles/setup.html).
+Before you can use the WORCS workflow, you have to install the required software, and register for a GitHub account.The setup process is documented in a vignette included with the `worcs` R-package, and also [available online](https://cjvanlissa.github.io/worcs/articles/setup.html).
 After preparing your system, you can follow the step-by-step procedure outlined below.
+<!--Barbara: Does this work on gitlab/bitbucket as well?-->
 
@@
 
-1. Create a new Private repository on GitHub. Do not change any of the default options. Copy the repository link to the clipboard. This link will look like https://github.com/yourname/yourrepo.git
+1. Create a new Private repository on GitHub.
+<!-- Why private!?!
+I would actually argue for a public repo, but if you do not feel comfortable doing so: another option would be to give an overview of advantages/disadvantages of private/public here, and make an argument for either case.-->
+Do not change any of the default options. Copy the repository link to the clipboard. This link will look like https://github.com/yourname/yourrepo.git
 2. In RStudio, click File > New Project > New directory > WORCS Project Template
     a. Paste the GitHub Repository address in the textbox
-    b. Keep the checkbox for `renv` checked if you want to document all R-package dependencies, and have them automatically loaded on any system (recommended).
+    b.
+Keep the checkbox for `renv` checked if you want to document all dependencies (recommended).
+The only reason not to do this is if you have dependencies outside of the R environment, or if you are using a more comprehensive solution for computational reproducibility, e.g., @peikertReproducibleDataAnalysis2019.
     c. Select a preregistration template
@@
 6. Optional: Render the preregistration to PDF, and upload it to AsPredicted.org or OSF.io as an attachment
-7.
-Optional: Add study Materials (to which you own the rights) to the repository.
-It is possible to solicit feedback (by opening a GitHub Issue) and acknowledge outside contributions (by accepting Pull requests)
+7. Optional: Add study Materials (only those to which you own the rights!) to the repository.
+It is possible to solicit feedback (by opening a GitHub Issue) and receive outside contributions (by accepting Pull requests)
+<!-- Barbara: NB this can only be done in a public repo; another argument for making a repo public! -->
 
@@
 12. Cite essential references with one at-symbol (`[@essentialref2020]`), and non-essential references with a double at-symbol (`[@nonessential2020]`).
+<!-- Barbara: Again I cannot see the difference... -->
 

### 2020-05-17 paper/Manuscript.Rmd
manual change to resolve conflicts

@@
   - name: "Aaron Peikert"
-    institution: "6"
+    institution: "3,6"
     orcid: "0000-0001-7813-818X"

### 2020-05-18 paper/Manuscript.Rmd

--- 2020-05-17 paper/Manuscript.Rmd
@@
 
-author: #"`r source('authors_yaml.R')`"
+author:
   - name: "Caspar J. Van Lissa"
     institution: "1,2"
-    orcid: "0000-0002-0808-5024"
     corresponding: yes
@@
     institution: "3,4"
-    orcid: "0000-0001-8765-6982"
   - name: "Loek Brinkman"
     institution: "2,5"
-    orcid: "0000-0003-3997-1173"
   - name: "Aaron Peikert"
-    institution: "6"
-    orcid: "0000-0001-7813-818X"
-  - name: "Barbara M.I. Vreede"
+    institution: "3,6"
+  - name: "Barbara Vreede"
     institution: "7"
-    orcid: "0000-0002-5023-4601"
 
-affiliation: #"`r source('affiliation_yaml.R')`"
+affiliation:
   - id: "1"
@@
   - id: "6"
-    institution: "Test"
+    institution: "Humboldt-Universitaet zu Berlin"
   - id: "7"
@@
 # Introduction
-<!--
-Barbara:
-A trick I learned in collaborative writing with git is to make sure each new sentence is on a new line.
-This way, you can effectively use a git diff to see changes in individual sentences instead of having to scour through an entire paragraph.
-I'll add this, but only in the paragraphs I've changed, so you do not get inundated with differences!
--->
-The (social) sciences are amidst a paradigm shift towards open science. 
+
+The social sciences are amidst a paradigm shift towards open science.
 Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [as discussed by @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
@@
 
-The TOP-guidelines are primarily geared towards top-down systemic change by journals, funders, and organizations.
+The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
 However, large organizations are slow to change.
 For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
-<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 -->
-At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
+<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 --> At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
 Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support.
@@
 This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
-This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions, and where such requirements do exist, this workflow will help fulfill them.
+This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions: If such requirements exist, this workflow will help fulfill them.
 
@@
 ```
-
-## Prerequisites
-<!--Barbara: I'm moving up this section and editing it a bit, to introduce more explicitly that worcs was developed for R.-->
-WORCS has focused on developing a workflow for R, instead of for SPSS or Python.
-<!-- Barbara: same as for Git, I could not find a citation. Maybe cite the website: www.r-project.org -->
-Several arguments support the choice to work in R.
-First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science.
-Second, many tools for an open science workflow are already implemented in R.
-A vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
-Wrappers in R are often available for tools developed in other programming languages.
-<!-- LB: Perhaps give some pointers to learn R (e.g. R4DataScience)-->
-Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
-Finally, R is highly interoperable:
-Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.
-
-Working with R is simplified immensely by using the RStudio integrated development engine (IDE) for R.
-<!-- Barbara: 
-https://support.rstudio.com/hc/en-us/articles/206212048-Citing-RStudio says:
-To cite RStudio in publications, you can get the latest citation information by running the command RStudio.Version() in a recent version of RStudio IDE.
-
-I did this, which gave:
-
-  @Manual{,
-    title = {RStudio: Integrated Development Environment for R},
-    author = {{RStudio Team}},
-    organization = {RStudio, Inc.},
-    address = {Boston, MA},
-    year = {2016},
-    url = {http://www.rstudio.com/},
-  }
--->
-Many of the tools used in WORCS are embedded directly into the user interface of RStudio.
-RStudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
-This makes RStudio a comprehensive solution for open science research projects.
-Another important feature of RStudio is project management.
-A project bundles writing, analyses, data, references, etcetera, into a self-contained folder, that can be uploaded entirely to GitHub and downloaded by future users.
-The `worcs` R-package installs a new RStudio project template.
-When a new project is initialized from this template, much of the bookkeeping required to set up an open science project is performed automatically.
 
@@
 
-The second solution is *version control*:
-Maintaining an indelible log of every change to all project files.
+The second solution is *version control*: Maintaining an indelible log of every change to all project files.
 Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax [see also @blischakQuickIntroductionVersion2016].
 Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions.
-WORCS instead uses the *explicit* version control software "Git". Git tracks changes to files, and stores these changes when the user makes a "commit".
-<!--Barbara: I have looked for a way to cite Git but did not find one...
-Seems good practice especially in this paper as an example to cite all software, but for Git I just cannot find the way.
-Maybe cite the website: www.git-scm.com -->
+WORCS instead uses the *explicit* version control software "Git".
+Git tracks changes to files, and stores these changes when the user makes a "commit".
 Git retains a complete log of all commits, and users can compare changes between different commits, or go back to a previous version of the code (for example, after making a mistake, or to replicate a previous version of the results).
@@
 
-## Considering filetypes
-A key consideration when developing a research project is what filetypes to use.
+## Additional considerations
+
+A few additional considerations were considered in developing WORCS.
+The first consideration is to develop a workflow for R, instead of for SPSS or Python.
+Several arguments support the choice to work in R.
+First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science.
+Second, all of the aforementioned tools for an open science workflow are already implemented in R.
+Many statistical methods are first implemented in R, as there is a vibrant online community developing new R-packages and providing support and tutorials for existing ones.
+Wrappers in R are often available for tools developed in other programming languages.
+<!-- LB: Perhaps give some pointers to learn R (e.g.
+R4DataScience)-->Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
+Finally, R is highly interoperable: Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.
+
+Working with R is simplified immensely by using the RStudio user interface for R.
+Many of the tools used in WORCS are embedded directly into the visual user interface of RStudio.
+RStudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
+This makes RStudio a comprehensive solution for open science research projects.
+Another important feature of RStudio is project management.
+A project bundles writing, analyses, data, references, etcetera, into a self-contained folder, that can be uploaded entirely to GitHub and downloaded by future users.
+The `worcs` R-package installs a new RStudio project template.
+When a new project is initialized from this template, much of the bookkeeping required to set up an open science project is performed automatically.
+
+Another key consideration is what filetypes to use.
 A case can be made to use text-based files only (or as much as possible), instead of binary files.
@@
 Each change reflects a word or line of code added or removed.
-Binary files, on the other hand, should not be version controlled with Git, as individual changes in the encrypted file are uninterpretable.
-Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub. Uploading large binary files is frowned upon.
+Binary files, on the other hand, should not be version controlled.
+The changes in the encrypted file are uninterpretable.
+Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub.
+Uploading large binary files is frowned upon.
 GitHub also allows visitors to peruse text-based files online, and conveniently renders certain filetypes.
 For example, GitHub highlights syntax in `.R` and `.Rmd` files, and displays `.csv` files as spreadsheets.
-
-<!--Barbara: Maybe write here something about writing manuscripts on a line-by-line basis, so that version control can keep track of changes more accurately than if a whole paragraph is identified as altered?-->
 
@@
 <!-- LB:  this section comes at a weird place in the flow of the paper.
-Barbara: agreed! I would remove this altogether.-->
-Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct.
+-->Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct.
 Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.
@@
 Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count.
-Researchers can overcome this impediment by preparing two versions of the manuscript:
-One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
+Researchers can overcome this impediment by preparing two versions of the manuscript: One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
 The print version should reference the online version, so interested readers can find the comprehensive reference list.
 Step 20 of the WORCS procedure suggests uploading the online version to a preprint server.
-This is important because most preprint servers - [at least the ones listed by OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
+This is important because most preprint servers - [at least the ones hosted by OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
 This means that authors will receive credit for cited work; even if they are cited only in the online version.
@@
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
-<!-- Barbara: I see no difference between these two symbols...-->
 
@@
 The first solution is publishing a *checksum* of the raw data upon collection.
-A checksum is the result of an algorithm applied to a file, and represents the contents of this file in 32-characters.
-Any change to the underlying file results in a different checksum. Thus, one can use a checksum to verify that the contents of a file remain unchanged.
-If researchers publish the checksum of their raw data, in addition to the complete analysis code, then they can prove to an independent auditer that running the public analysis code on the private dataset with the correct checksum results in the published results of their work.
+A checksum represents the contents of a file in a 32-character summary.
+Any change to the underlying file results in a different checksum.
+Thus, one can use a checksum to verify that the contents of a file remain unchanged.
+If researchers publish the checksum of their raw data and the complete analysis code, then they can prove to an independent auditer that running the public analysis code on the private dataset with the correct checksum results in the published results of their work.
 
@@
 The ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) functionality on GitHub is a particularly user-friendly way to do this.
-<!--Barbara: I'm not sure this is the case.
-Pull requests are useful if someone wants to change or add to a project, but this would not involve a change to an existing project, but rather a new project/analysis with "old" data.
-I would remove this suggestion altogether.-->
 
@@
 Once the GitHub repository is made public, people will have access to a checksum for the raw data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
-
-<!--Barbara: Is it indeed the case that data by default is loaded as a .csv?
-How about (raw) text data; can worcs accommodate a .txt file, for instance (in case of text analysis)?
-Or should this all be done in prepare_data.R?-->
 
@@
 Private repositories can be made public at a later date, along with their entire time-stamped history.
-<!-- Barbara: Pull requests are not anonymous either, aren't they?-->
 
@@
 This topic is beyond the scope of the present paper, but many user-friendly methods for simulating data are available in `R`, including the package [`simstudy`](https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html).
-<!-- citation:
-
-  @Manual{,
-    title = {simstudy: Simulation of Study Data},
-    author = {Keith Goldfeld},
-    year = {2020},
-    note = {R package version 0.1.16},
-    url = {https://CRAN.R-project.org/package=simstudy},
-  }
-  -->
 When using the WORCS-workflow, it is recommended to preregister, if possible, a `.R` script with the planned analyses, along with a verbal, conceptual description.
@@
 
-Before you can use the WORCS workflow, you have to install the required software, and register for a GitHub account.The setup process is documented in a vignette included with the `worcs` R-package, and also [available online](https://cjvanlissa.github.io/worcs/articles/setup.html).
+Before you can use the WORCS workflow, you have to install the required software, and register for a GitHub account.
+The setup process is documented in a vignette included with the `worcs` R-package, and also [available online](https://cjvanlissa.github.io/worcs/articles/setup.html).
 After preparing your system, you can follow the step-by-step procedure outlined below.
-<!--Barbara: Does this work on gitlab/bitbucket as well?-->
 
@@
 
-1. Create a new Private repository on GitHub.
-<!-- Why private!?!
-I would actually argue for a public repo, but if you do not feel comfortable doing so: another option would be to give an overview of advantages/disadvantages of private/public here, and make an argument for either case.-->
-Do not change any of the default options. Copy the repository link to the clipboard. This link will look like https://github.com/yourname/yourrepo.git
+1. Create a new Private repository on GitHub. Do not change any of the default options. Copy the repository link to the clipboard. This link will look like https://github.com/yourname/yourrepo.git
 2. In RStudio, click File > New Project > New directory > WORCS Project Template
     a. Paste the GitHub Repository address in the textbox
-    b.
-Keep the checkbox for `renv` checked if you want to document all dependencies (recommended).
-The only reason not to do this is if you have dependencies outside of the R environment, or if you are using a more comprehensive solution for computational reproducibility, e.g., @peikertReproducibleDataAnalysis2019.
+    b. Keep the checkbox for `renv` checked if you want to document all R-package dependencies, and have them automatically loaded on any system (recommended).
     c. Select a preregistration template
@@
 6. Optional: Render the preregistration to PDF, and upload it to AsPredicted.org or OSF.io as an attachment
-7. Optional: Add study Materials (only those to which you own the rights!) to the repository.
-It is possible to solicit feedback (by opening a GitHub Issue) and receive outside contributions (by accepting Pull requests)
-<!-- Barbara: NB this can only be done in a public repo; another argument for making a repo public! -->
+7.
+Optional: Add study Materials (to which you own the rights) to the repository.
+It is possible to solicit feedback (by opening a GitHub Issue) and acknowledge outside contributions (by accepting Pull requests)
 
@@
 12. Cite essential references with one at-symbol (`[@essentialref2020]`), and non-essential references with a double at-symbol (`[@nonessential2020]`).
-<!-- Barbara: Again I cannot see the difference... -->
 
@@
 
+## Outlook
+
+<!-- not sure where this repro integration section belongs -->
+The `worcs`-package integrates several reproducibility tools into a convenient workflow by assuming reasonable standards that fulfil most researches needs.
+Currently, however, if a researcher prefers to use other tools that are conceptually similar, they have to leave WORCS behind.
+We currently integrate WORCS into a modular framework of reproducibility tools for R [@repro], which will serve as the backend for WORCS.
+This ongoing integration into a unifying framework does not compromise the ease of use for the researcher.
+However, it gives them more flexibility to satisfy the demands of their research projects while meeting the TOP-guidelines.
+
 \newpage

### 2020-05-18 paper/Manuscript.Rmd
Integrate Aaron's edits

@@
 A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of open science.
-<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example here is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019.
-Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5.
-Peikert and Brandmaier's workflow builds upon the same general principles making it fully compatible with the solution proposed here.
-Their workflow aims to encompass even the most sophisticated analytic strategies and is hence relatively complex itself.
-WORCS greatly simplifies matters by assuming that no other software then R Packages are needed <!-- difference docker -->and that that a single data file is sufficient <!-- difference make -->.
-If both assumptions are met, and they routinely are, this eschews much overhead of Peikert and Brandmaier's workflow resulting in a much more lightweight approach to safeguard computational reproducibility under *most circumstances*.
-WORCS addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines, while being compatible with existing approaches.
+<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019.
+Because that workflow focuses on computational reproducibility, it only addresses TOP-guidelines 2, 3, and 5.
+Moreover, in striving to ensure strict computational reproducibility for even the most sophisticated analyses, Peikert and Brandmaier's workflow ends up being relatively complex to implement.
+WORCS is designed to be a simple and sufficient solution for projects that can be conducted entirely within R.
+It eschews much overhead of Peikert and Brandmaier's workflow, resulting in a much more lightweight approach to safeguard computational reproducibility under *most circumstances*.
+Importantly, WORCS builds upon the same general principles as Peikert and Brandmaier, which means the two workflows are compatible.
+Finally, WORCS addresses a unique issue not covered by other existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines, while being compatible with existing approaches.
 
@@
 These solutions differ in user-friendliness and effectiveness.
-If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine, that can be stored like a sort of time capsule and identically reinstated on different systems.
+If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine that can be stored like a sort of time capsule, and identically reinstated on different systems.
 This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data).
@@
 
+## Future developments
+
+The WORCS procedure and accompanying R-package provide a user-friendly and lightweight workflow for open, reproducible research in R, that meets all TOP-guidelines.
+It is beyond the scope of WORCS to support the incorporation of tools outside of the R-environment, or to containerize a project so that it can be identically reinstated on a different system or virtual machine.
+One important area of future development is to enable these extensions of the workflow.
+To this end, we are integrating WORCS into a modular framework of reproducibility tools for R [@repro].
+This will make it possible for researchers to enhance a WORCS project with strict reproducibility tools like 'Docker' and 'Make', or vice versa, to initiate a WORCS project in a strictly reproducible environment.
+This combines the strengths of WORCS with those of @peikertReproducibleDataAnalysis2019.
+
 <!-- Open peer review? -->
@@
 The accompanying R-package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.
-
-## Outlook
-
-<!-- not sure where this repro integration section belongs -->
-The `worcs`-package integrates several reproducibility tools into a convenient workflow by assuming reasonable standards that fulfil most researches needs.
-Currently, however, if a researcher prefers to use other tools that are conceptually similar, they have to leave WORCS behind.
-We currently integrate WORCS into a modular framework of reproducibility tools for R [@repro], which will serve as the backend for WORCS.
-This ongoing integration into a unifying framework does not compromise the ease of use for the researcher.
-However, it gives them more flexibility to satisfy the demands of their research projects while meeting the TOP-guidelines.
 

### 2020-05-18 paper/Manuscript.Rmd

--- 2020-05-17 paper/Manuscript.Rmd
@@
 A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of open science.
-<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example here is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019.
-Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5.
-Peikert and Brandmaier's approach overlaps, to some extent, with the solution proposed here, but focuses on achieving strict computational reproducibility by conducting analyses on a virtual machine, that can be stored like a sort of time capsule and identically reinstated on different systems.
-The workflow is excellent, but difficult to implement for non-programmers.
-WORCS instead uses a much more lightweight approach to safeguard computational reproducibility under *most circumstances*.
-This should be sufficient for most users, but can also be combined with Peikert and Brandmaier's approach if strict computational reproducibility is essential.
-WORCS addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines, while being compatible with existing approaches.
+<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019.
+Because that workflow focuses on computational reproducibility, it only addresses TOP-guidelines 2, 3, and 5.
+Moreover, in striving to ensure strict computational reproducibility for even the most sophisticated analyses, Peikert and Brandmaier's workflow ends up being relatively complex to implement.
+WORCS is designed to be a simple and sufficient solution for projects that can be conducted entirely within R.
+It eschews much overhead of Peikert and Brandmaier's workflow, resulting in a much more lightweight approach to safeguard computational reproducibility under *most circumstances*.
+Importantly, WORCS builds upon the same general principles as Peikert and Brandmaier, which means the two workflows are compatible.
+Finally, WORCS addresses a unique issue not covered by other existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines, while being compatible with existing approaches.
 
@@
 These solutions differ in user-friendliness and effectiveness.
-If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine.
+If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine that can be stored like a sort of time capsule, and identically reinstated on different systems.
 This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data).
-However, it is very difficult to set up.
 <!--For small projects, or novice users || Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project.

### 2020-05-18 paper/Manuscript.Rmd
marijn changes

@@
 Step 20 of the WORCS procedure suggests uploading the online version to a preprint server.
-This is important because most preprint servers - [at least the ones listed by OSF.io](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
+This is important because most preprint servers - at least those hosted by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
 This means that authors will receive credit for cited work; even if they are cited only in the online version.

### 2020-05-18 paper/Manuscript.Rmd
Incorporate changes

@@
 # Introduction
-<!--
-Barbara:
-A trick I learned in collaborative writing with git is to make sure each new sentence is on a new line.
-This way, you can effectively use a git diff to see changes in individual sentences instead of having to scour through an entire paragraph.
-I'll add this, but only in the paragraphs I've changed, so you do not get inundated with differences!
--->
+
 The (social) sciences are amidst a paradigm shift towards open science. 
@@
 
-<!-- LB: Although Open Science is advocated by many, there is no clear definition of Open Science (ref multiple definitions).
-An influential operationalisation of Open Science has been de TOP guidlines.
--->
-One of the most impactful contributions to open science has been the development of the TOP-guidelines [@nosekPromotingOpenResearch2015a], and the effort to urge publishers to adopt them.
+Although Open Science is advocated by many, there is no clear definition of Open Science (LOEK? ref multiple definitions).
+The TOP-guidelines have been influential as a concrete operationalisation of Open Science principles [@nosekPromotingOpenResearch2015a].
 These guidelines describe eight standards for open science: 1) Comprehensive citation of literature, data, materials, and methods; 2) sharing data, 3) sharing the code required to reproduce analyses, 4) sharing new research materials, and 5) sharing details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results.
@@
 Data sharing is important for computational reproducibility and secondary analysis.
-Computational reproducibility means that a third party can exactly recreate the results from the raw data, using the published analysis code.
+Computational reproducibility means that a third party can exactly recreate the results from the original data, using the published analysis code.
 Secondary analysis means that a third party can conduct sensitivity analyses, explore alternative explanations, or even use existing data to answer a different research question.
@@
 Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns.
-For example, the GDPR prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU.
-Different rules may apply to anonymized or pseudonimized data.
-Before sharing any data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.
+For example, one concern that applies when working with human participant data, is pseudonimization.
+It is crucial to pseudonimize data as soon as they are collected, by removing or deleting any sensitive personal information and contact details.
+Pseudononimized data can often be shared if participants have provided informed consent to that effect.
+Nevertheles, it is important to note that the European GDPR prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU, unless it offers an "adequate level of protection".
+Although different rules may apply to pseudonimized data, there are many repositories that are GDPR compliant, such as the European servers of the Open Science Framework.
+Different Universities, countries, and funding bodies also have their own repositories that are complient with local legistation.
+Before sharing any human participant data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.
 
@@
 can I indicate a point in my workflow (e.g.
-after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it.
-Any data processing steps should be documented in the code, instead of applied to the raw data.
-This way, the code will be a pipeline from the raw data to the published results.
+after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->
+If data cannot be shared, researchers should aim to safeguard the potential for computational reproducibility and secondary analysis as much as possible.
+The `worcs` package uses two solutions to accomplish this goal.
+The first solution is to publish a *checksum* of the original data file.
+Think of a checksum as a 32-character summary of the contents of a file, or alternatively, a "one word" description of the file.
+Any change to the file will result in a different checksum. 
+Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
+When data cannot be shared, the risk of fraud or abuse can be mitigated by publishing a checksum for the original data, in addition to the complete analysis code.
+Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditer that running the public analysis code on the private data results in the published results of their work.
 
-When data cannot be shared, researchers should aim to safeguard the potential for computational reproducibility and secondary analysis as much as possible.
-WORCS uses two solutions to accomplish this goal.
-The first solution is publishing a *checksum* of the raw data upon collection.
-A checksum is the result of an algorithm applied to a file, and represents the contents of this file in 32-characters.
-Any change to the underlying file results in a different checksum. Thus, one can use a checksum to verify that the contents of a file remain unchanged.
-If researchers publish the checksum of their raw data, in addition to the complete analysis code, then they can prove to an independent auditer that running the public analysis code on the private dataset with the correct checksum results in the published results of their work.
-
-The second solution is to share a synthetic dataset with similar characteristics to the real data.
-Synthetic data mimic the level of measurement and (conditional) distributions of the real data.
-Sharing synthetic data allows any third party to 1) verify that the published code runs, 2) debug the code, and 3) write valid code for alternative analyses.
-However, it is important to know that complex multivariate relationships present in the real data are often lost in synthetic data.
+The second solution implemented in `worcs` is to share a synthetic dataset with similar characteristics to the real data.
+Synthetic data mimic the level of measurement and (conditional) distributions of the real data [see @nowokSynthpopBespokeCreation2016].
+Sharing synthetic data allows any third party to 1) verify that the published code works, 2) debug the code, and 3) write valid code for alternative analyses.
+It is important to note that complex multivariate relationships present in the real data are often lost in synthetic data.
 Thus, findings from the real data might not be replicated in the synthetic data, and findings in the synthetic data *should not be substantively interpreted*.
@@
 A third party can write analysis code based on the synthetic data, and send it to the authors who evaluate it on the real data and send back the results.
-The ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) functionality on GitHub is a particularly user-friendly way to do this.
-<!--Barbara: I'm not sure this is the case.
-Pull requests are useful if someone wants to change or add to a project, but this would not involve a change to an existing project, but rather a new project/analysis with "old" data.
-I would remove this suggestion altogether.-->
 
-The procedure for data sharing in WORCS is as follows.
-First, researchers load the data into R.
-This process should be documented in the file `prepare_data.R`, which is automatically created by the WORCS-template.
-This file is intended to document any steps that must be taken to get your data *into* R, and that are not intended to be repeated when re-running the analyses.
-Use this file to report any preprocessing required to obtain a neat `data.frame`.
-For example, if the data was originally in SPSS format, this file might just contain the code required to read the SPSS file.
+### Processing data in WORCS
 
-Then, researchers must decide whether to use open or closed data, and call one of these convenience functions:
+Regardless of whether the data will be open or closed, it is important that the raw data will be left unchanged as much as possible.
+This eliminates human error.
+Any alterations to the data - including processing steps and even error corrections - should be documented in the code, instead of applied to the raw data.
+This way, the code will be a pipeline from the raw data to the published results.
+Every `worcs` project contains an empty R-script called `prepare_data.R`.
+As soon as the data are collected, researchers should use this file to document the steps necessary to load the data into R, pseudonimize it, and prepare it for analysis.
+For example, if the data was originally in SPSS format with IP addresses and GPS location, this file might just contain the code required to read the SPSS file, and to remove those columns of sensitive personal information.
+As soon as the data are pseudonimized and processed into a 'tidy' format [i.e., data that can be stored as a spreadsheet, see @wickhamTidyData2014], the researcher should version control some indelible record of the raw data, and avoid making any further changes to the version controlled file.
 
-* `open_data(data)`: This function stores the data object (a `data.frame` or `matrix`) in a spreadsheet file called `data.csv`, stores a checksum for the file, and creates a codebook.
-All of these files should now be (manually) added to the 'Git' repository, committed, and pushed to GitHub.
+The `worcs` package offers two functions for version controlling a record of the data: One for open, and one for closed data:
+
+* `open_data(data)`: This function stores the `data` object (a `data.frame` or `matrix`) in a spreadsheet file (called `data.csv` by default).
+This file should now be (manually) added to the 'Git' repository, committed, and pushed to GitHub.
 Once the GitHub repository is made public, these data will be open to the public.
-* `closed_data(data)`: This function stores the data in a local file called `data.csv`, which is added to the `.gitignore` file so it cannot be accidentally added to the 'Git' repository.
-The function also creates a synthetic dataset called `synthetic_data.csv`.
+Assume that, once you do this, the data cannot be un-shared.
+* `closed_data(data)`: This function also stores the data in a local file (called `data.csv` by default), but that file is added to the `.gitignore` file so it cannot be accidentally added to the 'Git' repository.
+The function also adds a checksum for the original data in the `.worcs` project file, and creates a synthetic dataset (called `synthetic_data.csv` by default).
 The user should commit all changed files and push them to GitHub.
-Once the GitHub repository is made public, people will have access to a checksum for the raw data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
+Once the GitHub repository is made public, people will have access to a checksum for the original data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
 
-<!--Barbara: Is it indeed the case that data by default is loaded as a .csv?
-How about (raw) text data; can worcs accommodate a .txt file, for instance (in case of text analysis)?
-Or should this all be done in prepare_data.R?-->
+The final line of the `prepare_data.R` file should thus be either `open_data()` or `closed_data()`.
+The analysis code, on the other hand, should begin with `load_data()`.
+This function loads the real data if it is available on the user's computer, and otherwise, loads the synthetic data.
+This will have the effect that third party users who copy the GitHub repository will automatically load the synthetic dataset, whereas the study authors, who have the real data stored locally, will automatically load the real data.
+This function makes it possible for reviewers, coauthors, and auditors, to write analysis scripts without requiring access to the original data.
+They can simply start the script with `load_data()`, and write their code based on the synthetic data.
+Then, they can submit their code to you - by email or as a ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) on GitHub.
+When you run their code on your system, `load_data()` will detect the original data, and use that to run their script.
+You can then return the results to the third party.
 
-Some users may intend to share their data through a platform with restricted access (like an institutional repository), but not publically through GitHub.
-In this case, it is recommended to use `closed_data()`, and to manually upload the original data `data.csv` to the restricted access platform.
-If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the GitHub repository.
+As of version 0.1.1, the `worcs` package only stores data as `.csv` files.
+As explained above, this is intentional, because `.csv` files are human- and machine readable, and because data typically need to be in a spreadsheet format for analysis.
+Many types of data can be represented as a spreadsheet - text corpora can have one row per document, and EEG- or EKG waveforms can have one row per measurement occasion.
+The `prepare_data.R` file should document any steps required to convert these data into tabular format.
+If it is data must be stored in a different file type anyway, readers are encouraged to follow the development of the `repro` package [@repro], which will enable such enhancements.
 
-The `prepare_data.R` file describes the data wrangling process to derive a spreadsheet ready for analysis.
-The final line of this file will often be `open_data()` or `closed_data()`.
-The code for analysis of the data, on the other hand, will typically begin with `load_data()`.
-This function loads the real data if it is available (i.e., the file `data.csv` exists, and the checksum is correct), and otherwise, loads the synthetic data.
-This will have the effect that third party users who copy the GitHub repository will automatically load the synthetic dataset, whereas the study authors, who have the real data stored locally, will automatically load the real data.
+Some users may intend to make their data openly available through a restricted access platform, such as an institutional repository, but not publically through GitHub.
+In this case, it is recommended to use `closed_data()`, and to manually upload the original `data.csv` file to the restricted access platform.
+If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the GitHub repository as an [Add-on](https://help.osf.io/hc/en-us/categories/360001550973-Add-ons).
 
@@
 4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
-4. Collect data and either add to repository, or register hash of raw datafile
+4. Collect data and either add to repository, or register hash of original datafile
 5.

### 2020-05-18 paper/Manuscript.Rmd
Another update

@@
 For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
-<!-- A more recent preprint reported that many journals do not explicitly state their policies about  klebelPeerReviewPreprint2020 -->
 At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
@@
 ## Prerequisites
-<!--Barbara: I'm moving up this section and editing it a bit, to introduce more explicitly that worcs was developed for R.-->
-WORCS has focused on developing a workflow for R, instead of for SPSS or Python.
-<!-- Barbara: same as for Git, I could not find a citation. Maybe cite the website: www.r-project.org -->
+
+Although the principles underlying this workflow are universal, WORCS has been optimized for R [@rcoreteamLanguageEnvironmentStatistical2020].
 Several arguments support the choice to work in R.
 First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science.
-Second, many tools for an open science workflow are already implemented in R.
-A vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
-Wrappers in R are often available for tools developed in other programming languages.
+Second, as of this writing, R is the most advanced programming language in terms of the implementation of tools required for an open science workflow.
+Third, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
+
 <!-- LB: Perhaps give some pointers to learn R (e.g. R4DataScience)-->
@@
 Finally, R is highly interoperable:
-Packages are available to load nearly every imaginable filetype, code written in other programming languages [including Python, @allaireMarkdownPythonEngine] can be called from R, and output can be written to many file types, including DOCX and PDF (in APA style), and HTML format.
+Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
+Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R [including C++, Fortran, and Python, @allaireMarkdownPythonEngine].
 
@@
 -->
-Many of the tools used in WORCS are embedded directly into the user interface of RStudio.
-RStudio automates and streamlines many tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
+Most of the tools used in WORCS are embedded directly into the user interface of RStudio.
+RStudio automates and streamlines tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
 This makes RStudio a comprehensive solution for open science research projects.
@@
 The `worcs` R-package installs a new RStudio project template.
-When a new project is initialized from this template, much of the bookkeeping required to set up an open science project is performed automatically.
+When a new project is initialized from this template, the bookkeeping required to set up an open science project is performed automatically.
 

### 2020-05-18 paper/Manuscript.Rmd
Mention github alternatives

@@
 
-Working with R is simplified immensely by using the RStudio integrated development engine (IDE) for R.
-<!-- Barbara: 
-https://support.rstudio.com/hc/en-us/articles/206212048-Citing-RStudio says:
-To cite RStudio in publications, you can get the latest citation information by running the command RStudio.Version() in a recent version of RStudio IDE.
-
-I did this, which gave:
-
-  @Manual{,
-    title = {RStudio: Integrated Development Environment for R},
-    author = {{RStudio Team}},
-    organization = {RStudio, Inc.},
-    address = {Boston, MA},
-    year = {2016},
-    url = {http://www.rstudio.com/},
-  }
--->
+Working with R is simplified immensely by using the RStudio integrated development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
 Most of the tools used in WORCS are embedded directly into the user interface of RStudio.
@@
 The first of these is *dynamic document generation*.
-Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses.
-The text is automatically formatted as an APA-style manuscript [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables.
-Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text.
-Although there is a slight learning curve to transitioning to DDG, the investment really pays off:
+Dynamic document generation (DDG) means writing scientific reports in a format
+that interleaves written reports with blocks of computer code used to conduct
+the analyses. The text is automatically formatted as an APA-style manuscript
+[thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and
+results of the code blocks are automatically generated and insterted in the
+text, or rendered as Figures and Tables. Dynamic document generation supercedes
+the classical approach of using separate programs to write prose and conduct
+analyses, and then manually copy-pasting analysis results into the text.
+Although there is a slight learning curve to transitioning to DDG, the
+investment really pays off:
 
-* The time saved from painstakingly copy-pasting output and manually formatting text soon outweighs the investment of switching to a new program
-* Human error in copy-pasting results is eliminated
-* When revisions require major changes to the analyses, all results, Figures and Tables are automatically updated
-* Flexibility in output formats means that a manuscript can be rendered to presentation format, or even to a website. 
-* Reproducibility is guaranteed, and can be verified by reviewers and/or readers, because the code is run each time the document is compiled.
+* Time saved from painstakingly copy-pasting output and manually formatting text
+soon outweighs the investment of switching to a new program
+* Human error in manually copying results is eliminated
+* When revisions require major changes to the analyses, all results, Figures and
+Tables are automatically updated
+* Flexibility in output formats means that a manuscript can be rendered to
+presentation format, or even to a website
+* Reproducibility is guaranteed, and can be verified by reviewers and/or
+readers, because the code is run each time the document is compiled.
 
-In sum, while writing academic papers in a programming environment might seem counter-intuitive at first, this approach is much more amenable to the needs of academics than most word processing software.
-It prevents mistakes, and saves time.
+In sum, while writing academic papers in a programming environment might seem
+counter-intuitive at first, this approach is much more amenable to the needs of
+academics than most word processing software. It prevents mistakes, and saves
+time. We recommend centering a research project around one dynamically generated
+(RMarkdown) document, which includes all analysis code. Longer scripts can be
+stored in `.R` files, and called from the main document using the `source()`
+function.
 
 The second solution is *version control*:
-Maintaining an indelible log of every change to all project files.
-Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax [see also @blischakQuickIntroductionVersion2016].
-Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions.
-WORCS instead uses the *explicit* version control software "Git". Git tracks changes to files, and stores these changes when the user makes a "commit".
+Maintaining an indelible log of every change to all project files. Version 
+control is a near-essential tool for scientific reproducibility, as anyone
+learns who has had the experience of accidentally deleting a crucial file, or of
+being unable to reproduce analyses because they ran some analyses interactively
+instead of using syntax [see also @blischakQuickIntroductionVersion2016].
+Many scientists use some form of *implicit* version control; for example, by
+renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"),
+tracking changes in word processing software, or using cloud hosting services
+that retain backups of previous versions.
+WORCS instead uses the *explicit* version control software 'Git'
+(www.git-scm.com). Git tracks
+changes to files, and stores these changes when the user makes a "commit" (a 
+snapshot of the version controlled files).
 <!--Barbara: I have looked for a way to cite Git but did not find one... 
-Seems good practice especially in this paper as an example to cite all software, but for Git I just cannot find the way.
-Maybe cite the website: www.git-scm.com -->
-Git retains a complete log of all commits, and users can compare changes between different commits, or go back to a previous version of the code (for example, after making a mistake, or to replicate a previous version of the results).
+Seems good practice especially in this paper as an example to cite all software,
+but for Git I just cannot find the way. Maybe cite the website: www.git-scm.com
+CJ: Other papers discussing scientific merits of Git also just give a URL -->
+Git retains a complete log of all commits, and users can compare changes between
+different commits, or go back to a previous version of the code (for example,
+after making a mistake, or to replicate a previous version of the results).
 A project version controlled with Git is called a "Repository", or Repo.
-It is also possible to prevent some files from being version controlled - such as privacy sensitive data.
-The `.gitignore` file is a list of files that should not be version-controlled, and thus exist only on the user's private computer.
+Git only version controls files explicitly committed by the user. Moreover, it
+is possible to prevent files from being version controlled - which is useful for
+privacy sensitive data.
+The `.gitignore` file lists files that should not be version-controlled.
+These files thus exist only on the user's private computer.
 
-<!--In statistical consultations, I often hear that researchers are unable to reproduce specific results because they only worked with a "very specific" version of the syntax - usually a combination of -->
-
-The functionality of Git is amplified by services such as GitHub (other services are available).
-GitHub is best understood as a cloud storage service with social networking functionality.
-The "cloud storage" aspect of GitHub works as follows: You can "clone" a local Git repository to the GitHub website, as a backup or research archive.
-The "social network" aspect comes into play when the project is set to "public": This allows other researchers to peruse the repository on the GitHub website to see how the work was done, clone it to their own computer to replicate the original work or apply the methods to their own data, open "Issues" to ask questions or give feedback on the project, or even send a "Pull request" with suggested changes to the text or code for your consideration.
-Git and GitHub shine as tools for collaboration, because different people can simultaneously work on different parts of a project, and their changes can be compared and automatically merged on the website.
-Even on solo projects, working with Git/GitHub has many benefits: Staying organized, being able to start a new study with a clone of an old, similar repository, or splitting off an "experimental branch" to try something new, while retaining the ability to revert to a previous state of the project, or to merge the experimental branch.
+The functionality of 'Git' is amplified by services such as 'GitHub'
+(https://github.com; other similar services include 'GitLab', https://gitlab.com
+and 'Bitbucket', https://bitbucket.org). 'GitHub' is best understood as a cloud
+storage service with social networking functionality. The cloud storage aspect
+of 'GitHub' works as follows: You can "clone" (copy) a local 'Git' repository to
+the 'GitHub' website, as a backup or research archive. The social network aspect
+comes into play when a repository on 'GitHub' is set to "public": This allows 
+other researchers to peruse the repository and see how the work was done, clone
+it to their own computer to replicate the original work or apply the methods to
+their own data, open "Issues" to ask questions or give feedback on the project,
+or even send a "Pull request" with suggested changes to the text or code for
+your consideration. 'Git' and 'GitHub' shine as tools for collaboration, because
+different people can simultaneously work on different parts of a project, and
+their changes can be compared and automatically merged on the website. Even on
+solo projects, working with 'Git'/'GitHub' has many benefits: Staying organized,
+being able to start a new study with a clone of an old, similar repository, or
+splitting off an "experimental branch" to try something new, while retaining the
+ability to revert to a previous state of the project, or to merge the 
+experimental branch.
 

### 2020-05-18 paper/Manuscript.Rmd
updatesss

@@
 The functionality of 'Git' is amplified by services such as 'GitHub'
-(https://github.com; other similar services include 'GitLab', https://gitlab.com
-and 'Bitbucket', https://bitbucket.org). 'GitHub' is best understood as a cloud
+(https://github.com. 'GitHub' is best understood as a cloud
 storage service with social networking functionality. The cloud storage aspect
@@
 splitting off an "experimental branch" to try something new, while retaining the
-ability to revert to a previous state of the project, or to merge the 
-experimental branch.
+ability to "revert" (return) to a previous state of the project, or to "merge"
+(incorporate) the experimental branch. Although this paper focuses on 'GitHub',
+`worcs` (as of version 0.1.2) supports alternative services such as 'GitLab'
+(https://gitlab.com) and 'Bitbucket' (https://bitbucket.org), and has a vignette
+detailing instructions for these platforms.
 
-The third solution is *dependency management*: Keeping track of exactly what software was used to conduct the analyses.
-<!-- Being able to replicate analyses requires not only access to the data and analysis code, but also to the software used to evaluate that code.--> At first glance, it might seem sufficient to state that analyses were conducted in *Program X*.
-However, every program is susceptible to changes and updates.
-R-packages in particular are updated frequently, because they are Open Source, and there is an active community of developers contributing functionality and bugfixes.
-Potentially, any update could change the results of the code.
-For example, R-version 3.6.0 fixed a bug in the random number generator.
-This means that any analyses that use random numbers, such as bootstrapping, cannot be strictly replicated from previous versions.
-Or, for example, R-version 4.0.0 changed a default parameter that affects how data are loaded from files.
+The third solution is *dependency management*: Keeping track of exactly what
+software was used to conduct the analyses. At first glance, it might seem
+sufficient to state that analyses were conducted in *Program X*. However, every
+program is susceptible to changes, updates, and bugfixes. R-packages in
+particular are updated frequently, because they are Open Source, and there is an
+active community of developers contributing functionality and bugfixes. 
+Potentially, any update could change the results of the code. For example, 
+R-version 3.6.0 fixed a bug in the random number generator. This means that any
+analyses that use random numbers, such as bootstrapping, cannot be strictly
+replicated from previous versions. Or, for example, R-version 4.0.0 changed a
+default parameter that affects how data are loaded from files.
 

### 2020-05-18 paper/Manuscript.Rmd
updae

@@
 after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->
-If data cannot be shared, researchers should aim to safeguard the potential for computational reproducibility and secondary analysis as much as possible.
-The `worcs` package uses two solutions to accomplish this goal.
+If data can be shared openly, it can simply be committed it to the 'Git'
+repository, along with the analysis code. This way, others can download the
+entire repository and reproduce the analyses from start to finish. 
+If data cannot be shared, researchers should aim to safeguard the potential for
+computational reproducibility and secondary analysis as much as possible. The
+`worcs` package uses two solutions to accomplish this goal.
 The first solution is to publish a *checksum* of the original data file.
-Think of a checksum as a 32-character summary of the contents of a file, or alternatively, a "one word" description of the file.
+Think of a checksum as a 32-character summary, or as a "one word" description,
+of the contents of a file.
 Any change to the file will result in a different checksum. 

### 2020-05-19 paper/Manuscript.Rmd

@@
     institution: "7"
+  - name: "Anna-Lena Lamprecht"
+    institution: "2,8"
     orcid: "0000-0002-5023-4601"
@@
 
-The TOP-guidelines are primarily geared towards top-down systemic change by journals, funders, and organizations.
+The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
 However, large organizations are slow to change.
-For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
+For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that
+they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
 At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
 Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support.
-The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement.
+The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement [REF Loek paper].
 This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
-This workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions, and where such requirements do exist, this workflow will help fulfill them.
+The WORCS workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions:
+If such requirements exist, this workflow will help fulfill them.
 
@@
 
+<!-- AL: Similar to my comment on the title, in this section it needs to make clear which scope of related work it covers.
+For example, in the life sciences there has also been a lot of work on practices for increasing reproducibility, with automated workflows, containerization, Research Objects, provenance protocols, and a lot more.
+I don't think it's necessary for this paper to discuss this, but that is under the provision that it defines its scope better.
+-->
 ## Prerequisites
@@
 Related: Can you make a schematic overview of the WORCS?overview-->
-Many solutions exist to ensure computational reproducibility.
-These solutions typically work by enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs.
-These solutions differ in user-friendliness and effectiveness.
-If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine that can be stored like a sort of time capsule, and identically reinstated on different systems.
-This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data).
-<!--For small projects, or novice users || Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project.
-This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space.
-Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library.
-In conclusion, `packrat` is somewhat unwieldy.
-The package `checkpoint` offers an extremely lightweight solution: Instead of storing all software used, it simply records the calendar date, and then installs the version of all required packages that was available on the central R repository `CRAN` on that date from an archival website.
-Thus, all that is needed to restore a project is the date on which it was conducted.
-This is very user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN` packages, and cannot mix package versions from different dates.
+Many solutions exist to ensure computational reproducibility. These solutions typically work by
+enveloping your research project in a distinct "environment" that only has access to programs that
+are explicitly installed, and maintaining a record of these programs. These solutions differ in
+user-friendliness and effectiveness. If strict reproducibility is required, one might use the
+aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all
+analyses inside a "container": An environment that behaves like a virtual computer that can be
+stored like a sort of time capsule, and identically reinstated on a user computer, or in the cloud.
+This is effective, and also preserves software outside the R environment (e.g., preprocessing tools
+for neuroimaging data). However, it can be difficult to set up for novice users - although a package
+is in development to facilitate this transition [@repo]. The cloud-based research collaboration platform
+'Code Ocean' is also based on Docker, for example. It is user-friendly, but it is not clear whether the
+["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/) is fully
+GDPR compliant - it seems to focus more on limiting liability than on compliance. This is a risk with
+any cloud-based solution. The advantage of using a local system is that data sharing is optional, and
+WORCS capitalizes on this flexibility.
+
+A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software
+used in a project to a dedicated library that can be stored and copied with the project. This solution
+is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages
+must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space.
+Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address
+the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is
+somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing
+all software used, it simply records the calendar date, and then installs the version of all required
+packages that was available on the central R repository `CRAN` on that date from an archival website.
+Thus, all that is needed to restore a project is the date on which it was conducted. This is very
+user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN`
+packages, and cannot mix package versions from different dates.
 
@@
 
-It is easiest to mark the distinction between essential and non-essential references from the start, instead of going back to cut references prior to publication.
-Therefore, the `worcs` package distinguishes between the traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference.
+It is easiest to mark the distinction between essential and non-essential references from the start, instead of going
+back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the
+traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference.
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
-<!-- Barbara: I see no difference between these two symbols...-->
 
@@
 Once the GitHub repository is made public, people will have access to a checksum for the original data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
+<!-- Maybe remark that in the "closed" case, a the data should be backupped somewhere else, as GitHub will not do the job in this setting. -->
 

### 2020-05-19 paper/Manuscript.Rmd
Add R for data science book

@@
 Second, as of this writing, R is the most advanced programming language in terms of the implementation of tools required for an open science workflow.
-Third, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
-
-<!-- LB: Perhaps give some pointers to learn R (e.g. R4DataScience)-->
-Third, at the time of writing, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
+Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
+Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
 Finally, R is highly interoperable:
@@
 Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R [including C++, Fortran, and Python, @allaireMarkdownPythonEngine].
+There are excellent free resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].
 

### 2020-05-19 paper/Manuscript.Rmd
some text on WORCS and DORA and FAIR

@@
 
+WORCS encourages and simplifies the adoption of Open Science principles in daily scientific work.
+It helps to make all research objects that are created through the scientific process (data and code, but also manuscripts) open and publicly assessible.
+This enables other researchers to reproduce results, but also provides a proper basis for research evaluation according to the San Francisco Declaration on Research Assessment (DORA) [https://sfdora.org/], which is increasingly adopted by research institutions around the globe.
+For example, reviewers of grant proposals have immediate access to previous work of a researcher or a group of researchers, and can assess research quality based on content rather than relying on spurious surrogate indicators like personal h-indexes, journal impact factors and conference rankings.
+The detailed version control and commit tracking of github and similar platforms furthermore make it possible to assess the contributions made by different researchers, which can for example play a role in hiring or promotion procedures.
+
+The FAIR Guiding Principles for scientific data management and stewardship [https://doi.org/10.1038/sdata.2016.18] advocate that all digital research objects should be Findable, Accessible, Interoperable and Reusable.
+They are increasingly adopted by research organizations and funders as a requirement towards more open, transparent and reproducible science.
+The WORCS workflow includes a step to publish the original raw data (if possible) or alternatively a synthetic data set in the (public) project repository, thus making the research data or a workable substitute accessible.
+WORCS uses CSV as an open and widely used, tabular format for data representation, thus enabling interoperability with other data sets and software.
+The R scripts for analyzing the data are naturally part of the same repository, facilitating reproducibility of obtained results by other researchers.
+
+Next to data, also research software should be FAIR [https://doi.org/10.3233/DS-190026].
+WORCS itself sets a good example and follows the recommendations for FAIR research software [https://fair-software.eu/]: It is hosted on a public repository with version control (github), it has a licence (GPL v3.0), it is registered in a community registry (CRAN), it enables the citation of the software (<!-- this paper.
+anything else?
+consider adding a cff file -->), and a software quality checklist was used for its development (<!-- which one?
+-->).
+By connecting the created R projects to github by default, it also guides its users through the first steps towards making their software FAIR, and lowers the threshold for taking the remaining steps.
+
 \newpage

### 2020-05-19 paper/Manuscript.Rmd
mention gitlab/bitbucket

@@
 The functionality of 'Git' is amplified by services such as 'GitHub'
-(https://github.com. 'GitHub' is best understood as a cloud
-storage service with social networking functionality. The cloud storage aspect
+(https://github.com). 'GitHub' is best understood as a cloud storage service
+with social networking functionality. The cloud storage aspect
 of 'GitHub' works as follows: You can "clone" (copy) a local 'Git' repository to
@@
 (incorporate) the experimental branch. Although this paper focuses on 'GitHub',
-`worcs` (as of version 0.1.2) supports alternative services such as 'GitLab'
-(https://gitlab.com) and 'Bitbucket' (https://bitbucket.org), and has a vignette
-detailing instructions for these platforms.
+the two main alternative 'Git' cloud storage services, 'GitLab'
+(https://gitlab.com) and 'Bitbucket' (https://bitbucket.org), can interchangeably
+be used in `worcs` projects (as of version 0.1.2). The package has a 
+[vignette]() with additional instructions for using these platforms.
 

### 2020-05-19 paper/Manuscript.Rmd
finish writing FAIR

--- 2020-05-19 paper/Manuscript.Rmd
@@
 ability to "revert" (return) to a previous state of the project, or to "merge"
-(incorporate) the experimental branch. Although this paper focuses on 'GitHub',
-`worcs` (as of version 0.1.2) supports alternative services such as 'GitLab'
-(https://gitlab.com) and 'Bitbucket' (https://bitbucket.org), and has a vignette
-detailing instructions for these platforms.
+(incorporate) the experimental branch. This paper focuses on 'GitHub', which is
+the most widely used 'Git' cloud storage services. However, as of version 0.1.2,
+`worcs` projects can interchangeably use the two main alternative services,
+'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). The
+package has a [vignette]() with additional instructions for using these platforms.
 
@@
 
+WORCS encourages and simplifies the adoption of Open Science principles in daily scientific work.
+It helps researchers make all research output created throughout the scientific process - not just manuscripts, but data, code, and methods - open and publicly assessible.
+This enables other researchers to reproduce results, facilitates cumulative science by allowing others to make direct use of these research objects.
+
+Sharing all research output also provides a proper basis for research evaluation according to the San Francisco Declaration on Research Assessment (DORA; https://sfdora.org/), which increasingly plays a role in grant funding and hiring or promotion procedures.
+Direct access to research objects allows funding bodies and hiring committies to evaluate research quality based on content rather than relying on spurious surrogate indicators like journal impact factors, conference rankings, and h-indexes.
+The detailed version control and commit tracking of github and similar platforms furthermore make it possible to assess the relative contributions made by different researchers.
+
+The FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a] advocate that research data should be 
+Findable, Accessible, Interoperable and Reusable. Moreover, these principles are
+increasingly applied to other types of research output 
+[@lamprechtFAIRPrinciplesResearch2019]. Research
+organizations and funders use the FAIR principles as standard for open, 
+transparent and reproducible science.
+The WORCS workflow helps users meet the FAIR principles by making all research objects -
+including (origina) data and code - accessible in a (public) repository. 
+The `worcs` package additionally
+facilitates meeting the FAIR principles in several ways. First, each `worcs` project has a YAML file
+(https://yaml.org/) that lists the research objects in the project, which will allow web crawlers 
+to index `worcs` projects. Second, a codebook is automatically generated for each data file stored
+using `open_data()` or `closed_data()`. We encourage users to elaborate on the default codebook by adding
+variable description and categories, which might - in the future - enable indexing data by topic area. 
+Relatedly, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/].
+It is hosted on a public version controlled repository ('GitHub'), it has an open source licence (GPL v3.0), is registered in a community registry (CRAN), it enables the citation of the software (using the `citation("worcs")` command, or by citing this paper), and followed the CII Best Practices software quality checklist during development.
+
 \newpage

### 2020-05-19 paper/Manuscript.Rmd
Rewrite intro

@@
 
-The (social) sciences are amidst a paradigm shift towards open science. 
-Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [as discussed by @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
-However, open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012].
-Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012].
-Capitalizing on these advances has the potential to accellerate scientific progress [see also @coyneReplicationInitiativesWill2016].
+The (social) sciences are arguably past the tipping point of the paradigm shift
+towards open science.
+Several immediate causes stirred up support for this transition; including
+highly publicized cases of scientific fraud [as discussed by
+@leveltFailingScienceFraudulent2012], increasing awareness of questionable
+research practices and their consequences for the credibility of findings [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
+[@shroutPsychologyScienceKnowledge2018].
+However, open science is not merely a cure for this crisis - it is also an
+opportunity [@adolphOpenBehavioralScience2012].
+Open science helps researchers to more easily conduct reliable, cumulative, and 
+collaborative science [@nosekScientificUtopiaOpening2012].
+Capitalizing on these advances has the potential to accellerate scientific
+progress [see also @coyneReplicationInitiativesWill2016].
 
-This paper introduces WORCS: A Workflow for Open Reproducible Code in Science.
-WORCS is a lightweight approach to open science and computational reproducibility.
-It provides a step-by-step procedure that researchers can follow to make an entire research project Open and Reproducible.
-WORCS is based on best practices, and can be used either in parallel to, or in absence of, top-down requirements by journals.
-The main goal of WORCS is to lower the threshold for adopting an open science workflow by providing the present tutorial, along with an R-package with user-friendly support functions, and an RStudio project template.
+Many individual researchers are motivated to adopt best practices for open
+science and enjoy these benefits. And yet, the question of "how" to 
+adopt open science practices can be daunting. Making the transition effectively
+requires researchers to become knowledgeable about different open science 
+challenges and solutions, and to become proficient with new and unfamiliar tools.
+This paper is designed to ease that transition by providing a simple workflow 
+that meets most requirements open science, based on best practices, and 
+supported by user-friendly software in R: The Worfkflow for Open Reproducible 
+Code in Science (WORCS). WORCS is a lightweight approach to open science and
+computational reproducibility. The workflow (Figure \@ref(fig:workflow)) 
+constitutes a step-by-step procedure that researchers can follow to make a 
+research project Open and Reproducible. WORCS is compatible with open science
+requirements already implemented by journals and institutions, and will help 
+fulfill them. It can also be used to produce work in accordance with best 
+practices in the absence of top-down support. The WORCS workflow is supported by
+an R-package with user-friendly support functions, and an RStudio project
+template that is, effectively, a "one click solution". This paper introduces
+the WORCS workflow, discusses how it meets best practices for open science, and
+illustrates the use of the `worcs` package.
 
@@
 
-The remainder of the paper is structured as follows.
+<!-- AL: The remainder of the paper is structured as follows.
 ...
-<!-- AL: Or some variation of this obligatory end-of-intro paragraph.
+Or some variation of this obligatory end-of-intro paragraph.
 As the paper touches on a lot of aspects, from principal to very practical, it would be very useful to provide some guidance here.
--->
+CJ: I was always taught that you should not explain your structure in the paper - you should just structure the paper. If the structure is unclear, make it better. -->
+<!--openness, transparency, and reproducible science -->
 
-# Defining open science
+# Meeting the requirements of Open Science
 
-Although Open Science is advocated by many, there is no clear definition of Open Science (LOEK? ref multiple definitions).
-The TOP-guidelines have been influential as a concrete operationalisation of Open Science principles [@nosekPromotingOpenResearch2015a].
-These guidelines describe eight standards for open science: 1) Comprehensive citation of literature, data, materials, and methods; 2) sharing data, 3) sharing the code required to reproduce analyses, 4) sharing new research materials, and 5) sharing details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results.
-WORCS defines the goals of open science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the subject of a research project, not to its execution.
-We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).
+Although Open Science is advocated by many, it does not have a unitary definition.
+Instead, several guidelines exist for practices that are considered "open".
+The TOP-guidelines are one of the most influential concrete operationalisations
+of Open Science principles [@nosekPromotingOpenResearch2015a].
+These guidelines describe eight standards for open science: 1) Comprehensive
+citation of literature, data, materials, and methods; 2) sharing data, 3)
+sharing the code required to reproduce analyses, 4) sharing new research
+materials, and 5) sharing details of the design and analysis; 6)
+pre-registration of studies before data collection, and 7) pre-registration of
+the analysis plan prior to analysis; and 8) replication of published results.
+WORCS defines the goals of open science in terms of these guidelines, and is
+designed to facilitate meeting each of these guidelines, with one exception: We
+do not address replication of published results, because replication relates to
+the subject of a research project, not to its execution.
+We group the remaining seven guidelines into three categories: Citation (1),
+sharing (2-5), and preregistration (6-7).
+
+WORCS meets the TOP-guidelines in a way that is amenable to the 
+FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a]. These principles advocate that research
+data should be Findable, Accessible, Interoperable and Reusable. However, they are
+increasingly applied as a standard for 
+to other types of research output as well
+[@lamprechtFAIRPrinciplesResearch2019].
+WORCS helps users meet the FAIR principles by making all research objects -
+including (origina) data and code - accessible in a (public) repository. 
+The `worcs` package additionally
+facilitates meeting the FAIR principles in several ways. First, each `worcs`
+project has a YAML file (https://yaml.org/) that lists the research objects in
+the project, which will allow web crawlers 
+to index `worcs` projects. Second, a codebook is automatically generated for
+each data file stored
+using `open_data()` or `closed_data()`. We encourage users to elaborate on the
+default codebook by adding
+variable description and categories, which might - in the future - enable
+indexing data by topic area. 
+Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/].
+It is hosted on a public version controlled repository ('GitHub'), it has an open source licence (GPL v3.0), is registered in a community registry (CRAN), it enables the citation of the software (using the `citation("worcs")` command, or by citing this paper), and followed the CII Best Practices software quality checklist during development.
+
+Sharing all research output also provides a proper basis for research evaluation
+according to the San Francisco Declaration on Research Assessment (DORA;
+https://sfdora.org/), which plays an increasing role in grant funding, hiring, 
+and promotion procedures. Direct access to research objects allows stakeholders
+to evaluate research quality based on content rather than relying on spurious
+surrogate indicators like journal impact factors, conference rankings, and
+h-indexes. The detailed version control and commit tracking of 'GitHub' and
+similar platforms furthermore make it possible to assess the relative
+contributions made by different researchers.
 
@@
 they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
-At the same time, there is substantial momentum for grassroots change [@nosekScientificUtopiaOpening2012].
-Many individual researchers are willing to adopt best practices for open science, even in the absence of institutional support.
-The formation of "Open Science Communities" at universities around the globe illustrates this grassroots movement [REF Loek paper].
-This paper is designed as an instrument for grassroots change, that can help individual scientists to meet the principles of open science.
-The WORCS workflow can be used regardless of any top-down open science requirements already implemented by journals and institutions:
-If such requirements exist, this workflow will help fulfill them.
+At the same time, <!-- edit -->
+The formation of "Open Science Communities" at universities around the globe
+illustrates the momentum of this grassroots movement
+[@brinkmanWidescaleAdoptionOpen2020].
 
@@
 
-Sharing all research output also provides a proper basis for research evaluation according to the San Francisco Declaration on Research Assessment (DORA; https://sfdora.org/), which increasingly plays a role in grant funding and hiring or promotion procedures.
-Direct access to research objects allows funding bodies and hiring committies to evaluate research quality based on content rather than relying on spurious surrogate indicators like journal impact factors, conference rankings, and h-indexes.
-The detailed version control and commit tracking of github and similar platforms furthermore make it possible to assess the relative contributions made by different researchers.
-
-The FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a] advocate that research data should be 
-Findable, Accessible, Interoperable and Reusable. Moreover, these principles are
-increasingly applied to other types of research output 
-[@lamprechtFAIRPrinciplesResearch2019]. Research
-organizations and funders use the FAIR principles as standard for open, 
-transparent and reproducible science.
-The WORCS workflow helps users meet the FAIR principles by making all research objects -
-including (origina) data and code - accessible in a (public) repository. 
-The `worcs` package additionally
-facilitates meeting the FAIR principles in several ways. First, each `worcs` project has a YAML file
-(https://yaml.org/) that lists the research objects in the project, which will allow web crawlers 
-to index `worcs` projects. Second, a codebook is automatically generated for each data file stored
-using `open_data()` or `closed_data()`. We encourage users to elaborate on the default codebook by adding
-variable description and categories, which might - in the future - enable indexing data by topic area. 
-Relatedly, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/].
-It is hosted on a public version controlled repository ('GitHub'), it has an open source licence (GPL v3.0), is registered in a community registry (CRAN), it enables the citation of the software (using the `citation("worcs")` command, or by citing this paper), and followed the CII Best Practices software quality checklist during development.
-
 \newpage

### 2020-05-19 paper/Manuscript.Rmd
Rewrite paragraph about existing approaches

@@
 
-## Goal of this paper
+<!--## Goal of this paper
 
@@
 At the same time, <!-- edit -->
-The formation of "Open Science Communities" at universities around the globe
+<!--The formation of "Open Science Communities" at universities around the globe
 illustrates the momentum of this grassroots movement
-[@brinkmanWidescaleAdoptionOpen2020].
+[@brinkmanWidescaleAdoptionOpen2020]. -->
 
@@
 
-There have been several previous efforts to promote grass-roots adoption of open science principles.
-Each of these efforts has a different scope, strengths, and limitations that set it apart from WORCS.
-For example, there are "signalling solutions"; guidelines to structure and incentivize disclosure about open science practices.
-Specifically, @aalbersbergMakingScienceTransparent2018 suggested publishing a "TOP-statement" as supplemental material, which discloses the authors' adherence to open science principles.
-Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based Transparency Checklist that authors can complete online to generate a report.
-Such signalling solutions are very easy to adopt, and they address TOP-guidelines 1-7.
-Many journals now also offer authors the opportunity to earn "badges" for adhering to open science guidelines [@kidwellBadgesAcknowledgeOpen2016].
-These signalling solutions help structure authors' disclosures about, and incentivize adherence to, open science practices.
+There have been several previous efforts to promote grass-roots adoption of open
+science principles. Each of these efforts has a different scope, strengths, and
+limitations that set it apart from WORCS. For example, there are "signalling
+solutions"; guidelines to structure and incentivize disclosure about open
+science practices. Specifically, @aalbersbergMakingScienceTransparent2018
+suggested publishing a "TOP-statement" as supplemental material, which discloses
+the authors' adherence to open science principles. Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based
+Transparency Checklist that authors can complete online to generate a report.
+Such signalling solutions are very easy to adopt, and they address
+TOP-guidelines 1-7. Many journals now also offer authors the opportunity to earn
+"badges" for adhering to open science guidelines
+[@kidwellBadgesAcknowledgeOpen2016]. These signalling solutions help structure
+authors' disclosures about, and incentivize adherence to, open science practices.
 
-A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of open science.
-<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019.
-Because that workflow focuses on computational reproducibility, it only addresses TOP-guidelines 2, 3, and 5.
-Moreover, in striving to ensure strict computational reproducibility for even the most sophisticated analyses, Peikert and Brandmaier's workflow ends up being relatively complex to implement.
-WORCS is designed to be a simple and sufficient solution for projects that can be conducted entirely within R.
-It eschews much overhead of Peikert and Brandmaier's workflow, resulting in a much more lightweight approach to safeguard computational reproducibility under *most circumstances*.
-Importantly, WORCS builds upon the same general principles as Peikert and Brandmaier, which means the two workflows are compatible.
-Finally, WORCS addresses a unique issue not covered by other existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines, while being compatible with existing approaches.
+A different class of solutions instead focuses on the practical issue of *how*
+researchers can meet the requirements of open science.
+<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example is the workflow for
+reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
+excellent workflow strives to ensure strict computational reproducibility for
+even the most sophisticated analyses. It has some limitations, however: In
+focusing on reproducibility, however, it only addresses TOP-guidelines 2, 3, and
+5. Moreover, in aiming for strict computational reproducibility it ends up being
+relatively complex to implement. WORCS builds upon the same general principles
+as Peikert and Brandmaier, which means the two workflows are compatible. What 
+sets WORCS apart is that is an easy and sufficient solution for projects that
+can be conducted entirely within R, resulting in a much more lightweight
+approach to safeguard computational reproducibility under *most circumstances*.
+WORCS also addresses a unique issue not covered by other existing solutions,
+namely to provide a workflow most conducive to satisfying the TOP-guidelines and
+FAIR principles, while being compatible with existing requirements.
 

### 2020-05-19 paper/Manuscript.Rmd
Update writing

@@
 
-Although the principles underlying this workflow are universal, WORCS has been optimized for R [@rcoreteamLanguageEnvironmentStatistical2020].
+Although the principles underlying this workflow are universal, WORCS has been
+designed for R [@rcoreteamLanguageEnvironmentStatistical2020].
 Several arguments support the choice to work in R.
-First, R and all of it extensions are free and Open Source, and Open Source software should be the tool of choice for open science.
-Second, as of this writing, R is the most advanced programming language in terms of the implementation of tools required for an open science workflow.
+First, R and all of it extensions are free and open source software, which
+should be the tool of choice for open science.
+Second, as of this writing, R is the most advanced programming language in terms
+of the implementation of tools required for an open science workflow.
 Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
@@
 Another important feature of RStudio is project management.
-A project bundles writing, analyses, data, references, etcetera, into a self-contained folder, that can be uploaded entirely to GitHub and downloaded by future users.
-The `worcs` R-package installs a new RStudio project template.
-When a new project is initialized from this template, the bookkeeping required to set up an open science project is performed automatically.
+A project bundles writing, analyses, data, references, etcetera, into a
+self-contained folder, that can be uploaded entirely to 'GitHub' and downloaded
+by future users. The `worcs` R-package installs a new RStudio project template.
+When a new project is initialized from this template, the bookkeeping required
+to set up an open science project is performed automatically. The `worcs` 
+package has a [vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)
+on how to set up your system with all of the required tools.
 
@@
 WORCS relies on several free, open source software solutions.
-The first of these is *dynamic document generation*.
-Dynamic document generation (DDG) means writing scientific reports in a format
-that interleaves written reports with blocks of computer code used to conduct
-the analyses. The text is automatically formatted as an APA-style manuscript
-[thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and
-results of the code blocks are automatically generated and insterted in the
-text, or rendered as Figures and Tables. Dynamic document generation supercedes
+The first is *dynamic document generation* (DDG): Writing scientific reports in
+a format that interleaves written reports with blocks of code used to conduct
+the analyses. The text is automatically formatted; the `worcs` package can 
+output manuscripts as an APA-style paper [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], or as a host of other scientific formats 
+[see @allaireRticlesArticleFormats2020], or as a plain 'Markdown' format for
+'GitHub'. When the text is formatted, the results of the code blocks are
+automatically generated and insterted in the text, or rendered as Figures and
+Tables. Dynamic document generation supercedes
 the classical approach of using separate programs to write prose and conduct
@@
 snapshot of the version controlled files).
-<!--Barbara: I have looked for a way to cite Git but did not find one... 
-Seems good practice especially in this paper as an example to cite all software,
-but for Git I just cannot find the way. Maybe cite the website: www.git-scm.com
-CJ: Other papers discussing scientific merits of Git also just give a URL -->
 Git retains a complete log of all commits, and users can compare changes between
@@
 active community of developers contributing functionality and bugfixes. 
-Potentially, any update could change the results of the code. For example, 
+Potentially, any update could change the results of the code. <!--For example, 
 R-version 3.6.0 fixed a bug in the random number generator. This means that any
@@
 replicated from previous versions. Or, for example, R-version 4.0.0 changed a
-default parameter that affects how data are loaded from files.
+default parameter that affects how data are loaded from files.-->
 
@@
 is in development to facilitate this transition [@repo]. The cloud-based research collaboration platform
-'Code Ocean' is also based on Docker, for example. It is user-friendly, but it is not clear whether the
+'Code Ocean' is also based on Docker. It is very user-friendly, but it is not clear whether the
 ["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/) is fully
 GDPR compliant - it seems to focus more on limiting liability than on compliance. This is a risk with
-any cloud-based solution. The advantage of using a local system is that data sharing is optional, and
-WORCS capitalizes on this flexibility.
+any cloud-computing solution.
 
-A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software
-used in a project to a dedicated library that can be stored and copied with the project. This solution
-is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages
+For analyses conducted entirely in R, a slightly less sophisticated solution is offered by the R-package `packrat`, which installs all packages
+used in a project to a dedicated library that can be stored and copied with the project. This has the disadvantage that for each project, all packages
 must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space.
-Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address
+Moreover, 'GitHub' is not designed to store your entire project library, so `packrat` does not address
 the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is
 somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing
-all software used, it simply records the calendar date, and then installs the version of all required
+all R-packages used, it simply records the calendar date, and when another user loads your project, it installs the version of all required
 packages that was available on the central R repository `CRAN` on that date from an archival website.
-Thus, all that is needed to restore a project is the date on which it was conducted. This is very
-user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN`
+Thus, all that is needed to restore a project is the date on which it was conducted. The main limitation is that this solution is limited to `CRAN`
 packages, and cannot mix package versions from different dates.
 
-A recently developed package strikes a great balance between user-friendliness and effectiveness: `renv`, developed by the team behind RStudio.
-This package maintains a text-based log file of which packages are used, in what versions, and where they were installed from (e.g., CRAN, Bioconductor, GitHub).
-Being text-based, this log can be version controlled with Git.
-On a local computer, `renv` installs all packages used in a cache that is shared between all projects that use `renv`.
-Thus, if one of your projects uses a version of a package that is already in the cache, it does not need to be installed again.
-This overcomes the limitations of long installation times and large space requirements of `renv`.
-When someone else loads your project, `renv` will install all of the required packages onto a cache on their computer.
-Whereas `packrat` copies the source code of all packages, `renv` obtains the packages from their original repositories.
-Thus, as long as these repositories used maintain their archives, this solution is very lightweight and effective.
-It is not foolproof, but some have argued that all solutions for computational reproducibility have a limited shelf life [@brownHowLearnedStop2017].
-Ultimately, the best solution might be to use a good-enough solution, like `renv`, and acknowledge that all code requires some maintenance if you want to reproduce it in the future.
+A recently developed package strikes a great balance between user-friendliness 
+and effectiveness: `renv`, developed by the team behind 'RStudio'. This package
+maintains a text-based, human-readable log of the packages used, their version
+numbers, and where they were installed from (e.g., 'CRAN', 'Bioconductor', 'GitHub').
+This text-based log can be version controlled with 'Git'. `renv` installs all
+packages in a cache that is shared between all projects that use `renv`. Thus,
+if one of your projects uses a package that is already in the cache, it does not
+need to be installed again. This overcomes the limitations of long installation
+times and large space requirements of `packrat`. When someone else loads your 
+project, `renv` will install all of the required packages onto a cache on their
+computer. Whereas `packrat` copies the source code of all packages, `renv`
+obtains the packages from their original repositories. Thus, as long as these
+repositories are maintained, this solution is very lightweight and effective.
+It is not foolproof, but all solutions for computational
+reproducibility have a limited shelf life [@brownHowLearnedStop2017].
+Ultimately, the best solution might be to use a good-enough solution, like
+`renv`, and acknowledge that all code requires some maintenance if you want to
+reproduce it in the future.
 
-## Considering filetypes
+## Text-based files are better
+
 A key consideration when developing a research project is what filetypes to use.
-A case can be made to use text-based files only (or as much as possible), instead of binary files.
-Text-based files can be read by machines and humans alike, without requiring special (licensed) software.
-Binary files, such as Word (`.docx`) or SPSS (`.sav`, `.spo`),  must be decoded by software first.
-Git is designed to keep track of changes to the content of text-based files.
-Each change reflects a word or line of code added or removed.
-Binary files, on the other hand, should not be version controlled with Git, as individual changes in the encrypted file are uninterpretable.
-Moreover, text-based files are often smaller than binary files, which means they take up less space on cloud hosting services, such as GitHub. Uploading large binary files is frowned upon.
-GitHub also allows visitors to peruse text-based files online, and conveniently renders certain filetypes.
-For example, GitHub highlights syntax in `.R` and `.Rmd` files, and displays `.csv` files as spreadsheets.
+A case can be made to use text-based files only (or as much as possible),
+instead of binary files.
+Text-based files can be read by machines and humans alike.
+Binary files, such as 'Word' (`.docx`) or 'SPSS' (`.sav`, `.spo`) files,
+must be decoded by (commercial) software first.
+'Git' is designed to track any changes to the content of text-based files.
+Each change reflects a line of text added or removed. This means that, when
+writing a version-controlled paper in plain-text, it is recommended to insert a
+line break after every sentence (instead of after every paragraph). That way,
+the change log will indicate which specific sentence was edited.
+Binary files should not be version controlled with 'Git',
+as changes in the encrypted file are uninterpretable. Binary files are often 
+also larger than text-based files, which means they take up more space on cloud
+hosting services, such as 'GitHub'. For those reasons, uploading large binary
+files to these platforms is frowned upon.
+'GitHub' does allow visitors to peruse text-based files online, and
+conveniently renders certain filetypes: For example, 'GitHub' displays `.md`
+(Markdown) files as web pages, and `.csv` files as spreadsheets. This is why
+`worcs` uses Markdown and `.csv` extensively.
 

### 2020-05-19 paper/Manuscript.Rmd
Last update of 19-5

@@
 
-<!--Barbara: Maybe write here something about writing manuscripts on a line-by-line basis, so that version control can keep track of changes more accurately than if a whole paragraph is identified as altered?-->
-
-## Open science as opportunity
-
-<!-- LB:  this section comes at a weird place in the flow of the paper. 
-Barbara: agreed! I would remove this altogether.-->
-Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct.
-Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.
-
 # How WORCS relates to the TOP-guidelines
@@
 
-TOP encourages comprehensive citation of literature, data, materials, methods, and software.
-In principle, researchers can meet this requirement by simply citing every reference used.
-Unfortunately, citation of data and software is less commonplace than citation of literature and materials.
-Crediting these resources is important, because it incentivizes data sharing and the development of open-source software, and supports the open science efforts of others.
+TOP encourages comprehensive citation of literature, data, materials, methods,
+and software. In principle,
+researchers can meet this requirement by simply
+citing every reference used. Unfortunately, citation of data and software is
+less commonplace than citation of literature and materials. Crediting these
+resources is important, because it incentivizes data sharing and the development
+of open-source software, supports the open science efforts of others, and helps
+researchers receive credit for *all* of their research output, in line with
+[DORA](https://sfdora.org/).
 
-<!--Various initiatives are underway to innovate how scientific output is measured and valued. https://www.nwo.nl/beleid/statement+waarderen+en+belonen+van+wetenschappers --> 
-To facilitate citing datasets, researchers sometimes publish *data papers*; documents that detail the procedure, sample, and codebook.
-Specialized journals, such as the [*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/), aid in the publication of these data papers.
-For smaller projects, researchers sometimes publish their data in an online repository - like GitHub - along with a text file with the preferred citation for the data (which can be a substantive paper), and the license that applies to the data, such as Creative Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or [BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/).
-When in doubt, one can always contact the data creators and ask what the preferred citation is.
+To facilitate citing datasets, researchers sometimes publish *data papers*;
+documents that detail the procedure, sample, and codebook. Specialized journals,
+such as the
+[*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/),
+aid in the publication of these data papers. For smaller projects, researchers
+sometimes publish their data in an online repository - like 'GitHub' - along
+with a text file with the preferred citation for the data (which can be a
+substantive paper), and the license that applies to the data, such as Creative
+Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or
+[BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/). When in doubt,
+one can always contact the data creators and ask what the preferred citation is.
 
@@
 https://citation-file-format.github.io/ is a file format to include software citation information into source code repositories.-->
-When using R, software (R packages) can be conveniently cited by calling `citation("packagename")`.
-This returns an APA-style reference, and a BibTeX entry.
-However, users of the WORCS project template for RStudio do not need to manually cite R-packages.
+When using R, software (R packages) can be conveniently cited by calling
+`citation("packagename")`. This returns an APA-style reference, and a BibTeX
+entry. In addition to the manual citation of important packages, the `worcs`
+project template for RStudio offers a comprehensive record of all dependencies.
 WORCS generates a template manuscript in APA6 style, based on the `papaja` package [@austPapajaPrepareReproducible2020], which automatically cites all packages used.

### 2020-05-20 paper/Manuscript.Rmd
Add automatic version number

@@
 ;) -->
-    | Version 0.1.0 (Feb 8, 2020)
+    | `r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`
 
@@
 The Journal of Open Source Software (JOSS) is a way of turning software directly into a publication, without the need of writing an actual paper about it (the review is about the code).
-https://citation-file-format.github.io/ is a file format to include software citation information into source code repositories.-->
+https://citation-file-format.github.io/ is a file format to include software citation information into source code repositories.
+CJ: I don't think this is so relevant, because WORCS is geared towards doing research,
+not writing software! If people use JOSS for R-packages, like I do for tidyLPA,
+then that citation will show up when users run citation("tidyLPA") --> 
 When using R, software (R packages) can be conveniently cited by calling

### 2020-05-20 R/misc.R

@@ -50,3 +50,8 @@ read_as_utf <- function(..., encoding = "UTF-8"){
   if(is.null(Args[["encoding"]])) Args[["encoding"]] <- encoding
   do.call(readLines, Args)
 }
+
+cran_version <- function(x = packageVersion("worcs")){
+  gsub("(\\d+\\.\\d+\\.\\d+).*", "\\1", as.character(x), perl = TRUE)
+}
+

### 2020-05-20 paper/Manuscript.Rmd
Update manuscript and workflow vignette

@@
 back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the
-traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference.
+traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference.
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
@@
 Thus, it is possible to preregister a study simply by pushing a text document with the study plans to GitHub.
-If the GitHub repository is Public, the preregistration is immediately visible to the world, and Reviewers can submit comments through pull requests.
-If the repository is Private, Reviewers can be invited as "Collaborators", although this removes their anonymity.
-Private repositories can be made public at a later date, along with their entire time-stamped history.
-<!-- Barbara: Pull requests are not anonymous either, aren't they?-->
+By tagging the commit as a preregistration (see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)), it is distinct from all other commits, and easily findable by people and programs.
 
-The advantages, disadvantages, and pitfalls for preregistering different types of studies have been extensively debated, and summarizing that debate falls outside of the scope of this paper, and has been addressed adequately elsewhere [e.g., @lindsayResearchPreregistration1012016].
-We take the pragmatic position that, in confirmatory (hypothesis-testing) research, it is beneficial to plan research projects before executing them, to preregister these plans, and adhere to them.
+The advantages, disadvantages, and pitfalls for preregistering different types of studies have been extensively debated, and that has been summarized elsewhere [e.g., @lindsayResearchPreregistration1012016].
+We take the pragmatic position that, in confirmatory (hypothesis-testing) research, it is beneficial to plan projects before executing them, to preregister these plans, and adhere to them.
 All deviations from this procedure should be disclosed; e.g., prior exposure to the data, whether direct (e.g., by computing summary statistics) or indirect (e.g., by reading papers using the same data); deviations from the analysis plan to handle unforseen contingencies, such as violations of model assumptions; or additional exploratory analyses.
@@
 
-1. Document study plans in a `preregistration.Rmd` file, and planned analyses in a `.R` file.
+1. Document study plans in a `preregistration.Rmd` file, and optionally, planned analyses in a `.R` file.
 2. Commit these documents to the local Git repository, and push them to GitHub
@@
   + The `.Rmd` file can be Knitted to a PDF form before uploading to a repository
+
+If the GitHub repository is Public, the preregistration is immediately visible to the world, and reviewers (both formal reviewers designated by a journal, and informal reviewers recruited by other means) can submit comments through pull requests.
+If the repository is Private, Reviewers can be invited as "Collaborators".
+Private repositories can be made public at a later date, along with their entire time-stamped history.
+It is important to note that contributing to the repository in any way will void reviewers' anonymity, as their contributions will be linked to a user name.
 
@@
 
-Before you can use the WORCS workflow, you have to install the required software, and register for a GitHub account.The setup process is documented in a vignette included with the `worcs` R-package, and also [available online](https://cjvanlissa.github.io/worcs/articles/setup.html).
-After preparing your system, you can follow the step-by-step procedure outlined below.
-<!--Barbara: Does this work on gitlab/bitbucket as well?-->
+Before you can use the WORCS workflow, you have to install the required software.
+You have to execute this setup procedure only once for each computer you intend to use `worcs` on.
+The procedure should take about 30 minutes, and it is documented in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
+After preparing your system, you can use the WORCS workflow for all of your projects.
+This step-by-step procedure is detailed in [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
+In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r cran_version()`) below.
+We highly recommend users to refer to [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html) for the most up-to-date procedure (subject to improvements).
 
-## Phase 1: Study design
-
-1. Create a new Private repository on GitHub.
-<!-- Why private!?!
-I would actually argue for a public repo, but if you do not feel comfortable doing so: another option would be to give an overview of advantages/disadvantages of private/public here, and make an argument for either case.-->
-Do not change any of the default options. Copy the repository link to the clipboard. This link will look like https://github.com/yourname/yourrepo.git
-2. In RStudio, click File > New Project > New directory > WORCS Project Template
-    a. Paste the GitHub Repository address in the textbox
-    b. Keep the checkbox for `renv` checked if you want to document all R-package dependencies, and have them automatically loaded on any system (recommended).
-    c. Select a preregistration template
-    d. Click Open in new session, and Create project
-3. Write the preregistration `preregistration.Rmd`
-4.
-In the top-right corner of RStudio, select the Git tab, select the checkboxes next to all files, and click the Commit button.
-Write an informative message for the commit, e.g., "Preregistration", again click Commit, and then click the green Push arrow to send your commit to GitHub
-5. Go to the GitHub repository for this project, and tag the Commit as a preregistration
-6. Optional: Render the preregistration to PDF, and upload it to AsPredicted.org or OSF.io as an attachment
-7. Optional: Add study Materials (only those to which you own the rights!) to the repository.
-It is possible to solicit feedback (by opening a GitHub Issue) and receive outside contributions (by accepting Pull requests)
-<!-- Barbara: NB this can only be done in a public repo; another argument for making a repo public! -->
-
-## Phase 2: Data analysis
-
-8. Read the data into R, and document this procedure in `prepare_data.R`
-9. Use `open_data()` or `closed_data()` to store the data
-10. Write the manuscript in `Manuscript.Rmd`, using code chunks to perform the analyses.
-11. Regularly commit your progress to the Git repository; ideally, after completing each small and clearly defined task. Use informative commit messages. Push the commits to GitHub.
-12. Cite essential references with one at-symbol (`[@essentialref2020]`), and non-essential references with a double at-symbol (`[@nonessential2020]`).
-<!-- Barbara: Again I cannot see the difference... -->
-
-## Phase 3: Paper submission
-
-13. Save the state of the project library (all packages used), call `renv::snapshot()`. This updates the lockfile, `renv.lock`.
-14.
-Add an open science statement to the manuscript, which references the online repository for the project.
-This reference can be masked for blind review.
-The open science statement should indicate whether additional references are available in the online repository; whether data, code, materials, and study design details are shared, and whether a pre-registration is available.
-For further guidance, see @aalbersbergMakingScienceTransparent2018.
-Example:
-    > In the spirit of open science, an online repository is available at XXX, which contains more comprehensive citations, [the data/a synthetic data file], analysis code, the research materials used, details about the study design, and a tagged commit with a preregistration.
-
-15. To render the paper with essential citations only for submission, change the line `knit: worcs::cite_all` to `knit: worcs::cite_essential`. Then, press the Knit button to generate a PDF
-
-## Phase 4: Publication
-
-Note that some researchers take the following steps at an earlier stage.
-However, for newcomers to open science, who are unsure about potential implications for the publication of their work, it might be prudent to perform these actions after acceptance, and before publication:
-
-16. Make the GitHub repository public
-17. [Create an OSF project](https://help.osf.io/hc/en-us/articles/360019737594-Create-a-Project); although you may have already done this in Step 6.
-18. [Connect your GitHub repository to the OSF project](https://help.osf.io/hc/en-us/articles/360019929813-Connect-GitHub-to-a-Project)
-19. Add an open science statement to the manuscript, with a link to the OSF project
-20.
-Optional: [Publish preprint in a not-for-profit preprint repository such as PsyArchiv, and connect it to your existing OSF project](https://help.osf.io/hc/en-us/articles/360019930533-Upload-a-Preprint)
-    + Check [Sherpa Romeo](http://sherpa.ac.uk/romeo/index.php) to be sure that your intended outlet allows the publication of preprints; many journals do, nowadays - and if they do not, it is worth considering other outlets.
-
-<!-- 
-4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
--->
+```{r results="asis", echo = FALSE, eval = TRUE}
+get_workflow <- function(filename = "../vignettes/workflow.Rmd"){
+  txt <- readLines(filename)
+  txt <- txt[grepl("<!--S:", txt, fixed = T)|grepl("^###", txt)]
+  txt <- gsub("(<!--S: |-->.*$)", "", txt)
+  txt <- gsub("^#", "", txt)
+  out <- vector("character")
+  for(i in seq_along(txt)){
+    if(startsWith(txt[i], "#")){
+      out <- append(out, c("", txt[i], ""))
+    } else {
+      out <- append(out, txt[i])
+    }
+  }
+  cat(out, sep = "\n")
+}
+get_workflow()
+```
 
@@
 
-Another important challenge is managing collaborations when only the lead author uses Rmarkdown, and the coauthors use Word.
-In this case, it is possible to Knit the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf` in the manuscript's YAML header to `output: papaja::apa6_docx`.
-There are some limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
-When soliting feedback from co-authors, ask them to use Tracked Changes and comment bubbles in Word.
-<!-- LB: The work-around with track-changes in Word files is doable, but not very satisfying.
-I think you can also use Overleaf instead of Word (user friendly and web based) for bidirectional communication, but haven’t tried this myself.
-I didn’t know about ‘redoc’, will check it out.-->Then, manually<!--this should really be avoided if possible!
-no fun...
---> incorporate their changes into the `manuscript.Rmd` file.
-In most cases, this is the most user-friendly approach, as most lead authors would review changes by co-authors anyway.
-A second approach is to ask collaborators to work in plain text.
-In this case, send collaborators the `manuscript.Rmd` file, and ask them to open (and save) it in Word or Notepad as a plain text file.
-When they send it back, make sure any changes to your local file are committed, and then simply overwrite your local version with their file.
-In RStudio, select the file in the Git tab, and click the Diff button to examine what changes the collaborator has made relative to your last committed version.
-Finally, @peikertReproducibleDataAnalysis2019 mention that the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] might be useful for collaborating with Word users.
-This package enables bidirectional conversion between Word and Rmarkdown, with support for tracked changes.
-However, a simpler solution (i.e., asking collaborators to work in plain text) might be sufficient.
+Another important challenge is managing collaborations when only the lead author
+uses Rmarkdown, and the coauthors use Word. In this case, it is possible to Knit
+the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf`
+in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some
+limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
+When soliting feedback from co-authors, ask them to use Tracked Changes and
+comment bubbles in Word. <!-- LB: Overleaf?-->Then, manually incorporate their
+changes into the `manuscript.Rmd` file. In most cases, this is the most
+user-friendly approach, as most lead authors would review changes by co-authors
+anyway. A second approach is to ask collaborators to work in plain text. In this
+case, send collaborators the `manuscript.Rmd` file, and ask them to open (and
+save) it in Word or Notepad as a plain text file. When they send it back, make
+sure any changes to your local file are committed, and then simply overwrite
+your local version with their file. In RStudio, select the file in the Git tab,
+and click the Diff button to examine what changes the collaborator has made
+relative to your last committed version. Finally,
+@peikertReproducibleDataAnalysis2019 mention the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] as a
+potentially useful way to collaborate with Word users. This package enables
+bidirectional conversion between Word and
+Rmarkdown, with support for tracked changes. However, `redoc` is currently in
+suspended development, and for many users, a simple solution (i.e.,
+asking collaborators to work in plain text) might be sufficient.
 
-If all collaborators are committed to using WORCS, they can [Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo) from the lead author on GitHub, [clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository), make their own changes, and [send a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) to incorporate their changes.
-Working this way is extremely conducive to scientific collaboration [@ramGitCanFacilitate2013].
-Moreover, being familiar with GitHub opens doors to new forms of collaboration: In the open source software community, continuous peer review and voluntary collaborative acts by strangers who are interested in a project are commonplace [see @adolphOpenBehavioralScience2012].
-This kind of collaboration is budding in scientific software development as well; for example, I <!-- AL: which of the meanwhile several authors?
-:) -->  became a co-author on several R-packages after submitting pull requests with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018].
-It is also possible to invite such collaboration by [opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues) for tasks that still need to be accomplished, and tag known Collaborators to address them, or invite external collaborators to contribute their expertise.
+If all collaborators are committed to using WORCS, they can 
+[Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo)
+from the lead author on GitHub,
+[clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository),
+make their own changes, and
+[send a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request)
+to incorporate their changes. Working this way is extremely conducive to
+scientific collaboration [@ramGitCanFacilitate2013]. Recall that, when using 
+'Git' for collaborative writing, it is recommended to insert a
+line break after every sentence so that the change log will indicate which
+specific sentence was edited, and to prevent
+["merge conflicts"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/resolving-a-merge-conflict-on-github)
+when two authors edit the same line. The resulting document will be rendered to
+PDF without spurious line breaks.
+
+Moreover, being familiar
+with GitHub opens doors to new forms of collaboration: In the open source
+software community, continuous peer review and voluntary collaborative acts by
+strangers who are interested in a project are commonplace
+[see @adolphOpenBehavioralScience2012].  This kind of collaboration is budding
+in scientific software development as well; for example, the lead author of this
+paper became a co-author on several R-packages after submitting pull requests
+with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018], and two co-authors of this
+paper became involved by contributing pull requests to `worcs`. It is also
+possible to invite such collaboration by 
+[opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues)
+for tasks that still need to be accomplished, and tag known Collaborators to
+address them, or invite external collaborators to contribute their expertise.
 

### 2020-05-20 paper/Manuscript.Rmd
Fix references

@@
 Code in Science (WORCS). WORCS is a lightweight approach to open science and
-computational reproducibility. The workflow (Figure \@ref(fig:workflow)) 
+computational reproducibility. The workflow (Figure <!--\@ref(fig:workflow)-->) 
 constitutes a step-by-step procedure that researchers can follow to make a 
@@
 for neuroimaging data). However, it can be difficult to set up for novice users - although a package
-is in development to facilitate this transition [@repo]. The cloud-based research collaboration platform
+is in development to facilitate this transition [@repro]. The cloud-based research collaboration platform
 'Code Ocean' is also based on Docker. It is very user-friendly, but it is not clear whether the
@@
 back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the
-traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@`), used to cite a non-essential reference.
+traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference.
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
@@
 
-```{r scholarbib, echo = FALSE}
+```{r, scholarbib, echo = FALSE}
 knitr::include_graphics("scholar_bib.png")
 ```
-
-<!--
-Future: Explain why each reference is cited? How best to include this?
-* Note
-* In a function
-* In a separate document-->
 

### 2020-05-21 paper/Manuscript.Rmd
added tryCatch around worcs:::cran_version() in case it cannot be called

@@
 ;) -->
-    | `r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`
+    | `r paste0("Version: ", tryCatch({worcs:::cran_version()}, error=function(err) 'NA'), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`
 
@@
 This step-by-step procedure is detailed in [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
-In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r cran_version()`) below.
+In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r
+tryCatch({worcs:::cran_version()}, error = function(err) 'NA' )`) below.
 We highly recommend users to refer to [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html) for the most up-to-date procedure (subject to improvements).

### 2020-05-21 paper/Manuscript.Rmd
Try to improve structure

--- 2020-05-20 paper/Manuscript.Rmd
@@
   step-by-step procedure that researchers can follow to make an entire project
-  Open and Reproducible. It is based on best practices, and can be used either
+  open and reproducible. It is based on best practices, and can be used either
   in parallel to, or in absence of, top-down requirements by journals. WORCS
@@
 constitutes a step-by-step procedure that researchers can follow to make a 
-research project Open and Reproducible. WORCS is compatible with open science
+research project open and reproducible. WORCS is compatible with open science
 requirements already implemented by journals and institutions, and will help 
 fulfill them. It can also be used to produce work in accordance with best 
-practices in the absence of top-down support. The WORCS workflow is supported by
-an R-package with user-friendly support functions, and an RStudio project
-template that is, effectively, a "one click solution". This paper introduces
-the WORCS workflow, discusses how it meets best practices for open science, and
-illustrates the use of the `worcs` package.
+practices in the absence of top-down support. Although the principles underlying
+the WORCS procedure are universal, they have been implemented for R users in the
+R-package `worcs`, which offers an RStudio project template that is, effectively,
+a "one click solution". This paper introduces the WORCS workflow, discusses how
+it meets best practices for open science, and illustrates the use of the `worcs`
+package.
 
@@
 FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a]. These principles advocate that research
-data should be Findable, Accessible, Interoperable and Reusable. However, they are
-increasingly applied as a standard for 
-to other types of research output as well
+data should be Findable, Accessible, Interoperable and Reusable. They are
+increasingly applied as a standard for other types of research output as well
 [@lamprechtFAIRPrinciplesResearch2019].
 WORCS helps users meet the FAIR principles by making all research objects -
-including (origina) data and code - accessible in a (public) repository. 
+including data and code - accessible in a repository that can be made public. 
 The `worcs` package additionally
@@
 indexing data by topic area. 
-Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/].
-It is hosted on a public version controlled repository ('GitHub'), it has an open source licence (GPL v3.0), is registered in a community registry (CRAN), it enables the citation of the software (using the `citation("worcs")` command, or by citing this paper), and followed the CII Best Practices software quality checklist during development.
+Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is 
+hosted on the public version controlled repository 'GitHub', it has an open
+source licence (GPL v3.0), is registered in a community registry (CRAN), it
+enables the citation of the software (using the `citation("worcs")` command, or
+by citing this paper), and followed the CII Best Practices software quality
+checklist during development.
 
-Sharing all research output also provides a proper basis for research evaluation
+Sharing all research objects, as advocated in WORCS, also provides a proper
+basis for research evaluation
 according to the San Francisco Declaration on Research Assessment (DORA;
@@
 
-## Existing solutions
-
-There have been several previous efforts to promote grass-roots adoption of open
-science principles. Each of these efforts has a different scope, strengths, and
-limitations that set it apart from WORCS. For example, there are "signalling
-solutions"; guidelines to structure and incentivize disclosure about open
-science practices. Specifically, @aalbersbergMakingScienceTransparent2018
-suggested publishing a "TOP-statement" as supplemental material, which discloses
-the authors' adherence to open science principles. Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based
-Transparency Checklist that authors can complete online to generate a report.
-Such signalling solutions are very easy to adopt, and they address
-TOP-guidelines 1-7. Many journals now also offer authors the opportunity to earn
-"badges" for adhering to open science guidelines
-[@kidwellBadgesAcknowledgeOpen2016]. These signalling solutions help structure
-authors' disclosures about, and incentivize adherence to, open science practices.
-
-A different class of solutions instead focuses on the practical issue of *how*
-researchers can meet the requirements of open science.
-<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example is the workflow for
-reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
-excellent workflow strives to ensure strict computational reproducibility for
-even the most sophisticated analyses. It has some limitations, however: In
-focusing on reproducibility, however, it only addresses TOP-guidelines 2, 3, and
-5. Moreover, in aiming for strict computational reproducibility it ends up being
-relatively complex to implement. WORCS builds upon the same general principles
-as Peikert and Brandmaier, which means the two workflows are compatible. What 
-sets WORCS apart is that is an easy and sufficient solution for projects that
-can be conducted entirely within R, resulting in a much more lightweight
-approach to safeguard computational reproducibility under *most circumstances*.
-WORCS also addresses a unique issue not covered by other existing solutions,
-namely to provide a workflow most conducive to satisfying the TOP-guidelines and
-FAIR principles, while being compatible with existing requirements.
-
-```{r, echo = FALSE, eval = FALSE}
-tab1 <- data.frame(
-  Reference = c("Aalbers et al., 2018", "Peikert & Brandmaier, 2019"),
-  Scope = c("Disclose adherence to open science principles", "Completely reproducible analysis"),
-  Difficulty = c("Easy", "Hard"),
-  Effectiveness = c("Low", "High"),
-  "TOP-guidelines" = c("1-7", "2,3,5")
-)
-apa_table(tab1)
-```
-
-<!-- AL: Similar to my comment on the title, in this section it needs to make clear which scope of related work it covers.
-For example, in the life sciences there has also been a lot of work on practices for increasing reproducibility, with automated workflows, containerization, Research Objects, provenance protocols, and a lot more.
-I don't think it's necessary for this paper to discuss this, but that is under the provision that it defines its scope better.
--->
 ## Prerequisites
@@
 `worcs` uses Markdown and `.csv` extensively.
+
+# Introducing the workflow 
+<!-- AL: Agreeing with LB's ealier comment, it would be good to discuss this overall workflow before going into the details. -->
+
+## Preparing your system
+
+Before you can use the WORCS workflow, you have to install the required software.
+You have to execute this setup procedure only once for each computer you intend to use `worcs` on.
+The procedure should take about 30 minutes, and it is documented in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
+After preparing your system, you can use the WORCS workflow for all of your projects.
+This step-by-step procedure is detailed in [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
+In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r cran_version()`) below.
+We highly recommend users to refer to [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html) for the most up-to-date procedure (subject to improvements).
+
+```{r results="asis", echo = FALSE, eval = TRUE}
+get_workflow <- function(filename = "../vignettes/workflow.Rmd"){
+  txt <- readLines(filename)
+  txt <- txt[grepl("<!--S:", txt, fixed = T)|grepl("^###", txt)]
+  txt <- gsub("(<!--S: |-->.*$)", "", txt)
+  txt <- gsub("^#", "", txt)
+  out <- vector("character")
+  for(i in seq_along(txt)){
+    if(startsWith(txt[i], "#")){
+      out <- append(out, c("", txt[i], ""))
+    } else {
+      out <- append(out, txt[i])
+    }
+  }
+  cat(out, sep = "\n")
+}
+get_workflow()
+```
 
@@
 
-# The workflow 
-<!-- AL: Agreeing with LB's ealier comment, it would be good to discuss this overall workflow before going into the details. -->
-
-## Preparing your system
-
-Before you can use the WORCS workflow, you have to install the required software.
-You have to execute this setup procedure only once for each computer you intend to use `worcs` on.
-The procedure should take about 30 minutes, and it is documented in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
-After preparing your system, you can use the WORCS workflow for all of your projects.
-This step-by-step procedure is detailed in [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
-In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r cran_version()`) below.
-We highly recommend users to refer to [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html) for the most up-to-date procedure (subject to improvements).
-
-```{r results="asis", echo = FALSE, eval = TRUE}
-get_workflow <- function(filename = "../vignettes/workflow.Rmd"){
-  txt <- readLines(filename)
-  txt <- txt[grepl("<!--S:", txt, fixed = T)|grepl("^###", txt)]
-  txt <- gsub("(<!--S: |-->.*$)", "", txt)
-  txt <- gsub("^#", "", txt)
-  out <- vector("character")
-  for(i in seq_along(txt)){
-    if(startsWith(txt[i], "#")){
-      out <- append(out, c("", txt[i], ""))
-    } else {
-      out <- append(out, txt[i])
-    }
-  }
-  cat(out, sep = "\n")
-}
-get_workflow()
-```
-
 # Discussion
@@
 This relatively light-weight workflow meets most of the requirements for open science, as detailed in the TOP-guidelines.
-However, several limitations and issues for future development remain.
 
-One potential challenge is the learning curve associated with the tools outlined in this paper.
-Learning to work with R, Rmarkdown, and Git/GitHub requires an initial time investment.
-The amount of time required is reduced by the availability of tutorials, such as this one, and other tutorials cited throughout this document.
-Moreover, the time investment tends to pay off.
-Working with R opens the door to many cutting edge analysis techniques.
-Working with Rmarkdown saves time and prevents mistakes by avoiding tedious copying of results into a text document.
-Working with Git/GitHub keeps projects organized, prevents accidental loss of work, enables integrating changes by collaborators in a non-destructive manner, and ensures that entire research projects are archived and can be accessed or copied by third parties.
+## Differences from existing solutions
+
+There have been several previous efforts to promote grass-roots adoption of open
+science principles. Each of these efforts has a different scope, strengths, and
+limitations that set it apart from WORCS. For example, there are "signalling
+solutions"; guidelines to structure and incentivize disclosure about open
+science practices. Specifically, @aalbersbergMakingScienceTransparent2018
+suggested publishing a "TOP-statement" as supplemental material, which discloses
+the authors' adherence to open science principles. Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based
+Transparency Checklist that authors can complete online to generate a report.
+Such signalling solutions are very easy to adopt, and they address
+TOP-guidelines 1-7. Many journals now also offer authors the opportunity to earn
+"badges" for adhering to open science guidelines
+[@kidwellBadgesAcknowledgeOpen2016]. These signalling solutions help structure
+authors' disclosures about, and incentivize adherence to, open science practices.
+
+A different class of solutions instead focuses on the practical issue of *how*
+researchers can meet the requirements of open science.
+<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example is the workflow for
+reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
+excellent workflow strives to ensure strict computational reproducibility for
+even the most sophisticated analyses. It has some limitations, however: In
+focusing on reproducibility, however, it only addresses TOP-guidelines 2, 3, and
+5. Moreover, in aiming for strict computational reproducibility it ends up being
+relatively complex to implement. WORCS builds upon the same general principles
+as Peikert and Brandmaier, which means the two workflows are compatible. What 
+sets WORCS apart is that is an easy and sufficient solution for projects that
+can be conducted entirely within R, resulting in a much more lightweight
+approach to safeguard computational reproducibility under *most circumstances*.
+WORCS also addresses a unique issue not covered by other existing solutions,
+namely to provide a workflow most conducive to satisfying the TOP-guidelines and
+FAIR principles, while being compatible with existing requirements.
+
+```{r, echo = FALSE, eval = FALSE}
+tab1 <- data.frame(
+  Reference = c("Aalbers et al., 2018", "Peikert & Brandmaier, 2019"),
+  Scope = c("Disclose adherence to open science principles", "Completely reproducible analysis"),
+  Difficulty = c("Easy", "Hard"),
+  Effectiveness = c("Low", "High"),
+  "TOP-guidelines" = c("1-7", "2,3,5")
+)
+apa_table(tab1)
+```
+
+<!-- AL: Similar to my comment on the title, in this section it needs to make clear which scope of related work it covers.
+For example, in the life sciences there has also been a lot of work on practices for increasing reproducibility, with automated workflows, containerization, Research Objects, provenance protocols, and a lot more.
+I don't think it's necessary for this paper to discuss this, but that is under the provision that it defines its scope better.
+-->
+
+## Limitations
+
+WORCS is intended to substantially reduce the threshold for adopting best 
+practices in open and reproducible research. However, several limitations and
+issues for future development remain. One potential challenge is the learning
+curve associated with the tools outlined in this paper. Learning to work with R,
+Rmarkdown, and Git/GitHub requires an initial time investment. The amount of
+time required is reduced by the availability of tutorials, such as this one, and
+other tutorials cited throughout this document. Moreover, the time investment
+tends to pay off. Working with R opens the door to many cutting edge analysis
+techniques. Working with Rmarkdown saves time and prevents mistakes by avoiding
+tedious copying of results into a text document. Working with Git/GitHub keeps
+projects organized, prevents accidental loss of work, enables integrating
+changes by collaborators in a non-destructive manner, and ensures that entire
+research projects are archived and can be accessed or copied by third parties.
 Thus, the time investment is eminently worthwhile.

### 2020-05-21 paper/Manuscript.Rmd

@@
 
-# Meeting the requirements of Open Science
+# What are Open Science practices?
 
@@
 
-## Prerequisites
+## Implementation in R
 
-Although the principles underlying this workflow are universal, WORCS has been
-designed for R [@rcoreteamLanguageEnvironmentStatistical2020].
+Although the principles underlying this workflow are universal, WORCS has 
+currently been implemented for R only [@rcoreteamLanguageEnvironmentStatistical2020].
 Several arguments support the choice to work in R.
@@
 # Introducing the workflow
-<!-- AL: Agreeing with LB's ealier comment, it would be good to discuss this overall workflow before going into the details. -->
 

### 2020-05-21 paper/Manuscript.Rmd
Create escape option for literal double at

@@
 back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the
-traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`@@`), used to cite a non-essential reference.
+traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`\@\@`), used to cite a non-essential reference.
 Users can render a manuscript to PDF either with, or without, comprehensive citations.
@@
 For more options, see the [Rmarkdown cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html).
-3. To indicate a *non-essential* citation, mark it with a double at-symbol: `@@nonessential2020`.
+3. To indicate a *non-essential* citation, mark it with a double at-symbol: `\@\@nonessential2020`.
 4. When Knitting the document, adapt the `knit` command in the YAML header.  

### 2020-05-23 paper/Manuscript.Rmd
Fixed figure problem
Make synthetic a bit more flexible wrt classes

@@
 
-```{r, scholarbib, echo = FALSE}
+```{r, scholarbib, echo = FALSE, fig.cap="Exporting a BibTex reference from Google Scholar"}
 knitr::include_graphics("scholar_bib.png")
@@
 
-The second solution implemented in `worcs` is to share a synthetic dataset with similar characteristics to the real data.
-Synthetic data mimic the level of measurement and (conditional) distributions of the real data [see @nowokSynthpopBespokeCreation2016].
-Sharing synthetic data allows any third party to 1) verify that the published code works, 2) debug the code, and 3) write valid code for alternative analyses.
-It is important to note that complex multivariate relationships present in the real data are often lost in synthetic data.
-Thus, findings from the real data might not be replicated in the synthetic data, and findings in the synthetic data *should not be substantively interpreted*.
-Still, sharing synthetic data facilitates data requests from third parties.
-A third party can write analysis code based on the synthetic data, and send it to the authors who evaluate it on the real data and send back the results.
+The second solution is to share a synthetic dataset with similar characteristics
+to the real data. Synthetic data mimic the level of measurement and
+(conditional) distributions of the real data
+[see @nowokSynthpopBespokeCreation2016]. Sharing synthetic data allows any third
+party to 1) verify that the published code works, 2) debug the code, and 3)
+write valid code for alternative analyses. It is important to note that complex
+multivariate relationships present in the real data are often lost in synthetic
+data. Thus, findings from the real data might not be replicated in the synthetic
+data, and findings in the synthetic data *should not be substantively interpreted*.
+Still, sharing synthetic data facilitates data requests from third parties. A
+third party can write analysis code based on the synthetic data, and send it to
+the authors who evaluate it on the real data and send back the results. The
+`worcs` package offers a simple but flexible function for synthetic data,
+`synthetic()`, which by default uses random forests to generate a model for
+synthesis (see @wrightRangerFastImplementation2015). For data synthesis to work
+properly, it is essential that the `class` (data type) of variables is defined
+correctly. As of Version `r cran_version()`, numeric, integer, and factor types
+are supported out of the box. Other types of variables should be converted to
+one of these types, or users can use a custom `model_expression` when calling
+`synthetic`.
 
@@
 Once the GitHub repository is made public, people will have access to a checksum for the original data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
+
 <!-- Maybe remark that in the "closed" case, a the data should be backupped somewhere else, as GitHub will not do the job in this setting. -->

### 2020-05-23 R/synthetic.R

@@ -42,6 +42,12 @@
 #' }
 #' Variables are thus imputed in order of occurrence in the \code{data.frame}.
 #' To impute in a different order, reorder the data.
+#'
+#' Note that, for data synthesis to work properly, it is essential that the
+#' \code{class} of variables is defined correctly. The default algorithm
+#' \code{\link[ranger]{ranger}} supports numeric, integer, and factor types.
+#' Other types of variables should be converted to one of these types, or users
+#' can use a custom \code{model_expression} when calling \code{synthetic}.
 #' @examples
 #' iris_syn <- synthetic(iris)
 #' iris_missings <- iris
@@ -134,9 +140,24 @@ synthetic.data.frame <- function(data,
   if(verbose) close(pb)
 
   # Check if variable classes are maintained
-  matching_types <- mapply(function(x, y){all(x == y)}, x = coltypes, y = lapply(xsynth, class))
-  if(any(!matching_types)){
-    warning("Column type of the synthetic data did not match column type of the original data for column: ", paste0(names(xsynth)[!matching_types], collapse = ", "))
+  for(thisvar in 1:ncol(data)){
+    #thisvar=6
+    if(!all(class(xsynth[[thisvar]]) == coltypes[[thisvar]])){
+      msg <- paste0("Synthetic variable '", names(data)[thisvar], "' did not have identical classes to its original counterpart.")
+      # Try to convert to correct class
+      convert_func <- paste0("as.", coltypes[[thisvar]])
+      convert_func <- convert_func[sapply(convert_func, exists)]
+      if(length(convert_func) > 0) convert_func <- convert_func[1]
+      newvar <- tryCatch({
+        do.call(convert_func, list(xsynth[[thisvar]]))
+      }, error = function(e){NULL})
+      if(!is.null(newvar)){
+        col_message(msg, " Attempted to convert to its original type. Check the input types of your variables, and check whether the data are synthesized correctly.", success = TRUE)
+        xsynth[[thisvar]] <- newvar
+      } else {
+        col_message(msg, " Failed to convert to its original type. Check the input types of your variables, and check whether the data are synthesized correctly.", success = FALSE)
+      }
+    }
   }
 
   # Restore missing values

### 2020-05-25 R/citations.R
Updates to functions

@@ -73,7 +73,7 @@ comprehensive_cite <- function(input, encoding = "UTF-8", ..., citeall = TRUE) {
 }
 
 .nonessential_to_normal <- function(text){
-  text <- gsub("@@", "@", text, fixed = TRUE)
+  text <- gsub("(?<!`)@@", "@", text, perl = TRUE)
   text <- gsub("\\@\\@", "@@", text, fixed = TRUE)
   text
 }

### 2020-05-25 R/misc.R

@@ -52,6 +52,9 @@ read_as_utf <- function(..., encoding = "UTF-8"){
 }
 
 cran_version <- function(x = packageVersion("worcs")){
-  gsub("(\\d+\\.\\d+\\.\\d+).*", "\\1", as.character(x), perl = TRUE)
+  tryCatch(
+    gsub("(\\d+\\.\\d+\\.\\d+).*", "\\1", as.character(x), perl = TRUE),
+    error = function(e){NA}
+  )
 }
 

### 2020-05-25 paper/Manuscript.Rmd
Update paper

@@
 title: |
-    | WORCS: A Workflow for Open Reproducible Code in the Social Sciences and Beyond <!-- AL: I strongly recommend to make the title more specific.
-"Science" is very general, and in the Anglo-American understanding actually refers to the beta sciences.
-Also, the abstract and the paper start with talking about the situation in the social sciences and derive the motivaion from there, so it would be good if the title would reflect this.
-The acronym can stay as it is.
-;) -->
-    | `r paste0("Version: ", tryCatch({worcs:::cran_version()}, error=function(err) 'NA'), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`
+    | WORCS: A Workflow for Open Reproducible Code in Science
+    | `r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`
 
@@
 abstract: |
-  The social sciences are amidst a paradigm shift towards open science. In part,
-  this transition has been fueled by cases of scientific fraud and increasing
-  awareness of questionable research practices and failed replications. However,
-  open science is not
-  merely a cure (or punishment) for this crisis - it is also an opportunity.
   Technological advances enable researchers to more easily conduct reliable,
   cumulative, collaborative science. Capitalizing on these advances has the
-  potential to accelerate scientific progress. This tutorial paper introduces a
-  workflow for open reproducible code in science (WORCS). WORCS is a 
-  step-by-step procedure that researchers can follow to make an entire project
-  open and reproducible. It is based on best practices, and can be used either
-  in parallel to, or in absence of, top-down requirements by journals. WORCS
-  lowers the threshold for adopting an open science workflow by providing this
-  tutorial, along with an R-package with user-friendly support functions, and an
-  RStudio project template.
+  potential to accelerate scientific progress. However, adopting open science
+  principles can be challenging and costly, because doing so requires
+  substantial conceptual education and training in the use of new tools. The 
+  goal of this paper is to remove these obstacles. We introduce the Workflow for
+  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
+  procedure that researchers can follow to make an entire project open and
+  reproducible. It is based on best practices, and can be used either in
+  parallel to, or in absence of, top-down requirements by journals,
+  institutions, and funding bodies. To further facilitate the transition, the
+  WORCS principles are implemented in an R-package, which offers an RStudio 
+  project template that is, effectively, a "one-click solution" for R users. 
+  This paper introduces the workflow, discusses how it meets different 
+  requirements of open science, and provides examples of the implementation of
+  `worcs` in R.
 
@@
 
-\vspace{\baselineskip}
+<!--\vspace{\baselineskip}-->
 
-<!-- Structure
-# Introduction
-# Defining open science
-* TOP guidelines
-# Goal of this paper
-* Grass-roots
-* open science as opportunity ?move
-    - Easier work
-    - Less mistakes
-    - Better collaboration
-    - Cumulative science
-    - Error correction
-# Introducing the tools
-# Setup - do this only once
-# WORCS - steps to follow for each project
-# Discussion
-* advantages
-* limitations
-* future directions
--->
-
-# Introduction
-
-The (social) sciences are arguably past the tipping point of the paradigm shift
+The (social) sciences are arguably past the tipping point of a paradigm shift
 towards open science.
-Several immediate causes stirred up support for this transition; including
+Several immediate causes rallied support for this transition, including
 highly publicized cases of scientific fraud [as discussed by
@@
 [@shroutPsychologyScienceKnowledge2018].
-However, open science is not merely a cure for this crisis - it is also an
-opportunity [@adolphOpenBehavioralScience2012].
-Open science helps researchers to more easily conduct reliable, cumulative, and 
-collaborative science [@nosekScientificUtopiaOpening2012].
+However, open science should not be seen as a cure (or punishment) for this
+crisis. As the late dr. Jonathan Tennant put it: "Open science is just good science"
+[@tennantOpenScienceJust2018].
+Moreover, open science creates opportunities for researchers to more easily
+conduct reliable, cumulative, and 
+collaborative science [@nosekScientificUtopiaOpening2012;
+@adolphOpenBehavioralScience2012]. Open science also
+promotes inclusivity, because it removes barriers for participation.
 Capitalizing on these advances has the potential to accellerate scientific
@@
 science and enjoy these benefits. And yet, the question of "how" to 
-adopt open science practices can be daunting. Making the transition effectively
+adopt open science practices can be daunting. Making the transition
 requires researchers to become knowledgeable about different open science 
@@
 This paper is designed to ease that transition by providing a simple workflow 
-that meets most requirements open science, based on best practices, and 
-supported by user-friendly software in R: The Worfkflow for Open Reproducible 
+that meets most requirements open science, is based on best practices, and 
+is supported by user-friendly software in R: The Worfkflow for Open Reproducible 
 Code in Science (WORCS). WORCS is a lightweight approach to open science and
@@
 requirements already implemented by journals and institutions, and will help 
-fulfill them. It can also be used to produce work in accordance with best 
-practices in the absence of top-down support. Although the principles underlying
+fulfill them. It can also be used in the absence of top-down support or guidelines
+to produce work in accordance with best practices.
+Although the principles underlying
 the WORCS procedure are universal, they have been implemented for R users in the
@@
 
-# What are Open Science practices?
+# Defining Open Science practices
 
-Although Open Science is advocated by many, it does not have a unitary definition.
+Although Open Science is advocated by many, it does not have a unitary definition
+[but, for an effort to ground open science in first principles, see 
+@tennantValuePropositionOpen2020].
 Instead, several guidelines exist for practices that are considered "open".
@@
 the subject of a research project, not to its execution.
-We group the remaining seven guidelines into three categories: Citation (1),
+We group the remaining seven guidelines into three categories: citation (1),
 sharing (2-5), and preregistration (6-7).
-
-WORCS meets the TOP-guidelines in a way that is amenable to the 
-FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a]. These principles advocate that research
-data should be Findable, Accessible, Interoperable and Reusable. They are
-increasingly applied as a standard for other types of research output as well
-[@lamprechtFAIRPrinciplesResearch2019].
-WORCS helps users meet the FAIR principles by making all research objects -
-including data and code - accessible in a repository that can be made public. 
-The `worcs` package additionally
-facilitates meeting the FAIR principles in several ways. First, each `worcs`
-project has a YAML file (https://yaml.org/) that lists the research objects in
-the project, which will allow web crawlers 
-to index `worcs` projects. Second, a codebook is automatically generated for
-each data file stored
-using `open_data()` or `closed_data()`. We encourage users to elaborate on the
-default codebook by adding
-variable description and categories, which might - in the future - enable
-indexing data by topic area. 
-Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is 
-hosted on the public version controlled repository 'GitHub', it has an open
-source licence (GPL v3.0), is registered in a community registry (CRAN), it
-enables the citation of the software (using the `citation("worcs")` command, or
-by citing this paper), and followed the CII Best Practices software quality
-checklist during development.
-
-Sharing all research objects, as advocated in WORCS, also provides a proper
-basis for research evaluation
-according to the San Francisco Declaration on Research Assessment (DORA;
-https://sfdora.org/), which plays an increasing role in grant funding, hiring, 
-and promotion procedures. Direct access to research objects allows stakeholders
-to evaluate research quality based on content rather than relying on spurious
-surrogate indicators like journal impact factors, conference rankings, and
-h-indexes. The detailed version control and commit tracking of 'GitHub' and
-similar platforms furthermore make it possible to assess the relative
-contributions made by different researchers.
 
@@
 
-Although the principles underlying this workflow are universal, WORCS has 
-currently been implemented for R only [@rcoreteamLanguageEnvironmentStatistical2020].
+Although the principles underlying this workflow are universal, WORCS has first
+been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
 Several arguments support the choice to work in R.
-First, R and all of it extensions are free and open source software, which
+First, R and all of it extensions are free open source software, which
 should be the tool of choice for open science.
@@
 Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
-Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R [including C++, Fortran, and Python, @allaireMarkdownPythonEngine].
-There are excellent free resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].
+Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
+Fortran, and Python [@allaireMarkdownPythonEngine]. There are excellent free
+resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].
 
-Working with R is simplified immensely by using the RStudio integrated development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
-Most of the tools used in WORCS are embedded directly into the user interface of RStudio.
-RStudio automates and streamlines tedious or complicated aspects of working with R, Rmarkdown, and Git/GitHub.
+Working with R is simplified immensely by using the RStudio integrated
+development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
+RStudio automates and streamlines tedious or complicated aspects of working with
+the tools used in WORCS, and many are embedded directly into the visual user
+interface.
 This makes RStudio a comprehensive solution for open science research projects.
@@
 
+### Dynamic document generation
+
 WORCS relies on several free, open source software solutions.
@@
 a format that interleaves written reports with blocks of code used to conduct
-the analyses. The text is automatically formatted; the `worcs` package can 
-output manuscripts as an APA-style paper [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], or as a host of other scientific formats 
-[see @allaireRticlesArticleFormats2020], or as a plain 'Markdown' format for
-'GitHub'. When the text is formatted, the results of the code blocks are
+the analyses. The text is automatically formatted as a scientific paper in 
+several potential styles. When the text is formatted, the results of the code
+blocks are
 automatically generated and insterted in the text, or rendered as Figures and
@@
 analyses, and then manually copy-pasting analysis results into the text.
+
 Although there is a slight learning curve to transitioning to DDG, the
-investment really pays off:
-
-* Time saved from painstakingly copy-pasting output and manually formatting text
-soon outweighs the investment of switching to a new program
-* Human error in manually copying results is eliminated
-* When revisions require major changes to the analyses, all results, Figures and
-Tables are automatically updated
-* Flexibility in output formats means that a manuscript can be rendered to
-presentation format, or even to a website
-* Reproducibility is guaranteed, and can be verified by reviewers and/or
-readers, because the code is run each time the document is compiled.
-
-In sum, while writing academic papers in a programming environment might seem
+investment really pays off. Time saved from painstakingly copy-pasting output
+and manually formatting text soon outweighs the investment of switching to a new
+program. Moreover, human error in manually copying results is eliminated.
+When revisions require major changes to the analyses, all results, Figures and
+Tables can be automatically updated. The flexibility in output formats also 
+means that a manuscript can be rendered to presentation format, or even to a
+website or blog post. Moreover, reproducibility is guaranteed, and can be
+verified by reviewers and/or readers, because the code is run each time the
+document is compiled. In sum, while writing academic papers in a programming environment might seem
 counter-intuitive at first, this approach is much more amenable to the needs of
 academics than most word processing software. It prevents mistakes, and saves
-time. We recommend centering a research project around one dynamically generated
-(RMarkdown) document, which includes all analysis code. Longer scripts can be
+time.
+
+In the R-implementation of WORCS, we recommend centering a research project
+around one dynamically generated 'RMarkdown' document, which includes all 
+analysis code. Longer scripts can be
 stored in `.R` files, and called from the main document using the `source()`
-function.
+function. When a reader or reviewer compiles this document, all code is run
+automatically, thus verifying computational replicability. The 'RMarkdown' 
+document can be automatically formatted in many styles, including
+APA style [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020],
+a host of other scientific formats 
+[see @allaireRticlesArticleFormats2020], and as plain 'Markdown' for 'GitHub'.
+
+### Version control
 
@@
 that retain backups of previous versions.
-WORCS instead uses the *explicit* version control software 'Git'
+
+WORCS recommends using the *explicit* version control software 'Git'
 (www.git-scm.com). Git tracks
@@
 privacy sensitive data.
-The `.gitignore` file lists files that should not be version-controlled.
+A `.gitignore` file lists files that should not be version-controlled.
 These files thus exist only on the user's private computer.
@@
 ability to "revert" (return) to a previous state of the project, or to "merge"
-(incorporate) the experimental branch. This paper focuses on 'GitHub', which is
-the most widely used 'Git' cloud storage services. However, as of version 0.1.2,
-`worcs` projects can interchangeably use the two main alternative services,
-'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). The
-package has a [vignette]() with additional instructions for using these platforms.
+(incorporate) the experimental branch.
+
+The R-implementation of WORCS uses the package `gert`, a 'Git' client for 'R'
+[@oomsGertSimpleGit2019], to connect a local project to a remote repository on
+'GitHub'. The `worcs` package also contains a user-friendly function to set
+the default user credentials, `git_user()`, and a function to add, commit, and
+push all changed files in one step, `git_update()`. It is worth noting that,
+although this paper focuses on 'GitHub', which is the most widely used 'Git'
+cloud storage service, two notable alternative services exist: 
+'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). As of
+version 0.1.2, the `worcs` template dialog accepts 'https' addresses of 
+repositories on both of these platforms interchangeably. The package has a
+[vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) with
+additional instructions for using these platforms. However,
+`worcs` has been most extensively tested with 'GitHub'.
+
+### Dependency management
 
@@
 sufficient to state that analyses were conducted in *Program X*. However, every
-program is susceptible to changes, updates, and bugfixes. R-packages in
-particular are updated frequently, because they are Open Source, and there is an
+program is susceptible to changes, updates, and bugfixes. Open Source software,
+in particular, is regularly updated because there is an
 active community of developers contributing functionality and bugfixes. 
-Potentially, any update could change the results of the code. <!--For example, 
-R-version 3.6.0 fixed a bug in the random number generator. This means that any
-analyses that use random numbers, such as bootstrapping, cannot be strictly
-replicated from previous versions. Or, for example, R-version 4.0.0 changed a
-default parameter that affects how data are loaded from files.-->
+Potentially, any such update could change the results of the code, thus rendering
+the analysis 'computationally non-reproducible'.
 
@@
 are explicitly installed, and maintaining a record of these programs. These solutions differ in
-user-friendliness and effectiveness. If strict reproducibility is required, one might use the
-aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all
-analyses inside a "container": An environment that behaves like a virtual computer that can be
-stored like a sort of time capsule, and identically reinstated on a user computer, or in the cloud.
-This is effective, and also preserves software outside the R environment (e.g., preprocessing tools
-for neuroimaging data). However, it can be difficult to set up for novice users - although a package
-is in development to facilitate this transition [@repro]. The cloud-based research collaboration platform
-'Code Ocean' is also based on Docker. It is very user-friendly, but it is not clear whether the
-["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/) is fully
-GDPR compliant - it seems to focus more on limiting liability than on compliance. This is a risk with
-any cloud-computing solution.
+user-friendliness and effectiveness. If strict reproducibility is required, users might choose to conduct all analyses in a 'Docker container': An environment that behaves like a
+virtual computer, which can be stored like a sort of time capsule, and
+identically reinstated on a user computer, or in the cloud.
+This is effective, and also preserves software outside the R environment (e.g.,
+preprocessing tools for neuroimaging data). An excellent Docker-based workflow
+was recently introduced by @peikertReproducibleDataAnalysis2019. This workflow
+is based on many of the same principles as WORCS, and the two approaches are
+conceptually and practically compatible. However, the workflow described by 
+Peikert and Brandmaier can be difficult to set up for novice users. An R-package
+is currently in development to facilitate this transition [@repro]. Another 
+user-friendly Docker-based workflow is offered by the cloud-based collaboration
+platform 'Digital Ocean'. Although Digital Ocean enjoys some popularity among 
+scientists, it is not clear whether this platform's
+["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/)
+is GDPR compliant - it seems to focus more on limiting liability than on
+compliance. The necessity to upload data for analysis is a risk with any
+cloud-computing solution when human participants are involved.
 
-For analyses conducted entirely in R, a slightly less sophisticated solution is offered by the R-package `packrat`, which installs all packages
-used in a project to a dedicated library that can be stored and copied with the project. This has the disadvantage that for each project, all packages
-must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space.
-Moreover, 'GitHub' is not designed to store your entire project library, so `packrat` does not address
-the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is
-somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing
-all R-packages used, it simply records the calendar date, and when another user loads your project, it installs the version of all required
-packages that was available on the central R repository `CRAN` on that date from an archival website.
-Thus, all that is needed to restore a project is the date on which it was conducted. The main limitation is that this solution is limited to `CRAN`
-packages, and cannot mix package versions from different dates.
-
-A recently developed package strikes a great balance between user-friendliness 
-and effectiveness: `renv`, developed by the team behind 'RStudio'. This package
-maintains a text-based, human-readable log of the packages used, their version
-numbers, and where they were installed from (e.g., 'CRAN', 'Bioconductor', 'GitHub').
-This text-based log can be version controlled with 'Git'. `renv` installs all
-packages in a cache that is shared between all projects that use `renv`. Thus,
-if one of your projects uses a package that is already in the cache, it does not
-need to be installed again. This overcomes the limitations of long installation
-times and large space requirements of `packrat`. When someone else loads your 
+For analyses conducted entirely in R, the recently released package `renv`
+strikes a great balance between user-friendliness  and effectiveness. Developed by
+the team behind 'RStudio', this package simply maintains a text-based,
+human-readable log of all packages used, their version numbers, and where they
+were installed from (e.g., 'CRAN', 'Bioconductor', 'GitHub'). As long as these
+repositories maintain their archives, this solution is reliable and very
+lightweight.
+The text-based log file can be version controlled with 'Git'. `renv`
+installs the logged packages in a cache that is shared between all
+projects on the computer. Thus, if one of your projects uses a package that is
+already in the cache, it does not
+need to be installed again. This prevents long installation
+times, large space requirements, and frustration having to re-install software
+that is already installed on the system. When someone else loads your 
 project, `renv` will install all of the required packages onto a cache on their
-computer. Whereas `packrat` copies the source code of all packages, `renv`
-obtains the packages from their original repositories. Thus, as long as these
-repositories are maintained, this solution is very lightweight and effective.
-It is not foolproof, but all solutions for computational
-reproducibility have a limited shelf life [@brownHowLearnedStop2017].
-Ultimately, the best solution might be to use a good-enough solution, like
-`renv`, and acknowledge that all code requires some maintenance if you want to
-reproduce it in the future.
+computer. 
+This is not a foolproof method to ensure strict computational reproducibility, but it is more user-friendly
+and lightweight than any alternative out there, requiring no technical skill to
+setup. When choosing an appropriate solution for dependency management, it is
+important to consider that all solutions have a limited shelf life
+[@brownHowLearnedStop2017].
+Ultimately, the best approach might be to use a "just good-enough" solution, and
+acknowledge that all code requires some maintenance if you want to
+reproduce it in the future. For projects in R, `renv` meets these requirements.
 
@@
 A key consideration when developing a research project is what filetypes to use.
-A case can be made to use text-based files only (or as much as possible),
+A case can be made to use text-based files exclusively (or as much as possible),
 instead of binary files.
@@
 Binary files, such as 'Word' (`.docx`) or 'SPSS' (`.sav`, `.spo`) files,
-must be decoded by (commercial) software first.
-'Git' is designed to track any changes to the content of text-based files.
-Each change reflects a line of text added or removed. This means that, when
+must be decoded first, and many require commercial software to do so.
+Text-based files are also required when using 'Git' for version control.
+In 'Git', each change reflects a line of text added or removed. This means that, when
 writing a version-controlled paper in plain-text, it is recommended to insert a
@@
 Binary files should not be version controlled with 'Git',
-as changes in the encrypted file are uninterpretable. Binary files are often 
-also larger than text-based files, which means they take up more space on cloud
-hosting services, such as 'GitHub'. For those reasons, uploading large binary
+as changes in the encrypted file are uninterpretable. Binary files are also often 
+larger than text-based files, which means they take up more space on cloud
+hosting services, such as 'GitHub'. For these reasons, uploading large binary
 files to these platforms is frowned upon.
-'GitHub' does allow visitors to peruse text-based files online, and
+'GitHub' encourages the use of text-based files by allowing visitors of their website to peruse
+text-based files online, and
 conveniently renders certain filetypes: For example, 'GitHub' displays `.md`
 (Markdown) files as web pages, and `.csv` files as spreadsheets. This is why
-`worcs` uses Markdown and `.csv` extensively.
+`worcs` uses text-based file types like Markdown and `.csv` extensively.
 
@@
 
-Before you can use the WORCS workflow, you have to install the required software.
-You have to execute this setup procedure only once for each computer you intend to use `worcs` on.
-The procedure should take about 30 minutes, and it is documented in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
-After preparing your system, you can use the WORCS workflow for all of your projects.
-This step-by-step procedure is detailed in [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
-In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r cran_version()`) below.
+Before you can use the R-implementation of the WORCS workflow, you have to install the required software.
+You only have to perform this setup procedure only once for each computer you
+intend to use `worcs` on. It should take about 30 minutes, and it is documented
+in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
+The setup vignette does not assume a prior installation of R, so it is suitable for novice users.
+After setting up your system, you can use the WORCS workflow for all of your
+projects. This step-by-step procedure is detailed in [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
+In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r worcs:::cran_version()`) below.
 We highly recommend users to refer to [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html) for the most up-to-date procedure (subject to improvements).
@@
 
-TOP encourages comprehensive citation of literature, data, materials, methods,
+The TOP-guidelines encourage comprehensive citation of literature, data, materials, methods,
 and software. In principle,
@@
 
-<!-- AL: Some general things about software citation that could be good to mention here: Tool or software papers are also common, often as short papers, sometimes called "Application Notes".
-The Journal of Open Source Software (JOSS) is a way of turning software directly into a publication, without the need of writing an actual paper about it (the review is about the code).
-https://citation-file-format.github.io/ is a file format to include software citation information into source code repositories.
-CJ: I don't think this is so relevant, because WORCS is geared towards doing research,
-not writing software! If people use JOSS for R-packages, like I do for tidyLPA,
-then that citation will show up when users run citation("tidyLPA") --> 
-When using R, software (R packages) can be conveniently cited by calling
-`citation("packagename")`. This returns an APA-style reference, and a BibTeX
-entry. In addition to the manual citation of important packages, the `worcs`
-project template for RStudio offers a comprehensive record of all dependencies.
-WORCS generates a template manuscript in APA6 style, based on the `papaja` package [@austPapajaPrepareReproducible2020], which automatically cites all packages used.
+References for software are sometimes provided within the software environment;
+for example, in R, package references can be found by
+calling `citation("packagename")`. This returns an APA-style reference, and a
+BibTeX entry. Software papers are also common, and are sometimes called
+"Application Notes". The [Journal of Open Source Software (JOSS)](https://joss.theoj.org/)
+offers programmers a way to generate a publication 
+based on well-documented software;
+in this process, the code itself is peer reviewed, and writing an actual paper is optional. For an example, see
+@rosenbergTidyLPAPackageEasily2018.
+Software citations can also sometimes be found 
+in the source code repository (e.g., on 'GitHub'), using the
+['Citation File Format'](https://citation-file-format.github.io/).
 
@@
 The print version should reference the online version, so interested readers can find the comprehensive reference list.
-Step 20 of the WORCS procedure suggests uploading the online version to a preprint server.
+The WORCS procedure suggests uploading the online version to a preprint server.
 This is important because most preprint servers - at least those hosted by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
@@
 
-It is easiest to mark the distinction between essential and non-essential references from the start, instead of going
-back to cut references prior to publication. Therefore, the `worcs` package distinguishes between the
-traditional at-symbol (`@`) used to cite a reference, and the "double at"-symbol (`\@\@`), used to cite a non-essential reference.
-Users can render a manuscript to PDF either with, or without, comprehensive citations.
+## Implementation in `worcs`
 
-The procedure for citation in WORCS is as follows:<!-- LB: this procedure is secondary to the main workflow. Consider changing the order (see main issue)-->
+The `worcs` package includes a [vignette on citation](https://cjvanlissa.github.io/worcs/articles/citing.html) to explain the
+fundamentals for novice users, and offer recommendations specific to `worcs`. We
+recommend using the free, open-source reference manager
+[Zotero](https://www.zotero.org/download/); it is feature-rich, 
+user-friendly, and highly interoperable with other reference managers. A 
+tutorial for using Zotero with RMarkdown exists  [here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-rmarkdown/).
 
-1. During writing, maintain a plain-text `.bib` file with the BibTeX references for all citations.
-    + You can make this file by hand; e.g., Figure \@ref(fig:scholarbib) shows how to obtain a BibTeX reference from Google Scholar; simply copy-paste each reference into the `.bib` file
-    + You can export a `.bib` file from most reference manager programs; the free, open-source reference manager [Zotero](https://www.zotero.org/download/) is excellent and user-friendly, and highly interoperable with other commercial reference managers.
-[Here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-rmarkdown/) is a tutorial for using Zotero with Rmarkdown.
-2.
-To cite a reference, use the `citekey` - the first word in the BibTeX entry for that reference.
-Insert it in the Rmarkdown file like so: `@yourcitekey2020`.
-For a parenthesized reference, use `[@citekeyone2020; @citekeytwo2020]`.
-For more options, see the [Rmarkdown cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html).
-3. To indicate a *non-essential* citation, mark it with a double at-symbol: `\@\@nonessential2020`.
-4. When Knitting the document, adapt the `knit` command in the YAML header.  
-  `knit: worcs::cite_all` renders all citations, and  
-  `knit: worcs::cite_essential` removes all *non-essential* citations.
-5.
-Optional: To be extremely thorough, you could make a "branch" of the GitHub repository for the print version of the manuscript.
-Only in this branch, you use the function `knit: worcs::cite_essential`.
-The procedure is documented in [this tutorial](http://rstudio-pubs-static.s3.amazonaws.com/142364_3b344a38149b465c8ebc9a8cd2eee3aa.html).
+The `worcs` package also offers original solutions for comprehensive citation.
+Firstly,
+during the writing process, authors can mark the distinction between essential
+and non-essential references. It is easier to do this from the start, instead of
+going back to cut non-essential references just prior to publication. Standard
+Markdown uses the at-symbol (`@`) to cite a reference. `worcs` additionally 
+reserves the "double at"-symbol (`@@`) to cite a non-essential references.
+Users can render the manuscript either with, or without, comprehensive citations
+by adapting the `knit` command in the front matter, setting it to
+`knit: worcs::cite_all` to render all citations, and to 
+`knit: worcs::cite_essential` to remove all *non-essential* citations.
 
-```{r, scholarbib, echo = FALSE, fig.cap="Exporting a BibTex reference from Google Scholar"}
-knitr::include_graphics("scholar_bib.png")
-```
+With regard to the citation of R-packages, it is worth noting that the `papaja` 
+package [@austPapajaPrepareReproducible2020], which provides the 'APA6'
+template used in `worcs`, also
+includes functions to automatically cite all packages used in the R-session, and
+add their references to the bibliography file.
 
@@
 
-Data sharing is important for computational reproducibility and secondary analysis.
-Computational reproducibility means that a third party can exactly recreate the results from the original data, using the published analysis code.
-Secondary analysis means that a third party can conduct sensitivity analyses, explore alternative explanations, or even use existing data to answer a different research question.
-From an open science perspective, data sharing is always desirable.
-From a practical point of view, it is not always possible.
-Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns.
-For example, one concern that applies when working with human participant data, is pseudonimization.
-It is crucial to pseudonimize data as soon as they are collected, by removing or deleting any sensitive personal information and contact details.
-Pseudononimized data can often be shared if participants have provided informed consent to that effect.
-Nevertheles, it is important to note that the European GDPR prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU, unless it offers an "adequate level of protection".
-Although different rules may apply to pseudonimized data, there are many repositories that are GDPR compliant, such as the European servers of the Open Science Framework.
-Different Universities, countries, and funding bodies also have their own repositories that are complient with local legistation.
-Before sharing any human participant data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.
+Data sharing is important for computational reproducibility and secondary
+analysis. Computational reproducibility means that a third party can exactly
+recreate the results from the original data, using the published analysis code.
+Secondary analysis means that a third party can conduct sensitivity analyses,
+explore alternative explanations, or even use existing data to answer a
+different research question. From an open science perspective, data sharing is
+always desirable. From a practical point of view, it is not always possible.
+Data sharing may be impeded by lack of storage space, legal constraints, ethical restrictions, absence
+of informed consent for data sharing, or privacy concerns. For example, one
+concern that applies when working with human participant data, is
+pseudonimization. 
+Many legislatures require researchers to 'de-identify' human subject data upon collection, by removing or deleting any sensitive personal information and
+contact details. However, when researchers have an imperfect understanding of the
+legal obligations and dispensations, they can end up over- or undersharing, and
+risk either placing participants at risk of being identified, or undermine the
+replicability of their own work and potential reuse value of their data [@phillipsDiscombobulationDeidentification2016].
+
+Once data are de-identified or pseudononimized, they can often be shared if participants have
+provided informed consent to that effect. Nevertheles, it is important to note
+that the European GDPR prohibits storing 'personal data' (information which can
+identify a natural person whether directly or indirectly) on a server outside
+the EU, unless it offers an "adequate level of protection". Although different
+rules may apply to pseudonimized data, there are many repositories that are
+GDPR compliant, such as the European servers of the Open Science Framework.
+Different Universities, countries, and funding bodies also have their own
+repositories that are complient with local legistation. Before sharing any
+human participant data, it is recommended to obtain approval from an internal
+(ethical) review board, guidance from Research Data Management Support, and
+informed consent from participants.
 
@@
 after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->
-If data can be shared openly, it can simply be committed it to the 'Git'
+If data can be shared openly, they can simply be pushed to a public 'Git'
 repository, along with the analysis code. This way, others can download the
@@
 If data cannot be shared, researchers should aim to safeguard the potential for
-computational reproducibility and secondary analysis as much as possible. The
-`worcs` package uses two solutions to accomplish this goal.
+computational reproducibility and secondary analysis as much as possible. WORCS
+recommends two solutions to accomplish this goal.
 The first solution is to publish a *checksum* of the original data file.
@@
 the authors who evaluate it on the real data and send back the results. The
-`worcs` package offers a simple but flexible function for synthetic data,
+`worcs` package offers a simple but flexible function to generate synthetic data,
 `synthetic()`, which by default uses random forests to generate a model for
@@
 properly, it is essential that the `class` (data type) of variables is defined
-correctly. As of Version `r cran_version()`, numeric, integer, and factor types
+correctly. As of `worcs` version `r worcs:::cran_version()`, numeric, integer, factor, and logical data
 are supported out of the box. Other types of variables should be converted to
@@
 
-Regardless of whether the data will be open or closed, it is important that the raw data will be left unchanged as much as possible.
-This eliminates human error.
-Any alterations to the data - including processing steps and even error corrections - should be documented in the code, instead of applied to the raw data.
+Regardless of whether the data will be open or closed, it is important that the
+raw data be left unchanged as much as possible. This eliminates human error.
+Any alterations to the data - including processing steps and even error 
+corrections - should be documented in the code, instead of applied to the raw data.
 This way, the code will be a pipeline from the raw data to the published results.
-Every `worcs` project contains an empty R-script called `prepare_data.R`.
-As soon as the data are collected, researchers should use this file to document the steps necessary to load the data into R, pseudonimize it, and prepare it for analysis.
-For example, if the data was originally in SPSS format with IP addresses and GPS location, this file might just contain the code required to read the SPSS file, and to remove those columns of sensitive personal information.
-As soon as the data are pseudonimized and processed into a 'tidy' format [i.e., data that can be stored as a spreadsheet, see @wickhamTidyData2014], the researcher should version control some indelible record of the raw data, and avoid making any further changes to the version controlled file.
+Thus, for instance, if data have been entered manually in a spreadsheet, and 
+a researcher discovers that in one of the variables, missing values were coded as
+`-99` whereas they were coded as `-9` in all other variables, then we would recommend adding
+a line of code to recode the missing values - thereby documenting the mistake - 
+instead of "fixing" it by editing the raw spreadsheet.
 
-The `worcs` package offers two functions for version controlling a record of the data: One for open, and one for closed data:
+When initializing a new `worcs` project, an empty R-script called `prepare_data.R` is generated.
+As soon as raw data are collected, researchers should use this file to document
+all teps necessary to load the 
+data into R, pseudonimize it, and prepare it for analysis. 
+For example, if the data was originally in SPSS format with IP addresses and GPS
+location, this file might just contain the code required to read the SPSS file,
+and to remove those columns of sensitive personal information. As soon as the
+data are pseudonimized and processed into a 'tidy' format [i.e., data that can
+be stored as a spreadsheet, see @wickhamTidyData2014], the researcher should
+version control some indelible record of the raw data.
 
-* `open_data(data)`: This function stores the `data` object (a `data.frame` or `matrix`) in a spreadsheet file (called `data.csv` by default).
-This file should now be (manually) added to the 'Git' repository, committed, and pushed to GitHub.
-Once the GitHub repository is made public, these data will be open to the public.
+The `worcs` package offers two functions for version controlling a record of the
+data: One for open, and one for closed data. Researchers should assume that the
+decision to make data open is irreversible. Thus, the decision should be made well
+before arriving at this stage in the workflow. If there is any doubt, it is prudent
+to proceed with closed data and make the data open once all doubt is assuaged.
+
+In `worcs`, researchers can call the function `open_data(data)` to make the data
+publicly available. This function stores the `data` object (a `data.frame` or
+`matrix`) in a spreadsheet file (called `data.csv` by default), and generates an
+accompanying codebook that can be viewed online.
+All changed files should now be (manually) added to the 'Git' repository, committed, and
+pushed to GitHub. Once the GitHub repository is made public, these data will be
+open to the public.
 Assume that, once you do this, the data cannot be un-shared.
-* `closed_data(data)`: This function also stores the data in a local file (called `data.csv` by default), but that file is added to the `.gitignore` file so it cannot be accidentally added to the 'Git' repository.
-The function also adds a checksum for the original data in the `.worcs` project file, and creates a synthetic dataset (called `synthetic_data.csv` by default).
-The user should commit all changed files and push them to GitHub.
-Once the GitHub repository is made public, people will have access to a checksum for the original data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.
+
+Alternatively, if the project requires data to remain closed, researchers can
+call `closed_data(data)`. This function also stores the data in a local `.csv` file
+and generates a codebook,
+but the original data file is added to `.gitignore` so it cannot be
+accidentally added to the 'Git' repository.
+The function also computes a checksum for the original data, and logs it in the
+`.worcs` project file. This checksum serves as the indelible record of the state 
+of the raw data.
+The function `closed_data()` also creates a synthetic dataset (called
+`synthetic_data.csv` by default).
+The user should push all changed files 'GitHub'.
+Once the 'GitHub' repository is made public, people will have access to a
+checksum for the original data that exist on your local machine, a codebook
+describing those data, and a synthetic copy of the data.
 
@@
 
-The final line of the `prepare_data.R` file should thus be either `open_data()` or `closed_data()`.
+As the purpose of the `prepare_data.R` file is to prepare data for sharing in a
+format as close to "raw data" as possible, the final line of this file
+should typically be either
+`open_data()` or `closed_data()`. 
 The analysis code, on the other hand, should begin with `load_data()`.
-This function loads the real data if it is available on the user's computer, and otherwise, loads the synthetic data.
-This will have the effect that third party users who copy the GitHub repository will automatically load the synthetic dataset, whereas the study authors, who have the real data stored locally, will automatically load the real data.
-This function makes it possible for reviewers, coauthors, and auditors, to write analysis scripts without requiring access to the original data.
-They can simply start the script with `load_data()`, and write their code based on the synthetic data.
-Then, they can submit their code to you - by email or as a ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) on GitHub.
-When you run their code on your system, `load_data()` will detect the original data, and use that to run their script.
-You can then return the results to the third party.
+This function loads the real data if it is available on the user's computer, and
+otherwise, loads the synthetic data. This will have the effect that third party
+users who copy a 'GitHub' repository with closed data will automatically load the synthetic
+dataset, whereas the study authors, who have the real data stored locally, will
+automatically load the real data. This function makes it possible for reviewers,
+coauthors, and auditors, to write analysis scripts without requiring access to
+the original data. They can simply start the script with `load_data()`, and
+write their code based on the synthetic data. Then, they can submit their code
+to you - by email or as a
+["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request)
+on GitHub. When you run their code on your system, `load_data()` will detect the
+original data, and use that to run their script. You can then return the results
+to the third party.
 
-As of version 0.1.1, the `worcs` package only stores data as `.csv` files.
-As explained above, this is intentional, because `.csv` files are human- and machine readable, and because data typically need to be in a spreadsheet format for analysis.
-Many types of data can be represented as a spreadsheet - text corpora can have one row per document, and EEG- or EKG waveforms can have one row per measurement occasion.
-The `prepare_data.R` file should document any steps required to convert these data into tabular format.
-If it is data must be stored in a different file type anyway, readers are encouraged to follow the development of the `repro` package [@repro], which will enable such enhancements.
+As of version `r worcs:::cran_version`, the `worcs` package only stores data as `.csv` files.
+As explained above, this is recommended, because `.csv` files are human- and
+machine readable, and because data typically need to be in a spreadsheet format
+for analysis. Many types of data can be represented as a spreadsheet - text
+corpora can have one row per document, and EEG- or EKG waveforms can have one
+row per measurement occasion. The `prepare_data.R` file should document any
+steps required to convert these data into tabular format. If it is data must be
+stored in a different file type anyway, readers are encouraged to follow the
+development of the `repro` package [@repro], which will enable such enhancements.
 
@@
 
-When writing a manuscript in Rmarkdown, analysis code is embedded in the prose of the paper.
-Thus, the TOP-guideline of sharing analysis code can be met simply by committing the source code of the manuscript to GitHub, and making this GitHub repository Public upon publication.
-If authors additionally use open data and a reproducible environment (default in WORCS), then a third party can simply replicate all analyses by copying the entire GitHub repository, and Knitting the manuscript on their local computer.
+When writing a manuscript created using dynamic document generation, analysis
+code is embedded in the prose of the paper. Thus, the TOP-guideline of sharing
+analysis code can be met simply by committing the source code of the manuscript
+to the 'Git' repository, and making this repository Public, e.g., on 'GitHub'.
+If authors additionally use open data and a reproducible environment (as suggested
+in WORCS), then a third party can simply replicate all analyses by copying the
+entire repository from the 'Git' hosting service, and Knitting the manuscript on
+their local computer.
 
-Aside from analysis code, the TOP-guidelines also encourage sharing new research materials, and details of the study design and analysis.
-These goals can be accomplished simply by placing any relevant documents in the RStudio project folder, committing them to the Git repository, and pushing to GitHub.
-As with any document version controlled in this way, it is advisable to use plain text only.
+Aside from analysis code, the TOP-guidelines also encourage sharing new research
+materials, and details of the study design and analysis. These goals can be
+accomplished by placing any such documents in the 'Git' repository folder,
+committing them, and pushing to a cloud hosting service. As with any document
+version controlled in this way, it is advisable to use plain text only.
 
@@
 
-Lindsay and colleagues [-@lindsayResearchPreregistration1012016] define preregistration as "creating a permanent record of your study plans before you look at the data.
-The plan is stored in a date-stamped, uneditable file in a secure online archive." Two such archives are well-known in the social sciences: AsPredicted.org, and OSF.io.
-However, GitHub also conforms to these standards.
-Thus, it is possible to preregister a study simply by pushing a text document with the study plans to GitHub.
-By tagging the commit as a preregistration (see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)), it is distinct from all other commits, and easily findable by people and programs.
+Lindsay and colleagues [-@lindsayResearchPreregistration1012016] define
+preregistration as "creating a permanent record of your study plans before you
+look at the data. The plan is stored in a date-stamped, uneditable file in a
+secure online archive." Two such archives are well-known in the social sciences:
+[AsPredicted.org](https://aspredicted.org/), and [OSF.io](https://osf.io/).
+However, 'Git' cloud hosting services also conform to these standards.
+Thus, it is possible to preregister a study simply by pushing a text document
+with the study plans to, e.g., 'GitHub'.
+By tagging the commit as a preregistration
+(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
+it is distinct from all other commits, and easily findable by people and programs.
 
-The advantages, disadvantages, and pitfalls for preregistering different types of studies have been extensively debated, and that has been summarized elsewhere [e.g., @lindsayResearchPreregistration1012016].
-We take the pragmatic position that, in confirmatory (hypothesis-testing) research, it is beneficial to plan projects before executing them, to preregister these plans, and adhere to them.
-All deviations from this procedure should be disclosed; e.g., prior exposure to the data, whether direct (e.g., by computing summary statistics) or indirect (e.g., by reading papers using the same data); deviations from the analysis plan to handle unforseen contingencies, such as violations of model assumptions; or additional exploratory analyses.
+The advantages, disadvantages, and pitfalls for preregistering different types
+of studies have been extensively debated elsewhere [see @lindsayResearchPreregistration1012016].
+For example, because the practice of preregistration has historically been closely
+tied to experimental research,
+it has been a matter of some debate whether secondary data
+analyses can be preregistered 
+[but see @westonRecommendationsIncreasingTransparency2019 for an excellent
+discussion of the topic]. Preregistration is the only way to ensure that a
+researcher is not "HARKing": Hypothesizing after the results are known
+[@kerrHARKingHypothesizingResults1998].
+When analyzing existing data, it is difficult to *prove* that a researcher did
+not have direct (or indirect, through collaborators or by reading studies using
+the same data) exposure to the data, before composing the preregistration.
+However, the question of proof is only relevant from a perspective of preventing
+scientific misconduct.
+Good faith preregistration efforts always improve the quality of deductive
+(theory-testing) research, because they avoid HARKing, ensure reliable
+significance tests, avoid overfitting noise in the data, and limit the number of
+forking paths researchers wander during data analysis
+[@gelmanStatisticalCrisisScience2014].
+In all cases, researchers should minimize exposure to the data, and disclose any
+exposure.
 
-WORCS uses the following approach to preregistration:
+WORCS takes the pragmatic position that, in deductive (hypothesis-testing) research,
+it is beneficial to plan projects before executing them, to preregister these
+plans, and adhere to them.
+All deviations from this procedure should be disclosed; e.g., prior exposure to
+the data, whether direct (e.g., by computing summary statistics) or indirect
+(e.g., by reading papers using the same data); deviations from the analysis plan
+to handle unforseen contingencies, such as violations of model assumptions;
+or additional exploratory analyses.
 
-1. Document study plans in a `preregistration.Rmd` file, and optionally, planned analyses in a `.R` file.
-2. Commit these documents to the local Git repository, and push them to GitHub
-3. On GitHub, [tag the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration". A tagged release helps others retrieve this commit.
-4. Optional: Upload the files as attachments to another pre-registration repository, e.g., AsPredicted.org or OSF.io.
-  + The `.Rmd` file can be Knitted to a PDF form before uploading to a repository
+WORCS recommends documenting a verbal, conceptual description of the study plans
+in a text-based file. Optionally, an analysis script can be preregistered to
+documenting the planned analyses.
+If any changes must be made to the analysis code after obtaining the data,
+one can refer to the conceptual description to justify the changes.
+The ideal preregistered analysis script consists of a complete analysis
+that can be evaluated once the data are obtained.
+This ideal is often unattainable, because the data present researchers with
+unanticipated challenges;
+e.g., assumptions are violated, or analyses work differently than expected.
+Some of these challenges can be avoided by simulating the data one expects to obtain,
+and writing the analysis syntax based on the simulated data.
+This topic is beyond the scope of the present paper, but many user-friendly
+methods for simulating data are available in most statistical programming languages.
+For instance, `R` users can use the package `simstudy` [@goldfeldSimstudySimulationStudy2020].
 
-If the GitHub repository is Public, the preregistration is immediately visible to the world, and reviewers (both formal reviewers designated by a journal, and informal reviewers recruited by other means) can submit comments through pull requests.
-If the repository is Private, Reviewers can be invited as "Collaborators".
-Private repositories can be made public at a later date, along with their entire time-stamped history.
-It is important to note that contributing to the repository in any way will void reviewers' anonymity, as their contributions will be linked to a user name.
+To preregister a project using 'Git', one can commit
+all preregistration documents to the local 'Git' repository, 
+and push them 
+to a remote 'Git' hosting service.
+By [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration" on the remote repository, it can be easily identified as
+such and retrieved. This simple but robust approach renders pre-registration through
+preregistration services, such as AsPredicted.org or OSF.io, optional.
+However, it is easy to take this optional step when the preregistration is
+written using dynamic document generation, as 
+recommended in WORCS: This file can be formatted as PDF and uploaded as an attachment
+to the formal preregistration service.
 
-### Preregistering study plans
+As soon as a project is preregistered on a Public 'Git' repository, it is
+visible to the world, and reviewers (both formal reviewers designated by a
+journal, and informal reviewers recruited by other means) can submit comments,
+e.g., through pull requests on 'GitHub'. If the repository is Private, Reviewers
+can be invited as "Collaborators". Private repositories can be made public at a
+later date, along with their entire time-stamped history and tagged releases.
+It is important to note that contributing to the repository in any way will void
+reviewers' anonymity, as their contributions will be linked to a user name.
 
-Several excellent preregistration templates are available in the R-package `prereg` [@austPreregMarkdownTemplates2019], including templates from organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
-When opening a new RStudio project with the WORCS project template, it is possible to select one of these preregistration templates, which will generate a file called `preregistration.Rmd`.
-This file should be used to document study plans, ideally prior to data collection.
+### Implementation of preregistration in `worcs`
 
-Because the practice of preregistration is strongly tied to the collection of primary data, it has been a matter of some debate whether secondary data analyses can be preregistered [but see @westonRecommendationsIncreasingTransparency2019 for an excellent discussion of the topic].
-Preregistration is the only way to ensure that a researcher is not "HARKing": Hypothesizing after the results are known [@kerrHARKingHypothesizingResults1998].
-When analyzing existing data, it is difficult to *prove* that a researcher did not have direct (or indirect, through collaborators or by reading studies using the same data) exposure to the data, before composing the preregistration.
-However, the question of proof is only relevant from a perspective of preventing scientific misconduct.
-Good faith preregistration efforts always improve the quality of deductive (theory-testing) research, because they avoid HARKing, ensure reliable significance tests, avoid overfitting noise in the data, and limit the number of forking paths researchers wander during data analysis [@gelmanStatisticalCrisisScience2014].
-In all cases, researchers should minimize exposure to the data, and disclose any exposure.
+Several excellent preregistration templates are available in the R-package
+`prereg` [@austPreregMarkdownTemplates2019], including templates from
+organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and
+from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
+When initializing an RStudio project with the `worcs` project template, one of
+these preregistration templates can be selected, which will generate a file
+called `preregistration.Rmd`. This file should be used to document study plans,
+ideally prior to data collection.
 
-### Preregistering analyses
+## Compatibility with other standards for open science
 
-The ideal preregistered analysis consists of a complete analysis script that can be evaluated once the data are obtained.
-This ideal is often unattainable, because the data present researchers with unanticipated challenges; e.g., assumptions are violated, or analyses work differently than expected.
-Some of these challenges can be avoided by simulating the data one expects to obtain, and writing the analysis syntax based on the simulated data.
-This topic is beyond the scope of the present paper, but many user-friendly methods for simulating data are available in `R`, including the package [`simstudy`](https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html).
-<!-- citation:
+Although WORCS was designed to meet the TOP-guidelines, it does so in a way that
+is compatible with other open science principles. Most notably, WORCS is 
+amenable to the
+FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a]. These principles advocate that research
+data should be Findable, Accessible, Interoperable and Reusable. They are
+increasingly applied as a standard for other types of research output as well
+[@lamprechtFAIRPrinciplesResearch2019].
+WORCS helps users meet the FAIR principles by making all research objects -
+including data and code - accessible in a repository that can be made public. 
+The `worcs` package additionally
+facilitates meeting the FAIR principles in several ways. First, each `worcs`
+project has a YAML file (https://yaml.org/) that lists the research objects in
+the project, which will allow web crawlers 
+to index `worcs` projects. Second, a codebook is automatically generated for
+each data file stored
+using `open_data()` or `closed_data()`. We encourage users to elaborate on the
+default codebook by adding
+variable description and categories, which would - in the future - enable
+indexing data by topic area. 
+Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is 
+hosted on the public version controlled repository 'GitHub', it has an open
+source licence (GPL v3.0), is registered in a community registry (CRAN), it
+enables the citation of the software (using the `citation("worcs")` command, or
+by citing this paper), and followed the CII Best Practices software quality
+checklist during development.
 
-  @Manual{,
-    title = {simstudy: Simulation of Study Data},
-    author = {Keith Goldfeld},
-    year = {2020},
-    note = {R package version 0.1.16},
-    url = {https://CRAN.R-project.org/package=simstudy},
-  }
-  -->
-When using the WORCS-workflow, it is recommended to preregister, if possible, a `.R` script with the planned analyses, along with a verbal, conceptual description.
-If any changes must be made to the R-script after obtaining the data, one can refer to the conceptual description to explain why these changes are necessary.
-
-<!--
-3. Write conceptual preregistration
-* https://osf.io/zab38/wiki/home/ Templates of OSF Registration Forms
-* R package prereg
-* mention other examples and cite van 't Veer, A.
-E., & Giner-Sorolla, R.
-(2016).
-Pre-registration in social psychology---A discussion and suggested template.
-Journal of Experimental Social Psychology, 67, 2--12.
-doi: 10.1016/j.jesp.2016.03.004
-3. Commit, push, and tag th as preregistration
-Preregistration: tag on GitHub
-https://stackoverflow.com/questions/18216991/create-a-tag-in-a-github-repository
-
-4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
-4. Collect data and either add to repository, or register hash of original datafile
-5.
-Conduct analyses and write; commit regularly (ideally, after completing a small and clearly defined task), with informative commit messages.
-Call renv::snapshot() to save the state of the project library to the lockfile (called renv.lock),
-6. Create OSF project, connect to GitHub, mention OSF page in manuscript
-7. Optional: Publish preprint in a not-for-profit preprint repository (check Sherpa Romeo) http://sherpa.ac.uk/romeo/index.php
-7. Submit
-8. Make repository public -->
+Sharing all research objects, as advocated in WORCS, also provides a proper
+basis for research evaluation
+according to the San Francisco Declaration on Research Assessment (DORA;
+https://sfdora.org/), which plays an increasing role in grant funding, hiring, 
+and promotion procedures. Direct access to research objects allows stakeholders
+to evaluate research quality based on content rather than relying on spurious
+surrogate indicators like journal impact factors, conference rankings, and
+h-indexes. The detailed version control and commit tracking of 'GitHub' and
+similar platforms furthermore make it possible to assess the relative
+contributions made by different researchers.
 
@@
 
-In this tutorial paper, we have presented a workflow for open reproducible code in science.
+In this tutorial paper, we have presented a workflow for open reproducible code
+in science.
 The workflow aims to lower the threshold for grass-roots adoption of open science principles.
-It is accompanied by an R-package with convenience functions, and a WORCS project template for RStudio.
-This relatively light-weight workflow meets most of the requirements for open science, as detailed in the TOP-guidelines.
+The workflow is supported by an R-package with an RStudio project template and 
+convenience functions.
+This relatively light-weight workflow meets most of the requirements for open science
+as detailed in the TOP-guidelines, and is compatible with other open science guidelines.
+The workflow helps researchers meet existing requirements for open science,
+and to reap the benefits of working openly and reproducibly, 
+even when such requirements are absent.
 
-## Differences from existing solutions
+## Comparing WORCS to existing solutions
 
@@
 authors' disclosures about, and incentivize adherence to, open science practices.
+WORCS has its own checklist,
 

### 2020-05-25 paper/Manuscript.Rmd
Further improvements

@@
 abstract: |
-  Technological advances enable researchers to more easily conduct reliable,
-  cumulative, collaborative science. Capitalizing on these advances has the
-  potential to accelerate scientific progress. However, adopting open science
+  Adopting open science
   principles can be challenging and costly, because doing so requires
@@
   parallel to, or in absence of, top-down requirements by journals,
-  institutions, and funding bodies. To further facilitate the transition, the
-  WORCS principles are implemented in an R-package, which offers an RStudio 
+  institutions, and funding bodies. To facilitate widespread adoption, the
+  WORCS principles are implemented in the R-package `worcs`, which offers an
+  RStudio 
   project template that is, effectively, a "one-click solution" for R users. 
@@
 [@tennantOpenScienceJust2018].
-Moreover, open science creates opportunities for researchers to more easily
+Open science creates opportunities for researchers to more easily
 conduct reliable, cumulative, and 
@@
 
-Many individual researchers are motivated to adopt best practices for open
+Many researchers are motivated to adopt best practices for open
 science and enjoy these benefits. And yet, the question of "how" to 
-adopt open science practices can be daunting. Making the transition
+apply open science practices can be daunting. Making the transition
 requires researchers to become knowledgeable about different open science 
 challenges and solutions, and to become proficient with new and unfamiliar tools.
-This paper is designed to ease that transition by providing a simple workflow 
-that meets most requirements open science, is based on best practices, and 
-is supported by user-friendly software in R: The Worfkflow for Open Reproducible 
+This paper is designed to ease that transition by presenting a simple workflow 
+that meets most requirements open science, and is based on best practices:
+The Worfkflow for Open Reproducible 
 Code in Science (WORCS). WORCS is a lightweight approach to open science and
@@
 to produce work in accordance with best practices.
-Although the principles underlying
-the WORCS procedure are universal, they have been implemented for R users in the
-R-package `worcs`, which offers an RStudio project template that is, effectively,
+Although the principles underlying this workflow are universal, WORCS has first
+been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
+The R-package `worcs` offers a project template for RStudio
+[@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
 a "one click solution". This paper introduces the WORCS workflow, discusses how
@@
 
-<!-- AL: I second LB's comment somewhere below that the overall WORCS workflow should be introduced much earlier, maybe already here.
-A diagrammatic overview/flowchart of the principal steps would be great.
-It should catch the reader's eye and give an immediate idea of what WORCS is about.
-Now it is indeed quite buried in the details that follow.
--->
-
-<!--
-thereby lowering the threshold for adoption.
-, and still ensures reproducibility in most circumstances. 
- It is easy to implement, thereby lowering the threshold
-
-* Grass-roots
-* open science as opportunity
-    - Easier work
-    - Less mistakes
-    - Better collaboration
-    - Cumulative science
-    - Error correction
-
-Benefits to individual scientists: Automation (after a learning curve), thus more productive, avoid mistakes, join projects, clone projects and make them your own.
-Collective benefits: Cumulative science, -->
-<!-- AL: These "value propositions" would also be good to discuss already in the introduction. Directly after the introduction of the WORCS overview would make sense, I think. -->
-
-<!-- AL: The remainder of the paper is structured as follows.
-...
-Or some variation of this obligatory end-of-intro paragraph.
-As the paper touches on a lot of aspects, from principal to very practical, it would be very useful to provide some guidance here.
-CJ: I was always taught that you should not explain your structure in the paper - you should just structure the paper. If the structure is unclear, make it better. -->
-<!--openness, transparency, and reproducible science -->
-
 # Defining Open Science practices
@@
 Although Open Science is advocated by many, it does not have a unitary definition
-[but, for an effort to ground open science in first principles, see 
+[for an effort to ground open science in first principles, see 
 @tennantValuePropositionOpen2020].
@@
 
-<!--## Goal of this paper
-
-The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
-However, large organizations are slow to change.
-For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that
-they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
-At the same time, <!-- edit -->
-<!--The formation of "Open Science Communities" at universities around the globe
-illustrates the momentum of this grassroots movement
-[@brinkmanWidescaleAdoptionOpen2020]. -->
-
-## Implementation in R
-
-Although the principles underlying this workflow are universal, WORCS has first
-been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
-Several arguments support the choice to work in R.
-First, R and all of it extensions are free open source software, which
-should be the tool of choice for open science.
-Second, as of this writing, R is the most advanced programming language in terms
-of the implementation of tools required for an open science workflow.
-Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
-Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
-Finally, R is highly interoperable:
-Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
-Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
-Fortran, and Python [@allaireMarkdownPythonEngine]. There are excellent free
-resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].
-
-Working with R is simplified immensely by using the RStudio integrated
-development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
-RStudio automates and streamlines tedious or complicated aspects of working with
-the tools used in WORCS, and many are embedded directly into the visual user
-interface.
-This makes RStudio a comprehensive solution for open science research projects.
-Another important feature of RStudio is project management.
-A project bundles writing, analyses, data, references, etcetera, into a
-self-contained folder, that can be uploaded entirely to 'GitHub' and downloaded
-by future users. The `worcs` R-package installs a new RStudio project template.
-When a new project is initialized from this template, the bookkeeping required
-to set up an open science project is performed automatically. The `worcs` 
-package has a [vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)
-on how to set up your system with all of the required tools.
-
 ## Introducing the tools
@@
 the analyses. The text is automatically formatted as a scientific paper in 
-several potential styles. When the text is formatted, the results of the code
+several potential styles. When the text is formatted, the code
 blocks are
-automatically generated and insterted in the text, or rendered as Figures and
+evaluated and their results are insterted in the text, or rendered as Figures and
 Tables. Dynamic document generation supercedes
@@
 analyses, and then manually copy-pasting analysis results into the text.
+The source code of this paper can serve as an example of DDG, and is
+[available here](https://github.com/cjvanlissa/worcs/blob/master/paper/Manuscript.Rmd).
 
@@
 a host of other scientific formats 
-[see @allaireRticlesArticleFormats2020], and as plain 'Markdown' for 'GitHub'.
+[see @allaireRticlesArticleFormats2020], and as plain 'Markdown'.
 
@@
 
-WORCS recommends using the *explicit* version control software 'Git'
+An integral part of WORCS is the *explicit* version control software 'Git'
 (www.git-scm.com). Git tracks
@@
 their own data, open "Issues" to ask questions or give feedback on the project,
-or even send a "Pull request" with suggested changes to the text or code for
+or even send a "Pull request" (a request to incorporate some code)
+with suggested changes to the text or code for
 your consideration. 'Git' and 'GitHub' shine as tools for collaboration, because
@@
 
+It is worth noting that,
+although this paper focuses on the most widely used 'Git'
+cloud storage service, 'GitHub', two notable alternatives exist: 
+'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). As of
+version 0.1.2, the `worcs` template dialog accepts 'https' addresses from 
+repositories hosted on either of these platforms interchangeably.
+The package has a
+[vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) with
+additional instructions for using alternative platforms. In 
+order to present the workflow in a platform-agnostic way, we will refer to such
+cloud hosting services as "remote repositories" in the remainder of this
+paper.
+However, `worcs` has been tested most extensively with 'GitHub'.
+
 The R-implementation of WORCS uses the package `gert`, a 'Git' client for 'R'
-[@oomsGertSimpleGit2019], to connect a local project to a remote repository on
-'GitHub'. The `worcs` package also contains a user-friendly function to set
+[@oomsGertSimpleGit2019], to connect a local project to a remote repository.
+The `worcs` package also contains a user-friendly function to set
 the default user credentials, `git_user()`, and a function to add, commit, and
-push all changed files in one step, `git_update()`. It is worth noting that,
-although this paper focuses on 'GitHub', which is the most widely used 'Git'
-cloud storage service, two notable alternative services exist: 
-'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). As of
-version 0.1.2, the `worcs` template dialog accepts 'https' addresses of 
-repositories on both of these platforms interchangeably. The package has a
-[vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) with
-additional instructions for using these platforms. However,
-`worcs` has been most extensively tested with 'GitHub'.
+push all changed files to a remote repository in one step, `git_update()`.
 
@@
 
-<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc.
-This could be a place where readers stumble and stop reading.
-Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective.
-[main issue] it should become clear earlier what the WORCS workflow entails: e.g.
-switching to working with a particular R Project Template.
-Discuss the ‘workflow' section earlier.
-Related: Can you make a schematic overview of the WORCS?overview-->
 Many solutions exist to ensure computational reproducibility. These solutions typically work by
@@
 larger than text-based files, which means they take up more space on cloud
-hosting services, such as 'GitHub'. For these reasons, uploading large binary
-files to these platforms is frowned upon.
-'GitHub' encourages the use of text-based files by allowing visitors of their website to peruse
-text-based files online, and
-conveniently renders certain filetypes: For example, 'GitHub' displays `.md`
-(Markdown) files as web pages, and `.csv` files as spreadsheets. This is why
-`worcs` uses text-based file types like Markdown and `.csv` extensively.
+hosting services. For these reasons, uploading large binary
+files to remote repositories is frowned upon. 
+WORCS encourages the use of
+text-based files for all purposes.
+It is worth noting that 'GitHub' also rewards the use of text-based files by
+rendering certain text-based filetypes: 
+For example, `.md`
+(Markdown) files are displayed as web pages, and `.csv` files as spreadsheets.
 
@@
 
-## Preparing your system
-
-Before you can use the R-implementation of the WORCS workflow, you have to install the required software.
-You only have to perform this setup procedure only once for each computer you
-intend to use `worcs` on. It should take about 30 minutes, and it is documented
-in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
-The setup vignette does not assume a prior installation of R, so it is suitable for novice users.
-After setting up your system, you can use the WORCS workflow for all of your
-projects. This step-by-step procedure is detailed in [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
-In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r worcs:::cran_version()`) below.
-We highly recommend users to refer to [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html) for the most up-to-date procedure (subject to improvements).
+We provide a conceptual outline of the WORCS procedure below, based on WORCS
+Version 0.1.2. 
+For users of the `worcs` R-package, we highly recommend referring to
+[the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
+which is more detailed and updated with future developments.
 
@@
 
-# How WORCS relates to the TOP-guidelines
+## The R-implementation of WORCS
+
+The R-package `worcs` automate most steps of the workflow
+for R users.
+Several arguments support the choice to implement this workflow in R.
+First, R and all of it extensions are free open source software, which
+should be the tool of choice for open science.
+Second, as of this writing, R is the most advanced programming language in terms
+of the implementation of tools required for an open science workflow.
+Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
+Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
+Finally, R is highly interoperable:
+Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
+Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
+Fortran, and Python [@allaireMarkdownPythonEngine]. There are excellent free
+resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].
+
+Working with R is simplified immensely by using the RStudio integrated
+development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
+RStudio automates and streamlines tedious or complicated aspects of working with
+the tools used in WORCS, and many are embedded directly into the visual user
+interface.
+This makes RStudio a comprehensive solution for open science research projects.
+Another important feature of RStudio is project management.
+A project bundles writing, analyses, data, references, etcetera, into a
+self-contained folder, that can be uploaded entirely to a 'Git' remote repository and downloaded
+by future users. The `worcs` R-package installs a new RStudio project template.
+When a new project is initialized from this template, the bookkeeping required
+to set up an open science project is performed automatically.
+
+### Preparing your system
+
+Before you can use the R-implementation of the WORCS workflow, you have to install the required software.
+You only have to perform this setup procedure only once for each computer you
+intend to use `worcs` on. It should take about 30 minutes, and it is documented
+in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
+The setup vignette does not assume a prior installation of R, so it is suitable for novice users.
+After setting up your system, you can use `worcs` for all of your
+projects. The R-implementation of the workflow introduced earlier is detailed,
+step-by-step, in
+[the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
+
+# How WORCS helps meet the TOP-guidelines
 
@@
 [*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/),
-aid in the publication of these data papers. For smaller projects, researchers
-sometimes publish their data in an online repository - like 'GitHub' - along
+aid in the publication of these data papers. For smaller projects, researchers often
+simply share the data in a remote repository, along
 with a text file with the preferred citation for the data (which can be a
@@
 Software citations can also sometimes be found 
-in the source code repository (e.g., on 'GitHub'), using the
+in the source code repository (e.g., on 'Git' remote repositories), using the
 ['Citation File Format'](https://citation-file-format.github.io/).
@@
 
-<!-- LB: As I reader, I would prefer to hear about data sharing options before the synthetic data part (which is only relevant to those who cannot share their data).
-E.g.
-can I indicate a point in my workflow (e.g.
-after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->
 If data can be shared openly, they can simply be pushed to a public 'Git'
@@
 All changed files should now be (manually) added to the 'Git' repository, committed, and
-pushed to GitHub. Once the GitHub repository is made public, these data will be
+pushed to the remote repository. Once the remote repository is made public, these data will be
 open to the public.
@@
 `synthetic_data.csv` by default).
-The user should push all changed files 'GitHub'.
-Once the 'GitHub' repository is made public, people will have access to a
+The user should push all changed files to the remote repository.
+Once the remote repository is made public, people will have access to a
 checksum for the original data that exist on your local machine, a codebook
@@
 otherwise, loads the synthetic data. This will have the effect that third party
-users who copy a 'GitHub' repository with closed data will automatically load the synthetic
+users who copy a remote repository with closed data will automatically load the synthetic
 dataset, whereas the study authors, who have the real data stored locally, will
@@
 to you - by email or as a
-["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request)
-on GitHub. When you run their code on your system, `load_data()` will detect the
+["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request).
+When you run their code on your system, `load_data()` will detect the
 original data, and use that to run their script. You can then return the results
@@
 
-As of version `r worcs:::cran_version`, the `worcs` package only stores data as `.csv` files.
+As of version 0.1.2, the `worcs` package only stores data as `.csv` files.
 As explained above, this is recommended, because `.csv` files are human- and
@@
 
-Some users may intend to make their data openly available through a restricted access platform, such as an institutional repository, but not publically through GitHub.
+Some users may intend to make their data openly available through a restricted access platform, such as an institutional repository, but not publicly through a 'Git' remote repository.
 In this case, it is recommended to use `closed_data()`, and to manually upload the original `data.csv` file to the restricted access platform.
-If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the GitHub repository as an [Add-on](https://help.osf.io/hc/en-us/categories/360001550973-Add-ons).
+If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the 'Git' remote repository as an [Add-on](https://help.osf.io/hc/en-us/categories/360001550973-Add-ons).
 
@@
 analysis code can be met simply by committing the source code of the manuscript
-to the 'Git' repository, and making this repository Public, e.g., on 'GitHub'.
+to the 'Git' repository, and making this remote repository Public.
 If authors additionally use open data and a reproducible environment (as suggested
@@
 Thus, it is possible to preregister a study simply by pushing a text document
-with the study plans to, e.g., 'GitHub'.
+with the study plans to a 'Git' remote repository.
 By tagging the commit as a preregistration
@@
 journal, and informal reviewers recruited by other means) can submit comments,
-e.g., through pull requests on 'GitHub'. If the repository is Private, Reviewers
-can be invited as "Collaborators". Private repositories can be made public at a
+e.g., through "Pull requests" on a 'Git' remote repository. 
+If the remote repository is Private, Reviewers
+can be invited as "Collaborators". Private remote repositories can be made public at a
 later date, along with their entire time-stamped history and tagged releases.
@@
 Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is 
-hosted on the public version controlled repository 'GitHub', it has an open
-source licence (GPL v3.0), is registered in a community registry (CRAN), it
+hosted on a version controlled public remote repository 'GitHub', has an open
+source licence (GPL v3.0), is registered in a community registry (CRAN),
 enables the citation of the software (using the `citation("worcs")` command, or
@@
 surrogate indicators like journal impact factors, conference rankings, and
-h-indexes. The detailed version control and commit tracking of 'GitHub' and
-similar platforms furthermore make it possible to assess the relative
+h-indexes. The detailed version control and commit tracking of 'Git' remote repositories
+furthermore make it possible to assess the relative
 contributions made by different researchers.
@@
 researchers can meet the requirements of open science.
-<!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example is the workflow for
+One notable example is the workflow for
 reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
@@
 
-```{r, echo = FALSE, eval = FALSE}
-tab1 <- data.frame(
-  Reference = c("Aalbers et al., 2018", "Peikert & Brandmaier, 2019"),
-  Scope = c("Disclose adherence to open science principles", "Completely reproducible analysis"),
-  Difficulty = c("Easy", "Hard"),
-  Effectiveness = c("Low", "High"),
-  "TOP-guidelines" = c("1-7", "2,3,5")
-)
-apa_table(tab1)
-```
-
-<!-- AL: Similar to my comment on the title, in this section it needs to make clear which scope of related work it covers.
-For example, in the life sciences there has also been a lot of work on practices for increasing reproducibility, with automated workflows, containerization, Research Objects, provenance protocols, and a lot more.
-I don't think it's necessary for this paper to discuss this, but that is under the provision that it defines its scope better.
--->
-
 ## Limitations
@@
 curve associated with the tools outlined in this paper. Learning to work with R,
-Rmarkdown, and Git/GitHub requires an initial time investment. The amount of
+Rmarkdown, and Git requires an initial time investment. The amount of
 time required is reduced by the availability of tutorials, such as this one, and
@@
 techniques. Working with Rmarkdown saves time and prevents mistakes by avoiding
-tedious copying of results into a text document. Working with Git/GitHub keeps
+tedious copying of results into a text document. Working with Git keeps
 projects organized, prevents accidental loss of work, enables integrating
@@
 
-Moreover, being familiar
-with GitHub opens doors to new forms of collaboration: In the open source
+Being familiar
+with 'Git' remote repositories like GitHub opens doors to new forms of collaboration:
+In the open source
 software community, continuous peer review and voluntary collaborative acts by

### 2020-05-25 paper/Manuscript.Rmd
Update paper and vignette

@@
   - name: "Caspar J. Van Lissa"
-    institution: "1,2"
-    orcid: "0000-0002-0808-5024"
+    affiliation: "1,2"
     corresponding: yes
@@
   - name: "Andreas M. Brandmaier"
-    institution: "3,4"
-    orcid: "0000-0001-8765-6982"
+    affiliation: "3,4"
+  - name: "Aaron Peikert"
+    affiliation: "3,5"
+  - name: "Barbara Vreede"
+    affiliation: "6"
+  - name: "Anna-Lena Lamprecht"
+    affiliation: "2,7"
+    email: "a.l.lamprecht@uu.nl"
   - name: "Loek Brinkman"
-    institution: "2,5"
-    orcid: "0000-0003-3997-1173"
-  - name: "Aaron Peikert"
-    institution: "3,6"
-    orcid: "0000-0001-7813-818X"
-  - name: "Barbara M.I. Vreede"
-    institution: "7"
-  - name: "Anna-Lena Lamprecht"
-    institution: "2,8"
-    orcid: "0000-0002-5023-4601"
+    affiliation: "2,8"
+  - name: "Marijn E. Struiksma"
+    affiliation: "2,9"
 
@@
   - id: "5"
+    institution: "Humboldt-Universitaet zu Berlin"
+  - id: "6"
+    institution: "University Library Utrecht"
+  - id: "7"
+    institution: "Utrecht University, Department of Information and Computing Sciences"
+  - id: "8"
     institution: "UMC Utrecht"
-  - id: "6"
-    institution: "Humboldt-Universitaet zu Berlin"
-  - id: "7"
-    institution: "University Library Utrecht"
-  - id: "8"
-    institution: "Utrecht University, Department of Information and Computing Sciences"
+  - id: "9"
+    institution: "Utrecht Institute of Linguistics OTS"
 
@@
   Adopting open science
-  principles can be challenging and costly, because doing so requires
+  principles can be challenging and time-intensive, because doing so requires
   substantial conceptual education and training in the use of new tools. The 
@@
 describing those data, and a synthetic copy of the data.
-
-<!-- Maybe remark that in the "closed" case, a the data should be backupped somewhere else, as GitHub will not do the job in this setting. -->
+Keep in mind that, when using closed data, the original data file exists only on
+the user's computer. Adequate provisions to back up the data, while respecting
+principles of responsible data stewardship, should be made.
 
@@
 Being familiar
-with 'Git' remote repositories like GitHub opens doors to new forms of collaboration:
+with 'Git' remote repositories opens doors to new forms of collaboration:
 In the open source

### 2020-05-25 paper/Manuscript.Rmd
update 2020-05-25 19:44:33

@@
 
-The WORCS procedure and accompanying R-package provide a user-friendly and lightweight workflow for open, reproducible research in R, that meets all TOP-guidelines.
-It is beyond the scope of WORCS to support the incorporation of tools outside of the R-environment, or to containerize a project so that it can be identically reinstated on a different system or virtual machine.
-One important area of future development is to enable these extensions of the workflow.
-To this end, we are integrating WORCS into a modular framework of reproducibility tools for R [@repro].
-This will make it possible for researchers to enhance a WORCS project with strict reproducibility tools like 'Docker' and 'Make', or vice versa, to initiate a WORCS project in a strictly reproducible environment.
-This combines the strengths of WORCS with those of @peikertReproducibleDataAnalysis2019.
+The WORCS procedure provides a user-friendly and
+lightweight workflow for open, reproducible research, that meets all
+TOP-guidelines. Nevertheless, there are clear directions for future developments.
+Firstly, although the workflow is currently implemented only in R, it is 
+conceptually relevant for users of other statistical programming languages, most
+notably Python [@10.5555/1593511].
+We welcome efforts to implement WORCS in other platforms.
+Secondly, even when a project is in R, it may have dependencies outside of the
+R environment that cannot be managed using `renv`.
+It is beyond the scope of the `worcs` package to support the incorporation of
+tools outside of R, or to containerize a project so that it can
+be identically reinstated on a different system or virtual machine.
+Therefore, another
+important area of future development is to enable such extensions of the
+workflow. To this end, we are collaborating with @repro on the development of
+`repro`, an R-package that will facilitate setting up reproducibility tools such
+as 'Make' and 'Docker'.
+This will make it possible for researchers
+to enhance a WORCS project with these strict reproducibility tools,
+or vice versa, to initiate a WORCS project within a strictly reproducible
+environment generated using `repro`. 
+This will combine the strengths of the WORCS with those of the workflow proposed
+by @peikertReproducibleDataAnalysis2019.
 

### 2020-05-25 paper/Manuscript.Rmd
update 2020-05-25 19:52:55

@@
   and coauthors are welcome. Please contact the lead author.
-  Acknowledgements: Jeroen Ooms (feedback, suggesting gert, and using Remotes
-  field).
+  We acknowledge Jeroen Ooms for offering feedback and suggesting gert.
 keywords          : "open science; reproducibility; r; dynamic document generation; version control"
-wordcount         : "5921"
+wordcount         : "8257"
 
@@
 Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
-Fortran, and Python [@allaireMarkdownPythonEngine]. There are excellent free
+Fortran, and Python [@allaireMarkdownPythonEngine2020]. There are excellent free
 resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].

### 2020-05-26 README.Rmd
update 2020-05-12 19:52:04
update 2020-05-12 20:03:31

@@
 
-# WORCS <a href='https://osf.io/zcvbs/'><img src='worcs_icon.png' align="right" height="139" /></a>
+# WORCS <a href='https://osf.io/zcvbs/'><img src='https://github.com/cjvanlissa/worcs/raw/master/docs/worcs_icon.png' align="right" height="139" /></a>
 
@@
 [![Travis build status](https://travis-ci.org/cjvanlissa/worcs.svg?branch=master)](https://travis-ci.org/cjvanlissa/worcs)
+[![CII Best Practices](https://bestpractices.coreinfrastructure.org/projects/3969/badge)](https://bestpractices.coreinfrastructure.org/projects/3969)
+[![CRAN status](https://www.r-pkg.org/badges/version/worcs)](https://CRAN.R-project.org/package=worcs)
+[![metacran downloads](https://cranlogs.r-pkg.org/badges/worcs)](https://cran.r-project.org/package=worcs)
 <!--[![Codecov test
@@
 
-WORCS is actively recruiting collaborators with valuable expertise, to improve both the workflow and the package.
-Relevant contributions warrant coauthorship to the paper.
+The WORCS paper and `worcs` package are open to contributions from collaborators with valuable expertise, to improve both the workflow and the package.
+Relevant contributions warrant coauthorship to the paper or package.
 Please contact the lead author at [c.j.vanlissa@uu.nl](mailto:c.j.vanlissa@uu.nl), or:
@@
 Code of Conduct v2.0](https://www.contributor-covenant.org/).
+Contributions to the package must adhere to the [tidyverse style guide](https://style.tidyverse.org/).
+When contributing code, please add tests for that contribution to the `tests/testthat` folder, and ensure that these tests pass on [Travis-CI](https://travis-ci.org/github/cjvanlissa/worcs).
 

### 2020-05-26 paper/Manuscript.Rmd
add grant info
another round of edits and suggestions

@@
 
+<!-- I would suggest to put OSCU as last item in the list of affiliations, as it is complementary to the primary institutional affiliations --> 
 affiliation:
@@
   substantial conceptual education and training in the use of new tools. The 
-  goal of this paper is to remove these obstacles. We introduce the Workflow for
+  goal of this paper is to lower the threshold to adoption. We introduce the Workflow for
   Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
-  procedure that researchers can follow to make an entire project open and
+  procedure that researchers can follow to make an entire <!-- Here I would include a word for describing the kind of project. "data science" maybe? --> project open and
   reproducible. It is based on best practices, and can be used either in
@@
 authornote: |
-  This is a preprint, not a peer-reviewed publication. Comments, contributions,
-  and coauthors are welcome. Please contact the lead author.
+  This is a preprint, not a peer-reviewed publication. Comments and
+  contributions are welcome. Please contact the lead author.
+  The lead author is supported by a NWO Veni grant (NWO grant number
+  VI.Veni.191G.090).
   We acknowledge Jeroen Ooms for offering feedback and suggesting gert.
-keywords          : "open science; reproducibility; r; dynamic document generation; version control"
+keywords          : "open science; reproducibility; r; dynamic document generation; version control, FAIR"
 wordcount         : "8257"
@@
 
-The (social) sciences are arguably past the tipping point of a paradigm shift
+# Introduction
+
+The (social) sciences <!-- Is the "(social)" necessary here?
+Does it add anything?
+How about just "science", or alternatively "research" or "academia"?
+--> are arguably past the tipping point of a paradigm shift
 towards open science.
@@
 However, open science should not be seen as a cure (or punishment) for this
-crisis. As the late dr. Jonathan Tennant put it: "Open science is just good science"
+crisis. As the late dr. <!-- Is it common to include the title? Maybe rather a description of who he was, his function? --> Jonathan Tennant put it: "Open science is just good science"
 [@tennantOpenScienceJust2018].
@@
 promotes inclusivity, because it removes barriers for participation.
-Capitalizing on these advances has the potential to accellerate scientific
+Capitalizing on these advances has the potential to accelerate scientific
 progress [see also @coyneReplicationInitiativesWill2016].
@@
 that meets most requirements open science, and is based on best practices:
-The Worfkflow for Open Reproducible 
+The Workflow for Open Reproducible 
 Code in Science (WORCS). WORCS is a lightweight approach to open science and
-computational reproducibility. The workflow (Figure <!--\@ref(fig:workflow)-->) 
+computational reproducibility <!-- here another good point to say something about the scope -->. The workflow (Figure <!--\@ref(fig:workflow)-->) 
 constitutes a step-by-step procedure that researchers can follow to make a 
@@
 
+WORCS relies on several free, open source software solutions.
+<!-- maybe give a short outline of this section here -->
+
 ### Dynamic document generation
 
-WORCS relies on several free, open source software solutions.
 The first is *dynamic document generation* (DDG): Writing scientific reports in
@@
 blocks are
-evaluated and their results are insterted in the text, or rendered as Figures and
-Tables. Dynamic document generation supercedes
+evaluated and their results are insterted in the text, or rendered as figures and
+tables. Dynamic document generation supersedes
 the classical approach of using separate programs to write prose and conduct
@@
 program. Moreover, human error in manually copying results is eliminated.
-When revisions require major changes to the analyses, all results, Figures and
-Tables can be automatically updated. The flexibility in output formats also 
+When revisions require major changes to the analyses, all results, figures and
+tables can be automatically updated. The flexibility in output formats also 
 means that a manuscript can be rendered to presentation format, or even to a
-website or blog post. Moreover, reproducibility is guaranteed, and can be
+website or blog post.
+Moreover, reproducibility is guaranteed <!-- that is a strong claim, and probably not true.
+For example, if the code calls a remote service, that might change its interface or discontinue operation, and the code will not be reproducible anymore - so, suggestion is to weeken this claim a bit...
+-->, and can be
 verified by reviewers and/or readers, because the code is run each time the
@@
 academics than most word processing software. It prevents mistakes, and saves
-time.
+time. <!-- https://www.scientificpaperofthefuture.org/ could be referenced here -->
 
@@
 being unable to reproduce analyses because they ran some analyses interactively
-instead of using syntax [see also @blischakQuickIntroductionVersion2016].
+instead of using syntax <!-- Also interactive analyses often use syntax, so this is not a good distinction.
+"...
+instead of storing commands in a script file" or something like that would be more precise --> [see also @blischakQuickIntroductionVersion2016].
 Many scientists use some form of *implicit* version control; for example, by
@@
 their own data, open "Issues" to ask questions or give feedback on the project,
-or even send a "Pull request" (a request to incorporate some code)
-with suggested changes to the text or code for
+or even send a "Pull request" with suggested changes to the text or code for
 your consideration. 'Git' and 'GitHub' shine as tools for collaboration, because
@@
 sufficient to state that analyses were conducted in *Program X*. However, every
-program is susceptible to changes, updates, and bugfixes. Open Source software,
+program is susceptible to changes, updates, and bugfixes. A lot of Open Source software,
 in particular, is regularly updated because there is an
@@
 ```
+<!-- Some questions on the workflow, for simplicity I put them here: 
+Step 6: "a script documenting the steps" - is that an executable script, or just documentation, or a script with comments? 
+Step 7: What about external data sources, actuall?
+For example, when I use the daily corona stats from Johns Hopkins university directly from their online source.
+Should they be copied or just a link provided?
+For FAIR (or at least F) data (with a PID) a link would suffice, but maybe for other sources a copy is advisable.
+Step 14: Is that optional, too? And could it not be done earlier, already in Phase 1?-->
 
@@
 of the implementation of tools required for an open science workflow.
+<!-- I would omit the second argument.
+It is highly debatable, and as it stands not more than a claim.
+Without having looked at the details, I would for example also see Python absolutely capable of doing this.
+The other four arguments are completely sufficient for motivating the focus on R, so IMHO no reason to open this discussion.
+-->
 Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
@@
 [DORA](https://sfdora.org/).
+<!-- Maybe due to restructuring, but so far DORA has not been mentioned in the paper.
+Easiest way to solve this would be to remove the subordinate clause at the end, and leave the discussion of DORA to the respective section later.
+-->
 
@@
 ['Citation File Format'](https://citation-file-format.github.io/).
+<!-- just today the Software Heritage Project announced https://www.softwareheritage.org/2020/05/26/citing-software-with-style/ - might also be worth mentioning here? -->
 
@@
 The WORCS procedure suggests uploading the online version to a preprint server.
-This is important because most preprint servers - at least those hosted by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
+This is important because most major preprint servers - at least those hosted <!-- are these "most"?
+not sure this remark is very accurate --> by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
 This means that authors will receive credit for cited work; even if they are cited only in the online version.
@@
 Markdown uses the at-symbol (`@`) to cite a reference. `worcs` additionally 
-reserves the "double at"-symbol (`@@`) to cite a non-essential references.
+reserves the "double at"-symbol (`@@`) to cite a non-essential reference.
 Users can render the manuscript either with, or without, comprehensive citations
@@
 As soon as raw data are collected, researchers should use this file to document
-all teps necessary to load the 
+all steps necessary to load the 
 data into R, pseudonimize it, and prepare it for analysis. 
@@
 `open_data()` or `closed_data()`. 
-The analysis code, on the other hand, should begin with `load_data()`.
+The analysis code, on the other hand, should begin with `load_data()`. <!-- Is this actually limited to one data file, or can it handle multiple ones? --> 
 This function loads the real data if it is available on the user's computer, and
@@
 steps required to convert these data into tabular format. If it is data must be
-stored in a different file type anyway, readers are encouraged to follow the
+stored in a different file type anyway <!-- something is wrong with the (beginning of) this sentence -->, readers are encouraged to follow the
 development of the `repro` package [@repro], which will enable such enhancements.
@@
 However, 'Git' cloud hosting services also conform to these standards.
+<!-- I am not familiar with these standard, but AFAIK the hosting services make no guarantees regarding long-term availablity of the repositories, in particular not about stable URLs pointing to repositories.
+They will not disappear suddenly or in the next few years, probably, but if it is important for preregistrations to be available "forever", then they should be made at a service that provides a permanent storage and identifier.
+-->
 Thus, it is possible to preregister a study simply by pushing a text document
@@
 in a text-based file. Optionally, an analysis script can be preregistered to
-documenting the planned analyses.
+document the planned analyses.
 If any changes must be made to the analysis code after obtaining the data,
@@
 If the remote repository is Private, Reviewers
-can be invited as "Collaborators". Private remote repositories can be made public at a
-later date, along with their entire time-stamped history and tagged releases.
+can be invited as "Collaborators". 
 It is important to note that contributing to the repository in any way will void
 reviewers' anonymity, as their contributions will be linked to a user name.
+Private remote repositories can be made public at a
+later date, along with their entire time-stamped history and tagged releases.
 
@@
 
-Although WORCS was designed to meet the TOP-guidelines, it does so in a way that
+WORCS was designed to meet the TOP-guidelines, and it does so in a way that
 is compatible with other open science principles. Most notably, WORCS is 
 amenable to the
-FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a]. These principles advocate that research
-data should be Findable, Accessible, Interoperable and Reusable. They are
-increasingly applied as a standard for other types of research output as well
-[@lamprechtFAIRPrinciplesResearch2019].
-WORCS helps users meet the FAIR principles by making all research objects -
+FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a].
+These principles advocate that digital research objects should be Findable, Accessible, Interoperable and Reusable.
+Initially mainly promoted as principles for data, they are increasingly applied as a standard for other types of research output as well, most notably software [@lamprechtFAIRPrinciplesResearch2019].
+WORCS helps users meet the FAIR principles by making all created research objects -
 including data and code - accessible in a repository that can be made public. 
@@
 indexing data by topic area. 
-Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is 
+Incidentally, `worcs` itself follows the recommendations for FAIR research software [see https://fair-software.eu/]. It is 
 hosted on a version controlled public remote repository 'GitHub', has an open
@@
 enables the citation of the software (using the `citation("worcs")` command, or
-by citing this paper), and followed the CII Best Practices software quality
+by citing this paper), and followed the CII Best Practices <!-- reference? --> software quality
 checklist during development.
@@
 authors' disclosures about, and incentivize adherence to, open science practices.
-WORCS has its own checklist,
+WORCS has its own checklist, <!-- incomplete sentence! -->
 
@@
 reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
-excellent workflow strives to ensure strict computational reproducibility for
+excellent <!-- I am not opposing this statement, but given that the two authors of this work are now co-authors on this very paper, it is a bit awkward to use this qualification...
+;) --> workflow strives to ensure strict computational reproducibility for
 even the most sophisticated analyses. It has some limitations, however: In
-focusing on reproducibility, however, it only addresses TOP-guidelines 2, 3, and
+focusing on reproducibility, it only addresses TOP-guidelines 2, 3, and
 5. Moreover, in aiming for strict computational reproducibility it ends up being
 relatively complex to implement. WORCS builds upon the same general principles
-as Peikert and Brandmaier, which means the two workflows are compatible. What 
+as Peikert and Brandmaier <!-- just noting that their names are frequently repeated.
+has their approach/tool/...
+a name that could be used instead?
+-->, which means the two workflows are compatible.
+What
 sets WORCS apart is that is an easy and sufficient solution for projects that
@@
 [opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues)
-for tasks that still need to be accomplished, and tag known Collaborators to
+for tasks that still need to be accomplished, and tag known collaborators to
 address them, or invite external collaborators to contribute their expertise.
@@
 
-The WORCS procedure provides a user-friendly and
+WORCS provides a user-friendly and
 lightweight workflow for open, reproducible research, that meets all
-TOP-guidelines. Nevertheless, there are clear directions for future developments.
+TOP-guidelines <!-- was it not with one exception? then maybe just replace the "all" in this sentence by "the" -->. Nevertheless, there are clear directions for future developments.
 Firstly, although the workflow is currently implemented only in R, it is 
@@
 environment generated using `repro`. 
-This will combine the strengths of the WORCS with those of the workflow proposed
+This will combine the strengths of WORCS with those of the workflow proposed
 by @peikertReproducibleDataAnalysis2019.
@@
 It helps researchers make all research output created throughout the scientific process - not just manuscripts, but data, code, and methods - open and publicly assessible.
-This enables other researchers to reproduce results, facilitates cumulative science by allowing others to make direct use of these research objects.
+This enables other researchers to reproduce results, and facilitates cumulative science by allowing others to make direct use of these research objects.
+<!-- Somehow this feels like very sudden and incomplete ending to me. I am sure you can add some more glorious last words. :) -->
 

### 2020-05-27 paper/Manuscript.Rmd
edits and comments on the final ms by Barbara

--- 2020-05-25 paper/Manuscript.Rmd
@@
 authornote: |
-  This is a preprint, not a peer-reviewed publication. Comments, contributions,
-  and coauthors are welcome. Please contact the lead author.
-  We acknowledge Jeroen Ooms for offering feedback and suggesting gert.
+  This is a preprint, not a peer-reviewed publication. Comments and
+  contributions are welcome. Please contact the lead author.
+  The lead author is supported by a NWO Veni grant (NWO grant number
+  VI.Veni.191G.090).
+  We acknowledge Jeroen Ooms for offering feedback and suggesting the use of the R package gert.
 keywords          : "open science; reproducibility; r; dynamic document generation; version control"
@@
 [@tennantOpenScienceJust2018].
+<!-- Perhaps change the citation and link to https://www.youtube.com/watch?v=UEEcwRUgQu8 instead.-->
 Open science creates opportunities for researchers to more easily
@@
 The functionality of 'Git' is amplified by services such as 'GitHub'
+<!-- After the first introduction, I would refrain from using quotes around Git, GitHub, etc.
+It makes sense when terms are first introduced, but after that they can be considered understood.-->
 (https://github.com). 'GitHub' is best understood as a cloud storage service
@@
 comes into play when a repository on 'GitHub' is set to "public": This allows 
-other researchers to peruse the repository and see how the work was done, clone
+other researchers to peruse the repository and see how the work was done; clone
 it to their own computer to replicate the original work or apply the methods to
-their own data, open "Issues" to ask questions or give feedback on the project,
+their own data; open "Issues" to ask questions or give feedback on the project,
 or even send a "Pull request" (a request to incorporate some code)
@@
 
-The R-implementation of WORCS uses the package `gert`, a 'Git' client for 'R'
+The R-implementation of WORCS uses the package `gert`, a 'Git' client for R
 [@oomsGertSimpleGit2019], to connect a local project to a remote repository.
@@
 must be decoded first, and many require commercial software to do so.
-Text-based files are also required when using 'Git' for version control.
-In 'Git', each change reflects a line of text added or removed. This means that, when
-writing a version-controlled paper in plain-text, it is recommended to insert a
-line break after every sentence (instead of after every paragraph). That way,
-the change log will indicate which specific sentence was edited.
-Binary files should not be version controlled with 'Git',
+It is not recommended to use Git for version control of binary files,
 as changes in the encrypted file are uninterpretable. Binary files are also often 
@@
 (Markdown) files are displayed as web pages, and `.csv` files as spreadsheets.
+
+The change log, or "diff" view in 'Git' is text-based, and reflects changes on a line-by-line basis. Thus, when writing a version-controlled paper in plain-text, it is recommended to start each 
+sentence on a new line.
+That way, the diff will indicate which specific sentence was edited.
+As Markdown does not parse a line break as the end of a paragraph, this practice does not disrupt the flow of a paragraph in the rendered version.
 
@@
 
-The R-package `worcs` automate most steps of the workflow
+The R-package `worcs` automates most steps of the workflow
 for R users.
@@
 First, R and all of it extensions are free open source software, which
-should be the tool of choice for open science.
+make it a tool of choice for open science.
 Second, as of this writing, R is the most advanced programming language in terms
@@
 Fortran, and Python [@allaireMarkdownPythonEngine2020]. There are excellent free
-resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].
+resources for learning R, such as the book 'R for Data Science' [@grolemundDataScience2017].
 
 Working with R is simplified immensely by using the RStudio integrated
-development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
+development environment (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
 RStudio automates and streamlines tedious or complicated aspects of working with
@@
 of open-source software, supports the open science efforts of others, and helps
-researchers receive credit for *all* of their research output, in line with
+researchers receive credit for *all* of their research output, in line with the San Francisco
+Declaration on Research Assessment, 
 [DORA](https://sfdora.org/).
@@
 When data cannot be shared, the risk of fraud or abuse can be mitigated by publishing a checksum for the original data, in addition to the complete analysis code.
-Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditer that running the public analysis code on the private data results in the published results of their work.
+Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditor that running the public analysis code on the private data results in the published results of their work.
 
@@
 version control some indelible record of the raw data.
+<!-- Tidy data is more than just data in a spreadsheet, but spreadsheet (or tabular) data is really all that is required by worcs to function, so you could opt to remove the reference to tidy data here and simply go with tabular/spreadsheet.
+That said, tidy data is a solid basis for data science, see also that reference, so if you want to keep this comment and reference, I would recommend to add something to the description of 'tidy data', e.g.
+"i.e.
+clean, tabular data that is formatted in a way that facilitates further analysis, see @wickhamTidyData2014"-->
 
@@
 In `worcs`, researchers can call the function `open_data(data)` to make the data
-publicly available. This function stores the `data` object (a `data.frame` or
-`matrix`) in a spreadsheet file (called `data.csv` by default), and generates an
+publicly available. This function stores the `data` object (e.g. a `data.frame` or
+`matrix` <!-- added e.g. because a tibble works here as well-->) in a spreadsheet file (called `data.csv` by default), and generates an
 accompanying codebook that can be viewed online.
-All changed files should now be (manually) added to the 'Git' repository, committed, and
+<!-- I personally don't know what a codebook is, maybe a reference or further explanation is in order? -->
+All changed files should now be (manually) <!-- the word 'manually' throws me.
+I think you mean that the git workflow has to be enacted on these files?
+I would say it is sufficient to state that this needs to be done; "manually" to me implies some point-and-drag action.
+Maybe a good time to refer (briefly) to the `git_update()` function?
+--> added to the 'Git' repository, committed, and
 pushed to the remote repository. Once the remote repository is made public, these data will be
 open to the public.
-Assume that, once you do this, the data cannot be un-shared.
+Assume that, once this is done, the data cannot be un-shared.
 
@@
 As the purpose of the `prepare_data.R` file is to prepare data for sharing in a
-format as close to "raw data" as possible, the final line of this file
+format as close to "raw data" as possible, <!-- This does not follow logically imho.
+If the purpose of the file is to share (close to) raw data, then the file *prior* to prepare_data.R should be shared; which is not the case.
+In fact, the script invites preprocessing (tidying) of the data before it is shared.
+The fact that this script ends with `open_data()` or `closed_data()` has to do with sharing, but not with the format.
+--> final line of this file
 should typically be either
 `open_data()` or `closed_data()`.
-The analysis code, on the other hand, should begin with `load_data()`.
-This function loads the real data if it is available on the user's computer, and
+
+After generating tidy, shareable data, this data is then loaded in the analysis code with the function `load_data()`.
+This function only loads the real data if it is available on the user's computer, and
 otherwise, loads the synthetic data. This will have the effect that third party
@@
 automatically load the real data. This function makes it possible for reviewers,
-coauthors, and auditors, to write analysis scripts without requiring access to
-the original data. They can simply start the script with `load_data()`, and
-write their code based on the synthetic data. Then, they can submit their code
-to you - by email or as a
-["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request).
-When you run their code on your system, `load_data()` will detect the
-original data, and use that to run their script. You can then return the results
-to the third party.
+coauthors, and others, to write analysis scripts without requiring access to
+the original data.
+A script that loads data with `load_data()`, based on the synthetic data, can be submitted to the original authors (for example through a 
+["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request)), and run in the presence of the original data.
+In this environment, `load_data()` will instead load the
+original data, and the results of this run can be returned to the third party.
 
@@
 As explained above, this is recommended, because `.csv` files are human- and
-machine readable, and because data typically need to be in a spreadsheet format
-for analysis. Many types of data can be represented as a spreadsheet - text
-corpora can have one row per document, and EEG- or EKG waveforms can have one
+machine readable, and because data typically need to be in a tabular format
+for analysis. Many types of data can be represented in tabular format - text
+<!-- I think I prefer the word 'tabular data' since spreadsheet implies functionality of the table's cells: https://en.wikipedia.org/wiki/Spreadsheet-->
+corpora can have one row per document, and EEG or EKG waveforms can have one
 row per measurement occasion. The `prepare_data.R` file should document any
-steps required to convert these data into tabular format. If it is data must be
-stored in a different file type anyway, readers are encouraged to follow the
+steps required to convert these data into tabular format.
+If this is not an appropriate format for the data, readers are encouraged to follow the
 development of the `repro` package [@repro], which will enable such enhancements.
@@
 version controlled in this way, it is advisable to use plain text only.
+<!-- Not entirely true. GitHub can certainly store documents in binary format, or images, or whatnot -- it's just that version control on them will be useless. -->
 
@@
 [AsPredicted.org](https://aspredicted.org/), and [OSF.io](https://osf.io/).
-However, 'Git' cloud hosting services also conform to these standards.
-Thus, it is possible to preregister a study simply by pushing a text document
-with the study plans to a 'Git' remote repository.
-By tagging the commit as a preregistration
-(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
-it is distinct from all other commits, and easily findable by people and programs.
 
@@
 [but see @westonRecommendationsIncreasingTransparency2019 for an excellent
-discussion of the topic]. Preregistration is the only way to ensure that a
-researcher is not "HARKing": Hypothesizing after the results are known
-[@kerrHARKingHypothesizingResults1998].
+discussion of the topic].
+
 When analyzing existing data, it is difficult to *prove* that a researcher did
@@
 However, the question of proof is only relevant from a perspective of preventing
-scientific misconduct.
+scientific misconduct. Preregistration is in a good way to show that a
+researcher is not "HARKing": Hypothesizing after the results are known
+[@kerrHARKingHypothesizingResults1998].
 Good faith preregistration efforts always improve the quality of deductive
@@
 [@gelmanStatisticalCrisisScience2014].
-In all cases, researchers should minimize exposure to the data, and disclose any
-exposure.
+<!-- remove because it follows in the next paragraph. In all cases, researchers should minimize exposure to the data, and disclose any
+exposure.-->
 
@@
 (e.g., by reading papers using the same data); deviations from the analysis plan
-to handle unforseen contingencies, such as violations of model assumptions;
+to handle unforeseen contingencies, such as violations of model assumptions;
 or additional exploratory analyses.
@@
 
-To preregister a project using 'Git', one can commit
-all preregistration documents to the local 'Git' repository, 
-and push them 
-to a remote 'Git' hosting service.
-By [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration" on the remote repository, it can be easily identified as
-such and retrieved. This simple but robust approach renders pre-registration through
+The `worcs` workflow facilitates preregistration, primarily by explicitly inviting a user to write the preregistration document (a `preregistration.Rmd` file is part of the generated toolkit of a `worcs` project).
+Within the workflow, the 'Git' cloud hosting services can be used to preregister a study simply by pushing a text document with the study plans to a 'Git' remote repository.
+By [tagging the commit](https://help.github.com/en/github/administering-a-repository/creating-releases) as a preregistration
+(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
+it is distinct from all other commits, and easily findable by people and machines.
+
+This simple but robust approach renders pre-registration through
 preregistration services, such as AsPredicted.org or OSF.io, optional.
@@
 ideally prior to data collection.
+<!-- Maybe move this section up, or move the part where I referred to this template down?-->
 
@@
 data should be Findable, Accessible, Interoperable and Reusable. They are
-increasingly applied as a standard for other types of research output as well
+increasingly applied as a standard for other types of digital research output as well
 [@lamprechtFAIRPrinciplesResearch2019].
-WORCS helps users meet the FAIR principles by making all research objects -
-including data and code - accessible in a repository that can be made public. 
+<!-- I'm not certain about the language here: FAIR is not a goal but rather a process; a spectrum...
+words like 'standard' and 'meet the principles' suggest a checklist of criteria, which is not really what FAIR is about...
+BUT 1.
+I tried for a while to find a better way of phrasing this and couldn't, and 2.
+if Anna-Lena signed off on this then who am I to judge :D -->
+WORCS helps users meet the FAIR principles by making all digital research objects -
+including data and code - accessible through a repository that can be made public.
+<!-- accessible != open. Data does not need to be present in the repository to be accessible through it.-->
 The `worcs` package additionally
@@
 indexing data by topic area. 
-Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is 
-hosted on a version controlled public remote repository 'GitHub', has an open
+Incidentally, `worcs` is itself an example research software following FAIR guidelines [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is 
+hosted on the version controlled public remote repository 'GitHub', has an open
 source licence (GPL v3.0), is registered in a community registry (CRAN),
@@
 by citing this paper), and followed the CII Best Practices software quality
-checklist during development.
+checklist during development(link?).
 
@@
 surrogate indicators like journal impact factors, conference rankings, and
-h-indexes. The detailed version control and commit tracking of 'Git' remote repositories
+h-indexes.
+<!-- Evaluators were always free to read  papers and not rely on these rankings.
+Imho the main benefit of open science and DORA is not that an evaluator can scour over your github — ain't nobody got time for that — but that all digital objects related to the research can be used and cited, attaining a broader impact profile than would be seen with papers alone.-->
+The detailed version control and commit-tracking of 'GitHub' remote repositories
 furthermore make it possible to assess the relative
@@
 authors' disclosures about, and incentivize adherence to, open science practices.
-WORCS has its own checklist,
+WORCS, too, has its own checklist that calls attention to a number of concrete items contributing to an open and reproducible research project.
 
@@
 limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
-When soliting feedback from co-authors, ask them to use Tracked Changes and
+When soliciting feedback from co-authors, ask them to use Track Changes and
 comment bubbles in Word. <!-- LB: Overleaf?-->Then, manually incorporate their

### 2020-05-28 paper/Manuscript.Rmd

--- 2020-05-25 paper/Manuscript.Rmd
@@
     affiliation: "2,9"
+    email: "m.struiksma@uu.nl"
 
@@
   - id: "9"
-    institution: "Utrecht Institute of Linguistics OTS"
+    institution: "Utrecht University, Department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS"
 
@@
 authornote: |
-  This is a preprint, not a peer-reviewed publication. Comments, contributions,
-  and coauthors are welcome. Please contact the lead author.
+  This is a preprint, not a peer-reviewed publication. Comments and
+  contributions are welcome. Please contact the lead author.
+  The lead author is supported by a NWO Veni grant (NWO grant number
+  VI.Veni.191G.090).
   We acknowledge Jeroen Ooms for offering feedback and suggesting gert.

### 2020-05-28 paper/Manuscript.Rmd
Integrate AL comments

@@
 
-<!-- I would suggest to put OSCU as last item in the list of affiliations, as it is complementary to the primary institutional affiliations --> 
 affiliation:
@@
   Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
-  procedure that researchers can follow to make an entire <!-- Here I would include a word for describing the kind of project. "data science" maybe? --> project open and
+  procedure that researchers can follow to make a research project open and
   reproducible. It is based on best practices, and can be used either in
   parallel to, or in absence of, top-down requirements by journals,
-  institutions, and funding bodies. To facilitate widespread adoption, the
+  institutions, and funding bodies. This workflow is primarily targeted towards
+  scholars working in R on a private computer,
+  conducting research that involves academic prose, analysis code,
+  and (optionally) tabular data - but it is flexible enough to accommodate many
+  other scenarios.
+  To facilitate widespread adoption, the
   WORCS principles are implemented in the R-package `worcs`, which offers an
@@
 
-# Introduction
-
-The (social) sciences <!-- Is the "(social)" necessary here?
-Does it add anything?
-How about just "science", or alternatively "research" or "academia"?
---> are arguably past the tipping point of a paradigm shift
+Academia is arguably past the tipping point of a paradigm shift
 towards open science.
@@
 However, open science should not be seen as a cure (or punishment) for this
-crisis. As the late dr. <!-- Is it common to include the title? Maybe rather a description of who he was, his function? --> Jonathan Tennant put it: "Open science is just good science"
+crisis. As the late dr. Jonathan Tennant put it: "Open science is just good science"
 [@tennantOpenScienceJust2018].
@@
 Code in Science (WORCS). WORCS is a lightweight approach to open science and
-computational reproducibility <!-- here another good point to say something about the scope -->. The workflow (Figure <!--\@ref(fig:workflow)-->) 
+computational reproducibility for scholars working in R on a private computer,
+conducting research that involves academic prose, analysis code,
+and (optionally) tabular data - but it is flexible enough to accommodate many
+other scenarios. The workflow (Figure \@ref(fig:workflow)) 
 constitutes a step-by-step procedure that researchers can follow to make a 
@@
 package.
+
+```{r workflow, fig.cap="Schematic illustration of the WORCS procedure"}
+knitr::include_graphics("workflow_graph/workflow.png")
+```
 
@@
 
-WORCS relies on several free, open source software solutions.
-<!-- maybe give a short outline of this section here -->
+WORCS relies on several free, open source software solutions which we will
+discuss before introducing the workflow.
 
@@
 means that a manuscript can be rendered to presentation format, or even to a
-website or blog post.
-Moreover, reproducibility is guaranteed <!-- that is a strong claim, and probably not true.
-For example, if the code calls a remote service, that might change its interface or discontinue operation, and the code will not be reproducible anymore - so, suggestion is to weeken this claim a bit...
--->, and can be
-verified by reviewers and/or readers, because the code is run each time the
-document is compiled. In sum, while writing academic papers in a programming environment might seem
+website or blog post. Moreover, the fact that code can be run each time the
+document is compiled encourages researchers to work reproducibly,
+and allows reviewers and/or readers verify reproducibility simply by re-compiling
+the document.
+In sum, while writing academic papers in a programming environment might seem
 counter-intuitive at first, this approach is much more amenable to the needs of
 academics than most word processing software. It prevents mistakes, and saves
-time. <!-- https://www.scientificpaperofthefuture.org/ could be referenced here -->
+time.
 
@@
 learns who has had the experience of accidentally deleting a crucial file, or of
-being unable to reproduce analyses because they ran some analyses interactively
-instead of using syntax <!-- Also interactive analyses often use syntax, so this is not a good distinction.
-"...
-instead of storing commands in a script file" or something like that would be more precise --> [see also @blischakQuickIntroductionVersion2016].
+being unable to reproduce analyses because they ran some analyses interactively,
+instead of documenting the syntax in a script file [see also @blischakQuickIntroductionVersion2016].
 Many scientists use some form of *implicit* version control; for example, by
@@
 ```
-<!-- Some questions on the workflow, for simplicity I put them here: 
-Step 6: "a script documenting the steps" - is that an executable script, or just documentation, or a script with comments? 
-Step 7: What about external data sources, actuall?
-For example, when I use the daily corona stats from Johns Hopkins university directly from their online source.
-Should they be copied or just a link provided?
-For FAIR (or at least F) data (with a PID) a link would suffice, but maybe for other sources a copy is advisable.
-Step 14: Is that optional, too? And could it not be done earlier, already in Phase 1?-->
 
@@
 should be the tool of choice for open science.
-Second, as of this writing, R is the most advanced programming language in terms
-of the implementation of tools required for an open science workflow.
-<!-- I would omit the second argument.
-It is highly debatable, and as it stands not more than a claim.
-Without having looked at the details, I would for example also see Python absolutely capable of doing this.
-The other four arguments are completely sufficient for motivating the focus on R, so IMHO no reason to open this discussion.
--->
+Second, all tools required for an open science workflow are implemented in R,
+most of these tools are directly accessible through the user interface of RStudio,
+and some of them are actively developed by the team behind RStudio.
 Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
@@
 in the source code repository (e.g., on 'Git' remote repositories), using the
-['Citation File Format'](https://citation-file-format.github.io/).
-<!-- just today the Software Heritage Project announced https://www.softwareheritage.org/2020/05/26/citing-software-with-style/ - might also be worth mentioning here? -->
+['Citation File Format'](https://citation-file-format.github.io/). The
+[*Software Heritage Project*](https://www.softwareheritage.org/2020/05/26/citing-software-with-style/)
+recently published a BibLaTeX software citation style that will further facilitate
+accurate citation of software in dynamically generated documents.
 
@@
 The WORCS procedure suggests uploading the online version to a preprint server.
-This is important because most major preprint servers - at least those hosted <!-- are these "most"?
-not sure this remark is very accurate --> by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
+This is important because most major preprint servers - including [arXiv.org](https://arxiv.org)
+and all preprint services hosted by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
 This means that authors will receive credit for cited work; even if they are cited only in the online version.
@@
 `open_data()` or `closed_data()`. 
-The analysis code, on the other hand, should begin with `load_data()`. <!-- Is this actually limited to one data file, or can it handle multiple ones? --> 
+The analysis code, on the other hand, should begin with `load_data()`.
 This function loads the real data if it is available on the user's computer, and
@@
 
+Note that these functions can be used to store multiple data files, if necessary.
 As of version 0.1.2, the `worcs` package only stores data as `.csv` files.
@@
 row per measurement occasion. The `prepare_data.R` file should document any
-steps required to convert these data into tabular format. If it is data must be
-stored in a different file type anyway <!-- something is wrong with the (beginning of) this sentence -->, readers are encouraged to follow the
-development of the `repro` package [@repro], which will enable such enhancements.
+steps required to convert these data into tabular format. 
+If this is not an appropriate format, readers are encouraged to follow the
+development of the `repro` package [@repro], which will offer greater support for
+reproducibly storing and retrieving diverse data formats.
 
@@
 However, 'Git' cloud hosting services also conform to these standards.
-<!-- I am not familiar with these standard, but AFAIK the hosting services make no guarantees regarding long-term availablity of the repositories, in particular not about stable URLs pointing to repositories.
-They will not disappear suddenly or in the next few years, probably, but if it is important for preregistrations to be available "forever", then they should be made at a service that provides a permanent storage and identifier.
--->
-Thus, it is possible to preregister a study simply by pushing a text document
-with the study plans to a 'Git' remote repository.
-By tagging the commit as a preregistration
-(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
-it is distinct from all other commits, and easily findable by people and programs.
+Thus, it is possible to preregister a study simply by committing a preregistration document
+to the local 'Git' repository, and pushing it to the remote repository.
+Subsequently [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases)
+as "Preregistration" on the remote repository renders it distinct from all other commits,
+and easily findable by people and programs.
+This approach is simple, quick, and robust.
+Moreover, it is compatible with formal preregistration through services
+such as AsPredicted.org or OSF.io:
+If the preregistration is
+written using dynamic document generation, as 
+recommended in WORCS, this file can be rendered to PDF and uploaded as an attachment
+to the formal preregistration service.
 
@@
 
-To preregister a project using 'Git', one can commit
-all preregistration documents to the local 'Git' repository, 
-and push them 
-to a remote 'Git' hosting service.
-By [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration" on the remote repository, it can be easily identified as
-such and retrieved. This simple but robust approach renders pre-registration through
-preregistration services, such as AsPredicted.org or OSF.io, optional.
-However, it is easy to take this optional step when the preregistration is
-written using dynamic document generation, as 
-recommended in WORCS: This file can be formatted as PDF and uploaded as an attachment
-to the formal preregistration service.
-
 As soon as a project is preregistered on a Public 'Git' repository, it is
@@
 Initially mainly promoted as principles for data, they are increasingly applied as a standard for other types of research output as well, most notably software [@lamprechtFAIRPrinciplesResearch2019].
+
 WORCS helps users meet the FAIR principles by making all created research objects -
@@
 indexing data by topic area.
+
 Incidentally, `worcs` itself follows the recommendations for FAIR research software [see https://fair-software.eu/]. It is 
@@
 enables the citation of the software (using the `citation("worcs")` command, or
-by citing this paper), and followed the CII Best Practices <!-- reference? --> software quality
-checklist during development.
+by citing this paper), and followed the [*CII Best Practices*](https://bestpractices.coreinfrastructure.org/en)
+software quality checklist during development.
 
@@
 reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
-excellent <!-- I am not opposing this statement, but given that the two authors of this work are now co-authors on this very paper, it is a bit awkward to use this qualification...
-;) --> workflow strives to ensure strict computational reproducibility for
+workflow strives to ensure strict computational reproducibility for
 even the most sophisticated analyses. It has some limitations, however: In
@@
 relatively complex to implement. WORCS builds upon the same general principles
-as Peikert and Brandmaier <!-- just noting that their names are frequently repeated.
-has their approach/tool/...
-a name that could be used instead?
--->, which means the two workflows are compatible.
-What
+as Peikert and Brandmaier,
+which means the two workflows are compatible. What 
 sets WORCS apart is that is an easy and sufficient solution for projects that
@@
 FAIR principles, while being compatible with existing requirements.
+
+One initiative that WORCS is fully compatible with, is the [*Scientific Paper of
+the Future*](https://www.scientificpaperofthefuture.org/). This organization encourages
+geoscientists to document data provenance and availability in public repositories,
+document software used,
+and document the analysis steps taken to derive the results. All of these goals
+could be met using WORCS.
 
@@
 lightweight workflow for open, reproducible research, that meets all
-TOP-guidelines <!-- was it not with one exception? then maybe just replace the "all" in this sentence by "the" -->. Nevertheless, there are clear directions for future developments.
+TOP-guidelines, except replication.
+Nevertheless, there are clear directions for future developments.
 Firstly, although the workflow is currently implemented only in R, it is 

### 2020-05-28 paper/Manuscript.Rmd
final comments

--- 2020-05-28 paper/Manuscript.Rmd
@@
 ---
-title: |
-    | WORCS: A Workflow for Open Reproducible Code in Science
-    | `r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`
+title             :  "WORCS: A Workflow for Open Reproducible Code in Science \\newline `r paste0('Version: ', worcs:::cran_version(), ', date: ', format(Sys.Date(), '%d %b,\ %Y'))`"
 
@@
     affiliation: "2,9"
-    email: "m.struiksma@uu.nl"
 
@@
   - id: "9"
-    institution: "Utrecht University, Department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS"
+    institution: "Utrecht Institute of Linguistics OTS"
 
@@
   substantial conceptual education and training in the use of new tools. The 
-  goal of this paper is to remove these obstacles. We introduce the Workflow for
-  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
+  goal of this paper is to remove these obstacles. We introduce the _Workflow for
+  Open Reproducible Code in Science_ (WORCS). WORCS constitutes a step-by-step
   procedure that researchers can follow to make an entire project open and
-  reproducible. It is based on best practices, and can be used either in
+  reproducible. It is based on current best practices, and can be used either in
   parallel to, or in absence of, top-down requirements by journals,
   institutions, and funding bodies. To facilitate widespread adoption, the
-  WORCS principles are implemented in the R-package `worcs`, which offers an
-  RStudio 
-  project template that is, effectively, a "one-click solution" for R users. 
+  WORCS principles are implemented in the R package `worcs`, which offers an
+  RStudio project template that is, effectively, a "one-click solution" for R users. 
   This paper introduces the workflow, discusses how it meets different 
@@
   VI.Veni.191G.090).
-  We acknowledge Jeroen Ooms for offering feedback and suggesting gert.
-keywords          : "open science; reproducibility; r; dynamic document generation; version control"
+  We acknowledge Jeroen Ooms for offering feedback and suggesting the `gert` package.
+keywords          : "open science; reproducibility; R; dynamic document generation; version control"
 wordcount         : "8257"
@@
 However, open science should not be seen as a cure (or punishment) for this
-crisis. As the late dr. Jonathan Tennant put it: "Open science is just good science"
+crisis. As the late Jonathan Tennant brilliantly put it: "Open science is just good science"
 [@tennantOpenScienceJust2018].
@@
 promotes inclusivity, because it removes barriers for participation.
-Capitalizing on these advances has the potential to accellerate scientific
+Capitalizing on these advances has the potential to accelerate scientific
 progress [see also @coyneReplicationInitiativesWill2016].
 
-Many researchers are motivated to adopt best practices for open
+Many researchers are motivated to adopt current best practices for open
 science and enjoy these benefits. And yet, the question of "how" to 
@@
 This paper is designed to ease that transition by presenting a simple workflow 
-that meets most requirements open science, and is based on best practices:
-The Worfkflow for Open Reproducible 
-Code in Science (WORCS). WORCS is a lightweight approach to open science and
+that meets most current requirements of open science, and is based on recent best practices: The _Workflow for Open Reproducible Code in Science_ (WORCS).
+ WORCS is a lightweight approach to open science and
 computational reproducibility. The workflow (Figure <!--\@ref(fig:workflow)-->) 
 constitutes a step-by-step procedure that researchers can follow to make a 
-research project open and reproducible. WORCS is compatible with open science
-requirements already implemented by journals and institutions, and will help 
+research project (including data, code, and a report) open and reproducible. WORCS is compatible with open science requirements already implemented by journals and institutions, and will help to
 fulfill them. It can also be used in the absence of top-down support or guidelines
-to produce work in accordance with best practices.
+to produce work in accordance with current best practices.
 Although the principles underlying this workflow are universal, WORCS has first
 been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
-The R-package `worcs` offers a project template for RStudio
+The R package `worcs` [@worcspackage2020] offers a project template for RStudio
 [@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
@@
 Although Open Science is advocated by many, it does not have a unitary definition
-[for an effort to ground open science in first principles, see 
-@tennantValuePropositionOpen2020].
-Instead, several guidelines exist for practices that are considered "open".
-The TOP-guidelines are one of the most influential concrete operationalisations
+[for a proposition to arrive at a unified notion, see 
+@tennantValuePropositionOpen2020].<!--AB: I wasn't sure what was meant with 'first principles'; I think it's more about the unification-->
+Instead, there are several guidelines for practices that are considered "open".
+The TOP-guidelines are one of the most influential operationalisations <!--AB: operationalisations are always concrete-->
 of Open Science principles [@nosekPromotingOpenResearch2015a].
@@
 do not address replication of published results, because replication relates to
-the subject of a research project, not to its execution.
+the subject of a research project, not to its organisation or execution.
 We group the remaining seven guidelines into three categories: citation (1),
@@
 
-Although there is a slight learning curve to transitioning to DDG, the
-investment really pays off. Time saved from painstakingly copy-pasting output
+Although there is a slight learning curve to transitioning to DDG, we strongly believe that the
+investment really pays off<!--AB: It's a claim after all, so let's say "we strongly believe"-->. Time saved from painstakingly copy-pasting output
 and manually formatting text soon outweighs the investment of switching to a new
 program. Moreover, human error in manually copying results is eliminated.
-When revisions require major changes to the analyses, all results, Figures and
-Tables can be automatically updated. The flexibility in output formats also 
+When revisions require major changes to the analyses, all results, figures and
+tables can be automatically updated. The flexibility in output formats also 
 means that a manuscript can be rendered to presentation format, or even to a
@@
 stored in `.R` files, and called from the main document using the `source()`
-function. When a reader or reviewer compiles this document, all code is run
+function or using a special syntax for calling labeled chunks of code. When a reader or reviewer compiles this document, all code is run
 automatically, thus verifying computational replicability. The 'RMarkdown' 
 document can be automatically formatted in many styles, including
-APA style [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020],
+APA style [thanks to the R package `papaja`, @austPapajaPrepareReproducible2020],
 a host of other scientific formats 
@@
 
-The second solution is *version control*:
+The second tool <!--AB: when saying solution, we need a problem--> is *version control*:
 Maintaining an indelible log of every change to all project files. Version 
@@
 cloud storage service, 'GitHub', two notable alternatives exist: 
-'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). As of
+'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org) <!--AB: https://gitea.io/ which is a clone of Gogs seems also pretty cool-->. As of
 version 0.1.2, the `worcs` template dialog accepts 'https' addresses from 
@@
 
-The third solution is *dependency management*: Keeping track of exactly what
+The third tool <!--AB: Same as before--> is *dependency management*: Keeping track of exactly what
 software was used to conduct the analyses. At first glance, it might seem
@@
 conceptually and practically compatible. However, the workflow described by 
-Peikert and Brandmaier can be difficult to set up for novice users. An R-package
+Peikert and Brandmaier can be difficult to set up for novice users. The accompanying R package `repro`
 is currently in development to facilitate this transition [@repro]. Another 
@@
 Version 0.1.2. 
-For users of the `worcs` R-package, we highly recommend referring to
+For users of the `worcs` R package, we highly recommend referring to
 [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
@@
 
-The R-package `worcs` automate most steps of the workflow
+The R package `worcs` automates most steps of the workflow
 for R users.
@@
 should be the tool of choice for open science.
-Second, as of this writing, R is the most advanced programming language in terms
+Second, as of this writing, R is among the <!--AB: I think python folks will not be happy about this comment if we say R is most advanced--> most advanced programming language in terms
 of the implementation of tools required for an open science workflow.
 Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
-Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
+Fourth, R is well-supported; a vibrant and inclusive online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
 Finally, R is highly interoperable:
@@
 Another important feature of RStudio is project management.
-A project bundles writing, analyses, data, references, etcetera, into a
+A project bundles writing, analyses, data, references, et cetera, into a
 self-contained folder, that can be uploaded entirely to a 'Git' remote repository and downloaded
-by future users. The `worcs` R-package installs a new RStudio project template.
+by future users. The `worcs` R package installs a new RStudio project template.
 When a new project is initialized from this template, the bookkeeping required
@@
 researchers receive credit for *all* of their research output, in line with
-[DORA](https://sfdora.org/).
+[DORA](https://sfdora.org/). <!--AB: spell out-->
 
@@
 
-With regard to the citation of R-packages, it is worth noting that the `papaja` 
+With regard to the citation of R packages, it is worth noting that the `papaja` 
 package [@austPapajaPrepareReproducible2020], which provides the 'APA6'
@@
 of the contents of a file.
-Any change to the file will result in a different checksum. 
+Any change^[It is possible that random changes to a file result in the same checksum but chances are low] <!--AB: I would really urge us to be precise about this.
+If checksums were really unique, we would have a really awesome compression algorithm but by applying the pigeonhole principle, it is easy to show that many large files must map onto the same short checksum--> to the file will result in a different checksum.
 Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
@@
 for analysis. Many types of data can be represented as a spreadsheet - text
-corpora can have one row per document, and EEG- or EKG waveforms can have one
-row per measurement occasion. The `prepare_data.R` file should document any
+corpora can have one row per document, and EEG- or ECG waveforms can have one
+row per channel. The `prepare_data.R` file should document any
 steps required to convert these data into tabular format. If it is data must be
@@
 
-Several excellent preregistration templates are available in the R-package
+Several excellent preregistration templates are available in the R package
 `prereg` [@austPreregMarkdownTemplates2019], including templates from
@@
 The workflow aims to lower the threshold for grass-roots adoption of open science principles.
-The workflow is supported by an R-package with an RStudio project template and 
+The workflow is supported by an R package with an RStudio project template and 
 convenience functions.
@@
 reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
-excellent workflow strives to ensure strict computational reproducibility for
+<!--AB: the 'excellent' seems preposterous now that we joined this project--> workflow strives to ensure strict computational reproducibility for
 even the most sophisticated analyses. It has some limitations, however: In
-focusing on reproducibility, however, it only addresses TOP-guidelines 2, 3, and
+focusing on reproducibility, it only addresses TOP-guidelines 2, 3, and
 5. Moreover, in aiming for strict computational reproducibility it ends up being
@@
 in scientific software development as well; for example, the lead author of this
-paper became a co-author on several R-packages after submitting pull requests
+paper became a co-author on several R packages after submitting pull requests
 with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018], and two co-authors of this
@@
 workflow. To this end, we are collaborating with @repro on the development of
-`repro`, an R-package that will facilitate setting up reproducibility tools such
+`repro`, an R package that will facilitate setting up reproducibility tools such
 as 'Make' and 'Docker'.
@@
 The step-by-step procedure outlined in this tutorial helps researchers make an entire research project Open and Reproducible.
-The accompanying R-package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.
+The accompanying R package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.
 

### 2020-05-28 paper/Manuscript.Rmd
incorporate barbara changes

--- 2020-05-28 paper/Manuscript.Rmd
@@
   VI.Veni.191G.090).
-  We acknowledge Jeroen Ooms for offering feedback and suggesting gert.
+  We acknowledge Jeroen Ooms for offering feedback and suggesting the use of the
+  R package gert.
 keywords          : "open science; reproducibility; r; dynamic document generation; version control, FAIR"
@@
 comes into play when a repository on 'GitHub' is set to "public": This allows 
-other researchers to peruse the repository and see how the work was done, clone
+other researchers to peruse the repository and see how the work was done; clone
 it to their own computer to replicate the original work or apply the methods to
-their own data, open "Issues" to ask questions or give feedback on the project,
+their own data; open "Issues" to ask questions or give feedback on the project,
 or even send a "Pull request" with suggested changes to the text or code for
@@
 must be decoded first, and many require commercial software to do so.
-Text-based files are also required when using 'Git' for version control.
-In 'Git', each change reflects a line of text added or removed. This means that, when
-writing a version-controlled paper in plain-text, it is recommended to insert a
-line break after every sentence (instead of after every paragraph). That way,
-the change log will indicate which specific sentence was edited.
-Binary files should not be version controlled with 'Git',
-as changes in the encrypted file are uninterpretable. Binary files are also often 
+The version control software 'Git' is also designed for use with text-based files.
+The 'Git' change log reflects changes to text-based files on a line-by-line basis.
+Binary files can be version controlled with 'Git', but 
+the change log for a binary file are uninterpretable.
+Binary files are also often 
 larger than text-based files, which means they take up more space on cloud
@@
 text-based files for all purposes.
-It is worth noting that 'GitHub' also rewards the use of text-based files by
-rendering certain text-based filetypes: 
+
+Two additional points are worth noting:
+First, 'Git' tracks line-by-line changes to text-based files,
+based on line breaks in the text.
+When writing a version-controlled paper in plain-text,
+it is therefore recommended to start each sentence on a new line.
+That way, the change log will indicate which specific sentence was edited, 
+instead of replacing an entire paragraph.
+As RMarkdown does not parse a single line break as the end of a paragraph,
+this practice does not disrupt the flow of a paragraph in the rendered version.
+Second, the remote repository service 'GitHub' rewards the use of text-based files by
+rendering certain text-based filetypes for online viewing: 
 For example, `.md`
 (Markdown) files are displayed as web pages, and `.csv` files as spreadsheets.
+The `worcs` implementation for R makes use of this fact by, for example,
+rendering a codebook for the data as `.md` so visitors of a 'Git' remote repository
+can view it as a web page.
 
@@
 First, R and all of it extensions are free open source software, which
-should be the tool of choice for open science.
+make it a tool of choice for open science.
 Second, all tools required for an open science workflow are implemented in R,
@@
 Fortran, and Python [@allaireMarkdownPythonEngine2020]. There are excellent free
-resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].
+resources for learning R, such as the book 'R for Data Science' [@grolemundDataScience2017].
 
@@
 of open-source software, supports the open science efforts of others, and helps
-researchers receive credit for *all* of their research output, in line with
+researchers receive credit for *all* of their research output, in line with the San Francisco
+Declaration on Research Assessment, 
 [DORA](https://sfdora.org/).
-<!-- Maybe due to restructuring, but so far DORA has not been mentioned in the paper.
-Easiest way to solve this would be to remove the subordinate clause at the end, and leave the discussion of DORA to the respective section later.
--->
 
@@
 
+When initializing a new `worcs` project, an empty R-script called `prepare_data.R` is generated.
+As soon as raw data are collected, researchers should use this file to document
+all steps necessary to load the 
+data into R, pseudonimize it, and prepare it for analysis.
+As data should be shared in a format as close to raw as possible, 
+this script should
+be as short as possible and only document the minimum necessary steps.
+For example, if the data was originally in SPSS format with IP addresses and GPS
+location, this file might just contain the code required to read the SPSS file,
+and to remove those columns of sensitive personal information. As soon as the
+data are pseudonimized and processed into tabular (spreadsheet-like) format,
+the researcher should
+version control some indelible record of the raw data.
+The concept of 'tidy' data
+is a helpful standard; i.e., clean, tabular data that is formatted in a way that facilitates further analysis [see @wickhamTidyData2014].
+
+One issue of concern is the practice of "fixing" mistakes in the data.
+It is not uncommon for researchers to perform such corrections manually in the raw
+data, 
+even if all other analysis steps are documented.
 Regardless of whether the data will be open or closed, it is important that the
@@
 
-When initializing a new `worcs` project, an empty R-script called `prepare_data.R` is generated.
-As soon as raw data are collected, researchers should use this file to document
-all steps necessary to load the 
-data into R, pseudonimize it, and prepare it for analysis. 
-For example, if the data was originally in SPSS format with IP addresses and GPS
-location, this file might just contain the code required to read the SPSS file,
-and to remove those columns of sensitive personal information. As soon as the
-data are pseudonimized and processed into a 'tidy' format [i.e., data that can
-be stored as a spreadsheet, see @wickhamTidyData2014], the researcher should
-version control some indelible record of the raw data.
-
 The `worcs` package offers two functions for version controlling a record of the
@@
 In `worcs`, researchers can call the function `open_data(data)` to make the data
-publicly available. This function stores the `data` object (a `data.frame` or
-`matrix`) in a spreadsheet file (called `data.csv` by default), and generates an
-accompanying codebook that can be viewed online.
-All changed files should now be (manually) added to the 'Git' repository, committed, and
-pushed to the remote repository. Once the remote repository is made public, these data will be
+publicly available. This function stores the `data` object (such as a `data.frame` or
+`matrix`) in a tabular data file (called `data.csv` by default).
+It also generates an
+accompanying codebook - 
+a human-readable document that serves as a legend for the tabular data file -
+that can be viewed online.
+All changed files should now be added to the 'Git' repository, committed, and
+pushed to the remote repository.
+This can be done through the 'Git' panel in RStudio, or by running the `worcs`
+function `git_update("commit message")`.
+Once the remote repository is made public, these data will be
 open to the public.
-Assume that, once you do this, the data cannot be un-shared.
+Assume that, once this is done, the data cannot be un-shared.
 
@@
 
-As the purpose of the `prepare_data.R` file is to prepare data for sharing in a
-format as close to "raw data" as possible, the final line of this file
+As the purpose of the `prepare_data.R` file is to prepare data for sharing,
+the final line of this file
 should typically be either
 `open_data()` or `closed_data()`.
-The analysis code, on the other hand, should begin with `load_data()`.
-This function loads the real data if it is available on the user's computer, and
-otherwise, loads the synthetic data. This will have the effect that third party
+
+After generating tidy, shareable data, this data is then loaded in the analysis code with the function `load_data()`.
+This function only loads the real data if it is available on the user's computer, 
+and otherwise, loads the synthetic data.
+This will have the effect that third party
 users who copy a remote repository with closed data will automatically load the synthetic
 dataset, whereas the study authors, who have the real data stored locally, will
-automatically load the real data. This function makes it possible for reviewers,
+automatically load the real data.
+This makes it possible for reviewers,
 coauthors, and auditors, to write analysis scripts without requiring access to
-the original data. They can simply start the script with `load_data()`, and
-write their code based on the synthetic data. Then, they can submit their code
+the original data.
+They can simply start the script with `load_data()`, and
+write their code based on the synthetic data.
+Then, they can submit their code
 to you - by email or as a
 ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request).
-When you run their code on your system, `load_data()` will detect the
+When you run their code on your system, `load_data()` will load the
 original data, and use that to run their script. You can then return the results
@@
 As explained above, this is recommended, because `.csv` files are human- and
-machine readable, and because data typically need to be in a spreadsheet format
-for analysis. Many types of data can be represented as a spreadsheet - text
+machine readable, and because data typically need to be in tabular format
+for analysis. Many types of data can be represented as a table - text
 corpora can have one row per document, and EEG- or EKG waveforms can have one
@@
 steps required to convert these data into tabular format. 
-If this is not an appropriate format, readers are encouraged to follow the
+If this is not an appropriate format for the data, readers are encouraged to follow the
 development of the `repro` package [@repro], which will offer greater support for
@@
 committing them, and pushing to a cloud hosting service. As with any document
-version controlled in this way, it is advisable to use plain text only.
+version controlled in this way, it is advisable (but not required) to use plain text only.
 
@@
 scientific misconduct.
+Preregistration is a good way to ensure that a
+researcher is not "HARKing": Hypothesizing after the results are known
+[@kerrHARKingHypothesizingResults1998].
+
 Good faith preregistration efforts always improve the quality of deductive
@@
 [@gelmanStatisticalCrisisScience2014].
-In all cases, researchers should minimize exposure to the data, and disclose any
-exposure.
 
@@
 plans, and adhere to them.
-All deviations from this procedure should be disclosed; e.g., prior exposure to
-the data, whether direct (e.g., by computing summary statistics) or indirect
-(e.g., by reading papers using the same data); deviations from the analysis plan
-to handle unforseen contingencies, such as violations of model assumptions;
+All deviations from this procedure should be disclosed.
+Researchers should minimize exposure to the data, and disclose any
+prior exposure, whether direct (e.g., by computing summary statistics) or indirect
+(e.g., by reading papers using the same data).
+Similarly, one can disclose any deviations from the analysis plan
+to handle unforeseen contingencies, such as violations of model assumptions;
 or additional exploratory analyses.
@@
 
-Several excellent preregistration templates are available in the R-package
+The `worcs` workflow facilitates preregistration,
+primarily by explicitly inviting a user to write the preregistration document.
+To this end, `worcs` imports several preregistration templates from the R-package
 `prereg` [@austPreregMarkdownTemplates2019], including templates from
@@
 ideally prior to data collection. 
+Within the workflow, the 'Git' cloud hosting services can be used to preregister a study simply by pushing a text document with the study plans to a 'Git' remote repository.
+By [tagging the commit](https://help.github.com/en/github/administering-a-repository/creating-releases) as a preregistration
+(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
+it is distinct from all other commits, and easily findable by people and machines.
 
@@
 These principles advocate that digital research objects should be Findable, Accessible, Interoperable and Reusable.
-Initially mainly promoted as principles for data, they are increasingly applied as a standard for other types of research output as well, most notably software [@lamprechtFAIRPrinciplesResearch2019].
+Initially mainly promoted as principles for data, they are increasingly applied as a standard for other types of research output as well,
+most notably software [@lamprechtFAIRPrinciplesResearch2019].
 
@@
 authors' disclosures about, and incentivize adherence to, open science practices.
-WORCS has its own checklist, <!-- incomplete sentence! -->
+WORCS, too, has its own checklist that calls attention to a number of concrete
+items contributing to an open and reproducible research project. Users of the
+`worcs` package can receive a badge, displayed on the `README.md` file,
+based on a semi-automated scoring of the checklist items, by calling the function
+`check_worcs()` within a `worcs` project directory.
 

### 2020-05-29 paper/Manuscript.Rmd
Update description and author note
Update authors in paper

@@
     affiliation: "3,4"
+  - name: "Loek Brinkman"
+    affiliation: "2,5"
+  - name: "Anna-Lena Lamprecht"
+    affiliation: "2,6"
+    email: "a.l.lamprecht@uu.nl"
   - name: "Aaron Peikert"
-    affiliation: "3,5"
+    affiliation: "3,7"
+  - name: "Marijn E. Struiksma"
+    affiliation: "2,8"
   - name: "Barbara Vreede"
-    affiliation: "6"
-  - name: "Anna-Lena Lamprecht"
-    affiliation: "2,7"
-    email: "a.l.lamprecht@uu.nl"
-  - name: "Loek Brinkman"
-    affiliation: "2,8"
-  - name: "Marijn E. Struiksma"
-    affiliation: "2,9"
-    email: "m.struiksma@uu.nl"
+    affiliation: "9"
 
@@
   - id: "5"
+    institution: "University Medical Center Utrecht"
+  - id: "6"
+    institution: "Utrecht University, Department of Information and Computing Sciences"
+  - id: "7"
     institution: "Humboldt-Universitaet zu Berlin"
-  - id: "6"
+  - id: "8"
+    institution: "Utrecht Institute of Linguistics OTS"
+  - id: "9"
     institution: "University Library Utrecht"
-  - id: "7"
-    institution: "Utrecht University, Department of Information and Computing Sciences"
-  - id: "8"
-    institution: "UMC Utrecht"
-  - id: "9"
-    institution: "Utrecht University, Department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS"
 
@@
   contributions are welcome. Please contact the lead author.
+  All authors of this paper, except the first, are listed in alphabetical order.
+  Author contributions are detailed in the document 'authors.csv' on the project
+  repository, at [https://github.com/cjvanlissa/worcs](https://github.com/cjvanlissa/worcs).
   The lead author is supported by a NWO Veni grant (NWO grant number

### 2020-05-29 paper/Manuscript.Rmd
AB changes

@@
   To facilitate widespread adoption, the
-  WORCS principles are implemented in the R-package `worcs`, which offers an
+  WORCS principles are implemented in the R package `worcs`, which offers an
   RStudio 
@@
   R package gert.
-keywords          : "open science; reproducibility; r; dynamic document generation; version control, FAIR"
+keywords          : "open science; reproducibility; r; dynamic document generation; version control"
 wordcount         : "8257"
@@
 
-Many researchers are motivated to adopt best practices for open
+Many researchers are motivated to adopt current best practices for open
 science and enjoy these benefits. And yet, the question of "how" to 
@@
 that meets most requirements open science, and is based on best practices:
-The Workflow for Open Reproducible 
-Code in Science (WORCS). WORCS is a lightweight approach to open science and
+The *Workflow for Open Reproducible Code in Science* (WORCS).
+WORCS is a lightweight approach to open science and
 computational reproducibility for scholars working in R on a private computer,
@@
 constitutes a step-by-step procedure that researchers can follow to make a 
-research project open and reproducible. WORCS is compatible with open science
+research project (including data, code, and a report) open and reproducible.
+WORCS is compatible with open science
 requirements already implemented by journals and institutions, and will help 
@@
 been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
-The R-package `worcs` offers a project template for RStudio
+The R package `worcs` [@worcspackage2020] offers a project template for RStudio
 [@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
@@
 Instead, several guidelines exist for practices that are considered "open".
-The TOP-guidelines are one of the most influential concrete operationalisations
+The TOP-guidelines are one of the most influential operationalisations
 of Open Science principles [@nosekPromotingOpenResearch2015a].
@@
 do not address replication of published results, because replication relates to
-the subject of a research project, not to its execution.
+the subject of a research project, not to its organization or execution.
 We group the remaining seven guidelines into three categories: citation (1),
@@
 
-Although there is a slight learning curve to transitioning to DDG, the
-investment really pays off. Time saved from painstakingly copy-pasting output
+Although there is a slight learning curve to transitioning to DDG, we strongly
+believe that the investment will pay off for users of the workflow, as it did for us.
+Time saved from painstakingly copy-pasting output
 and manually formatting text soon outweighs the investment of switching to a new
@@
 document can be automatically formatted in many styles, including
-APA style [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020],
+APA style [thanks to the R package `papaja`, @austPapajaPrepareReproducible2020],
 a host of other scientific formats 
@@
 although this paper focuses on the most widely used 'Git'
-cloud storage service, 'GitHub', two notable alternatives exist: 
-'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). As of
-version 0.1.2, the `worcs` template dialog accepts 'https' addresses from 
-repositories hosted on either of these platforms interchangeably.
-The package has a
-[vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) with
-additional instructions for using alternative platforms. In 
+remote repository service, 'GitHub', two notable alternatives exist: 
+'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org).
+In 
 order to present the workflow in a platform-agnostic way, we will refer to such
@@
 paper.
-However, `worcs` has been tested most extensively with 'GitHub'.
+As of
+version 0.1.2, `worcs` has been tested with these three remote repository services,
+but has been optimized for use with 'GitHub'.
+A [vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) is included
+with instructions for using alternative platforms. 
+When initializing a new project using the `worcs` project template,
+a 'https' remote repository address will usually work best with the authentication procedure outlined in the
+[setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
 
@@
 This is effective, and also preserves software outside the R environment (e.g.,
-preprocessing tools for neuroimaging data). An excellent Docker-based workflow
+preprocessing tools for neuroimaging data). A Docker-based workflow
 was recently introduced by @peikertReproducibleDataAnalysis2019. This workflow
@@
 conceptually and practically compatible. However, the workflow described by 
-Peikert and Brandmaier can be difficult to set up for novice users. An R-package
+Peikert and Brandmaier can be difficult to set up for novice users. 
+The accompanying R package `repro`
 is currently in development to facilitate this transition [@repro]. Another 
@@
 
-The R-package `worcs` automates most steps of the workflow
+The R package `worcs` automates most steps of the workflow
 for R users.
@@
 Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
-Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
+Fourth, R is well-supported;
+a vibrant and inclusive online community exists for R,
+developing new R methods and packages,
+and providing support and tutorials for existing ones.
 Finally, R is highly interoperable:
@@
 recommends two solutions to accomplish this goal.
-The first solution is to publish a *checksum* of the original data file.
+The first solution is to publish a *checksum* of the original data file [@rivest1992md5].
 Think of a checksum as a 32-character summary, or as a "one word" description,
 of the contents of a file.
-Any change to the file will result in a different checksum. 
+Any change to the file will result in a different checksum^[It is theoretically possible but improbable that random changes to a file will result in the same checksum.]. 
 Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
@@
 machine readable, and because data typically need to be in tabular format
-for analysis. Many types of data can be represented as a table - text
-corpora can have one row per document, and EEG- or EKG waveforms can have one
-row per measurement occasion. The `prepare_data.R` file should document any
+for analysis. Many types of data can be represented as a table; for example,
+text corpora can have one row per document, and EEG- or ECG waveforms can have a
+row per measurement occasion and a column per channel.
+The `prepare_data.R` file should document any
 steps required to convert these data into tabular format. 
@@
 primarily by explicitly inviting a user to write the preregistration document.
-To this end, `worcs` imports several preregistration templates from the R-package
+To this end, `worcs` imports several preregistration templates from the R package
 `prereg` [@austPreregMarkdownTemplates2019], including templates from

### 2020-05-29 README.Rmd
Update template

@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
-<!--[![CRAN status](https://www.r-pkg.org/badges/version/worcs)](https://cran.r-project.org/package=worcs)
-[![](https://cranlogs.r-pkg.org/badges/worcs)](https://cran.r-project.org/package=worcs)-->
+[![CRAN status](https://www.r-pkg.org/badges/version/worcs)](https://cran.r-project.org/package=worcs)
+[![](https://cranlogs.r-pkg.org/badges/worcs)](https://cran.r-project.org/package=worcs)
 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
@@
 
-> Van Lissa, C. J. (2020, February 11). WORCS: A Workflow for Open Reproducible Code in Science. https://doi.org/10.17605/OSF.IO/ZCVBS
+> Van Lissa, C.
+J., Brandmaier, A.
+M., Brinkman, L., Lamprecht, A., Peikert, A., , Struiksma, M.
+E., & Vreede, B.
+(2020, May 29).
+WORCS: A Workflow for Open Reproducible Code in Science.
+https://doi.org/10.17605/OSF.IO/ZCVBS
 

### 2020-05-29 R/worcs_checklist.R

@@ -19,7 +19,8 @@
 #' @keywords datasets
 #' @name worcs_checklist
 #' @usage data(worcs_checklist)
-#' @references by Van Lissa, Brandmaier, Brinkman, & Vreede (2020)
+#' @references Van Lissa, C. J., Brandmaier, A. M., Brinkman, L., Lamprecht, A.,
+#' Peikert, A., , Struiksma, M. E., & Vreede, B. (2020)
 #' <doi:10.17605/OSF.IO/ZCVBS>.
 #' @format A data frame with 15 rows and 5 variables.
 NULL

### 2020-05-29 R/worcs_project.R

@@ -99,7 +99,7 @@ worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregist
     copy_resources(which_files = c(
       "README.md",
       "prepare_data.R",
-      "worcs_badge.png"
+      "worcs_icon.png"
     ), path = path)
     col_message("Copying standard files.")
   }, error = function(e){
@@ -113,55 +113,63 @@ worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregist
   if(!manuscript == "none"){
     tryCatch({
       # Construct path to filename and create directory
-      man_dir <- file.path(path, "manuscript")
-      manuscript_file <- file.path(man_dir, "manuscript.Rmd")
-      dir.create(man_dir)
+      man_dir_rel <- "manuscript"
+      man_dir_abs <- file.path(path, man_dir_rel)
+      man_fn_rel <- file.path(man_dir_rel, "manuscript.Rmd")
+      man_fn_abs <- file.path(man_dir_abs, "manuscript.Rmd")
+      dir.create(man_dir_abs)
       switch(manuscript,
-             apa6 = create_man_papaja(manuscript_file, remote_repo = remote_repo),
-             acm_article = create_man_rticles(manuscript_file, "acm_article", remote_repo = remote_repo),
-             acs_article = create_man_rticles(manuscript_file, "acs_article", remote_repo = remote_repo),
-             aea_article = create_man_rticles(manuscript_file, "aea_article", remote_repo = remote_repo),
-             agu_article = create_man_rticles(manuscript_file, "agu_article", remote_repo = remote_repo),
-             amq_article = create_man_rticles(manuscript_file, "amq_article", remote_repo = remote_repo),
-             ams_article = create_man_rticles(manuscript_file, "ams_article", remote_repo = remote_repo),
-             asa_article = create_man_rticles(manuscript_file, "asa_article", remote_repo = remote_repo),
-             biometrics_article = create_man_rticles(manuscript_file, "biometrics_article", remote_repo = remote_repo),
-             copernicus_article = create_man_rticles(manuscript_file, "copernicus_article", remote_repo = remote_repo),
-             ctex = create_man_rticles(manuscript_file, "ctex", remote_repo = remote_repo),
-             elsevier_article = create_man_rticles(manuscript_file, "elsevier_article", remote_repo = remote_repo),
[…]

### 2020-05-29 paper/update_documentation.R

@@ -1,3 +1,4 @@
+# Include in makefile: Rscript -e 'source("paper/update_documentation.R")
 txt <- readLines("vignettes/workflow.Rmd")
 txt <- txt[grepl("<!--S:", txt, fixed = T)|grepl("^###", txt)]
 txt <- gsub("(<!--S: |-->.*$)", "", txt)
@@ -17,17 +18,15 @@ readme <- append(readme, out)
 
 worcs:::write_as_utf(readme, "inst/rstudio/templates/project/resources/README.md")
 
-reorder_numbers <- function(txt = readClipboard()){
-  num_line <- which(grepl("^\\d{1,}\\.", txt))
-  for(i in seq_along(num_line)){
-    this_line <- txt[num_line[i]]
-    the_num <- as.numeric(strsplit(this_line, "\\.")[[1]][1])
-    if(!the_num == i){
-      txt[num_line[i]] <- gsub(paste0("^", the_num), i, txt[num_line[i]])
-    }
-  }
-  txt
-}
-cat(get_workflow(), sep = "\n")
+# reorder_numbers <- function(txt = readClipboard()){
+#   num_line <- which(grepl("^\\d{1,}\\.", txt))
+#   for(i in seq_along(num_line)){
+#     this_line <- txt[num_line[i]]
+#     the_num <- as.numeric(strsplit(this_line, "\\.")[[1]][1])
+#     if(!the_num == i){
+#       txt[num_line[i]] <- gsub(paste0("^", the_num), i, txt[num_line[i]])
+#     }
+#   }
+#   txt
+# }
 
-reorder_numbers()

### 2020-05-29 README.Rmd
update 2020-05-29 11:52:21

@@
 [![CRAN status](https://www.r-pkg.org/badges/version/worcs)](https://cran.r-project.org/package=worcs)
-[![](https://cranlogs.r-pkg.org/badges/worcs)](https://cran.r-project.org/package=worcs)
+[![metacran downloads](https://cranlogs.r-pkg.org/badges/worcs)](https://cran.r-project.org/package=worcs)
 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
@@
 [![CII Best Practices](https://bestpractices.coreinfrastructure.org/projects/3969/badge)](https://bestpractices.coreinfrastructure.org/projects/3969)
-[![CRAN status](https://www.r-pkg.org/badges/version/worcs)](https://CRAN.R-project.org/package=worcs)
-[![metacran downloads](https://cranlogs.r-pkg.org/badges/worcs)](https://cran.r-project.org/package=worcs)
 <!--[![Codecov test

### 2020-05-29 paper/Manuscript.Rmd
Line endings
update 2020-05-29 16:54:47

@@
   Author contributions are detailed in the document 'authors.csv' on the project
-  repository, at [https://github.com/cjvanlissa/worcs](https://github.com/cjvanlissa/worcs).
+  repository, at https://github.com/cjvanlissa/worcs.
   The lead author is supported by a NWO Veni grant (NWO grant number
@@
 user-friendly, and highly interoperable with other reference managers. A 
-tutorial for using Zotero with RMarkdown exists  [here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-rmarkdown/).
+tutorial for using Zotero with RMarkdown exists  [here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-RMarkdown/).
 
@@
 curve associated with the tools outlined in this paper. Learning to work with R,
-Rmarkdown, and Git requires an initial time investment. The amount of
+RMarkdown, and Git requires an initial time investment. The amount of
 time required is reduced by the availability of tutorials, such as this one, and
@@
 tends to pay off. Working with R opens the door to many cutting edge analysis
-techniques. Working with Rmarkdown saves time and prevents mistakes by avoiding
+techniques. Working with RMarkdown saves time and prevents mistakes by avoiding
 tedious copying of results into a text document. Working with Git keeps
@@
 Another important challenge is managing collaborations when only the lead author
-uses Rmarkdown, and the coauthors use Word. In this case, it is possible to Knit
+uses RMarkdown, and the coauthors use Word. In this case, it is possible to Knit
 the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf`
@@
 bidirectional conversion between Word and
-Rmarkdown, with support for tracked changes. However, `redoc` is currently in
+RMarkdown, with support for tracked changes. However, `redoc` is currently in
 suspended development, and for many users, a simple solution (i.e.,

### 2020-05-30 R/github.R
update 2020-05-30 09:06:15

@@ -17,6 +17,8 @@ has_git <- function(){
 #' @param email Character. The email address you want to use with 'Git'.
 #' @return No return value. This function is called for its side effects.
 #' @rdname git_user
+#' @examples
+#' do.call(git_user, worcs:::get_user())
 #' @export
 #' @importFrom gert git_config_global_set
 git_user <- function(name, email){

### 2020-05-31 README.Rmd
Add graph to readme

@@
 
-For most users, the recommended starting point is to read [the preprint paper, available on the Open Science Framework](https://osf.io/zcvbs/).
+For most users, the recommended starting point is to [read the preprint paper](https://osf.io/zcvbs/),
+which introduces the WORCS workflow, explains the underlying tools, and illustrates how the `worcs` package can be used to create a new project that follows the workflow.
 
-The paper describes the WORCS workflow, and how the `worcs` package fits into that workflow.
+The workflow is illustrated below; the [workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html) describes each step in detail.
+
+```{r, echo = FALSE}
+knitr::include_graphics("paper/workflow_graph/workflow.png")
+```
+
+## Installing the package
+
+Before installing the package, please read [this vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), which explains how to set up your computer for `worcs`.
+
+After reading [the vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), you can install the development version of the `worcs` package from GitHub with:
+
+``` r
+if(!requireNamespace("remotes"))install.packages("remotes")
+remotes::install_github("cjvanlissa/worcs", dependencies = TRUE, update = "never")
+remotes::install_github("crsh/papaja", dependencies = TRUE, update = "never")
+tinytex::install_tinytex()
+worcs::git_user("your_name", "your_email")
+```
+
+## Citing WORCS
+
+You can cite WORCS using the following citation (please use the same citation for either the package, or the paper):
+
+> Van Lissa, C.
+J., Brandmaier, A.
+M., Brinkman, L., Lamprecht, A., Peikert, A., , Struiksma, M.
+E., & Vreede, B.
+(2020, May 29).
+WORCS: A Workflow for Open Reproducible Code in Science.
+https://doi.org/10.17605/OSF.IO/ZCVBS
 
@@
 
-## Installing the package
-
-Before installing the package, please read [this vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), which explains how to set up your computer for `worcs`.
-
-After reading [the vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), you can install the development version of the `worcs` package from GitHub with:
-
-``` r
-if(!requireNamespace("remotes"))install.packages("remotes")
-remotes::install_github("cjvanlissa/worcs", dependencies = TRUE, update = "never")
-remotes::install_github("crsh/papaja", dependencies = TRUE, update = "never")
-tinytex::install_tinytex()
-worcs::git_user("your_name", "your_email")
-```
-
-## Citing WORCS
-
-You can cite WORCS using the following citation (please use the same citation for either the package, or the paper):
-
-> Van Lissa, C.
-J., Brandmaier, A.
-M., Brinkman, L., Lamprecht, A., Peikert, A., , Struiksma, M.
-E., & Vreede, B.
-(2020, May 29).
-WORCS: A Workflow for Open Reproducible Code in Science.
-https://doi.org/10.17605/OSF.IO/ZCVBS
-
 ## Contributing and Contact Information

### 2020-05-31 README.Rmd
update 2020-05-31 20:19:17

@@
 remotes::install_github("cjvanlissa/worcs", dependencies = TRUE, update = "never")
-remotes::install_github("crsh/papaja", dependencies = TRUE, update = "never")
 tinytex::install_tinytex()
 worcs::git_user("your_name", "your_email")
+```
+
+If you intend to write APA style manuscripts, you can additionally install the `papaja` package:
+
+``` r
+remotes::install_github("crsh/papaja", dependencies = TRUE, update = "never")
+
 ```

### 2020-05-31 README.Rmd
update 2020-05-31 20:24:18
update 2020-05-31 20:28:51

@@
 ```{r, echo = FALSE}
-knitr::include_graphics("paper/workflow_graph/workflow.png")
+knitr::include_graphics("https://github.com/cjvanlissa/worcs/raw/master/paper/workflow_graph/workflow.png")
 ```
@@
 
-The WORCS paper and `worcs` package are open to contributions from collaborators with valuable expertise, to improve both the workflow and the package.
-Relevant contributions warrant coauthorship to the paper or package.
+The `worcs` package is open to contributions from collaborators with valuable expertise, to improve both the workflow and the software.
+Relevant contributions warrant coauthorship to the package.
 Please contact the lead author at [c.j.vanlissa@uu.nl](mailto:c.j.vanlissa@uu.nl), or:

### 2020-06-02 paper/Manuscript.Rmd
Figures in text

@@
 
-figsintext        : no
+figsintext        : yes
 figurelist        : no
@@
 
-```{r workflow, fig.cap="Schematic illustration of the WORCS procedure"}
+```{r workflow, fig.cap="Schematic illustration of the WORCS procedure", out.width="100%"}
 knitr::include_graphics("workflow_graph/workflow.png")

### 2020-06-02 paper/Manuscript_JSS.Rmd
Prepare manuscript for JSS submission

--- 2020-05-29 paper/Manuscript.Rmd
@@
 ---
-title: |
-    | WORCS: A Workflow for Open Reproducible Code in Science
-    | `r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`
-
-shorttitle        : "WORKFLOW FOR OPEN SCIENCE"
-
+documentclass: jss
 author:
-  - name: "Caspar J. Van Lissa"
-    affiliation: "1,2"
-    corresponding: yes
-    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
-    email: "c.j.vanlissa@uu.nl"
-  - name: "Andreas M. Brandmaier"
-    affiliation: "3,4"
-  - name: "Loek Brinkman"
-    affiliation: "2,5"
-  - name: "Anna-Lena Lamprecht"
-    affiliation: "2,6"
-    email: "a.l.lamprecht@uu.nl"
-  - name: "Aaron Peikert"
-    affiliation: "3,7"
-  - name: "Marijn E. Struiksma"
-    affiliation: "2,8"
-  - name: "Barbara Vreede"
-    affiliation: "9"
-
-affiliation:
-  - id: "1"
-    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
-  - id: "2"
-    institution: "Open Science Community Utrecht"
-  - id: "3"
-    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
-  - id: "4"
-    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
-  - id: "5"
-    institution: "University Medical Center Utrecht"
-  - id: "6"
-    institution: "Utrecht University, Department of Information and Computing Sciences"
-  - id: "7"
-    institution: "Humboldt-Universitaet zu Berlin"
-  - id: "8"
-    institution: "Utrecht Institute of Linguistics OTS"
-  - id: "9"
-    institution: "University Library Utrecht"
-
-abstract: |
+  - name: Caspar J. van Lissa
+    affiliation: 'Utrecht University'
+    address: Padualaan 14, 3584CH Utrecht, The Netherlands
+    email: c.j.vanlissa@uu.nl
+  - name: Andreas M. Brandmaier
+    affiliation: 'Max Planck Institute'
+  - name: Loek Brinkman
+    affiliation: 'UMC Utrecht \AND'
+  - name: Anna-Lena Lamprecht
+    affiliation: 'Utrecht University'
+    email: a.l.lamprecht@uu.nl
+  - name: Aaron Peikert
+    affiliation: Max Planck Institute
+  - name: Marijn E. Struiksma
+    affiliation: 'Utrecht University'
+  - name: Barbara Vreede
+    affiliation: 'Utrecht University'
+title:
+  formatted: >
+    WORCS: A Workflow for Open Reproducible Code in Science
+  short:      "WORKFLOW FOR OPEN SCIENCE"
+abstract: >
   Adopting open science
   principles can be challenging and time-intensive, because doing so requires
-  substantial conceptual education and training in the use of new tools. The 
-  goal of this paper is to lower the threshold to adoption. We introduce the Workflow for
-  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
+  substantial conceptual education and training in the use of new tools.
+  This paper introduces the Workflow for
+  Open Reproducible Code in Science (WORCS): A step-by-step
   procedure that researchers can follow to make a research project open and
-  reproducible. It is based on best practices, and can be used either in
+  reproducible, accompanied by an open source software implementation that 
+  automates these steps.
+  The 
+  purpose of the workflow is to lower the threshold for 
+  adoption of open science principles. 
+  It is based on best practices, and can be used either in
   parallel to, or in absence of, top-down requirements by journals,
-  institutions, and funding bodies. This workflow is primarily targeted towards
+  institutions, and funding bodies.
+  This workflow is primarily targeted towards
   scholars working in R on a private computer,
@@
   To facilitate widespread adoption, the
-  WORCS principles are implemented in the R package `worcs`, which offers an
+  WORCS principles are implemented in the R package \pkg{worcs}, which offers an
   RStudio 
-  project template that is, effectively, a "one-click solution" for R users. 
-  This paper introduces the workflow, discusses how it meets different 
-  requirements of open science, and provides examples of the implementation of
-  `worcs` in R.
-
-authornote: |
-  This is a preprint, not a peer-reviewed publication. Comments and
-  contributions are welcome. Please contact the lead author.
-  All authors of this paper, except the first, are listed in alphabetical order.
-  Author contributions are detailed in the document 'authors.csv' on the project
-  repository, at https://github.com/cjvanlissa/worcs.
-  The lead author is supported by a NWO Veni grant (NWO grant number
-  VI.Veni.191G.090).
-  We acknowledge Jeroen Ooms for offering feedback and suggesting the use of the
-  R package gert.
-keywords          : "open science; reproducibility; r; dynamic document generation; version control"
-wordcount         : "8257"
-
+  project template that is, effectively, a "one-click solution" for R users,
+  and utility functions for specific workflow steps.
+  This paper addresses the design of the workflow,
+  discusses how it meets different 
+  standards for open science, and provides examples of the implementation of
+  \pkg{worcs} in \proglang{R}.
+authornote: "blabla"
+keywords:
+  # at least one keyword must be supplied
+  formatted: [open science, reproducibility, r, dynamic document generation, version control]
+  plain:     [open science, reproducibility, r, dynamic document generation, version control]
+preamble: >
+  \usepackage{amsmath}
 bibliography      : ["worcs.bib"]  
-
-figsintext        : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-lineno            : yes
-mask              : no
-
-class             : "man"
-classoption       : "noextraspace"
-output            : papaja::apa6_pdf
 urlcolor          : blue
-knit              : worcs::cite_all
+output: rticles::jss_article
+knit: worcs::cite_all
 ---
@@
 }
+knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message= FALSE)
 ```
-
-```{r analysis_preferences}
-# Seed for random number generation
-set.seed(42)
-```
-
-\setlength{\parskip}{0em}
-
-<!--\vspace{\baselineskip}-->
 
@@
 WORCS is a lightweight approach to open science and
-computational reproducibility for scholars working in R on a private computer,
+computational reproducibility for scholars working in \proglang{R} on a private computer,
 conducting research that involves academic prose, analysis code,
@@
 Although the principles underlying this workflow are universal, WORCS has first
-been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
-The R package `worcs` [@worcspackage2020] offers a project template for RStudio
+been implemented for \proglang{R} users [@rcoreteamLanguageEnvironmentStatistical2020].
+The \proglang{R} package \pkg{worcs} [@worcspackage2020] offers a project template for RStudio
 [@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
 a "one click solution". This paper introduces the WORCS workflow, discusses how
-it meets best practices for open science, and illustrates the use of the `worcs`
+it meets best practices for open science, and illustrates the use of the \pkg{worcs}
 package.
 
-```{r workflow, fig.cap="Schematic illustration of the WORCS procedure"}
-knitr::include_graphics("workflow_graph/workflow.png")
+```{r workflow, fig.cap="Schematic illustration of the WORCS procedure.", out.width="100%"}
+#knitr::include_graphics("workflow_graph/workflow.png")
+knitr::include_graphics("workflow_graph/workflow.svg")
 ```
@@
 
-In the R-implementation of WORCS, we recommend centering a research project
-around one dynamically generated 'RMarkdown' document, which includes all 
+In the \proglang{R} implementation of WORCS, we recommend centering a research project
+around one dynamically generated \code{RMarkdown} document, which includes all 
 analysis code. Longer scripts can be
-stored in `.R` files, and called from the main document using the `source()`
+stored in \code{.R} files, and called from the main document using the \code{source()}
 function. When a reader or reviewer compiles this document, all code is run
-automatically, thus verifying computational replicability. The 'RMarkdown' 
+automatically, thus verifying computational replicability. The \code{RMarkdown} 
 document can be automatically formatted in many styles, including
-APA style [thanks to the R package `papaja`, @austPapajaPrepareReproducible2020],
+APA style [thanks to the \proglang{R} package \pkg{papaja}, @austPapajaPrepareReproducible2020],
 a host of other scientific formats 
-[see @allaireRticlesArticleFormats2020], and as plain 'Markdown'.
+[see @allaireRticlesArticleFormats2020], and as plain \code{Markdown}.
 
@@
 
-An integral part of WORCS is the *explicit* version control software 'Git'
+An integral part of WORCS is the *explicit* version control software Git
 (www.git-scm.com). Git tracks
@@
 privacy sensitive data.
-A `.gitignore` file lists files that should not be version-controlled.
+A \code{.gitignore} file lists files that should not be version-controlled.
 These files thus exist only on the user's private computer.
 
-The functionality of 'Git' is amplified by services such as 'GitHub'
-(https://github.com). 'GitHub' is best understood as a cloud storage service
+The functionality of Git is amplified by services such as GitHub
+(https://github.com). GitHub is best understood as a cloud storage service
 with social networking functionality. The cloud storage aspect
-of 'GitHub' works as follows: You can "clone" (copy) a local 'Git' repository to
-the 'GitHub' website, as a backup or research archive. The social network aspect
-comes into play when a repository on 'GitHub' is set to "public": This allows 
+of GitHub works as follows: You can "clone" (copy) a local Git repository to
+the GitHub website, as a backup or research archive. The social network aspect
+comes into play when a repository on GitHub is set to "public": This allows 
 other researchers to peruse the repository and see how the work was done; clone
@@
 or even send a "Pull request" with suggested changes to the text or code for
-your consideration. 'Git' and 'GitHub' shine as tools for collaboration, because
+your consideration. Git and GitHub shine as tools for collaboration, because
 different people can simultaneously work on different parts of a project, and
 their changes can be compared and automatically merged on the website. Even on
-solo projects, working with 'Git'/'GitHub' has many benefits: Staying organized,
+solo projects, working with Git/GitHub has many benefits: Staying organized,
 being able to start a new study with a clone of an old, similar repository, or
@@
 It is worth noting that,
-although this paper focuses on the most widely used 'Git'
-remote repository service, 'GitHub', two notable alternatives exist: 
-'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org).
+although this paper focuses on the most widely used Git
+remote repository service, GitHub, two notable alternatives exist: 
+GitLab (https://gitlab.com) and Bitbucket (https://bitbucket.org).
 In 
@@
 As of
-version 0.1.2, `worcs` has been tested with these three remote repository services,
-but has been optimized for use with 'GitHub'.
+version 0.1.2, \pkg{worcs} has been tested with these three remote repository services,
+but has been optimized for use with GitHub.
 A [vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) is included
 with instructions for using alternative platforms. 
-When initializing a new project using the `worcs` project template,
+When initializing a new project using the \pkg{worcs} project template,
 a 'https' remote repository address will usually work best with the authentication procedure outlined in the
@@
 
-The R-implementation of WORCS uses the package `gert`, a 'Git' client for 'R'
+The \proglang{R} implementation of WORCS uses the package \code{gert}, a Git client for \proglang{R}
 [@oomsGertSimpleGit2019], to connect a local project to a remote repository.
-The `worcs` package also contains a user-friendly function to set
-the default user credentials, `git_user()`, and a function to add, commit, and
-push all changed files to a remote repository in one step, `git_update()`.
+The \pkg{worcs} package also contains a user-friendly function to set
+the default user credentials, \code{git_user()}, and a function to add, commit, and
+push all changed files to a remote repository in one step, \code{git_update()}.
 
@@
 Potentially, any such update could change the results of the code, thus rendering
-the analysis 'computationally non-reproducible'.
+the analysis computationally non-reproducible.
 
@@
 are explicitly installed, and maintaining a record of these programs. These solutions differ in
-user-friendliness and effectiveness. If strict reproducibility is required, users might choose to conduct all analyses in a 'Docker container': An environment that behaves like a
+user-friendliness and effectiveness. If strict reproducibility is required, users might choose to conduct all analyses in a "Docker container": An environment that behaves like a
 virtual computer, which can be stored like a sort of time capsule, and
 identically reinstated on a user computer, or in the cloud.
-This is effective, and also preserves software outside the R environment (e.g.,
+This is effective, and also preserves software outside the \proglang{R} environment (e.g.,
 preprocessing tools for neuroimaging data). A Docker-based workflow
@@
 Peikert and Brandmaier can be difficult to set up for novice users. 
-The accompanying R package `repro`
+The accompanying \proglang{R} package \pkg{repro}
 is currently in development to facilitate this transition [@repro]. Another 
 user-friendly Docker-based workflow is offered by the cloud-based collaboration
-platform 'Digital Ocean'. Although Digital Ocean enjoys some popularity among 
+platform Digital Ocean. Although Digital Ocean enjoys some popularity among 
 scientists, it is not clear whether this platform's
@@
 
-For analyses conducted entirely in R, the recently released package `renv`
+For analyses conducted entirely in \proglang{R}, the recently released package \pkg{renv}
 strikes a great balance between user-friendliness  and effectiveness. Developed by
-the team behind 'RStudio', this package simply maintains a text-based,
+the team behind RStudio, this package simply maintains a text-based,
 human-readable log of all packages used, their version numbers, and where they
-were installed from (e.g., 'CRAN', 'Bioconductor', 'GitHub'). As long as these
+were installed from (e.g., CRAN, Bioconductor, GitHub). As long as these
 repositories maintain their archives, this solution is reliable and very
 lightweight.
-The text-based log file can be version controlled with 'Git'. `renv`
+The text-based log file can be version controlled with Git. \pkg{renv}
 installs the logged packages in a cache that is shared between all
@@
 that is already installed on the system. When someone else loads your 
-project, `renv` will install all of the required packages onto a cache on their
+project, \pkg{renv} will install all of the required packages onto a cache on their
 computer. 
@@
 acknowledge that all code requires some maintenance if you want to
-reproduce it in the future. For projects in R, `renv` meets these requirements.
+reproduce it in the future. For projects in \proglang{R}, \pkg{renv} meets these requirements.
 
@@
 Text-based files can be read by machines and humans alike.
-Binary files, such as 'Word' (`.docx`) or 'SPSS' (`.sav`, `.spo`) files,
+Binary files, such as Word (\code{.docx}) or SPSS (\code{.sav}, \code{.spo}) files,
 must be decoded first, and many require commercial software to do so.
-The version control software 'Git' is also designed for use with text-based files.
-The 'Git' change log reflects changes to text-based files on a line-by-line basis.
-Binary files can be version controlled with 'Git', but 
+The version control software Git is also designed for use with text-based files.
+The Git change log reflects changes to text-based files on a line-by-line basis.
+Binary files can be version controlled with Git, but 
 the change log for a binary file are uninterpretable.
@@
 Two additional points are worth noting:
-First, 'Git' tracks line-by-line changes to text-based files,
+First, Git tracks line-by-line changes to text-based files,
 based on line breaks in the text.
@@
 this practice does not disrupt the flow of a paragraph in the rendered version.
-Second, the remote repository service 'GitHub' rewards the use of text-based files by
+Second, the remote repository service GitHub rewards the use of text-based files by
 rendering certain text-based filetypes for online viewing: 
-For example, `.md`
-(Markdown) files are displayed as web pages, and `.csv` files as spreadsheets.
-The `worcs` implementation for R makes use of this fact by, for example,
-rendering a codebook for the data as `.md` so visitors of a 'Git' remote repository
+For example, \code{.md}
+(Markdown) files are displayed as web pages, and \code{.csv} files as spreadsheets.
+The \pkg{worcs} implementation for \proglang{R} makes use of this fact by, for example,
+rendering a codebook for the data as \code{.md} so visitors of a Git remote repository
 can view it as a web page.
@@
 Version 0.1.2. 
-For users of the `worcs` R package, we highly recommend referring to
+For users of the \pkg{worcs} \proglang{R} package, we highly recommend referring to
 [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
@@
 
-## The R-implementation of WORCS
-
-The R package `worcs` automates most steps of the workflow
-for R users.
-Several arguments support the choice to implement this workflow in R.
-First, R and all of it extensions are free open source software, which
+## The R implementation of WORCS
+
+The \proglang{R} package \pkg{worcs} automates most steps of the workflow
+for \proglang{R} users.
+Several arguments support the choice to implement this workflow in \proglang{R}.
+First, \proglang{R} and all of it extensions are free open source software, which
 make it a tool of choice for open science.
-Second, all tools required for an open science workflow are implemented in R,
+Second, all tools required for an open science workflow are implemented in \proglang{R},
 most of these tools are directly accessible through the user interface of RStudio,
 and some of them are actively developed by the team behind RStudio.
-Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
-Fourth, R is well-supported;
-a vibrant and inclusive online community exists for R,
-developing new R methods and packages,
+Third, \proglang{R} is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
+Fourth, \proglang{R} is well-supported;
+a vibrant and inclusive online community exists for \proglang{R},
+developing new \proglang{R} methods and packages,
 and providing support and tutorials for existing ones.
-Finally, R is highly interoperable:
+Finally, \proglang{R} is highly interoperable:
 Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
-Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
-Fortran, and Python [@allaireMarkdownPythonEngine2020]. There are excellent free
-resources for learning R, such as the book 'R for Data Science' [@grolemundDataScience2017].
-
-Working with R is simplified immensely by using the RStudio integrated
-development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
+Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from \proglang{R}, including \proglang{C++}, 
+\proglang{Fortran}, and \proglang{Python} [@allaireMarkdownPythonEngine2020]. There are excellent free
+resources for learning \proglang{R} [e.g., @grolemundDataScience2017].
+
+Working with \proglang{R} is simplified immensely by using the RStudio integrated
+development engine (IDE) for \proglang{R} [@rstudioteamRStudioIntegratedDevelopment2015].
 RStudio automates and streamlines tedious or complicated aspects of working with
@@
 A project bundles writing, analyses, data, references, et cetera, into a
-self-contained folder, that can be uploaded entirely to a 'Git' remote repository and downloaded
-by future users. The `worcs` R package installs a new RStudio project template.
+self-contained folder, that can be uploaded entirely to a Git remote repository and downloaded
+by future users. The \pkg{worcs} \proglang{R} package installs a new RStudio project template.
 When a new project is initialized from this template, the bookkeeping required
@@
 
-Before you can use the R-implementation of the WORCS workflow, you have to install the required software.
+Before you can use the \proglang{R} implementation of the WORCS workflow, you have to install the required software.
 You only have to perform this setup procedure only once for each computer you
-intend to use `worcs` on. It should take about 30 minutes, and it is documented
+intend to use \pkg{worcs} on. It should take about 30 minutes, and it is documented
 in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
-The setup vignette does not assume a prior installation of R, so it is suitable for novice users.
-After setting up your system, you can use `worcs` for all of your
-projects. The R-implementation of the workflow introduced earlier is detailed,
+The setup vignette does not assume a prior installation of \proglang{R}, so it is suitable for novice users.
+After setting up your system, you can use \pkg{worcs} for all of your
+projects. The \proglang{R} implementation of the workflow introduced earlier is detailed,
 step-by-step, in
@@
 References for software are sometimes provided within the software environment;
-for example, in R, package references can be found by
-calling `citation("packagename")`. This returns an APA-style reference, and a
+for example, in \proglang{R}, package references can be found by
+calling \code{citation("packagename")}. This returns an APA-style reference, and a
 BibTeX entry. Software papers are also common, and are sometimes called
@@
 Software citations can also sometimes be found 
-in the source code repository (e.g., on 'Git' remote repositories), using the
+in the source code repository (e.g., on Git remote repositories), using the
 ['Citation File Format'](https://citation-file-format.github.io/). The
@@
 
-## Implementation in `worcs`
-
-The `worcs` package includes a [vignette on citation](https://cjvanlissa.github.io/worcs/articles/citing.html) to explain the
-fundamentals for novice users, and offer recommendations specific to `worcs`. We
+## Implementation in worcs
+
+The \pkg{worcs} package includes a [vignette on citation](https://cjvanlissa.github.io/worcs/articles/citing.html) to explain the
+fundamentals for novice users, and offer recommendations specific to \pkg{worcs}. We
 recommend using the free, open-source reference manager
@@
 
-The `worcs` package also offers original solutions for comprehensive citation.
+The \pkg{worcs} package also offers original solutions for comprehensive citation.
 Firstly,
@@
 going back to cut non-essential references just prior to publication. Standard
-Markdown uses the at-symbol (`@`) to cite a reference. `worcs` additionally 
-reserves the "double at"-symbol (`@@`) to cite a non-essential reference.
+Markdown uses the at-symbol (\code{@}) to cite a reference. \pkg{worcs} additionally 
+reserves the "double at"-symbol (\code{@}) to cite a non-essential reference.
 Users can render the manuscript either with, or without, comprehensive citations
-by adapting the `knit` command in the front matter, setting it to
-`knit: worcs::cite_all` to render all citations, and to 
-`knit: worcs::cite_essential` to remove all *non-essential* citations.
-
-With regard to the citation of R packages, it is worth noting that the `papaja` 
-package [@austPapajaPrepareReproducible2020], which provides the 'APA6'
-template used in `worcs`, also
-includes functions to automatically cite all packages used in the R-session, and
+by adapting the \code{knit} command in the front matter, setting it to
+\code{knit: worcs::cite_all} to render all citations, and to 
+\code{knit: worcs::cite_essential} to remove all *non-essential* citations.
+
+With regard to the citation of \proglang{R} packages, it is worth noting that the \pkg{papaja} 
+package [@austPapajaPrepareReproducible2020], which provides the \code{APA6}
+template used in \pkg{worcs}, also
+includes functions to automatically cite all packages used in the \proglang{R} session, and
 add their references to the bibliography file.
@@
 
-If data can be shared openly, they can simply be pushed to a public 'Git'
+If data can be shared openly, they can simply be pushed to a public Git
 repository, along with the analysis code. This way, others can download the
@@
 the authors who evaluate it on the real data and send back the results. The
-`worcs` package offers a simple but flexible function to generate synthetic data,
-`synthetic()`, which by default uses random forests to generate a model for
+\pkg{worcs} package offers a simple but flexible function to generate synthetic data,
+\code{synthetic()}, which by default uses random forests to generate a model for
 synthesis (see @wrightRangerFastImplementation2015). For data synthesis to work
-properly, it is essential that the `class` (data type) of variables is defined
-correctly. As of `worcs` version `r worcs:::cran_version()`, numeric, integer, factor, and logical data
+properly, it is essential that the \code{class} (data type) of variables is defined
+correctly. As of \pkg{worcs} version \code{r worcs:::cran_version()}, numeric, integer, factor, and logical data
 are supported out of the box. Other types of variables should be converted to
-one of these types, or users can use a custom `model_expression` when calling
-`synthetic`.
+one of these types, or users can use a custom \code{model_expression} when calling
+\code{synthetic}.
 
@@
 
-When initializing a new `worcs` project, an empty R-script called `prepare_data.R` is generated.
+When initializing a new \pkg{worcs} project, an empty \proglang{R} script called \code{prepare_data.R} is generated.
 As soon as raw data are collected, researchers should use this file to document
 all steps necessary to load the 
-data into R, pseudonimize it, and prepare it for analysis.
+data into \proglang{R}, pseudonimize it, and prepare it for analysis.
 As data should be shared in a format as close to raw as possible, 
@@
 a researcher discovers that in one of the variables, missing values were coded as
-`-99` whereas they were coded as `-9` in all other variables, then we would recommend adding
+\code{-99} whereas they were coded as \code{-9} in all other variables, then we would recommend adding
 a line of code to recode the missing values - thereby documenting the mistake - 
@@
 
-The `worcs` package offers two functions for version controlling a record of the
+The \pkg{worcs} package offers two functions for version controlling a record of the
 data: One for open, and one for closed data. Researchers should assume that the
@@
 
-In `worcs`, researchers can call the function `open_data(data)` to make the data
-publicly available. This function stores the `data` object (such as a `data.frame` or
-`matrix`) in a tabular data file (called `data.csv` by default).
+In \pkg{worcs}, researchers can call the function \code{open_data(data)} to make the data
+publicly available. This function stores the \code{data} object (such as a \code{data.frame} or
+\code{matrix}) in a tabular data file (called \code{data.csv} by default).
 It also generates an
@@
 that can be viewed online.
-All changed files should now be added to the 'Git' repository, committed, and
+All changed files should now be added to the Git repository, committed, and
 pushed to the remote repository.
-This can be done through the 'Git' panel in RStudio, or by running the `worcs`
-function `git_update("commit message")`.
+This can be done through the Git panel in RStudio, or by running the \pkg{worcs}
+function \code{git_update("commit message")}.
 Once the remote repository is made public, these data will be
@@
 Alternatively, if the project requires data to remain closed, researchers can
-call `closed_data(data)`. This function also stores the data in a local `.csv` file
+call \code{closed_data(data)}. This function also stores the data in a local \code{.csv} file
 and generates a codebook,
-but the original data file is added to `.gitignore` so it cannot be
-accidentally added to the 'Git' repository.
+but the original data file is added to \code{.gitignore} so it cannot be
+accidentally added to the Git repository.
 The function also computes a checksum for the original data, and logs it in the
-`.worcs` project file. This checksum serves as the indelible record of the state 
+\code{.worcs} project file. This checksum serves as the indelible record of the state 
 of the raw data.
-The function `closed_data()` also creates a synthetic dataset (called
-`synthetic_data.csv` by default).
+The function \code{closed_data()} also creates a synthetic dataset (called
+\code{synthetic_data.csv} by default).
 The user should push all changed files to the remote repository.
@@
 
-As the purpose of the `prepare_data.R` file is to prepare data for sharing,
+As the purpose of the \code{prepare_data.R} file is to prepare data for sharing,
 the final line of this file
 should typically be either
-`open_data()` or `closed_data()`.
-
-After generating tidy, shareable data, this data is then loaded in the analysis code with the function `load_data()`.
+\code{open_data()} or \code{closed_data()}.
+
+After generating tidy, shareable data, this data is then loaded in the analysis code with the function \code{load_data()}.
 This function only loads the real data if it is available on the user's computer, 
@@
 the original data.
-They can simply start the script with `load_data()`, and
+They can simply start the script with \code{load_data()}, and
 write their code based on the synthetic data.
@@
 ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request).
-When you run their code on your system, `load_data()` will load the
+When you run their code on your system, \code{load_data()} will load the
 original data, and use that to run their script. You can then return the results
@@
 Note that these functions can be used to store multiple data files, if necessary.
-As of version 0.1.2, the `worcs` package only stores data as `.csv` files.
-As explained above, this is recommended, because `.csv` files are human- and
+As of version 0.1.2, the \pkg{worcs} package only stores data as \code{.csv} files.
+As explained above, this is recommended, because \code{.csv} files are human- and
 machine readable, and because data typically need to be in tabular format
@@
 row per measurement occasion and a column per channel.
-The `prepare_data.R` file should document any
+The \code{prepare_data.R} file should document any
 steps required to convert these data into tabular format. 
 If this is not an appropriate format for the data, readers are encouraged to follow the
-development of the `repro` package [@repro], which will offer greater support for
+development of the \pkg{repro} package [@repro], which will offer greater support for
 reproducibly storing and retrieving diverse data formats.
 
-Some users may intend to make their data openly available through a restricted access platform, such as an institutional repository, but not publicly through a 'Git' remote repository.
-In this case, it is recommended to use `closed_data()`, and to manually upload the original `data.csv` file to the restricted access platform.
-If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the 'Git' remote repository as an [Add-on](https://help.osf.io/hc/en-us/categories/360001550973-Add-ons).
+Some users may intend to make their data openly available through a restricted access platform, such as an institutional repository, but not publicly through a Git remote repository.
+In this case, it is recommended to use \code{closed_data()}, and to manually upload the original \code{data.csv} file to the restricted access platform.
+If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the Git remote repository as an [Add-on](https://help.osf.io/hc/en-us/categories/360001550973-Add-ons).
 
@@
 analysis code can be met simply by committing the source code of the manuscript
-to the 'Git' repository, and making this remote repository Public.
+to the Git repository, and making this remote repository Public.
 If authors additionally use open data and a reproducible environment (as suggested
 in WORCS), then a third party can simply replicate all analyses by copying the
-entire repository from the 'Git' hosting service, and Knitting the manuscript on
+entire repository from the Git hosting service, and Knitting the manuscript on
 their local computer.
@@
 materials, and details of the study design and analysis. These goals can be
-accomplished by placing any such documents in the 'Git' repository folder,
+accomplished by placing any such documents in the Git repository folder,
 committing them, and pushing to a cloud hosting service. As with any document
@@
 [AsPredicted.org](https://aspredicted.org/), and [OSF.io](https://osf.io/).
-However, 'Git' cloud hosting services also conform to these standards.
+However, Git cloud hosting services also conform to these standards.
 Thus, it is possible to preregister a study simply by committing a preregistration document
-to the local 'Git' repository, and pushing it to the remote repository.
+to the local Git repository, and pushing it to the remote repository.
 Subsequently [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases)
@@
 methods for simulating data are available in most statistical programming languages.
-For instance, `R` users can use the package `simstudy` [@goldfeldSimstudySimulationStudy2020].
-
-As soon as a project is preregistered on a Public 'Git' repository, it is
+For instance, \proglang{R} users can use the package \code{simstudy} [@goldfeldSimstudySimulationStudy2020].
+
+As soon as a project is preregistered on a Public Git repository, it is
 visible to the world, and reviewers (both formal reviewers designated by a
 journal, and informal reviewers recruited by other means) can submit comments,
-e.g., through "Pull requests" on a 'Git' remote repository. 
+e.g., through "Pull requests" on a Git remote repository. 
 If the remote repository is Private, Reviewers
@@
 
-### Implementation of preregistration in `worcs`
-
-The `worcs` workflow facilitates preregistration,
+### Implementation of preregistration in worcs
+
+The \pkg{worcs} workflow facilitates preregistration,
 primarily by explicitly inviting a user to write the preregistration document.
-To this end, `worcs` imports several preregistration templates from the R package
-`prereg` [@austPreregMarkdownTemplates2019], including templates from
+To this end, \pkg{worcs} imports several preregistration templates from the \proglang{R} package
+\code{prereg} [@austPreregMarkdownTemplates2019], including templates from
 organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and
 from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
-When initializing an RStudio project with the `worcs` project template, one of
+When initializing an RStudio project with the \pkg{worcs} project template, one of
 these preregistration templates can be selected, which will generate a file
-called `preregistration.Rmd`. This file should be used to document study plans,
+called \code{preregistration.Rmd}. This file should be used to document study plans,
 ideally prior to data collection. 
-Within the workflow, the 'Git' cloud hosting services can be used to preregister a study simply by pushing a text document with the study plans to a 'Git' remote repository.
+Within the workflow, the Git cloud hosting services can be used to preregister a study simply by pushing a text document with the study plans to a Git remote repository.
 By [tagging the commit](https://help.github.com/en/github/administering-a-repository/creating-releases) as a preregistration
@@
 including data and code - accessible in a repository that can be made public. 
-The `worcs` package additionally
-facilitates meeting the FAIR principles in several ways. First, each `worcs`
+The \pkg{worcs} package additionally
+facilitates meeting the FAIR principles in several ways. First, each \pkg{worcs}
 project has a YAML file (https://yaml.org/) that lists the research objects in
 the project, which will allow web crawlers 
-to index `worcs` projects. Second, a codebook is automatically generated for
+to index \pkg{worcs} projects. Second, a codebook is automatically generated for
 each data file stored
-using `open_data()` or `closed_data()`. We encourage users to elaborate on the
+using \code{open_data()} or \code{closed_data()}. We encourage users to elaborate on the
 default codebook by adding
@@
 
-Incidentally, `worcs` itself follows the recommendations for FAIR research software [see https://fair-software.eu/]. It is 
-hosted on a version controlled public remote repository 'GitHub', has an open
+Incidentally, \pkg{worcs} itself follows the recommendations for FAIR research software [see https://fair-software.eu/]. It is 
+hosted on a version controlled public remote repository GitHub, has an open
 source licence (GPL v3.0), is registered in a community registry (CRAN),
-enables the citation of the software (using the `citation("worcs")` command, or
+enables the citation of the software (using the \code{citation("worcs")} command, or
 by citing this paper), and followed the [*CII Best Practices*](https://bestpractices.coreinfrastructure.org/en)
@@
 surrogate indicators like journal impact factors, conference rankings, and
-h-indexes. The detailed version control and commit tracking of 'Git' remote repositories
+h-indexes. The detailed version control and commit tracking of Git remote repositories
 furthermore make it possible to assess the relative
@@
 The workflow aims to lower the threshold for grass-roots adoption of open science principles.
-The workflow is supported by an R package with an RStudio project template and 
+The workflow is supported by an \proglang{R} package with an RStudio project template and 
 convenience functions.
@@
 items contributing to an open and reproducible research project. Users of the
-`worcs` package can receive a badge, displayed on the `README.md` file,
+\pkg{worcs} package can receive a badge, displayed on the \code{README.md} file,
 based on a semi-automated scoring of the checklist items, by calling the function
-`check_worcs()` within a `worcs` project directory.
+\code{check_worcs()} within a \pkg{worcs} project directory.
 
@@
 sets WORCS apart is that is an easy and sufficient solution for projects that
-can be conducted entirely within R, resulting in a much more lightweight
+can be conducted entirely within \proglang{R}, resulting in a much more lightweight
 approach to safeguard computational reproducibility under *most circumstances*.
@@
 issues for future development remain. One potential challenge is the learning
-curve associated with the tools outlined in this paper. Learning to work with R,
+curve associated with the tools outlined in this paper. Learning to work with \proglang{R},
 RMarkdown, and Git requires an initial time investment. The amount of
@@
 other tutorials cited throughout this document. Moreover, the time investment
-tends to pay off. Working with R opens the door to many cutting edge analysis
+tends to pay off. Working with \proglang{R} opens the door to many cutting edge analysis
 techniques. Working with RMarkdown saves time and prevents mistakes by avoiding
@@
 uses RMarkdown, and the coauthors use Word. In this case, it is possible to Knit
-the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf`
-in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some
+the manuscript to Word (.docx), by changing the line \code{output: papaja::apa6_pdf}
+in the manuscript's YAML header to \code{output: papaja::apa6_docx}. There are some
 limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
@@
 comment bubbles in Word. <!-- LB: Overleaf?-->Then, manually incorporate their
-changes into the `manuscript.Rmd` file. In most cases, this is the most
+changes into the \code{manuscript.Rmd} file. In most cases, this is the most
 user-friendly approach, as most lead authors would review changes by co-authors
 anyway. A second approach is to ask collaborators to work in plain text. In this
-case, send collaborators the `manuscript.Rmd` file, and ask them to open (and
+case, send collaborators the \code{manuscript.Rmd} file, and ask them to open (and
 save) it in Word or Notepad as a plain text file. When they send it back, make
@@
 relative to your last committed version. Finally,
-@peikertReproducibleDataAnalysis2019 mention the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] as a
+@peikertReproducibleDataAnalysis2019 mention the [\code{redoc} package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] as a
 potentially useful way to collaborate with Word users. This package enables
 bidirectional conversion between Word and
-RMarkdown, with support for tracked changes. However, `redoc` is currently in
+RMarkdown, with support for tracked changes. However, \code{redoc} is currently in
 suspended development, and for many users, a simple solution (i.e.,
@@
 scientific collaboration [@ramGitCanFacilitate2013]. Recall that, when using 
-'Git' for collaborative writing, it is recommended to insert a
+Git for collaborative writing, it is recommended to insert a
 line break after every sentence so that the change log will indicate which
@@
 Being familiar
-with 'Git' remote repositories opens doors to new forms of collaboration:
+with Git remote repositories opens doors to new forms of collaboration:
 In the open source
@@
 in scientific software development as well; for example, the lead author of this
-paper became a co-author on several R packages after submitting pull requests
+paper became a co-author on several \proglang{R} packages after submitting pull requests
 with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018], and two co-authors of this
-paper became involved by contributing pull requests to `worcs`. It is also
+paper became involved by contributing pull requests to \pkg{worcs}. It is also
 possible to invite such collaboration by 
@@
 Nevertheless, there are clear directions for future developments.
-Firstly, although the workflow is currently implemented only in R, it is 
+Firstly, although the workflow is currently implemented only in \proglang{R}, it is 
 conceptually relevant for users of other statistical programming languages, most
-notably Python [@10.5555/1593511].
+notably \proglang{Python} [@10.5555/1593511].
 We welcome efforts to implement WORCS in other platforms.
-Secondly, even when a project is in R, it may have dependencies outside of the
-R environment that cannot be managed using `renv`.
-It is beyond the scope of the `worcs` package to support the incorporation of
-tools outside of R, or to containerize a project so that it can
+Secondly, even when a project is in \proglang{R}, it may have dependencies outside of the
+\proglang{R} environment that cannot be managed using \pkg{renv}.
+It is beyond the scope of the \pkg{worcs} package to support the incorporation of
+tools outside of \proglang{R}, or to containerize a project so that it can
 be identically reinstated on a different system or virtual machine.
@@
 workflow. To this end, we are collaborating with @repro on the development of
-`repro`, an R package that will facilitate setting up reproducibility tools such
-as 'Make' and 'Docker'.
+\pkg{repro}, an \proglang{R} package that will facilitate setting up reproducibility tools such
+as Make and Docker.
 This will make it possible for researchers
@@
 or vice versa, to initiate a WORCS project within a strictly reproducible
-environment generated using `repro`. 
+environment generated using \pkg{repro}. 
 This will combine the strengths of WORCS with those of the workflow proposed
@@
 The step-by-step procedure outlined in this tutorial helps researchers make an entire research project Open and Reproducible.
-The accompanying R package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.
+The accompanying \proglang{R} package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.
 
@@
 This enables other researchers to reproduce results, and facilitates cumulative science by allowing others to make direct use of these research objects.
-<!-- Somehow this feels like very sudden and incomplete ending to me. I am sure you can add some more glorious last words. :) -->
 
 \newpage
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-\setlength{\parskip}{0em}
-
-<div id = "refs"></div>
-\endgroup

### 2020-06-02 paper/Manuscript_JSS.Rmd
JSS Manuscript

@@
   standards for open science, and provides examples of the implementation of
-  \pkg{worcs} in \proglang{R}.
+  \pkg{worcs} in \proglang{R}. All source code for the \proglang{R} package,
+  and replication materials for the manuscript, are available at 
+  https://github.com/cjvanlissa/worcs.
 authornote: "blabla"
@@
 
-Incidentally, \pkg{worcs} itself follows the recommendations for FAIR research software [see https://fair-software.eu/]. It is 
+Incidentally, \pkg{worcs} itself follows the recommendations for [FAIR research software](https://fair-software.eu/). It is 
 hosted on a version controlled public remote repository GitHub, has an open

### 2020-06-03 paper/Manuscript.Rmd
Complete author information; fix Code Ocean

@@
 user-friendly Docker-based workflow is offered by the cloud-based collaboration
-platform 'Digital Ocean'. Although Digital Ocean enjoys some popularity among 
-scientists, it is not clear whether this platform's
-["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/)
-is GDPR compliant - it seems to focus more on limiting liability than on
-compliance. The necessity to upload data for analysis is a risk with any
+platform 'Code Ocean'. Although Code Ocean enjoys some popularity among 
+scientists, it is not clear whether this platform
+is GDPR compliant - as of this writing, its [terms of use](https://codeocean.com/terms-of-use) do not address the GDPR, nor is it clear where
+data are stored. The necessity to upload data for analysis is a risk with any
 cloud-computing solution when human participants are involved.

### 2020-06-04 paper/Manuscript.Rmd
recomend knitr::read_chunk()
closes #75

@@
 analysis code. Longer scripts can be
-stored in `.R` files, and called from the main document using the `source()`
-function. When a reader or reviewer compiles this document, all code is run
+stored in `.R` files, and included in the main document using the [`read_chunk()`
+function](https://github.com/yihui/knitr-examples/blob/master/113-externalization.Rmd) from the `knitr` package, which comes together with the `rmarkdown` package.
+When a reader or reviewer compiles this document, all code is run
 automatically, thus verifying computational replicability. The 'RMarkdown' 

### 2020-06-26 paper/Manuscript_JSS.Rmd

@@
 user-friendly Docker-based workflow is offered by the cloud-based collaboration
-platform Digital Ocean. Although Digital Ocean enjoys some popularity among 
-scientists, it is not clear whether this platform's
-["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/)
-is GDPR compliant - it seems to focus more on limiting liability than on
-compliance. The necessity to upload data for analysis is a risk with any
+platform 'Code Ocean'. Although Code Ocean enjoys some popularity among 
+scientists, it is not clear whether this platform
+is GDPR compliant - as of this writing, its [terms of use](https://codeocean.com/terms-of-use) do not address the GDPR, nor is it clear where
+data are stored.
+The necessity to upload data for analysis is a risk with any
 cloud-computing solution when human participants are involved.

### 2020-09-28 paper/Manuscript_DS.Rmd
Data Science submission

--- 2020-06-26 paper/Manuscript_JSS.Rmd
@@
 ---
-documentclass: jss
 author:
@@
   - name: Loek Brinkman
-    affiliation: 'UMC Utrecht \AND'
+    affiliation: 'UMC Utrecht'
   - name: Anna-Lena Lamprecht
@@
     affiliation: 'Utrecht University'
-title:
-  formatted: >
-    WORCS: A Workflow for Open Reproducible Code in Science
+title: "WORCS: A Workflow for Open Reproducible Code in Science"
 short:      "WORKFLOW FOR OPEN SCIENCE"
-abstract: >
+abstract: |
   Adopting open science
@@
   To facilitate widespread adoption, the
-  WORCS principles are implemented in the R package \pkg{worcs}, which offers an
+  WORCS principles are implemented in the R package `worcs`, which offers an
   RStudio 
@@
   standards for open science, and provides examples of the implementation of
-  \pkg{worcs} in \proglang{R}. All source code for the \proglang{R} package,
+  `worcs` in `R`. All source code for the `R` package,
   and replication materials for the manuscript, are available at 
@@
 urlcolor          : blue
-output: rticles::jss_article
+output: rmarkdown::html_document
 knit: worcs::cite_all
@@
 towards open science.
-Several immediate causes rallied support for this transition, including
-highly publicized cases of scientific fraud [as discussed by
+Support for this transition was first motivated by several
+highly publicized cases of scientific fraud [
 @leveltFailingScienceFraudulent2012], increasing awareness of questionable
-research practices and their consequences for the credibility of findings [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
+research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
 [@shroutPsychologyScienceKnowledge2018].
@@
 Capitalizing on these advances has the potential to accelerate scientific
-progress [see also @coyneReplicationInitiativesWill2016].
+progress [see also @coyneReplicationInitiativesWill2016],
+as has been aptly demonstrated by the role of open science in the response to the coronavirus pandemic [@sondervanCOVID19PandemicStresses2020].
 
 Many researchers are motivated to adopt current best practices for open
-science and enjoy these benefits. And yet, the question of "how" to 
-apply open science practices can be daunting. Making the transition
+science and enjoy these benefits. And yet, the question of how to 
+get started can be daunting. Making the transition
 requires researchers to become knowledgeable about different open science 
@@
 WORCS is a lightweight approach to open science and
-computational reproducibility for scholars working in \proglang{R} on a private computer,
+computational reproducibility for scholars working in `R` on a private computer,
 conducting research that involves academic prose, analysis code,
 and (optionally) tabular data - but it is flexible enough to accommodate many
-other scenarios. The workflow (Figure \@ref(fig:workflow)) 
+other scenarios. The workflow (Figure 1) 
 constitutes a step-by-step procedure that researchers can follow to make a 
-research project (including data, code, and a report) open and reproducible.
+research project open and reproducible.
 WORCS is compatible with open science
 requirements already implemented by journals and institutions, and will help 
-fulfill them. It can also be used in the absence of top-down support or guidelines
-to produce work in accordance with best practices.
-Although the principles underlying this workflow are universal, WORCS has first
-been implemented for \proglang{R} users [@rcoreteamLanguageEnvironmentStatistical2020].
-The \proglang{R} package \pkg{worcs} [@worcspackage2020] offers a project template for RStudio
+fulfill them.
+It can also be used to produce work in accordance with best practices
+in the absence of top-down support or guidelines.
+Although the principles underlying this workflow are universal, WORCS has been
+implemented in `R` [@rcoreteamLanguageEnvironmentStatistical2020].
+The `worcs` package [@worcspackage2020] offers a project template for RStudio
 [@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
-a "one click solution". This paper introduces the WORCS workflow, discusses how
-it meets best practices for open science, and illustrates the use of the \pkg{worcs}
-package.
-
-```{r workflow, fig.cap="Schematic illustration of the WORCS procedure.", out.width="100%"}
-#knitr::include_graphics("workflow_graph/workflow.png")
-knitr::include_graphics("workflow_graph/workflow.svg")
+a "one click solution". This paper discusses the principles underlying WORCS,
+introduces the workflow, addresses how WORCS meets best practices for open science,
+and illustrates the use of the `worcs` package.
+
+```{r workflow, fig.cap="*Figure 1*: Schematic illustration of the WORCS procedure.", out.width="100%"}
+knitr::include_graphics("workflow.png")
+#knitr::include_graphics("workflow_graph/workflow.svg")
 ```
@@
 Although Open Science is advocated by many, it does not have a unitary definition
-[for an effort to ground open science in first principles, see 
-@tennantValuePropositionOpen2020].
-Instead, several guidelines exist for practices that are considered "open".
-The TOP-guidelines are one of the most influential operationalisations
-of Open Science principles [@nosekPromotingOpenResearch2015a].
+[but see @tennantValuePropositionOpen2020].
+One of the most influential operationalisations
+of Open Science principles is known as the "TOP-guidelines" [@nosekPromotingOpenResearch2015a].
 These guidelines describe eight standards for open science: 1) Comprehensive
@@
 analyses, and then manually copy-pasting analysis results into the text.
-The source code of this paper can serve as an example of DDG, and is
-[available here](https://github.com/cjvanlissa/worcs/blob/master/paper/Manuscript.Rmd).
-
-Although there is a slight learning curve to transitioning to DDG, we strongly
-believe that the investment will pay off for users of the workflow, as it did for us.
+This paper is an example of DDG, and its source code is
+[available here](https://github.com/cjvanlissa/worcs/blob/master/paper/Manuscript_DS.Rmd).
+
+Although transitioning to DDG involves a slight learning curve, we strongly
+believe that the investment will pay off.
 Time saved from painstakingly copy-pasting output
@@
 When revisions require major changes to the analyses, all results, figures and
-tables can be automatically updated. The flexibility in output formats also 
+tables are automatically updated. The flexibility in output formats also 
 means that a manuscript can be rendered to presentation format, or even to a
-website or blog post. Moreover, the fact that code can be run each time the
-document is compiled encourages researchers to work reproducibly,
+website or blog post. Moreover, the fact that code is evaluated each time the
+document is compiled requires researchers to work reproducibly,
 and allows reviewers and/or readers verify reproducibility simply by re-compiling
 the document.
-In sum, while writing academic papers in a programming environment might seem
+While writing academic papers in a programming environment might seem
 counter-intuitive at first, this approach is much more amenable to the needs of
@@
 
-In the \proglang{R} implementation of WORCS, we recommend centering a research project
-around one dynamically generated \code{RMarkdown} document, which includes all 
+In the `R` implementation of WORCS, we recommend centering a research project
+around one dynamically generated `RMarkdown` document, which includes all 
 analysis code. Longer scripts can be
-stored in \code{.R} files, and called from the main document using the \code{source()}
+stored in `.R` files, and called from the main document using the `source()`
 function. When a reader or reviewer compiles this document, all code is run
-automatically, thus verifying computational replicability. The \code{RMarkdown} 
+automatically, thus verifying computational replicability. The `RMarkdown` 
 document can be automatically formatted in many styles, including
-APA style [thanks to the \proglang{R} package \pkg{papaja}, @austPapajaPrepareReproducible2020],
+APA style [thanks to the `R` package `papaja`, @austPapajaPrepareReproducible2020],
 a host of other scientific formats 
-[see @allaireRticlesArticleFormats2020], and as plain \code{Markdown}.
+[see @allaireRticlesArticleFormats2020], and as plain `Markdown`.
 
@@
 privacy sensitive data.
-A \code{.gitignore} file lists files that should not be version-controlled.
+A `.gitignore` file lists files that should not be version-controlled.
 These files thus exist only on the user's private computer.
 
-The functionality of Git is amplified by services such as GitHub
+The functionality of Git is enhanced by services such as GitHub
 (https://github.com). GitHub is best understood as a cloud storage service
 with social networking functionality. The cloud storage aspect
-of GitHub works as follows: You can "clone" (copy) a local Git repository to
+of GitHub allows you to "clone" (copy) a local Git repository to
 the GitHub website, as a backup or research archive. The social network aspect
@@
 
-It is worth noting that,
-although this paper focuses on the most widely used Git
-remote repository service, GitHub, two notable alternatives exist: 
-GitLab (https://gitlab.com) and Bitbucket (https://bitbucket.org).
-In 
-order to present the workflow in a platform-agnostic way, we will refer to such
-cloud hosting services as "remote repositories" in the remainder of this
-paper.
-As of
-version 0.1.2, \pkg{worcs} has been tested with these three remote repository services,
-but has been optimized for use with GitHub.
-A [vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) is included
-with instructions for using alternative platforms. 
-When initializing a new project using the \pkg{worcs} project template,
-a 'https' remote repository address will usually work best with the authentication procedure outlined in the
-[setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
-
-The \proglang{R} implementation of WORCS uses the package \code{gert}, a Git client for \proglang{R}
-[@oomsGertSimpleGit2019], to connect a local project to a remote repository.
-The \pkg{worcs} package also contains a user-friendly function to set
-the default user credentials, \code{git_user()}, and a function to add, commit, and
-push all changed files to a remote repository in one step, \code{git_update()}.
+Although GitHub is the most widely used
+remote repository, `worcs` [supports alternative services](https://cjvanlissa.github.io/worcs/articles/git_cloud.html).
+We will refer to all such services as "remote repositories".
+The `worcs` package uses `gert` [@oomsGertSimpleGit2019] to connect a local project to a remote repository. It also includes user-friendly functions to set
+user credentials, `git_user()`, and to add, commit, and
+push changed files to a remote repository in one step, `git_update()`.
 
@@
 sufficient to state that analyses were conducted in *Program X*. However, every
-program is susceptible to changes, updates, and bugfixes. A lot of Open Source software,
+program is susceptible to changes, updates, and bugfixes. Open Source software,
 in particular, is regularly updated because there is an
@@
 
-Many solutions exist to ensure computational reproducibility. These solutions typically work by
+Many solutions exist to ensure computational reproducibility, which differ in
+user-friendliness and effectiveness. 
+These solutions typically work by
 enveloping your research project in a distinct "environment" that only has access to programs that
-are explicitly installed, and maintaining a record of these programs. These solutions differ in
-user-friendliness and effectiveness. If strict reproducibility is required, users might choose to conduct all analyses in a "Docker container": An environment that behaves like a
-virtual computer, which can be stored like a sort of time capsule, and
+are explicitly installed, and maintaining a record of these programs.
+If strict reproducibility is required, users might choose to conduct all analyses in a "Docker container": An environment that behaves like a
+virtual computer, which can be stored like a time capsule, and
 identically reinstated on a user computer, or in the cloud.
-This is effective, and also preserves software outside the \proglang{R} environment (e.g.,
+This is effective, and also preserves software outside the `R` environment (e.g.,
 preprocessing tools for neuroimaging data). A Docker-based workflow
-was recently introduced by @peikertReproducibleDataAnalysis2019. This workflow
-is based on many of the same principles as WORCS, and the two approaches are
-conceptually and practically compatible. However, the workflow described by 
-Peikert and Brandmaier can be difficult to set up for novice users. 
-The accompanying \proglang{R} package \pkg{repro}
-is currently in development to facilitate this transition [@repro]. Another 
+that is conceptually and practically compatible with WORCS
+was recently introduced by @peikertReproducibleDataAnalysis2019.
+However, it is somewhat difficult to set up for novice users. 
+The `R` package `repro`
+is currently in development to facilitate adoption [@repro]. 
+Another 
 user-friendly Docker-based workflow is offered by the cloud-based collaboration
@@
 data are stored.
-The necessity to upload data for analysis is a risk with any
-cloud-computing solution when human participants are involved.
-
-For analyses conducted entirely in \proglang{R}, the recently released package \pkg{renv}
+The necessity to upload human participant data is a liability
+when using any
+cloud-computing solution for data analysis.
+
+For analyses conducted entirely in `R`, the recently released package `renv`
 strikes a great balance between user-friendliness  and effectiveness. Developed by
@@
 lightweight.
-The text-based log file can be version controlled with Git. \pkg{renv}
+The text-based log file can be version controlled with Git. `renv`
 installs the logged packages in a cache that is shared between all
@@
 that is already installed on the system. When someone else loads your 
-project, \pkg{renv} will install all of the required packages onto a cache on their
+project, `renv` will install all of the required packages onto a cache on their
 computer. 
@@
 [@brownHowLearnedStop2017].
-Ultimately, the best approach might be to use a "just good-enough" solution, and
+Ultimately, the best approach might be to use a "good-enough" solution, and
 acknowledge that all code requires some maintenance if you want to
-reproduce it in the future. For projects in \proglang{R}, \pkg{renv} meets these requirements.
+reproduce it in the future. For projects in `R`, `renv` meets these requirements.
 
@@
 A key consideration when developing a research project is what filetypes to use.
-A case can be made to use text-based files exclusively (or as much as possible),
+WORCS encourages the use of text-based files whenever possible,
 instead of binary files.
 Text-based files can be read by machines and humans alike.
-Binary files, such as Word (\code{.docx}) or SPSS (\code{.sav}, \code{.spo}) files,
-must be decoded first, and many require commercial software to do so.
-The version control software Git is also designed for use with text-based files.
-The Git change log reflects changes to text-based files on a line-by-line basis.
-Binary files can be version controlled with Git, but 
-the change log for a binary file are uninterpretable.
+Binary files, such as Word (`.docx`) or SPSS (`.sav`, `.spo`) files,
+must be decoded first, often requiring commercial software.
+Git is designed for use with text-based files, as it
+logs changes to text-based files on a line-by-line basis.
+Although binary files can be version controlled with Git, 
+their change log is uninterpretable.
 Binary files are also often 
@@
 files to remote repositories is frowned upon.
-WORCS encourages the use of
-text-based files for all purposes.
 
 Two additional points are worth noting:
-First, Git tracks line-by-line changes to text-based files,
-based on line breaks in the text.
-When writing a version-controlled paper in plain-text,
-it is therefore recommended to start each sentence on a new line.
+First, as Git tracks line-by-line changes,
+it is recommended to start each sentence on a new line.
 That way, the change log will indicate which specific sentence was edited, 
@@
 this practice does not disrupt the flow of a paragraph in the rendered version.
-Second, the remote repository service GitHub rewards the use of text-based files by
-rendering certain text-based filetypes for online viewing: 
-For example, \code{.md}
-(Markdown) files are displayed as web pages, and \code{.csv} files as spreadsheets.
-The \pkg{worcs} implementation for \proglang{R} makes use of this fact by, for example,
-rendering a codebook for the data as \code{.md} so visitors of a Git remote repository
-can view it as a web page.
+Second, the remote repository GitHub
+renders certain text-based filetypes for online viewing: 
+For example, `.md`
+(Markdown) files are displayed as web pages, and `.csv` files as spreadsheets.
+The `worcs` package uses this feature to, for example,
+render a codebook for the data as `.md` so it will be shown
+as a web page on GitHub.
 
@@
 Version 0.1.2. 
-For users of the \pkg{worcs} \proglang{R} package, we highly recommend referring to
+For users of the `worcs` `R` package, we highly recommend referring to
 [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
@@
 
-The \proglang{R} package \pkg{worcs} automates most steps of the workflow
-for \proglang{R} users.
-Several arguments support the choice to implement this workflow in \proglang{R}.
-First, \proglang{R} and all of it extensions are free open source software, which
+The `R` package `worcs` automates much of the workflow
+for `R` users.
+Several arguments support the choice to implement this workflow in `R`.
+First, `R` and all of it extensions are free open source software, which
 make it a tool of choice for open science.
-Second, all tools required for an open science workflow are implemented in \proglang{R},
-most of these tools are directly accessible through the user interface of RStudio,
+Second, all tools required for an open science workflow are implemented in `R`,
+most of these tools are directly accessible through the RStudio user interface,
 and some of them are actively developed by the team behind RStudio.
-Third, \proglang{R} is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
-Fourth, \proglang{R} is well-supported;
-a vibrant and inclusive online community exists for \proglang{R},
-developing new \proglang{R} methods and packages,
-and providing support and tutorials for existing ones.
-Finally, \proglang{R} is highly interoperable:
-Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
-Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from \proglang{R}, including \proglang{C++}, 
-\proglang{Fortran}, and \proglang{Python} [@allaireMarkdownPythonEngine2020]. There are excellent free
-resources for learning \proglang{R} [e.g., @grolemundDataScience2017].
-
-Working with \proglang{R} is simplified immensely by using the RStudio integrated
-development engine (IDE) for \proglang{R} [@rstudioteamRStudioIntegratedDevelopment2015].
+Third, `R` is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
+Fourth, `R` is well-supported by
+a vibrant and inclusive online community,
+which develops new methods and packages,
+and provides support and tutorials for existing ones.
+Finally, `R` is highly interoperable:
+Packages are available to read and write nearly every filetype,
+including DOCX, PDF (in APA style), and HTML. 
+Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from `R`, including `C++`, 
+`Fortran`, and `Python` [@allaireMarkdownPythonEngine2020]. There are excellent free
+resources for learning `R` [e.g., @grolemundDataScience2017].
+
+Working with `R` is simplified immensely by using the RStudio integrated
+development engine (IDE) for `R` [@rstudioteamRStudioIntegratedDevelopment2015].
 RStudio automates and streamlines tedious or complicated aspects of working with
@@
 self-contained folder, that can be uploaded entirely to a Git remote repository and downloaded
-by future users. The \pkg{worcs} \proglang{R} package installs a new RStudio project template.
+by future users. The `worcs` `R` package installs a new RStudio project template.
 When a new project is initialized from this template, the bookkeeping required
@@
 
-Before you can use the \proglang{R} implementation of the WORCS workflow, you have to install the required software.
-You only have to perform this setup procedure only once for each computer you
-intend to use \pkg{worcs} on. It should take about 30 minutes, and it is documented
+Before you can use the `R` implementation of the WORCS workflow,
+you have to install the required software.
+This 30 minute procedure is documented
 in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
-The setup vignette does not assume a prior installation of \proglang{R}, so it is suitable for novice users.
-After setting up your system, you can use \pkg{worcs} for all of your
-projects. The \proglang{R} implementation of the workflow introduced earlier is detailed,
+After setting up your system, you can use `worcs` for all of your
+projects. The `R` implementation of the workflow introduced earlier is detailed,
 step-by-step, in
@@
 References for software are sometimes provided within the software environment;
-for example, in \proglang{R}, package references can be found by
-calling \code{citation("packagename")}. This returns an APA-style reference, and a
+for example, in `R`, package references can be found by
+calling `citation("packagename")`. This returns an APA-style reference, and a
 BibTeX entry. Software papers are also common, and are sometimes called
@@
 
-The \pkg{worcs} package includes a [vignette on citation](https://cjvanlissa.github.io/worcs/articles/citing.html) to explain the
-fundamentals for novice users, and offer recommendations specific to \pkg{worcs}. We
+The `worcs` package includes a [vignette on citation](https://cjvanlissa.github.io/worcs/articles/citing.html) to explain the
+fundamentals for novice users, and offer recommendations specific to `worcs`. We
 recommend using the free, open-source reference manager
@@
 
-The \pkg{worcs} package also offers original solutions for comprehensive citation.
+The `worcs` package also offers original solutions for comprehensive citation.
 Firstly,
@@
 going back to cut non-essential references just prior to publication. Standard
-Markdown uses the at-symbol (\code{@}) to cite a reference. \pkg{worcs} additionally 
-reserves the "double at"-symbol (\code{@}) to cite a non-essential reference.
+Markdown uses the at-symbol (`@`) to cite a reference. `worcs` additionally 
+reserves the "double at"-symbol (`@`) to cite a non-essential reference.
 Users can render the manuscript either with, or without, comprehensive citations
-by adapting the \code{knit} command in the front matter, setting it to
-\code{knit: worcs::cite_all} to render all citations, and to 
-\code{knit: worcs::cite_essential} to remove all *non-essential* citations.
-
-With regard to the citation of \proglang{R} packages, it is worth noting that the \pkg{papaja} 
-package [@austPapajaPrepareReproducible2020], which provides the \code{APA6}
-template used in \pkg{worcs}, also
-includes functions to automatically cite all packages used in the \proglang{R} session, and
+by adapting the `knit` command in the front matter, setting it to
+`knit: worcs::cite_all` to render all citations, and to 
+`knit: worcs::cite_essential` to remove all *non-essential* citations.
+
+With regard to the citation of `R` packages, it is worth noting that the `papaja` 
+package [@austPapajaPrepareReproducible2020], which provides the `APA6`
+template used in `worcs`, also
+includes functions to automatically cite all packages used in the `R` session, and
 add their references to the bibliography file.
@@
 the authors who evaluate it on the real data and send back the results. The
-\pkg{worcs} package offers a simple but flexible function to generate synthetic data,
-\code{synthetic()}, which by default uses random forests to generate a model for
+`worcs` package offers a simple but flexible function to generate synthetic data,
+`synthetic()`, which by default uses random forests to generate a model for
 synthesis (see @wrightRangerFastImplementation2015). For data synthesis to work
-properly, it is essential that the \code{class} (data type) of variables is defined
-correctly. As of \pkg{worcs} version \code{r worcs:::cran_version()}, numeric, integer, factor, and logical data
+properly, it is essential that the `class` (data type) of variables is defined
+correctly. As of `worcs` version `r worcs:::cran_version()`, numeric, integer, factor, and logical data
 are supported out of the box. Other types of variables should be converted to
-one of these types, or users can use a custom \code{model_expression} when calling
-\code{synthetic}.
+one of these types, or users can use a custom `model_expression` when calling
+`synthetic`.
 
@@
 
-When initializing a new \pkg{worcs} project, an empty \proglang{R} script called \code{prepare_data.R} is generated.
+When initializing a new `worcs` project, an empty `R` script called `prepare_data.R` is generated.
 As soon as raw data are collected, researchers should use this file to document
 all steps necessary to load the 
-data into \proglang{R}, pseudonimize it, and prepare it for analysis.
+data into `R`, pseudonimize it, and prepare it for analysis.
 As data should be shared in a format as close to raw as possible, 
@@
 a researcher discovers that in one of the variables, missing values were coded as
-\code{-99} whereas they were coded as \code{-9} in all other variables, then we would recommend adding
+`-99` whereas they were coded as `-9` in all other variables, then we would recommend adding
 a line of code to recode the missing values - thereby documenting the mistake - 
@@
 
-The \pkg{worcs} package offers two functions for version controlling a record of the
+The `worcs` package offers two functions for version controlling a record of the
 data: One for open, and one for closed data. Researchers should assume that the
@@
 
-In \pkg{worcs}, researchers can call the function \code{open_data(data)} to make the data
-publicly available. This function stores the \code{data} object (such as a \code{data.frame} or
-\code{matrix}) in a tabular data file (called \code{data.csv} by default).
+In `worcs`, researchers can call the function `open_data(data)` to make the data
+publicly available. This function stores the `data` object (such as a `data.frame` or
+`matrix`) in a tabular data file (called `data.csv` by default).
 It also generates an
@@
 pushed to the remote repository.
-This can be done through the Git panel in RStudio, or by running the \pkg{worcs}
-function \code{git_update("commit message")}.
+This can be done through the Git panel in RStudio, or by running the `worcs`
+function `git_update("commit message")`.
 Once the remote repository is made public, these data will be
@@
 Alternatively, if the project requires data to remain closed, researchers can
-call \code{closed_data(data)}. This function also stores the data in a local \code{.csv} file
+call `closed_data(data)`. This function also stores the data in a local `.csv` file
 and generates a codebook,
-but the original data file is added to \code{.gitignore} so it cannot be
+but the original data file is added to `.gitignore` so it cannot be
 accidentally added to the Git repository.
 The function also computes a checksum for the original data, and logs it in the
-\code{.worcs} project file. This checksum serves as the indelible record of the state 
+`.worcs` project file. This checksum serves as the indelible record of the state 
 of the raw data.
-The function \code{closed_data()} also creates a synthetic dataset (called
-\code{synthetic_data.csv} by default).
+The function `closed_data()` also creates a synthetic dataset (called
+`synthetic_data.csv` by default).
 The user should push all changed files to the remote repository.
@@
 
-As the purpose of the \code{prepare_data.R} file is to prepare data for sharing,
+As the purpose of the `prepare_data.R` file is to prepare data for sharing,
 the final line of this file
 should typically be either
-\code{open_data()} or \code{closed_data()}.
-
-After generating tidy, shareable data, this data is then loaded in the analysis code with the function \code{load_data()}.
+`open_data()` or `closed_data()`.
+
+After generating tidy, shareable data, this data is then loaded in the analysis code with the function `load_data()`.
 This function only loads the real data if it is available on the user's computer, 
@@
 the original data.
-They can simply start the script with \code{load_data()}, and
+They can simply start the script with `load_data()`, and
 write their code based on the synthetic data.
@@
 ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request).
-When you run their code on your system, \code{load_data()} will load the
+When you run their code on your system, `load_data()` will load the
 original data, and use that to run their script. You can then return the results
@@
 Note that these functions can be used to store multiple data files, if necessary.
-As of version 0.1.2, the \pkg{worcs} package only stores data as \code{.csv} files.
-As explained above, this is recommended, because \code{.csv} files are human- and
+As of version 0.1.2, the `worcs` package only stores data as `.csv` files.
+As explained above, this is recommended, because `.csv` files are human- and
 machine readable, and because data typically need to be in tabular format
@@
 row per measurement occasion and a column per channel.
-The \code{prepare_data.R} file should document any
+The `prepare_data.R` file should document any
 steps required to convert these data into tabular format. 
 If this is not an appropriate format for the data, readers are encouraged to follow the
-development of the \pkg{repro} package [@repro], which will offer greater support for
+development of the `repro` package [@repro], which will offer greater support for
 reproducibly storing and retrieving diverse data formats.
@@
 Some users may intend to make their data openly available through a restricted access platform, such as an institutional repository, but not publicly through a Git remote repository.
-In this case, it is recommended to use \code{closed_data()}, and to manually upload the original \code{data.csv} file to the restricted access platform.
+In this case, it is recommended to use `closed_data()`, and to manually upload the original `data.csv` file to the restricted access platform.
 If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the Git remote repository as an [Add-on](https://help.osf.io/hc/en-us/categories/360001550973-Add-ons).
@@
 methods for simulating data are available in most statistical programming languages.
-For instance, \proglang{R} users can use the package \code{simstudy} [@goldfeldSimstudySimulationStudy2020].
+For instance, `R` users can use the package `simstudy` [@goldfeldSimstudySimulationStudy2020].
 
@@
 
-The \pkg{worcs} workflow facilitates preregistration,
+The `worcs` workflow facilitates preregistration,
 primarily by explicitly inviting a user to write the preregistration document.
-To this end, \pkg{worcs} imports several preregistration templates from the \proglang{R} package
-\code{prereg} [@austPreregMarkdownTemplates2019], including templates from
+To this end, `worcs` imports several preregistration templates from the `R` package
+`prereg` [@austPreregMarkdownTemplates2019], including templates from
 organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and
 from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
-When initializing an RStudio project with the \pkg{worcs} project template, one of
+When initializing an RStudio project with the `worcs` project template, one of
 these preregistration templates can be selected, which will generate a file
-called \code{preregistration.Rmd}. This file should be used to document study plans,
+called `preregistration.Rmd`. This file should be used to document study plans,
 ideally prior to data collection. 
@@
 including data and code - accessible in a repository that can be made public. 
-The \pkg{worcs} package additionally
-facilitates meeting the FAIR principles in several ways. First, each \pkg{worcs}
+The `worcs` package additionally
+facilitates meeting the FAIR principles in several ways. First, each `worcs`
 project has a YAML file (https://yaml.org/) that lists the research objects in
 the project, which will allow web crawlers 
-to index \pkg{worcs} projects. Second, a codebook is automatically generated for
+to index `worcs` projects. Second, a codebook is automatically generated for
 each data file stored
-using \code{open_data()} or \code{closed_data()}. We encourage users to elaborate on the
+using `open_data()` or `closed_data()`. We encourage users to elaborate on the
 default codebook by adding
@@
 
-Incidentally, \pkg{worcs} itself follows the recommendations for [FAIR research software](https://fair-software.eu/). It is 
+Incidentally, `worcs` itself follows the recommendations for [FAIR research software](https://fair-software.eu/). It is 
 hosted on a version controlled public remote repository GitHub, has an open
 source licence (GPL v3.0), is registered in a community registry (CRAN),
-enables the citation of the software (using the \code{citation("worcs")} command, or
+enables the citation of the software (using the `citation("worcs")` command, or
 by citing this paper), and followed the [*CII Best Practices*](https://bestpractices.coreinfrastructure.org/en)
@@
 The workflow aims to lower the threshold for grass-roots adoption of open science principles.
-The workflow is supported by an \proglang{R} package with an RStudio project template and 
+The workflow is supported by an `R` package with an RStudio project template and 
 convenience functions.
@@
 items contributing to an open and reproducible research project. Users of the
-\pkg{worcs} package can receive a badge, displayed on the \code{README.md} file,
+`worcs` package can receive a badge, displayed on the `README.md` file,
 based on a semi-automated scoring of the checklist items, by calling the function
-\code{check_worcs()} within a \pkg{worcs} project directory.
+`check_worcs()` within a `worcs` project directory.
 
@@
 sets WORCS apart is that is an easy and sufficient solution for projects that
-can be conducted entirely within \proglang{R}, resulting in a much more lightweight
+can be conducted entirely within `R`, resulting in a much more lightweight
 approach to safeguard computational reproducibility under *most circumstances*.
@@
 issues for future development remain. One potential challenge is the learning
-curve associated with the tools outlined in this paper. Learning to work with \proglang{R},
-RMarkdown, and Git requires an initial time investment. The amount of
-time required is reduced by the availability of tutorials, such as this one, and
-other tutorials cited throughout this document. Moreover, the time investment
-tends to pay off. Working with \proglang{R} opens the door to many cutting edge analysis
+curve associated with the tools outlined in this paper. Learning to work with `R`,
+RMarkdown, and Git takes time - although tutorials such as this one substantially reduce the learning curve.
+Moreover, the time investment
+tends to pay off. Working with `R` opens the door to many cutting edge analysis
 techniques. Working with RMarkdown saves time and prevents mistakes by avoiding
@@
 uses RMarkdown, and the coauthors use Word. In this case, it is possible to Knit
-the manuscript to Word (.docx), by changing the line \code{output: papaja::apa6_pdf}
-in the manuscript's YAML header to \code{output: papaja::apa6_docx}. There are some
+the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf`
+in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some
 limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
@@
 comment bubbles in Word. <!-- LB: Overleaf?-->Then, manually incorporate their
-changes into the \code{manuscript.Rmd} file. In most cases, this is the most
+changes into the `manuscript.Rmd` file. In most cases, this is the most
 user-friendly approach, as most lead authors would review changes by co-authors
 anyway. A second approach is to ask collaborators to work in plain text. In this
-case, send collaborators the \code{manuscript.Rmd} file, and ask them to open (and
+case, send collaborators the `manuscript.Rmd` file, and ask them to open (and
 save) it in Word or Notepad as a plain text file. When they send it back, make
@@
 and click the Diff button to examine what changes the collaborator has made
-relative to your last committed version. Finally,
-@peikertReproducibleDataAnalysis2019 mention the [\code{redoc} package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] as a
+relative to your last committed version. <!--Finally,
+@peikertReproducibleDataAnalysis2019 mention the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] as a
 potentially useful way to collaborate with Word users. This package enables
 bidirectional conversion between Word and
-RMarkdown, with support for tracked changes. However, \code{redoc} is currently in
+RMarkdown, with support for tracked changes. However, `redoc` is currently in
 suspended development, and for many users, a simple solution (i.e.,
-asking collaborators to work in plain text) might be sufficient.
+asking collaborators to work in plain text) might be sufficient.-->
 
@@
 in scientific software development as well; for example, the lead author of this
-paper became a co-author on several \proglang{R} packages after submitting pull requests
+paper became a co-author on several `R` packages after submitting pull requests
 with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018], and two co-authors of this
-paper became involved by contributing pull requests to \pkg{worcs}. It is also
+paper became involved by contributing pull requests to `worcs`. It is also
 possible to invite such collaboration by 
@@
 Nevertheless, there are clear directions for future developments.
-Firstly, although the workflow is currently implemented only in \proglang{R}, it is 
+Firstly, although the workflow is currently implemented only in `R`, it is 
 conceptually relevant for users of other statistical programming languages, most
-notably \proglang{Python} [@10.5555/1593511].
+notably `Python` [@10.5555/1593511].
 We welcome efforts to implement WORCS in other platforms.
-Secondly, even when a project is in \proglang{R}, it may have dependencies outside of the
-\proglang{R} environment that cannot be managed using \pkg{renv}.
-It is beyond the scope of the \pkg{worcs} package to support the incorporation of
-tools outside of \proglang{R}, or to containerize a project so that it can
+Secondly, even when a project is in `R`, it may have dependencies outside of the
+`R` environment that cannot be managed using `renv`.
+It is beyond the scope of the `worcs` package to support tools outside of `R`,
+or to containerize a project so that it can
 be identically reinstated on a different system or virtual machine.
-Therefore, another
-important area of future development is to enable such extensions of the
-workflow. To this end, we are collaborating with @repro on the development of
-\pkg{repro}, an \proglang{R} package that will facilitate setting up reproducibility tools such
-as Make and Docker.
-This will make it possible for researchers
-to enhance a WORCS project with these strict reproducibility tools,
-or vice versa, to initiate a WORCS project within a strictly reproducible
-environment generated using \pkg{repro}. 
-This will combine the strengths of WORCS with those of the workflow proposed
+The forthcoming package `repro` [@repro] will offer such extensions of the
+workflow, thus combining the strengths of WORCS with the solutions proposed
 by @peikertReproducibleDataAnalysis2019.
@@
 The step-by-step procedure outlined in this tutorial helps researchers make an entire research project Open and Reproducible.
-The accompanying \proglang{R} package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.
+The accompanying `R` package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.
 
@@
 
-\newpage
+# References

### 2020-09-29 tests/testthat/test-citations.R
Final commit before CRAN submission

@@ -89,16 +89,16 @@ myiris <- read.csv('iris.csv')
   # Render with all citations
   sink_out <- capture_output(cite_all(file_name))
 
-  contents <- readLines(gsub("Rmd$", "md", file_name))
+  contents <- readLines(normalizePath(gsub("\\.Rmd", "\\.md", file_name)))
 
   expect_true(any(contents == "Optional reference: @reference2020"))
 
   sink_out <- capture_output(cite_essential(file_name))
 
-  contents <- readLines(gsub("Rmd$", "md", file_name))
+  contents <- readLines(normalizePath(gsub("\\.Rmd", "\\.md", file_name)))
 
   expect_true(!any(contents == "Optional reference: @reference2020"))
-  expect_true(any(startsWith(contents[25:length(contents)], "myiris")))
+  expect_true(any(grepl("myiris", contents, fixed = TRUE)))
   expect_true(any(contents == "Optional reference:"))
 })
 

### 2020-10-08 inst/rstudio/templates/project/resources/pss.Rmd
Version 5

---
title           : "Title here"
shorttitle      : "Short title here"
date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
authors         : "Author names here"
---

<!-- PSS (Preregistration and Sharing Software) is a template that can be used
for pre-registrating confirmatory or exploratory studies.
For more details about the description please check the reference:

Krypotos, A. M., Klugkist, I., Mertens, G., & Engelhard, I. M. (2019).
A step-by-step guide on preregistration and effective data sharing for
psychopathology research. Journal of Abnormal Psychology, 128(6), 517-537.
https://psycnet.apa.org/doi/10.1037/abn0000424 -->

# Title

<!-- Here you can provide the project title. -->

## Authors

<!-- Here you can provide the authors' list. -->

## Affiliations

<!--  Authors' affiliations -->

# Background of the study

## Primary study/Secondary analyses

<!-- Is it a primary study (where data are collected) or a study 
where secondary analyses are performed on an existing data set? -->

## Does the study refer to a meta-analysis?

<!-- Mention whether the study refers to a meta-analysis.
In this case, you can ignore the irrelevant parts below
(e.g., number of participants). -->

## Research questions

<!-- Define the research question -->

## Study hypotheses

<!-- Define the study hypotheses -->

# Method

## Stimuli

<!-- Define the study stimuli -->

## Questionnaires

<!-- Define the study questionnaires, if used -->

## Equipment

<!-- Define the study equipment. This includes computer characteristics etc. -->

## Procedure

<!-- Define the study's procedure. -->

## Protocol

<!-- Define the study's protocol. -->

# Statistical analyses

## Participants

<!-- How many participants will be recruited? 
Has a power analysis been performed? -->

## Stopping rule

<!--  When will data accumulation end? -->

## Confirming hypotheses threshold

<!-- What is the threshold for confirming each of the research hypotheses? -->

## Disconfirming hypotheses threshold

<!-- What is the threeshold for rejecting each of the research hypotheses? -->

# Other

## Other (Optional)

<!-- Further comments -->

# References

<!-- Included references -->

## 
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
\noindent

### 2020-10-08 inst/rstudio/templates/project/resources/secondary.Rmd
Version 5

---
title           : "Title here"
shorttitle      : "Short title here"
date            : "date here"
authors           : "authors here"
---

<!-- This template aims provides a relative short format to preregister
secondary analyses on preexisting data
(much like the AsPredicted.org format for new data collection).
For more details about the description please check the reference:

Mertens, G., & Krypotos, A. M. (2019). Preregistration of analyses of
preexisting data. Psychologica Belgica, 59(1), 338-352. doi: 10.5334/pb.493 -->

# Title

<!-- Here you can provide the project title.  -->

## Authors

<!-- Here you can provide the authors' list, together with the affiliations. -->

## Affiliations

<!--  Authors' affiliations-->

# Study hypotheses

<!-- Provide a brief description of the relevant theory 
and formulate the hypotheses as precisely as possible. -->

# Operationalization

<!-- State exactly how the variables specified in each hypothesis 
will be operationalized.  -->

# Data source

<!-- Specify the source of the obtained data. Also provide 
information about the context of the data source and clarify 
whether the data has been previously published. -->

# Data request/access

<!-- Specify how the data will be requested or accessed. 
Clarify whether the data were already available and whether 
the dataset has been previously explored or analyzed. -->

# Exclusion criteria

<!-- Specify whether there were any criteria for the 
exclusions of certain datasets, observations or time points.-->

# Statistical analyses

<!-- Specify the statistical model that will be used to analyze the data. 
Be as specific as possible and avoid ambiguity. -->

# Hypotheses (dis-)confirmation

<!-- Specify exactly how the hypothesis will be evaluated. 
Give specific criteria relevant to the used analytical 
model and framework (e.g., alpha-values, Bayes Factor, RMSEA).-->

# Analysis validation

<!-- Indicate whether the proposed analyses have previously been 
validated on a subset of the data, or a simulated dataset. 
If so, provide the data files and analysis code.-->

# Timeline

<!-- Provide the (foreseen) dates for the different steps 
in this preregistration form.-->

# What is known about the data that could be relevant for the tested hypotheses?

<!-- What is known about the data that could be relevant for the tested
hypotheses? I.e., disclose any prior exposure to the data set, direct or
indirect, and specific information (e.g., knowing the mean of a variable) that
is relevant for your research question. -->

## Other (Optional)

<!-- Further comments -->

# References

<!-- Included references -->

## 
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
\noindent

### 2020-10-08 R/worcs_project.R

@@ -129,18 +129,9 @@ worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregist
 
   # Begin prereg
   if(!preregistration == "none"){
-    tryCatch({
-      draft(
-        file.path(path, "preregistration.Rmd"),
-        paste0(preregistration, "_prereg"),
-        package = "prereg",
-        create_dir = FALSE,
-        edit = FALSE
-      )
-      col_message("Creating preregistration files.")
-    }, error = function(e){
-      col_message("Creating preregistration files.", success = FALSE)
-    })
+    cl[[1L]] <- quote(worcs::add_preregistration)
+    names(cl)[which(names(cl) == "path")] <- "worcs_directory"
+    eval(cl, parent.frame())
   }
   # End prereg
 
@@ -489,4 +480,77 @@ add_manuscript <- function(worcs_directory = ".", manuscript = "APA6", remote_re
       col_message("Creating manuscript files.", success = FALSE)
     })
   # End manuscript
+}
+
+
+#' @title Add Rmarkdown preregistration
+#' @description Adds an Rmarkdown preregistration template to a 'worcs' project.
+#' @param worcs_directory Character, indicating the directory
+#' in which to create the manuscript files. Default: '.', which points to the
+#' current working directory.
+#' @param preregistration Character, indicating what template to use for the
+#' preregistration. Default: \code{"COS"}; use \code{"None"} to omit a
+#' preregistration. See Details for other available choices.
+#' @param verbose Logical. Whether or not to print messages to the console
+#' during project creation. Default: TRUE
+#' @param ... Additional arguments passed to and from functions.
+#' @return No return value. This function is called for its side effects.
+#' @details Available choices include the templates
+#' \code{"COS", "VantVeer", "Brandt", "AsPredicted"}, which are imported from the
+#' \code{\link[prereg]{cos_prereg}} package, and documented there. Furthermore,
+#' several unique templates are included with \code{worcs}:
+#' \itemize{
+#'   \item{\code{"PSS"}}{Preregistration and Sharing Software (Krypotos,
+#'   Klugkist, Mertens, & Engelhard, 2019)}
+#'   \item{\code{"Secondary"}}{Preregistrating for secondary analyses (Mertens &
+#'   Krypotos, 2019)}
+#' }
+#' @examples
+#' the_test <- "worcs_prereg"
+#' old_wd <- getwd()
[…]

### 2020-10-08 tests/testthat/test-add_preregistration.R

@@ -1,0 +1,24 @@
+test_that("preregistration can be generated", {
+  the_test <- "preregistration"
+  old_wd <- getwd()
+  test_dir <- file.path(tempdir(), the_test)
+  dir.create(test_dir)
+  setwd(test_dir)
+  file.create(file.path(tempdir(), the_test, ".worcs"))
+  on.exit(unlink(test_dir, recursive = TRUE), add = TRUE)
+
+  add_preregistration(
+    worcs_directory = ".",
+    preregistration = "vantveer"
+  )
+  expect_true(file.exists(".worcs"))
+  #expect_true(file.exists("manuscript/american-chemical-society.csl"))
+  expect_true(file.exists("preregistration.Rmd"))
+  file.remove("preregistration.Rmd")
+  add_preregistration(
+    worcs_directory = ".",
+    preregistration = "pss"
+  )
+  expect_true(file.exists("preregistration.Rmd"))
+  setwd(old_wd)
+})

### 2020-10-27 paper/reviews_ds.Rmd
Format reviewer comments

---
title       : "Manuscript title"
authors     : "First Author & Second Author"
journal     : "Journal title"
manuscript  : "Manuscript number"

class       : "draft"

output      : papaja::revision_letter_pdf
---

Dear Dr. Editor,

thank you very much for taking the time to consider our manuscript for publication at _`r rmarkdown::metadata$journal`_.
In the following we address your and each reviewers' concerns point-by-point.

# Reviewer \#1

Review #1 submitted on 19/Oct/2020 by Daniel Garijo ORCID logo https://orcid.org/0000-0003-0454-7145

* Overall Impression: Weak
* Suggested Decision: Reject
* Technical Quality of the paper: Good
* Presentation: Good
* Reviewer`s confidence: High
* Significance: High significance
* Background: Reasonable
* Novelty: Limited novelty
* Data availability: All used and produced data (if any) are FAIR and openly available in established data repositories
* Length of the manuscript: The length of this manuscript is about right

This paper describes worcs, a package and workflow to help make R code reproducible according to the TOP guidelines.

The paper describes a tutorial of the tool, and how it implements the TOP guidelines.

Reasons to accept:

- The paper is relevant to the journal and on a timely topic.
- The capabilities of the tool for synthetic data generation, connect to OSF, preprint publishing and dependency gathering show a lot of promise.
- All the software (except a 404 page in the help, which is minor) is available online; and the project seems maintained and well documented.

Reasons to reject:

\RC{There is no user evaluation, and therefore all the claims about worcs are unsupported by evidence. If the authors had included a compelling evaluation I would be in favor of accepting the paper.}

Response.

\Assignment{All}
\WorkInProgress
\Hard

\RC{This paper describes worcs, a package and workflow to help make R code reproducible according to the TOP guidelines.

The paper is well written, easy to follow and very relevant and timely for the journal. I think the capabilities of the worcs package for synthetic data generation, connect to osf, help with the preprints and capture dependencies are very valuable to the community. However, the main limitation of the paper is that the workflow and proposed package are not evaluated with users, and therefore the claims of the authors with respect of its capabilities are not supported by evidence. I would like to encourage the authors to gather user/community feedback and demonstrate with numbers that worcs can help scientists as they claim. Then, the paper would become a strong contribution for the journal.}

\Assignment{All}
\WorkInProgress
\Hard

\RC{Unsupported claims (beyond the ones derived from an absence of evaluation): The paper claims in many parts that worcs helps make an entire research project Open and Reproducible. However, at the same time there is also a claim for not supporting replication. Since replication is generally understood as a necessary steps towards reproducibility, I find this kind of contradictory. Maybe the claims can be modified stating that worcs helps documentation and understanding, hence supporting FAIR.}

\Assignment{All}
\WorkInProgress
\Hard

\RC{I am confused by the citation "Van Lissa, Caspar J., Aaron Peikert, and Andreas M. Brandmaier. 2020. “Worcs: Workflow for Open Reproducible Code in Science.”". Is this referring to another paper or the actual package? If it's another paper then what is the contribution of this publication?}

Response: Citation should be formatted as software.

\Assignment{CJ}
\WorkInProgress
\Easy

\RC{If the paper is generated as DDG, where are the functions that show it's capabilities? It is also not clear to me that some of the functionalities of Latex would be usable in RMarkdown, such as all the equations.}

Response: Add comment that LaTeX is supported in RMarkdown; showing functions (?) out of scope for present paper.

\Assignment{CJ}
\WorkInProgress
\Medium

\RC{It is not clear how worcs addresses depedency management. The section lists a series of tools that can be used, but the section is not very specific to worcs.}

Response: Not true; explain that worcs project automatically initializes `renv`.

\Assignment{CJ}
\WorkInProgress
\Easy

\RC{The workflow of releasing data products on GitHub is not ideal. First, because datasets of 100 MBs usually require special handling. Second, it doesn't follow very well the FAIR principles for findability and attribution of data. Instead, repositories like Zenodo or FigShare could be user to address data storage, with proper metadata and its corresponding citation. The authors seem to not describe very well what would happen in data science experiments where the data size is significant.}

\Assignment{BV, AB, AL, CJ}
\WorkInProgress
\Hard

\RC{In the 3 phases listed, it is not clear what worcs does and does not for users.}

\Assignment{All}
\WorkInProgress
\Medium

\RC{In some cases it looks like the authors claim that everyone should use R for open science. While R has a strong community behind, I believe there are others with very strong support (e.g. Python). I am not sure if suggesting a shift to R is the right approach. Note that this is the impression I got from reading the paper, maybe it was unintended by the authors.}

Response: Clearly define scope; feedback is welcome.

\Assignment{All}
\WorkInProgress
\Hard

\RC{Vignete on citation leads to a 404.}

\Assignment{CJ}
\WorkInProgress
\Easy

\RC{Synthetic data generation is fantastic, but it is not clear to me how does worcs actually keep the data consistent e.g. for training models. Does it follow a similar distribution to the source data?}

Response: Provide reference, clarify explanation in text.

\Assignment{CJ}
\WorkInProgress
\Each

\RC{Docker is deemed as complicated for novice users, but the setup vignete looks quite complex to me as well. Wouldn't a Docker container with worcs installed be easier to use? Maybe these aspects would be better highlighted in a user evaluation.}

Response: I've never used Docker. Andreas and Aaron - do you think you could clarify why worcs might have benefits over Docker? Should we make this proposed docker container available IN ADDITION to worcs as package?

\Assignment{AP, AB}
\WorkInProgress
\Medium

# Reviewer \#2
Submitted on 19/Oct/2020 by Remzi Celebi ORCID logo https://orcid.org/0000-0001-7769-4272

* Overall Impression: Good
* Suggested Decision: Accept
* Technical Quality of the paper: Good
* Presentation: Good
* Reviewer`s confidence: Medium
* Significance: High significance
* Background: Reasonable
* Novelty: Limited novelty
* Data availability: With exceptions that are admissible according to the data availability guidelines, all used and produced data are FAIR and openly available in established data repositories
* Length of the manuscript: The length of this manuscript is about right

The paper introduces the Workflow for Open Reproducible Code in Science (WORCS) , a workflow that complies with most of best practices for open sciences projects and demonstrates the use of the R worcs package for adapting WORCS workflow. The package provides a template for R users to create projects that follow open science principles.

Reasons to accept:

It provides a good discussion of TOP-guidelines (a set of open science principles) and how proposed workflow/package can be used to facilitate the adaptation of these guidelines.

Reasons to reject:

\RC{I would suggest the authors include a flowchart for users that shows the decision making process for which tools/operations should be used in WORCS. An example project would also help users better understand the package.}

\Assignment{ALL}
\WorkInProgress
\Medium

\RC{The claim that WORCS is amenable to the FAIR Guiding Principles should be justified. What are the solutions provided by WORCS that make a project interoperable and reusable? It would be improved with a more thorough discussion on how WORCS meets each FAIR principle.}

\Assignment{AL, BV, CJ}
\WorkInProgress
\Medium

\RC{Is WORCS providing any service for de-identification?}

Response: What would this entail? Is it within scope?

\Assignment{ALL}
\WorkInProgress
\Medium

# \Reviewer \#3

Submitted on 26/Oct/2020 by Anonymous

* Overall Impression: Weak
* Suggested Decision: Reject
* Technical Quality of the paper: Average
* Presentation: Good
* Reviewer`s confidence: High
* Significance: High significance
* Background: Reasonable
* Novelty: Limited novelty
* Data availability: All used and produced data (if any) are FAIR and openly available in established data repositories
* Length of the manuscript: The length of this manuscript is about right

The paper presents WORCS, a library that is implemented within R-studio with the purpose of promoting the creation of reproducible research. In particular, they assist the researchers in the tasks of generating the scholarly article, thereby managing a number of tasks that researchers have to deal with from the reformulation of their hypothesis to version control, to dependency management, and finally generating the scholarly article.

Reasons to accept:

On the positive side, the authors built their tool, with Open Science principles in mind. In doing so, they have pined the actions that need to be performed as part of the research process. For instance, steps such as the pre-registration of the objective of the study before starting the investigation is ignored by existing solutions.

Reasons to reject:

The objective of the author is ambiguous. That said, I do not think that WORCS provide sufficient features that make it stands out from state of the art tools, for the following reasons.

\RC{C1) The process as described by the author for conducting research is assumed to be linear. However, researchers often go back and forth on steps that they have already completed, and sometimes have to undo them completely and start from scratch. This does not seem to be tackled by the tool.}

Response: Opinions? I don't think this is entirely true. Nothing prevents researchers from going back and changing stuff? Or are there specific bottlenecks?

\Assignment{ALL}
\WorkInProgress
\Hard

\RC{C2) The gist of the tool seems to allow researchers to document their steps and generate a scholarly document that is ready for submission. In a sense, there is a similarity with existing notebooks such as Jupyter. So I was wondering why don’t the authors start from an existing popular tool such as Jupyter as the starting point (given the similarity in features that they would like to add), and extend Jupyter for instance with the new features such as dependency documentation and version control ?}

Response: Does anyone know Jupyter? What are our advantages?

\Assignment{ALL}
\WorkInProgress
\Medium

\RC{C3) Regarding Version control and dependency documentation, the authors uses in their system existing solution, namely git and env, respectively. While I did not criticize reusing existing solution, on the contrary, I don't believe that simply using such solution add a substantial value. For example, git capture changes at the level of the line of code/text, whereas a researcher needs an abstraction that informs him/her on the coarse-grained elements that are significant from the point of view of development and research steps so as to have a clear idea of where s/he is within the process and help him/her make the decision of the next steps to undertake.}

Response: Ideas?

\Assignment{BV, CJ}
\WorkInProgress
\Medium

\RC{To sum up, I think that the initial idea is interesting. However, I do not think that WORCS will be adopted by researchers or promote the state of the art of reproducibility, even if we focus on researchers who conduct their development and analysis using R.}

Response: Is this a fair point of criticism? "I do not thing that WORCS will be adopted"? Seems to me that the review should address the paper/software itself, not the reviewer's forecast of adoption. Then again, perhaps it is easily waylaid by referencing the download numbers (~2255 downloads now).

\Assignment{ALL}
\WorkInProgress
\Medium

# \Reviewer \#4

Submitted on 26/Oct/2020 by Katherine Wolstencroft https://orcid.org/0000-0002-1279-5133

* Overall Impression: Good
* Suggested Decision: Accept
* Technical Quality of the paper: Excellent
* Presentation: Good
* Reviewer`s confidence: Medium
* Significance: Moderate significance
* Background: Reasonable
* Novelty: Limited novelty
* Data availability: All used and produced data (if any) are FAIR and openly available in established data repositories
* Length of the manuscript: The length of this manuscript is about right

This paper describes a Workflow for Open Reproducible Code in Science (WORCS). It promotes reproducible and FAIR R code and complies with the TOP-guidelines. The workflow is presented with an R library and a Github repository containing worcs templates and all supporting materials for this manuscript.

Reasons to accept:

The paper is well written and clearly describes WORCS, which promotes reproducible and open code for scientific investigations. It complies with current best practices for open and reproducible code and it also follows the FAIR principles. Such initiatives help promote better open and reproducible science. The approach taken here is pragmatic and lightweight, which is necessary for large scale uptake. The authors also focus on ease of use and provide a "one-click solution".

WORCS is for scientists working with R with RMarkdown. It could therefore be argued that the utility is limited. However, R is widely used across science and the approach taken here could serve as an example for other software.

Reasons to reject:

\RC{The abstract of the paper states that the manuscript "provides examples of the implementation of worcs in R", so I was expecting science projects that had been described using worcs. If these examples exist, they should be referenced and linked to from the manuscript, as they would provide a better demonstration of the practical utility of the workflow and software than tutorial style examples.}

Response: Should we do this?

\Assignment{ALL}
\WorkInProgress
\Medium

\RC{There is a comparison in the paper to another R-based reproducible science solution (Peikert and Brandmaier (2019)). The authors discuss some of the differences between these approaches, but the comparison is limited. It is not clear to me how different these approaches actually are, apart from ease of use. If scientists are already familiar with R, is there a significant learning curve with either solution?}

\Assignment{AP, AB}
\WorkInProgress
\Medium

# Editor

Submitted by Tobias Kuhn on Tue, 10/27/2020 - 02:04, http://orcid.org/0000-0002-1267-0234

\RC{The reviewers agree that the paper has merit, but some of them also point to major shortcomings, in particular the lack of discussion of user adoption and of handling non-linear research processes. Note that according to our guidelines, resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse".}

Response:

\Assignment{AP, AB}
\WorkInProgress
\Medium

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2020-10-27 paper/reviews_ds.Rmd
Fix backticks

@@
 * Presentation: Good
-* Reviewer`s confidence: High
+* Reviewer's confidence: High
 * Significance: High significance
@@
 * Presentation: Good
-* Reviewer`s confidence: Medium
+* Reviewer's confidence: Medium
 * Significance: High significance
@@
 * Presentation: Good
-* Reviewer`s confidence: Medium
+* Reviewer's confidence: Medium
 * Significance: Moderate significance

### 2020-10-28 paper/Manuscript_DS.Rmd
explain why renv is more userfriendly than docker

@@
 project, `renv` will install all of the required packages onto a cache on their
-computer. 
-This is not a foolproof method to ensure strict computational reproducibility, but it is more user-friendly
+computer. As opposed to the alternatives `renv` is tightly integrated with the
+native R package management system and thus does not require the users to change their workflow of installing/removing and using packages.
+This is not a fail-safe method to ensure strict computational reproducibility, but it is more user-friendly
 and lightweight than any alternative out there, requiring no technical skill to

### 2020-10-28 paper/reviews_ds.Rmd
respond to reviewer concerning docker

@@
 
-Response: I've never used Docker.
-Andreas and Aaron - do you think you could clarify why worcs might have benefits over Docker?
-Should we make this proposed docker container available IN ADDITION to worcs as package?
-
+Response: Thank you for this thoughtful suggestion.
+We explained better why `renv` is more user friendly (because it builds upon the "normal" package workflow of R users).
+For users familiar with Docker, we have published an image on Docker Hub with all requirements of worcs installed and a Docker specific setup vignette, that is much shorter and given that you know how to use Docker easier.
 \Assignment{AP, AB}

### 2020-10-28 paper/Manuscript_DS.Rmd
hint at the docker setup

--- 2020-09-28 paper/Manuscript_DS.Rmd
@@
 By [tagging the commit](https://help.github.com/en/github/administering-a-repository/creating-releases) as a preregistration
-(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
+(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html)),
 it is distinct from all other commits, and easily findable by people and machines.

### 2020-10-28 paper/reviews_ds.Rmd
Address R1 comment about unsupported claims

--- 2020-10-27 paper/reviews_ds.Rmd
@@
 
-\Assignment{All}
+Response: We appreciate the Reviewer's comment, which seems to be rooted in a different definition of the terms "replication" and "reproducibility" than was intended by us.
+The lack of consistent definitions for these terms has been lamented by other authors (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018).
+As WORCS aims to meet the requirements set out in the TOP-guidelines, we adhere to the definitions of reproducibility and replication as used in the TOP-guidelines.
+This definition is formalized by Patil and colleagues: "We define reproducibility as re-performing the same analysis with the same code using a different analyst, and we define replicability as re-performing the experiment and collecting new data".
+The TOP-guidelines state that the first seven guidelines are intended to ensure the "reproducibility of the reported results based on the originating data".
+The eighth criterion is "not formally a transparency standard for authors, [but] addresses journal guidelines for consideration of independent replications for publication."
+
+We had taken notice of this definitional quagmire, but decided not to dwell on it in the present paper.
+This reviewer comment has made it abundantly clear that we need to explicitly define our terms.
+Thus, in order to address this comment, the revised manuscript now reads: REF.
+
+We wish to fully address this comment, including the criticism of unsupported claims - but it is a bit difficult to do so, given the fundamental misunderstanding about what constitutes "reproducibility".
+We have stated that worcs helps make an entire research project open and reproducible.
+It does so by:
+1) Providing a conceptual workflow that addresses all steps required to make a project open and reproducible
+2) Addressing common bottlenecks, such as the inability to share primary data, and offering solutions to overcome these hurdles
+3) Version controlling a project on Git(Hub)
+4) Creating a fully reproducible RMarkdown document
+5) Managing the package dependencies
+
+The Reviewer, by contrast, proposes that the main contribution of worcs is to help documentation and understanding.
+It is not immediately clear to us which components of worcs address these two points - except perhaps the readme.md that is created upon initialization of a new project.
+\Assignment{CJ, ALL}
 \WorkInProgress

### 2020-10-28 paper/reviews_ds.Rmd
Address R1 comment about no user evaluation

@@
 
-Response.
+Response: We appreciate the recommendation of performing a user evaluation, and this is certainly a next step in the development of worcs.
+However, as the Editor has indicated, 'resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse"'.
+We will thus first address the potential for re-use, and then we will address steps we have taken and intend to take with regard to user evaluation.
 
-\Assignment{All}
+First, there are several arguments to substantiate the potential for reuse of worcs:
+1) Worcs is implemented in an R-package
+2) This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network")
+3) The software meets all criteria of the Core Infrastructure Initiative "best practices" (CII Best Practices)
+4) The R-package uses only tools that are fully integrated within the RStudio IDE, thus enhancing ease of use
+5) The R-package is compatible with other reproducibility solutions, including Docker, drake, and make
+
+Second, there is some evidence to demonstrate that worcs is already being adopted:
+1) The worcs R-package has been downloaded 2255 times from CRAN (GitHub downloads are not tracked)
+2) The video tutorial for worcs has been viewed 356 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
+3) Worcs is the software used for reproducibility education at Utrecht University
+4) The lead author has given invited lectures on worcs at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
+5) The lead author has applied worcs in several projects:
+	* https://github.com/cjvanlissa/storm_meta_father_mother
+	* https://github.com/cjvanlissa/veni_sysrev
+	* https://github.com/cjvanlissa/placebo_adhd
+	* https://github.com/cjvanlissa/schumpe
+	* https://github.com/cjvanlissa/coping_covid
+
+Third, although we have not conducted a user evaluation, there is some evidence that worcs addresses users' needs:
+1) Evaluate worcshops!!
+2) The issues page on the worcs GitHub profile demonstrates that we have been very responsive in addressing users' feedback (as required by the aforementioned CII Best Practices):
+* https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed
+
+<!-- The goal of the present manuscript is to define the problem worcs intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools worcs uses to address this problem, explain how worcs uses these tools, and provide a step-by-step workflow.
+It is not clear what "claims" are made that are unsupported by evidence; -->
+
+\Assignment{CJ, BV, AL}
 \WorkInProgress

### 2020-10-28 paper/reviews_ds.Rmd
Address R1 comment user evaluation required

@@
 
-\Assignment{All}
-\WorkInProgress
-\Hard
+Response: In addressing this comment, we respectfully refer to our response to the preceding comment ("There is no user evaluation", R1), and again emphasize the Editor's comment that "resource papers do not need to have a full-blown user evaluation".
+We would also like to clarify that the goal of the present manuscript is to define the problem worcs intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools worcs uses to address this problem, explain how worcs uses these tools, and provide a step-by-step workflow.
+It is not clear what "claims" are made that are unsupported by evidence.
+In the revision, we have attempted to further clarify this goal from the paper onset: REF.
+
+We do take the suggestion of performing a user evaluation very seriously, but in our mind, this is a next step.
+First, the workflow must be published, then it can be taught, and subsequently evaluated.
+
+\Assignment{CJ}
+\Done
+\Easy
 

### 2020-10-28 paper/Manuscript_DS.Rmd
Add docker file

--- 2020-09-28 paper/Manuscript_DS.Rmd
@@
   - name: Loek Brinkman
-    affiliation: 'UMC Utrecht'
   - name: Anna-Lena Lamprecht
@@
 short:      "WORKFLOW FOR OPEN SCIENCE"
+shorttitle:      "WORKFLOW FOR OPEN SCIENCE"
 abstract: |

### 2020-10-28 paper/Manuscript_DS.Rmd

@@
 After setting up your system, you can use `worcs` for all of your
-projects. The `R` implementation of the workflow introduced earlier is detailed,
+projects. If you have Docker installed and you know how to use it, you may alternatively
+read [the stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html).
+The `R` implementation of the workflow introduced earlier is detailed,
 step-by-step, in

### 2020-10-28 paper/reviews_ds.Rmd
R1 Relevant citation for RMarkdown

@@
 
-Response: Add comment that LaTeX is supported in RMarkdown; showing functions (?) out of scope for present paper.
+Response: Add comment that LaTeX is supported in RMarkdown; showing functions (?) out of scope for present paper. Reference RMarkdown @xieMarkdownDefinitiveGuide2018
 

### 2020-10-28 paper/Manuscript_DS.Rmd
Tweak docker references

@@
 After setting up your system, you can use `worcs` for all of your
-projects. If you have Docker installed and you know how to use it, you may alternatively
-read [the stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html).
+projects.
+Alternatively, Docker users can follow [the stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html).
 The `R` implementation of the workflow introduced earlier is detailed,

### 2020-10-28 paper/Manuscript_DS.Rmd
Tweak docker references

@@
 This is effective, and also preserves software outside the `R` environment (e.g.,
-preprocessing tools for neuroimaging data). A Docker-based workflow
-that is conceptually and practically compatible with WORCS
-was recently introduced by @peikertReproducibleDataAnalysis2019.
-However, it is somewhat difficult to set up for novice users. 
-The `R` package `repro`
-is currently in development to facilitate adoption [@repro]. 
+preprocessing tools for neuroimaging data).
+The [stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html)
+describes how users may instantiate a Docker build running `worcs` and its dependencies.
+For more information on this topic,
+@peikertReproducibleDataAnalysis2019 describe 
+a Docker-based workflow for computational reproducibility that is
+conceptually and practically compatible with WORCS.
+One remaining challenge of Docker is that it is somewhat difficult to set up for novice users.
+The `worcs` package meets most reproducibility requirements while using only tools
+native to RStudio, and is therefore easier to adopt for `R` users.
+However, an `R` package is currently in development to facilitate Docker integration [`repro`, @repro].
 Another 

### 2020-10-31 paper/Manuscript_DS.Rmd
Rewrite reproducibility section

@@
 
-Many solutions exist to ensure computational reproducibility, which differ in
-user-friendliness and effectiveness. 
+Many solutions exist to ensure computational reproducibility,
+which differ in user-friendliness and effectiveness.
 These solutions typically work by 
-enveloping your research project in a distinct "environment" that only has access to programs that
-are explicitly installed, and maintaining a record of these programs.
-If strict reproducibility is required, users might choose to conduct all analyses in a "Docker container": An environment that behaves like a
+enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs.
+When choosing an appropriate solution for dependency management, there is a tradeoff between ease-of-use on the one hand, and robustness on the other.
+In making this tradeoff, it is important to consider that all solutions have a limited shelf life
+[@brownHowLearnedStop2017].
+Therefore, the best approach might be to use a "good-enough" solution, and
+acknowledge that all code requires some maintenance if you want to
+reproduce it in the future.
+
+For analyses conducted entirely in `R`, the recently released package `renv` [@usheyRenvProjectEnvironments2020]
+strikes a great balance between user-friendliness and robustness.
+Developed by the team behind RStudio, 
+this package simply maintains a text-based,
+human-readable log of all packages used, their version numbers, and where they
+were installed from (e.g., CRAN, Bioconductor, GitHub).
+This lightweight approach safeguards reproducibility as long as these repositories maintain their archives.
+The text-based log file can be version controlled with Git.
+The `renv` function `snapshot()` is used to add packages to the log file.
+This function identifies all packages loaded using `library()` calls in the scripts in the project directory.
+
+All solutions to computational reproducibility require users to make their dependencies explicit, often by reinstalling all software for each project in a controlled environment.
+This can lead to long installation times and large memory requirements.
+Similarly, `renv` requires users to call `install.packages()` for each library used - but it does not reinstall packages for each new project.
+Instead, packages reside in a cache that is shared between all projects on the computer.
+If a project uses a package that is already in the cache, it is not reinstalled, but linked from the cache.
+This improves the user experience by saving time and memory.
+When someone else loads your project, `renv` will install all of the required packages.
+As opposed to alternative solutions, `renv` is tightly integrated with the
+native R package management system.
+Therefore, it does not require the users to change their workflow of installing/removing and using packages, and requires no technical skill to setup.
+It is not a fail-safe method to ensure strict computational reproducibility,
+but it meets most reproducibility requirements using only tools
+native to RStudio, and is therefore easier to adopt for `R` users.
+
+If stricter reproducibility is required,
+users might choose to conduct all analyses in a "Docker container": An environment that behaves like a
 virtual computer, which can be stored like a time capsule, and
@@
 The [stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html)
-describes how users may instantiate a Docker build running `worcs` and its dependencies.
+describes how to instantiate a Docker build running `worcs` and its dependencies.
 For more information on this topic,
@@
 One remaining challenge of Docker is that it is somewhat difficult to set up for novice users.
-The `worcs` package meets most reproducibility requirements while using only tools
-native to RStudio, and is therefore easier to adopt for `R` users.
-However, an `R` package is currently in development to facilitate Docker integration [`repro`, @repro].
+To overcome this challenge, an `R` package is currently in development to facilitate Docker integration [`repro`, @repro].
 Another 
@@
 cloud-computing solution for data analysis.
-
-For analyses conducted entirely in `R`, the recently released package `renv`
-strikes a great balance between user-friendliness  and effectiveness. Developed by
-the team behind RStudio, this package simply maintains a text-based,
-human-readable log of all packages used, their version numbers, and where they
-were installed from (e.g., CRAN, Bioconductor, GitHub). As long as these
-repositories maintain their archives, this solution is reliable and very
-lightweight.
-The text-based log file can be version controlled with Git. `renv`
-installs the logged packages in a cache that is shared between all
-projects on the computer. Thus, if one of your projects uses a package that is
-already in the cache, it does not
-need to be installed again. This prevents long installation
-times, large space requirements, and frustration having to re-install software
-that is already installed on the system. When someone else loads your 
-project, `renv` will install all of the required packages onto a cache on their
-computer. As opposed to the alternatives `renv` is tightly integrated with the
-native R package management system and thus does not require the users to change their workflow of installing/removing and using packages.
-This is not a fail-safe method to ensure strict computational reproducibility, but it is more user-friendly
-and lightweight than any alternative out there, requiring no technical skill to
-setup. When choosing an appropriate solution for dependency management, it is
-important to consider that all solutions have a limited shelf life
-[@brownHowLearnedStop2017].
-Ultimately, the best approach might be to use a "good-enough" solution, and
-acknowledge that all code requires some maintenance if you want to
-reproduce it in the future. For projects in `R`, `renv` meets these requirements.
 

### 2020-10-31 paper/Manuscript_DS.Rmd
RStudio public benefit, encourage porting to other languages

@@
 
-The `R` package `worcs` automates much of the workflow
-for `R` users.
-Several arguments support the choice to implement this workflow in `R`.
+WORCS is, first and foremost, a conceptual workflow that could be implemented in any software environment.
+As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].
+Several arguments support the choice to implement this workflow in `R` first, although we encourage developers to port the software to other languages.
 First, `R` and all of it extensions are free open source software, which
@@
 and some of them are actively developed by the team behind RStudio.
-Third, `R` is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
-Fourth, `R` is well-supported by
+Third, unlike any competitor, RStudio is a Public Benefit Corporation, whose primary purpose is explicitly aligned with open science principles:
+
+> RStudio’s primary purpose is to create free and open-source software for data science, scientific research, and technical communication.
+This allows anyone with access to a computer to participate freely in a global economy that rewards data literacy; enhances the production and consumption of knowledge; and facilitates collaboration and reproducible research in science, education and industry.
+
+Fourth, `R` is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
+Fifth, `R` is well-supported by
 a vibrant and inclusive online community,

### 2020-10-31 paper/reviews_ds.Rmd
Update reviews_ds.Rmd
Response to R1: Why not Jupyter?

@@
 
-Response: Does anyone know Jupyter? What are our advantages?
+Response: We agree with the Reviewer that the conceptual workflow as explained in the paper could be implemented in any software environment, including Jupyter.
+In fact, we acknowledge this in the paragraph on *Future developments*: *"although the workflow is currently implemented only in `R`, it is conceptually relevant for users of other statistical programming languages, most notably `Python` [@10.5555/1593511].
+We welcome efforts to implement WORCS in other platforms."*.
 
-\Assignment{ALL}
+The simple and straightforward response to the Reviewer's question is that we implemented this workflow in R because we are R developers.
+But in the paper, we provide some more detailed arguments to substantiate a preference for R over other environments:
+
+> First, `R` and all of it extensions are free open source software, which
+make it a tool of choice for open science.
+Second, all tools required for an open science workflow are implemented in `R`,
+most of these tools are directly accessible through the RStudio user interface,
+and some of them are actively developed by the team behind RStudio.
+Third, unlike any competitor, RStudio is a Public Benefit Corporation, whose primary purpose is explicitly aligned with open science principles:  
+*RStudio’s primary purpose is to create free and open-source software for data science, scientific research, and technical communication.
+This allows anyone with access to a computer to participate freely in a global economy that rewards data literacy; enhances the production and consumption of knowledge; and facilitates collaboration and reproducible research in science, education and industry.*
+Fourth, `R` is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
+Fifth, `R` is well-supported by
+a vibrant and inclusive online community,
+which develops new methods and packages,
+and provides support and tutorials for existing ones.
+Finally, `R` is highly interoperable:
+Packages are available to read and write nearly every filetype,
+including DOCX, PDF (in APA style), and HTML. 
+Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from `R`, including `C++`, 
+`Fortran`, and `Python` [@allaireMarkdownPythonEngine2020]. There are excellent free
+resources for learning `R` [e.g., @grolemundDataScience2017].
+
+\Assignment{BV, CJ}
 \WorkInProgress

### 2020-10-31 paper/Manuscript_DS.Rmd
clarify public benefit corporation section

@@
 As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].
-Several arguments support the choice to implement this workflow in `R` first, although we encourage developers to port the software to other languages.
+Several arguments support the choice to implement this workflow in `R` and RStudio first, although we encourage developers to port the software to other languages.
 First, `R` and all of it extensions are free open source software, which
@@
 and some of them are actively developed by the team behind RStudio.
-Third, unlike any competitor, RStudio is a Public Benefit Corporation, whose primary purpose is explicitly aligned with open science principles:
+Third, RStudio is a Public Benefit Corporation, whose commitment to open science is codified in the company charter:
 

### 2020-10-31 paper/Manuscript_DS.Rmd
R3 C2: Non-linear

@@
 We provide a conceptual outline of the WORCS procedure below, based on WORCS
-Version 0.1.2. 
+Version 0.1.5.
+Note that, although the steps are numbered for reference purposes,
+we acknowledge that the process of conducting research is not always linear.
 For users of the `worcs` `R` package, we highly recommend referring to

### 2020-10-31 paper/reviews_ds.Rmd
R3 C2: Non-linear

@@
 
-Response: Opinions? I don't think this is entirely true. Nothing prevents researchers from going back and changing stuff? Or are there specific bottlenecks?
+Response: We agree with the Reviewer that researchers often go back and forth in their work.
+We did not intend to convey the impression that the process of conducting research is assumed to be linear.
+Of course, since we are presenting a workflow - there are a number of steps involved, which we have placed in a sensible order.
+For example, Study Design must logically occur prior to Writing and Analysis, and Publication can only logically occur after Writing.
+We also wish to emphasize that the "tool" (we assume the Reviewer is referring to the `worcs` R-package) does not enforce any kind of linearity.
+
+To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in the paper and the Workflow vignette, that the steps are not necessarily linear.
+Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have now added functions to increase the non-linear flexibility of the `worcs` package: `add_manuscript()` and `add_preregistration()`.
 

### 2020-10-31 paper/reviews_ds.Rmd
R3 C1: Non-linear

@@
 We did not intend to convey the impression that the process of conducting research is assumed to be linear.
-Of course, since we are presenting a workflow - there are a number of steps involved, which we have placed in a sensible order.
+Of course, since we are presenting a workflow - there are a number of steps involved, which we have placed in roughly chronological order and numbered for reference purposes.
 For example, Study Design must logically occur prior to Writing and Analysis, and Publication can only logically occur after Writing.
@@
 
-To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in the paper and the Workflow vignette, that the steps are not necessarily linear.
+To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in the paper and the Workflow vignette, that the steps are not necessarily linear:  *"Note that, although the steps are numbered for reference purposes, we acknowledge that the process of conducting research is not always linear."*.
 Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have now added functions to increase the non-linear flexibility of the `worcs` package: `add_manuscript()` and `add_preregistration()`.

### 2020-10-31 paper/Manuscript_DS.Rmd
R3 C3: Git not useful

@@
 An integral part of WORCS is the *explicit* version control software Git
-(www.git-scm.com). Git tracks
-changes to files, and stores these changes when the user makes a "commit" (a 
-snapshot of the version controlled files).
+(www.git-scm.com).
+Git tracks changes to files on a line-by-line basis.
+The user can sture these changes by making a "commit": a 
+snapshot of the version controlled files.
+Each "commit" can contain as many changes as desired;
+for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
 Git retains a complete log of all commits, and users can compare changes between
@@
 must be decoded first, often requiring commercial software.
-Git is designed for use with text-based files, as it
-logs changes to text-based files on a line-by-line basis.
 Although binary files can be version controlled with Git, 

### 2020-10-31 paper/reviews_ds.Rmd
R3 C3: Git not useful

@@
 
-\Assignment{ALL}
-\WorkInProgress
-\Hard
+\Assignment{CJ}
+\Done
+\Easy
 
@@
 
-Response: We agree with the Reviewer that the conceptual workflow as explained in the paper could be implemented in any software environment, including Jupyter.
+Response: There are two points we wish to address here.
+First, we would like to clarify that this paper introduces WORCS, a conceptual workflow for open science research projects.
+When the Reviewer refers to "the tool" - we assume this to mean the `worcs` R-package, which offers convenience functions to help R-users follow the WORCS-workflow.
+This is clarified in the Abstract and in the paper, e.g.:
+
+> Abstract: *This paper introduces the Workflow for Open Reproducible Code in Science (WORCS): A step-by-step procedure that researchers can follow to make a research project open and reproducible, accompanied by an open source software implementation that automates these steps.*
+
+> Manuscript, paragraph titled **The R implementation of WORCS**: *WORCS is, first and foremost, a conceptual workflow that could be implemented in any software environment.
+As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].*
+
+Second, in response to the main comment, we agree with the Reviewer that the conceptual workflow as explained in the paper could be implemented in any software environment, including Jupyter.
 In fact, we acknowledge this in the paragraph on *Future developments*: *"although the workflow is currently implemented only in `R`, it is conceptually relevant for users of other statistical programming languages, most notably `Python` [@10.5555/1593511].
@@
 
-The simple and straightforward response to the Reviewer's question is that we implemented this workflow in R because we are R developers.
-But in the paper, we provide some more detailed arguments to substantiate a preference for R over other environments:
+To address the Reviewer's question directly: We implemented this workflow in R because we are R developers.
+But in the paper, we also provide some arguments that compel our preference for R and RStudio as a choice for statistical computing in open science:
 
@@
 
-Response: Ideas?
+Response: Respectfully, we are not entirely sure what the criticism is here.
+First, there seems to be a misunderstanding about the scope of the paper, which is described in the Abstract: *"This paper introduces [a] step-by-step procedure that researchers can follow to make a research project open and reproducible, accompanied by an open source software implementation that automates these steps"*.
 
-\Assignment{BV, CJ}
-\WorkInProgress
+Our paper thus introduces a workflow for open science, based on objective criteria set out in the TOP-guidelines.
+Aside from the work in progress of our co-authors - which is focused on strict computational reproducibility and is compatible with the present workflow - this is one of the first efforts to translate the requirements of open science into a step-by-step workflow.
+When you search Google Scholar for "open science workflow", the first hit that actually describes a workflow is the present manuscript.
+
+Second, the Reviewer claims that using e.g.
+Git does not add substantial value - but this claim is in disagreement with published literature that makes a strong case for the value of using Git for research [see two references cited in the paper, @ramGitCanFacilitate2013;  @blischakQuickIntroductionVersion2016].
+
+Third, the Reviewer remarks that 
+"git capture changes at the level of the line of code/text, whereas a researcher needs an abstraction that informs him/her on the coarse-grained elements [...] so as to have a clear idea of where s/he is within the process and help him/her make the decision of the next steps to undertake".
+This comment suggests a fundamental misunderstanding of how Git works.
+Although Git tracks file changes on a line-by-line basis,
+any number of such changes can be committed together, thereby capturing the more "course-grained elements".
+For example, in our response to these reviews, we have created one Git commit for each Reviewer comment, and these commits contain changes to the action letter, manuscript, and vignettes.
+
+We do see the value in clarifying this point in the manuscript as well, and to address this comment, the manuscript now reads:
+
+> Git tracks changes to files on a line-by-line basis.
+The user can sture these changes by making a "commit": a snapshot of the version controlled files.
+Each "commit" can contain as many changes as desired; for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
+
+\Assignment{CJ}
+\Done
 \Medium

### 2020-10-31 paper/reviews_ds.Rmd
Improve R3 C3: Git not useful

@@
 Second, in response to the main comment, we agree with the Reviewer that the conceptual workflow as explained in the paper could be implemented in any software environment, including Jupyter.
-In fact, we acknowledge this in the paragraph on *Future developments*: *"although the workflow is currently implemented only in `R`, it is conceptually relevant for users of other statistical programming languages, most notably `Python` [@10.5555/1593511].
+In fact, we emphasize this in the paragraph on *Future developments*: *"although the workflow is currently implemented only in `R`, it is conceptually relevant for users of other statistical programming languages, most notably `Python` [@10.5555/1593511].
 We welcome efforts to implement WORCS in other platforms."*.
 
-To address the Reviewer's question directly: We implemented this workflow in R because we are R developers.
+To answer the Reviewer's question directly: We implemented this workflow in R because we are R developers.
 But in the paper, we also provide some arguments that compel our preference for R and RStudio as a choice for statistical computing in open science:
@@
 \Assignment{BV, CJ}
-\WorkInProgress
+\Done
 \Medium
@@
 Response: Respectfully, we are not entirely sure what the criticism is here.
-First, there seems to be a misunderstanding about the scope of the paper, which is described in the Abstract: *"This paper introduces [a] step-by-step procedure that researchers can follow to make a research project open and reproducible, accompanied by an open source software implementation that automates these steps"*.
-
-Our paper thus introduces a workflow for open science, based on objective criteria set out in the TOP-guidelines.
-Aside from the work in progress of our co-authors - which is focused on strict computational reproducibility and is compatible with the present workflow - this is one of the first efforts to translate the requirements of open science into a step-by-step workflow.
+First, as addressed in our response to the previous comment, there seems to be a misunderstanding about the scope of the paper.
+To clarify, we again state that our paper introduces a workflow for open science, based on objective criteria set out in the TOP-guidelines.
+Regarding novelty: Aside from the work in progress of our co-authors - which is focused on strict computational reproducibility and is compatible with the present workflow - this is one of the first efforts to translate the requirements of open science into a step-by-step workflow.
 When you search Google Scholar for "open science workflow", the first hit that actually describes a workflow is the present manuscript.

### 2020-10-31 paper/reviews_ds.Rmd
R3 C4: Will not be adopted

@@
 
-Second, there is some evidence to demonstrate that worcs is already being adopted:
-1) The worcs R-package has been downloaded 2255 times from CRAN (GitHub downloads are not tracked)
-2) The video tutorial for worcs has been viewed 356 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
-3) Worcs is the software used for reproducibility education at Utrecht University
-4) The lead author has given invited lectures on worcs at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
-5) The lead author has applied worcs in several projects:
+Second, there is some evidence to demonstrate that worcs is already being adopted (as of 31-10-2020):
+
+1. The preprint has been downloaded 561 times, since being published on 31-05-2020
+1. The `worcs` R-package has been downloaded 2401 times from CRAN, since being published on 18-05-2020
+1. The GitHub project has been forked by users unknown to us 4 times, and starred (= watched) 32 times
+1. The video tutorial for worcs has been viewed 362 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
+1. WORCS is used to teach reproducibility at Utrecht University
+1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
+1. The lead author has applied worcs in several research projects:
 	* https://github.com/cjvanlissa/storm_meta_father_mother
@@
 
-\RC{To sum up, I think that the initial idea is interesting.
+\RC{C4: To sum up, I think that the initial idea is interesting.
 However, I do not think that WORCS will be adopted by researchers or promote the state of the art of reproducibility, even if we focus on researchers who conduct their development and analysis using R.}
 
-Response: Is this a fair point of criticism?
-"I do not thing that WORCS will be adopted"?
-Seems to me that the review should address the paper/software itself, not the reviewer's forecast of adoption.
-Then again, perhaps it is easily waylaid by referencing the download numbers (~2255 downloads now).
+Response: We thank the Reviewer for acknowledging the idea to be interesting.
+Whether it will be adopted by researchers remains to be seen after publication.
+There is some prelimiary evidence that the workflow is being adopted (as of 31-10-2020):
 
-\Assignment{ALL}
-\WorkInProgress
-\Medium
+1. The preprint has been downloaded 561 times, since being published on 31-05-2020
+1. The `worcs` R-package has been downloaded 2401 times from CRAN, since being published on 18-05-2020
+1. The GitHub project has been forked by users unknown to us 4 times, and starred (= watched) 32 times
+1. The video tutorial for worcs has been viewed 362 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
+1. WORCS is used to teach reproducibility at Utrecht University
+1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
+1. The lead author has applied worcs in several research projects:
+	* https://github.com/cjvanlissa/storm_meta_father_mother
+	* https://github.com/cjvanlissa/veni_sysrev
+	* https://github.com/cjvanlissa/placebo_adhd
+	* https://github.com/cjvanlissa/schumpe
+	* https://github.com/cjvanlissa/coping_covid
+
+\Assignment{CJ}
+\Done
+\Easy
 

### 2020-10-31 paper/Manuscript_DS.Rmd
R4 C2: Compare with Peikert and Brandmaier

@@
 researchers can meet the requirements of open science.
-One notable example is the workflow for
-reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
-workflow strives to ensure strict computational reproducibility for
-even the most sophisticated analyses. It has some limitations, however: In
-focusing on reproducibility, it only addresses TOP-guidelines 2, 3, and
-5. Moreover, in aiming for strict computational reproducibility it ends up being
-relatively complex to implement. WORCS builds upon the same general principles
-as Peikert and Brandmaier,
-which means the two workflows are compatible. What 
-sets WORCS apart is that is an easy and sufficient solution for projects that
-can be conducted entirely within `R`, resulting in a much more lightweight
-approach to safeguard computational reproducibility under *most circumstances*.
-WORCS also addresses a unique issue not covered by other existing solutions,
-namely to provide a workflow most conducive to satisfying the TOP-guidelines and
-FAIR principles, while being compatible with existing requirements.
+A notable example is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019,
+which uses Docker to ensure strict computational reproducibility for
+even the most sophisticated analyses.
+There are some key differences with WORCS.
+First, with regard to the scope of the workflow: 
+WORCS is designed to address a unique issue not covered by other existing solutions,
+namely to provide a workflow most conducive to satisfying TOP-guidelines 1-7 while adhering to the FAIR principles.
+Peikert and Brandmaier instead focus primarily on computational reproducibility,
+which is relevant for TOP-guidelines 2, 3, and 5.
+Second, with regard to ease of use,
+WORCS aims to bring down the learning curve by adopting sensible defaults for any decisions to be made.
+Peikert and Brandmaier, by contrast, designed their workflow to be very flexible and comprehensive,
+thus requiring substantially more background knowledge and technical sophistication from users.
+To sum up, WORCS builds upon the same general principles
+as Peikert and Brandmaier, and the two workflows are compatible. What 
+sets WORCS apart is that it is a more lightweight
+approach to safeguard computational reproducibility under *most circumstances*, 
+which is easier to adopt for projects that can be conducted entirely within `R`.
 

### 2020-10-31 paper/reviews_ds.Rmd
R4 C2: Compare with Peikert and Brandmaier

@@
 
+We agree that this is an important point, and have rewritten the paragraph comparing these two solutions (as cited below).
+We hope that this adresses the Reviewer's question.
+However, as the workflow of Peikert and Brandmaier is still in the preprint stage, and the software implementation has not yet been developed - a more detailed comparison of these two approaches would be premature.
+
+> A different class of solutions instead focuses on the practical issue of *how*
+researchers can meet the requirements of open science.
+A notable example is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019,
+which uses Docker to ensure strict computational reproducibility for
+even the most sophisticated analyses.
+There are some key differences with WORCS.
+First, with regard to the scope of the workflow: 
+WORCS is designed to address a unique issue not covered by other existing solutions,
+namely to provide a workflow most conducive to satisfying TOP-guidelines 1-7 while adhering to the FAIR principles.
+Peikert and Brandmaier instead focus primarily on computational reproducibility,
+which is relevant for TOP-guidelines 2, 3, and 5.
+Second, with regard to ease of use,
+WORCS aims to bring down the learning curve by adopting sensible defaults for any decisions to be made.
+Peikert and Brandmaier, by contrast, designed their workflow to be very flexible and comprehensive,
+thus requiring substantially more background knowledge and technical sophistication from users.
+To sum up, WORCS builds upon the same general principles
+as Peikert and Brandmaier, and the two workflows are compatible. What 
+sets WORCS apart is that it is a more lightweight
+approach to safeguard computational reproducibility under *most circumstances*, 
+which is easier to adopt for projects that can be conducted entirely within `R`.
+
 \Assignment{AP, AB}

### 2020-10-31 paper/reviews_ds.Rmd
R4 C2: Compare with Peikert and Brandmaier

@@
 
-\RC{The abstract of the paper states that the manuscript "provides examples of the implementation of worcs in R", so I was expecting science projects that had been described using worcs.
+\RC{C1: The abstract of the paper states that the manuscript "provides examples of the implementation of worcs in R", so I was expecting science projects that had been described using worcs.
 If these examples exist, they should be referenced and linked to from the manuscript, as they would provide a better demonstration of the practical utility of the workflow and software than tutorial style examples.}
@@
 
-\RC{There is a comparison in the paper to another R-based reproducible science solution (Peikert and Brandmaier (2019)).
+\RC{C2: There is a comparison in the paper to another R-based reproducible science solution (Peikert and Brandmaier (2019)).
 The authors discuss some of the differences between these approaches, but the comparison is limited.

### 2020-11-01 README.Rmd
R4 C1: Add list of WORCS project examples

@@
 [![CRAN status](https://www.r-pkg.org/badges/version/worcs)](https://cran.r-project.org/package=worcs)
-[![metacran downloads](https://cranlogs.r-pkg.org/badges/worcs)](https://cran.r-project.org/package=worcs)
+[![CRAN RStudio mirror downloads](https://cranlogs.r-pkg.org/badges/grand-total/worcs?color=blue)](https://r-pkg.org/pkg/worcs)
 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
@@
 
+## Adoption of WORCS by users
+
+```{r, echo = FALSE, message=FALSE, warning=FALSE}
+library(gh)
+library(dlstats)
+dl <- dlstats::cran_stats("worcs")
+dl <- sum(dl$downloads)
+
+worcs_repo <- gh("/repos/cjvanlissa/worcs")
+
+repos <- gh("GET /search/code", username = "cjvanlissa", q="worcs_version filename:.worcs")
+
+worcs_repos <- data.frame(t(sapply(repos[["items"]], function(x){
+  c(owner = x$repository$owner$login,
+    name = x$repository$name,
+    url = x$repository$html_url
+  )
+})))
+
+worcs_repos <- worcs_repos[!((worcs_repos$owner %in% c("cjvanlissa", "cran") & worcs_repos$name == "worcs")| grepl("(penguin|\\btmp|demo|test)", tolower(worcs_repos$name))), ]
+
+tmp <- paste0(sapply(unique(worcs_repos$owner), function(x){
+  owner_repos <- worcs_repos[worcs_repos$owner == x, , drop = FALSE]
+  paste0("\n    - ", x, "  ",
+         paste0("\n        + [", owner_repos$name, "](", owner_repos$url, ")  ", collapse = ""))
+  }),
+  collapse = "")
+
+invited <- read.table("https://raw.githubusercontent.com/cjvanlissa/resume/master/invited.csv", sep = ",", header = TRUE)
+invited <- invited[grepl("(worcs|workflow)", tolower(invited$Topic)), c("Institute", "Year", "City", "Country", "Date")]
+invited <- paste0(apply(invited, 1, function(x){paste0("\n    - ", x[1], " (", x[2], ")  ")}), collapse = "")
+```
+
+As of `r Sys.Date()`, these are indicators of the adoption of `worcs` by users:
+
+1. The preprint has been downloaded 561 times, since being published on 31-05-2020
+1. The `worcs` R-package has been downloaded `r dl` times from CRAN, since being published on 18-05-2020
+1. The GitHub project has been forked `r worcs_repo$network_count` times, watched `r worcs_repo$subscribers_count` times, and starred `r worcs_repo$watchers` times
+1. The lead author has given invited lectures on WORCS at:`r invited`
+1. WORCS is currently used in the following GitHub repositories (sorted by user):`r tmp`
+
 ## Contributing and Contact Information

### 2020-11-01 paper/Manuscript_DS.Rmd
R4 C1: Add list of WORCS project examples

@@
   standards for open science, and provides examples of the implementation of
-  `worcs` in `R`. All source code for the `R` package,
-  and replication materials for the manuscript, are available at 
+  `worcs` in `R`. The source code for the `R` package and manuscript,
+  and a list of user examples of WORCS projects, are available at 
   https://github.com/cjvanlissa/worcs.
@@
 which is more detailed and updated with future developments.
+For an up-to-date list of sample `worcs` projects, see the [WORCS GitHub page](https://github.com/cjvanlissa/worcs).
 

### 2020-11-01 paper/reviews_ds.Rmd
R4 C1: Add list of WORCS project examples

@@
 
-Response: Should we do this?
+Response: We thank the Reviewer for this helpful suggestion!
+Although such user examples are not yet numerous, the lead author has six public WORCS repositories, and there are two public repositories by other users.
+In our opinion, it would be most useful to have a continuously cumulating list of example repositories, and link that list in the paper - rather than to publish a static list.
+Therefore, to address this comment, we have scripted a web scraper that searches GitHub for WORCS repositories (using the metadata tags created by the `worcs` package).
+We have embedded the scraper in the README.Rmd file on the worcs GitHub page, which is regularly updated.
+In the revised paper, we point readers to [this readme](https://github.com/cjvanlissa/worcs) for a list of user examples, in the Abstract, and when introducing the workflow.
 
-\Assignment{ALL}
-\WorkInProgress
-\Medium
+\Assignment{CJ}
+\Done
+\Hard
 
@@
 
-\Assignment{AP, AB}
+\Assignment{CJ, AP, AB}
 \WorkInProgress

### 2020-11-01 paper/reviews_ds.Rmd
Editor comment draft response

@@
 
-Response:
+Response: We thank the Editor and all Reviewers for their helpful comments on this manuscript.
+We have attempted to address all of these points in the action letter above.
+With regard to the specific point reiterated by the Editor:
+
+### User adoption
+
+We thank the Editor for pointing out that a full-blown user evaluation is not required for a resource paper.
+Of course, we take this Reviewer suggestion very seriously, and intend to conduct a user evaluation after the paper is published and users begin to adopt the workflow.
+In the meantime, we present the following evidence for WORCS' potential for reuse:
+
+TODO
+
+### Non-linear research processes
+
+Although Reviewer 3 claims that WORCS assumes a linear research process, this is incorrect.
+
+We agree with Reviewer 3 that researchers often go back and forth in their work. 
+We did not intend to convey the impression that the process of conducting research is assumed to be linear. 
+Of course, since we are presenting a workflow - there are a number of steps involved, which we have placed in roughly chronological order and numbered for reference purposes.
+For example, Study Design must logically occur prior to Writing and Analysis, and Publication can only logically occur after Writing.
+We also wish to emphasize that the "tool" (we assume the Reviewer is referring to the `worcs` R-package) does not enforce any kind of linearity.
+
+To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in the paper and the Workflow vignette, that the steps are not necessarily linear:  *"Note that, although the steps are numbered for reference purposes, we acknowledge that the process of conducting research is not always linear."*.
+Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have now added functions to make non-linear workflows even more convenient, specifically: `add_manuscript()` and `add_preregistration()`.
 

### 2020-11-02 paper/Manuscript_DS.Rmd
R1 C4 fix reference list

@@
 author:
-  - name: Caspar J. van Lissa
-    affiliation: 'Utrecht University'
-    address: Padualaan 14, 3584CH Utrecht, The Netherlands
-    email: c.j.vanlissa@uu.nl
-  - name: Andreas M. Brandmaier
-    affiliation: 'Max Planck Institute'
-  - name: Loek Brinkman
-  - name: Anna-Lena Lamprecht
-    affiliation: 'Utrecht University'
-    email: a.l.lamprecht@uu.nl
-  - name: Aaron Peikert
-    affiliation: Max Planck Institute
-  - name: Marijn E. Struiksma
-    affiliation: 'Utrecht University'
-  - name: Barbara Vreede
-    affiliation: 'Utrecht University'
+  name: |
+    [Caspar J. Van Lissa](c.j.vanlissa@uu.nl)$^{a,b}$
+    Andreas M. Brandmaier$^{c,d}$
+    Loek Brinkman$^{b,e}$
+    Anna-Lena Lamprecht$^{b,f}$
+    Aaron Peikert$^{c,g}$
+    Marijn E. Struiksma$^{b,h}$
+    Barbara Vreede$^{i}$
+
+  affiliation: | 
+      a: Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics  
+      b: Open Science Community Utrecht  
+      c: Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany  
+      d: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK  
+      e: UMC Utrecht  
+      f: Utrecht University, Department of Information and Computing Sciences  
+      g: Humboldt-Universitaet zu Berlin  
+      h: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
+      i: University Library Utrecht
 title: "WORCS: A Workflow for Open Reproducible Code in Science"
@@
 output: rmarkdown::html_document
+---
+<!--
+, "worcs_packages.bib"
+output: 
+  rmarkdown::html_document:
+    citation_package: biblatex Note: Citations are always handled by pandoc when format is not PDF
 knit: worcs::cite_all
----
+csl: apa.csl
+csl: apa.csl
+-->
 
@@
 implemented in `R` [@rcoreteamLanguageEnvironmentStatistical2020].
-The `worcs` package [@worcspackage2020] offers a project template for RStudio
+The `worcs` package [@worcs] offers a project template for RStudio
 [@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
@@
 
-# Defining Open Science practices
-
-Although Open Science is advocated by many, it does not have a unitary definition
-[but see @tennantValuePropositionOpen2020].
-One of the most influential operationalisations
-of Open Science principles is known as the "TOP-guidelines" [@nosekPromotingOpenResearch2015a].
-These guidelines describe eight standards for open science: 1) Comprehensive
+# Defining open science practices
+
+Although open science is advocated by many,
+it does not have a unitary definition.
+Throughout this paper, we adhere to the definition of open science as
+"the practice of science in such a way that others can collaborate and contribute,
+where research data, lab notes and other research processes are freely available,
+under terms that enable reuse, redistribution and reproduction of the research and its underlying data and methods" [@vicente-saezOpenScienceNow2018].
+In this definition, the term "science" refers to all scholarly disciplines.
+We further define the term *reproducible*, as used in the WORCS acronym, as being able to re-perform "the same analysis with the same code using a different analyst" [@patilVisualToolDefining2019].
+
+The "TOP-guidelines" are one of the most influential concrete operationalisations 
+of general open science principles [@nosekPromotingOpenResearch2015a].
+These guidelines describe eight standards for open science,
+the first seven of which "account for reproducibility of the reported results
+based on the originating data, and for sharing sufficient information to conduct an independent 
+replication" [@nosekPromotingOpenResearch2015a].
+In this context, 
+reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
+and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
+These guidelines are: 1) Comprehensive
 citation of literature, data, materials, and methods; 2) sharing data, 3)
@@
 pre-registration of studies before data collection, and 7) pre-registration of
-the analysis plan prior to analysis; and 8) replication of published results.
-WORCS defines the goals of open science in terms of these guidelines, and is
-designed to facilitate meeting each of these guidelines, with one exception: We
-do not address replication of published results, because replication relates to
-the subject of a research project, not to its organization or execution.
-We group the remaining seven guidelines into three categories: citation (1),
+the analysis plan prior to analysis.
+The eighth criterion is "not formally a transparency standard for authors", but "addresses 
+journal guidelines for consideration of independent replications for publication".
+WORCS defines the goals of open science in terms of the first seven TOP-guidelines,
+and the workflow and its `R` implementation are designed to facilitate compliance therewith.
+We group these guidelines into three categories: citation (1),
 sharing (2-5), and preregistration (6-7).
@@
 
-WORCS encourages and simplifies the adoption of Open Science principles in daily scientific work.
+WORCS encourages and simplifies the adoption of open science principles in daily scientific work.
 It helps researchers make all research output created throughout the scientific process - not just manuscripts, but data, code, and methods - open and publicly assessible.

### 2020-11-02 paper/reviews_ds.Rmd
R1 C4 fix reference list

@@
 
-\RC{There is no user evaluation, and therefore all the claims about worcs are unsupported by evidence. If the authors had included a compelling evaluation I would be in favor of accepting the paper.}
+\RC{C1 There is no user evaluation, and therefore all the claims about worcs are unsupported by evidence.
+If the authors had included a compelling evaluation I would be in favor of accepting the paper.}
 
@@
 
-\RC{This paper describes worcs, a package and workflow to help make R code reproducible according to the TOP guidelines.
+\RC{C2 This paper describes worcs, a package and workflow to help make R code reproducible according to the TOP guidelines.
 
@@
 
-\RC{Unsupported claims (beyond the ones derived from an absence of evaluation): The paper claims in many parts that worcs helps make an entire research project Open and Reproducible.
+\RC{C3 Unsupported claims (beyond the ones derived from an absence of evaluation): The paper claims in many parts that worcs helps make an entire research project Open and Reproducible.
 However, at the same time there is also a claim for not supporting replication.
@@
 
-Response: We appreciate the Reviewer's comment, which seems to be rooted in a different definition of the terms "replication" and "reproducibility" than was intended by us.
-The lack of consistent definitions for these terms has been lamented by other authors (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018).
-As WORCS aims to meet the requirements set out in the TOP-guidelines, we adhere to the definitions of reproducibility and replication as used in the TOP-guidelines.
-This definition is formalized by Patil and colleagues: "We define reproducibility as re-performing the same analysis with the same code using a different analyst, and we define replicability as re-performing the experiment and collecting new data".
-The TOP-guidelines state that the first seven guidelines are intended to ensure the "reproducibility of the reported results based on the originating data".
-The eighth criterion is "not formally a transparency standard for authors, [but] addresses journal guidelines for consideration of independent replications for publication."
+Response: We understand that our writing comes across as contradictory,
+and this apparent contradiction arises from a lack of explicit definitions of the terms "replication" and "reproducibility" in the original manuscript.
+We are aware of the oft-lamented lack of consistent definitions for these terms (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018).
+In the Revision, we have attempted to address this lack of clarity by including explicit definitions for these terms:
 
-We had taken notice of this definitional quagmire, but decided not to dwell on it in the present paper.
-This reviewer comment has made it abundantly clear that we need to explicitly define our terms.
-Thus, in order to address this comment, the revised manuscript now reads: REF.
+> Throughout this paper, we adhere to the definition of open science as
+"the practice of science in such a way that others can collaborate and contribute,
+where research data, lab notes and other research processes are freely available,
+under terms that enable reuse, redistribution and reproduction of the research and its underlying data and methods" [@vicente-saezOpenScienceNow2018].  
+[...] reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
+and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
 
-We wish to fully address this comment, including the criticism of unsupported claims - but it is a bit difficult to do so, given the fundamental misunderstanding about what constitutes "reproducibility".
+After making these definitions explicit, it becomes clear that the Reviewer's definition of replication as "a necessary steps towards reproducibility" is not consistent with the present paper.
+
+We further wish to clarify that the fact that we address guidelines 1-7, but not guideline 8 ("replication") is consistent with the intended purpose of the TOP-guidelines.
+Specifically, the first seven guidelines are intended to ensure the "reproducibility of the reported results based on the originating data" [direct quote from @nosekPromotingOpenResearch2015a].
+The eighth criterion, by contrast, is "not formally a transparency standard for authors, [but] addresses journal guidelines for consideration of independent replications for publication" (again a direct quote).
+To avoid further confusion or controversy, we now address this distinction between the first 7 guidelines and the final 8th explicitly in the revised manuscript:
+
+> These guidelines describe eight standards for open science,
+the first seven of which "account for reproducibility of the reported results
+based on the originating data, and for sharing sufficient information to conduct an independent 
+replication" [@nosekPromotingOpenResearch2015a].
+In this context, 
+reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
+and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
+These guidelines are: 1) Comprehensive
+citation of literature, data, materials, and methods; 2) sharing data, 3)
+sharing the code required to reproduce analyses, 4) sharing new research
+materials, and 5) sharing details of the design and analysis; 6)
+pre-registration of studies before data collection, and 7) pre-registration of
+the analysis plan prior to analysis.
+The eighth criterion is "not formally a transparency standard for authors", but "addresses 
+journal guidelines for consideration of independent replications for publication".
+WORCS defines the goals of open science in terms of the first seven TOP-guidelines,
+and the workflow and its `R` implementation are designed to facilitate compliance therewith.
+
+<!--We wish to fully address this comment, including the criticism of unsupported claims - but it is a bit difficult to do so, given the fundamental misunderstanding about what constitutes "reproducibility".
 We have stated that worcs helps make an entire research project open and reproducible.
@@
 The Reviewer, by contrast, proposes that the main contribution of worcs is to help documentation and understanding.
-It is not immediately clear to us which components of worcs address these two points - except perhaps the readme.md that is created upon initialization of a new project.
+It is not immediately clear to us which components of worcs address these two points - except perhaps the readme.md that is created upon initialization of a new project.-->
+
 \Assignment{CJ, ALL}
-\WorkInProgress
+\Done
 \Hard
@@
 
-Response: Citation should be formatted as software.
+Response: We apologize; it appears that there were numerous errors in the parsing of the reference list after rendering the manuscript to HTML - the preferred format of Data Science.
+The reference is indeed to the R-package.
+In the revised manuscript, we have repaired the references so that it is clear that this is a software package, et cetera.
 

### 2020-11-02 paper/Manuscript_DS.Rmd
R1 C5 rmarkdown reference

@@
 implemented in `R` [@rcoreteamLanguageEnvironmentStatistical2020].
-The `worcs` package [@worcs] offers a project template for RStudio
+The `worcs` package [@worcspackage2020] offers a project template for RStudio
 [@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
@@
 In the `R` implementation of WORCS, we recommend centering a research project
-around one dynamically generated `RMarkdown` document, which includes all 
-analysis code. Longer scripts can be
-stored in `.R` files, and called from the main document using the `source()`
-function. When a reader or reviewer compiles this document, all code is run
+around one dynamically generated `RMarkdown` document [@xieMarkdownDefinitiveGuide2018].
+RMarkdown is based on [`Markdown`](https://en.wikipedia.org/wiki/Markdown),
+which uses plain-text commands to typeset a document.
+It enhances this format with support for R-code, and is compatible with [LaTeX](https://www.latex-project.org/) math typesetting.
+For a full overview of the functionality of `RMarkdown`, see @xieMarkdownDefinitiveGuide2018.
+
+The RMarkdown document can incorporate blocks of analysis code,
+which can be used to render results directly to [Tables](https://bookdown.org/yihui/rmarkdown-cookbook/kable.html) and [Figures](http://zevross.com/blog/2017/06/19/tips-and-tricks-for-working-with-images-and-figures-in-r-markdown-documents/),
+which can be [dynamically referenced](https://bookdown.org/yihui/rmarkdown-cookbook/cross-ref.html) in the text.
+Long scripts can be stored in `.R` files,
+and included in the main document using the [`read_chunk()`
+function](https://github.com/yihui/knitr-examples/blob/master/113-externalization.Rmd) from the `knitr` package, which comes together with the `rmarkdown` package.
+When a reader or reviewer compiles this document, all code is run
 automatically, thus verifying computational replicability. The `RMarkdown` 

### 2020-11-02 paper/reviews_ds.Rmd
R1 C5 rmarkdown reference

@@
 
-\RC{I am confused by the citation "Van Lissa, Caspar J., Aaron Peikert, and Andreas M.
+\RC{C4: I am confused by the citation "Van Lissa, Caspar J., Aaron Peikert, and Andreas M.
 Brandmaier.
@@
 
-\RC{If the paper is generated as DDG, where are the functions that show it's capabilities?
+\RC{C5: If the paper is generated as DDG, where are the functions that show it's capabilities?
 It is also not clear to me that some of the functionalities of Latex would be usable in RMarkdown, such as all the equations.}
 
-Response: Add comment that LaTeX is supported in RMarkdown; showing functions (?) out of scope for present paper. Reference RMarkdown @xieMarkdownDefinitiveGuide2018
+Response: We understand that our introduction of the RMarkdown format was a bit too brief, and that readers may wonder whether specific functionalities of LaTeX would be usable in RMarkdown.
+To address this comment, the revised manuscript now reads:
+
+> In the `R` implementation of WORCS, we recommend centering a research project
+around one dynamically generated `RMarkdown` document [@xieMarkdownDefinitiveGuide2018].
+RMarkdown is based on [`Markdown`](https://en.wikipedia.org/wiki/Markdown),
+which uses plain-text commands to typeset a document.
+It enhances this format with support for R-code, and is compatible with [LaTeX](https://www.latex-project.org/) math typesetting.
+For a full overview of the functionality of `RMarkdown`, see @xieMarkdownDefinitiveGuide2018.
 

### 2020-11-03 paper/Manuscript_DS.Rmd
R1 C6 Dependency management

@@
 
-For analyses conducted entirely in `R`, the recently released package `renv` [@usheyRenvProjectEnvironments2020]
-strikes a great balance between user-friendliness and robustness.
+For dependency management, the `R` implementation of `worcs` relies on the recently released package `renv` [@usheyRenvProjectEnvironments2020].
+This package strikes a great balance between user-friendliness and robustness.
 Developed by the team behind RStudio, 
-this package simply maintains a text-based,
+`renv` maintains a text-based,
 human-readable log of all packages used, their version numbers, and where they
 were installed from (e.g., CRAN, Bioconductor, GitHub).
+This text-based log file can be version controlled with Git.
+When someone else loads your project,
+`renv` will install all of the required packages from this list.
 This lightweight approach safeguards reproducibility as long as these repositories maintain their archives.
-The text-based log file can be version controlled with Git.
-The `renv` function `snapshot()` is used to add packages to the log file.
-This function identifies all packages loaded using `library()` calls in the scripts in the project directory.
+The `worcs` project template in RStudio automatically sets up `renv` for dependency management if the checkbox labeled "renv" is selected during project creation.
 
@@
 This improves the user experience by saving time and memory.
-When someone else loads your project, `renv` will install all of the required packages.
 As opposed to alternative solutions, `renv` is tightly integrated with the
 native R package management system.
-Therefore, it does not require the users to change their workflow of installing/removing and using packages, and requires no technical skill to setup.
+It monitors the scripts in the project directory for `library()` calls,
+and notifies the user when it is necessary to update the dependencies file using the `snapshot()` function.
+Therefore, `renv` does not require the users to change their workflow of installing/removing and using packages, and requires little technical skill to setup.
 It is not a fail-safe method to ensure strict computational reproducibility,
 but it meets most reproducibility requirements using only tools
-native to RStudio, and is therefore easier to adopt for `R` users.
-
-If stricter reproducibility is required,
-users might choose to conduct all analyses in a "Docker container": An environment that behaves like a
+native to `R` and RStudio.
+
+When a project is conducted in a different software environment than `R`,
+or if an R-project has external dependencies not tracked by `renv`,
+then users might choose to use [Docker](https://www.docker.com/) for dependency management.  
+Docker instantiates a "Docker container": An environment that behaves like a
 virtual computer, which can be stored like a time capsule, and
 identically reinstated on a user computer, or in the cloud.
-This is effective, and also preserves software outside the `R` environment (e.g.,
-preprocessing tools for neuroimaging data).
+Although the use of Docker falls outside of the scope of the present paper,
+for existing users, a Docker build running `worcs` and its dependencies is available.
 The [stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html)
-describes how to instantiate a Docker build running `worcs` and its dependencies.
+explains how to instantiate this build.
 For more information on this topic,
@@
 conceptually and practically compatible with WORCS.
-One remaining challenge of Docker is that it is somewhat difficult to set up for novice users.
-To overcome this challenge, an `R` package is currently in development to facilitate Docker integration [`repro`, @repro].
-Another 
-user-friendly Docker-based workflow is offered by the cloud-based collaboration
-platform 'Code Ocean'. Although Code Ocean enjoys some popularity among 
+
+As Docker is somewhat challenging to set up for novice users,
+an `R` package is currently in development to facilitate Docker integration [`repro`, @repro].
+Another example of a Docker-based workflow is the cloud-based collaboration
+platform 'Code Ocean'.
+Although Code Ocean enjoys some popularity among 
 scientists, it is not clear whether this platform
@@
 data are stored.
-The necessity to upload human participant data is a liability
-when using any
-cloud-computing solution for data analysis.
+The necessity to upload human participant data for analysis is an important consideration and potential liability when using any cloud-computing solution.
 

### 2020-11-03 paper/reviews_ds.Rmd
R1 C6 Dependency management

@@
 class       : "draft"
-
+bibliography      : ["worcs.bib"]  
 output      : papaja::revision_letter_pdf
@@
 \Assignment{CJ}
-\WorkInProgress
+\Done
 \Medium
 
-\RC{It is not clear how worcs addresses depedency management. The section lists a series of tools that can be used, but the section is not very specific to worcs.}
+\RC{C6: It is not clear how worcs addresses depedency management. The section lists a series of tools that can be used, but the section is not very specific to worcs.}
 
-Response: Not true; explain that worcs project automatically initializes `renv`.
+Response: We apologize for the apparent lack of clarity of this section. 
+We have restructured and clarified this section to convey the following information:
+
+1) Introduce the general concept of dependency management
+2) Introduce `renv`, the solution for dependency management used by the R-implementation of `worcs`
+3) Acknowledge the more comprehensive approach to dependency management offered by Docker
+
+We have also, in response to another Reviewer comment, created a Docker Hub build that runs `worcs`.
+Therefore, as of this revision, both of the tools discussed in this section (i.e., `renv` and Docker) are directly relevant for WORCS-users, and the section should read less like a "series of tools that can be used".
+If the Reviewer so desires, we would also be willing to cut the brief discussion of Code Ocean - but this had been included at the request of an informal reviewer of this paper (Daniël Lakens).
 
 \Assignment{CJ}
-\WorkInProgress
+\Done
 \Easy
 
-\RC{The workflow of releasing data products on GitHub is not ideal.
+\RC{C7: The workflow of releasing data products on GitHub is not ideal.
 First, because datasets of 100 MBs usually require special handling.
@@
 The authors seem to not describe very well what would happen in data science experiments where the data size is significant.}
+
+BV: Can you please argue for why GitHub is equivalent to Zenodo for this purpose?
+AL: Can you address the reference to FAIR principles in relation to GitHub?
+CJ: I will address "big" data
 
@@
 \WorkInProgress
-\Each
+\Easy
 
@@
 
-# \Reviewer \#3
+# Reviewer \#3
 
@@
 
-# \Reviewer \#4
+# Reviewer \#4
 

### 2020-11-03 paper/Manuscript_DS.Rmd
R1 C8 What does worcs do?

@@
 Version 0.1.5.
+The conceptual workflow recommends actions to take 
+in the different phases of a research project, in order to work openly and transparently
+in compliance with the TOP-guidelines and best practices for open science.
 Note that, although the steps are numbered for reference purposes,
 we acknowledge that the process of conducting research is not always linear.
-For users of the `worcs` `R` package, we highly recommend referring to
+We refer users of the `worcs` implementation in `R` to
 [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
-which is more detailed and updated with future developments.
+which describes in more detail which steps are automated by the `R` package
+and which actions must be taken by the user.
+The vignette is also updated with future developments of the package.
+
 For an up-to-date list of sample `worcs` projects, see the [WORCS GitHub page](https://github.com/cjvanlissa/worcs).

### 2020-11-03 paper/reviews_ds.Rmd
R1 C8 What does worcs do?

@@
 
-\RC{In the 3 phases listed, it is not clear what worcs does and does not for users.}
+\RC{C8: In the 3 phases listed, it is not clear what worcs does and does not for users.}
 
-\Assignment{All}
-\WorkInProgress
+Response: It is not entirely clear to us whether the Reviewer is referring to the conceptual workflow ("WORCS"), or to the `R` implementation of WORCS, i.e., `worcs`.
+
+We suspect that the Reviewer means the `worcs` R-package, so to address this comment,
+the Revision now states the following:
+
+> For users of the `worcs` `R` package, we highly recommend referring to
+[the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
+which describes in more detail which steps are automated by the `R` package
+and which actions must be taken by the user.
+The vignette is also updated with future developments of the package.
+
+Just in case we misunderstood, and the Reviewer actually meant the conceptual workflow WORCS, we have also added a few sentences to clarify the purpose of the conceptual workflow:
+
+> We provide a conceptual outline of the WORCS procedure below, based on WORCS
+Version 0.1.5.
+The conceptual workflow recommends actions to take 
+in the different phases of a research project, in order to work openly and transparently
+in compliance with the TOP-guidelines and best practices for open science.
+Note that, although the steps are numbered for reference purposes,
+we acknowledge that the process of conducting research is not always linear.
+
+\Assignment{CJ}
+\Done
 \Medium

### 2020-11-03 paper/Manuscript_DS.Rmd
R1 C9 Should everyone use R?

@@
   procedure that researchers can follow to make a research project open and
-  reproducible, accompanied by an open source software implementation that 
-  automates these steps.
-  The 
-  purpose of the workflow is to lower the threshold for 
+  reproducible.
+  The purpose of the workflow is to lower the threshold for 
   adoption of open science principles.
@@
   institutions, and funding bodies.
-  This workflow is primarily targeted towards
-  scholars working in R on a private computer,
+  To facilitate widespread adoption, the WORCS principles have been implemented
+  in the R package `worcs`, which offers an RStudio 
+  project template and utility functions for specific workflow steps.
+  This paper introduces the conceptual workflow,
+  discusses how it meets different 
+  standards for open science, and provides examples of the implementation in `R`. 
+  This paper is primarily targeted towards
+  scholars conducting research projects in R on a private computer,
   conducting research that involves academic prose, analysis code,
-  and (optionally) tabular data - but it is flexible enough to accommodate many
-  other scenarios.
-  To facilitate widespread adoption, the
-  WORCS principles are implemented in the R package `worcs`, which offers an
-  RStudio 
-  project template that is, effectively, a "one-click solution" for R users,
-  and utility functions for specific workflow steps.
-  This paper addresses the design of the workflow,
-  discusses how it meets different 
-  standards for open science, and provides examples of the implementation of
-  `worcs` in `R`. The source code for the `R` package and manuscript,
+  and tabular data that does not require specialized computational
+  provisions.
+  However, the workflow is flexible enough to accommodate other scenarios,
+  and offers a sensible starting point for customized solutions.
+  The source code for the `R` package and manuscript,
   and a list of user examples of WORCS projects, are available at 
@@
 challenges and solutions, and to become proficient with new and unfamiliar tools.
-This paper is designed to ease that transition by presenting a simple workflow 
-that meets most requirements open science, and is based on best practices:
+This paper is designed to ease that transition by presenting a simple workflow,
+based on best practices,  
+that meets most requirements open science:
 The *Workflow for Open Reproducible Code in Science* (WORCS).
-WORCS is a lightweight approach to open science and
-computational reproducibility for scholars working in `R` on a private computer,
+
+This paper introduces the conceptual workflow (referred to in upper case, "WORCS"),
+and discusses its underlying principles and how it meets different standards for open science.
+The principles underlying this workflow are universal and largely platform independent.
+We also present a software implementation of WORCS for `R` users [@rcoreteamLanguageEnvironmentStatistical2020]:
+The `R` package `worcs` [referred to in monospace font, @worcspackage2020].
+This package offers a project template for RStudio
+[@rstudioteamRStudioIntegratedDevelopment2015],
+and several convenience functions to automate specific workflow steps.
+This paper is most relevant for scholars using R on a private computer to
 conducting research that involves academic prose, analysis code,
-and (optionally) tabular data - but it is flexible enough to accommodate many
-other scenarios. The workflow (Figure 1) 
+and tabular data.
+For other readers, it can serve as a primer on open science workflows,
+and provide a sensible starting point for a customized solution.
+
+The WORCS workflow (Figure 1) 
 constitutes a step-by-step procedure that researchers can follow to make a 
@@
 fulfill them.
-It can also be used to produce work in accordance with best practices
-in the absence of top-down support or guidelines.
-Although the principles underlying this workflow are universal, WORCS has been
-implemented in `R` [@rcoreteamLanguageEnvironmentStatistical2020].
-The `worcs` package [@worcspackage2020] offers a project template for RStudio
-[@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
-a "one click solution". This paper discusses the principles underlying WORCS,
-introduces the workflow, addresses how WORCS meets best practices for open science,
-and illustrates the use of the `worcs` package.
+It can also be used by individual authors to produce work in accordance with best practices
+in the absence of top-down support or guidelines (i.e., "grass roots" open science).
 
@@
 As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].
-Several arguments support the choice to implement this workflow in `R` and RStudio first, although we encourage developers to port the software to other languages.
+Several arguments support our choice to implement this workflow first in `R` and RStudio,
+although we encourage developers to port the software to other languages.
 First, `R` and all of it extensions are free open source software, which
@@
 `synthetic()`, which by default uses random forests to generate a model for
-synthesis (see @wrightRangerFastImplementation2015). For data synthesis to work
+synthesis (see @wrightRangerFastImplementation2015).
+
+Note that for data synthesis to work
 properly, it is essential that the `class` (data type) of variables is defined
 correctly. As of `worcs` version `r worcs:::cran_version()`, numeric, integer, factor, and logical data
-are supported out of the box. Other types of variables should be converted to
-one of these types, or users can use a custom `model_expression` when calling
-`synthetic`.
+are supported out of the box. 
+Other types of variables should be converted to one of these types.
+It is important to consider this when using special data types,
+such as the labeled vectors created by the `haven` package when reading an SPSS file.
+We further caution that the default random forests algorithm is computationally intensive,
+and may be unfeasible for use with "large" datasets.
+Users can provide use a custom `model_expression` and `predict_expression`
+to use a different algorithm when calling `synthetic()`.
+For example, users could use `lm()` as a `model_expression` to use linear regression,
+which preserves linear marginal relationships but can give rise to values out of range of the original data.
+Or users could call `sample()` as a `predict_expression` to bootstrap each variable,
+a very quick solution that maintains univariate distributions but loses all marginal relationships.
 
@@
 publicly available. This function stores the `data` object (such as a `data.frame` or
-`matrix`) in a tabular data file (called `data.csv` by default).
+`matrix`) in a tabular data file.
 It also generates an
@@
 of the raw data.
-The function `closed_data()` also creates a synthetic dataset (called
-`synthetic_data.csv` by default).
+The function `closed_data()` also calls the aforementioned `synthetic()` function to
+create a synthetic dataset.
 The user should push all changed files to the remote repository.

### 2020-11-03 paper/reviews_ds.Rmd
R1 C9 Should everyone use R?

@@
 
-\RC{In some cases it looks like the authors claim that everyone should use R for open science.
+\RC{C9: In some cases it looks like the authors claim that everyone should use R for open science.
 While R has a strong community behind, I believe there are others with very strong support (e.g.
@@
 
-Response: Clearly define scope; feedback is welcome.
+Response: We wish to emphasize that it was certainly unintended to claim that everyone should us R for open science.
+To address this comment, we have done two things: First, throughout the paper, we have attempted to clarify the distinction between the platform independent conceptual workflow "WORCS", and its implementation for R users, `worcs`.
+Second, throughout the paper, we encourage the implementation of WORCS in other languages.
+Most illustrative of these changes is the start of the paragraph on **The R implementation of WORCS**:
 
-\Assignment{All}
-\WorkInProgress
+> WORCS is, first and foremost, a conceptual workflow that could be implemented in any software environment.
+As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].
+Several arguments support our choice to implement this workflow first in `R` and RStudio, although we encourage developers to port the software to other languages.
+
+\Assignment{CJ}
+\Done
 \Hard

### 2020-11-03 paper/Manuscript_DS.Rmd
R1 C10 Fix url

@@
 
-The `worcs` package includes a [vignette on citation](https://cjvanlissa.github.io/worcs/articles/citing.html) to explain the
+The `worcs` package includes a [vignette on citation](https://cjvanlissa.github.io/worcs/articles/citation.html) to explain the
 fundamentals for novice users, and offer recommendations specific to `worcs`. We

### 2020-11-03 paper/reviews_ds.Rmd
R1 C10 Fix url

@@
 
-\RC{Vignete on citation leads to a 404.}
+\RC{C10: Vignete on citation leads to a 404.}
+
+Response: We apologize for the oversight; we have fixed this URL and checked all URLs in the manuscript for mistakes (none found).
+
+<!--
+check_urls <- function(x){
+  #x = "C:/Git_Repositories/worcs/paper/manuscript_ds.Rmd"
+  txt <- readLines(x)
+  mtch <- regmatches(txt,regexpr("(http[^ ]*)|(ftp[^ ]*)|(www\\.[^ ]*)",txt))
+  mtch <- gsub("[\\)\\.,;]+$", "", mtch)
+  #out <- sapply(mtch, RCurl::url.exists)
+  out <- sapply(mtch, httr::http_error)
+}
+-->
 
 \Assignment{CJ}
-\WorkInProgress
+\Done
 \Easy

### 2020-11-03 paper/reviews_ds.Rmd
streamlined display of bullet points make worcs render as package name or WORCS

@@
 First, there are several arguments to substantiate the potential for reuse of worcs:
-1) Worcs is implemented in an R-package
-2) This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network")
-3) The software meets all criteria of the Core Infrastructure Initiative "best practices" (CII Best Practices)
-4) The R-package uses only tools that are fully integrated within the RStudio IDE, thus enhancing ease of use
-5) The R-package is compatible with other reproducibility solutions, including Docker, drake, and make
+
+1. Worcs is implemented in an R-package
+2. This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network")
+3. The software meets all criteria of the Core Infrastructure Initiative "best practices" (CII Best Practices)
+4. The R-package uses only tools that are fully integrated within the RStudio IDE, thus enhancing ease of use
+5. The R-package is compatible with other reproducibility solutions, including Docker, drake, and make
 
@@
 1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
-1. The lead author has applied worcs in several research projects:
+1. The lead author has applied `worcs` in several research projects:
 	* https://github.com/cjvanlissa/storm_meta_father_mother
@@
 Third, although we have not conducted a user evaluation, there is some evidence that worcs addresses users' needs:
-1) Evaluate worcshops!!
-2) The issues page on the worcs GitHub profile demonstrates that we have been very responsive in addressing users' feedback (as required by the aforementioned CII Best Practices):
+
+1. Evaluate worcshops!!
+2. The issues page on the worcs GitHub profile demonstrates that we have been very responsive in addressing users' feedback (as required by the aforementioned CII Best Practices):
 * https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed

### 2020-11-03 paper/reviews_ds.Rmd
minor changes

@@
 
-1. Worcs is implemented in an R-package
+1. WORCS is implemented in an R-package
 2. This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network")
@@
 
-Second, there is some evidence to demonstrate that worcs is already being adopted (as of 31-10-2020):
+Second, there is some evidence to demonstrate that WORCS is already being adopted (as of 31-10-2020):
 
@@
 
-Third, although we have not conducted a user evaluation, there is some evidence that worcs addresses users' needs:
+Third, although we have not conducted a user evaluation, there is some evidence that `worcs` addresses users' needs:
 
@@
 Response: In addressing this comment, we respectfully refer to our response to the preceding comment ("There is no user evaluation", R1), and again emphasize the Editor's comment that "resource papers do not need to have a full-blown user evaluation".
-We would also like to clarify that the goal of the present manuscript is to define the problem worcs intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools worcs uses to address this problem, explain how worcs uses these tools, and provide a step-by-step workflow.
+We would also like to clarify that the goal of the present manuscript is to define the problem WORCS intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools WORCS uses to address this problem, explain how WORCS uses these tools, and provide a step-by-step workflow.
 It is not clear what "claims" are made that are unsupported by evidence.
@@
 We do take the suggestion of performing a user evaluation very seriously, but in our mind, this is a next step.
-First, the workflow must be published, then it can be taught, and subsequently evaluated.
+We have built WORCS based on established tools and best practices, and we have gathered preliminary positive feedback from the community through workshops and teaching.
 
@@
 We are aware of the oft-lamented lack of consistent definitions for these terms (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018).
-In the Revision, we have attempted to address this lack of clarity by including explicit definitions for these terms:
+In the revision, we have attempted to address this lack of clarity by including explicit definitions for these terms:
 

### 2020-11-03 paper/reviews_ds.Rmd
I think it is important to stress that not only the same code but the same data are used. I see the main difference in whether the same or new data are used.

@@
 under terms that enable reuse, redistribution and reproduction of the research and its underlying data and methods" [@vicente-saezOpenScienceNow2018].  
-[...] reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
+[...] reproducibility can be defined as "re-performing the same analysis with the same code and the same data using a different analyst",
 and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].

### 2020-11-03 paper/reviews_ds.Rmd
suggestion for responding to C7

@@
 
-After making these definitions explicit, it becomes clear that the Reviewer's definition of replication as "a necessary steps towards reproducibility" is not consistent with the present paper.
+After making these definitions explicit, it becomes clear that the Reviewer's definition of replication as "a necessary steps towards reproducibility" is not consistent with the definitions of the present paper.
 
@@
 In this context, 
-reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
+reproducibility can be defined as "re-performing the same analysis with the same code and the same data using a different analyst",
 and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
@@
 The reference is indeed to the R-package.
-In the revised manuscript, we have repaired the references so that it is clear that this is a software package, et cetera.
+In the revised manuscript, we have repaired the references so that it is clear that this is a software package.
 
@@
 The authors seem to not describe very well what would happen in data science experiments where the data size is significant.}
+
+We agree that GitHub is not an ideal platform for handling large data.
+However, it is ideal for handling version control of computer code, Markdown documents, and small to medium-sized data files.
+Larger files can be linked to git using git-annex (https://git-annex.branchable.com) and similar solutions.
+We now mention git-annex as a possible route to handle larger data files (TODO).
+We agree that it is important to make open data findable beyond just providing it freely.
+Digital object identifiers are the de-factor standard for making unique references to data sets.
+In line with the reviewer's suggestion, we suggest to use external providers to provide such DOIs while github is still the primary data storage.
+Our prefered route is to use github web hooks to link Zenodo and a WORCS-compatible repository.
+With this technique, Zenodo issues DOI for content hosted on github repositories including some nice extras, such as automatic badge DOI badges for the repository.
 

### 2020-11-03 paper/reviews_ds.Rmd
Fix function

--- 2020-11-03 paper/reviews_ds.Rmd
@@
 First, there are several arguments to substantiate the potential for reuse of worcs:
+1) Worcs is implemented in an R-package
+2) This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network")
+3) The software meets all criteria of the Core Infrastructure Initiative "best practices" (CII Best Practices)
+4) The R-package uses only tools that are fully integrated within the RStudio IDE, thus enhancing ease of use
+5) The R-package is compatible with other reproducibility solutions, including Docker, drake, and make
 
-1. WORCS is implemented in an R-package
-2. This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network")
-3. The software meets all criteria of the Core Infrastructure Initiative "best practices" (CII Best Practices)
-4. The R-package uses only tools that are fully integrated within the RStudio IDE, thus enhancing ease of use
-5. The R-package is compatible with other reproducibility solutions, including Docker, drake, and make
-
-Second, there is some evidence to demonstrate that WORCS is already being adopted (as of 31-10-2020):
+Second, there is some evidence to demonstrate that worcs is already being adopted (as of 31-10-2020):
 
@@
 1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
-1. The lead author has applied `worcs` in several research projects:
+1. The lead author has applied worcs in several research projects:
 	* https://github.com/cjvanlissa/storm_meta_father_mother
@@
 
-Third, although we have not conducted a user evaluation, there is some evidence that `worcs` addresses users' needs:
-
-1. Evaluate worcshops!!
-2. The issues page on the worcs GitHub profile demonstrates that we have been very responsive in addressing users' feedback (as required by the aforementioned CII Best Practices):
+Third, although we have not conducted a user evaluation, there is some evidence that worcs addresses users' needs:
+1) Evaluate worcshops!!
+2) The issues page on the worcs GitHub profile demonstrates that we have been very responsive in addressing users' feedback (as required by the aforementioned CII Best Practices):
 * https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed
@@
 Response: In addressing this comment, we respectfully refer to our response to the preceding comment ("There is no user evaluation", R1), and again emphasize the Editor's comment that "resource papers do not need to have a full-blown user evaluation".
-We would also like to clarify that the goal of the present manuscript is to define the problem WORCS intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools WORCS uses to address this problem, explain how WORCS uses these tools, and provide a step-by-step workflow.
+We would also like to clarify that the goal of the present manuscript is to define the problem worcs intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools worcs uses to address this problem, explain how worcs uses these tools, and provide a step-by-step workflow.
 It is not clear what "claims" are made that are unsupported by evidence.
@@
 We do take the suggestion of performing a user evaluation very seriously, but in our mind, this is a next step.
-We have built WORCS based on established tools and best practices, and we have gathered preliminary positive feedback from the community through workshops and teaching.
+First, the workflow must be published, then it can be taught, and subsequently evaluated.
 
@@
 We are aware of the oft-lamented lack of consistent definitions for these terms (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018).
-In the revision, we have attempted to address this lack of clarity by including explicit definitions for these terms:
+In the Revision, we have attempted to address this lack of clarity by including explicit definitions for these terms:
 
@@
   out <- sapply(mtch, httr::http_error)
+  out
 }

### 2020-11-03 paper/Manuscript_DS.Rmd
R1 C11: Synthetic data

@@
 synthesis (see @wrightRangerFastImplementation2015).
+This function is based on the work by @nowokSynthpopBespokeCreation2016 (see the `R` package `synthpop`),
+but with a simpler and more flexible user interface.
+Specifically, the `worcs` function `synthetic()` accepts any function that can be used to model
+marginal associations in the data.
+For further details, see the [function documentation](https://cjvanlissa.github.io/worcs/reference/synthetic.html).
 

### 2020-11-03 paper/reviews_ds.Rmd
R1 C11: Synthetic data

@@
 
-\RC{Synthetic data generation is fantastic, but it is not clear to me how does worcs actually keep the data consistent e.g.
+\RC{C11: Synthetic data generation is fantastic, but it is not clear to me how does worcs actually keep the data consistent e.g.
 for training models.
@@
 
-Response: Provide reference, clarify explanation in text.
+Response: We agree that synthetic data generation is exciting, but feel obliged to clarify that this is **not** a novel contribution of the present paper nor R-package. 
+As indicated in the paper, we use the algorithm by Nowok and colleagues,
+who explain the details of the method.
+We re-implemented the method in the `worcs` package because the `synthpop` package
+is rarely maintained (e.g., one of our pull requests took almost 10 months to be accepted),
+and because we wanted a more flexible, customizable interface.
+
+To answer the Reviewer's specific question, the synthetic data indeed follow similar distributions to the source data, as indicated in the paper:
+
+> Synthetic data mimic the level of measurement and
+(conditional) distributions of the real data
+[see @nowokSynthpopBespokeCreation2016].
+
+We have additionally clarified the paragraph that introduces the `synthetic()` function, and added further references to the methodological paper,
+and to the documentation of the `synthetic()` function (which summarizes the algorithm).
 

### 2020-11-03 R/synthetic.R

@@ -27,7 +27,8 @@
 #' @param verbose Logical, Default: TRUE. Whether to show a progress bar while
 #' running the algorithm and provide informative messages.
 #' @return A \code{data.frame} with synthetic data, based on \code{data}.
-#' @details This function uses a simple algorithm to generate a synthetic
+#' @details Based on the work by Nowok, Raab, and Dibben (2016),
+#' this function uses a simple algorithm to generate a synthetic
 #' dataset with similar characteristics to the original. The algorithm is as
 #' follows:
 #' \enumerate{
@@ -47,15 +48,76 @@
 #' \code{class} of variables is defined correctly. The default algorithm
 #' \code{\link[ranger]{ranger}} supports numeric, integer, and factor types.
 #' Other types of variables should be converted to one of these types, or users
-#' can use a custom \code{model_expression} when calling \code{synthetic}.
+#' can use a custom \code{model_expression} and \code{predict_expressio}
+#' when calling \code{synthetic}.
+#'
+#'
+#' Note that for data synthesis to work properly, it is essential that the
+#' \code{class} of variables is defined correctly. The default algorithm
+#' \code{\link[ranger]{ranger}} supports numeric, integer, factor, and logical
+#' data. Other types of variables should be converted to one of these types.
+#'
+#' Users can provide use a custom \code{model_expression} and
+#' \code{predict_expression} to use a different algorithm when calling
+#' \code{synthetic}.
+#'
+#' As demonstrated in the example, users could call \code{lm} as a
+#' \code{model_expression} to use
+#' linear regression, which preserves linear marginal relationships but can give
+#' rise to values out of range of the original data.
+#' Or users could call \code{sample} as a \code{predict_expression} to bootstrap
+#' each variable, a very quick solution that maintains univariate distributions
+#' but loses all marginal relationships. These examples are not exhaustive, and
+#' users can even create custom functions.
 #' @examples
+#' # Example using the iris dataset and default ranger algorithm
 #' iris_syn <- synthetic(iris)
+#'
+#' # Example using lm as prediction algorithm (only works for numeric variables)
+#' # note that, within the model_expression, a new data.frame is created because
[…]

### 2020-11-03 paper/reviews_ds.Rmd
R1 C12: Add docker image

@@
 We re-implemented the method in the `worcs` package because the `synthpop` package
-is rarely maintained (e.g., one of our pull requests took almost 10 months to be accepted),
+is rarely maintained (e.g., [one of our pull requests](https://github.com/bnowok/synthpop/pull/12) took almost 10 months to be accepted),
 and because we wanted a more flexible, customizable interface.
@@
 \Assignment{CJ}
-\WorkInProgress
+\Done
 \Easy
 
-\RC{Docker is deemed as complicated for novice users, but the setup vignete looks quite complex to me as well.
+\RC{C12: Docker is deemed as complicated for novice users, but the setup vignete looks quite complex to me as well.
 Wouldn't a Docker container with worcs installed be easier to use?
@@
 
-Response: Thank you for this thoughtful suggestion.
-We explained better why `renv` is more user friendly (because it builds upon the "normal" package workflow of R users).
-For users familiar with Docker, we have published an image on Docker Hub with all requirements of worcs installed and a Docker specific setup vignette, that is much shorter and given that you know how to use Docker easier.
+Response: We thank the Reviewer for this thoughtful suggestion.
+To address this comment, 
+we have published an image on Docker Hub with all requirements of `worcs` installed,
+and added a Docker-specific setup vignette to the package and website, which is referenced in the Setup pagragraph:
+
+> Before you can use the `R` implementation of the WORCS workflow,
+you have to install the required software.
+This 30 minute procedure is documented
+in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
+After setting up your system, you can use `worcs` for all of your
+projects.
+Alternatively, Docker users can follow [the stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html).
+
 \Assignment{AP, AB}
-\WorkInProgress
+\Done
 \Medium

### 2020-11-03 paper/reviews_ds.Rmd
R2 needs work

@@
 # Reviewer \#2
+
 Submitted on 19/Oct/2020 by Remzi Celebi ORCID logo https://orcid.org/0000-0001-7769-4272
@@
 
-\RC{I would suggest the authors include a flowchart for users that shows the decision making process for which tools/operations should be used in WORCS.
+\RC{C1: I would suggest the authors include a flowchart for users that shows the decision making process for which tools/operations should be used in WORCS.
 An example project would also help users better understand the package.}
@@
 
+Anna-Lena?
+
 \Assignment{AL, BV, CJ}
@@
 
-Response: What would this entail? Is it within scope?
+Response: It does not. It seems to us that this is outside of the scope of the present paper.
+If a compelling case could be made by the Reviewer to include such a service,
+we are open to the suggestion - but we would need to take on additional coauthors with the required expertise.
+<!--What would this entail? Is it within scope?-->
 

### 2020-11-03 paper/Manuscript_DS.Rmd
Some touches

@@
 Firstly, although the workflow is currently implemented only in `R`, it is 
-conceptually relevant for users of other statistical programming languages, most
-notably `Python` [@10.5555/1593511].
+conceptually relevant for users of other statistical programming languages, and we welcome efforts to implement WORCS in other platforms.
 We welcome efforts to implement WORCS in other platforms.

### 2020-11-03 paper/reviews_ds.Rmd
Some touches

@@
 
-\RC{C1) The process as described by the author for conducting research is assumed to be linear.
+\RC{C1: The process as described by the author for conducting research is assumed to be linear.
 However, researchers often go back and forth on steps that they have already completed, and sometimes have to undo them completely and start from scratch.
@@
 To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in the paper and the Workflow vignette, that the steps are not necessarily linear:  *"Note that, although the steps are numbered for reference purposes, we acknowledge that the process of conducting research is not always linear."*.
-Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have now added functions to increase the non-linear flexibility of the `worcs` package: `add_manuscript()` and `add_preregistration()`.
+Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have added more functions to increase the non-linear flexibility of the `worcs` package: `add_manuscript()` and `add_preregistration()`.
 
@@
 
-\RC{C2) The gist of the tool seems to allow researchers to document their steps and generate a scholarly document that is ready for submission.
+\RC{C2: The gist of the tool seems to allow researchers to document their steps and generate a scholarly document that is ready for submission.
 In a sense, there is a similarity with existing notebooks such as Jupyter.
@@
 Response: There are two points we wish to address here.
-First, we would like to clarify that this paper introduces WORCS, a conceptual workflow for open science research projects.
-When the Reviewer refers to "the tool" - we assume this to mean the `worcs` R-package, which offers convenience functions to help R-users follow the WORCS-workflow.
-This is clarified in the Abstract and in the paper, e.g.:
+First, the Reviewer focuses on "the tool" - which we assume refers to the R package.
+However, the main goal of this paper is to introduce the conceptual workflow.
+The R package is merely offered to make adoption of the conceptual workflow easier for R users.
+This distinction is clarified several places throughout the manuscript, and to prevent further confusion, we have clarified the distinction even further in the revision, e.g.:
 
-> Abstract: *This paper introduces the Workflow for Open Reproducible Code in Science (WORCS): A step-by-step procedure that researchers can follow to make a research project open and reproducible, accompanied by an open source software implementation that automates these steps.*
+Abstract:
 
-> Manuscript, paragraph titled **The R implementation of WORCS**: *WORCS is, first and foremost, a conceptual workflow that could be implemented in any software environment.
-As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].*
+> This paper introduces the Workflow for
+  Open Reproducible Code in Science (WORCS): A step-by-step
+  procedure that researchers can follow to make a research project open and
+  reproducible.  
+  [...]  
+  To facilitate widespread adoption, the WORCS principles have been implemented
+  in the R package `worcs`, which offers an RStudio 
+  project template and utility functions for specific workflow steps.
 
-Second, in response to the main comment, we agree with the Reviewer that the conceptual workflow as explained in the paper could be implemented in any software environment, including Jupyter.
-In fact, we emphasize this in the paragraph on *Future developments*: *"although the workflow is currently implemented only in `R`, it is conceptually relevant for users of other statistical programming languages, most notably `Python` [@10.5555/1593511].
-We welcome efforts to implement WORCS in other platforms."*.
+Manuscript, introduction:
+
+> This paper introduces the conceptual workflow (referred to in upper case, "WORCS"),
+and discusses its underlying principles and how it meets different standards for open science.
+The principles underlying this workflow are universal and largely platform independent.
+We also present a software implementation of WORCS for `R` users [@rcoreteamLanguageEnvironmentStatistical2020]:
+The `R` package `worcs` [referred to in monospace font, @worcspackage2020].
+
+Manuscript, paragraph titled **The R implementation of WORCS**:
+
+> WORCS is, first and foremost, a conceptual workflow that could be implemented in any software environment.
+As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].
+
+Second, in response to the question "why don’t the authors start from an existing popular tool such as Jupyter",
+<!--we are sure the Reviewer does not mean to imply that R is not a popular tool,
+or that R users should switch to Jupyter if they wish to make their research open and reproducible.
+To provide some context, on [ITJobsWatch](https://www.itjobswatch.co.uk),
+R is the 216th highest ranked job requirement, and Jupyter is 648th.-->
+we completely agree with the Reviewer that the conceptual workflow as explained in the paper could be implemented in any software environment, including Jupyter.
+In fact, we emphasize this in several places throughout the manuscript, for instance the paragraph titled **The R Implementation of WORCS**:
+
+> WORCS is, first and foremost, a conceptual workflow that could be implemented in any software environment.
+As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].
+Several arguments support our choice to implement this workflow first in `R` and RStudio,
+although we encourage developers to port the software to other languages.
+
+And in **Future developments**:
+
+> although the workflow is currently implemented only in `R`, it is conceptually relevant for users of other statistical programming languages, and we welcome efforts to implement WORCS in other platforms."
 
@@
 
-\RC{C3) Regarding Version control and dependency documentation, the authors uses in their system existing solution, namely git and env, respectively.
+\RC{C3: Regarding Version control and dependency documentation, the authors uses in their system existing solution, namely git and env, respectively.
 While I did not criticize reusing existing solution, on the contrary, I don't believe that simply using such solution add a substantial value.
@@
 Second, the Reviewer claims that using e.g.
-Git does not add substantial value - but this claim is in disagreement with published literature that makes a strong case for the value of using Git for research [see two references cited in the paper, @ramGitCanFacilitate2013;  @blischakQuickIntroductionVersion2016].
+Git does not add substantial value - but this claim is in disagreement with published literature (which we cite) that makes a strong case for the value of using Git for research [see @ramGitCanFacilitate2013;  @blischakQuickIntroductionVersion2016].
 
-Third, the Reviewer remarks that 
+Third, there seems to be some misunderstanding of the workings of existing solutions.
+For example, the Reviewer remarks that 
 "git capture changes at the level of the line of code/text, whereas a researcher needs an abstraction that informs him/her on the coarse-grained elements [...] so as to have a clear idea of where s/he is within the process and help him/her make the decision of the next steps to undertake".
@@
 
-We do see the value in clarifying this point in the manuscript as well, and to address this comment, the manuscript now reads:
+The Reviewer's comment does illustrate the importance of clarifying this point in the manuscript as well
+To address this comment, the revised manuscript now reads:
 

### 2020-11-04 paper/reviews_ds.Rmd
R1 C7 split up

@@
 
-\RC{C7: The workflow of releasing data products on GitHub is not ideal.
+\RC{C7a: The workflow of releasing data products on GitHub is not ideal.
 First, because datasets of 100 MBs usually require special handling.
-Second, it doesn't follow very well the FAIR principles for findability and attribution of data.
-Instead, repositories like Zenodo or FigShare could be user to address data storage, with proper metadata and its corresponding citation.
-The authors seem to not describe very well what would happen in data science experiments where the data size is significant.}
+[...] The authors seem to not describe very well what would happen in data science experiments where the data size is significant.}
 
-We agree that GitHub is not an ideal platform for handling large data.
-However, it is ideal for handling version control of computer code, Markdown documents, and small to medium-sized data files.
-Larger files can be linked to git using git-annex (https://git-annex.branchable.com) and similar solutions.
+We agree that GitHub is not an ideal platform for handling data larger than 100MB.
+However, 100MB is very large for tabular data; for example, if only integer numbers are stored that is $(100MB*1024^2)/(2^2) = 26214400$ unique values.
+
+GitHub is, however, ideal version controlling analysis code, RMarkdown documents, and small to medium-sized data files.
+Larger files can be linked to git using [git-annex](https://git-annex.branchable.com),
+[Git Large File Storage](https://git-lfs.github.com/), and similar solutions. 
 We now mention git-annex as a possible route to handle larger data files (TODO).
+
 We agree that it is important to make open data findable beyond just providing it freely.
-Digital object identifiers are the de-factor standard for making unique references to data sets.
-In line with the reviewer's suggestion, we suggest to use external providers to provide such DOIs while github is still the primary data storage.
+Digital object identifiers are the de-factor standard for making unique references to data sets. In line with the reviewer's suggestion,
+we suggest to use external providers to provide such DOIs while github is still the primary data storage. TODO!!!
+
 Our prefered route is to use github web hooks to link Zenodo and a WORCS-compatible repository.
@@
 
+on second thought, it’s probably worth mentioning to the reviewer that we know these options but defend the OSF-centric approach for our purposes.
+My impression is that the main challenge are the different expectations and solutions across disciplines that are currently evolving largely in parallel.
+If we acknowledge that there are alternatives, this may please the reviewers while we still can defend R + OSF for the behavioral and social sciences.
+Other disciplines may have different foci, e.g., python + jupyter + docker etc...
+and no single manuscript will be able to cover everything.
+
 \Assignment{BV, AB, AL, CJ}
@@
 \Hard
+
+\RC{C7b: Second, it doesn't follow very well the FAIR principles for findability and attribution of data.
+Instead, repositories like Zenodo or FigShare could be user to address data storage, with proper metadata and its corresponding citation.}
 

### 2020-11-05 paper/Manuscript_DS.Rmd
explain why worcs is easier

@@
 even the most sophisticated analyses.
-There are some key differences with WORCS.
-First, with regard to the scope of the workflow: 
+There are some decisive differences with WORCS.
+First, concerning the scope of the workflow: 
 WORCS is designed to address a unique issue not covered by other existing solutions,
-namely to provide a workflow most conducive to satisfying TOP-guidelines 1-7 while adhering to the FAIR principles.
+namely, to provide a workflow most conducive to satisfying TOP-guidelines 1-7 while adhering to the FAIR principles.
 Peikert and Brandmaier instead focus primarily on computational reproducibility,
 which is relevant for TOP-guidelines 2, 3, and 5.
-Second, with regard to ease of use,
-WORCS aims to bring down the learning curve by adopting sensible defaults for any decisions to be made.
+Second, WORCS is easier to use, as it only requires familiarity with R.
+WORCS was designed to tightly integrate into the R ecosystem, bringing down the learning curve by utilizing the users' knowledge of R.
 Peikert and Brandmaier, by contrast, designed their workflow to be very flexible and comprehensive,
@@
 To sum up, WORCS builds upon the same general principles
-as Peikert and Brandmaier, and the two workflows are compatible. What 
-sets WORCS apart is that it is a more lightweight
-approach to safeguard computational reproducibility under *most circumstances*, 
-which is easier to adopt for projects that can be conducted entirely within `R`.
+as Peikert and Brandmaier, and the two workflows are compatible.
+What sets WORCS apart is that it is more lightweight in terms of system and user requirements, but still ensures computational reproducibility under *most circumstances* for projects that can be conducted entirely within `R`.
 

### 2020-11-05 paper/reviews_ds.Rmd
explain why worcs is easier

@@
 We hope that this adresses the Reviewer's question.
-However, as the workflow of Peikert and Brandmaier is still in the preprint stage, and the software implementation has not yet been developed - a more detailed comparison of these two approaches would be premature.
+However, as the workflow of Peikert and Brandmaier is still in press, and the software implementation has not yet been fully developed - a more detailed comparison of these two approaches would be premature.
 

### 2020-11-05 paper/Manuscript_DS.Rmd
R1 C7a Big data

--- 2020-11-03 paper/Manuscript_DS.Rmd
@@
   conducting research that involves academic prose, analysis code,
-  and tabular data that does not require specialized computational
-  provisions.
+  and (optionally) tabular data.
   However, the workflow is flexible enough to accommodate other scenarios,
@@
 explore alternative explanations, or even use existing data to answer a
-different research question. From an open science perspective, data sharing is
+different research question. 
+If it is possible to share the data, these can simply be pushed to a Git remote
+repository, along with documentation (such as a codebook) and the analysis code.
+This way, others can download the
+entire repository and reproduce the analyses from start to finish. 
+From an open science perspective, data sharing is
 always desirable. From a practical point of view, it is not always possible.
-Data sharing may be impeded by lack of storage space, legal constraints, ethical restrictions, absence
-of informed consent for data sharing, or privacy concerns. For example, one
-concern that applies when working with human participant data, is
-pseudonimization. 
+
+Data sharing may be impeded by several concerns.
+One technical concern is whether data require special storage provisions,
+as is often the case with "big data".
+Most Git remote repositories do not support files larger than 100MB,
+but it is possible to enhance a Git repository with [Git Large File Storage](https://git-lfs.github.com/) to add remotely stored files as large as several GB.
+Files exceeding this size require custom solutions beyond the scope of this paper.
+
+Sharing human participant data additionally entails legal and ethical concerns.
+Before sharing any
+human participant data, it is recommended to obtain approval from qualified
+ethical and legal advisory organs, guidance from Research Data Management Support, and
+informed consent from participants.
 Many legislatures require researchers to 'de-identify' human subject data upon collection, by removing or deleting any sensitive personal information and
@@
 replicability of their own work and potential reuse value of their data [@phillipsDiscombobulationDeidentification2016].
-
-Once data are de-identified or pseudononimized, they can often be shared if participants have
-provided informed consent to that effect. Nevertheles, it is important to note
+Furthermore, it is important to note
 that the European GDPR prohibits storing 'personal data' (information which can
@@
 Different Universities, countries, and funding bodies also have their own
-repositories that are complient with local legistation. Before sharing any
-human participant data, it is recommended to obtain approval from an internal
-(ethical) review board, guidance from Research Data Management Support, and
-informed consent from participants.
-
-If data can be shared openly, they can simply be pushed to a public Git
-repository, along with the analysis code. This way, others can download the
-entire repository and reproduce the analyses from start to finish. 
+repositories that are complient with local legistation.
+In sum, data should only be shared after consult qualified legal and ethical advisory organs,
+obtaining informed consent from participants,
+and de-identifying or pseudononimizing data.
+
 If data cannot be shared, researchers should aim to safeguard the potential for
@@
 The first solution is to publish a *checksum* of the original data file [@rivest1992md5].
-Think of a checksum as a 32-character summary, or as a "one word" description,
-of the contents of a file.
-Any change to the file will result in a different checksum^[It is theoretically possible but improbable that random changes to a file will result in the same checksum.]. 
+Think of a checksum as a 32-character unique identifier^[It is theoretically possible but improbable that random changes to a file will result in the same checksum.],
+or as a lossy "summary" of the contents of a file.
+Any change to the file will result in a different checksum. 
 Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
@@
 
-The second solution is to share a synthetic dataset with similar characteristics
+A second solution is to share a synthetic dataset with similar characteristics
 to the real data. Synthetic data mimic the level of measurement and
@@
 Thus, the time investment is eminently worthwhile.
+
+Another limitation is the fact that no single workflow can suit all research projects.
+Not all projects conform to the steps outlined here;
+some may skip steps, add steps, or follow similar steps in a different order.
+In principle, WORCS does not enforce a linear order, and allows extentensive user flexibility.
+A related concern is that some projects might require specialized solutions outside of the scope of this paper.
+For example, some projects might use very large data files,
+rely on data that reside on a protected server,
+or use proprietary software dependencies that cannot be version controlled using `renv` or Docker.
+In such cases, the workflow can serve as a starting point for a custom approach;
+offering a primer on best practices, and a discussion of relevant factors to consider.
+We urge users to inform us of such challenges and custom solutions,
+so that we may add them to the [overview of WORCS-projects on GitHub](https://github.com/cjvanlissa/worcs), where they can serve as an example to others.
 

### 2020-11-05 paper/reviews_ds.Rmd
R1 C7a Big data

--- 2020-11-04 paper/reviews_ds.Rmd
@@
 
-We agree that GitHub is not an ideal platform for handling data larger than 100MB.
-However, 100MB is very large for tabular data; for example, if only integer numbers are stored that is $(100MB*1024^2)/(2^2) = 26214400$ unique values.
+Response: We have split this comment up to separately address the handling of large data, and findability (see below).
 
-GitHub is, however, ideal version controlling analysis code, RMarkdown documents, and small to medium-sized data files.
-Larger files can be linked to git using [git-annex](https://git-annex.branchable.com),
-[Git Large File Storage](https://git-lfs.github.com/), and similar solutions. 
-We now mention git-annex as a possible route to handle larger data files (TODO).
+With regard to the handling of large data,
+we agree that GitHub is not an ideal platform for handling data larger than 100MB.
+It is important to note, however, that there are no restrictions for sharing files up to that size.
+Further, take into consideration that 100MB is very large for tabular data;
+for example, if only integer numbers are stored that is $(100MB*1024^2)/(2^2) = 26214400$ unique values.
+In the social sciences, datasets rarely exceed this limit -
+and so for our readership,
+GitHub is as good a platform for sharing data as any.
+Moreover, GitHub is ideal for version controlling analysis code, RMarkdown documents, and small to medium-sized data files.
+In our opinion, the advantages of having **all** of a paper's resources in the same repository outweighs the advantages of encouraging every reader to additionally use specialized solutions for "big data" (which most will not need).
+
+To address this comment, we now explicitly address the problem of version controlling large files, and introduce a solution for larger files, which can be linked to git using
+[Git Large File Storage](https://git-lfs.github.com/):
+
+> Data sharing may be impeded by several concerns.
+One technical concern is whether data require special storage provisions,
+as is often the case with "big data".
+Most Git remote repositories do not support files larger than 100MB,
+but it is possible to enhance a Git repository with [Git Large File Storage](https://git-lfs.github.com/) to add remotely stored files as large as several GB.
+Files exceeding this size require custom solutions beyond the scope of this paper.
+
+We also added a paragraph to the Limitations section that discusses the fact that no single workflow is suitable for all situations:
+
+> Another limitation is the fact that no single workflow can suit all research projects.
+Not all projects conform to the steps outlined here;
+some may skip steps, add steps, or follow similar steps in a different order.
+In principle, WORCS does not enforce a linear order, and allows extentensive user flexibility.
+A related concern is that some projects might require specialized solutions outside of the scope of this paper.
+For example, some projects might use very large data files,
+rely on data that reside on a protected server,
+or use proprietary software dependencies that cannot be version controlled using `renv` or Docker.
+In such cases, the workflow can serve as a starting point for a custom approach;
+offering a primer on best practices, and a discussion of relevant factors to consider.
+We urge users to inform us of such challenges and custom solutions,
+so that we may add them to the [overview of WORCS-projects on GitHub](https://github.com/cjvanlissa/worcs), where they can serve as an example to others.
+
+\Assignment{BV, AB, AL, CJ}
+\WorkInProgress
+\Done
+
+\RC{C7b: Second, it doesn't follow very well the FAIR principles for findability and attribution of data.
+Instead, repositories like Zenodo or FigShare could be user to address data storage, with proper metadata and its corresponding citation.}
 
@@
 \Hard
-
-\RC{C7b: Second, it doesn't follow very well the FAIR principles for findability and attribution of data.
-Instead, repositories like Zenodo or FigShare could be user to address data storage, with proper metadata and its corresponding citation.}
 

### 2020-11-05 paper/Manuscript_DS.Rmd
minor tweak to Peikert's changes

@@
 which is relevant for TOP-guidelines 2, 3, and 5.
-Second, WORCS is easier to use, as it only requires familiarity with R.
-WORCS was designed to tightly integrate into the R ecosystem, bringing down the learning curve by utilizing the users' knowledge of R.
+Second, with regard to ease of use,
+WORCS aims to bring down the learning curve by adopting sensible defaults for any decisions to be made.
 Peikert and Brandmaier, by contrast, designed their workflow to be very flexible and comprehensive,
@@
 To sum up, WORCS builds upon the same general principles
-as Peikert and Brandmaier, and the two workflows are compatible.
-What sets WORCS apart is that it is more lightweight in terms of system and user requirements, but still ensures computational reproducibility under *most circumstances* for projects that can be conducted entirely within `R`.
+as Peikert and Brandmaier, and the two workflows are compatible. What sets WORCS apart is that it is more lightweight in terms of system and user requirements,
+thereby facilitating adoption, but still ensures computational reproducibility under *most circumstances*.
 

### 2020-11-05 paper/reviews_ds.Rmd
mention DOI

@@
 We agree that it is important to make open data findable beyond just providing it freely.
+Step 15 of the workflow, "Connect your ‘OSF’ project page to the ‘Git’ remote repository",
+was intended to meet this goal.
+However, we acknowledge that this requires further clarification.
+
+Particularly the addition of a robust citation for the data is an essential step that should have been mentioned in the manuscript.
+In the revision, we have addressed this by adding a step to the WORCS workflow, and discussing DOIs in the paper.
+
+Unfortunately, the Reviewer preference for Zenodo or FigShare is not explained.
+We consider these platforms to be somewhat interchangeable with using Git remote repositories - with one key distinction:
+A crucial advantage of storing the data on GitHub is that they will be bundled with the data documentation (codebook), analysis code, and manuscript.
+Thus, readers have access to **all** raw materials in one repository.
+Moreover, GitHub additionally offers user interaction with the materials through
+forking, cloning, opening issues, and sending pull requests.
+Thus, this platform is more conducive to collaboration.
+
 Digital object identifiers are the de-factor standard for making unique references to data sets. In line with the reviewer's suggestion,
 we suggest to use external providers to provide such DOIs while github is still the primary data storage. TODO!!!
-
-Our prefered route is to use github web hooks to link Zenodo and a WORCS-compatible repository.
-With this technique, Zenodo issues DOI for content hosted on github repositories including some nice extras, such as automatic badge DOI badges for the repository.
 

### 2020-11-05 paper/reviews_ds.Rmd
rewrite 7b

--- 2020-11-05 paper/reviews_ds.Rmd
@@
 
+A crucial advantage of storing the data on GitHub, relative to Zenodo or FigShare,
+is that they are bundled with the
+documentation (e.g., codebook), analysis code, and manuscript.
+Thus, readers have access to **all** materials required for replication in one repository.
+GitHub additionally offers user interaction with the materials through
+forking, cloning, opening issues, and sending pull requests -
+all useful tools for replication and collaboration.
+
 We agree that it is important to make open data findable beyond just providing it freely.
+
+The approach we advocate is to make all of a project's resources available in a 
+Step 15 of the workflow, "Connect your ‘OSF’ project page to the ‘Git’ remote repository",
+was intended to meet this goal.
+However, we acknowledge that this requires further clarification.
+
+Particularly the addition of a robust citation for the data is an essential step that should have been mentioned in the manuscript.
+In the revision, we have addressed this by adding a step to the WORCS workflow, and discussing DOIs in the paper.
+
+Unfortunately, the Reviewer preference for Zenodo or FigShare is not explained.
+We consider these platforms to be somewhat interchangeable with using Git remote repositories - with one key distinction:
+
 Digital object identifiers are the de-factor standard for making unique references to data sets. In line with the reviewer's suggestion,
@@
 
-Our prefered route is to use github web hooks to link Zenodo and a WORCS-compatible repository.
-With this technique, Zenodo issues DOI for content hosted on github repositories including some nice extras, such as automatic badge DOI badges for the repository.
-
 BV: Can you please argue for why GitHub is equivalent to Zenodo for this purpose?
 AL: Can you address the reference to FAIR principles in relation to GitHub?
-CJ: I will address "big" data
 
-on second thought, it’s probably worth mentioning to the reviewer that we know these options but defend the OSF-centric approach for our purposes.
+AB: "on second thought, it’s probably worth mentioning to the reviewer that we know these options but defend the OSF-centric approach for our purposes.
 My impression is that the main challenge are the different expectations and solutions across disciplines that are currently evolving largely in parallel.
@@
 Other disciplines may have different foci, e.g., python + jupyter + docker etc...
-and no single manuscript will be able to cover everything.
+and no single manuscript will be able to cover everything."
 

### 2020-11-05 paper/Manuscript_DS.Rmd
Fix Aaron's affiliation

@@
     Anna-Lena Lamprecht$^{b,f}$
-    Aaron Peikert$^{c,g}$
-    Marijn E. Struiksma$^{b,h}$
-    Barbara Vreede$^{i}$
+    Aaron Peikert$^{c}$
+    Marijn E. Struiksma$^{b,g}$
+    Barbara Vreede$^{h}$
 
@@
       f: Utrecht University, Department of Information and Computing Sciences  
-      g: Humboldt-Universitaet zu Berlin  
-      h: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
-      i: University Library Utrecht
+      g: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
+      h: University Library Utrecht
 title: "WORCS: A Workflow for Open Reproducible Code in Science"

### 2020-11-05 paper/reviews_ds.Rmd
Improve 7b

@@
 
-A crucial advantage of storing the data on GitHub, relative to Zenodo or FigShare,
-is that they are bundled with the
+We agree completely that it is important to make open data findable beyond just providing it freely.
+It is a pity that the Reviewer does not provide arguments to support a preference for Zenodo or FigShare over GitHub, because if there is a strong case to be made, we would be keen to include this in the paper.
+Without such arguments, as far as we are concerned, these repositories are largely interchangeable.
+
+We prefer storing the data on GitHub because they are then bundled with the
 documentation (e.g., codebook), analysis code, and manuscript.
@@
 
-We agree that it is important to make open data findable beyond just providing it freely.
+It is further worth noting that [GitHub is compatible with Zenodo](https://guides.github.com/activities/citable-code/);
+users can document GitHub repositories in Zenodo and generate a DOI.
+This is redundant with Step 15 of the workflow, however:
+Connecting the GitHub repository to an OSF project accomplishes effectively the same.
+We do acknowledge, however, 
+that this part of the workflow was not clearly documented in the previous version of the manuscript, 
+and that the use of DOIs warrants further clarification.
+
+In the revision, we have attempted to address this comment in two ways: First, by
+clarifying Step 15 in the WORCS workflow, specifically emphasizing the importance of generating a DOI. 
+Second, by discussing the use of DOIs in the paper.
 
@@
 was intended to meet this goal.
-However, we acknowledge that this requires further clarification.
-
-Particularly the addition of a robust citation for the data is an essential step that should have been mentioned in the manuscript.
-In the revision, we have addressed this by adding a step to the WORCS workflow, and discussing DOIs in the paper.
-
-Unfortunately, the Reviewer preference for Zenodo or FigShare is not explained.
-We consider these platforms to be somewhat interchangeable with using Git remote repositories - with one key distinction:
 

### 2020-11-08 README.Rmd
Incorporate Loek's comments

@@
 
-The `worcs` package is open to contributions from collaborators with valuable expertise, to improve both the workflow and the software.
-Relevant contributions warrant coauthorship to the package.
+We are always eager to receive user feedback and contributions to help us improve both the workflow and the software.
+Major contributions warrant coauthorship to the package.
 Please contact the lead author at [c.j.vanlissa@uu.nl](mailto:c.j.vanlissa@uu.nl), or:
 
-  - File a GitHub issue [here](https://github.com/cjvanlissa/worcs)
-  - Make a pull request
-    [here](https://github.com/cjvanlissa/worcs/pulls)
+  - [File a GitHub issue](https://github.com/cjvanlissa/worcs) for feedback, bug reports or feature requests
+  - [Make a pull request](https://github.com/cjvanlissa/worcs/pulls) to contribute your code or prose
 

### 2020-11-08 paper/Manuscript_DS.Rmd
Incorporate Loek's comments

@@
 for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
-Git retains a complete log of all commits, and users can compare changes between
-different commits, or go back to a previous version of the code (for example,
+Each commit is further tagged with a message, describing the changes.
+The `R` implementation of `worcs` contains a convenience function,
+`git_update("your commit message")`,
+which creates a new commit for all changes since the previous commit,
+labels it with "your commit message", and pushes these changes to the remote repository.
+It is, effectively, a Git "quick save" command.
+
+From an open science perspective, Git is particularly appealing because it
+retains a complete historical backlog of all commits.
+This can be used, for example, to verify that authors executed the steps outlined in a preregistration.
+Users can also compare changes between different commits,
+or go back to a previous version of the code (for example,
 after making a mistake, or to replicate a previous version of the results).

### 2020-11-08 paper/reviews_ds.Rmd
Update
Incorporate Loek's comments

@@
 
-Response: We appreciate the recommendation of performing a user evaluation, and this is certainly a next step in the development of worcs.
-However, as the Editor has indicated, 'resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse"'.
-We will thus first address the potential for re-use, and then we will address steps we have taken and intend to take with regard to user evaluation.
+Response: We appreciate the recommendation of performing a user evaluation, and this is certainly a next step in the development of WORCS.
+At this point, we have written this manuscript detailing the conceptual workflow, and released a software implementation for R-users.
+The next step, which we are now entering, is to build a user base, and then mount a user evaluation.
 
-First, there are several arguments to substantiate the potential for reuse of worcs:
+To address the Reviewer's comment, we have set up an evaluation questionnaire, and sent it to the attendees of our previous workshops on WORCS.
+At the time of writing, however, this has not yielded actionable points of improvement.
 
-1. WORCS is implemented in an R-package
-2. This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network")
-3. The software meets all criteria of the Core Infrastructure Initiative "best practices" (CII Best Practices)
+We do have a strong history of addressing user comments submitted through the GitHub issues page.
+WORCS adheres to the Core Infrastructure Initiative "best practices" (CII Best Practices),
+which requires that
+the channels through which users can provide feedback are clearly indicated,
+as they are [here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
+It is also evident from the [resolved tickets on GitHub](https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed) that we have been very responsive in actively incorporating user feedback - as also required by the CII Best Practices.
+
+We are committed to continue evaluating user experience, both actively through evaluation forms, and passively through the GitHub issues pages.
+We hope that this commitment, along with the evidence of our prior responsiveness to user feedback presented above, assuage the Reviewer's concerns.
+
+Finally, we wish to point out that the Editor has indicated that 'resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse"'.
+We therefore elaborate on the efforts we have made to ensure that WORCS is reusable, followed by evidence of its current adoption rates.
+
+We have made the following efforts to ensure potential for reuse:
+
+1. WORCS is implemented in an R-package, and thus readily available to R-users
+2. This R-package is available in a standardized repository, namely CRAN ("The Comprehensive R Archive Network"), and can thus be installed through default channels
+3.
+The software meets all criteria of the Core Infrastructure Initiative "best practices" (CII Best Practices), including clearly indicated channels for user feedback, and being responsive to that feedback
 4. The R-package uses only tools that are fully integrated within the RStudio IDE, thus enhancing ease of use
@@
 
-1. The preprint has been downloaded 561 times, since being published on 31-05-2020
-1. The `worcs` R-package has been downloaded 2401 times from CRAN, since being published on 18-05-2020
+1. [The preprint](https://psyarxiv.com/k4wde/) has been downloaded 568 times, since being published on 31-05-2020
+1. The `worcs` R-package has been downloaded 2474 times from CRAN, since being published on 18-05-2020
 1. The GitHub project has been forked by users unknown to us 4 times, and starred (= watched) 32 times
@@
 
-Third, although we have not conducted a user evaluation, there is some evidence that `worcs` addresses users' needs:
-
-1. Evaluate worcshops!!
-2. The issues page on the worcs GitHub profile demonstrates that we have been very responsive in addressing users' feedback (as required by the aforementioned CII Best Practices):
-* https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed
-
 <!-- The goal of the present manuscript is to define the problem worcs intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools worcs uses to address this problem, explain how worcs uses these tools, and provide a step-by-step workflow.
@@
 \Assignment{CJ, BV, AL}
-\WorkInProgress
+\Done
 \Hard
@@
 \Assignment{CJ}
-\WorkInProgress
+\Done
 \Easy
@@
 We agree completely that it is important to make open data findable beyond just providing it freely.
+This is achieved by linking the GitHub repository to an OSF.io project page;
+on OSF, the project receives a DOI.
+It is also possible to give a separate DOI to specific resources, such as data files.
+We agree that this was not emphasized enough in the manuscript, and have rewritten the revision to clarify.
+
 It is a pity that the Reviewer does not provide arguments to support a preference for Zenodo or FigShare over GitHub, because if there is a strong case to be made, we would be keen to include this in the paper.
@@
 
+Response: 
+F: Dit heeft weer met DOIs te maken. Ik denk dat als we het gebruik van OSF de default van worcs maken (die mijn opmerking R1C7a) dat we hier een sterk punt kunnen maken mbt findability. 
+A: Eigenlijk hetzelfde antwoord als bij Findability: dit wordt heel eenvouding als we de OSF route voor data storage de default maken. 
+I: Het feit dat we alleen gebruik maken van open-source software maakt alle projecten die worcs gebruiken nogal interoperable. 
+R: Dit is wel een belangrijkt punt: Kunnen we opnemen in de workflow dat worcs-projecten by default een CC-BY licentie krijgen?
+
 \Assignment{AL, BV, CJ}
@@
 
-Response: It does not. It seems to us that this is outside of the scope of the present paper.
-If a compelling case could be made by the Reviewer to include such a service,
-we are open to the suggestion - but we would need to take on additional coauthors with the required expertise.
+Response: WORCS does not provide a service for de-identification,
+and we consider this to be outside of the scope of the present paper.
+If the Reviewer feels differently, we are open to being convinced, however!
+Of course, since `worcs` is open source software, anybody is welcome to contribute this functionality in exchange for coauthorship, in line with the contribution guidelines [outlined here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
+The current team lacks relevant expertise to make this contribution.
 <!--What would this entail? Is it within scope?-->
@@
 \Assignment{ALL}
-\WorkInProgress
+\Done
 \Medium
@@
 Although Git tracks file changes on a line-by-line basis,
-any number of such changes can be committed together, thereby capturing the more "course-grained elements".
+such changes are typically grouped together in a "commit",
+which is how Git captures the more "course-grained elements" the Reviewer refers to.
 For example, in our response to these reviews, we have created one Git commit for each Reviewer comment, and these commits contain changes to the action letter, manuscript, and vignettes.
 
-The Reviewer's comment does illustrate the importance of clarifying this point in the manuscript as well
-To address this comment, the revised manuscript now reads:
+To address the Reviewer's comment, we now clarify this point in the revised manuscript:
 
-> Git tracks changes to files on a line-by-line basis.
-The user can sture these changes by making a "commit": a snapshot of the version controlled files.
-Each "commit" can contain as many changes as desired; for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
+> Git tracks changes to files on a line-by-line basis. The user can sture these changes by making a "commit": a 
+snapshot of the version controlled files.
+Each "commit" can contain as many changes as desired;
+for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
+Each commit is further tagged with a message, describing the changes.
+The `R` implementation of `worcs` contains a convenience function,
+`git_update("your commit message")`,
+which creates a new commit for all changes since the previous commit,
+labels it with "your commit message", and pushes these changes to the remote repository.
+It is, effectively, a Git "quick save" command.
 

### 2020-11-08 paper/reviews_ds.Rmd
number comments

@@
 
-\RC{The claim that WORCS is amenable to the FAIR Guiding Principles should be justified.
+\RC{C2: The claim that WORCS is amenable to the FAIR Guiding Principles should be justified.
 What are the solutions provided by WORCS that make a project interoperable and reusable?
@@
 
-\RC{Is WORCS providing any service for de-identification?}
+\RC{C3: Is WORCS providing any service for de-identification?}
 

### 2020-11-09 paper/Manuscript_DS.Rmd
Loek changes

@@
 replication" [@nosekPromotingOpenResearch2015a].
-In this context, 
-reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
-and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
 These guidelines are: 1) Comprehensive
@@
 journal guidelines for consideration of independent replications for publication".
+In this context, 
+reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
+and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
 WORCS defines the goals of open science in terms of the first seven TOP-guidelines,
@@
 
-WORCS relies on several free, open source software solutions which we will
+WORCS relies solely on a set of free, open source software solutions which we will
 discuss before introducing the workflow.
@@
 platform 'Code Ocean'.
-Although Code Ocean enjoys some popularity among 
-scientists, it is not clear whether this platform
-is GDPR compliant - as of this writing, its [terms of use](https://codeocean.com/terms-of-use) do not address the GDPR, nor is it clear where
-data are stored.
-The necessity to upload human participant data for analysis is an important consideration and potential liability when using any cloud-computing solution.
+This platform greatly simplifies setting up a reproducible computing environment with version control and dependency management.
+A key feature is that collaborators can log in to the server session,
+rather than each having to set up Docker on a local computer.
+One important consideration when using any cloud-computing solution
+is the necessity to upload human participant data for analysis.
+Such a platform would need to comply with relevant laws and ethical requirements.
+For example, as of this writing, the [terms of use](https://codeocean.com/terms-of-use) for Code Ocean do not address the GDPR,
+nor is it clear where data are stored.
+This is a potential liability for European scholars.
 
@@
 
+A final limitation is that, as WORCS is relatively new,
+the number of current users is still limited.
+We therefore do not have elaborate insights in user experiences or feedback.
+We strive to make the uptake of worcs as easy as possible,
+by means of this background paper, extensive documentation and vignettes for the `R` implementation,
+[online video tutorials](https://www.youtube.com/playlist?list=PLOQIOFZl7uYbTgVdUee9n3eebMNmJtV9y),
+webinars and [exemplar use cases](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users).
+We encourage our users to provide us with feedback, feature requests, or bug reports through the channels [indicated here](https://github.com/cjvanlissa/worcs/tree/master#contributing-and-contact-information).
+
 ## Future developments

### 2020-11-14 paper/Manuscript_DS.Rmd
Better address FAIR principles

@@
   adoption of open science principles.
-  It is based on best practices, and can be used either in
+  It is based on established best practices, and can be used either in
   parallel to, or in absence of, top-down requirements by journals,
@@
   discusses how it meets different 
-  standards for open science, and provides examples of the implementation in `R`. 
+  standards for open science, and addresses the functionality provided by the
+  `R` implementation, `worcs`. 
   This paper is primarily targeted towards
-  scholars conducting research projects in R on a private computer,
+  scholars conducting research projects in R,
   conducting research that involves academic prose, analysis code,
@@
   The source code for the `R` package and manuscript,
-  and a list of user examples of WORCS projects, are available at 
+  and a list of [user examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
   https://github.com/cjvanlissa/worcs.
@@
 
+# Introduction
+
 Academia is arguably past the tipping point of a paradigm shift
@@
 This paper is designed to ease that transition by presenting a simple workflow,
-based on best practices,  
+based on established best practices,  
 that meets most requirements open science:
@@
 and several convenience functions to automate specific workflow steps.
-This paper is most relevant for scholars using R on a private computer to
-conducting research that involves academic prose, analysis code,
+This paper is most relevant for scholars using R to
+conduct research that involves academic prose, analysis code,
 and tabular data.
@@
 
-The WORCS workflow (Figure 1) 
-constitutes a step-by-step procedure that researchers can follow to make a 
+The WORCS workflow constitutes a step-by-step procedure that researchers can follow to make a 
 research project open and reproducible.
+See Figure 1 for a schematic overview of the conceptual workflow,
+which shows the steps to take in different stages of the research process.
+The workflow is elaborated in greater detail below.
 WORCS is compatible with open science
@@
 Git tracks changes to files on a line-by-line basis.
-The user can sture these changes by making a "commit": a 
+The user can store these changes by making a "commit": a 
 snapshot of the version controlled files.
@@
 A `.gitignore` file lists files that should not be version-controlled.
-These files thus exist only on the user's private computer.
+These files thus exist only on the user's computer.
 
@@
 
-WORCS was designed to meet the TOP-guidelines, and it does so in a way that
-is compatible with other open science principles. Most notably, WORCS is 
-amenable to the
-FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a].
+Aside from the TOP Guidelines, the FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a] are an important standard for open science.
 These principles advocate that digital research objects should be Findable, Accessible, Interoperable and Reusable.
@@
 
-WORCS helps users meet the FAIR principles by making all created research objects -
-including data and code - accessible in a repository that can be made public. 
-The `worcs` package additionally
-facilitates meeting the FAIR principles in several ways. First, each `worcs`
-project has a YAML file (https://yaml.org/) that lists the research objects in
-the project, which will allow web crawlers 
-to index `worcs` projects. Second, a codebook is automatically generated for
-each data file stored
-using `open_data()` or `closed_data()`. We encourage users to elaborate on the
-default codebook by adding
+WORCS was not designed to comprehensively address these principles, but the workflow does facilitate meeting them.
+In terms of Findability, all files in GitHub projects are searchable by content and meta-data, both on the platform and through standard search engines.
+The workflow further recommends that users create a project page on the Open Science Framework – a recognized repository in the social sciences – and generate a DOI for this project.
+Additional DOIs can be generated through the OSF for specific resources, such as data sets.
+It is additionally or alternatively possible to connect the GitHub repository to Zenodo, which also offers DOIs and allows users to document projects with extensive metadata.
+Finally, each worcs project has a [YAML file] (https://yaml.org/) that lists the data objects (and other meta-data) in the project, which will allow web crawlers to index worcs projects.
+We use this metadata, for example, to identify public `worcs` projects on GitHub.
+Accessibility is primarily safeguarded by hosting all project files in a public GitHub repository, which can be downloaded as a ZIP archive, and cloned or forked using the Git protocol.
+[GitHub is committed to long term accessibility of these resources]( https://archiveprogram.github.com/).
+We further encourage users to make their data accessible when possible.
+When data must be closed, the `worcs` function `closed_data()` adds a paragraph to the project readme, indicating how readers can contact the author for access.
+
+With regard to interoperability, the workflow recommends using plain-text files whenever possible, which ensures that code and meta-data are human- and machine-readable.
+Furthermore, the `worcs` package is implemented in R, which is open source.
+To promote reusability, we recommend users to specify an appropriate license for their projects.
+The `worcs` package includes the CC licenses, and suggests a CC-BY 4.0 license by default.
+This license is featured in the GitHub repository, and we encourage users to also display it on their project’s OSF page.
+To further facilitate reusability, `worcs` automatically generates a codebook for
+each data file.
+We encourage users to elaborate on the default codebook by adding
 variable description and categories, which would - in the future - enable
@@
 
-Sharing all research objects, as advocated in WORCS, also provides a proper
+Sharing all research objects, as advocated in WORCS, also provides a thorough
 basis for research evaluation
@@
 lightweight workflow for open, reproducible research, that meets all
-TOP-guidelines, except replication.
+TOP-guidelines pertaining to reproducibility (1-7).
 Nevertheless, there are clear directions for future developments.

### 2020-11-17 paper/Manuscript_DS.Rmd
more extensive discussion of licensing

@@
   affiliation: | 
-      a: Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics  
-      b: Open Science Community Utrecht  
-      c: Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany  
-      d: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK  
-      e: UMC Utrecht  
-      f: Utrecht University, Department of Information and Computing Sciences  
-      g: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
-      h: University Library Utrecht
+      $a$: Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics  
+      $b$: Open Science Community Utrecht  
+      $c$: Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany  
+      $d$: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK  
+      $e$: UMC Utrecht  
+      $f$: Utrecht University, Department of Information and Computing Sciences  
+      $g$: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
+      $h$: University Library Utrecht
 title: "WORCS: A Workflow for Open Reproducible Code in Science"
@@
 research project open and reproducible.
-See Figure 1 for a schematic overview of the conceptual workflow,
-which shows the steps to take in different stages of the research process.
-The workflow is elaborated in greater detail below.
+These steps are elaborated in greater detail below.
+See Figure \@ref(fig:workflow) for a flowchart that highlights important steps in different stages of the research process.
 WORCS is compatible with open science
@@
 
-```{r workflow, fig.cap="*Figure 1*: Schematic illustration of the WORCS procedure.", out.width="100%"}
+```{r workflow, fig.cap="Schematic illustration of important steps in different stages of the WORCS procedure.", out.width="100%"}
 knitr::include_graphics("workflow.png")
@@
 We provide a conceptual outline of the WORCS procedure below, based on WORCS
-Version 0.1.5.
+Version 0.1.6.
 The conceptual workflow recommends actions to take 
@@
 The vignette is also updated with future developments of the package.
-
-For an up-to-date list of sample `worcs` projects, see the [WORCS GitHub page](https://github.com/cjvanlissa/worcs).
+For examples of how the workflow is used in practice,
+we maintain [a list of public `worcs` projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users) on the WORCS GitHub page.
 
@@
 `worcs` package offers a simple but flexible function to generate synthetic data,
-`synthetic()`, which by default uses random forests to generate a model for
-synthesis (see @wrightRangerFastImplementation2015).
+`synthetic()`.
 This function is based on the work by @nowokSynthpopBespokeCreation2016 (see the `R` package `synthpop`),
@@
 marginal associations in the data.
+By default, this function uses random forests, as implemented in the `ranger` package, to generate a model for synthesis [@wrightRangerFastImplementation2015].
+Note that the `ranger()` function does not handle all available data types in `R`.
+Users can call custom functions to build a prediction model for unsupported data types, using the argument `model_expression`.
 For further details, see the [function documentation](https://cjvanlissa.github.io/worcs/reference/synthetic.html).
@@
 In terms of Findability, all files in GitHub projects are searchable by content and meta-data, both on the platform and through standard search engines.
-The workflow further recommends that users create a project page on the Open Science Framework – a recognized repository in the social sciences – and generate a DOI for this project.
+The workflow further recommends that users create a project page on the Open Science Framework – a recognized repository in the social sciences.
+When doing so, it is important to generate a Digital Object Identifier (DOI) for this project.
+A DOI is a persistent way to identify and connect to an object on the internet, which is crucial for findability.
 Additional DOIs can be generated through the OSF for specific resources, such as data sets.
-It is additionally or alternatively possible to connect the GitHub repository to Zenodo, which also offers DOIs and allows users to document projects with extensive metadata.
+It is further worth noting that, optionally, [GitHub repositories can be connected to Zenodo](https://guides.github.com/activities/citable-code/):
+On Zenodo, users can store a snapshot of the repository, provide metadata, and generate a DOI for the project or specific resources.
+When a project is connected to the OSF, as recommended, these steps may be redundant.
 Finally, each worcs project has a [YAML file] (https://yaml.org/) that lists the data objects (and other meta-data) in the project, which will allow web crawlers to index worcs projects.
 We use this metadata, for example, to identify public `worcs` projects on GitHub.
+
 Accessibility is primarily safeguarded by hosting all project files in a public GitHub repository, which can be downloaded as a ZIP archive, and cloned or forked using the Git protocol.
@@
 To promote reusability, we recommend users to specify an appropriate license for their projects.
-The `worcs` package includes the CC licenses, and suggests a CC-BY 4.0 license by default.
+The `worcs` package includes the [Creative Commons (CC) licenses](https://creativecommons.org/licenses/), and suggests a CC-BY 4.0 license by default.
 This license is featured in the GitHub repository, and we encourage users to also display it on their project’s OSF page.
+Note that it is possible to license specific project resources separately under different licenses.
+For instance, whereas the CC-licenses are suitable for data, media, and academic papers - they are ill-suited for licensing software.
+It is thus prudent to consider separately licensing specific project resources,
+particularly when a project contains *original* source code (i.e., you write your own functions).
+For a more extensive discussion of appropriate licensing, see @peikertReproducibleDataAnalysis2019, @stoddenEnhancingReproducibilityComputational2016, and the website [choosealicense.com](https://choosealicense.com/).
+
 To further facilitate reusability, `worcs` automatically generates a codebook for
@@
 
-Incidentally, `worcs` itself follows the recommendations for [FAIR research software](https://fair-software.eu/). It is 
+Incidentally, the FAIR principles have been applied to software as well,
+and `worcs` follows these recommendations for [FAIR research software](https://fair-software.eu/). It is 
 hosted on a version controlled public remote repository GitHub, has an open

### 2020-11-17 paper/reviews_ds.Rmd
more extensive discussion of licensing

@@
 
-Dear Dr. Editor,
+Dear Dr. Kuhn,
 
-thank you very much for taking the time to consider our manuscript for publication at _`r rmarkdown::metadata$journal`_.
-In the following we address your and each reviewers' concerns point-by-point.
+We would like to thank you for providing us with the opportunity to revise our manuscript, "WORCS: A Workflow for Open Reproducible Code in Science".
+We sincerely appreciate the time you and the Reviewers have taken to read the manuscript, and the extensive and thought-provoking comments you have raised. 
+We hereby submit our revision of the manuscript for your consideration.
+We have attempted to address each of the issues raised by yourself and the Reviewers on a point-by-point basis, and detail the changes made in the action letter below.
+We would like to thank you and the Reviewers for your time, and we look forward to receiving your response.
+
+Sincerely, on behalve of all coauthors,
+Caspar van Lissa
 
@@
 To address the Reviewer's comment, we have set up an evaluation questionnaire, and sent it to the attendees of our previous workshops on WORCS.
-At the time of writing, however, this has not yielded actionable points of improvement.
+At the time of writing, responses are coming in.
+These are generally very favorable, and have not yet revealed major points of improvement.
 
-We do have a strong history of addressing user comments submitted through the GitHub issues page.
+We further want to point out that we have a strong history of addressing user comments submitted through the GitHub issues page.
 WORCS adheres to the Core Infrastructure Initiative "best practices" (CII Best Practices),
@@
 as they are [here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
-It is also evident from the [resolved tickets on GitHub](https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed) that we have been very responsive in actively incorporating user feedback - as also required by the CII Best Practices.
+It is also evident from the [resolved tickets on GitHub](https://github.com/cjvanlissa/worcs/issues?q=is%3Aissue+is%3Aclosed) that we have been very responsive in actively incorporating user feedback.
+This is another requirement of CII Best Practices.
 
-We are committed to continue evaluating user experience, both actively through evaluation forms, and passively through the GitHub issues pages.
-We hope that this commitment, along with the evidence of our prior responsiveness to user feedback presented above, assuage the Reviewer's concerns.
+We are committed to continue evaluating user experience, both actively through user evaluation forms, and passively through the GitHub issues pages.
+We hope that this commitment, along with the evidence of our prior responsiveness to user feedback presented above, help assuage the Reviewer's concerns.
 
 Finally, we wish to point out that the Editor has indicated that 'resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse"'.
-We therefore elaborate on the efforts we have made to ensure that WORCS is reusable, followed by evidence of its current adoption rates.
+We therefore also elaborate on the efforts we have made to ensure that WORCS is reusable, followed by evidence of its current adoption rates.
 
@@
 
-Second, there is some evidence to demonstrate that WORCS is already being adopted (as of 31-10-2020):
+There is also evidence that WORCS is already being adopted, even prior to publication (as of 14-11-2020):
 
-1. [The preprint](https://psyarxiv.com/k4wde/) has been downloaded 568 times, since being published on 31-05-2020
-1. The `worcs` R-package has been downloaded 2474 times from CRAN, since being published on 18-05-2020
-1. The GitHub project has been forked by users unknown to us 4 times, and starred (= watched) 32 times
-1. The video tutorial for worcs has been viewed 362 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
-1. WORCS is used to teach reproducibility at Utrecht University
-1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
-1. The lead author has applied `worcs` in several research projects:
-	* https://github.com/cjvanlissa/storm_meta_father_mother
-	* https://github.com/cjvanlissa/veni_sysrev
-	* https://github.com/cjvanlissa/placebo_adhd
-	* https://github.com/cjvanlissa/schumpe
-	* https://github.com/cjvanlissa/coping_covid
+1.  The preprint has been downloaded 582 times, since being published on
+    31-05-2020
+2.  The `worcs` R-package has been downloaded 2567 times from CRAN,
+    since being published on 18-05-2020
+3.  The GitHub project has been forked 8 times, watched 5 times, and
+    starred 40 times
+4.  The lead author has given invited lectures on WORCS at:
+      - Max-Planck-Institute for Human Development (2020)  
+      - Open Science Community Rotterdam (2020)  
+      - Eindhoven University of Technology (2020)  
+5.  WORCS is currently used in the following GitHub repositories (sorted
+    by user):
+      - cjvanlissa
+          - [veni\_sysrev](https://github.com/cjvanlissa/veni_sysrev)  
+          - [IHBT-SEM](https://github.com/cjvanlissa/IHBT-SEM)  
+          - [schumpe](https://github.com/cjvanlissa/schumpe)  
+          - [storm\_meta\_father\_mother](https://github.com/cjvanlissa/storm_meta_father_mother)  
+          - [placebo\_adhd](https://github.com/cjvanlissa/placebo_adhd)  
+          - [mpib](https://github.com/cjvanlissa/mpib)  
+          - [coping\_covid](https://github.com/cjvanlissa/coping_covid)  
+      - IndrajeetPatil
+          - [worcs\_practice](https://github.com/IndrajeetPatil/worcs_practice)  
+      - EstherPlomp
+          - [RDM-survey](https://github.com/EstherPlomp/RDM-survey)  
+      - lisallreiber
+          - [worcs\_ws\_empty](https://github.com/lisallreiber/worcs_ws_empty)  
+      - KimBaldry
+          - [BIO-MATE](https://github.com/KimBaldry/BIO-MATE)
 
@@
 
-Response: In addressing this comment, we respectfully refer to our response to the preceding comment ("There is no user evaluation", R1), and again emphasize the Editor's comment that "resource papers do not need to have a full-blown user evaluation".
+Response: In addressing this comment, we respectfully refer to our response to the preceding comment ("There is no user evaluation", R1 C1), and again emphasize the Editor's comment that "resource papers do not need to have a full-blown user evaluation".
 We would also like to clarify that the goal of the present manuscript is to define the problem WORCS intends to address (specifically; meeting the TOP-guidelines for Open Science), introduce the tools WORCS uses to address this problem, explain how WORCS uses these tools, and provide a step-by-step workflow.
@@
 
-We do take the suggestion of performing a user evaluation very seriously, but in our mind, this is a next step.
+We do take the suggestion of performing a user evaluation very seriously, but we also feel that it would be premature to do so even before the paper is published.
 We have built WORCS based on established tools and best practices, and we have gathered preliminary positive feedback from the community through workshops and teaching.
@@
 
-Response: We understand that our writing comes across as contradictory,
-and this apparent contradiction arises from a lack of explicit definitions of the terms "replication" and "reproducibility" in the original manuscript.
+Response: We thank the Reviewer for bringing this apparent contradiction to our attention.
+The misunderstanding arises from a lack of explicit definitions of the terms "replication" and "reproducibility" in the original manuscript.
 We are aware of the oft-lamented lack of consistent definitions for these terms (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018).
@@
 
-After making these definitions explicit, it becomes clear that the Reviewer's definition of replication as "a necessary steps towards reproducibility" is not consistent with the definitions of the present paper.
+After making these definitions explicit, it becomes clear that the Reviewer's definition of replication as "a necessary steps towards reproducibility" is not consistent with the definitions used in the present paper.
 
-We further wish to clarify that the fact that we address guidelines 1-7, but not guideline 8 ("replication") is consistent with the intended purpose of the TOP-guidelines.
+We further wish to clarify that the fact that we address guidelines 1-7, but not guideline 8 ("replication") is fully consistent with the intended purpose of the TOP-guidelines.
 Specifically, the first seven guidelines are intended to ensure the "reproducibility of the reported results based on the originating data" [direct quote from @nosekPromotingOpenResearch2015a].
@@
 replication" [@nosekPromotingOpenResearch2015a].
-In this context, 
-reproducibility can be defined as "re-performing the same analysis with the same code and the same data using a different analyst",
-and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
 These guidelines are: 1) Comprehensive
@@
 
-Response: We apologize; it appears that there were numerous errors in the parsing of the reference list after rendering the manuscript to HTML - the preferred format of Data Science.
+Response: We apologize; it appears that the reference list was not parsed correctly after rendering the manuscript to HTML - the preferred format of Data Science.
 The reference is indeed to the R-package.
@@
 3) Acknowledge the more comprehensive approach to dependency management offered by Docker
+
+The changes are extensive, so we will not quote them all here. Instead, we refer the Reviewer to the revised manuscript.
 
@@
 \Assignment{BV, AB, AL, CJ}
-\WorkInProgress
 \Done
+\Medium
 
@@
 
-We agree completely that it is important to make open data findable beyond just providing it freely.
-This is achieved by linking the GitHub repository to an OSF.io project page;
-on OSF, the project receives a DOI.
-It is also possible to give a separate DOI to specific resources, such as data files.
-We agree that this was not emphasized enough in the manuscript, and have rewritten the revision to clarify.
+Response: We agree completely that it is important to make open data findable beyond just providing it freely.
+In order to address the Reviewer's comment, we've revised the paper and the workflow vignette to emphasize more clearly that an essential step of the workflow is to connect the GitHub repository to a project page on the Open Science Framework - a standard repository for open research.
+We also devote more attention to the importance of DOIs; both for the project as a whole, and for specific resources, such as data files.
+Finally, we have elaborated on the discussion of the FAIR principles, emphasizing that WORCS is not designed to address these comprehensively, but pointing out specific ways in which WORCS is consistent with these guiding principles.
+Please see our response to Reviewer 2, comment 2 for more details.
 
-It is a pity that the Reviewer does not provide arguments to support a preference for Zenodo or FigShare over GitHub, because if there is a strong case to be made, we would be keen to include this in the paper.
+With regard to the specific platforms named by the Reviewer:
+We are not aware of arguments to support a strong preference for Zenodo or FigShare over GitHub.
 Without such arguments, as far as we are concerned, these repositories are largely interchangeable.
-
-We prefer storing the data on GitHub because they are then bundled with the
+One key advantage of storing the data on GitHub is that they are then bundled with the
 documentation (e.g., codebook), analysis code, and manuscript.
 Thus, readers have access to **all** materials required for replication in one repository.
+Moreover, the entire project can be readily updated from the console, and is version controlled.
 GitHub additionally offers user interaction with the materials through
@@
 users can document GitHub repositories in Zenodo and generate a DOI.
-This is redundant with Step 15 of the workflow, however:
-Connecting the GitHub repository to an OSF project accomplishes effectively the same.
-We do acknowledge, however, 
-that this part of the workflow was not clearly documented in the previous version of the manuscript, 
-and that the use of DOIs warrants further clarification.
-
-In the revision, we have attempted to address this comment in two ways: First, by
-clarifying Step 15 in the WORCS workflow, specifically emphasizing the importance of generating a DOI. 
-Second, by discussing the use of DOIs in the paper.
-
-The approach we advocate is to make all of a project's resources available in a 
-Step 15 of the workflow, "Connect your ‘OSF’ project page to the ‘Git’ remote repository",
-was intended to meet this goal.
-
-Digital object identifiers are the de-factor standard for making unique references to data sets. In line with the reviewer's suggestion,
-we suggest to use external providers to provide such DOIs while github is still the primary data storage. TODO!!!
-
-BV: Can you please argue for why GitHub is equivalent to Zenodo for this purpose?
-AL: Can you address the reference to FAIR principles in relation to GitHub?
-
-AB: "on second thought, it’s probably worth mentioning to the reviewer that we know these options but defend the OSF-centric approach for our purposes.
-My impression is that the main challenge are the different expectations and solutions across disciplines that are currently evolving largely in parallel.
-If we acknowledge that there are alternatives, this may please the reviewers while we still can defend R + OSF for the behavioral and social sciences.
-Other disciplines may have different foci, e.g., python + jupyter + docker etc...
-and no single manuscript will be able to cover everything."
+We now explicitly mention this option in the paper and workflow vignette.
+We also mention the potential redundancy with Steps 15-16 of the workflow, however:
+Connecting the GitHub repository to an OSF project and generating a DOI accomplishes effectively the same.
 
 \Assignment{BV, AB, AL, CJ}
-\WorkInProgress
+\Done
 \Hard
@@
 
-> For users of the `worcs` `R` package, we highly recommend referring to
+> We refer users of the `worcs` implementation in `R` to
 [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
@@
 > We provide a conceptual outline of the WORCS procedure below, based on WORCS
-Version 0.1.5.
+Version 0.1.6.
 The conceptual workflow recommends actions to take 
@@
 
-Response: We wish to emphasize that it was certainly unintended to claim that everyone should us R for open science.
+Response: We wish to emphasize that it was certainly unintended to suggest that everyone should us R for open science.
 To address this comment, we have done two things: First, throughout the paper, we have attempted to clarify the distinction between the platform independent conceptual workflow "WORCS", and its implementation for R users, `worcs`.
@@
 As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].
-Several arguments support our choice to implement this workflow first in `R` and RStudio, although we encourage developers to port the software to other languages.
+Several arguments support our choice to implement this workflow first in `R` and RStudio,
+although we encourage developers to port the software to other languages.
 
@@
 
-\Assignment{ALL}
-\WorkInProgress
+Response: We thank the Reviewer for these suggestions.
+Regarding the flowchart - we had already included one in the previous version of the manuscript.
+To address this comment, we have clarified the references in the text to this flowchart:
+
+> The WORCS workflow constitutes a step-by-step procedure that researchers can follow to make a 
+research project open and reproducible.
+These steps are elaborated in greater detail below.
+See Figure 1 for a flowchart that highlights important steps in different stages of the research process.
+
+Regarding the example project, we have now curated a list of all public `worcs` projects on GitHub, and we reference this list in several places in the manuscript, including:
+
+**Abstract**
+
+> The source code for the `R` package and manuscript,
+  and a list of [user examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
+  https://github.com/cjvanlissa/worcs.
+
+**Introducing the workflow**
+
+> For examples of how the workflow is used in practice,
+we maintain [a list of public `worcs` projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users) on the WORCS GitHub page.
+
+\Assignment{CJ}
+\Done
 \Medium
@@
 
-Anna-Lena?
+Response: To address this comment, we have made three changes:
 
-Response: 
-F: Dit heeft weer met DOIs te maken. Ik denk dat als we het gebruik van OSF de default van worcs maken (die mijn opmerking R1C7a) dat we hier een sterk punt kunnen maken mbt findability. 
-A: Eigenlijk hetzelfde antwoord als bij Findability: dit wordt heel eenvouding als we de OSF route voor data storage de default maken. 
-I: Het feit dat we alleen gebruik maken van open-source software maakt alle projecten die worcs gebruiken nogal interoperable. 
-R: Dit is wel een belangrijkt punt: Kunnen we opnemen in de workflow dat worcs-projecten by default een CC-BY licentie krijgen?
+1) We have changed the workflow, emphasizing the importance of creating a project on the OSF and connecting it to the GitHub repository, and emphasizing the importance of DOIs and a license
+2) We have rewritten the paragraph addressing the FAIR principles, with a more thorough discussion of how WORCS can help address each FAIR principle.
+We also state explicitly that *WORCS was not designed to comprehensively address these principles, but the workflow does facilitate meeting them.*
+3) We have made minor changes to the `worcs` package, e.g., to include information on the project readme about how readers can obtain access to data.
+
+This paragraph now reads:
+
+> Aside from the TOP Guidelines, the FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a] are an important standard for open science.
+These principles advocate that digital research objects should be Findable, Accessible, Interoperable and Reusable.
+Initially mainly promoted as principles for data, they are increasingly applied as a standard for other types of research output as well,
+most notably software [@lamprechtFAIRPrinciplesResearch2019].  
+  WORCS was not designed to comprehensively address these principles, but the workflow does facilitate meeting them.
+In terms of Findability, all files in GitHub projects are searchable by content and meta-data, both on the platform and through standard search engines.
+The workflow further recommends that users create a project page on the Open Science Framework – a recognized repository in the social sciences.
+When doing so, it is important to generate a Digital Object Identifier (DOI) for this project.
+A DOI is a persistent way to identify and connect to an object on the internet, which is crucial for findability.
+Additional DOIs can be generated through the OSF for specific resources, such as data sets.
+It is further worth noting that, optionally, [GitHub repositories can be connected to Zenodo](https://guides.github.com/activities/citable-code/):
+On Zenodo, users can store a snapshot of the repository, provide metadata, and generate a DOI for the project or specific resources.
+When a project is connected to the OSF, as recommended, these steps may be redundant.
+Finally, each worcs project has a [YAML file] (https://yaml.org/) that lists the data objects (and other meta-data) in the project, which will allow web crawlers to index worcs projects.
+We use this metadata, for example, to identify public `worcs` projects on GitHub.  
+  Accessibility is primarily safeguarded by hosting all project files in a public GitHub repository, which can be downloaded as a ZIP archive, and cloned or forked using the Git protocol.
+[GitHub is committed to long term accessibility of these resources]( https://archiveprogram.github.com/).
+We further encourage users to make their data accessible when possible.
+When data must be closed, the `worcs` function `closed_data()` adds a paragraph to the project readme, indicating how readers can contact the author for access.  
+  With regard to interoperability, the workflow recommends using plain-text files whenever possible, which ensures that code and meta-data are human- and machine-readable.
+Furthermore, the `worcs` package is implemented in R, which is open source.
+To promote reusability, we recommend users to specify an appropriate license for their projects.
+The `worcs` package includes the CC licenses, and suggests a CC-BY 4.0 license by default.
+This license is featured in the GitHub repository, and we encourage users to also display it on their project’s OSF page.
+To further facilitate reusability, `worcs` automatically generates a codebook for
+each data file.
+We encourage users to elaborate on the default codebook by adding
+variable description and categories, which would - in the future - enable
+indexing data by topic area.
 
 \Assignment{AL, BV, CJ}
-\WorkInProgress
+\Done
 \Medium
@@
 If the Reviewer feels differently, we are open to being convinced, however!
-Of course, since `worcs` is open source software, anybody is welcome to contribute this functionality in exchange for coauthorship, in line with the contribution guidelines [outlined here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
-The current team lacks relevant expertise to make this contribution.
+
+The current team lacks relevant expertise to make this contribution, but
+since `worcs` is open source software, anybody is welcome to contribute this functionality.
+Such a contribution would warrant coauthorship on the package, in line with the contribution guidelines [outlined here](https://github.com/cjvanlissa/worcs#contributing-and-contact-information).
 <!--What would this entail? Is it within scope?-->
@@
 
-To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in the paper and the Workflow vignette, that the steps are not necessarily linear:  *"Note that, although the steps are numbered for reference purposes, we acknowledge that the process of conducting research is not always linear."*.
+To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in two places in the paper, and in the Workflow vignette included with the `worcs` package, that the steps are not necessarily linear:
+
+**Introducing the workflow:**
+
+> Note that, although the steps are numbered for reference purposes,
+we acknowledge that the process of conducting research is not always linear.
+
+**Limitations:**
+
+> Another limitation is the fact that no single workflow can suit all research projects.
+Not all projects conform to the steps outlined here;
+some may skip steps, add steps, or follow similar steps in a different order.
+In principle, WORCS does not enforce a linear order, and allows extentensive user flexibility.
+
 Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have added more functions to increase the non-linear flexibility of the `worcs` package: `add_manuscript()` and `add_preregistration()`.
@@
 To clarify, we again state that our paper introduces a workflow for open science, based on objective criteria set out in the TOP-guidelines.
-Regarding novelty: Aside from the work in progress of our co-authors - which is focused on strict computational reproducibility and is compatible with the present workflow - this is one of the first efforts to translate the requirements of open science into a step-by-step workflow.
+Regarding novelty: Aside from the work-in-press of our co-authors - which is focused on strict computational reproducibility and is compatible with the present workflow - this is one of the first efforts to translate the requirements of open science into a step-by-step workflow.
 When you search Google Scholar for "open science workflow", the first hit that actually describes a workflow is the present manuscript.
@@
 
-> Git tracks changes to files on a line-by-line basis. The user can sture these changes by making a "commit": a 
+> Git tracks changes to files on a line-by-line basis.
+The user can store these changes by making a "commit": a 
 snapshot of the version controlled files.
@@
 Whether it will be adopted by researchers remains to be seen after publication.
-There is some prelimiary evidence that the workflow is being adopted (as of 31-10-2020):
+We can offer some prelimiary evidence that the workflow is already being adopted (as of 14-11-2020):
 
-1. The preprint has been downloaded 561 times, since being published on 31-05-2020
-1. The `worcs` R-package has been downloaded 2401 times from CRAN, since being published on 18-05-2020
-1. The GitHub project has been forked by users unknown to us 4 times, and starred (= watched) 32 times
-1. The video tutorial for worcs has been viewed 362 times, and liked 18 times (https://www.youtube.com/watch?v=uzjpN_yFeUU)
-1. WORCS is used to teach reproducibility at Utrecht University
-1. The lead author has given invited lectures on WORCS at the Erasmus University Rotterdam, Eindhoven University of Technology, and Max Planck Institute, Berlin
-1. The lead author has applied worcs in several research projects:
-	* https://github.com/cjvanlissa/storm_meta_father_mother
-	* https://github.com/cjvanlissa/veni_sysrev
-	* https://github.com/cjvanlissa/placebo_adhd
-	* https://github.com/cjvanlissa/schumpe
-	* https://github.com/cjvanlissa/coping_covid
+1.  The preprint has been downloaded 582 times, since being published on
+    31-05-2020
+2.  The `worcs` R-package has been downloaded 2567 times from CRAN,
+    since being published on 18-05-2020
+3.  The GitHub project has been forked 8 times, watched 5 times, and
+    starred 40 times
+4.  The lead author has given invited lectures on WORCS at:
+      - Max-Planck-Institute for Human Development (2020)  
+      - Open Science Community Rotterdam (2020)  
+      - Eindhoven University of Technology (2020)  
+5.  WORCS is currently used in the following GitHub repositories (sorted
+    by user):
+      - cjvanlissa
+          - [veni\_sysrev](https://github.com/cjvanlissa/veni_sysrev)  
+          - [IHBT-SEM](https://github.com/cjvanlissa/IHBT-SEM)  
+          - [schumpe](https://github.com/cjvanlissa/schumpe)  
+          - [storm\_meta\_father\_mother](https://github.com/cjvanlissa/storm_meta_father_mother)  
+          - [placebo\_adhd](https://github.com/cjvanlissa/placebo_adhd)  
+          - [mpib](https://github.com/cjvanlissa/mpib)  
+          - [coping\_covid](https://github.com/cjvanlissa/coping_covid)  
+      - IndrajeetPatil
+          - [worcs\_practice](https://github.com/IndrajeetPatil/worcs_practice)  
+      - EstherPlomp
+          - [RDM-survey](https://github.com/EstherPlomp/RDM-survey)  
+      - lisallreiber
+          - [worcs\_ws\_empty](https://github.com/lisallreiber/worcs_ws_empty)  
+      - KimBaldry
+          - [BIO-MATE](https://github.com/KimBaldry/BIO-MATE)
 
@@
 Response: We thank the Reviewer for this helpful suggestion!
-Although such user examples are not yet numerous, the lead author has six public WORCS repositories, and there are two public repositories by other users.
+Although such user examples are not yet numerous, the lead author has six public WORCS repositories, and there are four public repositories by other users.
 In our opinion, it would be most useful to have a continuously cumulating list of example repositories, and link that list in the paper - rather than to publish a static list.
@@
 We have embedded the scraper in the README.Rmd file on the worcs GitHub page, which is regularly updated.
-In the revised paper, we point readers to [this readme](https://github.com/cjvanlissa/worcs) for a list of user examples, in the Abstract, and when introducing the workflow.
+In the revised paper, we point readers to [a list of public `worcs` projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users) in the Abstract, and when introducing the workflow.
 
@@
 
-We agree that this is an important point, and have rewritten the paragraph comparing these two solutions (as cited below).
-We hope that this adresses the Reviewer's question.
-However, as the workflow of Peikert and Brandmaier is still in press, and the software implementation has not yet been fully developed - a more detailed comparison of these two approaches would be premature.
+We agree that this is an important point.
+We have attemptedto address it by rewriting the paragraph comparing these two solutions (as cited below).
+We hope that this is satisfactory.
+At this point in time, this is all we can say about the comparison of the two workflows, because the workflow of Peikert and Brandmaier is still in press, and the software implementation has not yet been fully developed.
 
@@
 \Assignment{CJ, AP, AB}
-\WorkInProgress
+\Done
 \Medium
@@
 Of course, we take this Reviewer suggestion very seriously, and intend to conduct a user evaluation after the paper is published and users begin to adopt the workflow.
-In the meantime, we present the following evidence for WORCS' potential for reuse:
-
-TODO
+Please refer to our response to Reviewer 1's comment 1, where we discuss our efforts to ensure the *potential for reuse*, and provide preliminary evidence for actual reuse/uptake of the workflow.
 
@@
 
-To address the Reviewer's comment, we have made two changes: First, we now explicitly state, in the paper and the Workflow vignette, that the steps are not necessarily linear:  *"Note that, although the steps are numbered for reference purposes, we acknowledge that the process of conducting research is not always linear."*.
-Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have now added functions to make non-linear workflows even more convenient, specifically: `add_manuscript()` and `add_preregistration()`.
+To address Reviewer 3's comment, we have made two changes: First, we now explicitly state, in two places in the paper, and in the Workflow vignette included with the `worcs` package, that the steps are not necessarily linear:
+
+**Introducing the workflow:**
+
+> Note that, although the steps are numbered for reference purposes,
+we acknowledge that the process of conducting research is not always linear.
+
+**Limitations:**
+
+> Another limitation is the fact that no single workflow can suit all research projects.
+Not all projects conform to the steps outlined here;
+some may skip steps, add steps, or follow similar steps in a different order.
+In principle, WORCS does not enforce a linear order, and allows extentensive user flexibility.
+
+Second, while it was always possible to manually add a manuscript or preregistration at a later point in time after creating a new project in RStudio - we have added more functions to increase the non-linear flexibility of the `worcs` package: `add_manuscript()` and `add_preregistration()`.
 
 \Assignment{AP, AB}
-\WorkInProgress
+\Done
 \Medium

### 2020-11-17 paper/reviews_ds.Rmd
remove REF in action letter

@@
 It is not clear what "claims" are made that are unsupported by evidence.
-In the revision, we have attempted to further clarify this goal from the paper onset: REF.
+In the revision, we have attempted to further clarify this goal from the paper onset:
 
-We do take the suggestion of performing a user evaluation very seriously, but we also feel that it would be premature to do so even before the paper is published.
+**Abstract**:
+
+> This paper introduces the Workflow for
+  Open Reproducible Code in Science (WORCS): A step-by-step
+  procedure that researchers can follow to make a research project open and
+  reproducible.
+  The purpose of the workflow is to lower the threshold for 
+  adoption of open science principles.
+  It is based on established best practices, and can be used either in
+  parallel to, or in absence of, top-down requirements by journals,
+  institutions, and funding bodies.
+  To facilitate widespread adoption, the WORCS principles have been implemented
+  in the R package `worcs`, which offers an RStudio 
+  project template and utility functions for specific workflow steps.
+  This paper introduces the conceptual workflow,
+  discusses how it meets different 
+  standards for open science, and addresses the functionality provided by the
+  `R` implementation, `worcs`.
+
+We again emphasize that we take the recommendation to perform user evaluation very seriously, but we also feel that it would be premature to do so even before the paper is published.
 We have built WORCS based on established tools and best practices, and we have gathered preliminary positive feedback from the community through workshops and teaching.

### 2020-11-17 paper/Manuscript_DS.Rmd
Fix reference to FOSTER definition

@@
 where research data, lab notes and other research processes are freely available,
-under terms that enable reuse, redistribution and reproduction of the research and its underlying data and methods" [@vicente-saezOpenScienceNow2018].
-In this definition, the term "science" refers to all scholarly disciplines.
+under terms that enable reuse, redistribution and reproduction of the research and its underlying data and methods" [the [FOSTER Open Science Definition](https://www.fosteropenscience.eu/foster-taxonomy/open-science-definition), as discussed in @sonjabezjakOpenScienceTraining2018].
+In this context, the term "science" refers to all scholarly disciplines.
 We further define the term *reproducible*, as used in the WORCS acronym, as being able to re-perform "the same analysis with the same code using a different analyst" [@patilVisualToolDefining2019].

### 2020-11-17 paper/Manuscript_DS.Rmd
Further discuss GitHub releases

@@
 (www.git-scm.com).
-Git tracks changes to files on a line-by-line basis.
+A project version controlled with Git is called a "repository", or repo.
+Git tracks changes to files in the repository on a line-by-line basis.
 The user can store these changes by making a "commit": a 
@@
 after making a mistake, or to replicate a previous version of the results).
-A project version controlled with Git is called a "Repository", or Repo.
 Git only version controls files explicitly committed by the user. Moreover, it
@@
 of GitHub allows you to "clone" (copy) a local Git repository to
-the GitHub website, as a backup or research archive. The social network aspect
-comes into play when a repository on GitHub is set to "public": This allows 
+the GitHub website, as a backup or research archive.
+The GitHub platform offers additional functionality over Git. 
+One such feature is that specific stages in the lifecycle of a project
+can be [tagged as a "release"](https://help.github.com/en/github/administering-a-repository/creating-releases):
+A named downloadable snapshot of a specific state of the project,
+that is prominently featured on the GitHub project page.
+The WORCS procedure encourages users to create releases to demarcate project milestones
+such as preregistration and submission to a journal.
+Releases are also useful to mark the submission of a revised manuscript, and publication.
+
+The social network aspect of GitHub
+comes into play when a repository is made "public": This allows 
 other researchers to peruse the repository and see how the work was done; clone

### 2020-11-17 paper/reviews_ds.Rmd
Further discuss GitHub releases

@@
 
+The GitHub function of creating "releases" also helps mark the "course-grained" progress of a research project.
+In the WORCS procedure, we encourage taggin the preprint and submitted version of a manuscript as a "release" on GitHub.
+Such releases are prominently featured on the GitHub project page, and are accompanied by a downloadable archive of the historic state of the project.
+To further address the Reviewer's comment, we now mention this in the manuscript text as well:
+
+> The GitHub platform offers additional functionality over Git. 
+One such feature is that specific stages in the lifecycle of a project
+can be [tagged as a "release"](https://help.github.com/en/github/administering-a-repository/creating-releases):
+A named downloadable snapshot of a specific state of the project,
+that is prominently featured on the GitHub project page.
+The WORCS procedure encourages users to create releases to demarcate project milestones
+such as preregistration and submission to a journal.
+Releases are also useful to mark the submission of a revised manuscript, and publication.
+
 \Assignment{CJ}

### 2020-11-17 paper/Manuscript.Rmd
add middle initials
sorry I forgot this before!
fix urls

--- 2020-06-03 paper/Manuscript.Rmd
@@
     affiliation: "2,8"
-  - name: "Barbara Vreede"
+  - name: "Barbara M.I. Vreede"
     affiliation: "9"
@@
 to the local 'Git' repository, and pushing it to the remote repository.
-Subsequently [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases)
+Subsequently [taging the release](https://docs.github.com/en/free-pro-team@latest/github/administering-a-repository/managing-releases-in-a-repository)
 as "Preregistration" on the remote repository renders it distinct from all other commits,
@@
 Within the workflow, the 'Git' cloud hosting services can be used to preregister a study simply by pushing a text document with the study plans to a 'Git' remote repository.
-By [tagging the commit](https://help.github.com/en/github/administering-a-repository/creating-releases) as a preregistration
+By [tagging the commit as a release](https://docs.github.com/en/free-pro-team@latest/github/administering-a-repository/managing-releases-in-a-repository), labelled "preregistration"
 (see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),

### 2020-11-17 paper/Manuscript_DS.Rmd
fix urls

@@
 One such feature is that specific stages in the lifecycle of a project
-can be [tagged as a "release"](https://help.github.com/en/github/administering-a-repository/creating-releases):
+can be [tagged as a "release"](https://docs.github.com/en/free-pro-team@latest/github/administering-a-repository/managing-releases-in-a-repository):
 A named downloadable snapshot of a specific state of the project,
@@
 to the local Git repository, and pushing it to the remote repository.
-Subsequently [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases)
+Subsequently [taging the release](https://docs.github.com/en/free-pro-team@latest/github/administering-a-repository/managing-releases-in-a-repository)
 as "Preregistration" on the remote repository renders it distinct from all other commits,
@@
 Within the workflow, the Git cloud hosting services can be used to preregister a study simply by pushing a text document with the study plans to a Git remote repository.
-By [tagging the commit](https://help.github.com/en/github/administering-a-repository/creating-releases) as a preregistration
+By [tagging the commit](https://docs.github.com/en/free-pro-team@latest/github/administering-a-repository/managing-releases-in-a-repository) as a preregistration
 (see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html)),

### 2020-11-17 paper/reviews_ds.Rmd
fix url function

@@
   mtch <- gsub("[\\)\\.,;]+$", "", mtch)
+  mtch <- unique(mtch)
   #out <- sapply(mtch, RCurl::url.exists)

### 2020-11-17 paper/Manuscript_DS.Rmd
update 2020-11-17 11:54:27

@@
 for existing users, a Docker build running `worcs` and its dependencies is available.
-The [stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html)
+The [stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/setup-docker.html)
 explains how to instantiate this build.
@@
 projects.
-Alternatively, Docker users can follow [the stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html).
+Alternatively, Docker users can follow [the stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/setup-docker.html).
 The `R` implementation of the workflow introduced earlier is detailed,

### 2020-11-17 paper/reviews_ds.Rmd
update 2020-11-17 11:54:27

@@
 ---
-title       : "Manuscript title"
-authors     : "First Author & Second Author"
-journal     : "Journal title"
-manuscript  : "Manuscript number"
+title       : "WORCS: A Workflow for Open Reproducible Code in Science"
+authors     : |
+    [Caspar J. Van Lissa](c.j.vanlissa@uu.nl),
+    Andreas M. Brandmaier,
+    Loek Brinkman,
+    Anna-Lena Lamprecht,
+    Aaron Peikert,
+    Marijn E. Struiksma, &
+    Barbara Vreede
+journal     : "Data Science"
 
@@
 output      : papaja::revision_letter_pdf
+urlcolor          : blue
 ---
@@
 projects.
-Alternatively, Docker users can follow [the stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html).
+Alternatively, Docker users can follow [the stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/setup-docker.html).
 

### 2020-11-17 paper/reviews_ds.Rmd
add different versions of action letter

@@
 bibliography      : ["worcs.bib"]  
-output      : papaja::revision_letter_pdf
+output      : html_document
 urlcolor          : blue
@@
 
-\RC{C1 There is no user evaluation, and therefore all the claims about worcs are unsupported by evidence.
-If the authors had included a compelling evaluation I would be in favor of accepting the paper.}
+## C1 There is no user evaluation, and therefore all the claims about worcs are unsupported by evidence. If the authors had included a compelling evaluation I would be in favor of accepting the paper.
 
@@
 
-\RC{C3 Unsupported claims (beyond the ones derived from an absence of evaluation): The paper claims in many parts that worcs helps make an entire research project Open and Reproducible.
+## C3 Unsupported claims (beyond the ones derived from an absence of evaluation): The paper claims in many parts that worcs helps make an entire research project Open and Reproducible.
 However, at the same time there is also a claim for not supporting replication.
 Since replication is generally understood as a necessary steps towards reproducibility, I find this kind of contradictory.
-Maybe the claims can be modified stating that worcs helps documentation and understanding, hence supporting FAIR.}
+Maybe the claims can be modified stating that worcs helps documentation and understanding, hence supporting FAIR.
 
@@
 
-\RC{C4: I am confused by the citation "Van Lissa, Caspar J., Aaron Peikert, and Andreas M.
+## C4: I am confused by the citation "Van Lissa, Caspar J., Aaron Peikert, and Andreas M.
 Brandmaier.
@@
 Is this referring to another paper or the actual package?
-If it's another paper then what is the contribution of this publication?}
+If it's another paper then what is the contribution of this publication?
 
@@
 
-\RC{C5: If the paper is generated as DDG, where are the functions that show it's capabilities?
-It is also not clear to me that some of the functionalities of Latex would be usable in RMarkdown, such as all the equations.}
+## C5: If the paper is generated as DDG, where are the functions that show it's capabilities?
+It is also not clear to me that some of the functionalities of Latex would be usable in RMarkdown, such as all the equations.
 
@@
 
-\RC{C6: It is not clear how worcs addresses depedency management. The section lists a series of tools that can be used, but the section is not very specific to worcs.}
+## C6: It is not clear how worcs addresses depedency management. The section lists a series of tools that can be used, but the section is not very specific to worcs.
 
@@
 
-\RC{C7a: The workflow of releasing data products on GitHub is not ideal.
+## C7a: The workflow of releasing data products on GitHub is not ideal.
 First, because datasets of 100 MBs usually require special handling.
-[...] The authors seem to not describe very well what would happen in data science experiments where the data size is significant.}
+[...] The authors seem to not describe very well what would happen in data science experiments where the data size is significant.
 
@@
 
-\RC{C7b: Second, it doesn't follow very well the FAIR principles for findability and attribution of data.
-Instead, repositories like Zenodo or FigShare could be user to address data storage, with proper metadata and its corresponding citation.}
+## C7b: Second, it doesn't follow very well the FAIR principles for findability and attribution of data.
+Instead, repositories like Zenodo or FigShare could be user to address data storage, with proper metadata and its corresponding citation.
 
@@
 
-\RC{C8: In the 3 phases listed, it is not clear what worcs does and does not for users.}
+## C8: In the 3 phases listed, it is not clear what worcs does and does not for users.
 
@@
 
-\RC{C9: In some cases it looks like the authors claim that everyone should use R for open science.
+## C9: In some cases it looks like the authors claim that everyone should use R for open science.
 While R has a strong community behind, I believe there are others with very strong support (e.g.
@@
 I am not sure if suggesting a shift to R is the right approach.
-Note that this is the impression I got from reading the paper, maybe it was unintended by the authors.}
+Note that this is the impression I got from reading the paper, maybe it was unintended by the authors.
 
@@
 
-\RC{C10: Vignete on citation leads to a 404.}
+## C10: Vignete on citation leads to a 404.
 
@@
   out <- sapply(mtch, httr::http_error)
-  out
+  out[out]
 }
@@
 
-\RC{C11: Synthetic data generation is fantastic, but it is not clear to me how does worcs actually keep the data consistent e.g.
+## C11: Synthetic data generation is fantastic, but it is not clear to me how does worcs actually keep the data consistent e.g.
 for training models.
-Does it follow a similar distribution to the source data?}
+Does it follow a similar distribution to the source data?
 
@@
 
-\RC{C12: Docker is deemed as complicated for novice users, but the setup vignete looks quite complex to me as well.
+## C12: Docker is deemed as complicated for novice users, but the setup vignete looks quite complex to me as well.
 Wouldn't a Docker container with worcs installed be easier to use?
-Maybe these aspects would be better highlighted in a user evaluation.}
+Maybe these aspects would be better highlighted in a user evaluation.
 
@@
 
-\RC{C1: I would suggest the authors include a flowchart for users that shows the decision making process for which tools/operations should be used in WORCS.
-An example project would also help users better understand the package.}
+## C1: I would suggest the authors include a flowchart for users that shows the decision making process for which tools/operations should be used in WORCS.
+An example project would also help users better understand the package.
 
@@
 
-\RC{C2: The claim that WORCS is amenable to the FAIR Guiding Principles should be justified.
+## C2: The claim that WORCS is amenable to the FAIR Guiding Principles should be justified.
 What are the solutions provided by WORCS that make a project interoperable and reusable?
-It would be improved with a more thorough discussion on how WORCS meets each FAIR principle.}
+It would be improved with a more thorough discussion on how WORCS meets each FAIR principle.
 
@@
 
-\RC{C3: Is WORCS providing any service for de-identification?}
+## C3: Is WORCS providing any service for de-identification?
 
@@
 
-\RC{C1: The process as described by the author for conducting research is assumed to be linear.
+## C1: The process as described by the author for conducting research is assumed to be linear.
 However, researchers often go back and forth on steps that they have already completed, and sometimes have to undo them completely and start from scratch.
-This does not seem to be tackled by the tool.}
+This does not seem to be tackled by the tool.
 
@@
 
-\RC{C2: The gist of the tool seems to allow researchers to document their steps and generate a scholarly document that is ready for submission.
+## C2: The gist of the tool seems to allow researchers to document their steps and generate a scholarly document that is ready for submission.
 In a sense, there is a similarity with existing notebooks such as Jupyter.
-So I was wondering why don’t the authors start from an existing popular tool such as Jupyter as the starting point (given the similarity in features that they would like to add), and extend Jupyter for instance with the new features such as dependency documentation and version control ?}
+So I was wondering why don’t the authors start from an existing popular tool such as Jupyter as the starting point (given the similarity in features that they would like to add), and extend Jupyter for instance with the new features such as dependency documentation and version control ?
 
@@
 
-\RC{C3: Regarding Version control and dependency documentation, the authors uses in their system existing solution, namely git and env, respectively.
+## C3: Regarding Version control and dependency documentation, the authors uses in their system existing solution, namely git and env, respectively.
 While I did not criticize reusing existing solution, on the contrary, I don't believe that simply using such solution add a substantial value.
-For example, git capture changes at the level of the line of code/text, whereas a researcher needs an abstraction that informs him/her on the coarse-grained elements that are significant from the point of view of development and research steps so as to have a clear idea of where s/he is within the process and help him/her make the decision of the next steps to undertake.}
+For example, git capture changes at the level of the line of code/text, whereas a researcher needs an abstraction that informs him/her on the coarse-grained elements that are significant from the point of view of development and research steps so as to have a clear idea of where s/he is within the process and help him/her make the decision of the next steps to undertake.
 
@@
 
-\RC{C4: To sum up, I think that the initial idea is interesting.
-However, I do not think that WORCS will be adopted by researchers or promote the state of the art of reproducibility, even if we focus on researchers who conduct their development and analysis using R.}
+## C4: To sum up, I think that the initial idea is interesting.
+However, I do not think that WORCS will be adopted by researchers or promote the state of the art of reproducibility, even if we focus on researchers who conduct their development and analysis using R.
 
@@
 
-\RC{C1: The abstract of the paper states that the manuscript "provides examples of the implementation of worcs in R", so I was expecting science projects that had been described using worcs.
-If these examples exist, they should be referenced and linked to from the manuscript, as they would provide a better demonstration of the practical utility of the workflow and software than tutorial style examples.}
+## C1: The abstract of the paper states that the manuscript "provides examples of the implementation of worcs in R", so I was expecting science projects that had been described using worcs.
+If these examples exist, they should be referenced and linked to from the manuscript, as they would provide a better demonstration of the practical utility of the workflow and software than tutorial style examples.
 
@@
 
-\RC{C2: There is a comparison in the paper to another R-based reproducible science solution (Peikert and Brandmaier (2019)).
+## C2: There is a comparison in the paper to another R-based reproducible science solution (Peikert and Brandmaier (2019)).
 The authors discuss some of the differences between these approaches, but the comparison is limited.
 It is not clear to me how different these approaches actually are, apart from ease of use.
-If scientists are already familiar with R, is there a significant learning curve with either solution?}
+If scientists are already familiar with R, is there a significant learning curve with either solution?
 
@@
 
-\RC{The reviewers agree that the paper has merit, but some of them also point to major shortcomings, in particular the lack of discussion of user adoption and of handling non-linear research processes.
-Note that according to our guidelines, resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse".}
+## The reviewers agree that the paper has merit, but some of them also point to major shortcomings, in particular the lack of discussion of user adoption and of handling non-linear research processes.
+Note that according to our guidelines, resource papers do not need to have a full-blown user evaluation, but we expect "sound evidence of its (potential for) reuse".
 

### 2020-11-21 R/codebook.R
Recommend SSH for git remote everywhere

@@ -1,6 +1,8 @@
 #' @title Create codebook for a dataset
 #' @description Creates a codebook for a dataset in 'R Markdown' format, and
-#' renders it to 'markdown' for 'GitHub'. Users can customize the 'R Markdown'
+#' renders it to 'markdown' for 'GitHub'. A codebook contains metadata and
+#' documentation for a data file.
+#' We urge users to customize the automatically generated 'R Markdown'
 #' document and re-knit it, for example, to add a paragraph with details on
 #' the data collection procedures. The variable descriptives are stored in
 #' a \code{.csv} file, which can be edited in 'R' or a spreadsheet program.
@@ -107,6 +109,9 @@ make_codebook <-
       write(
         c(
           file_contents,
+          "",
+          "A codebook contains documentation and metadata describing the contents, structure, and layout of a data file.",
+          "",
           "## Dataset description",
           paste0("The data contains ", dm[1], " cases and ", dm[2], " variables."),
           "",

### 2020-11-21 R/github.R

@@ -219,3 +219,16 @@ git_update <- function(message = paste0("update ", Sys.time()),
     }, error = function(e){col_message("Could not push local commits to remote repository.", success = FALSE)})
   )
 }
+
+
+parse_repo <- function(remote_repo, verbose = TRUE){
+  valid_repo <- grepl("^git@.+?\\..+?:.+?/.+?(\\.git)?$", remote_repo) | grepl("^https://.+?\\..+?/.+?/.+?(\\.git)?$", remote_repo)
+  if(!valid_repo){
+    col_message("Not a valid 'Git' remote repository address: ", remote_repo, success = FALSE, verbose = verbose)
+    return(NULL)
+  }
+  repo_url <- gsub("(^.+?@)(.*)$", "\\2", remote_repo)
+  repo_url <- gsub("(\\..+?):", "\\1/", repo_url)
+  repo_url <- gsub("\\.git$", "", repo_url)
+  gsub("^(https://)?", "https://", repo_url)
+}

### 2020-11-21 R/worcs_project.R

@@ -31,12 +31,11 @@ recommend_data <- c('library("worcs")',
 #' <https://creativecommons.org/licenses/>.
 #' @param use_renv Logical, indicating whether or not to use 'renv' to make the
 #' project reproducible. Default: TRUE. See \code{\link[renv]{init}}.
-#' @param remote_repo Character, 'https' link to the remote repository for
-#' this project. This link should have the form \code{https://[...].git}.
+#' @param remote_repo Character, 'SSH' address of the remote repository for
+#' this project. This link should have the form \code{git@[...].git}.
 #' If a valid remote repository link is provided, a commit will
 #' be made containing the 'README.md' file, and will be pushed to the remote
-#' repository. Default: 'https'. When no 'https' address is provided, an 'SSH'
-#' address of the form \code{git@[...].git} is also accepted.
+#' repository. Default: 'git@'.
 #' @param verbose Logical. Whether or not to print messages to the console
 #' during project creation. Default: TRUE
 #' @param ... Additional arguments passed to and from functions.
@@ -51,7 +50,7 @@ recommend_data <- c('library("worcs")',
 #'               preregistration = "None",
 #'               add_license = "None",
 #'               use_renv = FALSE,
-#'               remote_repo = "https")
+#'               remote_repo = "git@")
 #' setwd(old_wd)
 #' unlink(file.path(tempdir(), the_test))
 #' @rdname worcs_project
@@ -62,7 +61,7 @@ recommend_data <- c('library("worcs")',
 #' @importFrom prereg vantveer_prereg
 #' @importFrom methods formalArgs
 # @importFrom renv init
-worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregistration = "COS", add_license = "CC_BY_4.0", use_renv = TRUE, remote_repo = "https", verbose = TRUE, ...) {
+worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregistration = "COS", add_license = "CC_BY_4.0", use_renv = TRUE, remote_repo = "git@", verbose = TRUE, ...) {
   cl <- match.call(expand.dots = FALSE)
 
   # collect inputs
@@ -207,8 +206,9 @@ worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregist
   }
 
   # Connect to remote repo if possible
-  if(use_git & endsWith(remote_repo, ".git") &
-     (startsWith(remote_repo, "https://") | startsWith(remote_repo, "git@"))){
[…]

### 2020-11-30 paper/preprint.Rmd
add separate preprint with papaja layout

--- 2020-11-17 paper/Manuscript_DS.Rmd
@@
 ---
+title: |
+    | WORCS: A Workflow for Open Reproducible Code in Science
+    | `r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`
+
+shorttitle        : "WORKFLOW FOR OPEN SCIENCE"
+
 author:
-  name: |
-    [Caspar J. Van Lissa](c.j.vanlissa@uu.nl)$^{a,b}$
-    Andreas M. Brandmaier$^{c,d}$
-    Loek Brinkman$^{b,e}$
-    Anna-Lena Lamprecht$^{b,f}$
-    Aaron Peikert$^{c}$
-    Marijn E. Struiksma$^{b,g}$
-    Barbara Vreede$^{h}$
-
-  affiliation: | 
-      $a$: Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics  
-      $b$: Open Science Community Utrecht  
-      $c$: Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany  
-      $d$: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK  
-      $e$: UMC Utrecht  
-      $f$: Utrecht University, Department of Information and Computing Sciences  
-      $g$: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
-      $h$: University Library Utrecht
-title: "WORCS: A Workflow for Open Reproducible Code in Science"
-short:      "WORKFLOW FOR OPEN SCIENCE"
-shorttitle:      "WORKFLOW FOR OPEN SCIENCE"
+  - name: "Caspar J. Van Lissa"
+    affiliation: "1,2"
+    corresponding: yes
+    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
+    email: "c.j.vanlissa@uu.nl"
+  - name: "Andreas M. Brandmaier"
+    affiliation: "3,4"
+  - name: "Loek Brinkman"
+    affiliation: "2,5"
+  - name: "Anna-Lena Lamprecht"
+    affiliation: "2,6"
+    email: "a.l.lamprecht@uu.nl"
+  - name: "Aaron Peikert"
+    affiliation: "3"
+  - name: "Marijn E. Struiksma"
+    affiliation: "2,7"
+  - name: "Barbara M.I. Vreede"
+    affiliation: "8"
+
+affiliation:
+  - id: "1"
+    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
+  - id: "2"
+    institution: "Open Science Community Utrecht"
+  - id: "3"
+    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
+  - id: "4"
+    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
+  - id: "5"
+    institution: "University Medical Center Utrecht"
+  - id: "6"
+    institution: "Utrecht University, Department of Information and Computing Sciences"
+  - id: "7"
+    institution: "Utrecht Institute of Linguistics OTS"
+  - id: "8"
+    institution: "University Library Utrecht"
+
 abstract: |
@@
   https://github.com/cjvanlissa/worcs.
-authornote: "blabla"
-keywords:
-  # at least one keyword must be supplied
-  formatted: [open science, reproducibility, r, dynamic document generation, version control]
-  plain:     [open science, reproducibility, r, dynamic document generation, version control]
-preamble: >
-  \usepackage{amsmath}
+
+authornote: |
+  This is a preprint, not a peer-reviewed publication. Comments and
+  contributions are welcome. Please contact the lead author.
+  All authors of this paper, except the first, are listed in alphabetical order.
+  Author contributions are detailed in the document 'authors.csv' on the project
+  repository, at https://github.com/cjvanlissa/worcs.
+  The lead author is supported by a NWO Veni grant (NWO grant number
+  VI.Veni.191G.090).
+  We acknowledge Jeroen Ooms for offering feedback and suggesting the use of the
+  R package gert.
+keywords          : "open science; reproducibility; r; dynamic document generation; version control"
+wordcount         : "8257"
+
 bibliography      : ["worcs.bib"]
+
+figsintext        : yes
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+lineno            : yes
+mask              : no
+
+class             : "man"
+classoption       : "noextraspace"
+output            : papaja::apa6_pdf
 urlcolor          : blue
-output: rmarkdown::html_document
+knit              : worcs::cite_all
 ---
-<!--
-, "worcs_packages.bib"
-output: 
-  rmarkdown::html_document:
-    citation_package: biblatex Note: Citations are always handled by pandoc when format is not PDF
-knit: worcs::cite_all
-csl: apa.csl
-csl: apa.csl
--->
 

### 2021-02-02 paper/camera_ready/Manuscript_DS.Rmd
update 2021-02-02 12:15:00

--- 2020-11-17 paper/Manuscript_DS.Rmd
@@
   name: |
-    [Caspar J. Van Lissa](c.j.vanlissa@uu.nl)$^{a,b}$
-    Andreas M. Brandmaier$^{c,d}$
-    Loek Brinkman$^{b,e}$
-    Anna-Lena Lamprecht$^{b,f}$
-    Aaron Peikert$^{c}$
-    Marijn E. Struiksma$^{b,g}$
-    Barbara Vreede$^{h}$
+    [Caspar J.
+Van Lissa](c.j.vanlissa@uu.nl ) <a href="https://orcid.org/0000-0002-1825-0097"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{*,a,b}$
+    Andreas M.
+Brandmaier <a href="https://orcid.org/0000-0001-8765-6982"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{c,d}$
+    Loek Brinkman <a href="https://orcid.org/0000-0003-3997-1173"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{b,e}$
+    Anna-Lena Lamprecht <a href="https://orcid.org/0000-0003-1953-5606"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{b,f}$
+    Aaron Peikert <a href="https://orcid.org/0000-0001-7813-818X"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{c}$
+    Marijn E.
+Struiksma <a href="https://orcid.org/0000-0002-1166-1424"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{b,g}$
+    Barbara M.I.
+Vreede <a href="https://orcid.org/0000-0002-5023-4601"><img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> $^{h,i}$
 
   affiliation: | 
-      $a$: Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics  
+      $*$: Corresponding author. E-mail: c.j.vanlissa@uu.nl  
+      $a$: Department of Methodology & Statistics, Utrecht University, The Netherlands  
       $b$: Open Science Community Utrecht  
@@
       $d$: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK  
-      $e$: UMC Utrecht  
-      $f$: Utrecht University, Department of Information and Computing Sciences  
-      $g$: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
-      $h$: University Library Utrecht
+      $e$: University Medical Center, Utrecht University, The Netherlands  
+      $f$: Department of Information and Computing Sciences, Utrecht University, The Netherlands  
+      $g$: Utrecht Institute of Linguistics OTS, Utrecht University  
+      $h$: University Library, Utrecht University and Netherlands eScience Center  
+      $i$: Netherlands eScience Center
 title: "WORCS: A Workflow for Open Reproducible Code in Science"
@@
   https://github.com/cjvanlissa/worcs.
+
+  Keywords: open science, reproducibility, r, dynamic document generation, version control
 authornote: "blabla"
@@
 urlcolor          : blue
-output: rmarkdown::html_document
+output            : rmarkdown::html_document
 ---
@@
 Another important challenge is managing collaborations when only the lead author
-uses RMarkdown, and the coauthors use Word. In this case, it is possible to Knit
+uses RMarkdown, and the coauthors use Word. When using the `papaja` package to
+write APA-style papers, it is possible to Knit
 the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf`

### 2021-02-02 paper/camera_ready/preprint.Rmd
update 2021-02-02 12:15:00

--- 2020-11-30 paper/preprint.Rmd
@@
 title: |
-    | WORCS: A Workflow for Open Reproducible Code in Science
-    | `r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`
+    | WORCS: A Workflow for Open Reproducible Code in Science <!--`r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`-->
 
@@
 author:
-  - name: "Caspar J. Van Lissa"
-    affiliation: "1,2"
-    corresponding: yes
+  - name: "Caspar J. Van Lissa \\orcidicon{0000-0002-1825-0097}"
+    affiliation: "*,a,b"
     address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
     email: "c.j.vanlissa@uu.nl"
-  - name: "Andreas M. Brandmaier"
-    affiliation: "3,4"
-  - name: "Loek Brinkman"
-    affiliation: "2,5"
-  - name: "Anna-Lena Lamprecht"
-    affiliation: "2,6"
+  - name: "Andreas M. Brandmaier \\orcidicon{0000-0001-8765-6982}"
+    affiliation: "c,d"
+  - name: "Loek Brinkman \\orcidicon{0000-0003-3997-1173}"
+    affiliation: "b,e"
+  - name: "Anna-Lena Lamprecht \\orcidicon{0000-0002-5023-4601}"
+    affiliation: "b,f"
     email: "a.l.lamprecht@uu.nl"
-  - name: "Aaron Peikert"
-    affiliation: "3"
-  - name: "Marijn E. Struiksma"
-    affiliation: "2,7"
-  - name: "Barbara M.I. Vreede"
-    affiliation: "8"
+  - name: "Aaron Peikert \\orcidicon{0000-0001-7813-818X}"
+    affiliation: "c"
+  - name: "Marijn E. Struiksma \\orcidicon{0000-0002-1166-1424}"
+    affiliation: "b,g"
+  - name: "Barbara M.I. Vreede \\orcidicon{0000-0002-5023-4601}"
+    affiliation: "h,i"
 
 affiliation:
-  - id: "1"
-    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
-  - id: "2"
+  - id: "a"
+    institution: "Department of Methodology & Statistics, Utrecht University, The Netherlands"
+  - id: "b"
     institution: "Open Science Community Utrecht"
-  - id: "3"
+  - id: "c"
     institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
-  - id: "4"
+  - id: "d"
     institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
-  - id: "5"
-    institution: "University Medical Center Utrecht"
-  - id: "6"
-    institution: "Utrecht University, Department of Information and Computing Sciences"
-  - id: "7"
-    institution: "Utrecht Institute of Linguistics OTS"
-  - id: "8"
-    institution: "University Library Utrecht"
+  - id: "e"
+    institution: "University Medical Center, Utrecht University, The Netherlands"
+  - id: "f"
+    institution: "Department of Information and Computing Sciences, Utrecht University, The Netherlands"
+  - id: "g"
+    institution: "Utrecht Institute of Linguistics OTS, Utrecht University"
+  - id: "h"
+    institution: "University Library, Utrecht University and Netherlands eScience Center"
+  - id: "i"
+    institution: "Netherlands eScience Center"
 
@@
   Adopting open science
-  principles can be challenging and time-intensive, because doing so requires
-  substantial conceptual education and training in the use of new tools.
+  principles can be challenging, requiring
+  conceptual education and training in the use of new tools.
   This paper introduces the Workflow for
@@
   reproducible.
-  The purpose of the workflow is to lower the threshold for 
+  This workflow intends to lower the threshold for 
   adoption of open science principles.
@@
   conducting research that involves academic prose, analysis code,
-  and (optionally) tabular data.
+  and tabular data.
   However, the workflow is flexible enough to accommodate other scenarios,
-  and offers a sensible starting point for customized solutions.
+  and offers a starting point for customized solutions.
   The source code for the `R` package and manuscript,
-  and a list of [user examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
+  and a list of [examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
   https://github.com/cjvanlissa/worcs.
@@
 authornote: |
-  This is a preprint, not a peer-reviewed publication. Comments and
-  contributions are welcome. Please contact the lead author.
   All authors of this paper, except the first, are listed in alphabetical order.
@@
   R package gert. 
+  *Corresponding author. E-mail: c.j.vanlissa@uu.nl
 keywords          : "open science; reproducibility; r; dynamic document generation; version control"
-wordcount         : "8257"
 
@@
 classoption       : "noextraspace"
-output            : papaja::apa6_pdf
+output:
+  papaja::apa6_pdf:
+    includes:
+      in_header: "preamble.tex"
 urlcolor          : blue

### 2021-02-02 R/codebook.R

@@ -31,17 +31,19 @@
 #' successful. This function is mostly called for its side effect of rendering
 #' an 'R Markdown' codebook.
 #' @examples
-#' library(rmarkdown)
-#' library(knitr)
-#' filename <- tempfile("codebook", fileext = ".Rmd")
-#' make_codebook(iris, filename = filename, csv_file = NULL)
-#' unlink(c(
-#'   ".worcs",
-#'   filename,
-#'   gsub("\\.Rmd", "\\.md", filename),
-#'   gsub("\\.Rmd", "\\.html", filename),
-#'   gsub("\\.Rmd", "_files", filename)
-#' ), recursive = TRUE)
+#' if(rmarkdown::pandoc_available("1.14")){
+#'   library(rmarkdown)
+#'   library(knitr)
+#'   filename <- tempfile("codebook", fileext = ".Rmd")
+#'   make_codebook(iris, filename = filename, csv_file = NULL)
+#'   unlink(c(
+#'     ".worcs",
+#'     filename,
+#'     gsub("\\.Rmd", "\\.md", filename),
+#'     gsub("\\.Rmd", "\\.html", filename),
+#'     gsub("\\.Rmd", "_files", filename)
+#'   ), recursive = TRUE)
+#' }
 #' @rdname codebook
 #' @export
 #' @importFrom rmarkdown draft render

### 2021-02-02 R/save_load.R

@@ -25,7 +25,7 @@
 #' setwd(test_dir)
 #' worcs:::write_worcsfile(".worcs")
 #' df <- iris[1:5, ]
-#' open_data(df, codebook = "bla.Rmd")
+#' open_data(df, codebook = NULL)
 #' setwd(old_wd)
 #' unlink(test_dir, recursive = TRUE)
 #' @seealso open_data closed_data save_data

### 2021-02-02 tests/testthat/test-codebook.R

@@ -1,40 +1,42 @@
 test_that("codebook works for iris", {
+  if (rmarkdown::pandoc_available("1.14")){
+    old_wd <- getwd()
+    test_dir <- file.path(tempdir(), "codebook")
+    dir.create(test_dir)
+    setwd(test_dir)
 
-  old_wd <- getwd()
-  test_dir <- file.path(tempdir(), "codebook")
-  dir.create(test_dir)
-  setwd(test_dir)
+    worcs::make_codebook(iris)
 
-  worcs::make_codebook(iris)
+    expect_true(file.exists("codebook.Rmd"))
 
-  expect_true(file.exists("codebook.Rmd"))
-  if (rmarkdown::pandoc_available("1.12.3")){
     expect_true(file.exists("codebook.md"))
-  }
-  expect_true(file.exists("codebook.csv"))
-  contents <- readLines("codebook.Rmd", encoding = "UTF-8")
-  expect_true(any(contents == "The data contains 150 cases and 5 variables."))
 
-  setwd(old_wd)
-  unlink(test_dir, recursive = TRUE)
+    expect_true(file.exists("codebook.csv"))
+    contents <- readLines("codebook.Rmd", encoding = "UTF-8")
+    expect_true(any(contents == "The data contains 150 cases and 5 variables."))
+
+    setwd(old_wd)
+    unlink(test_dir, recursive = TRUE)
+  }
 })
 
 if(TRUE){
 test_that("codebook works for PlantGrowth with missings", {
-  df <- CO2
-  df[c(7:8), 1] <- NA
-  df[c(1, 25), 2] <- NA
+  if (rmarkdown::pandoc_available("1.14")){
+    df <- CO2
+    df[c(7:8), 1] <- NA
+    df[c(1, 25), 2] <- NA
 
-  old_wd <- getwd()
-  test_dir <- file.path(tempdir(), "codebook")
-  dir.create(test_dir)
-  setwd(test_dir)
+    old_wd <- getwd()
+    test_dir <- file.path(tempdir(), "codebook")
+    dir.create(test_dir)
+    setwd(test_dir)
 
-  worcs::make_codebook(df)
-  expect_true(file.exists("codebook.Rmd"))
-  if (rmarkdown::pandoc_available("1.12.3")){
+    worcs::make_codebook(df)
+    expect_true(file.exists("codebook.Rmd"))
+
     expect_true(file.exists("codebook.md"))
-  }
+
 
[…]

### 2021-02-02 README.Rmd
update 2020-11-17 11:54:27
update 2020-12-16 17:39:45
update 2021-02-02 12:15:00
update 2021-02-02 12:22:03

@@
 
-For most users, the recommended starting point is to [read the preprint paper](https://osf.io/zcvbs/),
+For most users, the recommended starting point is to [read the paper](https://osf.io/zcvbs/), currently in press in [Data Science](https://content.iospress.com/journals/data-science/Pre-press/Pre-press),
 which introduces the WORCS workflow, explains the underlying tools, and illustrates how the `worcs` package can be used to create a new project that follows the workflow.
@@
 E., & Vreede, B.
-(2020, May 29).
+(in press).
 WORCS: A Workflow for Open Reproducible Code in Science.
+Data Science, in press.
 https://doi.org/10.17605/OSF.IO/ZCVBS
@@
 
-1. The preprint has been downloaded 561 times, since being published on 31-05-2020
+1. The preprint has been downloaded 702 times, since being published on 31-05-2020 (currently in press in [Data Science](https://content.iospress.com/journals/data-science/Pre-press/Pre-press))
 1. The `worcs` R-package has been downloaded `r dl` times from CRAN, since being published on 18-05-2020

### 2021-02-03 paper/reviews_ds.Rmd

@@
 
-Sincerely, on behalve of all coauthors,
+Sincerely, on behalf of all coauthors,
 Caspar van Lissa
@@
 With regard to the specific platforms named by the Reviewer:
-We are not aware of arguments to support a strong preference for Zenodo or FigShare over GitHub.
+The nature of a repository created by worcs is initially a living project, for which the worcs package automates the workflow.
+This workflow includes version control through git, for which remote storage at Github (or similar websites) is a logical extension that fits the workflow.
+<!--We are not aware of arguments to support a strong preference for Zenodo or FigShare over GitHub.
 Without such arguments, as far as we are concerned, these repositories are largely interchangeable.
-One key advantage of storing the data on GitHub is that they are then bundled with the
+BV I disagree with this: they are very different repositories that do very different things, and for our purposes GH is clearly the one to go to.
+It's not a toss-up!-->
+One key advantage of storing the data on GitHub as well, as part of the project, is that they are then bundled with the
 documentation (e.g., codebook), analysis code, and manuscript.

### 2021-02-03 README.Rmd
try to add github actions CI

@@
 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
-[![Travis build status](https://travis-ci.org/cjvanlissa/worcs.svg?branch=master)](https://travis-ci.org/cjvanlissa/worcs)
+[![check-standard](https://github.com/cjvanlissa/worcs/workflows/check-standard/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
 [![CII Best Practices](https://bestpractices.coreinfrastructure.org/projects/3969/badge)](https://bestpractices.coreinfrastructure.org/projects/3969)
@@
 Contributions to the package must adhere to the [tidyverse style guide](https://style.tidyverse.org/).
-When contributing code, please add tests for that contribution to the `tests/testthat` folder, and ensure that these tests pass on [Travis-CI](https://travis-ci.org/github/cjvanlissa/worcs).
+When contributing code, please add tests for that contribution to the `tests/testthat` folder, and ensure that these tests pass<!-- on [Travis-CI](https://travis-ci.org/github/cjvanlissa/worcs)-->.
 

### 2021-02-03 README.Rmd
fix badge

@@
 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
-[![check-standard](https://github.com/cjvanlissa/worcs/workflows/check-standard/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
+[![R-CMD-check](https://github.com/cjvanlissa/worcs/workflows/R-CMD-check/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
 [![CII Best Practices](https://bestpractices.coreinfrastructure.org/projects/3969/badge)](https://bestpractices.coreinfrastructure.org/projects/3969)

### 2021-02-03 README.Rmd
Add codecov

--- 2021-02-02 README.Rmd
@@
 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
-[![Travis build status](https://travis-ci.org/cjvanlissa/worcs.svg?branch=master)](https://travis-ci.org/cjvanlissa/worcs)
+[![R-CMD-check](https://github.com/cjvanlissa/worcs/workflows/R-CMD-check/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
 [![CII Best Practices](https://bestpractices.coreinfrastructure.org/projects/3969/badge)](https://bestpractices.coreinfrastructure.org/projects/3969)
-<!--[![Codecov test
-coverage](https://codecov.io/gh/cjvanlissa/worcs/branch/master/graph/badge.svg)](https://codecov.io/gh/cjvanlissa/worcs?branch=master)-->
+[![test-coverage](https://github.com/cjvanlissa/worcs/workflows/test-coverage/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
 <!--[![DOI](http://joss.theoj.org/papers/10.21105/joss.00978/status.svg)](https://doi.org/10.21105/joss.00978)-->
@@
 Contributions to the package must adhere to the [tidyverse style guide](https://style.tidyverse.org/).
-When contributing code, please add tests for that contribution to the `tests/testthat` folder, and ensure that these tests pass on [Travis-CI](https://travis-ci.org/github/cjvanlissa/worcs).
+When contributing code, please add tests for that contribution to the `tests/testthat` folder, and ensure that these tests pass in the [GitHub Actions panel](https://github.com/cjvanlissa/worcs/actions/workflows/R-CMD-check).
 

### 2021-02-05 README.Rmd
Add codecov percentage badge

@@
 [![R-CMD-check](https://github.com/cjvanlissa/worcs/workflows/R-CMD-check/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
+[![codecov](https://codecov.io/gh/cjvanlissa/worcs/branch/master/graph/badge.svg?token=7S9XKDRT4M)](https://codecov.io/gh/cjvanlissa/worcs)
+[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-v2.0%20adopted-ff69b4.svg)](code_of_conduct.md)
 [![CII Best Practices](https://bestpractices.coreinfrastructure.org/projects/3969/badge)](https://bestpractices.coreinfrastructure.org/projects/3969)
-[![test-coverage](https://github.com/cjvanlissa/worcs/workflows/test-coverage/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
 <!--[![DOI](http://joss.theoj.org/papers/10.21105/joss.00978/status.svg)](https://doi.org/10.21105/joss.00978)-->

### 2021-03-17 README.Rmd
Update readme

@@
 
-1. The preprint has been downloaded 702 times, since being published on 31-05-2020 (currently in press in [Data Science](https://content.iospress.com/journals/data-science/Pre-press/Pre-press))
+1. The preprint has been downloaded 776 times, since being published on 31-05-2020 (currently in press in [Data Science](https://content.iospress.com/journals/data-science/Pre-press/Pre-press))
 1. The `worcs` R-package has been downloaded `r dl` times from CRAN, since being published on 18-05-2020

### 2021-11-17 README.Rmd
update 2021-11-17 15:21:16

@@
 
-For most users, the recommended starting point is to [read the paper](https://osf.io/zcvbs/), currently in press in [Data Science](https://content.iospress.com/journals/data-science/Pre-press/Pre-press),
+For most users, the recommended starting point is to [read the paper, published in Data Science](https://content.iospress.com/articles/data-science/ds210031),
 which introduces the WORCS workflow, explains the underlying tools, and illustrates how the `worcs` package can be used to create a new project that follows the workflow.
@@
 WORCS: A Workflow for Open Reproducible Code in Science.
-Data Science, in press.
-https://doi.org/10.17605/OSF.IO/ZCVBS
+Data Science, 2021.
+Data Science, vol.
+4, no.
+1, pp.
+29-49.
+DOI: 10.3233/DS-210031.
 
@@
 
-1. The preprint has been downloaded 776 times, since being published on 31-05-2020 (currently in press in [Data Science](https://content.iospress.com/journals/data-science/Pre-press/Pre-press))
+1. The preprint has been downloaded 1372 times, since being published on 31-05-2020
+1. The paper in [Data Science](https://content.iospress.com/articles/data-science/ds210031) has been cited 13 times
 1. The `worcs` R-package has been downloaded `r dl` times from CRAN, since being published on 18-05-2020
 1. The GitHub project has been forked `r worcs_repo$network_count` times, watched `r worcs_repo$subscribers_count` times, and starred `r worcs_repo$watchers` times
-1. The lead author has given invited lectures on WORCS at:`r invited`
-1. WORCS is currently used in the following GitHub repositories (sorted by user):`r tmp`
+1. The lead author has given invited lecturegs on WORCS at:`r invited`
+1. WORCS is currently used in the following public itHub repositories (sorted by user):`r tmp`
 

### 2022-06-23 R/worcs_project.R
change papaja to CRAN version

@@ -397,8 +397,8 @@ nice_tab <- function(tab){
 #' "mdpi_article", "mnras_article", "oup_article", "peerj_article",
 #' "plos_article", "pnas_article", "rjournal_article", "rsos_article",
 #' "sage_article", "sim_article", "springer_article", "tf_article"}.
-#' For more information about \code{APA6}, see the 'papaja' package, at
-#' <https://github.com/crsh/papaja>.
+#' For more information about \code{APA6}, see the
+#' \code{\link[papaja:papaja]{papaja}} package.
 #' For more information about \code{github_document}, see
 #' \code{\link[rmarkdown]{github_document}}. The remaining formats are
 #' documented in the 'rticles' package.

### 2022-06-23 R/worcs_project.R
update 2022-06-23 18:40:38

@@ -455,6 +455,8 @@ nice_tab <- function(tab){
 # )
 # out <- paste0("#' ", out)
 # cat(out, sep = '\n', file = "clipboard")
+## !!!ALSO ADD THEM TO THE PROJECT TEMPLATE FILE WORCS.DCF!!!!
+# cat(p, sep = ', ', file = "clipboard")
 #' @return No return value. This function is called for its side effects.
 #' @examples
 #' the_test <- "worcs_manuscript"
@@ -554,6 +556,8 @@ add_manuscript <- function(worcs_directory = ".", manuscript = "APA6", remote_re
 # )
 # out <- paste0("#' ", out)
 # cat(out, sep = '\n', file = "clipboard")
+## !!!ALSO ADD THEM TO THE PROJECT TEMPLATE FILE WORCS.DCF!!!!
+# cat(p, sep = ', ', file = "clipboard")
 #' \itemize{
 #'   \item{\code{'PSS'}}{Preregistration and Sharing Software (Krypotos,
 #'   Klugkist, Mertens, & Engelhard, 2019)}

### 2022-06-24 README.Rmd
change papaja to CRAN version
update 2022-06-23 18:40:38
update 2022-06-24 22:22:21

@@
 ``` r
-remotes::install_github("crsh/papaja", dependencies = TRUE, update = "never")
+install.packages("papaja", dependencies = TRUE, update = "never")
 
@@
 E., & Vreede, B.
-(in press).
+(2021).
 WORCS: A Workflow for Open Reproducible Code in Science.
-Data Science, 2021.
+Data Science.
 Data Science, vol.
@@
 
-## Adoption of WORCS by users
+<!-- ## Adoption of WORCS by users -->
 
-```{r, echo = FALSE, message=FALSE, warning=FALSE}
+```{r, eval = FALSE, echo = FALSE, message=FALSE, warning=FALSE}
 library(gh)
@@
 
-worcs_repo <- gh("/repos/cjvanlissa/worcs")
+# worcs_repo <- gh("/repos/cjvanlissa/worcs")
+# 
+# repos <- gh("GET /search/code", username = "cjvanlissa", q="worcs_version filename:.worcs")
+# 
+# worcs_repos <- data.frame(t(sapply(repos[["items"]], function(x){
+#   c(owner = x$repository$owner$login,
+#     name = x$repository$name,
+#     url = x$repository$html_url
+#   )
+# })))
 
-repos <- gh("GET /search/code", username = "cjvanlissa", q="worcs_version filename:.worcs")
+# worcs_repos <- worcs_repos[!((worcs_repos$owner %in% c("cjvanlissa", "cran") & worcs_repos$name == "worcs")| grepl("(penguin|\\btmp|demo|test)", tolower(worcs_repos$name))), ]
 
-worcs_repos <- data.frame(t(sapply(repos[["items"]], function(x){
-  c(owner = x$repository$owner$login,
-    name = x$repository$name,
-    url = x$repository$html_url
-  )
-})))
+# tmp <- paste0(sapply(unique(worcs_repos$owner), function(x){
+#   owner_repos <- worcs_repos[worcs_repos$owner == x, , drop = FALSE]
+#   paste0("\n    - ", x, "  ",
+#          paste0("\n        + [", owner_repos$name, "](", owner_repos$url, ")  ", collapse = ""))
+#   }),
+#   collapse = "")
 
-worcs_repos <- worcs_repos[!((worcs_repos$owner %in% c("cjvanlissa", "cran") & worcs_repos$name == "worcs")| grepl("(penguin|\\btmp|demo|test)", tolower(worcs_repos$name))), ]
-
-tmp <- paste0(sapply(unique(worcs_repos$owner), function(x){
-  owner_repos <- worcs_repos[worcs_repos$owner == x, , drop = FALSE]
-  paste0("\n    - ", x, "  ",
-         paste0("\n        + [", owner_repos$name, "](", owner_repos$url, ")  ", collapse = ""))
-  }),
-  collapse = "")
-
-invited <- read.table("https://raw.githubusercontent.com/cjvanlissa/resume/master/invited.csv", sep = ",", header = TRUE)
-invited <- invited[grepl("(worcs|workflow)", tolower(invited$Topic)), c("Institute", "Year", "City", "Country", "Date")]
-invited <- paste0(apply(invited, 1, function(x){paste0("\n    - ", x[1], " (", x[2], ")  ")}), collapse = "")
+# invited <- read.table("https://raw.githubusercontent.com/cjvanlissa/resume/master/invited.csv", sep = ",", header = TRUE)
+# invited <- invited[grepl("(worcs|workflow)", tolower(invited$Topic)), c("Institute", "Year", "City", "Country", "Date")]
+# invited <- paste0(apply(invited, 1, function(x){paste0("\n    - ", x[1], " (", x[2], ")  ")}), collapse = "")
 ```
 
-As of `r Sys.Date()`, these are indicators of the adoption of `worcs` by users:
+<!-- As of `r Sys.Date()`, these are indicators of the adoption of `worcs` by users: -->
 
-1. The preprint has been downloaded 1372 times, since being published on 31-05-2020
-1. The paper in [Data Science](https://content.iospress.com/articles/data-science/ds210031) has been cited 13 times
-1. The `worcs` R-package has been downloaded `r dl` times from CRAN, since being published on 18-05-2020
-1. The GitHub project has been forked `r worcs_repo$network_count` times, watched `r worcs_repo$subscribers_count` times, and starred `r worcs_repo$watchers` times
-1. The lead author has given invited lecturegs on WORCS at:`r invited`
-1. WORCS is currently used in the following public itHub repositories (sorted by user):`r tmp`
+<!-- 1. The preprint has been downloaded 1372 times, since being published on 31-05-2020 -->
+<!-- 1. The paper in [Data Science](https://content.iospress.com/articles/data-science/ds210031) has been cited 13 times -->
+<!-- 1. The `worcs` R-package has been downloaded r dl times from CRAN, since being published on 18-05-2020 -->
+<!-- 1. The GitHub project has been forked r worcs_repo$network_count times, watched r worcs_repo$subscribers_count` times, and starred r worcs_repo$watchers times -->
+<!-- 1. The lead author has given invited lecturegs on WORCS at: r invited -->
+<!-- 1. WORCS is currently used in the following public itHub repositories (sorted by user): r tmp -->
 

### 2022-06-24 R/check_installation.R

@@ -1,0 +1,135 @@
+#' @title Check worcs dependencies
+#' @description This function checks that all worcs dependencies are correctly
+#' installed, and suggests how to remedy any missing dependencies.
+#' @param what Character vector indicating which dependencies to check. Default:
+#' `"all"`. Options include: `c("dependencies", "git_user", "libgit2",
+#' "git_init", "git_add", "git_commit", "SSH", "tinytex")`.
+#' @return Logical, indicating whether all checks passed or not.
+#' @examples
+#' check_worcs_installation("dependencies")
+#' @rdname check_worcs_installation
+#' @export
+#' @importFrom gert user_is_configured libgit2_config git_init git_add git_commit
+#' @importFrom credentials ssh_key_info
+#' @importFrom tinytex pdflatex
+#' @importFrom utils packageDescription
+check_worcs_installation <- function(what = "all"){
+  pass <- list()
+  errors <- list()
+  if(any(c("all", "dependencies") %in% what)){
+    thesedeps <- get_deps()
+    available <- installed.packages()[, "Package"]
+    if(all(thesedeps %in% available)){
+      pass[["dependencies"]] <- TRUE
+    } else {
+      errors[["dependencies"]] <- thesedeps[!thesedeps %in% available]
+      errors[["dependencies"]] <- paste0("The following packages are not installed, run install.packages() for: ", paste0(errors[["dependencies"]], collapse = ", "))
+      pass[["dependencies"]] <- FALSE
+    }
+  }
+
+
+  # Git ---------------------------------------------------------------------
+
+  if(any(c("all", "git") %in% what)){
+    # Check user
+    pass[["git_user"]] <- gert::user_is_configured()
+    if(!pass[["git_user"]]) errors[["git_user"]] <- "No user configured; please run worcs::git_user(yourname, youremail, overwrite = TRUE)"
+
+    # Check libgit for SSH
+    libgit <- gert::libgit2_config()
+    pass[["libgit2"]] <- !is.null(libgit[["version"]])
+    if(pass[["libgit2"]]){
+      pass[["libgit2"]] <- libgit$ssh
+    }
+    if(!pass[["libgit2"]]){
+      errors[["libgit2"]] <- "libgit2 is not properly installed and you may not be able to use the SSH protocol to connect to Git remote repositories."
[…]

### 2022-06-24 R/codebook.R

@@ -31,7 +31,7 @@
 #' successful. This function is mostly called for its side effect of rendering
 #' an 'R Markdown' codebook.
 #' @examples
-#' if(rmarkdown::pandoc_available("1.14")){
+#' if(rmarkdown::pandoc_available("2.0")){
 #'   library(rmarkdown)
 #'   library(knitr)
 #'   filename <- tempfile("codebook", fileext = ".Rmd")

### 2022-06-24 R/descriptives.R

@@ -218,17 +218,18 @@ skew_kurtosis.default <-
     }
   }
 
-
+#' @importFrom usethis ui_oops ui_done
 col_message <- function (..., col = 30, success = TRUE, verbose = TRUE){
   if(verbose){
     txt <- do.call(paste0, list(...))
     # Check if this function is called from within an rmarkdown document.
     # If that is the case, the colorized messages can cause knitting errors.
     if(!any(grepl("rmarkdown", unlist(lapply(sys.calls(), `[[`, 1)), fixed = TRUE))){
-      cat(paste0(ifelse(success, "\033[0;32mv  \033[0m", "\033[0;31mX  \033[0m"),
-                 "\033[0;", col, "m", txt, "\033[0m", "\n"))
-    } else {
-      cat(paste0(ifelse(success, "v  ", "X  "), txt, "\n"))
+      if(success){
+        usethis::ui_done(txt)
+      } else {
+        usethis::ui_oops(txt)
+      }
     }
   }
 }

### 2022-06-24 R/worcs_project.R

@@ -464,7 +464,7 @@ nice_tab <- function(tab){
 #' dir.create(file.path(tempdir(), the_test))
 #' file.create(file.path(tempdir(), the_test, ".worcs"))
 #' add_manuscript(file.path(tempdir(), the_test),
-#'               manuscript = "github_document")
+#'               manuscript = "None")
 #' setwd(old_wd)
 #' unlink(file.path(tempdir(), the_test))
 #' @rdname add_manuscript
@@ -494,18 +494,25 @@ add_manuscript <- function(worcs_directory = ".", manuscript = "APA6", remote_re
       dir.create(man_dir_abs)
       if(manuscript == "apa6"){
         create_man_papaja(man_fn_abs, remote_repo = remote_repo)
-      } else {
-        if("rticles" %in% rownames(installed.packages())){
-          all_rticles <- ls(asNamespace("rticles"))
-          all_rticles <- all_rticles[endsWith(all_rticles, "_article")]
-          if(manuscript %in% all_rticles){
-            create_man_rticles(man_fn_abs, manuscript, remote_repo = remote_repo)
-          } else {
-            create_man_github(man_fn_abs, remote_repo = remote_repo)
-          }
-        } else {
-          create_man_github(man_fn_abs, remote_repo = remote_repo)
-        }
+      }
+      # all_rticles <- ls(asNamespace("rticles"))
+      # all_rticles <- all_rticles[endsWith(all_rticles, "_article")]
+      # dput(all_rticles, "clipboard")
+      if(manuscript %in% c("acm_article", "acs_article", "aea_article", "agu_article",
+                           "ajs_article", "amq_article", "ams_article", "arxiv_article",
+                           "asa_article", "bioinformatics_article", "biometrics_article",
+                           "copernicus_article", "ctex_article", "elsevier_article", "frontiers_article",
+                           "glossa_article", "ieee_article", "ims_article", "informs_article",
+                           "iop_article", "isba_article", "jasa_article", "jedm_article",
+                           "joss_article", "jss_article", "lipics_article", "mdpi_article",
+                           "mnras_article", "oup_article", "peerj_article", "pihph_article",
+                           "plos_article", "pnas_article", "rjournal_article", "rsos_article",
+                           "rss_article", "sage_article", "sim_article", "springer_article",
+                           "tf_article", "trb_article", "wellcomeor_article")){
+        create_man_rticles(man_fn_abs, manuscript, remote_repo = remote_repo)
[…]

### 2022-06-24 tests/testthat/test-codebook.R

@@ -1,5 +1,5 @@
 test_that("codebook works for iris", {
-  if (rmarkdown::pandoc_available("1.14")){
+  if (rmarkdown::pandoc_available("2.0")){
     old_wd <- getwd()
     test_dir <- file.path(tempdir(), "codebook")
     dir.create(test_dir)
@@ -22,7 +22,7 @@ test_that("codebook works for iris", {
 
 if(TRUE){
 test_that("codebook works for PlantGrowth with missings", {
-  if (rmarkdown::pandoc_available("1.14")){
+  if (rmarkdown::pandoc_available("2.0")){
     df <- CO2
     df[c(7:8), 1] <- NA
     df[c(1, 25), 2] <- NA

### 2022-06-25 README.Rmd
update 2022-06-25 11:08:06

@@
 
-Before installing the package, please read [this vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), which explains how to set up your computer for `worcs`.
+Before installing the package, please read [this vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), which explains how to set up your computer for `worcs` (using the CRAN version).
 
-After reading [the vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), you can install the development version of the `worcs` package from GitHub with:
+If you know what you're doing and you wish to install the development version of the `worcs` package from GitHub instead, you can use:
 
 ``` r
-if(!requireNamespace("remotes"))install.packages("remotes")
+if(!requireNamespace("remotes")) install.packages("remotes")
 remotes::install_github("cjvanlissa/worcs", dependencies = TRUE, update = "never")
-tinytex::install_tinytex()
-worcs::git_user("your_name", "your_email")
-```
-
-If you intend to write APA style manuscripts, you can additionally install the `papaja` package:
-
-``` r
-install.packages("papaja", dependencies = TRUE, update = "never")
-
 ```

### 2022-06-25 R/check_installation.R

@@ -14,6 +14,7 @@
 #' @importFrom tinytex pdflatex
 #' @importFrom utils packageDescription
 #' @importFrom gh gh_token
+#' @importFrom renv consent
 check_worcs_installation <- function(what = "all"){
   pass <- list()
   errors <- list()
@@ -138,6 +139,20 @@ check_worcs_installation <- function(what = "all"){
 
     if(!pass[["rmarkdown_pdf"]]) errors[["rmarkdown_pdf"]] <- "Rmarkdown could not render a PDF file."
   }
+
+
+  # renv --------------------------------------------------------------------
+
+  if(any(c("all", "renv") %in% what)){
+
+    pass[["renv_consent"]] <- !inherits(try({
+      { sink(tempfile()); tmp <- invisible(renv::consent()); sink(); }
+      if(!isTRUE(tmp)) stop()
+    }, silent = TRUE), "try-error")
+
+    if(!pass[["renv_consent"]]) errors[["renv_consent"]] <- "renv does not have consent yet; run renv::consent(provided = TRUE)"
+  }
+
   # Show results ------------------------------------------------------------
 
 

### 2022-07-16 R/save_load.R
update 2022-07-16 14:35:26

@@ -236,10 +236,10 @@ save_data <- function(data,
 
   if(create_labels){
     col_message("Storing value labels in '", fn_labels, "'.", success = TRUE, verbose = verbose)
-    Args_labs <- cl["data"]
-    Args_labs$variables <- names(data)[sapply(data, inherits, what = "factor")]
-    Args_labs$filename <- fn_write_labels
-    do.call(make_labels, Args_labs)
+    make_labels(data = data,
+                variables = names(data)[sapply(data, inherits, what = "factor")],
+                fn_write_labels
+                )
     # Add to worcs
     to_worcs <- list(filename = fn_worcs,
                      "data" = list(list("labels" = value_labels)),

### 2022-07-16 R/synthetic.R

@@ -117,7 +117,7 @@
 #'                            missingness_expression = imp_fun(data))
 #' @references Nowok, B., Raab, G.M and Dibben, C. (2016).
 #' synthpop: Bespoke creation of synthetic data in R. Journal of Statistical
-#' Software, 74(11), 1-26. <doi:10.18637/jss.v074.i11>.
+#' Software, 74(11), 1-26. \doi{10.18637/jss.v074.i11}.
 #' @rdname synthetic
 #' @export
 synthetic <- function(data,

### 2022-07-16 R/worcs_checklist.R

@@ -21,6 +21,6 @@
 #' @usage data(worcs_checklist)
 #' @references Van Lissa, C. J., Brandmaier, A. M., Brinkman, L., Lamprecht, A.,
 #' Peikert, A., , Struiksma, M. E., & Vreede, B. (2021)
-#' <doi:10.3233/DS-210031>.
+#' \doi{10.3233/DS-210031}.
 #' @format A data frame with 15 rows and 5 variables.
 NULL

### 2022-07-16 tests/testthat/test-codebook.R

@@ -4,7 +4,7 @@ test_that("codebook works for iris", {
     test_dir <- file.path(tempdir(), "codebook")
     dir.create(test_dir)
     setwd(test_dir)
-
+    data("iris")
     worcs::make_codebook(iris)
 
     expect_true(file.exists("codebook.Rmd"))

### 2022-07-16 tests/testthat/test-descriptives.R

@@ -1,3 +1,4 @@
+data("iris")
 tmp <- descriptives(iris)
 
 test_that("descriptives creates correct output", {

### 2022-07-16 R/synthetic.R
update 2022-07-16 20:05:08

@@ -70,6 +70,7 @@
 #' but loses all marginal relationships. These examples are not exhaustive, and
 #' users can even create custom functions.
 #' @examples
+#' \dontrun{
 #' # Example using the iris dataset and default ranger algorithm
 #' iris_syn <- synthetic(iris)
 #'
@@ -82,12 +83,12 @@
 #'                                 data = data.frame(.outcome = y,
 #'                                 xsynth)),
 #'           predict_expression = predict(model, newdata = xsynth))
-#'
+#' }
 #' # Example using bootstrapping:
 #' synthetic(iris,
 #'           model_expression = NULL,
 #'           predict_expression = sample(y, size = length(y), replace = TRUE))
-#'
+#' \dontrun{
 #' # Example with missing data, no imputation
 #' iris_missings <- iris
 #' for(i in 1:10){
@@ -115,6 +116,7 @@
 #' # Then, call synthetic() with this function as missingness_expression:
 #' iris_miss_syn <- synthetic(iris_missings,
 #'                            missingness_expression = imp_fun(data))
+#' }
 #' @references Nowok, B., Raab, G.M and Dibben, C. (2016).
 #' synthpop: Bespoke creation of synthetic data in R. Journal of Statistical
 #' Software, 74(11), 1-26. \doi{10.18637/jss.v074.i11}.

### 2022-07-19 R/check_installation.R
fix check_worcs_installation() and replace references to SSH with HTTPS

@@ -31,7 +31,9 @@ check_worcs_installation <- function(what = "all") {
       stop("Argument 'what' does not refer to any valid checks.")
     }
   }
-  out <- lapply(checkfuns, do.call, args = list())
+  out <- lapply(checkfuns, function(thisfun){
+    eval(str2lang(paste0("worcs::", thisfun, "()")))})
+
   worcs_checkres <- list(pass = do.call(c, lapply(out, `[[`, "pass")),
                          errors = do.call(c, lapply(out, `[[`, "errors")))
   class(worcs_checkres) <- c("worcs_check", class(worcs_checkres))
@@ -144,7 +146,7 @@ check_github <- function() {
   pass[["github_pat"]] <- isFALSE(gh::gh_token() == "")
   if (!pass[["github_pat"]])
     errors[["github_pat"]] <-
-    "You do not have a personal access token for GitHub; if you intend to use GitHub, consider creating one following the instructions at https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/creating-a-personal-access-token. Then, run credentials::set_github_pat('YourPAT') in R."
+    "You do not have a personal access token for GitHub; to fix this, run usethis::create_github_token() to create a PAT on GitHub; copy it, then run gitcreds::gitcreds_set() and paste the PAT when asked."
 
   # GitHub SSH
   temp <- tempfile()
@@ -155,9 +157,10 @@ check_github <- function() {
   output <- readLines(temp)
   pass[["github_ssh"]] <-
     isTRUE(any(grepl("success", output, fixed = TRUE)))
+  # Maybe check if *any* type of authentication is possible
   if (!pass[["github_ssh"]])
     errors[["github_ssh"]] <-
-    "Could not authenticate GitHub via SSH; if you intend to use GitHub, consult https://happygitwithr.com/rstudio-git-github.html"
+    "Could not authenticate GitHub via SSH, but that's OK. We recommend using a Personal Access Token (PAT). If you intend to use SSH with GitHub, consult https://happygitwithr.com/rstudio-git-github.html"
   out <- list(pass = pass, errors = errors)
   class(out) <- c("worcs_check", class(out))
   return(out)
@@ -177,7 +180,7 @@ check_ssh <- function() {
 
   if (!pass[["ssh"]])
     errors[["ssh"]] <-
-    "Could not find a valid SSH key; please consult https://happygitwithr.com/ssh-keys.html"
[…]

### 2022-07-19 R/worcs_project.R

@@ -29,11 +29,13 @@ recommend_data <- c('library("worcs")',
 #' <https://creativecommons.org/licenses/>.
 #' @param use_renv Logical, indicating whether or not to use 'renv' to make the
 #' project reproducible. Default: TRUE. See \code{\link[renv]{init}}.
-#' @param remote_repo Character, 'SSH' address of the remote repository for
-#' this project. This link should have the form \code{git@[...].git}.
+#' @param remote_repo Character, address of the remote repository for
+#' this project. This link should have the form
+#' \code{https://github.com[username][repo].git} (preferred) or
+#' \code{git@[...].git} (if using SSH).
 #' If a valid remote repository link is provided, a commit will
 #' be made containing the 'README.md' file, and will be pushed to the remote
-#' repository. Default: 'git@'.
+#' repository. Default: 'https'.
 #' @param verbose Logical. Whether or not to print messages to the console
 #' during project creation. Default: TRUE
 #' @param ... Additional arguments passed to and from functions.
@@ -48,7 +50,7 @@ recommend_data <- c('library("worcs")',
 #'               preregistration = "None",
 #'               add_license = "None",
 #'               use_renv = FALSE,
-#'               remote_repo = "git@")
+#'               remote_repo = "https")
 #' setwd(old_wd)
 #' unlink(file.path(tempdir(), the_test))
 #' @rdname worcs_project
@@ -59,7 +61,7 @@ recommend_data <- c('library("worcs")',
 #' @importFrom prereg vantveer_prereg
 #' @importFrom methods formalArgs
 # @importFrom renv init
-worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregistration = "cos_prereg", add_license = "CC_BY_4.0", use_renv = TRUE, remote_repo = "git@", verbose = TRUE, ...) {
+worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregistration = "cos_prereg", add_license = "CC_BY_4.0", use_renv = TRUE, remote_repo = "https", verbose = TRUE, ...) {
   cl <- match.call(expand.dots = FALSE)
 
   # collect inputs

### 2023-06-19 R/endpoint.R
update 2023-06-19 14:01:57.027623

@@ -1,0 +1,169 @@
+#' @title Add endpoint to WORCS project
+#' @description Add a specific endpoint to the WORCS project file. Endpoints are
+#' files that are expected to be exactly reproducible (e.g., a manuscript,
+#' figure, table, et cetera). Reproducibility is checked by ensuring the
+#' endpoint's checksum is unchanged.
+#' @param filename Character, indicating the file to be tracked as endpoint.
+#' Default: NULL.
+#' @param worcs_directory Character, indicating the WORCS project directory
+#' to which to save data. The default value "." points to the current directory.
+#' Default: '.'
+#' @param verbose Logical. Whether or not to print status messages to the
+#' console. Default: TRUE
+#' @param ... Additional arguments.
+#' @return No return value. This function is called for its side effects.
+#' @examples
+#' # Create directory to run the example
+#' old_wd <- getwd()
+#' test_dir <- file.path(tempdir(), "add_endpoint")
+#' dir.create(test_dir)
+#' setwd(test_dir)
+#' write_worcsfile(".worcs")
+#' writeLines("test", "test.txt")
+#' add_endpoint("test.txt")
+#' # Cleaning example directory
+#' setwd(old_wd)
+#' unlink(test_dir, recursive = TRUE)
+#' @rdname add_endpoint
+#' @seealso
+#'  \code{\link[worcs]{snapshot_endpoints}}
+#'  \code{\link[worcs]{check_endpoints}}
+#' @export
+#' @importFrom yaml read_yaml
+add_endpoint <- function(filename = NULL, worcs_directory = ".", verbose = TRUE, ...){
+  dn_worcs <- dirname(check_recursive(file.path(normalizePath(worcs_directory),
+                                                        ".worcs")))
+  fn_worcs <- file.path(dn_worcs, ".worcs")
+  worcsfile <- yaml::read_yaml(fn_worcs)
+  endpoints <- worcsfile[["endpoints"]]
+
+  # if(is.null(entry_point)){
+  #   if(is.null(worcsfile[["entry_point"]])){
+  #     stop("No 'entry_point' specified, and the project contains no existing entry points.")
+  #   } else {
+  #     if(length(worcsfile[["entry_point"]]) > 1){
+  #       stop("No 'entry_point' specified, and the project contains multiple entry points. Specify one of the following: ", paste0("'", names(worcsfile[["entry_point"]])))
+  #     }
+  #   }
+  #
[…]

### 2023-06-19 R/zzz.R

@@ -1,7 +1,7 @@
 # @importFrom rstudioapi versionInfo
 .onAttach <- function(libname, pkgname) {
   #correct_version <- versionInfo()$version >= "1.1.28"
-  print_message <- "\033[0;34mWelcome to WORCS: Workflow for Open Reproducible Code in Science. Please read the tutorial before using this package, and consider citing it:\n  Van Lissa and colleagues (2020) <doi:10.17605/OSF.IO/ZCVBS>\033[0m"
+  print_message <- "\033[0;34mWelcome to WORCS: Workflow for Open Reproducible Code in Science. Please read the tutorial before using this package, and consider citing it:\n  Van Lissa and colleagues (2020) <doi:10.3233/DS-210031>\033[0m"
   if(!has_git()){
     print_message <- paste0(print_message, "\n\033[0;31mCould not find a working installation of 'Git', which is required to safeguard the transparency and reproducibility of your project. Please connect 'Git' by following the steps described in this vignette:\n  vignette('setup', package = 'worcs')\033[0m")
   }

### 2023-06-19 README.Rmd
update 2023-06-19 14:58:18.055905

@@
 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
-[![R-CMD-check](https://github.com/cjvanlissa/worcs/workflows/R-CMD-check/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
-[![codecov](https://codecov.io/gh/cjvanlissa/worcs/branch/master/graph/badge.svg?token=7S9XKDRT4M)](https://codecov.io/gh/cjvanlissa/worcs)
+[![R-CMD-check](https://github.com/cjvanlissa/worcs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cjvanlissa/worcs/actions/workflows/R-CMD-check.yaml)
+[![Codecov test coverage](https://codecov.io/gh/cjvanlissa/worcs/branch/master/graph/badge.svg)](https://app.codecov.io/gh/cjvanlissa/worcs?branch=master)
 [![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-v2.0%20adopted-ff69b4.svg)](code_of_conduct.md)

### 2023-07-21 R/endpoint.R
update 2023-07-21 21:26:59.608573

@@ -181,3 +181,19 @@ check_endpoints <- function(worcs_directory = ".", verbose = TRUE, ...){
   }
   return(invisible(all(replicates)))
 }
+
+#' @title Set up GitHub Actions to Check Endpoints
+#' @description Sets up a GitHub Action to perform continuous integration (CI)
+#' for a WORCS project. CI automatically evaluates `check_endpoints()`
+#' at each push or pull request.
+#' @return No return value. This function is called for its side effects.
+#' @seealso
+#'  \code{\link[usethis]{use_github_action}}
+#'  \code{\link[worcs]{add_endpoint}}
+#'  \code{\link[worcs]{check_endpoints}}
+#' @export
+#' @importFrom usethis use_github_action
+github_action_check_endpoints <- function(){
+
+  usethis::use_github_action(url = "https://github.com/cjvanlissa/actions/blob/main/worcs_endpoints.yaml", badge = TRUE)
+}

### 2023-07-21 R/save_load.R

@@ -453,7 +453,7 @@ check_metadata <- function(x, codebook, value_labels){
 store_checksum <- function(filename, entry_name = filename, worcsfile = ".worcs") {
   # Compute checksum on loaded data to ensure conformity
   cs <- digest(object = filename, file = TRUE)
-  checkworcs(dirname(filename), iserror = FALSE)
+  checkworcs(dirname(worcsfile), iserror = FALSE)
   checksums <- list(cs)
   names(checksums) <- entry_name
   do.call(write_worcsfile,

### 2023-09-12 R/recipes.R
change implementation of detecting binary files

@@ -65,6 +65,8 @@ add_recipe <- function(worcs_directory = ".", recipe = "rmarkdown::render('manus
 #' Default: '.'
 #' @param verbose Logical. Whether or not to print status messages to the
 #' console. Default: `TRUE`
+#' @param check_endpoints Logical. Whether or not to call `check_endpoints()`
+#' after reproducing the recipe. Default: `TRUE`
 #' @param ... Additional arguments.
 #' @return No return value. This function is called for its side effects.
 #' @examples
@@ -75,7 +77,6 @@ add_recipe <- function(worcs_directory = ".", recipe = "rmarkdown::render('manus
 #' setwd(test_dir)
 #' file.create(".worcs")
 #' worcs:::add_recipe(recipe = 'writeLines("test", "test.txt")')
-#' worcs:::reproduce()
 #' # Cleaning example directory
 #' setwd(old_wd)
 #' unlink(test_dir, recursive = TRUE)

### 2023-09-12 R/save_load.R

@@ -463,16 +463,11 @@ cs_fun <- function(filename, worcsfile = ".worcs"){
   dn_worcs <- dirname(check_recursive(file.path(normalizePath(worcsfile))))
   fn_rel <- path_rel_worcs(filename, dn_worcs)
   tryCatch({
-    git_record <- tryCatch(system2("git", paste0('-C "', dirname(worcsfile), '" ls-files --eol'), stdout = TRUE), error = function(e){ stop() }, warning = function(w){
-      message("This worcs project is not version controlled with Git.")
-      stop()
-    })
-    git_record <- git_record[grepl(fn_rel, git_record, fixed = TRUE)]
-    git_record <- strsplit(git_record[1], split = "\\s+")[[1]][c(1:2)]
-    if(isTRUE(any(grepl("/-text", git_record, fixed = TRUE) | grepl("/lf", git_record, fixed = TRUE) | grepl("/cr", git_record, fixed = TRUE)))){
+    if(is_binary(fn_rel)){
+      digest::digest(filename, file = TRUE)
+    } else {
       stop()
     }
-    digest::digest(filename, file = TRUE)
   }, error = function(e){
     suppressWarnings(digest::digest(paste0(readLines(filename), collapse = ""), serialize = FALSE, file = FALSE))
   })
@@ -685,4 +680,9 @@ path_rel_worcs <- function(fn, dn_worcs = NULL, worcs_directory = "."){
     stop("File path must be inside of the worcs project file.", call. = FALSE)
   }
   return(do.call(file.path, as.list(fn[-seq_along(dn_worcs)])))
+}
+
+is_binary <- function(x){
+  out <- tryCatch(system2("file", args = paste0("--mime '", x, "'"), stdout = TRUE), error = function(e){ stop() })
+  return(isFALSE(grepl(": text/", out, fixed = TRUE)))
 }

### 2023-09-12 tests/testthat/test-reproduce.R

@@ -8,13 +8,13 @@ test_that("add_recipe is evaluated by reproduce()", {
   on.exit({setwd(old_wd); unlink(test_dir, recursive = TRUE)}, add = TRUE)
 
   add_recipe(recipe = 'writeLines("test", "test.txt")')
-  reproduce()
+  reproduce(check_endpoints = FALSE)
   expect_true(file.exists("test.txt"))
   file.remove(file.path(test_dir, ".worcs"))
   worcs:::write_worcsfile(file.path(test_dir, ".worcs"))
 
   add_recipe(recipe = 'Rscript -e """writeLines("test", "test.txt")"""', terminal = TRUE)
-  reproduce()
+  reproduce(check_endpoints = FALSE)
   expect_true(file.exists("test.txt"))
 
 })

### 2023-09-13 tests/testthat/test-can_clone.R
add test for gert::clone

@@ -1,0 +1,18 @@
+test_that("gert can clone repos", {
+  the_test <- "gert_clone"
+  old_wd <- getwd()
+  test_dir <- file.path(tempdir(), the_test)
+  dir.create(test_dir)
+  setwd(test_dir)
+  on.exit({setwd(old_wd); unlink(test_dir, recursive = TRUE)}, add = TRUE)
+
+  gert::git_clone("https://github.com/cjvanlissa/actions.git", path = test_dir)
+  # gert::git_init(path = test_dir)
+  # if(isFALSE(gert::user_is_configured())){
+  #   gert::git_config_set(name = "user.name", value = "testuser", repo = ".")
+  #   gert::git_config_set(name = "user.email", value = "c.j.vanlissa@tilburguniversity.edu", repo = ".")
+  # }
+  expect_true(file.exists("worcs_reproduce.yaml"))
+})
+
+

### 2024-01-24 R/targets.R
update 2024-01-24 18:09:14.138723

@@ -1,0 +1,71 @@
+#' @title Add targets to WORCS Project
+#' @description Add a computational pipeline to a `worcs` project using the
+#' `targets` and `tarchetypes` packages (which must be installed). See those
+#' packages for extensive documentation.
+#' @param worcs_directory Character, indicating the WORCS project directory
+#' to which to save data. The default value "." points to the current directory.
+#' Default: '.'
+#' @param verbose Logical. Whether or not to print status messages to the
+#' console. Default: `TRUE`
+#' @param ... Arguments passed to `targets::use_targets()`.
+#' @return No return value. This function is called for its side effects.
+#' @examples
+#' # Create directory to run the example
+#' old_wd <- getwd()
+#' test_dir <- file.path(tempdir(), "targets")
+#' dir.create(test_dir)
+#' setwd(test_dir)
+#' file.create(".worcs")
+#' add_targets()
+#' # Cleaning example directory
+#' setwd(old_wd)
+#' unlink(test_dir, recursive = TRUE)
+#' @rdname add_targets
+#' @export
+add_targets <- function (worcs_directory = ".", verbose = TRUE, ...){
+  checkworcs(worcs_directory, iserror = TRUE)
+  fn_worcs <- path_abs_worcs(".worcs", worcs_directory)
+  if (!file.exists(fn_worcs)) {
+    stop(".worcs file not found.")
+  }
+  if(!(requireNamespace("targets", quietly = TRUE) & requireNamespace("tarchetypes", quietly = TRUE))) {
+    c(requireNamespace("targets", quietly = TRUE) & requireNamespace("tarchetypes", quietly = TRUE))
+    worcs:::col_message("Could not find required packages; please run ",
+
+                        paste0(c(c("install.packages('targets')", "")[requireNamespace("targets", quietly = TRUE)+1L], c("install.packages('tarchetypes')", "")[requireNamespace("tarchetypes", quietly = TRUE)+1L]), collapse = "; "),
+                        " then try again."
+                        , success = FALSE
+    )
+    return()
+  } else {
+
+    worcs_file <- yaml::read_yaml(fn_worcs)
+    tryCatch({
+      targets::use_targets(...)
+      col_message("Added targets to project.")
+    }, error = function(e){col_message("Could not add targets to project.", success = FALSE)})
[…]

### 2024-01-24 R/targets.R
update 2024-01-24 19:07:20.321432

@@ -30,7 +30,7 @@ add_targets <- function (worcs_directory = ".", verbose = TRUE, ...){
   }
   if(!(requireNamespace("targets", quietly = TRUE) & requireNamespace("tarchetypes", quietly = TRUE))) {
     c(requireNamespace("targets", quietly = TRUE) & requireNamespace("tarchetypes", quietly = TRUE))
-    worcs:::col_message("Could not find required packages; please run ",
+    col_message("Could not find required packages; please run ",
 
                         paste0(c(c("install.packages('targets')", "")[requireNamespace("targets", quietly = TRUE)+1L], c("install.packages('tarchetypes')", "")[requireNamespace("tarchetypes", quietly = TRUE)+1L]), collapse = "; "),
                         " then try again."
@@ -50,7 +50,7 @@ add_targets <- function (worcs_directory = ".", verbose = TRUE, ...){
 
     to_worcs <- list(filename = fn_worcs, modify = TRUE)
     # Change entry point
-    if(file.exists(worcs:::path_abs_worcs("run.r", worcs_directory = worcs_directory))){
+    if(file.exists(path_abs_worcs("run.r", worcs_directory = worcs_directory))){
       col_message("Setting entry point to 'run.r'.", verbose = verbose)
       to_worcs$entry_point <- "run.r"
       to_worcs$recipe <- list(recipe = "source('run.r')", terminal = FALSE)
@@ -63,21 +63,21 @@ add_targets <- function (worcs_directory = ".", verbose = TRUE, ...){
     do.call(write_worcsfile, to_worcs)
 
     # Add manuscript to pipeline
-    if(file.exists(worcs:::path_abs_worcs("manuscript/manuscript.rmd", worcs_directory = worcs_directory)) & file.exists(worcs:::path_abs_worcs("_targets.R", worcs_directory = worcs_directory))){
+    if(file.exists(path_abs_worcs("manuscript/manuscript.rmd", worcs_directory = worcs_directory)) & file.exists(path_abs_worcs("_targets.R", worcs_directory = worcs_directory))){
       # First, add manuscript to pipeline
-      lnz <- readLines(worcs:::path_abs_worcs("_targets.R", worcs_directory = worcs_directory))
+      lnz <- readLines(path_abs_worcs("_targets.R", worcs_directory = worcs_directory))
       if(all(tail(lnz, 2) == c("  )", ")"))){
         col_message("Adding rmarkdown manuscript to targets pipeline.", verbose = verbose)
[…]

### 2024-02-03 README.Rmd

--- 2022-06-25 README.Rmd
@@
 
-# WORCS <a href='https://osf.io/zcvbs/'><img src='https://github.com/cjvanlissa/worcs/raw/master/docs/worcs_icon.png' align="right" height="139" /></a>
+# WORCS <a href="https://cjvanlissa.github.io/worcs/"><img src="man/figures/logo.png" align="right" height="120" alt="worcs website" /></a>
 
 <!-- README.md is generated from README.Rmd. Please edit that file -->
+
+<!-- badges: start --->
 [![CRAN status](https://www.r-pkg.org/badges/version/worcs)](https://cran.r-project.org/package=worcs)
 [![CRAN RStudio mirror downloads](https://cranlogs.r-pkg.org/badges/grand-total/worcs?color=blue)](https://r-pkg.org/pkg/worcs)
-[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
-[![R-CMD-check](https://github.com/cjvanlissa/worcs/workflows/R-CMD-check/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
-[![codecov](https://codecov.io/gh/cjvanlissa/worcs/branch/master/graph/badge.svg?token=7S9XKDRT4M)](https://codecov.io/gh/cjvanlissa/worcs)
+[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html#maturing)
+[![R-CMD-check](https://github.com/cjvanlissa/worcs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cjvanlissa/worcs/actions/workflows/R-CMD-check.yaml)
+[![Codecov test coverage](https://app.codecov.io/gh/cjvanlissa/worcs/branch/master/graph/badge.svg)](https://app.codecov.io/gh/cjvanlissa/worcs?branch=master)
 [![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-v2.0%20adopted-ff69b4.svg)](code_of_conduct.md)
@@
 <!--[![DOI](http://joss.theoj.org/papers/10.21105/joss.00978/status.svg)](https://doi.org/10.21105/joss.00978)-->
+<!-- badges: end --->
 

### 2024-09-13 tests/testthat/test-targets.R
update 2024-09-13 10:55:01.014339

@@ -62,7 +62,7 @@ test_that("targets works with target markdown", {
   expect_true(file.exists("_targets.html"))
   file.remove("_targets.html")
   worcs::reproduce()
-  expect_true(file.exists("_targets.html"))
+  if(!file.exists("_targets.html")) stop()
 
   setwd(old_wd)
 })

### 2024-09-13 inst/rstudio/templates/project/resources/_targets.rmd

---
  title: "Target Markdown"
  output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Target Markdown is a powerful R Markdown interface for reproducible analysis pipelines, and the chapter at https://books.ropensci.org/targets/markdown.html walks through it in detail. This R Markdown report the example from the chapter. Try it out in both interactive and non-interactive modes, either by running the code chunks in different ways or setting the `tar_interactive` chunk option.

# Setup

If you are using old versions of `targets` (<= 0.7.0) and/or `knitr` (<= 1.33), you will need to load the `targets` package in the R Markdown document in order for Target Markdown code chunks to work.

```{r}
library(targets)
```

Near the top of the document, you may also wish to remove the `_targets_r` directory previously written by non-interactive runs of the report. Otherwise, your pipeline may contain superfluous targets.

```{r}
library(targets)
tar_unscript()
```

# Targets

Our first target borrows the `airquality` dataset built into base R.

```{targets raw-data}
tar_target(raw_data, airquality)
```

Our next targets preprocess the data, make a histogram, and fit a model.

```{targets downstream-targets}
list(
  tar_target(data, {raw_data[complete.cases(airquality), ]}),
  tar_target(hist, hist(data$Ozone))
)
```

Set the `tar_simple` chunk option to `TRUE` to define a single target with the command in the code chunk. The chunk below only contains `biglm(Ozone ~ Wind + Temp, data)` in the source, but because `tar_simple` is `TRUE`, it is shorthand for `tar_target(name = fit, command = biglm(Ozone ~ Wind + Temp, data))`. All other arguments to `tar_target()` are set to their default values (configurable with `tar_option_set()`).

```{targets fit, tar_simple = TRUE}
lm(Ozone ~ Wind + Temp, data)
```

# Pipeline

If you ran all the `{targets}` chunks in non-interactive mode, then your R scripts are set up to run the pipeline.

```{r}
tar_make()
```

# Output

You can retrieve results from the `_targets/` data store using `tar_read()` or `tar_load()`.

```{r, message = FALSE}
tar_read(fit)
```

```{r}
tar_read(hist)
```

At this point, you can go back and run `{targets}` chunks in interactive mode without interfering with the code or data of the non-interactive pipeline.

### 2024-12-29 R/add_synthetic.R
update 2024-12-29 13:11:08.818181

@@ -97,38 +97,45 @@ add_synthetic <- function(data,
 
 
   # Synthetic data
-  col_message("Storing synthetic data in '", fn_write_synth_rel, "' and updating the checksum in '.worcs'.", verbose = verbose)
+  tryCatch({
+    if(verbose) cli::cli_process_start("Storing synthetic data in {.val fn_write_synth_rel} and updating the checksum in {.val .worcs}.")
 
-
-  # Obtain save_expression from the worcs_file
-  save_expression <- worcs_file$data[[original_name]][["save_expression"]]
-  # If there is no save_expression, this is a legacy worcs_file.
-  # Use the default save expression of previous worcs versions.
-  if(is.null(save_expression)){
-    save_expression <- "write.csv(data, filename, row.names = FALSE)"
-  }
-  # Create an environment in which to evaluate the save_expression, in which
-  # filename is an object with value equal to fn_write_synth
-  save_env <- new.env()
-  assign(x = "filename", value = fn_write_synth_abs, envir = save_env)
-  out <- eval(parse(text = save_expression), envir = save_env)
-  # Add info to worcs_file
-  to_worcs$data[[original_name]]$synthetic <- fn_write_synth_rel
-  store_checksum(fn_write_synth_abs, entry_name = fn_write_synth_rel, worcsfile = fn_worcs)
-  write_gitig(fn_gitig, paste0("!", basename(fn_synthetic)))
-  col_message("Updating '.gitignore'.", verbose = verbose)
-  fn_readme <- path_abs_worcs("README.md", dn_worcs)
-  if(file.exists(fn_readme)){
-    lnz <- readLines(fn_readme)
-    if(!any(grepl("Synthetic data with similar", lnz, fixed = TRUE))){
-      update_textfile(filename = fn_readme,
-                      txt = "Synthetic data with similar characteristics to the original data have been provided. Using the function load_data() will load these synthetic data when the original data are unavailable. Note that these synthetic data cannot be used to reproduce the original results. However, it does allow users to run the code and, optionally, generate valid code that can be evaluated using the original data by the project authors.",
-                      next_to = "Some of the data used in this project are not publically available.",
-                      verbose = verbose)
+    # Obtain save_expression from the worcs_file
+    save_expression <- worcs_file$data[[original_name]][["save_expression"]]
+    # If there is no save_expression, this is a legacy worcs_file.
+    # Use the default save expression of previous worcs versions.
+    if(is.null(save_expression)){
[…]

### 2024-12-29 R/codebook.R

@@ -70,7 +70,7 @@ make_codebook <-
             category = NA,
             description = NA)
     if (file.exists(filename)) {
-      col_message(paste0("Removing previous version of '", filename, "'."), verbose = verbose)
+      col_message(paste0("Removing previous version of {.val filename}."), verbose = verbose)
       invisible(file.remove(filename))
     }
     draft(

### 2024-12-29 R/descriptives.R

@@ -218,17 +218,22 @@ skew_kurtosis.default <-
     }
   }
 
+
+is_quiet <- function() {
+  isTRUE(getOption("usethis.quiet", default = FALSE))
+}
+
 #' @importFrom usethis ui_oops ui_done
-col_message <- function (..., col = 30, success = TRUE, verbose = TRUE){
+col_message <- function (..., col = 30, success = TRUE, verbose = !getOption("usethis.quiet", default = FALSE)){
   if(verbose){
     txt <- do.call(paste0, list(...))
     # Check if this function is called from within an rmarkdown document.
     # If that is the case, the colorized messages can cause knitting errors.
     if(!any(grepl("rmarkdown", unlist(lapply(sys.calls(), `[[`, 1)), fixed = TRUE))){
       if(success){
-        usethis::ui_done(txt)
+        cli::cli_bullets(text = c("v" = txt))
       } else {
-        usethis::ui_oops(txt)
+        cli::cli_bullets(text = c("x" = txt))
       }
     }
   }

### 2024-12-29 R/github.R

@@ -260,7 +260,12 @@ git_remote_create <- function(name, private = TRUE){
   tryCatch({
     cli::cli_process_start("Creating GitHub repository '{git_usrnm}/{name}'")
     if(git_usrnm == "username") stop()
-    invisible(gh::gh("POST /user/repos", name = name, private = c("false", "true")[as.integer(private)+1]))
+    if(private){
+      invisible(gh::gh("POST /user/repos", name = name, private = "true"))
+    } else {
+      invisible(gh::gh("POST /user/repos", name = name))
+    }
+
     cli::cli_process_done() },
     error = function(err) {
       cli::cli_process_failed()

### 2024-12-31 R/worcs_project.R
update 2024-12-31 14:02:42.771659

@@ -23,21 +23,26 @@ recommend_data <- c('library("worcs")',
 #' \code{\link[prereg:prereg]{prereg}} package. For more information, see
 #' \code{\link{add_preregistration}}.
 #' @param add_license Character, indicating what license to include.
-#' Default: 'CC_BY_4.0'. Available options include:
-#' \code{"CC_BY_4.0", "CC_BY-SA_4.0", "CC_BY-NC_4.0", "CC_BY-NC-SA_4.0",
-#' "CC_BY-ND_4.0", "CC_BY-NC-ND_4.0", "None"}. For more information, see
-#' <https://creativecommons.org/licenses/>.
+#' Default: 'ccby'. Available options include:
+#' \code{c("cc0", "ccby", "gpl", "gpl3", "agpl", "agpl3", "apache", "apl2",
+#' "lgpl", "mit", "proprietary", "None"}. For more information, see
+#' \code{\link[usethis]{use_cc0_license}}.
 #' @param use_renv Logical, indicating whether or not to use 'renv' to make the
 #' project reproducible. Default: TRUE. See \code{\link[renv]{init}}.
 #' @param use_targets Logical, indicating whether or not to use 'targets' to
-#' create a Make-like pipeline. Default: FALSE See \code{\link[targets]{targets-package}}.
-#' @param remote_repo Character, address of the remote repository for
-#' this project. This link should have the form
+#' create a Make-like pipeline.
+#' Default: FALSE See \code{\link[targets]{targets-package}}.
+#' @param remote_repo Character, URL of, or name for, the remote repository for
+#' this project. If a URL of an existing repository is specified, it should have
+#' the form
 #' \code{https://github.com[username][repo].git} (preferred) or
 #' \code{git@[...].git} (if using SSH).
-#' If a valid remote repository link is provided, a commit will
+#' Alternatively, a name for a new repository can be provided. If a 'GitHub'
+#' user is authenticated on your device, this repository will be created on your
+#' account.
+#' Finally, a commit will
 #' be made containing the 'README.md' file, and will be pushed to the remote
-#' repository. Default: 'https'.
+#' repository. Default: 'https', which results in no repository being created.
 #' @param verbose Logical. Whether or not to print messages to the console
 #' during project creation. Default: TRUE
 #' @param ... Additional arguments passed to and from functions.
@@ -228,29 +233,15 @@ worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregist
   }
 
[…]

### 2025-01-08 README.Rmd
update 2025-01-08 15:29:03.686524

@@
 [![R-CMD-check](https://github.com/cjvanlissa/worcs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cjvanlissa/worcs/actions/workflows/R-CMD-check.yaml)
-[![Codecov test coverage](https://app.codecov.io/gh/cjvanlissa/worcs/branch/master/graph/badge.svg)](https://app.codecov.io/gh/cjvanlissa/worcs?branch=master)
 [![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-v2.0%20adopted-ff69b4.svg)](code_of_conduct.md)
 [![CII Best Practices](https://bestpractices.coreinfrastructure.org/projects/3969/badge)](https://bestpractices.coreinfrastructure.org/projects/3969)
-<!--[![DOI](http://joss.theoj.org/papers/10.21105/joss.00978/status.svg)](https://doi.org/10.21105/joss.00978)-->
 <!-- badges: end --->

### 2025-01-10 R/worcs_project.R
update 2025-01-10 09:47:57.727134

@@ -226,15 +226,12 @@ worcs_project <- function(path = "worcs_project", manuscript = "APA6", preregist
   }
 
   # Connect to remote repo if possible
-  valid_repo <- isFALSE(remote_repo == "https")
-  #
-  # repo_url <- parse_repo(remote_repo = remote_repo, verbose = verbose)
-  # valid_repo <- !is.null(repo_url)
+  valid_repo <- isFALSE(is.null(remote_repo) | remote_repo == "https")
+
   if(use_git & valid_repo){
     out_git <- git_connect_or_create(repo = path, remote_repo = remote_repo)
     out_push <- try(git_push(remote = "origin", repo = path, verbose = FALSE))
-    # out_test <- git_remote_test(repo = path)
-    # col_message("Could not connect to a remote 'GitHub' repository. You are working with a local 'Git' repository only.", success = FALSE, verbose = verbose)
+
   }
   if("GCtorture" %in% ls()) rm("GCtorture")
 }

### 2025-01-10 tests/testthat/helper_usethis.R

@@ -1,3 +1,14 @@
+skip_if_not_pandoc <- function(ver = NULL) {
+  if (!pandoc_available(ver)) {
+    msg <- if (is.null(ver)) {
+      "Pandoc is not available"
+    } else {
+      sprintf("Version of Pandoc is lower than %s.", ver)
+    }
+    skip(msg)
+  }
+}
+
 # copied from usethis:
 # https://github.com/r-lib/usethis/blob/b2e894eb6d1d7f3312a783db3bb03a7cc309ba61/tests/testthat/helper.R
 library(usethis)

### 2025-01-10 tests/testthat/test-checksum_markdown.R

@@ -1,20 +1,20 @@
 test_that("checksum works from within Rmarkdown", {
-  the_test <- "cs_markdown"
-  old_wd <- getwd()
-  test_dir <- file.path(tempdir(), the_test)
-  dir.create(test_dir)
-  setwd(test_dir)
-  on.exit(unlink(test_dir, recursive = TRUE), add = TRUE)
 
-
+  skip_if_not_pandoc("2.0")
+  scoped_temporary_project()
+  test_dir = "."
   worcs:::write_worcsfile(file.path(test_dir, ".worcs"))
   dat <- iris[1:4,1:4]
   open_data(dat)
   cs_correct <- read_yaml(".worcs")$checksums[["dat.csv"]]
+
   expect_true(worcs:::cs_fun(filename = file.path(test_dir, "dat.csv"), worcsfile = file.path(test_dir, ".worcs")) == cs_correct)
+
   dir.create(file.path(test_dir, "manuscript"))
   setwd(file.path(test_dir, "manuscript"))
-  expect_true(worcs:::cs_fun(filename = file.path(test_dir, "dat.csv"), worcsfile = file.path(test_dir, ".worcs")) == cs_correct)
+
+  # expect_true(worcs:::cs_fun(filename = file.path(test_dir, "dat.csv"), worcsfile = file.path(test_dir, ".worcs")) == cs_correct)
+  #
   man_txt <- c("---", "title: \"Untitled\"", "output: github_document", "---",
                "", "```{r setup}", "library(\"worcs\")", "load_data()", "```", "", "`r worcs:::cs_fun(file.path(dirname(worcs:::check_recursive(file.path(normalizePath('.')))), 'dat.csv'), file.path(dirname(worcs:::check_recursive(file.path(normalizePath('.')))), '.worcs'))`")
   writeLines(man_txt, "manuscript.Rmd")
@@ -22,6 +22,5 @@ test_that("checksum works from within Rmarkdown", {
   md_checksum <- tail(readLines("manuscript.md"), 1)
   expect_true(md_checksum == cs_correct)
 
-  setwd(old_wd)
 })
 

### 2025-01-10 tests/testthat/test-citations.R

@@ -1,7 +1,7 @@
 library(rmarkdown)
 
 test_that("documents with spaces in names can be rendered", {
-
+  skip_if_not_pandoc("2.0")
   the_test <- "citeall"
   dir_name <- file.path(tempdir(), the_test)
   if(dir.exists(dir_name)) unlink(dir_name, recursive = TRUE)

### 2025-01-10 tests/testthat/test-dir_of_save_data.R

@@ -10,7 +10,7 @@ test_that("data stored in correct dir", {
   add_manuscript(
     worcs_directory = ".",
     manuscript = "github_document",
-    remote_repo = "bla"
+    remote_repo = "https"
   )
   data("iris")
   df1 <- iris[1:5, 1:4]

### 2025-01-10 tests/testthat/test-github.R

@@ -2,6 +2,7 @@ if(FALSE){
 library(targets)
 test_that("targets works with apa6", {
   scoped_temporary_project()
+  skip_if_not_pandoc("2.0")
   # browser()
   unlink("R", recursive = TRUE)
   worcs::worcs_project(path = ".",
@@ -52,6 +53,7 @@ test_that("targets works with apa6", {
 
 
 test_that("targets works with target markdown", {
+  skip_if_not_pandoc("2.0")
   scoped_temporary_project()
   unlink("R", recursive = TRUE)
   worcs::worcs_project(path = ".",

### 2025-01-10 tests/testthat/test-targets.R

@@ -1,94 +1,96 @@
-library(targets)
-test_that("targets works with apa6", {
-  scoped_temporary_project()
-  # browser()
-  unlink("R", recursive = TRUE)
-  worcs::worcs_project(path = ".",
-                       manuscript = "github_document",
-                       preregistration = "None",
-                       add_license = "None",
-                       use_renv = FALSE,
-                       use_targets = TRUE,
-                       use_git = FALSE
-                       )
-  lnz <- readLines("_targets.R")
-  lnz[grep("tarchetypes::", lnz, fixed = TRUE)] <- gsub(")", ", cue = tar_cue(mode = 'always'))", lnz[grep("tarchetypes::", lnz, fixed = TRUE)], fixed = TRUE)
-  writeLines(lnz, "_targets.R")
-  # It seems like it's necessary to first render individual output chunks,
-  # then render the manuscript,
-  # and THEN it's possible to render the manuscript using tar_make()
+if(require(targets)){
+  test_that("targets works with apa6", {
+    scoped_temporary_project()
+    skip_if_not_pandoc("2.0")
+    # browser()
+    unlink("R", recursive = TRUE)
+    worcs::worcs_project(path = ".",
+                         manuscript = "github_document",
+                         preregistration = "None",
+                         add_license = "None",
+                         use_renv = FALSE,
+                         use_targets = TRUE,
+                         use_git = FALSE
+    )
+    lnz <- readLines("_targets.R")
+    lnz[grep("tarchetypes::", lnz, fixed = TRUE)] <- gsub(")", ", cue = tar_cue(mode = 'always'))", lnz[grep("tarchetypes::", lnz, fixed = TRUE)], fixed = TRUE)
+    writeLines(lnz, "_targets.R")
+    # It seems like it's necessary to first render individual output chunks,
+    # then render the manuscript,
+    # and THEN it's possible to render the manuscript using tar_make()
 
-  tryCatch(targets::tar_make(), error = function(e){}, warning = function(w){})
-  # expect_true(file.exists("manuscript/manuscript.html"))
-  tryCatch(rmarkdown::render("manuscript/manuscript.Rmd"), error = function(e){}, warning = function(w){})
-  expect_true(file.exists("manuscript/manuscript.html"))
-  if(file.exists("manuscript/manuscript.html")){
-    file.remove("manuscript/manuscript.html")
-  }
-  cat(readLines("_targets.R"))
[…]

### 2025-01-10 README.Rmd
update 2025-01-10 09:52:05.955697

@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.alt="A diagram of the WORCS workflow."}
 knitr::include_graphics("https://github.com/cjvanlissa/worcs/raw/master/paper/workflow_graph/workflow.png")
